### 2015-09-02 submission.tex
first commit

\documentclass[conference]{IEEEtran}

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf

\else

\fi

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}
\usepackage{algorithm}
%\usepackage{algorithm2e}
\usepackage{algpseudocode}
\usepackage{booktabs}
\usepackage{flushend}
\usepackage{url}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{graphicx}
%\usepackage{caption}
%\usepackage[labelfont=bf]{caption}
\usepackage[font=bf,labelsep=space]{caption}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
%\title{Network contention aware batch Scheduling on Torus-Connected Systems}
\title{ \textcolor{red}{Deja Vu: A New Topology-aware HPC Batch Scheduler with Workload Precognition} }
% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{

\IEEEauthorblockN{Xu Yang\IEEEauthorrefmark{1}, John Jenkins\IEEEauthorrefmark{2}, Misbah Mubarak\IEEEauthorrefmark{2}, Robert B. Ross\IEEEauthorrefmark{2}, Zhiling Lan\IEEEauthorrefmark{1}}
\IEEEauthorblockA{\IEEEauthorrefmark{1}Department of Computer Science,
Illinois Institute of Technology,
Chicago, Illinois, USA 60616\\
xyang56@hawk.iit.edu, lan@iit.edu}

\IEEEauthorblockA{\IEEEauthorrefmark{2}Mathematics and Computer Science Division, Argonne National Laboratory,
Argonne, IL, USA 60439\\
\{jenkins,rross\}@mcs.anl.gov, mmubarak@anl.gov}
}

% make the title area
\maketitle

\begin{abstract}

It has been widely recoginzed that network contension between concurrently running jobs in HPC system is one of the primary causes of performance variability. Optimizing job allocation and avoiding network sharing have been proved to be effective for alleviating network contension. In this work, we propose a new job scheduler framework for HPC system, which not only take job's size and expected runtime as primary submission parameters, but also an extra one that indicate job's dominate communication pattern(CP). Using this extra parameter CP can help our new scheduler framework enhance both job and system's performance in three ways. First, communication pattern indicates whether the job's performance rely heavily on the network or not. Second, with the precongnition of job's dominate communication pattern, the scheduler can provide job with better resource allocation. Third, when multiple jobs are dispatched to the system, knowing their communication pattern can help avoid network contension. Our extensive experimental results show that the new scheduling framework can greatly improve both system and user's job performance.

\end{abstract}

\IEEEpeerreviewmaketitle

\section{Introduction} 
\label{sec: intro}

The batch scheduler has always been a undervalued component in HPC system. As an interface between user and system, the batch scheduler usually do nothing more than simply take the submitted job from user and hand over to the system for execution. User are required to provide information about system resource needed by his job, such as job size (i.e. how many nodes are needed to run the job) and job's expected runtime (i.e. how long does user expected his job take to complete) upon job's submission. The batch scheduler keeps updated about system resource availability, when it being informed there are enough number of nodes becomes available, the scheduler picks up a job from the job queue and dispatch it to the system.

Most existing batch schedulers take so few initiative  because it been hold under the obscurantism policy. There are two observations about the ignored information that batch scheduler should have been informed. First, the workload on HPC system tend to be repetitive. We analyzed the workload on Mira, an IBM Blue/Gene Q system installed in Argonne National Lab, found that most jobs require 2k nodes and about 14 particular jobs consist of the majority its workload. In other word, Mira accommodates mostly some particular 14 jobs and most jobs have the same size. The second thing that batch scheduler should know is job's communication pattern. When a user submit his job, he at least has a vague idea about the most dominant communication pattern of his job (i.e. All-to-All, Nearest Neighbor, Collective).

In this work, we propose a new HPC batch scheduler framework, which can capture the workload signature introduced by repetitively submitted jobs and  integrates network simulation feedback control. In spite of the input parameters such as job size and expected runtime that traditional batch scheduler requires, our new framework also takes job's communication pattern (CP) as an extra parameter, which can be utilized when assigning job with its preferred allocation. Being aware of job's CP has several adavantages for improving both job and system performance. For example, CP indicates the better allocation for the job and thus, bath scheduler can get the job prefered allocation without embezzling extra system resources.

Unlike the traditional batch scheduler that is agnostic about upcoming jobs in the workload, our new framework is aware of the repetitiveness of HPC workload and take advantage of this workload feature for making the optimal scheduling/allocation decision for upcoming jobs. To endow the new framework with this ability, we implement a memory mechanism that keep record of the preference about allocation of jobs with different CPs.

To receive the feedback about resource availability from system, the new framework will keep real time records about the both network and node availability. The new framework will be informed with information such as how many nodes are available, layout and geometry of those available nodes  (contiguous or non-contiguous, ), and their network resources (connnected by torus or mesh). Our new framework will also collect certain information for each job when the job running on some its allocated resources, such as job's actual running time, network resource utilization, to determine whether its allocated resources is the optimal regarding to its communication pattern. These feedback information from the runtime system will be utilized for job's furture repetitive submission.

With more detailed information (i.e. job size, expected runtime, CP) about job provided by user, combined with feedback about job's previous execution and resource availability from system, the new scheduler framework can make better scheduling/allocation decisions. And with the new memory mechanism, our new framework is equiped with self-learning capability and the more jobs it handled, the more talented it becomes.

The major contributions of this paper:
\begin{itemize}
    \item We analyze the repetitiveness as one of signatures for the workload trace collected from Mira, an IBM Blue Gene/Q system at Argonne National Lab.

    \item We analyze the behavior of four most prevalent communication patterns among HPC jobs, under the circumstance of different allocation.

    \item We are the first to propose that HPC batch scheduler should take job's communication pattern (CP) as extra input, which brings benefit for both job and system.

    \item Our new scheduling framework establish a feedback control link between system and job scheduling module. (simple feedback control)

\end{itemize}

\section{System Observations}
\label{sec: observation}

Our work is based on two major observations about HPC system. First, the job submissions on HPC tend to be repetitive. The same application could be submitted to HPC system repetitively with different different input data or configuration. Second, most applications running on HPC system are highly communication intensive, and each application has different dominant communication patterns during its different execution phases. In this section, we will describe this the repetitiveness of job submission and applications' communication patterns.

\subsection{Repetitive Job Submission}
\label{sec: repetitive job submission}

The same application could run repetitively on HPC system. This is due to users from some scientific areas have to run the same application multiple times with different configurations. For example, some bioinformatics applications need to conduct the same computation i.e. protein sequence alignment, for a large sets of protein sequences. From system's perspective, all these submissions are the same (same computation instructions and communication phases). Therefore, we can observe the same job multiple times with slightly different input parameter, and all these jobs usually belong to the same user. To substantiate our claim, we analyze the workload trace in 2014 from Mira, an IBM Blue/Gene Q system. As it shown in figure \ref{fig: repetitiveness of Mira},  we list the number of total submissions from five most active users and the percentage of the repetitive ones.

\begin{figure}[h!] 
  \centering
  \includegraphics[width=0.38\textwidth]{Pic/JobInfo/repe-mira-2}
   \caption{ The repetitive job submissions on Mira in 2014. The percentage of repetitive submissions from certain user can be as much as 90\%. }
   \label{fig: repetitiveness of Mira}
\end{figure}

\subsection{Communication Pattern}
\label{sec:communication pattern matrix}

In this section, we analysis four communication patterns that are most prevalent in scientific computation.
They are All-to-All, Nearest Neighbor, Collective and Random\cite{roth}. All these four communication patterns are so typical that Message Passing Interface(MPI), the most popular parallel programming scheme, provides explicit support\cite{Pjesivac} \cite{thakur}.

A parallel application usuall conforms to a combination of several basic communication patterns. At its different execution phases, the application's communication behavior may follow different basic patterns repectively. There are many profiling tools available to capture the communication behavior from parallel applications, such as Tuning and Analysis Utilities(TAU)\cite{tau}, mpiP\cite{mpip} and ScalaTrace\cite{scala}. They can keep track of how much data was transferred between each pair of processes in the application and how long those transfer operations took. With the help from such profiling tools, we can get a matrix that summarizes the communication between processes within the application. Such matrix could contain information about

\begin{itemize}
 \item how many times each process send/receive data to each other process using MPI point-to-point operations.
 \item the amount of data has been sent/received in each MPI operation between processes.
 \item the amount of data has been sent/received by each process using MPI collective operation

\end{itemize}

that can be calibrated by these three communication patterns, we suppose each has \emph{n} parallel processes, we use $m_{i,j}$ to denote the amount of data sent from the \emph{i}th process to \emph{j}th process via network. Communication pattern matrix $M$ is used to characterizing the total amount of data transferred among all the processes within each application.

\subsubsection{All to All}
\label{sec:all-to-all}

Within parallel application that conforms to All-to-All (A2A) communication pattern, each process sends distinct data to all other processes. The implementation of this A2A communication in MPI is MPI\_Allgather. The communication pattern matrix of this category $M_{a2a}$ like this

\begin{displaymath}
\begin{bmatrix}
  0 & m_{1,2} & m_{1,3} & \cdots & m_{1,n-1} & m_{1,n}\\
  m_{2,1} & 0 & m_{2,3} &\cdots & m_{2,n-1}  & m_{2,n}\\
  m_{3,1} & m_{3,2} & 0 &\cdots & m_{3,n-1}  & m_{3,n}\\
  \vdots  & \vdots & \vdots & \ddots & \vdots & \vdots \\
  m_{n-1,1} & m_{n-1,2} & m_{n-1,3} & \cdots & 0 & m_{n-1,n}\\
  m_{n,1} & m_{n,2} & m_{n,3} & \cdots & m_{n,n-1} & 0
\end{bmatrix}
\end{displaymath}

In $M_{a2a}$, $m_{i,j} = 0$, for every $i=j$, all the elements on diagonal are 0. This is because process doesn't send data via network to itself. As we can see $M_{a2a}$ has a high density and its rank is $n$. One of the representative benchmark application that conforms to A2A communication is DNS3D from NPB3. DNS3D is highly sensitive to network and node allocaiton, since during each step, it executes three Fourier transforms for 3-D scalar variables\cite{zhou-ipdps}.
 
\subsubsection{Nearest Neighbor}
\label{sec:nearest neighbor}

Nearest Neighbor (NN) communication pattern is very common in scientific computations. This is because lots of physical systems are simulated by a set of elements arranged in grid. Cell-based AMR (Adaptive Mesh Refinement) cosmology simulation is one of this kind. Wu et al. in \cite{jingjin} present an ART application that implement cell-based AMR conforms to diagonal dominant NN communication pattern. Its communication pattern matrix $M_{nn}$ like

\begin{displaymath}
 \begin{bmatrix}
  0 & m_{1,2} & 0 & \cdots & 0 & 0\\
  m_{2,1} & 0 & m_{2,3} &\cdots & 0  & 0\\
  0 & m_{3,2} & 0 &\cdots & 0  & 0\\
  \vdots  & \vdots & \vdots & \ddots & \vdots & \vdots \\
  0 & 0 & 0 & \cdots & 0 & m_{n-1, n}\\
  0 & 0 & 0 & \cdots & m_{n, n-1} & 0
 \end{bmatrix} 
\end{displaymath}

In $M_{nn}$, $m_{i,j} = 0,$ when $|i-j|\neq 1$, all the elements are equal to zero except those on upper or lower diagonal. Matrix $M_{nn}$ is very sparse and its rank is n-1;

\subsubsection{Collective}
\label{sec:one-to-allandall-to-one}
\textcolor{red}{This senction is out of scope now, since we don't have MPI collective results. Every following section talk about Collectives are also obsolete.}

Collective communication is a frequently used component in parallel programming schemes like MPI.  One-to-All (O2A) and All-to-One (A2O) are chosen as typical collective operations in this paper. During a O2A operation, one process (usually process with rank ``root'') sends data to all other processes. The implementation of this O2A communication in MPI is MPI\_Bcast and MPI\_Scatter. So, the matrix $M_{o2a}$ is like

\begin{displaymath}
 \begin{bmatrix}
  0 & 0 &\cdots & 0 &\cdots & 0\\
  0 & 0 &\cdots & 0 &\cdots & 0\\
  \vdots  & \vdots & \ddots & \vdots & \ddots & \vdots \\
  m_{i,1} & m_{i,2} & \cdots & m_{i,j} &\cdots & m_{i,n}\\
  \vdots  & \vdots & \ddots & \vdots & \ddots & \vdots \\
  0 & 0 & \cdots 0 &\cdots & 0
\end{bmatrix} 
\end{displaymath}

And another commonly used collective communication is All-to-One(A2O), which is the inverse of One-to-All. Instead of spreading data from one process to many others, data in this A2O communication are sent from many processes and gathered to one single process. The implementation of this A2O communication in MPI is MPI\_Gather. O2A communication is so powerful for many parallel algorithms, such as parallel sorting and searching. The matrix $M_{a2o}$ like

\begin{displaymath}
\begin{bmatrix}
  0 & 0 &\cdots & m_{1,j} &\cdots & 0\\
  0 & 0 &\cdots & m_{2,j} &\cdots & 0\\
  \vdots  & \vdots & \ddots & \vdots & \ddots & \vdots \\
  0 & 0 & \cdots & m_{i,j} &\cdots & 0\\
  \vdots  & \vdots & \ddots & \vdots & \ddots & \vdots \\
  0 & 0 & \cdots & m_{n, j} &\cdots & 0
\end{bmatrix}  
\end{displaymath}

For O2A and A2O collective communication, their matrixes have all zero elements except only one row or one column. The rank of both matrixes' is 1.\\

\subsubsection{Hybrid Communication Pattern}

In real life, parallel scientific computation may conform to different communication at different phases. During the start phases, applications tend to do some collective communications, such as O2A to distribute data among tasks and in during the end phases, A2O are used to collect comuptation results from every tasks. Some applications may conforms to more than one kind of dominant communication pattern, such as pF3D, which is a laser-plasma interaction code that performs local All-to-All communication over sub-communicators and Nearest Neighbor communication over a 3D process grid\cite{langer}. There is no real application that exclusively conforms to certain single communication pattern. In order to represent real application that with different communication pattern during the its different execution phases, its communication pattern matrix can be formulated like

\begin{equation}
  M_{hyb} = \alpha_{1}\cdot M_{a2a}+\alpha_{2}\cdot M_{nn}+\alpha_{3}\cdot M_{a2o}+\alpha_{4}\cdot M_{o2a}
\end{equation}
And 
\begin{equation}
  \sum_{i=1}^{4}\alpha_{i} \equiv 1
\end{equation}

Here, $\alpha_{i}$ indicates the percentage of communication amount that conforms to each particular communication pattern.

\subsection{Distance Matrix}
\label{sec:desitance matrix}

In HPC system, batch scheduler plays a important role in job scheduling and system resource management. Batch scheduler is responsible for assigning each job the amount of resource (usually refers to compute nodes) it required. The amount of compute nodes a job required refers to the job's size. Suppose the scheduler provide a job, with size $n$, a set of node $S = \{N_{1}, N_{2},...,N_{n}\}$. If there is a direct link connected between node $N_{i}$ and $N_{j}$, the distance between them $d_{i,j} = 1$. Thus, we can use a distance matrix $D$ to characterizing the distance between nodes within each allocation. Here is what distance matrix $D$ looks like

\begin{displaymath}
 D = \begin{bmatrix}
  0 & d_{1,2} & d_{1,3} & \cdots & d_{1,n-1} & d_{1,n}\\
  d_{2,1} & 0 & d_{2,3} &\cdots & d_{2,n-1}  & d_{2,n}\\
  d_{3,1} & d_{3,2} & 0 &\cdots & 0  & 0\\
  \vdots  & \vdots & \vdots & \ddots & \vdots & \vdots \\
  d_{n-1,1} & d_{n-1,2} & d_{n-1,3} & \cdots & 0 & d_{n-1,n}\\
  d_{n,1} & d_{n,2} & d_{n,3} & \cdots & d_{n,n-1} & 0
 \end{bmatrix} 
\end{displaymath}

Apparently, the distance between two nodes within each allocation highly depends on system network topology.\\

\subsubsection{k-ary n-cube Distance Matrix}
\label{sec:k-ary n-cube distance matrix}

As the most popular interconnect nework adopted in current supercomputers, torus (e.g., k-ary n-cube) network has many attractive features like short diameter and low bisection bandwidth \cite{zhou-ipdps}. Torus interconnected network has been adopted by IBM Blue Gene and Cray XT series machines\cite{Dong-SC11}. In this paper, we use torus connected system to demonstrate the effectiveness of our new scheduling framework. Suppose the system network topology is k-ary n-cube,  the diameter of this topology is $\frac{nk}{2}$, and the average distance between any two nodes in this topology is $\frac{nk}{4}$. So, $d_{i,j}$ in the distance matrix $D$ of any allocation made on k-ary n-cube will be

\begin{equation}
1\leq d_{i,j}\leq \frac{nk}{2}
\end{equation}

And the average value of $d_{i,j}$ is $\frac{nk}{4}$.

Jobs running on Mira, the BG/Q machine and ANL, can specify its allocation topology when been submitted. It can choose between with or without torus connection. However, allocation without torus connection will suffer a degradation in terms of distance matrix, $d_{i,j}$ will become

\begin{equation}
  1\leq d_{i,j} \leq k\cdot n
\end{equation}

\section{Problem Formulation}
\label{sec:problem formulation}

Although there are lots work about finding the optimal mapping for assigning each process to each node \cite{hpdc14}\cite{jingjin}, in this work, the batch scheduler will assign each job the optimal node allocation with consideration of their communication pattern. The traditional scheduler always try to give each job a compact allocation or the node allocation with network resource that more than enough. We will demonstrate that such generous allocation is unnecessary for job with different communication pattern.

Given a job with communication matrix $M$ and its allocation distance matrix is $D$, the metric hop-byte $M\cdot D$ is usually used to quantify the network cost. The objective of the scheduler is try to find each job a node allocation, whose distance matrix $D$ is most similar to the transpose of communication pattern matrix $M^{\rm T}$, with the objective of

\begin{equation}
\min \sum_{i=1}^{n} \sum_{j=1}^{n} M \cdot D
\end{equation}

The traditional scheduler always try to give each job a compact (in terms of shape) allocation or an allocation with superabundant network resource so as to minimize every $d_{i,j}$, which further leads to minimum $M\cdot D$. However, if we check the characteristic of those three typical communication pattern in section\ref{sec:communication pattern matrix}, we can see finding a allocation with minimized every $d_{i,j}$ in distance matrix $D$ cost way too much effort. \textbf{This is because $\min_{i,j \in \{1,n\}} M \cdot D $ depends only on the elements in D at the position of the transpose of $M$.}

Take the One-to-All communication pattern for an example, providing a job conforms to O2A communication pattern with node allocation with minimized every $d_{i,j}$ in $D$ would be such a huge waste, since those $d_{i,j}$ that matter in this scenario reside in only one column. Thus, our scheduler only need to find a allocation within which, exist one node that has short distance to all the other nodes.

\section{Simulation Validation}
\label{sec:simulation validation}

With the help from parallel discrete event driven simulation, simulating large scale interconnect network with high fidelity becomes possible\cite{ross}. Jason et al. from Argonne National Lab develop this simulation toolkit CODES, which provides torus and dragonfly network simulation with high fidelity\cite{Jason-2011}\cite{mubarak-sc2012}.

We conduct massive simulation to validate our new scheduler framework model. Inorder to get a high fidelity of the network performance under workloads with different communication patterns, we use CODES, a parallel discrete event driven simulation toolkit from Argonne National Lab, which can provide torus connected network simulation with high fidelity\cite{Jason-2011}. Torus network model in CODES is built based on several assumptions. The model uses deterministic static routing for its simplicity and data injection rate follows the Poisson Process with mean arrival rate of $\lambda$ \cite{Jason-2011}. In our simulation, each node continuously generates a Poisson Stream of packets, the size of which is $\delta$. The total amount of data generated by each node is $\Delta$, here

\begin{equation}
\Delta = \sum_{j=1}^{n} m_{i, j} \quad i \in {1,n}
\end{equation}

\subsection{Simulation Settings}
\label{sec:simulation settings}
\textcolor{red}{The scale of network and job would be changed based on the CODES simulation}

For a job requires $n$ nodes submitted to 3D torus-connected system, it gets an allocation of various possible shapes, which could be 3D balanced-cube, 3D unbalanced-cube, 2D balanced-mesh, 2D unbalanced-mesh, 1D list. Figure \ref{fig: allocation} shows some possible allocations for a 64-node job. We use vector $v$ to denote the shape of an allocation. The vector of 3D balanced-cube allocation (in Figure \ref{fig: allocation} as red) is $v=\{4,4,4\}$, 3D unbalanced-cube allocation is $v=\{8,4,2\}$(in Figure \ref{fig: allocation} as green), 2D balanced-mesh allocation is $v=\{8,8,1\}$ (in Figure \ref{fig: allocation} as blue).

\begin{figure}[h!] 
  \centering
  \includegraphics[width=0.48\textwidth]{Pic/allocationshape/allocation}
  \caption{64-node job with different allocation shapes. Red, Green and Blue represents 3D balance, 3D unbalance and 2D balance allocation respectively. }
  \label{fig: allocation}
\end{figure}

The workload traces we used for evaluating our new scheduling framework consist of jobs with size variation from 64 to 16K. Due to space limitation, we only present the results for job size 64 and 512, both conforms to three different communication patterns. We suppose the scale of the torus network simulated in this experiment is set to $32 \times 32 \times 32 $, $32,768$ nodes in total, thus all the possible allocation for job with size 64 and 1024 are shown in table \ref{table:allocationvector}. Since we construct a symmetric 3D torus network in our simulation (the bandwidth of each link in x, y, x direction are the same), we suppose the allocation with $v=\{x, y, z\}$ is the same compared to all the other allocations with the dimension of the permutation of \{x, y, z\} in terms of jobs performance.

\begin{table}
\caption{Different allocation vectors for job with size 64-node and 512-node}
\label{table:allocationvector}
\begin{center}
\begin{tabular}{llllr}
\toprule
\multicolumn{4}{c}{Item} \\
\cmidrule(r){2-5}
Job Size    & 3D-B & 3D-UB & 2D-B & 2D-UB  \\
\midrule
64-node      & \{4,4,4\}    & \{8,4,2\}   & \{8,8,1\}  & \{16,4,1\}      \\
             &       &     &      & \{32,2,1\} \\
\midrule
512-node    & \{8,8,8\}   & \{16,8,4\} &     &       \\
            &       & \{16,16,2\} &    &  \{32,16,1\}      \\
            &  & \{32,8,2\} & &\\
            &  & \{32,4,4\} & &\\
\midrule
1024-node    &    & \{16,8,8\} &   \{32,32,1\}  &       \\
            &       & \{16,16,4\} &    &       \\
            &  & \{32,8,4\} & &\\
            &  & \{32,16,2\} & &\\

 \bottomrule
\end{tabular}
\end{center}
\end{table}

In our simulation, we set the packet size $\delta$ to 256 Byte. The amount of data transferred by each node pair $m_{i,j}$ is set to 2KB, which indicates $2KB/256B = 8$ packets will be transfterred between each node pair through the job's execution. The number of packets transferred between each node pair $\sigma = m_{i,j}/\delta$. The job size in the simulation is set to $n$.

\begin{table}
\caption{Nomenclature}
\label{table:nomenclature}
\begin{center}
  \begin{tabular}{ l | c }
    \hline \hline
    Symbol 	& Description \\ \hline \hline
    $\delta$ 	& packet size. \\ \hline
    $\sigma$    & the number of packets transferred between each node pair. \\ \hline
    $\Delta$    & the total amount of data generated by each node, $\Delta = \delta \cdot \sigma$. \\ \hline
    $N $   	& System size, i.e., the number of nodes in the system.\\ \hline
    $n$    & job size, i.e., the number of nodes required by the job.\\ \hline
    $M$    & the communication pattern matrix of each job.\\ \hline
    $m_{i,j}$ 	& the amount of data transferred between process $p_i$ and $p_j$. \\ \hline
    $D$    & the distance matrix of the allocation of each job. \\ \hline
    $d_{i,j}$    & the distance between node $i$ and $j$. \\ \hline
    \hline
  \end{tabular}
\end{center}
\end{table}

\subsection{Results}
\label{sec:results}
Given the system is $32\times 32 \times 32$ symmetric 3D torus, table \ref{table:allocationvector} shows all the possible allocation vectors for jobs with 64/512-node. Table \ref{tab: 64/512node-job} shows the normalized runtime of these two jobs conform to four different communication patterns with each allocation vector. As shown in table \ref{table:allocationvector}, there could be multiple allocation vectors for the same shape, such as the 512-node job could have 4 different allocation vectors with 3D unbalance shape. The average value of differnt allocation vectors for the same shape is presented in table \ref{tab: 64/512node-job}.

\begin{table}[ht]
\begin{center}
\caption{job conforms to four different communication patterns running 3D torus system. In each row, the values on the upper level are the results for 64-node job, the lower level are for 512-node job.
\textcolor{red}{Results in this table would be provided by CODES simulation, job scale depends on the job traces we decide to use}} 
\label{tab: 64/512node-job}
\begin{tabular}{l c c c c c} 
\toprule % Top horizontal line
\toprule
&\multicolumn{5}{c}{Allocation} \\
\cmidrule(l){2-6}
Communication Pattern $M$ & 3D-B & 3D-UB & 2D-B & 2D-UB & NC \\ % Column names row

\midrule % In-table horizontal line
$M_{a2a}$   & 1  &   &    &   &  \\ % Content row 1
           & 1  &   &    &   &  \\ % Content row 1
\midrule
$M_{nn}$  & 1  &   &    &   &  \\ % Content row 1
            & 1  &   &    &   &   \\ % Content row 1
\midrule
$M_{o2a}$   & 1  &   &    &   &  \\ % Content row 1
            & 1  &   &    &   & \\ % Content row 1
\midrule
$M_{a2o}$    & 1  &   &    &   & \\ % Content row 1
             & 1  &   &    &   &  \\ % Content row 1
% \midrule
% $M_{hyb}$   & 1 & 1.24 & 1.37  & 1.66 \\ % Content row 1
%             &  &  &    \\ % Content row 1
\midrule
\bottomrule % Bottom horizontal line
\end{tabular}
\end{center}
\end{table}

\section{Scheduling Mechanism}
\label{sec:scheduling mechanism}

Figure \ref{fig:overall design} shows the overall design of our new scheduling framework. It takes not only job size, user estimated runtime, but also job's communication pattern (CP) as input parameters. The CP here is provided by user, which indicate the most dominant communication pattern of the user job, such as All-to-All, Nearest Neighbor, Collective, ect. The detailed amount of communication data is neither necessary nor possible in this communication pattern. As we prove in section \ref{sec:communication pattern matrix}, job with different CP has different preference about network requirement, hence different preference about allocations. For example, job with All-to-All communication pattern requires more network resource, which means a compact or torus allcoation will guarantee a better performance. However, job conforms to Nearest Neighbor communication pattern won't fully utilize the torus network or compact allocation, which means a mesh allocation will serve it equally well.

\begin{figure}[h!]     
\centering  
\includegraphics[width=0.48\textwidth]{Pic/design/overalldesign}  
\caption{Overview of our new scheduling framework. It takes not only job size, user estimated runtime, but also job's communication pattern matrix as input and combines these user provide information with system information to make optimal prioritizing and allocation decisions. It also maintains hashtable, which provides quick reference for making the specific allocation for incoming job with distinct communication pattern. }
\label{fig:overall design}
\end{figure}

Since HPC workload tend to be repetive, our new scheduling framework keep record of historic jobs information to provide quick reference for making the preferable allocation for incoming job with distinct communication pattern. When user submit his job, our new scheduler will first look for job's user name and executable file name in historic record to check whether this job been submitted before. Then the scheduler gets the job's input parameter \{S, CP, ERT\} (job size, Expected RunTime, CommunicationPattern),  and check in the historic data to identify its preferred allocation. As time goes by, our new framework will gradually build up a comprehensive record of all the possible job submission, which is helpful to provide quick reference for the future jobs. Due to the repetitiveness of the workload, the size of this record will grows rapidly at first, then gently grow for a while and come to stagnation, as shown in figure
\ref{fig:table size growth}.

\begin{figure}[h!]     
\centering  
\includegraphics[width=0.48\textwidth]{Pic/design/tablesize}  
\caption{The size growth of historic submitted job records.}
\label{fig:table size growth}
\end{figure}

 We implement a record mechanism as the central component in our framework to keep track of the pattern in workload repetitiveness, hence make predictive optimizations in terms of both job and system performance. Firgure \ref{fig:scheduling mechanism} shows the overview of our design. The scheduler start to build the table since the first submitted job. When user submit a job, the scheduler get job's size and communication pattern, then check the record to find the best allocation for the job. On the other hand, the available resource will be categorized and organized into different types, such as 3D balance, 3D unbalance, 2D balance, 2D unbalance (toruse or mesh). All these available resource will be updated with real-time system feedback information. When a job is finished, the resource it released can be merge into the neighboring one (if the neighbor is also available) to come up with a bigger chunk. So the 2D resource can become into 3D by merging with newly released ones, so does mesh to torus.

\begin{figure}[h!]     
\centering  
\includegraphics[width=0.48\textwidth]{Pic/design/referencetable}  
\caption{overview of scheduling mechanism}
\label{fig:scheduling mechanism}
\end{figure}

With the feedback from system, our scheduling framework will become knowledgeable about the preference of jobs with different communication pattern. The feedback information keep the batch scheduler updated with the latest status of the system such as the number of available nodes, their layout and geometry,  when hotspot or bottleneck come up in network, etc. The scheduler schedule/dispatch the following jobs adaptively based these information.

Our new scheduling framework is designed based on the knowledge about the precognition of workload. The scheduler works periodically and take 3 steps in each scheduling cycle. The scheduling cycle can be predefined by a fixed time interval, or triggered by events like job completion. The 3 steps within each scheduling cycle are

\begin{itemize}

    \item \textbf{Update resource lists from system feedback}. The resource management module take the feedback information form system and refresh the resource lists every time there is a job completes. The resource released by the job can be merged with any other available resource adjacent to it, in order to form up more compact resources. This is similar to Garbage Collection mechanism in memory management. 1D or 2D resource can become 3D resource and unbalance resource become balance resource after being merged with others. The resource list is well organized when the update is finished, thus, better job scheduling

    \item \textbf{Job prioritization and allocation}. With job's information like size, expected runtime, dominant communication pattern provided upon submission and the well organized resource list maintained by resource management module, our scheduler will make decision about timing and location of each job being dispatched to the system. The scheduling policy will be discussed in detail in \ref{sec:scheduling policy}.

    \item \textbf{Update reservation}. When a job couldn't get its optimal reousrce allocation and any available suboptimal allocation can not compensate the job's performance lose, the job has an option to make reservation for its optimal resource. Therefore, each time a job got dispatched, the scheduler need to update the reservation, delete those that been satisfied and reserve for the following jobs if necessary.

\end{itemize}

\section{Scheduling Policy}
\label{sec:scheduling policy}

We designed and implemented two scheduling algorithms, Rackless and Cautious, for our new scheduler framework. Both of them take advantage of the scheduler's workload precongnition ability and can deliver comparable performance improvement 
as against the traditional schedulers. The comparision results are presented in \ref{sec:scheduling results}.

Before we introduce our scheduling policy, we will first give explanation about workload and system. Suppose the system has 3D torus network and the available resource is consist of a set of 3D or 2D sub-torus chunks, which can be denoted by $\{v\}$, each $v_i = \{x_i, y_i, z_i\}$. The size of each chunk $v_i$, $S_{v_i} = x_i\cdot y_i\cdot z_i $.

The job set in the scheduling window is denoted by $\{J\}$, each $J_i$ has size $S_{J_i}$. Due to the repetitiveness of the workload, $S_{J_i} \in \{n_1, n_2, .., n_k\}$, where $k \leq 5$. 
We can make two reasonable assumption about $\{v\}$ and $\{J\}$.

\begin{equation}
\label{eq:mod0}
\exists S_{v_i} > S_{v_j} \quad then\quad S_{v_i} \bmod S_{v_j} \equiv 0
\end{equation}

\begin{equation}
\label{eq:nofrag}
\min S_{J_i} \leq \min S_{v_i} 
\end{equation}

The reason for both Equation \ref{eq:mod0} and \ref{eq:nofrag} stand is the repetitiveness of workload.

\subsection{Scheduling Policy 1: Reckless}
\label{sec:reckless}

Reckless is designed to be responsive. The overhead for Reckless to make scheduling decision is very low, such that it can be competent when system has high job arrival rate. Reckless first pick a job from the scheduling window and check according the job's size and communication pattern if its optimal resource is availbale. The job get run when enough optimal resource is available. However, if there is no enough optimal resource and only enough sub-optimal availale, Reckless will make reservation if job with ATA communication pattern and run job with non-ATA communication pattern anyway. And if there is no enough resource for any job in the scheduling window, Reckless will hold and skip the current scheduling cycle, wait for more resource got released and make scheduling in the next cycle.

\begin{algorithm}
  \label{alg:scheudling1}
  \caption{Reckless}
  \begin{algorithmic}
    \State \{J\}, Job set in the scheudling window
    \State \{v\}, available system resource sets
    \While{\{J\} is not empty}

        \If{optimal $v_i$ is available for $j_i$}
            \State run $j_i$ on $v_i$;
            \State delete $j_i$ from \{J\};
            \State update reservation;
        \EndIf

        \If{$\exists \bar{v_i}$ satisfy $j_i$ AND $\bar{v_i}$ is not optimal}
            \If{$j_i$ is ATA dominant}
                \State Wait and Reserve $\hat{v_i}$;
            \EndIf
            \If{$j_i$ is not ATA dominant}
                \State run $j_i$ on $\bar{v_i}$;
                \State delete $j_i$ from \{J\};
                \State update reservation;
            \EndIf
        \EndIf

        \If{Available \{$v_i$\} $leq$ $forall$$j_i$ requirement}
            \State hold;
        \EndIf

    \EndWhile
  \end{algorithmic}
\end{algorithm}

\subsection{Scheduling Policy 2: Cautious}
\label{sec:cautious}

Our second scheduling algorithm named Cautious. The difference between Reckless and Cautious is that, Cautious always make evaluation before it assign job to sub-optimal resources. The most intuitive evaluation metric is job's turn-around time. As shown in \ref{alg: Evaluate}, it will compare the runtime for job on the sub-optimal resource, and the time for waiting optimal resource plus the runtime for job on optimal resource. The result will decide whether the job will run on the sub-optimal resource promptly or wait and reserve for its optimal resource.

The other evaluation metrics for Cautious to make decision in terms of system is the Lose of Capacity(LOC). As we shown before, the optimal resource always requires embazzle more network resource, which further leads to system fragmentation, and thus low system utilization. The system administrator is free to choose the evaluation metrics here based on the system and workload features.

\begin{algorithm}
  \label{alg:scheudling2}
  \caption{Cautious}
  \begin{algorithmic}
    \State \{J\}, Job set in the scheudling window
    \State \{v\}, available system resource sets
    \While{\{J\} is not empty}

        \If{optimal $v_i$ is available for $j_i$}
            \State run $j_i$ on $v_i$;
            \State delete $j_i$ from \{J\};
            \State update reservation;
        \EndIf

        \If{$\exists \bar{v_i}$ satisfy $j_i$ AND $\bar{v_i}$ is not optimal}
            \If{Evaluate($j_i$, $\bar{v_i}$)}
                \State run $j_i$ on $\bar{v_i}$;
                \State delete $j_i$ from \{J\};
                \State update reservation;
            \Else
                \State Wait and Reserve $\hat{v_i}$;
                \State Choose from \{J\}-$j_i$ do Backfilling;
            \EndIf
        \EndIf

        \If{Available \{$v_i$\} $\leq$ $\forall$$j_i$ requirement}
            \State hold;
        \EndIf

    \EndWhile

  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \label {alg: Evaluate}
  \caption{Evaluate($j_i$, $\bar{v_i}$)}
  \begin{algorithmic}
    \State Get $\bar{t}$, the time for job $j_i$ to complete on suboptimal resource $\bar{v_i}$;
    \State Get $\hat{t}$, the time for job $j_i$ to complete on optimal resource $\hat{v_i}$;
    \State Get $\delta{t}$, the time job $j_i$ wait for optimal resource $\hat{v_i}$;
    \If{$\delta{t} + \hat{t} \leq \bar{t}$}
        \State return False;
    \Else
        \State return True;
    \EndIf

  \end{algorithmic}
\end{algorithm}

Both Reckless and Cautious has its pros and cons. Reckless is responsive and efficient, but it make scheduling decision lack of thoughts. The job may suffer a long run time when it assign to sub-optimal resource and can lead to potential system fragmentation. Cautious, on the other hand, spends time for evaluation if there is only sub-optimal resource for the job.

\section{Evaluation Methodology}
\label{sec: evaluation methodology}

We conduct a series of experiments using the traces described in Section \ref{sec:job traces} to evaluate our design as against two most frequently used scheduling/allocation policies. Both of them use FCFS/EASY backfilling for job prioritizing. When it comes to job allocation, the first use contiguous/compact allocation strategy (\textbf{CC}), which means the nodes within the allocation it assigned to every job are spacially contiguous and compact. The partition-based allocation strategy used in IBM Blue Gene series of machines are of this kind. The second use non-contiguous allocation strategy(\textbf{NC}), which means it provide jobs with allocation without consideration about the adjacency of nodes. The nodes within each allocation could be scattered all over the system. Machines from Cray Inc. use similar allocation strategy like \textbf{NC}. Both the traditional \textbf{NC} and \textbf{CC} are agnostic about workload. In the rest of the paper, we use \emph{Reckless}, \emph{Cautious}, \emph{CC} and \emph{NC} to denote our scheduling policies and the default ones.

\subsection{Evaluation Metrics} 
\label{sec:metrics}
We use three scheduling metrics for the evaluation of our scheduling framework as against the default ones. They have been chosen for the quantified analysis of both system and job's performance.

\begin{itemize}

  \item \emph{System Utilization Rate}.  System Utilization Rate can be calculated in the following way. Suppose $T$ is the total elapsed time for $J$ jobs, $t_e$  the completion time for job $i$ and $t_s$ be its the start time, and $n_i$ be the size of job $i$, then system utilization rate is calculated as

  \begin{eqnarray}\displaystyle
  \dfrac{\sum_{0\leq i\leq J}{(t_e-t_s) \cdot n_i}}{N \cdot T}
  \end{eqnarray}

  \item \emph{Average Wait Time}. For each job, its wait time refers to the time elapsed between the moment it is submitted and the moment it is dispatched to run. This metric is calculated as the average across all the jobs submitted to the system. The batch scheduler should cut job's wait time down as much as possible.

   \begin{eqnarray}\displaystyle
    T_{wait} = t_{start}-t_{arrive}
  \end{eqnarray}

	\item \emph{Average Job Response time}. Response time refers to the amount of time it take when each job is submitted until it ends, which equals to its wait time plus its run time.

  \begin{eqnarray}\displaystyle
    T_{response} = T_{run}+T_{wait}
  \end{eqnarray}

\end{itemize}

A good batch scheduler should always balance system's performance and user job's performance. On one hand, the batch scheduler should be as responsive as possible, so that user job won't wait too long to get served. On the other hand, the batch scheduler should always guarantee a relative high and stable system uilization so that system resource can be fully utilized.

\subsection{CQSim: Trace-based Scheduling Simulator}
%\textbf{\textcolor{red}{This subsection is Self Plagiarized from Cluster14!!!!}}\\

Simulation is an integral part of our evaluation of various allocation policies
as well as their aggregate effects on system utilization, job's wait time and
response time. We have developed a simulator named CQSim to evaluate our design
at scale. The simulator is written in Python, and is formed by several modules
such as job module, node module, scheduling policy module, etc. Each module is
implemented as a class. The design principles are reusability, extensibility,
and efficiency. The simulator takes job events from the trace, and an event could be
job submission, start, end, and other events. Based on these events, the
simulator emulates job submission, allocation, and execution based on specific
scheduling and allocation policies. CQsim is open source, and is available to the
community \cite{Cqsim}.  \\

\subsection{Job Traces}
\label{sec:job traces}
%\textbf{\textcolor{red}{This subsection is Self Plagiarized from Cluster14!!!!}}\\

In this work, we use two real workload traces collected from production
supercomputers to evaluate our design. The objective of using multiple traces is
to quantify the performance of our design when dealing jobs and systems with
different characteristics. The first trace we used is from a machine named Blue Horizon at
the San Diego Supercomputer Center (denoted as SDSC-BLUE in the paper), which
contains 4,830 jobs. The second trace we used is from a IBM Blue Gene/P system
named Intrepid at Argonne National Laboratory (denoted as ANL-Intrepid in the paper)
\cite{par}. This trace contains 2,612 jobs. Figure \ref{fig:jobsizeintowtraces}
summarizes job size distribution of these traces. ANL-Intrepid is used to
represent capability computing where jobs require a large amount of computing
nodes for solving large-scale problems, whereas SDSC-BLUE is used to represent
capacity computing where the system is utilized to solve a large number of
small-sized problems.

\begin{figure}[h!] 
  \centering
  \includegraphics[width=0.38\textwidth]{Pic/JobInfo/job_dist}
   \caption{Job size distribution of ANL-Intrepid and SDSC-BLUE }
   \label{fig:jobsizeintowtraces}
\end{figure}

\section{Experiment Results}
\label{sec:scheduling results}

\textcolor{red}{all the figs in this section are obsolete!!!}

\begin{figure}[h!] 
  \centering
  \includegraphics[width=0.48\textwidth]{Pic/experiment/summary/utilization}
  \caption{System Utilization got from CC, NC, Greedy, Cautious respectively}
  \label{fig: uti}
\end{figure}

\begin{figure}[h!] 
  \centering
  \includegraphics[width=0.48\textwidth]{Pic/experiment/summary/awt-vs-cc}
  \caption{Jobs' average wait time reduction got by comparing CC as against our design.}
  \label{fig: uti}
\end{figure}

\begin{figure}[h!] 
  \centering
  \includegraphics[width=0.48\textwidth]{Pic/experiment/summary/art-vs-cc}
  \caption{Jobs' average response time reduction got by comparing CC as against our design.}
  \label{fig: uti}
\end{figure}

\begin{figure}[h!] 
  \centering
  \includegraphics[width=0.48\textwidth]{Pic/experiment/summary/awt-vs-nc}
  \caption{Jobs' average wait time reduction got by comparing NC as against our design.}
  \label{fig: uti}
\end{figure}

\begin{figure}[h!] 
  \centering
  \includegraphics[width=0.48\textwidth]{Pic/experiment/summary/art-vs-nc}
  \caption{Jobs' average response time reduction got by comparing NC as against our design.}
  \label{fig: uti}
\end{figure}

\section{Related Work}
\label{sec:related_work}
Batch scheduling on HPC system has been a hot topic getting constant discussion since the 1990's.
Feitelson et al proposed the most widely used job scheduling policy which is First Come, First Serve (FCFS) combined with 
EASY backfilling\cite{feit}. FCFS/EASY backfilling policy greatly improves HPC system's utilization and cuts the wait time 
of first job in the waiting queue. Since then, many studies seek to refine this classic scheduling paradigm. Tang et al. 
tried to improve the accuracy of user's estimated job runtime and walltime so as to enhance the efficiency of backfilling and  minimize the system fragmentation\cite{wei-ipdps2010} \cite{wei-jpdc2013} \cite{wei-ipdps2011}. % And there are some other variation of FCFS/EASY backfilling proposed to optimize system performance in terms of power consumption and energy cost \cite{zhou}\cite{yang-sc13}.

Only with the information about job size and estimated run time seems to be insufficient for batch scheduler do make the 
optimal scheduling/allocaiton decision. Some work proposed that the batch scheduler should be aware of some information 
about system network information in order to fully utilize system resources.Leung et al. presented allocation strategies 
based on space filling curves and one dimensional packing \cite{leung}. The problem with using space filling curve is that  
it can only be applied to system with the scale of $2^n$ nodes in each dimension, which is a ideal case that not existing 
in morden HPC systems. Albing et al. conducted study about the allocation strategies that the Cray
Application Level Placement Scheduler (ALPS) used \cite{carl-cug}. The job allocation in Cray Linux Environment (CLE) 
operating system is managed by ALPS, which simply works off an ordered node list, however that is ordered.

Based on their 
work, Yang et al. presented a window-based locality-aware job scheduling design for torus-connected system
\cite{yang-cluster14}. In this new scheduling framework, several jobs are taken into  consideration at the same time for job 
prioritizing and resource allocation. A list of slots is maintained to preserve node contiguity information for
resource allocation. Each job been assigned into one of these slot so that the locality can be mostly preserved.

Rosenthal et al. found there is limited benefit for many applications by
increasing network bandwidth \cite{rosenthal}. They employed a subset of the
CORAL mini-applications that represent U.S. Department of Energy workload and
leverage multirail networkings to evaluate the improvement of these applications
caused by increased network bandwidth. The applications that send mostly small
messages or larger messages asynchronously are not bandwidth bounded, hence,
benefit only slightly or not at all from increased bandwidth.

Hoefler et al. propose to use performance modeling techniques to analysis 
factors that impact the performance of parallel scientific applications 
\cite{hoefler-modeling}. However, as the scale of HPC continue grows, the 
interference of concurrently running jobs is getting worse, which is hard to be 
quantified by performance profiling tools.

Bogdan et al provide a set of guidelines how to configure a network with 
Dragonfly topology for workload with Nearest Neighbor communication 
pattern\cite{Bogdan-hpdc14}. They first derived a theoretical model of Nearest 
Neighbor communication performance on Dragonfly network that can predict network 
bottleneck. Then, they use simulation to validate the correctness of their 
theoretical model.

Dong et al describe IBM Blue Gene/Q's 5D torus interconnect
network\cite{Dong-SC11}. They developed simple benchmarks that conforms to four
different communication patterns, namely ping-pong, nearest neighbor, broadcast
and allredcue, to demonstrate the effectiveness of this highly parallel 5D torus
network.

\section{Conclusions}
\label{sec:conclusion}

\section*{Acknowledgment}
\label{sec: ack}
%We appreciate the valuable comments and suggestions from the anonymous reviewers. 
The work at Illinois Institute of Technology is supported in part by
US National Science Foundation grant CNS-1320125. The work at Argonne is
supported in part by the U.S. Department of Energy (DOE), Office of Science,
under Contract DE-AC02-06CH11357.

\begin{thebibliography}{10}
 %=====start of scheduling papers===== 
\bibitem{feit}
D.~Feitelson and A.~Weil.
\newblock Utilization and predictability in scheduling the {IBM} {SP2} with
  backfilling.
\newblock In {\em International Parallel and Distributed Processing Symposium}, 1998.

\bibitem{wei-ipdps2010}
W.~Tang, N.~Desai, D.~Buettner, and Z.~Lan.
\newblock Analyzing and adjusting user runtime estimates to improve job
  scheduling on the {Blue} {Gene/P}.
\newblock In {\em 2010 IEEE International Symposium on Parallel Distributed
  Processing}, 2010.

\bibitem{wei-ipdps2011}
W.~Tang, Z.~Lan, N.~Desai, D.~Buettner, and Y.~Yu.
\newblock Reducing fragmentation on torus-connected supercomputers.
\newblock In {\em 2011 IEEE International Symposium on Parallel Distributed
  Processing Symposium}, 2011.

\bibitem{wei-jpdc2013}
W.~Tang, N.~Desai, D.~Buettner, and Z.~Lan. 
\newblock Job Scheduling With Adjusted Runtime Estimates on Production Supercomputers.
\newblock Journal of Parallel and Distributed Computing (JPDC), 73(7):926-938, 2013.

\bibitem{zhou}
Z.~Zhou, Z.~Lan, W.~Tang, and N.~Desai.
\newblock Reducing energy costs for {IBM Blue} {Gene/P} via {Power-Aware} job
  scheduling.
\newblock In {\em 17th Workshop on Job Scheduling Strategies for Parallel
  Processing}, 2013.

\bibitem{yang-sc13}
X. ~Yang, Z.~Zhou, S.~Wallace Z.~Lan, W.~Tang, S. ~Coghlan and Mike. ~Papka.
\newblock Integrating Dynamic Pricing of Electricity into Energy Aware Scheduling for HPC Systems. 
\newblock In {Proceedings of the International Conference on High Performance Computing, Networking, Storage and Analysis(SC), 2013 International Conference for}
\newblock SC '13,
\newblock 2013,
\newblock Salt Lake City, Utah

\bibitem{Cqsim}
{\textcolor{red} {Cqsim: An event-driven simulator}}.
\newblock http://bluesky.cs.iit.edu/cqsim

%=====start of allocation aware papers===== 
\bibitem{leung}
Leung, V.J.; Arkin, E.M.; Bender, M.A.; Bunde, D.; Johnston, J.; 
Alok Lal; Mitchell, J.S.B.; Phillips, C.; Seiden, S.S., 
\newblock Processor allocation on Cplant: achieving general processor locality using one-dimensional allocation strategies
\newblock In {Proceedings of 2002 IEEE International Conference on Cluster Computing, 2002}.  pages 296--304, 2002

\bibitem{carl-cug}
Carl Albing and Mark Baker
\newblock ALPS,Topology, and Performance: A Comparison of Linear Orderings for 
Application Placement in a 3D torus.
\newblock Presented at the CUG 2010, Edinburgh, Scotland, UK, 2010.

\bibitem{yang-cluster14}
Xu Yang and Zhou Zhou and Wei Tang and Xingwu Zheng and Jia Wang and Zhiling Lan.
\newblock Cluster Computing (CLUSTER), 2014 IEEE International Conference on,
\newblock Balancing job performance with system performance via locality-aware scheduling on torus-connected systems
\newblock Sept, 2014,
\newblock pp.140-148

%+++++++++++++++++++++++
\bibitem{mpip}
\newblock mpiP: Lightweight, Scalable MPI Profiling.
\newblock http://mpip.sourceforge.net, 
\newblock 2013.

\bibitem{tau}
S. Shende and A. Malony. 
\newblock The TAU parallel performance system. 
\newblock International Journal of High Performance Computing Applications, 20(2):287–311, 2006.

\bibitem{scala}
X. Wu, F. Mueller, S. Pakin, 
\newblock "Automatic Generation of Executable Communication Specifications from Parallel Applications", 
\newblock In {Proceedings of the International Conference on High Performance Computing, Networking, Storage and Analysis(SC), 2011 International Conference for}
\newblock SC '11,
\newblock 2011,pages 12-21.

%+++++++++++++++++

%=====start of network papers===== 
\bibitem{rosenthal}
E. Rosenthal and E.A. Leon
\newblock Characterizing Application Sensitivity to Network Performance.
\newblock  In {Proceedings of SC14: International
Conference for High Performance Computing, Networking, Storage and Analysis},
2014

\bibitem{hoefler-modeling}
T. Hoefler and W. Gropp and M. Snir and W. Kramer
\newblock Performance Modeling for Systematic Performance Tuning,
\newblock 2011
\newblock Nov.
\newblock In {Proceedings of International Conference for High Performance Computing, Networking, Storage and Analysis (SC'11), SotP Session,
}

\bibitem{roth}
P. Roth and J. Meredith and J. Vetter,
\newblock "Automated Characterization of Parallel Application Communication Patterns",
In Proceedings of the 25rd International Symposium on High-performance Parallel and Distributed Computing
\newblock HPDC '15,
\newblock 2015,

%=====end of application/mapping papers===== 
\bibitem{Pjesivac}
Pje\v{s}ivac-Grbovi\'{c}, Jelena and Angskun, Thara and Bosilca, George and Fagg, Graham E. and Gabriel, Edgar and Dongarra, Jack J.
\newblock Performance Analysis of MPI Collective Operations,
\newblock Cluster Computing,
\newblock June 2007,
\newblock volume 10,
\newblock numpages 17,

\bibitem{thakur}
Thakur, Rajeev and Gropp, William D.
\newblock Improving the performance of collective operations in MPICH,
\newblock Recent Advances in Parallel Virtual Machine and Message Passing Interface,
\newblock year 2003,
\newblock Springer

\bibitem{abhinav-ipdps15}
Abhinav Bhatele, Andrew R. Titus, Jayaraman J. Thiagarajan, Nikhil Jain, Todd Gamblin,
Peer-Timo Bremer, Martin Schulz† and Laxmikant V. Kale,
\newblock "Identifying the Culprits behind Network Congestion",
\newblock In {\em 2015 IEEE International Symposium on Parallel Distributed Processing Symposium}, 2015.

\bibitem{langer}
S. Langer, A. Bhatele, and C. H. Still,
\newblock “pF3D simulations of laser-plasma interactions in National Ignition Facility experiments,” 
\newblock Computing in Science and Engineering, vol. 99, Aug. 2014, lLNL- JRNL-648736. 
\newblock [Online]. Available: http://doi.ieeecomputersociety.org/ 10.1109/MCSE.2014.79

%\bibitem{abhatele-ipdps08}
%A. Bhatele, L.V.  Kale
%\newblock Application-specific topology-aware mapping for three dimensional topologies.
%\newblock In {\em Proceedings of Workshop on Large-Scale Parallel Processing (held as part of IPDPS' 08)} 2008
%
%\bibitem{abhinav-sc13} 
%Bhatele, Abhinav and Mohror, Kathryn and Langer, Steven
%H. and Isaacs, Katherine E. 
%\newblock There Goes the Neighborhood: Performance Degradation Due to Nearby Jobs 
%\newblock In {Proceedings of SC13: International
%Conference for High Performance Computing, Networking, Storage and Analysis},
%2013

\bibitem{hpdc14}
Prisacari, Bogdan and Rodriguez, German and Heidelberger, Philip and Chen, Dong and Minkenberg, Cyriel and Hoefler, Torsten
\newblock Efficient Task Placement and Routing of Nearest Neighbor Exchanges in Dragonfly Networks
\newblock In Proceedings of the 23rd International Symposium on High-performance Parallel and Distributed Computing
\newblock HPDC '14,
\newblock 2014,
\newblock Vancouver, BC, Canada

\bibitem{jingjin}
Wu, Jingjin and Lan, Zhiling and Xiong, Xuanxing and Gnedin, Nickolay Y. and Kravtsov, Andrey V.
\newblock Hierarchical Task Mapping of Cell-based AMR Cosmology Simulations
\newblock Proceedings of the International Conference on High Performance Computing, Networking, Storage and Analysis
\newblock SC '12,
\newblock 2012,
\newblock Salt Lake City, Utah

\bibitem{zhou-ipdps}
Z. Zhou, X. Yang, Z. Lan, P. Rich, W. Tang, V. Morozov, and N. Desai.
\newblock Improving Batch Scheduling on Blue Gene/Q by Relaxing 5D Torus Network Allocation Constraints,
\newblock Proceedings of 29th IEEE International Parallel \& Distributed Processing 
Symposium,
\newblock IPDPS'15, 2015.
\newblock Hyderabad, INDIA

\bibitem{Bogdan-hpdc14}
Prisacari, Bogdan and Rodriguez, German and Heidelberger, Philip and Chen, Dong and Minkenberg, Cyriel and Hoefler, Torsten
\newblock Efficient Task Placement and Routing of Nearest Neighbor Exchanges in Dragonfly Networks,
\newblock Proceedings of the 23rd International Symposium on High-performance Parallel and Distributed Computing,
\newblock HPDC '14,
\newblock 2014,
\newblock  Vancouver, BC, Canada,

\bibitem{Dong-SC11}
Dong Chen; Eisley, N.A.; Heidelberger, P.; Senger, R.M.; Sugawara, Y.; Kumar,
S.; Salapura, V.; Satterfield, D.L.; Steinmacher-Burow, B.; Parker, J.J., 
\newblock "The IBM Blue Gene/Q interconnection network and message unit," 
\newblock High Performance Computing, Networking, Storage and Analysis (SC), 2011 International 
Conference,
\newblock pp.1,10, 
\newblock 12-18 Nov. 2011

\bibitem{Jason-2011}
Jason Cope, Ning Liu, Sam Lang, Phil Carns, Chris Carothers, Robert Ross.
\newblock Proceedings of the Workshop on Emerging Supercomputing Technologies 2011,
\newblock 2011

%===simulation papers =====

\bibitem{ross}
C. Carothers, D. Bauer, and S. Pearce.
\newblock ROSS: A High-Performance, Low memory, Modular Time Warp System. 
\newblock Journal of Parallel and Distributed Computing,
\newblock (62):1648–1669, 2002.

\bibitem{mubarak-sc2012}
Mubarak, M.; Carothers, C.D.; Ross, R.; Carns, P.,
\newblock "Modeling a Million-Node Dragonfly Network Using Massively Parallel Discrete-Event Simulation," 
\newblock High Performance Computing, Networking, Storage and Analysis (SCC), 
\newblock 2012 SC Companion: 
\newblock pp.366,376, 10-16 Nov. 2012

\end{thebibliography}

\end{document}

### 2015-09-15 submission.tex
add figure folder

@@
 \ifCLASSINFOpdf
-
 \else
@@
 % can use linebreaks \\ within to get better formatting as desired
-%\title{Network contention aware batch Scheduling on Torus-Connected Systems}
-\title{ \textcolor{red}{Deja Vu: A New Topology-aware HPC Batch Scheduler with Workload Precognition} }
+\title{Intelligent HPC Job Scheduling with Communication Pattern Awareness}
 % author names and affiliations
@@
 \IEEEauthorblockN{Xu Yang\IEEEauthorrefmark{1}, John Jenkins\IEEEauthorrefmark{2}, Misbah Mubarak\IEEEauthorrefmark{2}, Robert B. Ross\IEEEauthorrefmark{2}, Zhiling Lan\IEEEauthorrefmark{1}}
+
 \IEEEauthorblockA{\IEEEauthorrefmark{1}Department of Computer Science,
@@
 It has been widely recoginzed that network contension between concurrently running jobs in HPC system is one of the primary causes of performance variability.
-Optimizing job allocation and avoiding network sharing have been proved to be effective for alleviating network contension.
-In this work, we propose a new job scheduler framework for HPC system, which not only take job's size and expected runtime as primary submission parameters, but also an extra one that indicate job's dominate communication pattern(CP).
-Using this extra parameter CP can help our new scheduler framework enhance both job and system's performance in three ways.
-First, communication pattern indicates whether the job's performance rely heavily on the network or not.
-Second, with the precongnition of job's dominate communication pattern, the scheduler can provide job with better resource allocation.
-Third, when multiple jobs are dispatched to the system, knowing their communication pattern can help avoid network contension.
-Our extensive experimental results show that the new scheduling framework can greatly improve both system and user's job performance.
+Optimizing job allocation and avoiding network sharing have been proved to be effective for alleviating network contension and performance degradation.
+In this work, we propose a new job communication pattern aware allocation strategy which serve as a critical module in our future HPC batch scheduler.
+The novelty of our new strategy is that it makes allocation decision with not only system topology and resource availability information, but also with job's communication pattern information.
+The information scheduler gets about job communication pattern can help it to decide whether the job’s performance relies heavily on the network or not, and what the potential inter/intra-job interference would be.
+Thus, the scheduler can pick the preferable resource allocation based on specific job's communication pattern and preserve the locality of allocation in a better way.
+Using traces collected from three typical parallel applications of DOE, we validate the effectiveness of our new desgin.
+The idea presented in this paper is widely applicable: HPC parallel application has specific dominant communication pattern, allocating resource with communication pattern awareness can preserve locality in a bettwer way on systems with differnt network topologies.
 
@@
 
-The batch scheduler has always been a undervalued component in HPC system.
-As an interface between user and system, the batch scheduler usually do nothing more than simply take the submitted job from user and hand over to the system for execution.
-User are required to provide information about system resource needed by his job, such as job size (i.e.
-how many nodes are needed to run the job) and job's expected runtime (i.e.
-how long does user expected his job take to complete) upon job's submission.
-The batch scheduler keeps updated about system resource availability, when it being informed there are enough number of nodes becomes available, the scheduler picks up a job from the job queue and dispatch it to the system.
+The scale of supercomputer keeps growing from petascale to exascale, to accomodate the the demand of insatiable computing power from scientific research areas.
+Production systems contain hundreds of thousands of processors serve as irreplaceable research vehicle for scientific problems with increasing size and complexity.
+Supercomputers employed in a shared way to accomodate many jobs running concurrently\ref{fig: mira activity} , which will also improve system utilization.
+These jobs share the system infrastructure such as network, I/O bandwidth, which will inevitably cause contention over the shared resource.
+As supercomputer continues to evolve, these shared resources are tend to become the bottleneck for performance.
+
+\begin{figure}[h!] 
+  \centering
+  \includegraphics[width=0.38\textwidth]{figs/mira}
+   \caption{The activity of Mira, IBM BlueGene/Q system at Argonne National Lab. Many jobs runninng concurrently on the system. Different jobs represented by specific colors. }
+   \label{fig: mira activity}
+\end{figure}
+
+One of the most prominant problem comes into researcher's attention is the network contention between concurrently running jobs.
+The network sharing between concurrently running jobs cause communication variability, which reslut in jobs running slower and longer waiting time for results.
+The delay of currently running jobs will increase the queueing time of the following submitted jobs, thus leads to low system throughput, low utilization and high energy cost.
+The network sharing can even cause interference between neighbouring jobs, abort the the running job unexpectedly, or even worse, wrong science results.
+
+This adverse effect of network sharing can be mitigated by providing the job with better node allocation.
+The batch scheduler is responsible for assigning job with the amount of node it required upon submission.
+Currently two scheduling strategies are commonly used on torus-connected systems.
+One is so called partition based systems, where the scheduler assigns each user job a compact and contiguous set of computing nodes.
+IBM Blue Gene series systems fall into this category.
+This strategy is in favor of application’s performance by providing it with isolated partition and exclusive network connections, thus reducing network contention caused by concurrently running jobs sharing network bandwidth.
+However, this strategy can cause internal fragmentation (when more nodes are allocated to a job than it requests) and external fragmentation(when sufficient nodes are available for a request, but they can not be allocated contiguously), therefore leading to poor system performance (e.g., low system utilization and long waiting time for job).
+The other is non-contiguous allocation system, where free nodes are assigned to user job no matter whether they are contiguous or not.
+Cray XT/XE series systems fall into this category.
+Non-contiguous allocation eliminates internal and external fragmentation as seen in partition-based systems, thereby leading to high system utilization.
+Nevertheless, it introduces other problems such as scattering application processes all over the system and causing both inter-job contention and intra-job contention.
+The non-contiguous node allocation can make inter-process communication less efficient and cause network contention among concurrently running jobs, thereby resulting in poor job performance especially for those communication-intensive jobs.
+
+Partition-based allocation achieves good job performance by sacrificing system performance (e.g., poor system utiliza- tion), whereas non-contiguous allocation can result in better system performance but could severely degrade performance of user jobs (e.g, prolonged wait-time and run-time).
+Supercomputers keep growing in size, its network become more sophisticated then ever before and applications also increase its complexity.
+How to provide application with better allocation that can balance job performance with system performance become a prominant problem need a solution before HPC enters era of exasacle computing.
+In this work, we propose a guideline for future batch scheduling system, a smart way to make allocation with the awareness of job's communication pattern.
 
 Most existing batch schedulers take so few initiative because it been hold under the obscurantism policy.
-There are two observations about the ignored information that batch scheduler should have been informed.
-First, the workload on HPC system tend to be repetitive.
-We analyzed the workload on Mira, an IBM Blue/Gene Q system installed in Argonne National Lab, found that most jobs require 2k nodes and about 14 particular jobs consist of the majority its workload.
+Whether it use partition-based or non-contigous policy, means it chooses the benefit of either system or job over the other.
+Some propose compromised policies like buddy system\cite{buddy system paper} still make the allocation for job blindly without knowing job's communication information.
+
+Although HPC applications complexity continue to grow, people could get more knowledge about the applications communication information with the help from various parallel application profiling tools, such as TAU\cite{tau}, MPIP\cite{mpip} and DUMPI\cite{sst}.
+Job's communication trace and performance analysis can be get by using such tool at runtime.
+
+The off-line study of job's communication traces could be helpful to develop better allocation policies for batch scheduler to make on-line scheduling/allocating decisions.
+This is because the communication pattern of parallel applications on HPC system can be summarized into a couple of categories\cite{hpdc2015-ornl}.
+The deep understanding of one application can be helpful to all other applications that conforms to the same communication pattern.
+Job's communication traces study would be extreme helpful to leadership computing facility too.
+The set of applications running on those leadership computing facilities is usually stable, and the workloads from those HPC systems is usually has great repetitiveness.
+We analyzed the workload on Mira, an IBM Blue/Gene Q system at Argonne Leadership Computing Facility, found that most jobs require 2k nodes and about 14 particular jobs consist of the majority its workload\ref{fig: repetitiveness of Mira}.
 In other word, Mira accommodates mostly some particular 14 jobs and most jobs have the same size.
-The second thing that batch scheduler should know is job's communication pattern.
-When a user submit his job, he at least has a vague idea about the most dominant communication pattern of his job (i.e.
-All-to-All, Nearest Neighbor, Collective).
-
-In this work, we propose a new HPC batch scheduler framework, which can capture the workload signature introduced by repetitively submitted jobs and  integrates network simulation feedback control.
-In spite of the input parameters such as job size and expected runtime that traditional batch scheduler requires, our new framework also takes job's communication pattern (CP) as an extra parameter, which can be utilized when assigning job with its preferred allocation.
-Being aware of job's CP has several adavantages for improving both job and system performance.
-For example, CP indicates the better allocation for the job and thus, bath scheduler can get the job prefered allocation without embezzling extra system resources.
-
-Unlike the traditional batch scheduler that is agnostic about upcoming jobs in the workload, our new framework is aware of the repetitiveness of HPC workload and take advantage of this workload feature for making the optimal scheduling/allocation decision for upcoming jobs.
-To endow the new framework with this ability, we implement a memory mechanism that keep record of the preference about allocation of jobs with different CPs.
-
-To receive the feedback about resource availability from system, the new framework will keep real time records about the both network and node availability.
-The new framework will be informed with information such as how many nodes are available, layout and geometry of those available nodes  (contiguous or non-contiguous, ), and their network resources (connnected by torus or mesh).
-Our new framework will also collect certain information for each job when the job running on some its allocated resources, such as job's actual running time, network resource utilization, to determine whether its allocated resources is the optimal regarding to its communication pattern.
-These feedback information from the runtime system will be utilized for job's furture repetitive submission.
-
-With more detailed information (i.e.
-job size, expected runtime, CP) about job provided by user, combined with feedback about job's previous execution and resource availability from system, the new scheduler framework can make better scheduling/allocation decisions.
-And with the new memory mechanism, our new framework is equiped with self-learning capability and the more jobs it handled, the more talented it becomes.
-
-The major contributions of this paper:
+
+In this work, we present a guideline for future batch scheduler to make allocation with awareness about job's communication pattern.
+Rather than choose blindly between partition-base or non-contigous allocation, the new allocation mechanism make allocation based on job's communication pattern.
+With detail analysis about application's communication pattern, we can get a clear view of its communication topology graph, data transfer between processes, and decide which processes are tightly connected to which and conduct intensive communication.
+The "local community" consists of such tightly coupled processes shuld get compact alloction.
+Then, the allocation mechanism should make allocations that only needs to guarantee the locality of the subset of the nodes where those "local community" should reside on.
+
+We first pick three signature applications from DOE Design Forward Project\cite{design forward webpage} as examples to conduct deep study about their communication patterns.
+We identify the "local community" in these applications and assign them allocation accrodingly.
+The advantage of making allocation is obvious.
+First, compared with partition-based policy, our communication pattern aware policy is more flexible, it doesn't require the system to provide a big partition that can accomodate the whole application, just small set of compact nodes enough for all the "local community" in the application.
+On the other hand, our communication pattern aware allocation policy won't destroy the locality of application's communication pattern.
+On the contrary, it provides small compact node sest for the the "local communities" in the application will preserve the communication locality in the maximum extent.
+
+We use a sophisticated simulation toolkit named CODES, Co-Design of Multi-layer Exascale Storage Architecture, from Argonne National Lab as research vehicle to evaluate the performance of our communication pattern aware allocation policy on torus network.
+CODES enables the exploration of simulating different HPC networks with high fidelity\cite{Jason-2011}.
+We extend CODES with a Job Mapping API that support different allocation policies on torus network.
+
+We also conduct a case study with the results we got from communication pattern aware allocation policy evaluation with real trace from production system.\textcolor{red}{Not Sure About this part!!!}
+
+The rest of this paper orgainzed as follows.
+Section\ref{sec:application study} gives a detailed study of the three applications from DOE Desgin Forward Project.
+Section\ref{sec:simulation} talks about the simulation platform and Section
+
+\section{Application Study}
+\label{sec:application study}
+
+A parallel application usuall conforms to a combination of several basic communication patterns.
+At its different execution phases, the application's communication behavior may follow different basic patterns repectively.
+When we look into the data flow during the application execution, most parallel application start with broadcast operation to distribute the data from root process to other processes, followed by a series of computation and communication that comforms to certain pattern, which is usually the dominant part of application's execution.
+And before the application come to completion, all the working process will return their results to the root dirctly or hierarchically.
+In this work, we focus on the dominant communication pattern because that's where the application spend most of its running time.
+We believe making optimized allocation regarding application's dominant communication pattern would greatly benefit both the application and the system.
+In this section, we analysis two applications whose dominant communication patterns are most prevalent in scientific computation.
+
+There are many profiling tools available to capture the communication behavior from parallel applications, such as Tuning and Analysis Utilities(TAU)\cite{tau}, mpiP\cite{mpip}.
+They can help analysis parallel application, providing information like the percentage of different MPI operations of the applications, communication topology, the amount data transferred between processes.
+
+In this work, we provide detailed analysis about the communication pattern of two DOE MiniApps.
+MiniApps are reduced proxy applications that encapsulate the salient performance of larger full size applications\cite{miniapp}.
+Since we only focus on the dominant communication pattern of parallel applications, these MiniApps are perfect candidates for analysis the application's performance on different allocations or on new network architectures.
+
+\subsection{AMG}
+\label{sec:amg}
+AMG is the MiniApp of Algebraic Multigrid Solver, which is a parallel algebraic multigrid solver for linear systems arising from problems on unstructured mesh physics packages.
+It has been derived directly from the BoomerAMG solver that is being developed in the Center for Applied Scientific Computing (CASC) at LLNL\cite{amg}.
+The dominant communication pattern is this regional communication with decreasing message size for different parts of the multigrid v-cycle.
+
+\begin{figure}[h!] 
+  \centering
+  \includegraphics[width=0.38\textwidth]{figs/amg/amg_ct}
+   \caption{Communication Topology of AMG with 27 MPI Ranks }
+   \label{fig: amg communication topology}
+\end{figure}
+
+\begin{figure}[h!] 
+  \centering
+  \includegraphics[width=0.38\textwidth]{figs/amg/amg_data_transfer}
+   \caption{The amount of data each rank transferred in AMG. }
+   \label{fig: amg data trans}
+\end{figure}
+
+Figure \ref{fig: amg communication topology} shows the communication topology of a small scale AMG with 27 MPI Ranks.
+Here we prefer to show the communication topology with small scale version of the MiniApp because the dominant communication pattern of the application doesn't change with its scale.
+We can make the observation from the upper-left of the figure that each rank in AMG has intense communication between 3 other ranks, such as rank 0 between rank 1, 3 and 9, rank 1 between 2, 4 and 10, etc.
+And this relation is also symmetric along the main diagnal.
+The dominant communication pattern is quite obvious when we identify this relation between ranks.
+
+Figure \ref{fig: amg data trans} shows the amount of data each rank transferred in AMG.
+The blue line shows that total data amount transferred from each rank, while the red shows the amount of received data and green shows the amount of send data.
+The first thing we can spot from the figure is that for each rank, the amount of data sent and received are basically the same, which also well explain the symmetry of the application's dominant communication pattern.
+And the second thing we can find about AMG is that the data transfer amount also has a symmetric pattern.
+Rank 0 has the same amout of data transfer as Rank 26, Rank 1 as of Rank 25, etc.
+
+\subsection{Crystal Router}
+\label{sec:crystalrouter}
+
+The second MiniApp we have is Crystal Router, which is the extracted communication kernel of the full application Nek5000.
+Nek5000\cite{crystalrouter} is a spectral element CFD application developed at Argonne National Laboratory.
+It features spectral element multigrid solvers coupled with a highly scalable, parallel coarse-grid solver that widely used for projects including ocean current modeling, thermal hydraulics of reactor cores, and spatiotemporal chaos.
+The MiniApp of Nek5000, Crystal Router demonstrates the "many-to-many" communication pattern through scalable multi-stage communication process.
+
+\begin{figure}[h!] 
+  \centering
+  \includegraphics[width=0.38\textwidth]{figs/cr/cr10_ct}
+   \caption{Communication Topology of Crystal Router with 10 MPI Ranks }
+   \label{fig: crystalrouter10 communication topology}
+\end{figure}
+
+\begin{figure}[h!] 
+  \centering
+  \includegraphics[width=0.38\textwidth]{figs/cr/cr100_ct}
+   \caption{Communication Topology of Crystal Router with 100 MPI Ranks }
+   \label{fig: crystalrouter100 communication topology}
+\end{figure}
+
+\begin{figure}[h!] 
+  \centering
+  \includegraphics[width=0.38\textwidth]{figs/cr/cr_data_transfer}
+   \caption{The amount of data each rank transferred in Crystal Router. }
+   \label{fig: cr data trans}
+\end{figure}
+
+Here, we provide the communication topology graph of Crystal Router with 10-rank and 100-rank respectively.
+As we can see from figure \ref{fig: crystalrouter10 communication topology}, every 5-rank is a compact community, where data transfer is very intensive.
+Beside the "5-rank community", there is "pair-wise" global data transfer in Crystal Router.
+Obviously, there is data transfer alone the diagnal.
+When the scale of Crystal Router increase to 100, the "5-rank community" is still the dominant pattern in a multi-stage manner.
+The same pattern can be observed from the global data transfer.
+Our observation about the dominant communication pattern of Crystal Router is the multi-stage "5-rank community" and "pair-wise" global data transfer happen recursively as the scale of the application increase.
+
+Figure \ref{fig: cr data trans} shows the amount of data each rank transferred in Crystal Router.
+The locality pattern of "5-rank community" can still be spotted in this figure.
+The spiky shape occurs in the frequency of every 5 rank.
+Unlike AMG, the there is variance between the amount of data send and received for each rank.
+This is due to the specific feature of Crystal Router's communication pattern.
+
+\subsection{MultiGrid}
+\label{sec:multigrid}
+
+Differe from AMG, MultiGrid is geometric multigrid v-cycle from production elliptic solver BoxLib, a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes.
+MultiGrid conforms to 3D lattice communication pattern with decreasing message size and collectives for different parts of the multigrid v-cycle.
+It is widely used for structured grid physics packages.
+
+\begin{figure}[h!] 
+  \centering
+  \includegraphics[width=0.38\textwidth]{figs/mg/mg125_ct}
+   \caption{Communication Topology of MultiGrid with 125 MPI Ranks }
+   \label{fig: multigrid125 communication topology}
+\end{figure}
+
+\begin{figure}[h!] 
+  \centering
+  \includegraphics[width=0.38\textwidth]{figs/mg/mg_data_transfer}
+   \caption{The amount of data each rank transferred in MultiGrid. }
+   \label{fig: cr data trans}
+\end{figure}
+
+\section{Simulation Platform}
+\label{sec:simulation}
+
+A simulation toolkit named CODES (Co-Design of Multi-layer Exascale Storage Architecture) from Argonne National Lab enables the exploration of simulating different HPC networks with high fidelity\cite{Jason-2011}\cite{mubarak-sc2012}.
+CODES is built on top of Rensselaer Optimistic Simulation System (ROSS) parallel discrete-event simulator, which is capable of processing billions of events per second on leadership-class supercomputers\cite{ross}.
+CODES support both torus and dragonfly network with high fidelity flit-level simulation.
+In this work, we only use torus network to show the impact of node allocation to job with different dominant communication patterns.
+CODES has this network workload component that capable of conducting trace-driven simulation.
+It can take real MPI application trace generated by SST DUMPI\cite{sst} to drive CODES network models.
+
+Torus networks have been extensively used in the current generation of supercomputers because of their linear scaling on per-node cost and competitive communication performance.
+The topology of torus network is k-ary n-cube, with $k^n$ nodes in total arranged in an n-dimensional grid having k nodes in each dimension.
+Each node has 2$\times$ n direct linked neighbor nodes.
+
+For a job requires $n$ nodes submitted to non-partition based 3D torus-connected system, it gets an allocation of various possible shapes, which could be 3D balanced-cube, 3D unbalanced-cube, 2D mesh, or even 1D list.
+Figure \ref{fig: allocation} shows some possible allocations for a 64-node job.
+3D balanced-cube allocation (in Figure \ref{fig: allocation} as red) is $\{4,4,4\}$, 3D unbalanced-cube allocation is $\{8,4,2\}$(in Figure \ref{fig: allocation} as green), 2D mesh allocation is $\{8,8,1\}$ (in Figure \ref{fig: allocation} as blue).
+
+\begin{figure}[h!] 
+  \centering
+  \includegraphics[width=0.48\textwidth]{Pic/allocationshape/allocation}
+  \caption{64-node job with different allocation shapes. Red, Green and Blue represents 3D balanced, 3D unbalanced and 2D allocation respectively. }
+  \label{fig: allocation}
+\end{figure}
+
+We implemented a Job Mapping API for CODES, which can provide flexible job allocation strategies.
+The new API can help us to explore the impact of different allocation strategies to job with specific communication pattern.
+We have choosen two publicly available HPC network traces provided by the Design Forward Program\cite{design forward webpage}.
+The detailed analysis of these two traces are present in Section\ref{sec:amg} and \ref{sec:crystalrouter}.
+
+\subsection{Allocation Simulation}
+\label{sec:alloc sim}
+
+In this section, we analysis the impact of different allocation strategies on two real applications, AMG and Grystal Router.
+The torus network performance is determined by its dimensionality and link bandwidth.
+We evaluate the data transfer time of AMG and Crystal Router on torus network with different dimensionality and link bandwidth configurations in section \ref{sec: dimensionality study}.
+Then, we will show applications performance on 3D torus network with different allocation strategies in section \ref{sec: alloc strategy study}.
+
+\subsection{Torus Dimensionality}
+\label{sec: dimensionality study}
+
+As the increase of the dimensionality of torus netowrk, so does the number of links connected with each node.
+The increased aggregated bandwidth of each node will definitely reduce the data transfer time of each rank in the application.
+Figure xxx shows the performance of the AMG and CrystalRouter on a 2K node torus network model with a 3D torus (16x16x8), a 5D torus (8x4x4x4x4), and a 7D torus (4x4x4x4x2x2x2).
+The bandwidth between nodes is 2GiB/s in one direction, thus, the aggregated bandwidth is 12GiB/s per node in 3D torus, 20GiB/s per node in 5D torus, and 28GiB/s per node in 7D torus.
+As we can see from the figure, communication time of both applications decrease as the dimensionality of the network increase.
+The aggregated bandwidth of each node can accelerate the transfer of data.
+This can justify the fidelity and consistency of the torus network model.
+
+\begin{figure}[h!] 
+  \centering
+  \includegraphics[width=0.38\textwidth]{figs/torusdimen}
+   \caption{The Percentage of maximum time spent by each MPI rank to send data over 3D, 5D and 7D torus networks. }
+   \label{fig: torus dimen study}
+\end{figure}
+
+\section{Communication Pattern Aware Allocation Policy }
+In this section, we will explore the performance of three applications on different allocations in a 3D torus network.
+First, we analysis the impact from different allocation shape, then we will present the results by using our communication aware allocation policy.
+
+\subsection{Different Allocation Shape}
+As we shown in Fig\ref{fig: allocation}, there could be three differet allocation shape on 3D torus network, namely 3D balanced, 3D un-balanced, and 2D.
+In this section, we will show the performance of three applications on these three different shape allocations.
+
+\subsection{Allocation Strategies Study}
+\label{sec: alloc strategy study}
+
+We proposed two allocation strategies, Compact and Chunk to study the impact of different allocation the application's performance.
+The Compact allocation always provides application with a cubical allocation on 3D torus, regardless application's communication pattern.
+This cubical allocation will keep the locality of the application, all the processes reside on it will have the shortest pair-wise distance.
+
+The Chunk allocation simulation the scenario of non-contiguous allocation strategy employed in production system like Cray XT machines.
+The traditional non-contiguous allocation simply picks available nodes in spite of their locations and assign them to application.
+The motivation of non-contiguous allocation act like this is to maximize system utilization, while being blind to application's communication pattern would greatly prolong application's communication time.
+Although Chunk allocation provides application with non-contigous allocation too, it assign the chunk of nodes in consideration of application's communication pattern.
+For example, we know the dominant communication pattern of CrystalRouter in section \ref{sec:crystalrouter}, every 5 rank form a community, where intensive communication happens between them.
+With such knowledge knowledge about application's communication pattern, the Chunk allocation will provide CrystalRouter with allocation that consists of small chunks, the size of which will be at least 5 nodes.
+
+Figure xxx shows the performance of AMG and CrystalRouter on 3D torus with Compact and Chunk allocation.
+
+\iffalse
+
+\section{Scheduling Mechanism}
+\label{sec:scheduling mechanism}
+
+Figure \ref{fig:overall design} shows the overall design of our new scheduling framework.
+It takes not only job size, user estimated runtime, but also job's communication pattern (CP) as input parameters.
+The CP here is provided by user, which indicate the most dominant communication pattern of the user job.
+The detailed amount of communication data is neither necessary nor possible in this communication pattern.
+As we show in section \ref{sec:alloc sim} , job with different CP has different preference about network requirement, hence different preference about allocations.
+For example, job with All-to-All communication pattern requires more network resource, which means a compact or torus allcoation will guarantee a better performance.
+However, job conforms to Nearest Neighbor communication pattern won't fully utilize the torus network or compact allocation, which means a mesh allocation will serve it equally well.
+
+\begin{figure}[h!]     
+\centering  
+\includegraphics[width=0.48\textwidth]{Pic/design/overalldesign}  
+\caption{Overview of our new scheduling framework.
+It takes not only job size, user estimated runtime, but also job's communication pattern matrix as input and combines these user provide information with system information to make optimal prioritizing and allocation decisions.
+It also maintains hashtable, which provides quick reference for making the specific allocation for incoming job with distinct communication pattern.
+}
+\label{fig:overall design}
+\end{figure}
+
+Since HPC workload tend to be repetive, our new scheduling framework keep record of historic jobs information to provide quick reference for making the preferable allocation for incoming job with distinct communication pattern.
+When user submit his job, our new scheduler will first look for job's user name and executable file name in historic record to check whether this job been submitted before.
+Then the scheduler gets the job's input parameter \{S, CP, ERT\} (job size, Expected RunTime, CommunicationPattern),  and check in the historic data to identify its preferred allocation.
+As time goes by, our new framework will gradually build up a comprehensive record of all the possible job submission, which is helpful to provide quick reference for the future jobs.
+Due to the repetitiveness of the workload, the size of this record will grows rapidly at first, then gently grow for a while and come to stagnation, as shown in figure
+\ref{fig:table size growth}.
+
+%\begin{figure}[h!]     
+%\centering  
+%\includegraphics[width=0.48\textwidth]{Pic/design/tablesize}  
+%\caption{The size growth of historic submitted job records.}
+%\label{fig:table size growth}
+%\end{figure}
+
+We implement a record mechanism as the central component in our framework to keep track of the pattern in workload repetitiveness, hence make predictive optimizations in terms of both job and system performance.
+Firgure \ref{fig:scheduling mechanism} shows the overview of our design.
+The scheduler start to build the table since the first submitted job.
+When user submit a job, the scheduler get job's size and communication pattern, then check the record to find the best allocation for the job.
+On the other hand, the available resource will be categorized and organized into different types, such as 3D balance, 3D unbalance, 2D balance, 2D unbalance (toruse or mesh).
+All these available resource will be updated with real-time system feedback information.
+When a job is finished, the resource it released can be merge into the neighboring one (if the neighbor is also available) to come up with a bigger chunk.
+So the 2D resource can become into 3D by merging with newly released ones, so does mesh to torus.
+
+%\begin{figure}[h!]     
+%\centering  
+%\includegraphics[width=0.48\textwidth]{Pic/design/referencetable}  
+%\caption{overview of scheduling mechanism}
+%\label{fig:scheduling mechanism}
+%\end{figure}
+
+With the feedback from system, our scheduling framework will become knowledgeable about the preference of jobs with different communication pattern.
+The feedback information keep the batch scheduler updated with the latest status of the system such as the number of available nodes, their layout and geometry,  when hotspot or bottleneck come up in network, etc.
+The scheduler schedule/dispatch the following jobs adaptively based these information.
+
+Our new scheduling framework is designed based on the knowledge about the precognition of workload.
+The scheduler works periodically and take 3 steps in each scheduling cycle.
+The scheduling cycle can be predefined by a fixed time interval, or triggered by events like job completion.
+The 3 steps within each scheduling cycle are
+
 \begin{itemize}
-    \item We analyze the repetitiveness as one of signatures for the workload trace collected from Mira, an IBM Blue Gene/Q system at Argonne National Lab.
-
-    \item We analyze the behavior of four most prevalent communication patterns among HPC jobs, under the circumstance of different allocation.
-
-    \item We are the first to propose that HPC batch scheduler should take job's communication pattern (CP) as extra input, which brings benefit for both job and system.
-
-    \item Our new scheduling framework establish a feedback control link between system and job scheduling module. (simple feedback control)
+
+    \item \textbf{Update resource lists from system feedback}.
+The resource management module take the feedback information form system and refresh the resource lists every time there is a job completes.
+The resource released by the job can be merged with any other available resource adjacent to it, in order to form up more compact resources.
+This is similar to Garbage Collection mechanism in memory management.
+1D or 2D resource can become 3D resource and unbalance resource become balance resource after being merged with others.
+The resource list is well organized when the update is finished, thus, better job scheduling
+
+    \item \textbf{Job prioritization and allocation}.
+With job's information like size, expected runtime, dominant communication pattern provided upon submission and the well organized resource list maintained by resource management module, our scheduler will make decision about timing and location of each job being dispatched to the system.
+The scheduling policy will be discussed in detail in \ref{sec:scheduling policy}.
+
+    \item \textbf{Update reservation}.
+When a job couldn't get its optimal reousrce allocation and any available suboptimal allocation can not compensate the job's performance lose, the job has an option to make reservation for its optimal resource.
+Therefore, each time a job got dispatched, the scheduler need to update the reservation, delete those that been satisfied and reserve for the following jobs if necessary.
 
@@
 
-\section{System Observations}
-\label{sec: observation}
-
-Our work is based on two major observations about HPC system.
-First, the job submissions on HPC tend to be repetitive.
-The same application could be submitted to HPC system repetitively with different different input data or configuration.
-Second, most applications running on HPC system are highly communication intensive, and each application has different dominant communication patterns during its different execution phases.
-In this section, we will describe this the repetitiveness of job submission and applications' communication patterns.
+\section{Scheduling Policy}
+\label{sec:scheduling policy}
+
+We designed and implemented two scheduling algorithms, Rackless and Cautious, for our new scheduler framework.
+Both of them take advantage of the scheduler's workload precongnition ability and can deliver comparable performance improvement
+as against the traditional schedulers. The comparision results are presented in \ref{sec:scheduling results}.
+
+Before we introduce our scheduling policy, we will first give explanation about workload and system.
+Suppose the system has 3D torus network and the available resource is consist of a set of 3D or 2D sub-torus chunks, which can be denoted by $\{v\}$, each $v_i = \{x_i, y_i, z_i\}$.
+The size of each chunk $v_i$, $S_{v_i} = x_i\cdot y_i\cdot z_i $.
+
+The job set in the scheduling window is denoted by $\{J\}$, each $J_i$ has size $S_{J_i}$. Due to the repetitiveness of the workload, $S_{J_i} \in \{n_1, n_2, .., n_k\}$, where $k \leq 5$. 
+We can make two reasonable assumption about $\{v\}$ and $\{J\}$.
+
+\begin{equation}
+\label{eq:mod0}
+\exists S_{v_i} > S_{v_j} \quad then\quad S_{v_i} \bmod S_{v_j} \equiv 0
+\end{equation}
+
+\begin{equation}
+\label{eq:nofrag}
+\min S_{J_i} \leq \min S_{v_i} 
+\end{equation}
+
+The reason for both Equation \ref{eq:mod0} and \ref{eq:nofrag} stand is the repetitiveness of workload.
 
@@
 \end{figure}
-
-\subsection{Communication Pattern}
-\label{sec:communication pattern matrix}
-
-In this section, we analysis four communication patterns that are most prevalent in scientific computation.
-They are All-to-All, Nearest Neighbor, Collective and Random\cite{roth}.
-All these four communication patterns are so typical that Message Passing Interface(MPI), the most popular parallel programming scheme, provides explicit support\cite{Pjesivac} \cite{thakur}.
-
-A parallel application usuall conforms to a combination of several basic communication patterns.
-At its different execution phases, the application's communication behavior may follow different basic patterns repectively.
-There are many profiling tools available to capture the communication behavior from parallel applications, such as Tuning and Analysis Utilities(TAU)\cite{tau}, mpiP\cite{mpip} and ScalaTrace\cite{scala}.
-They can keep track of how much data was transferred between each pair of processes in the application and how long those transfer operations took.
-With the help from such profiling tools, we can get a matrix that summarizes the communication between processes within the application.
-Such matrix could contain information about
-
-\begin{itemize}
- \item how many times each process send/receive data to each other process using MPI point-to-point operations.
- \item the amount of data has been sent/received in each MPI operation between processes.
- \item the amount of data has been sent/received by each process using MPI collective operation
-
-\end{itemize}
-
-that can be calibrated by these three communication patterns, we suppose each has \emph{n} parallel processes, we use $m_{i,j}$ to denote the amount of data sent from the \emph{i}th process to \emph{j}th process via network.
-Communication pattern matrix $M$ is used to characterizing the total amount of data transferred among all the processes within each application.
-
-\subsubsection{All to All}
-\label{sec:all-to-all}
-
-Within parallel application that conforms to All-to-All (A2A) communication pattern, each process sends distinct data to all other processes.
-The implementation of this A2A communication in MPI is MPI\_Allgather.
-The communication pattern matrix of this category $M_{a2a}$ like this
-
-\begin{displaymath}
-\begin{bmatrix}
-  0 & m_{1,2} & m_{1,3} & \cdots & m_{1,n-1} & m_{1,n}\\
-  m_{2,1} & 0 & m_{2,3} &\cdots & m_{2,n-1}  & m_{2,n}\\
-  m_{3,1} & m_{3,2} & 0 &\cdots & m_{3,n-1}  & m_{3,n}\\
-  \vdots  & \vdots & \vdots & \ddots & \vdots & \vdots \\
-  m_{n-1,1} & m_{n-1,2} & m_{n-1,3} & \cdots & 0 & m_{n-1,n}\\
-  m_{n,1} & m_{n,2} & m_{n,3} & \cdots & m_{n,n-1} & 0
-\end{bmatrix}
-\end{displaymath}
-
-In $M_{a2a}$, $m_{i,j} = 0$, for every $i=j$, all the elements on diagonal are 0.
-This is because process doesn't send data via network to itself.
-As we can see $M_{a2a}$ has a high density and its rank is $n$.
-One of the representative benchmark application that conforms to A2A communication is DNS3D from NPB3.
-DNS3D is highly sensitive to network and node allocaiton, since during each step, it executes three Fourier transforms for 3-D scalar variables\cite{zhou-ipdps}.
- 
-\subsubsection{Nearest Neighbor}
-\label{sec:nearest neighbor}
-
-Nearest Neighbor (NN) communication pattern is very common in scientific computations.
-This is because lots of physical systems are simulated by a set of elements arranged in grid.
-Cell-based AMR (Adaptive Mesh Refinement) cosmology simulation is one of this kind.
-Wu et al.
-in \cite{jingjin} present an ART application that implement cell-based AMR conforms to diagonal dominant NN communication pattern.
-Its communication pattern matrix $M_{nn}$ like
-
-\begin{displaymath}
- \begin{bmatrix}
-  0 & m_{1,2} & 0 & \cdots & 0 & 0\\
-  m_{2,1} & 0 & m_{2,3} &\cdots & 0  & 0\\
-  0 & m_{3,2} & 0 &\cdots & 0  & 0\\
-  \vdots  & \vdots & \vdots & \ddots & \vdots & \vdots \\
-  0 & 0 & 0 & \cdots & 0 & m_{n-1, n}\\
-  0 & 0 & 0 & \cdots & m_{n, n-1} & 0
- \end{bmatrix} 
-\end{displaymath}
-
-In $M_{nn}$, $m_{i,j} = 0,$ when $|i-j|\neq 1$, all the elements are equal to zero except those on upper or lower diagonal. Matrix $M_{nn}$ is very sparse and its rank is n-1;
-
-\subsubsection{Collective}
-\label{sec:one-to-allandall-to-one}
-\textcolor{red}{This senction is out of scope now, since we don't have MPI collective results. Every following section talk about Collectives are also obsolete.}
-
-Collective communication is a frequently used component in parallel programming schemes like MPI.
-One-to-All (O2A) and All-to-One (A2O) are chosen as typical collective operations in this paper.
-During a O2A operation, one process (usually process with rank ``root'') sends data to all other processes.
-The implementation of this O2A communication in MPI is MPI\_Bcast and MPI\_Scatter.
-So, the matrix $M_{o2a}$ is like
-
-\begin{displaymath}
- \begin{bmatrix}
-  0 & 0 &\cdots & 0 &\cdots & 0\\
-  0 & 0 &\cdots & 0 &\cdots & 0\\
-  \vdots  & \vdots & \ddots & \vdots & \ddots & \vdots \\
-  m_{i,1} & m_{i,2} & \cdots & m_{i,j} &\cdots & m_{i,n}\\
-  \vdots  & \vdots & \ddots & \vdots & \ddots & \vdots \\
-  0 & 0 & \cdots 0 &\cdots & 0
-\end{bmatrix} 
-\end{displaymath}
-
-And another commonly used collective communication is All-to-One(A2O), which is the inverse of One-to-All.
-Instead of spreading data from one process to many others, data in this A2O communication are sent from many processes and gathered to one single process.
-The implementation of this A2O communication in MPI is MPI\_Gather.
-O2A communication is so powerful for many parallel algorithms, such as parallel sorting and searching.
-The matrix $M_{a2o}$ like
-
-\begin{displaymath}
-\begin{bmatrix}
-  0 & 0 &\cdots & m_{1,j} &\cdots & 0\\
-  0 & 0 &\cdots & m_{2,j} &\cdots & 0\\
-  \vdots  & \vdots & \ddots & \vdots & \ddots & \vdots \\
-  0 & 0 & \cdots & m_{i,j} &\cdots & 0\\
-  \vdots  & \vdots & \ddots & \vdots & \ddots & \vdots \\
-  0 & 0 & \cdots & m_{n, j} &\cdots & 0
-\end{bmatrix}  
-\end{displaymath}
-
-For O2A and A2O collective communication, their matrixes have all zero elements except only one row or one column. The rank of both matrixes' is 1.\\
-
-\subsubsection{Hybrid Communication Pattern}
-
-In real life, parallel scientific computation may conform to different communication at different phases.
-During the start phases, applications tend to do some collective communications, such as O2A to distribute data among tasks and in during the end phases, A2O are used to collect comuptation results from every tasks.
-Some applications may conforms to more than one kind of dominant communication pattern, such as pF3D, which is a laser-plasma interaction code that performs local All-to-All communication over sub-communicators and Nearest Neighbor communication over a 3D process grid\cite{langer}.
-There is no real application that exclusively conforms to certain single communication pattern.
-In order to represent real application that with different communication pattern during the its different execution phases, its communication pattern matrix can be formulated like
-
-\begin{equation}
-  M_{hyb} = \alpha_{1}\cdot M_{a2a}+\alpha_{2}\cdot M_{nn}+\alpha_{3}\cdot M_{a2o}+\alpha_{4}\cdot M_{o2a}
-\end{equation}
-And 
-\begin{equation}
-  \sum_{i=1}^{4}\alpha_{i} \equiv 1
-\end{equation}
-
-Here, $\alpha_{i}$ indicates the percentage of communication amount that conforms to each particular communication pattern.
-
-\subsection{Distance Matrix}
-\label{sec:desitance matrix}
-
-In HPC system, batch scheduler plays a important role in job scheduling and system resource management.
-Batch scheduler is responsible for assigning each job the amount of resource (usually refers to compute nodes) it required.
-The amount of compute nodes a job required refers to the job's size.
-Suppose the scheduler provide a job, with size $n$, a set of node $S = \{N_{1}, N_{2},...,N_{n}\}$.
-If there is a direct link connected between node $N_{i}$ and $N_{j}$, the distance between them $d_{i,j} = 1$.
-Thus, we can use a distance matrix $D$ to characterizing the distance between nodes within each allocation.
-Here is what distance matrix $D$ looks like
-
-\begin{displaymath}
- D = \begin{bmatrix}
-  0 & d_{1,2} & d_{1,3} & \cdots & d_{1,n-1} & d_{1,n}\\
-  d_{2,1} & 0 & d_{2,3} &\cdots & d_{2,n-1}  & d_{2,n}\\
-  d_{3,1} & d_{3,2} & 0 &\cdots & 0  & 0\\
-  \vdots  & \vdots & \vdots & \ddots & \vdots & \vdots \\
-  d_{n-1,1} & d_{n-1,2} & d_{n-1,3} & \cdots & 0 & d_{n-1,n}\\
-  d_{n,1} & d_{n,2} & d_{n,3} & \cdots & d_{n,n-1} & 0
- \end{bmatrix} 
-\end{displaymath}
-
-Apparently, the distance between two nodes within each allocation highly depends on system network topology.\\
-
-\subsubsection{k-ary n-cube Distance Matrix}
-\label{sec:k-ary n-cube distance matrix}
-
-As the most popular interconnect nework adopted in current supercomputers, torus (e.g., k-ary n-cube) network has many attractive features like short diameter and low bisection bandwidth \cite{zhou-ipdps}.
-Torus interconnected network has been adopted by IBM Blue Gene and Cray XT series machines\cite{Dong-SC11}.
-In this paper, we use torus connected system to demonstrate the effectiveness of our new scheduling framework.
-Suppose the system network topology is k-ary n-cube,  the diameter of this topology is $\frac{nk}{2}$, and the average distance between any two nodes in this topology is $\frac{nk}{4}$.
-So, $d_{i,j}$ in the distance matrix $D$ of any allocation made on k-ary n-cube will be
-
-\begin{equation}
-1\leq d_{i,j}\leq \frac{nk}{2}
-\end{equation}
-
-And the average value of $d_{i,j}$ is $\frac{nk}{4}$.
-
-Jobs running on Mira, the BG/Q machine and ANL, can specify its allocation topology when been submitted.
-It can choose between with or without torus connection.
-However, allocation without torus connection will suffer a degradation in terms of distance matrix, $d_{i,j}$ will become
-
-\begin{equation}
-  1\leq d_{i,j} \leq k\cdot n
-\end{equation}
-
-\section{Problem Formulation}
-\label{sec:problem formulation}
-
-Although there are lots work about finding the optimal mapping for assigning each process to each node \cite{hpdc14}\cite{jingjin}, in this work, the batch scheduler will assign each job the optimal node allocation with consideration of their communication pattern.
-The traditional scheduler always try to give each job a compact allocation or the node allocation with network resource that more than enough.
-We will demonstrate that such generous allocation is unnecessary for job with different communication pattern.
-
-Given a job with communication matrix $M$ and its allocation distance matrix is $D$, the metric hop-byte $M\cdot D$ is usually used to quantify the network cost.
-The objective of the scheduler is try to find each job a node allocation, whose distance matrix $D$ is most similar to the transpose of communication pattern matrix $M^{\rm T}$, with the objective of
-
-\begin{equation}
-\min \sum_{i=1}^{n} \sum_{j=1}^{n} M \cdot D
-\end{equation}
-
-The traditional scheduler always try to give each job a compact (in terms of shape) allocation or an allocation with superabundant network resource so as to minimize every $d_{i,j}$, which further leads to minimum $M\cdot D$.
-However, if we check the characteristic of those three typical communication pattern in section\ref{sec:communication pattern matrix}, we can see finding a allocation with minimized every $d_{i,j}$ in distance matrix $D$ cost way too much effort.
-\textbf{This is because $\min_{i,j \in \{1,n\}} M \cdot D $ depends only on the elements in D at the position of the transpose of $M$.}
-
-Take the One-to-All communication pattern for an example, providing a job conforms to O2A communication pattern with node allocation with minimized every $d_{i,j}$ in $D$ would be such a huge waste, since those $d_{i,j}$ that matter in this scenario reside in only one column.
-Thus, our scheduler only need to find a allocation within which, exist one node that has short distance to all the other nodes.
-
-\section{Simulation Validation}
-\label{sec:simulation validation}
-
-With the help from parallel discrete event driven simulation, simulating large scale interconnect network with high fidelity becomes possible\cite{ross}.
-Jason et al.
-from Argonne National Lab develop this simulation toolkit CODES, which provides torus and dragonfly network simulation with high fidelity\cite{Jason-2011}\cite{mubarak-sc2012}.
-
-We conduct massive simulation to validate our new scheduler framework model.
-Inorder to get a high fidelity of the network performance under workloads with different communication patterns, we use CODES, a parallel discrete event driven simulation toolkit from Argonne National Lab, which can provide torus connected network simulation with high fidelity\cite{Jason-2011}.
-Torus network model in CODES is built based on several assumptions.
-The model uses deterministic static routing for its simplicity and data injection rate follows the Poisson Process with mean arrival rate of $\lambda$ \cite{Jason-2011}.
-In our simulation, each node continuously generates a Poisson Stream of packets, the size of which is $\delta$.
-The total amount of data generated by each node is $\Delta$, here
-
-\begin{equation}
-\Delta = \sum_{j=1}^{n} m_{i, j} \quad i \in {1,n}
-\end{equation}
-
-\subsection{Simulation Settings}
-\label{sec:simulation settings}
-\textcolor{red}{The scale of network and job would be changed based on the CODES simulation}
-
-For a job requires $n$ nodes submitted to 3D torus-connected system, it gets an allocation of various possible shapes, which could be 3D balanced-cube, 3D unbalanced-cube, 2D balanced-mesh, 2D unbalanced-mesh, 1D list.
-Figure \ref{fig: allocation} shows some possible allocations for a 64-node job.
-We use vector $v$ to denote the shape of an allocation.
-The vector of 3D balanced-cube allocation (in Figure \ref{fig: allocation} as red) is $v=\{4,4,4\}$, 3D unbalanced-cube allocation is $v=\{8,4,2\}$(in Figure \ref{fig: allocation} as green), 2D balanced-mesh allocation is $v=\{8,8,1\}$ (in Figure \ref{fig: allocation} as blue).
-
-\begin{figure}[h!] 
-  \centering
-  \includegraphics[width=0.48\textwidth]{Pic/allocationshape/allocation}
-  \caption{64-node job with different allocation shapes. Red, Green and Blue represents 3D balance, 3D unbalance and 2D balance allocation respectively. }
-  \label{fig: allocation}
-\end{figure}
-
-The workload traces we used for evaluating our new scheduling framework consist of jobs with size variation from 64 to 16K.
-Due to space limitation, we only present the results for job size 64 and 512, both conforms to three different communication patterns.
-We suppose the scale of the torus network simulated in this experiment is set to $32 \times 32 \times 32 $, $32,768$ nodes in total, thus all the possible allocation for job with size 64 and 1024 are shown in table \ref{table:allocationvector}.
-Since we construct a symmetric 3D torus network in our simulation (the bandwidth of each link in x, y, x direction are the same), we suppose the allocation with $v=\{x, y, z\}$ is the same compared to all the other allocations with the dimension of the permutation of \{x, y, z\} in terms of jobs performance.
-
-\begin{table}
-\caption{Different allocation vectors for job with size 64-node and 512-node}
-\label{table:allocationvector}
-\begin{center}
-\begin{tabular}{llllr}
-\toprule
-\multicolumn{4}{c}{Item} \\
-\cmidrule(r){2-5}
-Job Size    & 3D-B & 3D-UB & 2D-B & 2D-UB  \\
-\midrule
-64-node      & \{4,4,4\}    & \{8,4,2\}   & \{8,8,1\}  & \{16,4,1\}      \\
-             &       &     &      & \{32,2,1\} \\
-\midrule
-512-node    & \{8,8,8\}   & \{16,8,4\} &     &       \\
-            &       & \{16,16,2\} &    &  \{32,16,1\}      \\
-            &  & \{32,8,2\} & &\\
-            &  & \{32,4,4\} & &\\
-\midrule
-1024-node    &    & \{16,8,8\} &   \{32,32,1\}  &       \\
-            &       & \{16,16,4\} &    &       \\
-            &  & \{32,8,4\} & &\\
-            &  & \{32,16,2\} & &\\
-
- \bottomrule
-\end{tabular}
-\end{center}
-\end{table}
-
-In our simulation, we set the packet size $\delta$ to 256 Byte.
-The amount of data transferred by each node pair $m_{i,j}$ is set to 2KB, which indicates $2KB/256B = 8$ packets will be transfterred between each node pair through the job's execution.
-The number of packets transferred between each node pair $\sigma = m_{i,j}/\delta$.
-The job size in the simulation is set to $n$.
-
-\begin{table}
-\caption{Nomenclature}
-\label{table:nomenclature}
-\begin{center}
-  \begin{tabular}{ l | c }
-    \hline \hline
-    Symbol 	& Description \\ \hline \hline
-    $\delta$ 	& packet size. \\ \hline
-    $\sigma$    & the number of packets transferred between each node pair. \\ \hline
-    $\Delta$    & the total amount of data generated by each node, $\Delta = \delta \cdot \sigma$. \\ \hline
-    $N $   	& System size, i.e., the number of nodes in the system.\\ \hline
-    $n$    & job size, i.e., the number of nodes required by the job.\\ \hline
-    $M$    & the communication pattern matrix of each job.\\ \hline
-    $m_{i,j}$ 	& the amount of data transferred between process $p_i$ and $p_j$. \\ \hline
-    $D$    & the distance matrix of the allocation of each job. \\ \hline
-    $d_{i,j}$    & the distance between node $i$ and $j$. \\ \hline
-    \hline
-  \end{tabular}
-\end{center}
-\end{table}
-
-\subsection{Results}
-\label{sec:results}
-Given the system is $32\times 32 \times 32$ symmetric 3D torus, table \ref{table:allocationvector} shows all the possible allocation vectors for jobs with 64/512-node.
-Table \ref{tab: 64/512node-job} shows the normalized runtime of these two jobs conform to four different communication patterns with each allocation vector.
-As shown in table \ref{table:allocationvector}, there could be multiple allocation vectors for the same shape, such as the 512-node job could have 4 different allocation vectors with 3D unbalance shape.
-The average value of differnt allocation vectors for the same shape is presented in table \ref{tab: 64/512node-job}.
-
-\begin{table}[ht]
-\begin{center}
-\caption{job conforms to four different communication patterns running 3D torus system.
-In each row, the values on the upper level are the results for 64-node job, the lower level are for 512-node job.
-\textcolor{red}{Results in this table would be provided by CODES simulation, job scale depends on the job traces we decide to use}} 
-\label{tab: 64/512node-job}
-\begin{tabular}{l c c c c c} 
-\toprule % Top horizontal line
-\toprule
-&\multicolumn{5}{c}{Allocation} \\
-\cmidrule(l){2-6}
-Communication Pattern $M$ & 3D-B & 3D-UB & 2D-B & 2D-UB & NC \\ % Column names row
-
-\midrule % In-table horizontal line
-$M_{a2a}$   & 1  &   &    &   &  \\ % Content row 1
-           & 1  &   &    &   &  \\ % Content row 1
-\midrule
-$M_{nn}$  & 1  &   &    &   &  \\ % Content row 1
-            & 1  &   &    &   &   \\ % Content row 1
-\midrule
-$M_{o2a}$   & 1  &   &    &   &  \\ % Content row 1
-            & 1  &   &    &   & \\ % Content row 1
-\midrule
-$M_{a2o}$    & 1  &   &    &   & \\ % Content row 1
-             & 1  &   &    &   &  \\ % Content row 1
-% \midrule
-% $M_{hyb}$   & 1 & 1.24 & 1.37  & 1.66 \\ % Content row 1
-%             &  &  &    \\ % Content row 1
-\midrule
-\bottomrule % Bottom horizontal line
-\end{tabular}
-\end{center}
-\end{table}
-
-\section{Scheduling Mechanism}
-\label{sec:scheduling mechanism}
-
-Figure \ref{fig:overall design} shows the overall design of our new scheduling framework.
-It takes not only job size, user estimated runtime, but also job's communication pattern (CP) as input parameters.
-The CP here is provided by user, which indicate the most dominant communication pattern of the user job, such as All-to-All, Nearest Neighbor, Collective, ect.
-The detailed amount of communication data is neither necessary nor possible in this communication pattern.
-As we prove in section \ref{sec:communication pattern matrix}, job with different CP has different preference about network requirement, hence different preference about allocations.
-For example, job with All-to-All communication pattern requires more network resource, which means a compact or torus allcoation will guarantee a better performance.
-However, job conforms to Nearest Neighbor communication pattern won't fully utilize the torus network or compact allocation, which means a mesh allocation will serve it equally well.
-
-\begin{figure}[h!]     
-\centering  
-\includegraphics[width=0.48\textwidth]{Pic/design/overalldesign}  
-\caption{Overview of our new scheduling framework.
-It takes not only job size, user estimated runtime, but also job's communication pattern matrix as input and combines these user provide information with system information to make optimal prioritizing and allocation decisions.
-It also maintains hashtable, which provides quick reference for making the specific allocation for incoming job with distinct communication pattern.
-}
-\label{fig:overall design}
-\end{figure}
-
-Since HPC workload tend to be repetive, our new scheduling framework keep record of historic jobs information to provide quick reference for making the preferable allocation for incoming job with distinct communication pattern.
-When user submit his job, our new scheduler will first look for job's user name and executable file name in historic record to check whether this job been submitted before.
-Then the scheduler gets the job's input parameter \{S, CP, ERT\} (job size, Expected RunTime, CommunicationPattern),  and check in the historic data to identify its preferred allocation.
-As time goes by, our new framework will gradually build up a comprehensive record of all the possible job submission, which is helpful to provide quick reference for the future jobs.
-Due to the repetitiveness of the workload, the size of this record will grows rapidly at first, then gently grow for a while and come to stagnation, as shown in figure
-\ref{fig:table size growth}.
-
-\begin{figure}[h!]     
-\centering  
-\includegraphics[width=0.48\textwidth]{Pic/design/tablesize}  
-\caption{The size growth of historic submitted job records.}
-\label{fig:table size growth}
-\end{figure}
-
- We implement a record mechanism as the central component in our framework to keep track of the pattern in workload repetitiveness, hence make predictive optimizations in terms of both job and system performance.
-Firgure \ref{fig:scheduling mechanism} shows the overview of our design.
-The scheduler start to build the table since the first submitted job.
-When user submit a job, the scheduler get job's size and communication pattern, then check the record to find the best allocation for the job.
-On the other hand, the available resource will be categorized and organized into different types, such as 3D balance, 3D unbalance, 2D balance, 2D unbalance (toruse or mesh).
-All these available resource will be updated with real-time system feedback information.
-When a job is finished, the resource it released can be merge into the neighboring one (if the neighbor is also available) to come up with a bigger chunk.
-So the 2D resource can become into 3D by merging with newly released ones, so does mesh to torus.
-
-\begin{figure}[h!]     
-\centering  
-\includegraphics[width=0.48\textwidth]{Pic/design/referencetable}  
-\caption{overview of scheduling mechanism}
-\label{fig:scheduling mechanism}
-\end{figure}
-
-With the feedback from system, our scheduling framework will become knowledgeable about the preference of jobs with different communication pattern.
-The feedback information keep the batch scheduler updated with the latest status of the system such as the number of available nodes, their layout and geometry,  when hotspot or bottleneck come up in network, etc.
-The scheduler schedule/dispatch the following jobs adaptively based these information.
-
-Our new scheduling framework is designed based on the knowledge about the precognition of workload.
-The scheduler works periodically and take 3 steps in each scheduling cycle.
-The scheduling cycle can be predefined by a fixed time interval, or triggered by events like job completion.
-The 3 steps within each scheduling cycle are
-
-\begin{itemize}
-
-    \item \textbf{Update resource lists from system feedback}.
-The resource management module take the feedback information form system and refresh the resource lists every time there is a job completes.
-The resource released by the job can be merged with any other available resource adjacent to it, in order to form up more compact resources.
-This is similar to Garbage Collection mechanism in memory management.
-1D or 2D resource can become 3D resource and unbalance resource become balance resource after being merged with others.
-The resource list is well organized when the update is finished, thus, better job scheduling
-
-    \item \textbf{Job prioritization and allocation}.
-With job's information like size, expected runtime, dominant communication pattern provided upon submission and the well organized resource list maintained by resource management module, our scheduler will make decision about timing and location of each job being dispatched to the system.
-The scheduling policy will be discussed in detail in \ref{sec:scheduling policy}.
-
-    \item \textbf{Update reservation}.
-When a job couldn't get its optimal reousrce allocation and any available suboptimal allocation can not compensate the job's performance lose, the job has an option to make reservation for its optimal resource.
-Therefore, each time a job got dispatched, the scheduler need to update the reservation, delete those that been satisfied and reserve for the following jobs if necessary.
-
-\end{itemize}
-
-\section{Scheduling Policy}
-\label{sec:scheduling policy}
-
-We designed and implemented two scheduling algorithms, Rackless and Cautious, for our new scheduler framework.
-Both of them take advantage of the scheduler's workload precongnition ability and can deliver comparable performance improvement
-as against the traditional schedulers. The comparision results are presented in \ref{sec:scheduling results}.
-
-Before we introduce our scheduling policy, we will first give explanation about workload and system.
-Suppose the system has 3D torus network and the available resource is consist of a set of 3D or 2D sub-torus chunks, which can be denoted by $\{v\}$, each $v_i = \{x_i, y_i, z_i\}$.
-The size of each chunk $v_i$, $S_{v_i} = x_i\cdot y_i\cdot z_i $.
-
-The job set in the scheduling window is denoted by $\{J\}$, each $J_i$ has size $S_{J_i}$. Due to the repetitiveness of the workload, $S_{J_i} \in \{n_1, n_2, .., n_k\}$, where $k \leq 5$. 
-We can make two reasonable assumption about $\{v\}$ and $\{J\}$.
-
-\begin{equation}
-\label{eq:mod0}
-\exists S_{v_i} > S_{v_j} \quad then\quad S_{v_i} \bmod S_{v_j} \equiv 0
-\end{equation}
-
-\begin{equation}
-\label{eq:nofrag}
-\min S_{J_i} \leq \min S_{v_i} 
-\end{equation}
-
-The reason for both Equation \ref{eq:mod0} and \ref{eq:nofrag} stand is the repetitiveness of workload.
 
@@
 
+\fi
+
 \section{Experiment Results}
 \label{sec:scheduling results}
-
-\textcolor{red}{all the figs in this section are obsolete!!!}
-
-\begin{figure}[h!] 
-  \centering
-  \includegraphics[width=0.48\textwidth]{Pic/experiment/summary/utilization}
-  \caption{System Utilization got from CC, NC, Greedy, Cautious respectively}
-  \label{fig: uti}
-\end{figure}
-
-\begin{figure}[h!] 
-  \centering
-  \includegraphics[width=0.48\textwidth]{Pic/experiment/summary/awt-vs-cc}
-  \caption{Jobs' average wait time reduction got by comparing CC as against our design.}
-  \label{fig: uti}
-\end{figure}
-
-\begin{figure}[h!] 
-  \centering
-  \includegraphics[width=0.48\textwidth]{Pic/experiment/summary/art-vs-cc}
-  \caption{Jobs' average response time reduction got by comparing CC as against our design.}
-  \label{fig: uti}
-\end{figure}
-
-\begin{figure}[h!] 
-  \centering
-  \includegraphics[width=0.48\textwidth]{Pic/experiment/summary/awt-vs-nc}
-  \caption{Jobs' average wait time reduction got by comparing NC as against our design.}
-  \label{fig: uti}
-\end{figure}
-
-\begin{figure}[h!] 
-  \centering
-  \includegraphics[width=0.48\textwidth]{Pic/experiment/summary/art-vs-nc}
-  \caption{Jobs' average response time reduction got by comparing NC as against our design.}
-  \label{fig: uti}
-\end{figure}
+%
+%\textcolor{red}{all the figs in this section are obsolete!!!}
+%
+%\begin{figure}[h!] 
+%  \centering
+%  \includegraphics[width=0.48\textwidth]{Pic/experiment/summary/utilization}
+%  \caption{System Utilization got from CC, NC, Greedy, Cautious respectively}
+%  \label{fig: uti}
+%\end{figure}
+%
+%
+%\begin{figure}[h!] 
+%  \centering
+%  \includegraphics[width=0.48\textwidth]{Pic/experiment/summary/awt-vs-cc}
+%  \caption{Jobs' average wait time reduction got by comparing CC as against our design.}
+%  \label{fig: uti}
+%\end{figure}
+%
+%
+%\begin{figure}[h!] 
+%  \centering
+%  \includegraphics[width=0.48\textwidth]{Pic/experiment/summary/art-vs-cc}
+%  \caption{Jobs' average response time reduction got by comparing CC as against our design.}
+%  \label{fig: uti}
+%\end{figure}
+%
+%\begin{figure}[h!] 
+%  \centering
+%  \includegraphics[width=0.48\textwidth]{Pic/experiment/summary/awt-vs-nc}
+%  \caption{Jobs' average wait time reduction got by comparing NC as against our design.}
+%  \label{fig: uti}
+%\end{figure}
+%
+%
+%\begin{figure}[h!] 
+%  \centering
+%  \includegraphics[width=0.48\textwidth]{Pic/experiment/summary/art-vs-nc}
+%  \caption{Jobs' average response time reduction got by comparing NC as against our design.}
+%  \label{fig: uti}
+%\end{figure}
 
@@
 
+\bibitem{design forward webpage}
+Department of Energy, “Characterization of the DOE Mini-apps.” [Online].
+\newblock Available: http://portal.nersc.gov/project/CAL/trace.htm
+\newblock (Accessed on: Sept 16th. 2015)
+
+\bibitem{miniapp}
+Bronson Messer
+\newblock "Using a Developing MiniApp to Compare Platform Characteristics on Cray Systems"
+\newblock In Proceedings of Cray User Group 2012
+
+\bibitem{amg}
+V. E. Henson and U. M. Yang, 
+\newblock "BoomerAMG: A Parallel Algebraic Multigrid Solver and Preconditioner",
+\newblock Appl. Num. Math. 41 (2002), pp. 155-177. UCRL-JC-141495.
+
+\bibitem{crystalrouter}
+P. Fischer, J. Lottes, D. Pointer, and A. Siegel, 
+\newblock “Petascale algorithms for reactor hydrodynamics,”
+\newblock Journal of Physics: Conference Series, vol. 125, no. 1, p. 012076, 2008.
+
 %+++++++++++++++++
@@
 
+\bibitem{hpdc2015-ornl}
+Philip. C.R, Jeremy S. M, etc
+\newblock "Automated Characterization of Parallel Application Communication Patterns"
+\newblock HPDC'15
+
 \bibitem{Dong-SC11}
@@
 \bibitem{ross}
-C. Carothers, D. Bauer, and S. Pearce.
-\newblock ROSS: A High-Performance, Low memory, Modular Time Warp System. 
-\newblock Journal of Parallel and Distributed Computing,
-\newblock (62):1648–1669, 2002.
+P. D. Barnes, C. D. Carothers, D. R. Jefferson, and J. M. LaPre, 
+\newblock “Warp speed: executing time warp on 1,966,080 cores,”
+\newblock in Proc. of the 2013 ACM SIGSIM Conf. on Principles of Advanced Discrete Simulation (PADS),
+\newblock May 2013, pp. 327–336.
 
@@
 
+\bibitem{sst}
+Sandia National Labs, “SST DUMPI trace library.” [Online].
+\newblock Available: http://sst.sandia.gov/using dumpi.html 
+\newblock (Accessed on: Sept 3rd, 2015)
+
 \end{thebibliography}

### 2015-09-16 submission.tex
add fig for dimensionality study and alloc-shape study

@@
   \centering
-  \includegraphics[width=0.38\textwidth]{figs/amg/amg_ct}
+  \includegraphics[width=0.38\textwidth]{figs/appstudy/amg/amg_ct}
    \caption{Communication Topology of AMG with 27 MPI Ranks }
@@
   \centering
-  \includegraphics[width=0.38\textwidth]{figs/amg/amg_data_transfer}
+  \includegraphics[width=0.38\textwidth]{figs/appstudy/amg/amg_data_transfer}
    \caption{The amount of data each rank transferred in AMG. }
@@
   \centering
-  \includegraphics[width=0.38\textwidth]{figs/cr/cr10_ct}
+  \includegraphics[width=0.38\textwidth]{figs/appstudy/cr/cr10_ct}
    \caption{Communication Topology of Crystal Router with 10 MPI Ranks }
@@
   \centering
-  \includegraphics[width=0.38\textwidth]{figs/cr/cr100_ct}
+  \includegraphics[width=0.38\textwidth]{figs/appstudy/cr/cr100_ct}
    \caption{Communication Topology of Crystal Router with 100 MPI Ranks }
@@
   \centering
-  \includegraphics[width=0.38\textwidth]{figs/cr/cr_data_transfer}
+  \includegraphics[width=0.38\textwidth]{figs/appstudy/cr/cr_data_transfer}
    \caption{The amount of data each rank transferred in Crystal Router. }
@@
   \centering
-  \includegraphics[width=0.38\textwidth]{figs/mg/mg125_ct}
+  \includegraphics[width=0.38\textwidth]{figs/appstudy/mg/mg125_ct}
    \caption{Communication Topology of MultiGrid with 125 MPI Ranks }
@@
   \centering
-  \includegraphics[width=0.38\textwidth]{figs/mg/mg_data_transfer}
+  \includegraphics[width=0.38\textwidth]{figs/appstudy/mg/mg_data_transfer}
    \caption{The amount of data each rank transferred in MultiGrid. }
@@
   \centering
-  \includegraphics[width=0.38\textwidth]{figs/torusdimen}
-   \caption{The Percentage of maximum time spent by each MPI rank to send data over 3D, 5D and 7D torus networks. }
-   \label{fig: torus dimen study}
-\end{figure}
+  \includegraphics[width=0.38\textwidth]{figs/dimenstudy/maxtime}
+   \caption{The Percentage of Maximum Time spent to send data over 3D, 5D and 7D torus networks by AMG1728, CrystalRouter1000, MultiGrid1000.}
+   \label{fig: dimenstudy-maxtime}
+\end{figure}
+
+\begin{figure}[h!] 
+  \centering
+  \includegraphics[width=0.38\textwidth]{figs/dimenstudy/avgtime}
+   \caption{The Percentage of Average Time spent to send data over 3D, 5D and 7D torus networks by AMG1728, CrystalRouter1000, MultiGrid1000.}
+   \label{fig: dimenstudy-avgtime}
+\end{figure}
+
+\textcolor{green}{cr1000 on 7D result is missing, will be added soon.}
+\textcolor{red}{We can also provide Per-Rank time spent for data transfer. not sure if necessay.}
 
@@
 In this section, we will show the performance of three applications on these three different shape allocations.
+
+There are lots of research work try to design fancy allocation algorithms to provide application with most compact allocation, like using space filling curve (SFC) to index high dimentional cube into one dimensional list.
+Allocation by cutting chunks of nodes from this one dimensional list will guarantee cubical allocation.
+However, try to provide cubical allocation without considering application's communication pattern could cause backfire.
+
+\textcolor{red}{Detail explanation will be add for Fig\ref{fig: shapestudy-avgtime}-Fig\ref{fig: shapestudy-mg125}.}
+
+\begin{figure}[h!] 
+  \centering
+  \includegraphics[width=0.38\textwidth]{figs/shapestudy/avgtime}
+   \caption{The Percentage of Average Time spent to send data by AMG216, CrystalRouter100, MultiGrid125 with 2D, 3D-unbalanced, 3D-balanced allocation.}
+   \label{fig: shapestudy-avgtime}
+\end{figure}
+
+\begin{figure}[h!] 
+  \centering
+  \includegraphics[width=0.38\textwidth]{figs/shapestudy/maxtime}
+   \caption{The Percentage of Max Time spent to send data by AMG216, CrystalRouter100, MultiGrid125 with 2D, 3D-unbalanced, 3D-balanced allocation.}
+   \label{fig: shapestudy-maxtime}
+\end{figure}
+
+\begin{figure}[h!] 
+  \centering
+  \includegraphics[width=0.38\textwidth]{figs/shapestudy/amg216_rank_msgtime}
+   \caption{(AMG216) Time spent by each rank to send data on 2D, 3D-unbalanced, 3D-balanced allocation. }
+   \label{fig: shapestudy-amg216}
+\end{figure}
+
+\begin{figure}[h!] 
+  \centering
+  \includegraphics[width=0.38\textwidth]{figs/shapestudy/cr100_rank_msgtime}
+   \caption{(CrystalRouter100) Time spent by each rank to send data on on 2D, 3D-unbalanced, 3D-balanced allocation.}
+   \label{fig: shapestudy-cr100}
+\end{figure}
+
+\begin{figure}[h!] 
+  \centering
+  \includegraphics[width=0.38\textwidth]{figs/shapestudy/mg125_rank_msgtime}
+   \caption{(MultiGrid125) Time spent by each rank to send data on 2D, 3D-unbalanced, 3D-balanced allocation.}
+   \label{fig: shapestudy-mg125}
+\end{figure}
 

### 2015-09-16 submission.tex
add color notation about imcomplete part

@@
 % can use linebreaks \\ within to get better formatting as desired
-\title{Intelligent HPC Job Scheduling with Communication Pattern Awareness}
+\title{Smart HPC Job Allocation with Communication Pattern Awareness}
 % author names and affiliations
@@
 This can justify the fidelity and consistency of the torus network model.
+\textcolor{green}{cr1000 on 7D result is missing in Fig\ref{fig: dimenstudy-avgtime} and \ref{fig: dimenstudy-maxtime} , will be added soon.}
 
@@
 
-\textcolor{green}{cr1000 on 7D result is missing, will be added soon.}
+\textcolor{blue}{sensitivity study of 5D torus with 2GB,4GB,8GB will be added here. }
+
 \textcolor{red}{We can also provide Per-Rank time spent for data transfer. not sure if necessay.}
@@
 
-Figure xxx shows the performance of AMG and CrystalRouter on 3D torus with Compact and Chunk allocation.
+Figure xxx shows the performance of AMG, CrystalRouter and MultiGrid on 3D torus with Compact and Chunk allocation.
+
+\textcolor{red}{The results and figures for this part is not ready yet!!!}
 
@@
 
-\fi
-
 \section{Experiment Results}
 \label{sec:scheduling results}
-%
-%\textcolor{red}{all the figs in this section are obsolete!!!}
-%
-%\begin{figure}[h!] 
-%  \centering
-%  \includegraphics[width=0.48\textwidth]{Pic/experiment/summary/utilization}
-%  \caption{System Utilization got from CC, NC, Greedy, Cautious respectively}
-%  \label{fig: uti}
-%\end{figure}
-%
-%
-%\begin{figure}[h!] 
-%  \centering
-%  \includegraphics[width=0.48\textwidth]{Pic/experiment/summary/awt-vs-cc}
-%  \caption{Jobs' average wait time reduction got by comparing CC as against our design.}
-%  \label{fig: uti}
-%\end{figure}
-%
-%
-%\begin{figure}[h!] 
-%  \centering
-%  \includegraphics[width=0.48\textwidth]{Pic/experiment/summary/art-vs-cc}
-%  \caption{Jobs' average response time reduction got by comparing CC as against our design.}
-%  \label{fig: uti}
-%\end{figure}
-%
-%\begin{figure}[h!] 
-%  \centering
-%  \includegraphics[width=0.48\textwidth]{Pic/experiment/summary/awt-vs-nc}
-%  \caption{Jobs' average wait time reduction got by comparing NC as against our design.}
-%  \label{fig: uti}
-%\end{figure}
-%
-%
-%\begin{figure}[h!] 
-%  \centering
-%  \includegraphics[width=0.48\textwidth]{Pic/experiment/summary/art-vs-nc}
-%  \caption{Jobs' average response time reduction got by comparing NC as against our design.}
-%  \label{fig: uti}
-%\end{figure}
+
+\fi
 

### 2015-09-16 submission.tex
add zhou as coauthor

@@
 
-\IEEEauthorblockN{Xu Yang\IEEEauthorrefmark{1}, John Jenkins\IEEEauthorrefmark{2}, Misbah Mubarak\IEEEauthorrefmark{2}, Robert B. Ross\IEEEauthorrefmark{2}, Zhiling Lan\IEEEauthorrefmark{1}}
+\IEEEauthorblockN{Xu Yang\IEEEauthorrefmark{1}, John Jenkins\IEEEauthorrefmark{2}, Misbah Mubarak\IEEEauthorrefmark{2}, Robert B.
+Ross\IEEEauthorrefmark{2}, Zhou Zhou\IEEEauthorrefmark{1}, Zhiling Lan\IEEEauthorrefmark{1}}
 
@@
 Chicago, Illinois, USA 60616\\
-xyang56@hawk.iit.edu, lan@iit.edu}
+{xyang56, zzhou}@hawk.iit.edu, lan@iit.edu}
 

### 2015-09-16 submission.tex
remove the commented par

@@
 
-\iffalse
-
-\section{Scheduling Mechanism}
-\label{sec:scheduling mechanism}
-
-Figure \ref{fig:overall design} shows the overall design of our new scheduling framework.
-It takes not only job size, user estimated runtime, but also job's communication pattern (CP) as input parameters.
-The CP here is provided by user, which indicate the most dominant communication pattern of the user job.
-The detailed amount of communication data is neither necessary nor possible in this communication pattern.
-As we show in section \ref{sec:alloc sim} , job with different CP has different preference about network requirement, hence different preference about allocations.
-For example, job with All-to-All communication pattern requires more network resource, which means a compact or torus allcoation will guarantee a better performance.
-However, job conforms to Nearest Neighbor communication pattern won't fully utilize the torus network or compact allocation, which means a mesh allocation will serve it equally well.
-
-\begin{figure}[h!]     
-\centering  
-\includegraphics[width=0.48\textwidth]{Pic/design/overalldesign}  
-\caption{Overview of our new scheduling framework.
-It takes not only job size, user estimated runtime, but also job's communication pattern matrix as input and combines these user provide information with system information to make optimal prioritizing and allocation decisions.
-It also maintains hashtable, which provides quick reference for making the specific allocation for incoming job with distinct communication pattern.
-}
-\label{fig:overall design}
-\end{figure}
-
-Since HPC workload tend to be repetive, our new scheduling framework keep record of historic jobs information to provide quick reference for making the preferable allocation for incoming job with distinct communication pattern.
-When user submit his job, our new scheduler will first look for job's user name and executable file name in historic record to check whether this job been submitted before.
-Then the scheduler gets the job's input parameter \{S, CP, ERT\} (job size, Expected RunTime, CommunicationPattern),  and check in the historic data to identify its preferred allocation.
-As time goes by, our new framework will gradually build up a comprehensive record of all the possible job submission, which is helpful to provide quick reference for the future jobs.
-Due to the repetitiveness of the workload, the size of this record will grows rapidly at first, then gently grow for a while and come to stagnation, as shown in figure
-\ref{fig:table size growth}.
-
-%\begin{figure}[h!]     
-%\centering  
-%\includegraphics[width=0.48\textwidth]{Pic/design/tablesize}  
-%\caption{The size growth of historic submitted job records.}
-%\label{fig:table size growth}
-%\end{figure}
-
-We implement a record mechanism as the central component in our framework to keep track of the pattern in workload repetitiveness, hence make predictive optimizations in terms of both job and system performance.
-Firgure \ref{fig:scheduling mechanism} shows the overview of our design.
-The scheduler start to build the table since the first submitted job.
-When user submit a job, the scheduler get job's size and communication pattern, then check the record to find the best allocation for the job.
-On the other hand, the available resource will be categorized and organized into different types, such as 3D balance, 3D unbalance, 2D balance, 2D unbalance (toruse or mesh).
-All these available resource will be updated with real-time system feedback information.
-When a job is finished, the resource it released can be merge into the neighboring one (if the neighbor is also available) to come up with a bigger chunk.
-So the 2D resource can become into 3D by merging with newly released ones, so does mesh to torus.
-
-%\begin{figure}[h!]     
-%\centering  
-%\includegraphics[width=0.48\textwidth]{Pic/design/referencetable}  
-%\caption{overview of scheduling mechanism}
-%\label{fig:scheduling mechanism}
-%\end{figure}
-
-With the feedback from system, our scheduling framework will become knowledgeable about the preference of jobs with different communication pattern.
-The feedback information keep the batch scheduler updated with the latest status of the system such as the number of available nodes, their layout and geometry,  when hotspot or bottleneck come up in network, etc.
-The scheduler schedule/dispatch the following jobs adaptively based these information.
-
-Our new scheduling framework is designed based on the knowledge about the precognition of workload.
-The scheduler works periodically and take 3 steps in each scheduling cycle.
-The scheduling cycle can be predefined by a fixed time interval, or triggered by events like job completion.
-The 3 steps within each scheduling cycle are
-
-\begin{itemize}
-
-    \item \textbf{Update resource lists from system feedback}.
-The resource management module take the feedback information form system and refresh the resource lists every time there is a job completes.
-The resource released by the job can be merged with any other available resource adjacent to it, in order to form up more compact resources.
-This is similar to Garbage Collection mechanism in memory management.
-1D or 2D resource can become 3D resource and unbalance resource become balance resource after being merged with others.
-The resource list is well organized when the update is finished, thus, better job scheduling
-
-    \item \textbf{Job prioritization and allocation}.
-With job's information like size, expected runtime, dominant communication pattern provided upon submission and the well organized resource list maintained by resource management module, our scheduler will make decision about timing and location of each job being dispatched to the system.
-The scheduling policy will be discussed in detail in \ref{sec:scheduling policy}.
-
-    \item \textbf{Update reservation}.
-When a job couldn't get its optimal reousrce allocation and any available suboptimal allocation can not compensate the job's performance lose, the job has an option to make reservation for its optimal resource.
-Therefore, each time a job got dispatched, the scheduler need to update the reservation, delete those that been satisfied and reserve for the following jobs if necessary.
-
-\end{itemize}
-
-\section{Scheduling Policy}
-\label{sec:scheduling policy}
-
-We designed and implemented two scheduling algorithms, Rackless and Cautious, for our new scheduler framework.
-Both of them take advantage of the scheduler's workload precongnition ability and can deliver comparable performance improvement
-as against the traditional schedulers. The comparision results are presented in \ref{sec:scheduling results}.
-
-Before we introduce our scheduling policy, we will first give explanation about workload and system.
-Suppose the system has 3D torus network and the available resource is consist of a set of 3D or 2D sub-torus chunks, which can be denoted by $\{v\}$, each $v_i = \{x_i, y_i, z_i\}$.
-The size of each chunk $v_i$, $S_{v_i} = x_i\cdot y_i\cdot z_i $.
-
-The job set in the scheduling window is denoted by $\{J\}$, each $J_i$ has size $S_{J_i}$. Due to the repetitiveness of the workload, $S_{J_i} \in \{n_1, n_2, .., n_k\}$, where $k \leq 5$. 
-We can make two reasonable assumption about $\{v\}$ and $\{J\}$.
-
-\begin{equation}
-\label{eq:mod0}
-\exists S_{v_i} > S_{v_j} \quad then\quad S_{v_i} \bmod S_{v_j} \equiv 0
-\end{equation}
-
-\begin{equation}
-\label{eq:nofrag}
-\min S_{J_i} \leq \min S_{v_i} 
-\end{equation}
-
-The reason for both Equation \ref{eq:mod0} and \ref{eq:nofrag} stand is the repetitiveness of workload.
-
-\subsection{Repetitive Job Submission}
-\label{sec: repetitive job submission}
-
-The same application could run repetitively on HPC system.
-This is due to users from some scientific areas have to run the same application multiple times with different configurations.
-For example, some bioinformatics applications need to conduct the same computation i.e.
-protein sequence alignment, for a large sets of protein sequences.
-From system's perspective, all these submissions are the same (same computation instructions and communication phases).
-Therefore, we can observe the same job multiple times with slightly different input parameter, and all these jobs usually belong to the same user.
-To substantiate our claim, we analyze the workload trace in 2014 from Mira, an IBM Blue/Gene Q system.
-As it shown in figure \ref{fig: repetitiveness of Mira},  we list the number of total submissions from five most active users and the percentage of the repetitive ones.
-
-\begin{figure}[h!] 
-  \centering
-  \includegraphics[width=0.38\textwidth]{Pic/JobInfo/repe-mira-2}
-   \caption{ The repetitive job submissions on Mira in 2014. The percentage of repetitive submissions from certain user can be as much as 90\%. }
-   \label{fig: repetitiveness of Mira}
-\end{figure}
-
-\subsection{Scheduling Policy 1: Reckless}
-\label{sec:reckless}
-
-Reckless is designed to be responsive.
-The overhead for Reckless to make scheduling decision is very low, such that it can be competent when system has high job arrival rate.
-Reckless first pick a job from the scheduling window and check according the job's size and communication pattern if its optimal resource is availbale.
-The job get run when enough optimal resource is available.
-However, if there is no enough optimal resource and only enough sub-optimal availale, Reckless will make reservation if job with ATA communication pattern and run job with non-ATA communication pattern anyway.
-And if there is no enough resource for any job in the scheduling window, Reckless will hold and skip the current scheduling cycle, wait for more resource got released and make scheduling in the next cycle.
-
-\begin{algorithm}
-  \label{alg:scheudling1}
-  \caption{Reckless}
-  \begin{algorithmic}
-    \State \{J\}, Job set in the scheudling window
-    \State \{v\}, available system resource sets
-    \While{\{J\} is not empty}
-
-        \If{optimal $v_i$ is available for $j_i$}
-            \State run $j_i$ on $v_i$;
-            \State delete $j_i$ from \{J\};
-            \State update reservation;
-        \EndIf
-
-        \If{$\exists \bar{v_i}$ satisfy $j_i$ AND $\bar{v_i}$ is not optimal}
-            \If{$j_i$ is ATA dominant}
-                \State Wait and Reserve $\hat{v_i}$;
-            \EndIf
-            \If{$j_i$ is not ATA dominant}
-                \State run $j_i$ on $\bar{v_i}$;
-                \State delete $j_i$ from \{J\};
-                \State update reservation;
-            \EndIf
-        \EndIf
-
-        \If{Available \{$v_i$\} $leq$ $forall$$j_i$ requirement}
-            \State hold;
-        \EndIf
-
-    \EndWhile
-  \end{algorithmic}
-\end{algorithm}
-
-\subsection{Scheduling Policy 2: Cautious}
-\label{sec:cautious}
-
-Our second scheduling algorithm named Cautious.
-The difference between Reckless and Cautious is that, Cautious always make evaluation before it assign job to sub-optimal resources.
-The most intuitive evaluation metric is job's turn-around time.
-As shown in \ref{alg: Evaluate}, it will compare the runtime for job on the sub-optimal resource, and the time for waiting optimal resource plus the runtime for job on optimal resource.
-The result will decide whether the job will run on the sub-optimal resource promptly or wait and reserve for its optimal resource.
-
-The other evaluation metrics for Cautious to make decision in terms of system is the Lose of Capacity(LOC).
-As we shown before, the optimal resource always requires embazzle more network resource, which further leads to system fragmentation, and thus low system utilization.
-The system administrator is free to choose the evaluation metrics here based on the system and workload features.
-
-\begin{algorithm}
-  \label{alg:scheudling2}
-  \caption{Cautious}
-  \begin{algorithmic}
-    \State \{J\}, Job set in the scheudling window
-    \State \{v\}, available system resource sets
-    \While{\{J\} is not empty}
-
-        \If{optimal $v_i$ is available for $j_i$}
-            \State run $j_i$ on $v_i$;
-            \State delete $j_i$ from \{J\};
-            \State update reservation;
-        \EndIf
-
-        \If{$\exists \bar{v_i}$ satisfy $j_i$ AND $\bar{v_i}$ is not optimal}
-            \If{Evaluate($j_i$, $\bar{v_i}$)}
-                \State run $j_i$ on $\bar{v_i}$;
-                \State delete $j_i$ from \{J\};
-                \State update reservation;
-            \Else
-                \State Wait and Reserve $\hat{v_i}$;
-                \State Choose from \{J\}-$j_i$ do Backfilling;
-            \EndIf
-        \EndIf
-
-        \If{Available \{$v_i$\} $\leq$ $\forall$$j_i$ requirement}
-            \State hold;
-        \EndIf
-
-    \EndWhile
-
-  \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}
-  \label {alg: Evaluate}
-  \caption{Evaluate($j_i$, $\bar{v_i}$)}
-  \begin{algorithmic}
-    \State Get $\bar{t}$, the time for job $j_i$ to complete on suboptimal resource $\bar{v_i}$;
-    \State Get $\hat{t}$, the time for job $j_i$ to complete on optimal resource $\hat{v_i}$;
-    \State Get $\delta{t}$, the time job $j_i$ wait for optimal resource $\hat{v_i}$;
-    \If{$\delta{t} + \hat{t} \leq \bar{t}$}
-        \State return False;
-    \Else
-        \State return True;
-    \EndIf
-
-  \end{algorithmic}
-\end{algorithm}
-
-Both Reckless and Cautious has its pros and cons.
-Reckless is responsive and efficient, but it make scheduling decision lack of thoughts.
-The job may suffer a long run time when it assign to sub-optimal resource and can lead to potential system fragmentation.
-Cautious, on the other hand, spends time for evaluation if there is only sub-optimal resource for the job.
-
-\section{Evaluation Methodology}
-\label{sec: evaluation methodology}
-
-We conduct a series of experiments using the traces described in Section \ref{sec:job traces} to evaluate our design as against two most frequently used scheduling/allocation policies.
-Both of them use FCFS/EASY backfilling for job prioritizing.
-When it comes to job allocation, the first use contiguous/compact allocation strategy (\textbf{CC}), which means the nodes within the allocation it assigned to every job are spacially contiguous and compact.
-The partition-based allocation strategy used in IBM Blue Gene series of machines are of this kind.
-The second use non-contiguous allocation strategy(\textbf{NC}), which means it provide jobs with allocation without consideration about the adjacency of nodes.
-The nodes within each allocation could be scattered all over the system.
-Machines from Cray Inc.
-use similar allocation strategy like \textbf{NC}.
-Both the traditional \textbf{NC} and \textbf{CC} are agnostic about workload.
-In the rest of the paper, we use \emph{Reckless}, \emph{Cautious}, \emph{CC} and \emph{NC} to denote our scheduling policies and the default ones.
-
-\subsection{Evaluation Metrics} 
-\label{sec:metrics}
-We use three scheduling metrics for the evaluation of our scheduling framework as against the default ones. They have been chosen for the quantified analysis of both system and job's performance.
-
-\begin{itemize}
-
-  \item \emph{System Utilization Rate}.
-System Utilization Rate can be calculated in the following way.
-Suppose $T$ is the total elapsed time for $J$ jobs, $t_e$  the completion time for job $i$ and $t_s$ be its the start time, and $n_i$ be the size of job $i$, then system utilization rate is calculated as
-
-  \begin{eqnarray}\displaystyle
-  \dfrac{\sum_{0\leq i\leq J}{(t_e-t_s) \cdot n_i}}{N \cdot T}
-  \end{eqnarray}
-
-  \item \emph{Average Wait Time}.
-For each job, its wait time refers to the time elapsed between the moment it is submitted and the moment it is dispatched to run.
-This metric is calculated as the average across all the jobs submitted to the system.
-The batch scheduler should cut job's wait time down as much as possible.
-
-   \begin{eqnarray}\displaystyle
-    T_{wait} = t_{start}-t_{arrive}
-  \end{eqnarray}
-
-	\item \emph{Average Job Response time}. Response time refers to the amount of time it take when each job is submitted until it ends, which equals to its wait time plus its run time.
-
-  \begin{eqnarray}\displaystyle
-    T_{response} = T_{run}+T_{wait}
-  \end{eqnarray}
-
-\end{itemize}
-
-A good batch scheduler should always balance system's performance and user job's performance.
-On one hand, the batch scheduler should be as responsive as possible, so that user job won't wait too long to get served.
-On the other hand, the batch scheduler should always guarantee a relative high and stable system uilization so that system resource can be fully utilized.
-
-\subsection{CQSim: Trace-based Scheduling Simulator}
-%\textbf{\textcolor{red}{This subsection is Self Plagiarized from Cluster14!!!!}}\\
-
-Simulation is an integral part of our evaluation of various allocation policies
-as well as their aggregate effects on system utilization, job's wait time and
-response time. We have developed a simulator named CQSim to evaluate our design
-at scale. The simulator is written in Python, and is formed by several modules
-such as job module, node module, scheduling policy module, etc. Each module is
-implemented as a class. The design principles are reusability, extensibility,
-and efficiency. The simulator takes job events from the trace, and an event could be
-job submission, start, end, and other events. Based on these events, the
-simulator emulates job submission, allocation, and execution based on specific
-scheduling and allocation policies. CQsim is open source, and is available to the
-community \cite{Cqsim}.  \\
-
-\subsection{Job Traces}
-\label{sec:job traces}
-%\textbf{\textcolor{red}{This subsection is Self Plagiarized from Cluster14!!!!}}\\
-
-In this work, we use two real workload traces collected from production
-supercomputers to evaluate our design. The objective of using multiple traces is
-to quantify the performance of our design when dealing jobs and systems with
-different characteristics. The first trace we used is from a machine named Blue Horizon at
-the San Diego Supercomputer Center (denoted as SDSC-BLUE in the paper), which
-contains 4,830 jobs. The second trace we used is from a IBM Blue Gene/P system
-named Intrepid at Argonne National Laboratory (denoted as ANL-Intrepid in the paper)
-\cite{par}. This trace contains 2,612 jobs. Figure \ref{fig:jobsizeintowtraces}
-summarizes job size distribution of these traces. ANL-Intrepid is used to
-represent capability computing where jobs require a large amount of computing
-nodes for solving large-scale problems, whereas SDSC-BLUE is used to represent
-capacity computing where the system is utilized to solve a large number of
-small-sized problems.
-
-\begin{figure}[h!] 
-  \centering
-  \includegraphics[width=0.38\textwidth]{Pic/JobInfo/job_dist}
-   \caption{Job size distribution of ANL-Intrepid and SDSC-BLUE }
-   \label{fig:jobsizeintowtraces}
-\end{figure}
-
-\section{Experiment Results}
-\label{sec:scheduling results}
-
-\fi
-
 \section{Related Work}

### 2015-09-20 submission.tex
redraw and update figs

@@
 % can use linebreaks \\ within to get better formatting as desired
-\title{Smart HPC Job Allocation with Communication Pattern Awareness}
+\title{Study of Intra/Inter-Application Interference for Smart Job Allocation on HPC system}
 % author names and affiliations
@@
 Since we only focus on the dominant communication pattern of parallel applications, these MiniApps are perfect candidates for analysis the application's performance on different allocations or on new network architectures.
+We choose three MiniApps, namely, Algebraic Multigrid Solver(AMG), Geometric MultiGrid(MultiGrid) and CrystalRouter.
+Their communication patterns can be representative for most MiniApps listed on the DOE Design Forward Webpage\cite{design forward webpage}.
+The communication pattern figures we present here are generated with the IPM\cite{ipm} data from \cite{design forward webpage}.
 
@@
   \includegraphics[width=0.38\textwidth]{figs/appstudy/amg/amg_ct}
-   \caption{Communication Topology of AMG with 27 MPI Ranks }
+   \caption{Rank-Rank communication topology of AMG with 27 mpi ranks. This figure is redrawed by collected data from\cite{design forward webpage}.   }
    \label{fig: amg communication topology}
@@
   \includegraphics[width=0.38\textwidth]{figs/appstudy/cr/cr10_ct}
-   \caption{Communication Topology of Crystal Router with 10 MPI Ranks }
+   \caption{Rank-Rank communication topology of CrystalRouter with 10 mpi ranks. This figure is redrawed by collected data from\cite{design forward webpage}. }
    \label{fig: crystalrouter10 communication topology}
@@
   \includegraphics[width=0.38\textwidth]{figs/appstudy/cr/cr100_ct}
-   \caption{Communication Topology of Crystal Router with 100 MPI Ranks }
+   \caption{Rank-Rank communication topology of CrystalRouter with 100 mpi ranks. This figure is redrawed by collected data from\cite{design forward webpage}. }
    \label{fig: crystalrouter100 communication topology}
@@
   \includegraphics[width=0.38\textwidth]{figs/appstudy/cr/cr_data_transfer}
-   \caption{The amount of data each rank transferred in Crystal Router. }
+   \caption{The amount of data each rank transferred in CrystalRouter. }
    \label{fig: cr data trans}
@@
   \includegraphics[width=0.38\textwidth]{figs/appstudy/mg/mg125_ct}
-   \caption{Communication Topology of MultiGrid with 125 MPI Ranks }
+   \caption{Rank-Rank communication topology of MultiGrid with 125 mpi ranks. This figure is redrawed by collected data from\cite{design forward webpage}. }
    \label{fig: multigrid125 communication topology}
@@
 
-\section{Simulation Platform}
+\section{Research Vehicle}
 \label{sec:simulation}
@@
 
-\section{Communication Pattern Aware Allocation Policy }
+\section{Interference Analysis}
 In this section, we will explore the performance of three applications on different allocations in a 3D torus network.
@@
 
+\bibitem{ipm}
+David Skinner, Nicholas Wright, Karl Fuerlinger and Kathy Yelick
+\newblock Integrated Performance Montoring(IPM)
+\newblock Available http://ipm-hpc.sourceforge.net/overview.html
+\newblock (Accessed on: Sept 16th. 2015)
+
 \bibitem{miniapp}

### 2015-09-23 submission.tex
the outline of version 2

@@
 \usepackage{amsmath}
+%\usepackage{subfig}
 \usepackage{graphicx}
-%\usepackage{caption}
+\usepackage{caption}
+\usepackage{subcaption}
 %\usepackage[labelfont=bf]{caption}
@@
     \centering
-  \includegraphics[width=0.38\textwidth]{figs/mira}
-   \caption{The activity of Mira, IBM BlueGene/Q system at Argonne National Lab. Many jobs runninng concurrently on the system. Different jobs represented by specific colors. }
-   \label{fig: mira activity}
+    \begin{subfigure}[t]{0.2\textwidth}
+        \centering
+        \includegraphics[height=1.2in]{figs/goodallocation}
+        \caption{God.}
+        \label{fig:overview_sub1}
+    \end{subfigure}%
+    \hspace{1em}%
+    \begin{subfigure}[t]{0.2\textwidth}
+        \centering
+        \includegraphics[height=1.2in]{figs/badallocation}
+        \caption{Bad.}
+        \label{fig:overview_sub2}
+    \end{subfigure}%
+   \caption{Indication of multiple jobs running on Cray system. Many jobs runninng concurrently on the system. Different jobs represented by specific colors.}
+   \label{fig: overview}
 \end{figure}
+
+%\begin{figure}[h!] 
+%  \centering
+%  \includegraphics[width=0.38\textwidth]{figs/mira}
+%   \caption{The activity of Mira, IBM BlueGene/Q system at Argonne National Lab. Many jobs runninng concurrently on the system. Different jobs represented by specific colors. }
+%   \label{fig: mira activity}
+%\end{figure}
 
@@
 
-\begin{figure}[h!] 
-  \centering
-  \includegraphics[width=0.38\textwidth]{figs/appstudy/amg/amg_ct}
-   \caption{Rank-Rank communication topology of AMG with 27 mpi ranks. This figure is redrawed by collected data from\cite{design forward webpage}.   }
-   \label{fig: amg communication topology}
-\end{figure}
-
-\begin{figure}[h!] 
-  \centering
-  \includegraphics[width=0.38\textwidth]{figs/appstudy/amg/amg_data_transfer}
-   \caption{The amount of data each rank transferred in AMG. }
-   \label{fig: amg data trans}
-\end{figure}
-
 Figure \ref{fig: amg communication topology} shows the communication topology of a small scale AMG with 27 MPI Ranks.
@@
 
+\begin{figure}[h!] 
+  \centering
+  \includegraphics[width=0.38\textwidth]{figs/appstudy/amg/amg_pip}
+   \caption{Rank-Rank communication topology of AMG with 27 mpi ranks. This figure is redrawed by collected data from\cite{design forward webpage}.   }
+   \label{fig: amg communication topology}
+\end{figure}
+
+\begin{figure}[h!] 
+  \centering
+  \includegraphics[width=0.38\textwidth]{figs/appstudy/amg/amg_data_transfer}
+   \caption{The amount of data each rank transferred in AMG. }
+   \label{fig: amg data trans}
+\end{figure}
+
 \subsection{Crystal Router}
@@
   \centering
-  \includegraphics[width=0.38\textwidth]{figs/appstudy/cr/cr10_ct}
-   \caption{Rank-Rank communication topology of CrystalRouter with 10 mpi ranks. This figure is redrawed by collected data from\cite{design forward webpage}. }
-   \label{fig: crystalrouter10 communication topology}
-\end{figure}
-
-\begin{figure}[h!] 
-  \centering
-  \includegraphics[width=0.38\textwidth]{figs/appstudy/cr/cr100_ct}
+  \includegraphics[width=0.38\textwidth]{figs/appstudy/cr/cr_pip}
    \caption{Rank-Rank communication topology of CrystalRouter with 100 mpi ranks. This figure is redrawed by collected data from\cite{design forward webpage}. }
@@
 
-Here, we provide the communication topology graph of Crystal Router with 10-rank and 100-rank respectively.
-As we can see from figure \ref{fig: crystalrouter10 communication topology}, every 5-rank is a compact community, where data transfer is very intensive.
+As we can see from figure \ref{fig: crystalrouter100 communication topology}, every 5-rank is a compact community, where data transfer is very intensive.
 Beside the "5-rank community", there is "pair-wise" global data transfer in Crystal Router.
@@
   \centering
-  \includegraphics[width=0.38\textwidth]{figs/appstudy/mg/mg125_ct}
+  \includegraphics[width=0.38\textwidth]{figs/appstudy/mg/mg_pip}
    \caption{Rank-Rank communication topology of MultiGrid with 125 mpi ranks. This figure is redrawed by collected data from\cite{design forward webpage}. }
@@
 
-Torus networks have been extensively used in the current generation of supercomputers because of their linear scaling on per-node cost and competitive communication performance.
-The topology of torus network is k-ary n-cube, with $k^n$ nodes in total arranged in an n-dimensional grid having k nodes in each dimension.
-Each node has 2$\times$ n direct linked neighbor nodes.
-
-For a job requires $n$ nodes submitted to non-partition based 3D torus-connected system, it gets an allocation of various possible shapes, which could be 3D balanced-cube, 3D unbalanced-cube, 2D mesh, or even 1D list.
-Figure \ref{fig: allocation} shows some possible allocations for a 64-node job.
-3D balanced-cube allocation (in Figure \ref{fig: allocation} as red) is $\{4,4,4\}$, 3D unbalanced-cube allocation is $\{8,4,2\}$(in Figure \ref{fig: allocation} as green), 2D mesh allocation is $\{8,8,1\}$ (in Figure \ref{fig: allocation} as blue).
-
-\begin{figure}[h!] 
-  \centering
-  \includegraphics[width=0.48\textwidth]{Pic/allocationshape/allocation}
-  \caption{64-node job with different allocation shapes. Red, Green and Blue represents 3D balanced, 3D unbalanced and 2D allocation respectively. }
-  \label{fig: allocation}
-\end{figure}
-
 We implemented a Job Mapping API for CODES, which can provide flexible job allocation strategies.
@@
 
-\subsection{Allocation Simulation}
-\label{sec:alloc sim}
-
-In this section, we analysis the impact of different allocation strategies on two real applications, AMG and Grystal Router.
 The torus network performance is determined by its dimensionality and link bandwidth.
-We evaluate the data transfer time of AMG and Crystal Router on torus network with different dimensionality and link bandwidth configurations in section \ref{sec: dimensionality study}.
-Then, we will show applications performance on 3D torus network with different allocation strategies in section \ref{sec: alloc strategy study}.
-
-\subsection{Torus Dimensionality}
-\label{sec: dimensionality study}
-
 As the increase of the dimensionality of torus netowrk, so does the number of links connected with each node.
 The increased aggregated bandwidth of each node will definitely reduce the data transfer time of each rank in the application.
-Figure xxx shows the performance of the AMG and CrystalRouter on a 2K node torus network model with a 3D torus (16x16x8), a 5D torus (8x4x4x4x4), and a 7D torus (4x4x4x4x2x2x2).
+Figure \ref{fig: dimenstudy} shows the performance of the AMG, CrystalRouter and MultiGrid on a 2K node torus network model with a 3D torus (16x16x8), a 5D torus (8x4x4x4x4), and a 7D torus (4x4x4x4x2x2x2).
 The bandwidth between nodes is 2GiB/s in one direction, thus, the aggregated bandwidth is 12GiB/s per node in 3D torus, 20GiB/s per node in 5D torus, and 28GiB/s per node in 7D torus.
-As we can see from the figure, communication time of both applications decrease as the dimensionality of the network increase.
-The aggregated bandwidth of each node can accelerate the transfer of data.
+As we can see from these figures, communication time of both applications decrease as the dimensionality of the network increase.
+The increased aggregated bandwidth of each node can accelerate the transfer of data.
 This can justify the fidelity and consistency of the torus network model.
-\textcolor{green}{cr1000 on 7D result is missing in Fig\ref{fig: dimenstudy-avgtime} and \ref{fig: dimenstudy-maxtime} , will be added soon.}
+The same pattern can be found in Figure \ref{fig:bandwidthstudy}, as the bandwidth increase from 2GiB/s, to 4GiB/s and 8GiB/s in the 5D torus network, the data transfer time can be greatly reduced.
 
@@
     \centering
-  \includegraphics[width=0.38\textwidth]{figs/dimenstudy/maxtime}
-   \caption{The Percentage of Maximum Time spent to send data over 3D, 5D and 7D torus networks by AMG1728, CrystalRouter1000, MultiGrid1000.}
-   \label{fig: dimenstudy-maxtime}
+    \begin{subfigure}[t]{0.2\textwidth}
+        \centering
+        \includegraphics[height=1.2in]{figs/dimenstudy/avgtime}
+        \caption{Normalized Average Time.}
+        \label{fig:1sub1}
+    \end{subfigure}%
+    \hspace{1em}%
+    \begin{subfigure}[t]{0.2\textwidth}
+        \centering
+        \includegraphics[height=1.2in]{figs/dimenstudy/maxtime}
+        \caption{Normalized Max Time.}
+        \label{fig:1sub2}
+    \end{subfigure}%
+   \caption{Normalized time spent to send data over 3D, 5D and 7D torus networks by AMG1728, CrystalRouter1000, MultiGrid1000.}
+   \label{fig: dimenstudy}
 \end{figure}
@@
     \centering
-  \includegraphics[width=0.38\textwidth]{figs/dimenstudy/avgtime}
-   \caption{The Percentage of Average Time spent to send data over 3D, 5D and 7D torus networks by AMG1728, CrystalRouter1000, MultiGrid1000.}
-   \label{fig: dimenstudy-avgtime}
+    \begin{subfigure}[t]{0.2\textwidth}
+        \centering
+        \includegraphics[height=1.3in]{figs/bandwidthstudy/avgtime}
+        \caption{Normalized Average Time.}
+        \label{fig:2sub1}
+    \end{subfigure}%
+    \hspace{1em}%
+    \begin{subfigure}[t]{0.2\textwidth}
+        \centering
+        \includegraphics[height=1.3in]{figs/bandwidthstudy/maxtime}
+        \caption{Normalized Max Time.}
+        \label{fig:2sub2}
+    \end{subfigure}%
+   \caption{Normalized time spent to send data over 5D torus with 2GB/s, 4GB/s and 8GB/s bandwidth by AMG1728, CrystalRouter1000, MultiGrid1000.}
+   \label{fig:bandwidthstudy}
 \end{figure}
 
-\textcolor{blue}{sensitivity study of 5D torus with 2GB,4GB,8GB will be added here. }
-
-\textcolor{red}{We can also provide Per-Rank time spent for data transfer. not sure if necessay.}
+%\begin{figure}[h!] 
+%  \centering
+%  \includegraphics[width=0.38\textwidth]{figs/bandwidthstudy/maxtime}
+%   \caption{The Normalized Max Time spent to send data over 5D torus with 2GB/s, 4GB/s and 8GB/s bandwidth by AMG1728, CrystalRouter1000, MultiGrid1000.}
+%   \label{fig: bandwidthstudy-maxtime}
+%\end{figure}
+%
+%\begin{figure}[h!] 
+%  \centering
+%  \includegraphics[width=0.38\textwidth]{figs/bandwidthstudy/avgtime}
+%   \caption{The Normalized Average Time spent to send data over 5D torus with 2GB/s, 4GB/s and 8GB/s bandwidth by AMG1728, CrystalRouter1000, MultiGrid1000.}
+%   \label{fig: bandwidthstudy-avgtime}
+%\end{figure}
 
 \section{Interference Analysis}
-In this section, we will explore the performance of three applications on different allocations in a 3D torus network.
-First, we analysis the impact from different allocation shape, then we will present the results by using our communication aware allocation policy.
+\textbf{presents results of single app running with con/non-cont allocation}
+In this section, we will explore the communication interference.
+First, we will analysis the application's intra-job interference by assigning each job different shapes of allocation.
+All these allcoation are contigous in their dimensionality.
 
-\subsection{Different Allocation Shape}
-As we shown in Fig\ref{fig: allocation}, there could be three differet allocation shape on 3D torus network, namely 3D balanced, 3D un-balanced, and 2D.
-In this section, we will show the performance of three applications on these three different shape allocations.
+Torus networks have been extensively used in the current generation of supercomputers because of their linear scaling on per-node cost and competitive communication performance.
+The topology of torus network is k-ary n-cube, with $k^n$ nodes in total arranged in an n-dimensional grid having k nodes in each dimension.
+Each node has 2$\times$ n direct linked neighbor nodes.
 
-There are lots of research work try to design fancy allocation algorithms to provide application with most compact allocation, like using space filling curve (SFC) to index high dimentional cube into one dimensional list.
+\subsection{Intra-Job Interference Analysis}
+\label{sec: introjob}
+
+To study the intra-job interference of each job, we simulate each job running on 3D torus network with different allocation shapes.
+We assign each application with three different shapes allocation, i.e.
+3D balanced-cube, 3D unbalanced-cube, 2D mesh.
+Figure \ref{figs/cont/allocation} shows some possible allocations for a 64-node job.
+3D balanced allocation is $\{4,4,4\}$ as red, 3D unbalanced allocation is $\{8,4,2\}$ as green, 2D allocation is $\{8,8,1\}$ as blue.
+
+Figure \ref{fig: differentshape-time} shows the data transfter time of three applications running with three different allocation ships.
+
+%There are lots of research work try to design fancy allocation algorithms to provide application with most compact allocation, like using space filling curve (SFC) to index high dimentional cube into one dimensional list.
 Allocation by cutting chunks of nodes from this one dimensional list will guarantee cubical allocation.
@@
 
-\textcolor{red}{Detail explanation will be add for Fig\ref{fig: shapestudy-avgtime}-Fig\ref{fig: shapestudy-mg125}.}
+\begin{figure}[h!]
+    \centering
+    \begin{subfigure}[t]{0.2\textwidth}
+        \centering
+        \includegraphics[height=1.2in]{figs/intra-job/shapestudy/avgtime}
+        \caption{Normalized Average Time.}
+        \label{fig:shape-sub1}
+    \end{subfigure}%
+    \hspace{1em}%
+    \begin{subfigure}[t]{0.2\textwidth}
+        \centering
+        \includegraphics[height=1.2in]{figs/intra-job/shapestudy/maxtime}
+        \caption{Normalized Max Time.}
+        \label{fig:shape-sub2}
+    \end{subfigure}%
+   \caption{Normalized time spent to send data by AMG216, CrystalRouter100, MultiGrid125 with 2D, 3D-unbalanced, 3D-balanced allocation.}
+   \label{fig: differentshape-time}
+\end{figure}
+
+%\begin{figure}[h!] 
+%  \centering
+%  \includegraphics[width=0.38\textwidth]{figs/shapestudy/amg216_rank_msgtime}
+%   \caption{(AMG216) Time spent by each rank to send data on 2D, 3D-unbalanced, 3D-balanced allocation. }
+%   \label{fig: shapestudy-amg216}
+%\end{figure}
+%
+%\begin{figure}[h!]
+%  \centering
+%  \includegraphics[width=0.38\textwidth]{figs/shapestudy/cr100_rank_msgtime}
+%   \caption{(CrystalRouter100) Time spent by each rank to send data on on 2D, 3D-unbalanced, 3D-balanced allocation.}
+%   \label{fig: shapestudy-cr100}
+%\end{figure}
+%
+%\begin{figure}[h!] 
+%  \centering
+%  \includegraphics[width=0.38\textwidth]{figs/shapestudy/mg125_rank_msgtime}
+%   \caption{(MultiGrid125) Time spent by each rank to send data on 2D, 3D-unbalanced, 3D-balanced allocation.}
+%   \label{fig: shapestudy-mg125}
+%\end{figure}
+Then, we simulate each application running on 3D torus with contiguous allocation and non-contiguous allocation.
+For non-contiguous allocation we set the chunk size to 10 in order to preserve application's locality.
+Figure \ref{fig: non-cont-vs-cont} shows the data transfer time of application running with contiguous and non-contiguous allocation.
 
@@
   \centering
-  \includegraphics[width=0.38\textwidth]{figs/shapestudy/avgtime}
-   \caption{The Percentage of Average Time spent to send data by AMG216, CrystalRouter100, MultiGrid125 with 2D, 3D-unbalanced, 3D-balanced allocation.}
-   \label{fig: shapestudy-avgtime}
+  \includegraphics[width=0.38\textwidth]{figs/intra-job/non-cont-vs-cont}
+   \caption{Three Applications running exclusively on 3D torus with contigous and non-contiguous allocation}
+   \label{fig: non-cont-vs-cont}
 \end{figure}
+
+\subsection{Inter-Job Interference Analysis}
+\label{sec: interjob ference study}
+
+\begin{figure}[t!]
+    \centering
+    \begin{subfigure}[t]{0.2\textwidth}
+        \centering
+        \includegraphics[height=1.2in]{figs/cont/allocation}
+        \caption{Contiguous}
+        \label{fig:cont_sub1}
+    \end{subfigure}%
+    \hspace{1em}%
+    \begin{subfigure}[t]{0.2\textwidth}
+        \centering
+        \includegraphics[height=1.2in]{figs/inter-job/noncont/sharing}
+        \caption{Non-contiguous}
+        \label{fig:noncont_sub1}
+    \end{subfigure}%
+%    \hspace{1em}%
+%    \begin{subfigure}[t]{0.2\textwidth}
+%        \centering
+%        \includegraphics[height=1.2in]{figs/inter-job/noncont/exclusive}
+%        \caption{"Exclusive Mode".}
+%        \label{fig:noncont_sub2}
+%    \end{subfigure}
+    \caption{(a) Illustrates the contiguous allocation in different shape. (b) Illustrates the non-contiguous allocation with chunk of size 4.}
+\end{figure}
+
+In order to analysis the interference between concurrently runing jobs, we simulate three application running concurrently on the same 3D network.
+We assign each application with a non-contiguous set of small chunk of nodes, 'chunk' refers to a small number(like 4) of adjacent nodes as shown in Figure \ref{fig:noncont_sub1}.
+The chunk belongs to one application is next to each other.
+The "chunk" size is critical to the application's performance.
+Figure\ref{fig: chunk_size} shows the results of applications' data transfer time with different chunk size.
 
@@
   \centering
-  \includegraphics[width=0.38\textwidth]{figs/shapestudy/maxtime}
-   \caption{The Percentage of Max Time spent to send data by AMG216, CrystalRouter100, MultiGrid125 with 2D, 3D-unbalanced, 3D-balanced allocation.}
-   \label{fig: shapestudy-maxtime}
+  \includegraphics[width=0.38\textwidth]{figs/inter-job/noncont/chunk_size}
+   \caption{The Normalized Time spent to send data by AMG216, CrystalRouter100, MultiGrid125 with different chunk size.}
+   \label{fig: chunk_size}
 \end{figure}
-
-\begin{figure}[h!] 
-  \centering
-  \includegraphics[width=0.38\textwidth]{figs/shapestudy/amg216_rank_msgtime}
-   \caption{(AMG216) Time spent by each rank to send data on 2D, 3D-unbalanced, 3D-balanced allocation. }
-   \label{fig: shapestudy-amg216}
-\end{figure}
-
-\begin{figure}[h!] 
-  \centering
-  \includegraphics[width=0.38\textwidth]{figs/shapestudy/cr100_rank_msgtime}
-   \caption{(CrystalRouter100) Time spent by each rank to send data on on 2D, 3D-unbalanced, 3D-balanced allocation.}
-   \label{fig: shapestudy-cr100}
-\end{figure}
-
-\begin{figure}[h!] 
-  \centering
-  \includegraphics[width=0.38\textwidth]{figs/shapestudy/mg125_rank_msgtime}
-   \caption{(MultiGrid125) Time spent by each rank to send data on 2D, 3D-unbalanced, 3D-balanced allocation.}
-   \label{fig: shapestudy-mg125}
-\end{figure}
-
-\subsection{Allocation Strategies Study}
-\label{sec: alloc strategy study}
-
-We proposed two allocation strategies, Compact and Chunk to study the impact of different allocation the application's performance.
-The Compact allocation always provides application with a cubical allocation on 3D torus, regardless application's communication pattern.
-This cubical allocation will keep the locality of the application, all the processes reside on it will have the shortest pair-wise distance.
-
-The Chunk allocation simulation the scenario of non-contiguous allocation strategy employed in production system like Cray XT machines.
-The traditional non-contiguous allocation simply picks available nodes in spite of their locations and assign them to application.
-The motivation of non-contiguous allocation act like this is to maximize system utilization, while being blind to application's communication pattern would greatly prolong application's communication time.
-Although Chunk allocation provides application with non-contigous allocation too, it assign the chunk of nodes in consideration of application's communication pattern.
-For example, we know the dominant communication pattern of CrystalRouter in section \ref{sec:crystalrouter}, every 5 rank form a community, where intensive communication happens between them.
-With such knowledge knowledge about application's communication pattern, the Chunk allocation will provide CrystalRouter with allocation that consists of small chunks, the size of which will be at least 5 nodes.
-
-Figure xxx shows the performance of AMG, CrystalRouter and MultiGrid on 3D torus with Compact and Chunk allocation.
-
-\textcolor{red}{The results and figures for this part is not ready yet!!!}
 
@@
 \label{sec:related_work}
-Batch scheduling on HPC system has been a hot topic getting constant discussion since the 1990's.
-Feitelson et al proposed the most widely used job scheduling policy which is First Come, First Serve (FCFS) combined with 
-EASY backfilling\cite{feit}. FCFS/EASY backfilling policy greatly improves HPC system's utilization and cuts the wait time 
-of first job in the waiting queue. Since then, many studies seek to refine this classic scheduling paradigm. Tang et al. 
-tried to improve the accuracy of user's estimated job runtime and walltime so as to enhance the efficiency of backfilling and  minimize the system fragmentation\cite{wei-ipdps2010} \cite{wei-jpdc2013} \cite{wei-ipdps2011}.
-% And there are some other variation of FCFS/EASY backfilling proposed to optimize system performance in terms of power consumption and energy cost \cite{zhou}\cite{yang-sc13}.
 
-Only with the information about job size and estimated run time seems to be insufficient for batch scheduler do make the 
-optimal scheduling/allocaiton decision. Some work proposed that the batch scheduler should be aware of some information 
-about system network information in order to fully utilize system resources.Leung et al. presented allocation strategies 
-based on space filling curves and one dimensional packing \cite{leung}. The problem with using space filling curve is that  
-it can only be applied to system with the scale of $2^n$ nodes in each dimension, which is a ideal case that not existing 
-in morden HPC systems. Albing et al. conducted study about the allocation strategies that the Cray
-Application Level Placement Scheduler (ALPS) used \cite{carl-cug}. The job allocation in Cray Linux Environment (CLE) 
-operating system is managed by ALPS, which simply works off an ordered node list, however that is ordered.
+P1: tools for monitoring and profiling
 
-Based on their 
-work, Yang et al. presented a window-based locality-aware job scheduling design for torus-connected system
-\cite{yang-cluster14}. In this new scheduling framework, several jobs are taken into  consideration at the same time for job 
-prioritizing and resource allocation. A list of slots is maintained to preserve node contiguity information for
-resource allocation. Each job been assigned into one of these slot so that the locality can be mostly preserved.
+P2: communication pattern
 
-Rosenthal et al. found there is limited benefit for many applications by
-increasing network bandwidth \cite{rosenthal}. They employed a subset of the
-CORAL mini-applications that represent U.S. Department of Energy workload and
-leverage multirail networkings to evaluate the improvement of these applications
-caused by increased network bandwidth. The applications that send mostly small
-messages or larger messages asynchronously are not bandwidth bounded, hence,
-benefit only slightly or not at all from increased bandwidth.
+P3: Interference between concurrently running jobs
 
-Hoefler et al. propose to use performance modeling techniques to analysis 
-factors that impact the performance of parallel scientific applications 
-\cite{hoefler-modeling}. However, as the scale of HPC continue grows, the 
-interference of concurrently running jobs is getting worse, which is hard to be 
-quantified by performance profiling tools.
-
-Bogdan et al provide a set of guidelines how to configure a network with 
-Dragonfly topology for workload with Nearest Neighbor communication 
-pattern\cite{Bogdan-hpdc14}. They first derived a theoretical model of Nearest 
-Neighbor communication performance on Dragonfly network that can predict network 
-bottleneck. Then, they use simulation to validate the correctness of their 
-theoretical model.
-
-Dong et al describe IBM Blue Gene/Q's 5D torus interconnect
-network\cite{Dong-SC11}. They developed simple benchmarks that conforms to four
-different communication patterns, namely ping-pong, nearest neighbor, broadcast
-and allredcue, to demonstrate the effectiveness of this highly parallel 5D torus
-network.
+P4: Job allocation(little bit)
 

### 2015-10-12 submission.tex
draft version V5

@@
 \usepackage{amsmath}
+\usepackage{textcomp}
 %\usepackage{subfig}
@@
 % can use linebreaks \\ within to get better formatting as desired
-\title{Study of Intra/Inter-Application Interference for Smart Job Allocation on HPC system}
+\title{Study of Intra- and Inter-Job Interference for Improving Allocation on HPC Systems}
 % author names and affiliations
@@
 Chicago, Illinois, USA 60616\\
-{xyang56, zzhou}@hawk.iit.edu, lan@iit.edu}
+\{xyang56, zzhou\}@hawk.iit.edu, lan@iit.edu}
 
@@
 
-It has been widely recoginzed that network contension between concurrently running jobs in HPC system is one of the primary causes of performance variability.
-Optimizing job allocation and avoiding network sharing have been proved to be effective for alleviating network contension and performance degradation.
-In this work, we propose a new job communication pattern aware allocation strategy which serve as a critical module in our future HPC batch scheduler.
-The novelty of our new strategy is that it makes allocation decision with not only system topology and resource availability information, but also with job's communication pattern information.
-The information scheduler gets about job communication pattern can help it to decide whether the job’s performance relies heavily on the network or not, and what the potential inter/intra-job interference would be.
-Thus, the scheduler can pick the preferable resource allocation based on specific job's communication pattern and preserve the locality of allocation in a better way.
-Using traces collected from three typical parallel applications of DOE, we validate the effectiveness of our new desgin.
-The idea presented in this paper is widely applicable: HPC parallel application has specific dominant communication pattern, allocating resource with communication pattern awareness can preserve locality in a bettwer way on systems with differnt network topologies.
+It has been widely recognized that network contention between concurrently running jobs in HPC system is one of the primary causes of performance variability.
+Optimizing job allocation and avoiding network sharing are proven techniques to alleviating network contention and performance degradation.
+In this work, we propose a novel job communication pattern aware allocation strategy which serve as a critical module in our future HPC batch scheduler.
+The novelty of our new strategy is that it makes allocation decisions based on job’s communication pattern communication pattern in addition to system topology and resource availability information.
+The scheduler uses job communication patterns information to decide whether a job’s performance relies heavily on the network or not, and what the potential inter/intra-job interference would be for a given allocation.
+Thus, the scheduler can pick a preferable resource allocation based on a specific job's communication pattern.
+Using traces have been collected from three representative parallel applications, we validate the effectiveness of our new design.
+The idea presented in this paper is widely applicable: HPC parallel applications have common communication patterns, and allocating resource with communication pattern awareness can improve performance on systems with a variety of network topologies.
 
@@
 
-The scale of supercomputer keeps growing from petascale to exascale, to accomodate the the demand of insatiable computing power from scientific research areas.
-Production systems contain hundreds of thousands of processors serve as irreplaceable research vehicle for scientific problems with increasing size and complexity.
-Supercomputers employed in a shared way to accomodate many jobs running concurrently\ref{fig: mira activity} , which will also improve system utilization.
-These jobs share the system infrastructure such as network, I/O bandwidth, which will inevitably cause contention over the shared resource.
-As supercomputer continues to evolve, these shared resources are tend to become the bottleneck for performance.
+The scale of supercomputers keeps growing to accommodate the the demand for computing power required by scientific research areas.
+Supercomputers contain hundreds of thousands of nodes, and serve as irreplaceable research vehicle for scientific problems with increasing size and complexity.
+Supercomputers are employed as shared resource to accommodate many jobs running concurrently\ref{fig: overview} and maintain high system utilization.
+These jobs share the system infrastructure such as network and I/O bandwidth, and inevitably there is contention over the shared resources.
+As supercomputers continue to evolve, these shared resources are increasingly the bottleneck for performance.
 
@@
         \includegraphics[height=1.2in]{figs/goodallocation}
-        \caption{God.}
+        \caption{Contiguous}
         \label{fig:overview_sub1}
@@
         \includegraphics[height=1.2in]{figs/badallocation}
-        \caption{Bad.}
+        \caption{Non-contiguous}
         \label{fig:overview_sub2}
     \end{subfigure}%
-   \caption{Indication of multiple jobs running on Cray system. Many jobs runninng concurrently on the system. Different jobs represented by specific colors.}
+   \caption{Indication of multiple jobs running concurrently with different allocations.
+Different job represented by specific color.
+a) shows the effect of contiguous allocation, which would alleviate the inter-job interference.
+b) shows non-contiguous allocation, which definitely introduce both intra and inter-job interference.
+}
    \label{fig: overview}
@@
 
-One of the most prominant problem comes into researcher's attention is the network contention between concurrently running jobs.
-The network sharing between concurrently running jobs cause communication variability, which reslut in jobs running slower and longer waiting time for results.
-The delay of currently running jobs will increase the queueing time of the following submitted jobs, thus leads to low system throughput, low utilization and high energy cost.
-The network sharing can even cause interference between neighbouring jobs, abort the the running job unexpectedly, or even worse, wrong science results.
+One of the most prominent  problems is network contention between concurrently running jobs.
+The network sharing between concurrently running jobs causes communication variability that results in jobs running slower\cite{abhinav-sc13}.
+The delay of currently running jobs increases the queueing time of the following submitted jobs, thus leading to low system throughput, low utilization\cite{jose-ipdps15}.
 
-This adverse effect of network sharing can be mitigated by providing the job with better node allocation.
-The batch scheduler is responsible for assigning job with the amount of node it required upon submission.
-Currently two scheduling strategies are commonly used on torus-connected systems.
-One is so called partition based systems, where the scheduler assigns each user job a compact and contiguous set of computing nodes.
-IBM Blue Gene series systems fall into this category.
-This strategy is in favor of application’s performance by providing it with isolated partition and exclusive network connections, thus reducing network contention caused by concurrently running jobs sharing network bandwidth.
-However, this strategy can cause internal fragmentation (when more nodes are allocated to a job than it requests) and external fragmentation(when sufficient nodes are available for a request, but they can not be allocated contiguously), therefore leading to poor system performance (e.g., low system utilization and long waiting time for job).
-The other is non-contiguous allocation system, where free nodes are assigned to user job no matter whether they are contiguous or not.
-Cray XT/XE series systems fall into this category.
+This adverse effect of network sharing can be mitigated by providing jobs with better node allocations.
+The batch scheduler is responsible for assigning job the amount of nodes it requires.
+Currently, two allocation strategies are commonly used on torus-connected systems.
+One is so called partition based allocation, where the scheduler assigns each user job a compact and contiguous set of computing nodes.
+IBM Blue Gene series systems use this strategy.
+This strategy favors application performance by providing application with isolated partitions and exclusive network connections, thus reducing network contention caused by concurrently running jobs.
+However, this strategy can cause internal fragmentation (when more nodes are allocated to a job than it requests) and external fragmentation (when sufficient nodes are available for a request, but they can not be allocated contiguously), therefore leading to low system utilization than otherwise possible.
+The other strategy is non-contiguous allocation, where free nodes are assigned to user jobs whether they are contiguous or not.
+Cray XT/XE series systems use this strategy.
 Non-contiguous allocation eliminates internal and external fragmentation as seen in partition-based systems, thereby leading to high system utilization.
-Nevertheless, it introduces other problems such as scattering application processes all over the system and causing both inter-job contention and intra-job contention.
-The non-contiguous node allocation can make inter-process communication less efficient and cause network contention among concurrently running jobs, thereby resulting in poor job performance especially for those communication-intensive jobs.
+However, it introduces other problems, for example scattering application processes all over the system, causing both inter-job and intra-job contention.
+The non-contiguous node allocation can significantly reduce job performance, especially for those communication-intensive jobs.
 
-Partition-based allocation achieves good job performance by sacrificing system performance (e.g., poor system utiliza- tion), whereas non-contiguous allocation can result in better system performance but could severely degrade performance of user jobs (e.g, prolonged wait-time and run-time).
-Supercomputers keep growing in size, its network become more sophisticated then ever before and applications also increase its complexity.
-How to provide application with better allocation that can balance job performance with system performance become a prominant problem need a solution before HPC enters era of exasacle computing.
-In this work, we propose a guideline for future batch scheduling system, a smart way to make allocation with the awareness of job's communication pattern.
-
-Most existing batch schedulers take so few initiative because it been hold under the obscurantism policy.
-Whether it use partition-based or non-contigous policy, means it chooses the benefit of either system or job over the other.
+%\textcolor{red}{
+%Partition-based allocation achieves good job performance by sacrificing system utilization, whereas non-contiguous allocation can result in better system utilization but could severely degrade performance of user jobs.
+What is needed is a more flexible strategy that provides allocations that minimize intra/inter-job contention but allow greater system utilization.
+In this work, we propose a guideline for future batch scheduling systems, a smart way to allocate resources with the awareness of jobs communication patterns.
+%
+%Most existing batch schedulers take so few initiative because it been hold under the obscurantism policy.
+Whether it use partition-based or non-contiguous policy, means it chooses the benefit of either system or job over the other.
 Some propose compromised policies like buddy system\cite{buddy system paper} still make the allocation for job blindly without knowing job's communication information.
-
-Although HPC applications complexity continue to grow, people could get more knowledge about the applications communication information with the help from various parallel application profiling tools, such as TAU\cite{tau}, MPIP\cite{mpip} and DUMPI\cite{sst}.
+%
+%Although HPC applications complexity continue to grow, people could get more knowledge about the applications communication information with the help from various parallel application profiling tools, such as TAU\cite{tau}, MPIP\cite{mpip} and DUMPI\cite{sst}.
 Job's communication trace and performance analysis can be get by using such tool at runtime.
-
-The off-line study of job's communication traces could be helpful to develop better allocation policies for batch scheduler to make on-line scheduling/allocating decisions.
+%
+%The off-line study of job's communication traces could be helpful to develop better allocation policies for batch scheduler to make on-line scheduling/allocating decisions.
 This is because the communication pattern of parallel applications on HPC system can be summarized into a couple of categories\cite{hpdc2015-ornl}.
@@
 In other word, Mira accommodates mostly some particular 14 jobs and most jobs have the same size.
+%}
 
-In this work, we present a guideline for future batch scheduler to make allocation with awareness about job's communication pattern.
-Rather than choose blindly between partition-base or non-contigous allocation, the new allocation mechanism make allocation based on job's communication pattern.
-With detail analysis about application's communication pattern, we can get a clear view of its communication topology graph, data transfer between processes, and decide which processes are tightly connected to which and conduct intensive communication.
-The "local community" consists of such tightly coupled processes shuld get compact alloction.
-Then, the allocation mechanism should make allocations that only needs to guarantee the locality of the subset of the nodes where those "local community" should reside on.
+%=========come back and work from here=============
+Rather than choose blindly between partition-based or non-contiguous allocation, the new allocation strategy choses resources based on a job's communication pattern.
+With detail analysis of application's communication patterns, a clear view of its communication topology is obtained shows which processes are tightly connected and conduct intensive communication.
+The ``local community" consists of such tightly coupled processes should get a compact allocation.
+Then, the allocation strategy only needs to guarantee the locality of the subset of the nodes where those ``local communities" would reside.
 
-We first pick three signature applications from DOE Design Forward Project\cite{design forward webpage} as examples to conduct deep study about their communication patterns.
-We identify the "local community" in these applications and assign them allocation accrodingly.
-The advantage of making allocation is obvious.
-First, compared with partition-based policy, our communication pattern aware policy is more flexible, it doesn't require the system to provide a big partition that can accomodate the whole application, just small set of compact nodes enough for all the "local community" in the application.
-On the other hand, our communication pattern aware allocation policy won't destroy the locality of application's communication pattern.
-On the contrary, it provides small compact node sest for the the "local communities" in the application will preserve the communication locality in the maximum extent.
+In this work, we focus on the study of impact of different allocation strategies on intra/inter-job interference on Torus network.
+Torus networks have been extensively used in the current generation of supercomputers because of their linear scaling on per-node cost and competitive communication performance.
+On the recent Top500 list, 6 of the top 10 supercomputers use a high-radix torus-interconnected network \cite{top500}.
+The current generation of IBM Blue Gene/Q (BG/Q) supercomputer, such as Mira at Argonne Nation Lab and Sequoia at  Lawrence Livermore National Laboratory, has its nodes connected in a 5D torus network \cite{bgq}.
+The K computer from Japan uses the “Tofu” system, which is a 6D mesh/torus topology \cite{tofu}.
+Titan, a Cray XK7 supercomputer located at the Oak Ridge Leadership Computing Facility (OLCF), has nodes connected in a 3D torus within the compute partition \cite{titan}.
+Although the analysis are based on torus network, the idea conveyed from this work is applicable to networks with different topologies.
 
-We use a sophisticated simulation toolkit named CODES, Co-Design of Multi-layer Exascale Storage Architecture, from Argonne National Lab as research vehicle to evaluate the performance of our communication pattern aware allocation policy on torus network.
-CODES enables the exploration of simulating different HPC networks with high fidelity\cite{Jason-2011}.
-We extend CODES with a Job Mapping API that support different allocation policies on torus network.
+We selected three signature applications from DOE Design Forward Project\cite{design forward webpage} as examples to conduct deep study about their communication patterns.
+We use a sophisticated simulation toolkit named CODES, Co-Design of Multi-layer Exascale Storage Architecture\cite{Jason-2011}, from Argonne National Laboratory as research vehicle to evaluate the performance of our communication pattern aware allocation policy on torus networks.
+We analysis each application's performance when it running on torus networks with different dimensionality and interconnected bandwidth.
+Then, we extend CODES with a Job Mapping API that support different allocation policies on torus network.
+Different allocation strategies on torus network are studied with the help of this new Job Mapping API.
+We analysis the intra/inter-job interference by simulating those three applications running exclusively or concurrently on torus network with different allocation strategies.
+According to results of our comprehensive exploration, making allocation with consideration about job's communication pattern is a very promising allocation strategy.
+We believe the idea presented in this work would be the guideline for the design of future HPC batch job scheduler and resource management module.
 
-We also conduct a case study with the results we got from communication pattern aware allocation policy evaluation with real trace from production system.\textcolor{red}{Not Sure About this part!!!}
-
-The rest of this paper orgainzed as follows.
-Section\ref{sec:application study} gives a detailed study of the three applications from DOE Desgin Forward Project.
-Section\ref{sec:simulation} talks about the simulation platform and Section
+The rest of this paper organized as follows.
+Section \ref{sec:application study} gives a detailed study of the three applications from DOE Design Forward Project.
+Section \ref{sec:codes} talks about CODEs as research vehicle for our work.
+Section \ref{sec:torus} shows the performance analysis of three applications on different torus networks.
+Section \ref{sec:interference} provides detailed analysis about the intra/inter-job interference between three applications on torus network.
+Section \ref{sec:discussion} talks about the idea of smart allocation strategy with job communication pattern awareness.
+Section \ref{sec:related_work} talks about the related work in the area and conclusion will be presented in Section \ref{sec:conclusion}.
 
@@
 
-A parallel application usuall conforms to a combination of several basic communication patterns.
-At its different execution phases, the application's communication behavior may follow different basic patterns repectively.
-When we look into the data flow during the application execution, most parallel application start with broadcast operation to distribute the data from root process to other processes, followed by a series of computation and communication that comforms to certain pattern, which is usually the dominant part of application's execution.
-And before the application come to completion, all the working process will return their results to the root dirctly or hierarchically.
-In this work, we focus on the dominant communication pattern because that's where the application spend most of its running time.
-We believe making optimized allocation regarding application's dominant communication pattern would greatly benefit both the application and the system.
-In this section, we analysis two applications whose dominant communication patterns are most prevalent in scientific computation.
+%A parallel application usually conforms to a combination of several basic communication patterns\cite{hpdc2015-ornl}.
+At its different execution phases, the application's communication behavior may follow different basic patterns respectively.
+\textcolor{blue}{When we look into the data flow during application execution, most parallel applications start with broadcast operation to distribute the data from root process to other processes\textcolor{red}{ref needed here}, followed by a series of computation and communication that conforms to certain pattern, which is usually the dominant part of application's execution.
+Before the application come to completion, all the working process will return their results to the root directly or hierarchically.} Parallel application's communication pattern usually involves lots of factors, such as communication intensity, operation-to-operation dependencies and the ``critical path" in its communication topology graph, etc.
+In this work, we focus on the locality of application's rank-to-rank communication topology graph.
 
-There are many profiling tools available to capture the communication behavior from parallel applications, such as Tuning and Analysis Utilities(TAU)\cite{tau}, mpiP\cite{mpip}.
+A parallel application usually conforms to a combination of several basic communication patterns\cite{hpdc2015-ornl}.
+At its different execution phases, the application's communication behavior may follow different basic patterns respectively.
+Collective communication is often used for data distribution and collection among processes.
+Nearest Neighbor communication is commonly seen in lots of structured and unstructured mesh physics packages.
+There are many profiling tools available to capture communication patterns from parallel applications, such as Tuning and Analysis Utilities(TAU)\cite{tau}, mpiP\cite{mpip}.
 They can help analysis parallel application, providing information like the percentage of different MPI operations of the applications, communication topology, the amount data transferred between processes.
 
-In this work, we provide detailed analysis about the communication pattern of two DOE MiniApps.
+In this work, we provide detailed analysis about the communication pattern of three DOE MiniApps.
 MiniApps are reduced proxy applications that encapsulate the salient performance of larger full size applications\cite{miniapp}.
@@
 We choose three MiniApps, namely, Algebraic Multigrid Solver(AMG), Geometric MultiGrid(MultiGrid) and CrystalRouter.
-Their communication patterns can be representative for most MiniApps listed on the DOE Design Forward Webpage\cite{design forward webpage}.
+Their representative communication patterns that are commonly seen in the HPC system workload.
 The communication pattern figures we present here are generated with the IPM\cite{ipm} data from \cite{design forward webpage}.
@@
 \label{sec:amg}
+
+\begin{figure}[t!]
+    \centering
+    \begin{subfigure}[t]{0.22\textwidth}
+        \centering
+        \includegraphics[height=1.2in]{figs/appstudy/amg/amg_pip}
+        \caption{Communication Topology}
+        \label{fig: amg communication topology}
+    \end{subfigure}%
+    ~
+    \begin{subfigure}[t]{0.22\textwidth}
+        \centering
+        \includegraphics[height=1.2in]{figs/appstudy/amg/amg_data_transfer}
+        \caption{Data Volume}
+        \label{fig: amg data trans}
+    \end{subfigure}
+    \caption{AMG. (a) rank-to-rank communication topology graph. (b) the data amount send/received by each rank in AMG.}
+\end{figure}
+
 AMG is the MiniApp of Algebraic Multigrid Solver, which is a parallel algebraic multigrid solver for linear systems arising from problems on unstructured mesh physics packages.
@@
 And the second thing we can find about AMG is that the data transfer amount also has a symmetric pattern.
-Rank 0 has the same amout of data transfer as Rank 26, Rank 1 as of Rank 25, etc.
+Rank 0 has the same amount of data transfer as Rank 26, Rank 1 as of Rank 25, etc.
 
-\begin{figure}[h!] 
-  \centering
-  \includegraphics[width=0.38\textwidth]{figs/appstudy/amg/amg_pip}
-   \caption{Rank-Rank communication topology of AMG with 27 mpi ranks. This figure is redrawed by collected data from\cite{design forward webpage}.   }
-   \label{fig: amg communication topology}
-\end{figure}
-
-\begin{figure}[h!] 
-  \centering
-  \includegraphics[width=0.38\textwidth]{figs/appstudy/amg/amg_data_transfer}
-   \caption{The amount of data each rank transferred in AMG. }
-   \label{fig: amg data trans}
-\end{figure}
+AMG represents a group of applications whose dominant communication pattern is Nearest Neighbor.
+Applications with such similar dominant communication pattern are like PARTISN, SNAP\cite{design forward webpage}.
 
@@
 
+\begin{figure}[t!]
+    \centering
+    \begin{subfigure}[t]{0.22\textwidth}
+        \centering
+        \includegraphics[height=1.2in]{figs/appstudy/cr/cr_pip}
+        \caption{Communication Topology}
+        \label{fig: cr communication topology}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[t]{0.22\textwidth}
+        \centering
+        \includegraphics[height=1.2in]{figs/appstudy/cr/cr_data_transfer}
+        \caption{Data Volume}
+        \label{fig: cr data trans}
+    \end{subfigure}
+    \caption{CrystalRouter. (a) rank-to-rank communication topology graph. (b) the data amount send/received by each rank in CrystalRouter. }
+\end{figure}
+
 The second MiniApp we have is Crystal Router, which is the extracted communication kernel of the full application Nek5000.
@@
 It features spectral element multigrid solvers coupled with a highly scalable, parallel coarse-grid solver that widely used for projects including ocean current modeling, thermal hydraulics of reactor cores, and spatiotemporal chaos.
-The MiniApp of Nek5000, Crystal Router demonstrates the "many-to-many" communication pattern through scalable multi-stage communication process.
+The communication kernel of Nek5000, CrystalRouter demonstrates the "many-to-many" communication pattern through scalable multi-stage communication process.
+\textcolor{yellow}{The way CrystalRouter works is nodes compute for a while, then synchronize and communicate, continually alternating between these two types of activities.}
 
-\begin{figure}[h!] 
-  \centering
-  \includegraphics[width=0.38\textwidth]{figs/appstudy/cr/cr_pip}
-   \caption{Rank-Rank communication topology of CrystalRouter with 100 mpi ranks. This figure is redrawed by collected data from\cite{design forward webpage}. }
-   \label{fig: crystalrouter100 communication topology}
-\end{figure}
-
-\begin{figure}[h!] 
-  \centering
-  \includegraphics[width=0.38\textwidth]{figs/appstudy/cr/cr_data_transfer}
-   \caption{The amount of data each rank transferred in CrystalRouter. }
-   \label{fig: cr data trans}
-\end{figure}
-
-As we can see from figure \ref{fig: crystalrouter100 communication topology}, every 5-rank is a compact community, where data transfer is very intensive.
-Beside the "5-rank community", there is "pair-wise" global data transfer in Crystal Router.
+As we can see from Figure \ref{fig: cr communication topology}, every 5-rank is a compact community, where data transfer is very intensive.
+Beside the "5-rank community", there is "pair-wise" global data transfer in CrystalRouter.
 Obviously, there is data transfer alone the diagnal.
-When the scale of Crystal Router increase to 100, the "5-rank community" is still the dominant pattern in a multi-stage manner.
+\textcolor{yellow}{This is because messages would be buffered on their node of origin until a synchronization point was reached when a single system call sent every message to its destination in one pass.} When the scale of Crystal Router increase to 100, the "5-rank community" is still the dominant pattern in a hierarchical manner.
 The same pattern can be observed from the global data transfer.
-Our observation about the dominant communication pattern of Crystal Router is the multi-stage "5-rank community" and "pair-wise" global data transfer happen recursively as the scale of the application increase.
+Our observation about the dominant communication pattern of CrystalRouter is the multi-stage "5-rank community" and "pair-wise" global data transfer happen hierarchically as the scale of the application increase.
 
@@
 
+CrystalRouter represents a group of applications whose dominant communication is a hybrid of multi-stage local and hierarchical global communication.
+
 \subsection{MultiGrid}
 \label{sec:multigrid}
+Different from AMG, MultiGrid is geometric multigrid v-cycle from production elliptic solver BoxLib, a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes.
+MultiGrid conforms to Many-to-Many communication pattern with decreasing message size and collectives for different parts of the multigrid v-cycle.
+It is widely used for structured grid physics packages.
+Figure\ref{fig: mg communication topology} shows the communication topology of MultiGrid and Figure\ref{fig: mg data trans} shows the the amount of data transferred between ranks in MultiGrid.
 
-Differe from AMG, MultiGrid is geometric multigrid v-cycle from production elliptic solver BoxLib, a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes.
-MultiGrid conforms to 3D lattice communication pattern with decreasing message size and collectives for different parts of the multigrid v-cycle.
-It is widely used for structured grid physics packages.
-
-\begin{figure}[h!] 
+\begin{figure}[t!]
     \centering
-  \includegraphics[width=0.38\textwidth]{figs/appstudy/mg/mg_pip}
-   \caption{Rank-Rank communication topology of MultiGrid with 125 mpi ranks. This figure is redrawed by collected data from\cite{design forward webpage}. }
-   \label{fig: multigrid125 communication topology}
+    \begin{subfigure}[t]{0.22\textwidth}
+        \centering
+        \includegraphics[height=1.2in]{figs/appstudy/mg/mg_pip}
+        \caption{Communication Topology}
+        \label{fig: mg communication topology}
+    \end{subfigure}
+    ~
+    \begin{subfigure}[t]{0.22\textwidth}
+        \centering
+        \includegraphics[height=1.2in]{figs/appstudy/mg/mg_data_transfer}
+        \caption{Data Volume}
+        \label{fig: mg data trans}
+    \end{subfigure}
+    \caption{MultiGrid. (a) rank-to-rank communication topology graph. (b) the data amount send/received by each rank in MultiGrid.}
 \end{figure}
 
-\begin{figure}[h!] 
-  \centering
-  \includegraphics[width=0.38\textwidth]{figs/appstudy/mg/mg_data_transfer}
-   \caption{The amount of data each rank transferred in MultiGrid. }
-   \label{fig: cr data trans}
-\end{figure}
+MultiGrid represents a group of applications whose dominant communication pattern is Many-to-Many (some literatures also refer it as All-to-All \cite{hpdc2015-ornl}).
+
+%\begin{figure*}[t!]
+%    \centering
+%    \begin{subfigure}[t]{0.32\textwidth}
+%        \centering
+%        \includegraphics[height=1.2in]{figs/appstudy/amg/amg_pip}
+%        \caption{AMG}
+%        \label{fig: amg communication topology}
+%    \end{subfigure}%
+%    ~ 
+%    \begin{subfigure}[t]{0.32\textwidth}
+%        \centering
+%        \includegraphics[height=1.2in]{figs/appstudy/cr/cr_pip}
+%        \caption{CrystalRouter}
+%        \label{fig: cr communication topology}
+%    \end{subfigure}
+%    ~
+%    \begin{subfigure}[t]{0.32\textwidth}
+%        \centering
+%        \includegraphics[height=1.2in]{figs/appstudy/mg/mg_pip}
+%        \caption{MultiGrid}
+%        \label{fig: mg communication topology}
+%    \end{subfigure}
+%    \caption{Caption place holder}
+%\end{figure*}
+%
+%\begin{figure*}[t!]
+%    \centering
+%    \begin{subfigure}[t]{0.32\textwidth}
+%        \centering
+%        \includegraphics[height=1.2in]{figs/appstudy/amg/amg_data_transfer}
+%        \caption{AMG DATA}
+%        \label{fig: amg data trans}
+%    \end{subfigure}
+%    ~
+%    \begin{subfigure}[t]{0.32\textwidth}
+%        \centering
+%        \includegraphics[height=1.2in]{figs/appstudy/cr/cr_data_transfer}
+%        \caption{CR DATA}
+%        \label{fig: cr data trans}
+%    \end{subfigure}
+%    ~
+%    \begin{subfigure}[t]{0.32\textwidth}
+%        \centering
+%        \includegraphics[height=1.2in]{figs/appstudy/mg/mg_data_transfer}
+%        \caption{MG DATA}
+%        \label{fig: mg data trans}
+%    \end{subfigure}
+%    \caption{Caption place holder}
+%\end{figure*}
+
+%We summarize the benchmarks and MiniApps on DOE design forward webpage\cite{design forward webpage}, as shown in Table \ref{tab: app summary}.
+According to their dominant communication patterns, there are three categories, namely Nearest Neighbor, All-to-All and 3D Lattice.
+The three applications studies in this work are representative for different categories.
+
+%\begin{table}[ht]
+%\begin{center}
+%\caption{ Summary of dominant communication pattern of MiniApps from DOE design forward project.} 
+%\label{tab: app summary}
+%\begin{tabular}{l c c c c} 
+%\toprule % Top horizontal line
+%\toprule
+%\multicolumn{2}{c}{Applications} \\
+%%\cmidrule(l){2-3}
+%\textbf{Dominant CP}  &  &  & &\\ % Column names row
+%\midrule % In-table horizontal line
+%\textbf{Nearest Neighbor}   &AMG    & SNAP  & PARTISN &\\ % Content row 1
+%
+%\midrule
+%\textbf{All-to-All}  &CrystalRouter    &BoxLib AMR    &MiniFE\\ % Content row 1
+%
+%\midrule
+%\textbf{3D Lattice}   &MultiGrid    & MiniDFT    &FillBoundary \\ % Content row 1
+%       
+%\midrule
+%\bottomrule % Bottom horizontal line
+%\end{tabular}
+%\end{center}
+%\end{table}
+
+%\textcolor{red}{Apps Summary} These three applications have representative communication patterns that are commonly seen in the HPC system workload.
+AMG represents a group of applications whose dominant communication pattern is Nearest Neighbor.
+MultiGrid represents applications whose dominant communication pattern is Many-to-Many (some literatures also refer it to All-to-All\cite{hpdc2015-ornl}).
+CrystalRouter represents those applications whose dominant communication is a hybrid of multi-stage local communication and hierarchical global communication.
 
 \section{Research Vehicle}
-\label{sec:simulation}
+\label{sec:codes}
+
+It is impossible to accurately study the interference between concurrently running jobs with different allocation strategies on production systems.
+There are a couple of reasons for that.
+First, the allocation strategy used on production machine is part of the system software, which can not be changed by user.
+Even the system administrator is not authorized to make that change.
+Second, it is unrealistic to reserve the system exclusively to run the same job with desired allocation without interference then compare the results to that in the presence of interference.
+Due to those reasons, we resort to simulation for this work.
+
+There are couple of simulation platform to study the behavior of large scale systems.
+SimGrid from Inria is a versatile simulation tool sets that provide simulation for Grids, Clouds and HPC systems\cite{simgrid}.
+The Structural Simulation Toolkit (SST) from Sandia National Lab was developed to explore innovations in highly concurrent systems where the ISA, microarchitecture, and memory interact with the programming model and communications system\cite{sst}.
 
@@
 It can take real MPI application trace generated by SST DUMPI\cite{sst} to drive CODES network models.
+We implemented a Job Mapping API for CODES, which can provide flexible job allocation strategies.
+The new API can help us to explore the impact of different allocation strategies to those three applications in Section \ref{sec:application study}
 
-We implemented a Job Mapping API for CODES, which can provide flexible job allocation strategies.
-The new API can help us to explore the impact of different allocation strategies to job with specific communication pattern.
-We have choosen two publicly available HPC network traces provided by the Design Forward Program\cite{design forward webpage}.
-The detailed analysis of these two traces are present in Section\ref{sec:amg} and \ref{sec:crystalrouter}.
+\section{Experiment}
+\label{sec:experiment results}
 
+CODES can provide Torus network model with good scalability and high fidelity\cite{codes-torus}.
+In this section, we will first study the communication behavior of three applications on torus network with different dimensionality and bandwidth configurations.
+Then, we study the intra- and inter-job interference with these three applications by running them exclusively and concurrently with different allocation strategies.
+
+\subsection{Configuration Study }
+\label{sec: config study}
+
+The topology of torus network is k-ary n-cube, with $k^n$ nodes in total arranged in an n-dimensional grid having k nodes in each dimension.
+Each node has 2$\times$ n direct linked neighbor nodes.
 The torus network performance is determined by its dimensionality and link bandwidth.
-As the increase of the dimensionality of torus netowrk, so does the number of links connected with each node.
+As the increase of the dimensionality of torus network, so does the number of links connected with each node.
 The increased aggregated bandwidth of each node will definitely reduce the data transfer time of each rank in the application.
-Figure \ref{fig: dimenstudy} shows the performance of the AMG, CrystalRouter and MultiGrid on a 2K node torus network model with a 3D torus (16x16x8), a 5D torus (8x4x4x4x4), and a 7D torus (4x4x4x4x2x2x2).
-The bandwidth between nodes is 2GiB/s in one direction, thus, the aggregated bandwidth is 12GiB/s per node in 3D torus, 20GiB/s per node in 5D torus, and 28GiB/s per node in 7D torus.
-As we can see from these figures, communication time of both applications decrease as the dimensionality of the network increase.
+Figure \ref{fig:dimensionality-study} shows the performance of the AMG, CrystalRouter and MultiGrid on a 2K node torus network model with a 3D torus (16x16x8), a 5D torus (8x4x4x4x4), and a 7D torus (4x4x4x4x2x2x2).
+The bandwidth of direct link between nodes is 2GiB/s in each direction, thus, the aggregated bandwidth is 12GiB/s per node in 3D torus, 20GiB/s per node in 5D torus, and 28GiB/s per node in 7D torus.
+
+We can see from Figure \ref{fig:dimensionality-study} that the data transfer time of those three applications are greatly reduced as the dimensionality increases.
 The increased aggregated bandwidth of each node can accelerate the transfer of data.
 This can justify the fidelity and consistency of the torus network model.
-The same pattern can be found in Figure \ref{fig:bandwidthstudy}, as the bandwidth increase from 2GiB/s, to 4GiB/s and 8GiB/s in the 5D torus network, the data transfer time can be greatly reduced.
+We allocate each rank within each application linearly on the 3D torus.
+However, due to the variance of data transfer volume of each rank, AMG's data transfer time has great dispersion, shown as the outliers above the boxes in Figure \ref{fig:dimen-amg}.
+There is no such great dispersion for CrystalRouter and MultiGrid, shown in Figure \ref{fig:dimen-cr} and \ref{fig:dimen-mg}, since no great variance of data transfer amount of each rank in these applications.
 
-\begin{figure}[h!]
+\begin{figure*}[t!]
     \centering
-    \begin{subfigure}[t]{0.2\textwidth}
+    \begin{subfigure}[t]{0.32\textwidth}
         \centering
-        \includegraphics[height=1.2in]{figs/dimenstudy/avgtime}
-        \caption{Normalized Average Time.}
-        \label{fig:1sub1}
+        \includegraphics[height=1.5in]{figs/dimenstudy/amg_box}
+        \caption{AMG}
+        \label{fig:dimen-amg}
     \end{subfigure}%
     \hspace{1em}%
-    \begin{subfigure}[t]{0.2\textwidth}
+    \begin{subfigure}[t]{0.32\textwidth}
         \centering
-        \includegraphics[height=1.2in]{figs/dimenstudy/maxtime}
-        \caption{Normalized Max Time.}
-        \label{fig:1sub2}
+        \includegraphics[height=1.5in]{figs/dimenstudy/cr_box}
+        \caption{CrystalRouter}
+        \label{fig:dimen-cr}
     \end{subfigure}%
-   \caption{Normalized time spent to send data over 3D, 5D and 7D torus networks by AMG1728, CrystalRouter1000, MultiGrid1000.}
-   \label{fig: dimenstudy}
-\end{figure}
+    \begin{subfigure}[t]{0.32\textwidth}
+        \centering
+        \includegraphics[height=1.5in]{figs/dimenstudy/mg_box}
+        \caption{MultiGrid}
+        \label{fig:dimen-mg}
+    \end{subfigure}%
+   \caption{Data Transfer Time of AMG, CrystalRouter and MultiGrid on 3D, 5D and 7D torus network.
+The network bandwidth increases as the dimensionality grows.
+Due to the data volume sent/received by each rank greatly varies, there is great dispersion of AMG's data transfer time, shown as outliers above the boxes in Figure \ref{fig:dimen-amg}.}
+   \label{fig:dimensionality-study}
+\end{figure*}
 
-\begin{figure}[h!]
+Then we use fixed dimensionality to study the impact of increased bandwidth.
+We run these three applications on 5D torus with direct link bandwidth increases from 2GiB/s, to 4GiB/s and 8GiB/s.
+As shown in Figure \ref{fig:bandwidth-study}, the data transfer time of each application can be greatly reduced as the bandwidth increases.
+
+\begin{figure*}[t!]
     \centering
-    \begin{subfigure}[t]{0.2\textwidth}
+    \begin{subfigure}[t]{0.32\textwidth}
         \centering
-        \includegraphics[height=1.3in]{figs/bandwidthstudy/avgtime}
-        \caption{Normalized Average Time.}
-        \label{fig:2sub1}
+        \includegraphics[height=1.5in]{figs/bandwidthstudy/amg_bw_box}
+        \caption{AMG}
+        \label{fig:bdwstudy-amg}
     \end{subfigure}%
     \hspace{1em}%
-    \begin{subfigure}[t]{0.2\textwidth}
+    \begin{subfigure}[t]{0.32\textwidth}
         \centering
-        \includegraphics[height=1.3in]{figs/bandwidthstudy/maxtime}
-        \caption{Normalized Max Time.}
-        \label{fig:2sub2}
+        \includegraphics[height=1.5in]{figs/bandwidthstudy/cr_bw_box}
+        \caption{CrystalRouter}
+        \label{fig:bdwstudy-cr}
     \end{subfigure}%
-   \caption{Normalized time spent to send data over 5D torus with 2GB/s, 4GB/s and 8GB/s bandwidth by AMG1728, CrystalRouter1000, MultiGrid1000.}
-   \label{fig:bandwidthstudy}
-\end{figure}
+    \begin{subfigure}[t]{0.32\textwidth}
+        \centering
+        \includegraphics[height=1.5in]{figs/bandwidthstudy/mg_bw_box}
+        \caption{MultiGrid}
+        \label{fig:bdwstudy-mg}
+    \end{subfigure}%
+   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 5D torus network with 2Gb/s, 4Gb/s and 8Gb/s direct link bandwidth. }
+   \label{fig:bandwidth-study}
+\end{figure*}
 
-%\begin{figure}[h!] 
-%  \centering
-%  \includegraphics[width=0.38\textwidth]{figs/bandwidthstudy/maxtime}
-%   \caption{The Normalized Max Time spent to send data over 5D torus with 2GB/s, 4GB/s and 8GB/s bandwidth by AMG1728, CrystalRouter1000, MultiGrid1000.}
-%   \label{fig: bandwidthstudy-maxtime}
-%\end{figure}
-%
-%\begin{figure}[h!] 
-%  \centering
-%  \includegraphics[width=0.38\textwidth]{figs/bandwidthstudy/avgtime}
-%   \caption{The Normalized Average Time spent to send data over 5D torus with 2GB/s, 4GB/s and 8GB/s bandwidth by AMG1728, CrystalRouter1000, MultiGrid1000.}
-%   \label{fig: bandwidthstudy-avgtime}
-%\end{figure}
+We also study the impact of dimensionality with fixed aggregated bandwidth.
+We use 3D, 5D and 7D torus network with the same per-node aggregated bandwidth, which is 28Gb/s.
+With the same aggregated bandwidth, higher dimensionality means shorter diameter and pair-wise distance between nodes.
+The bandwidth between nodes in each torus network are shown in Table \ref{tab: fix-bandwidth}
 
-\section{Interference Analysis}
-\textbf{presents results of single app running with con/non-cont allocation}
-In this section, we will explore the communication interference.
-First, we will analysis the application's intra-job interference by assigning each job different shapes of allocation.
-All these allcoation are contigous in their dimensionality.
+\begin{table}[ht]
+\begin{center}
+\caption{Torus networks with same aggregated bandwidth} 
+\label{tab: fix-bandwidth}
+\begin{tabular}{l c c c} 
+\toprule % Top horizontal line
+\toprule
+&\multicolumn{3}{c}{Bandwidth } \\
+\cmidrule(l){2-4}
+Dimension  & direct & aggregate &\\ % Column names row
+\midrule % In-table horizontal line
+3D      & 4.67Gb/s  & 28Gb/s  & \\  % Content row 1
+\midrule % In-table horizontal line
+5D    & 2.8Gb/s  & 28Gb/s  &\\ % Content row 1
+\midrule % In-table horizontal line
+7D    & 2Gb/s & 28Gb/s & \\
+\midrule
+\bottomrule % Bottom horizontal line
+\end{tabular}
+\end{center}
+\end{table}
 
-Torus networks have been extensively used in the current generation of supercomputers because of their linear scaling on per-node cost and competitive communication performance.
-The topology of torus network is k-ary n-cube, with $k^n$ nodes in total arranged in an n-dimensional grid having k nodes in each dimension.
-Each node has 2$\times$ n direct linked neighbor nodes.
+As we see from Figure \ref{fig:samebd-amg}, the quartiles of ``3D" box is a little bit higher (1\%) than that of ``5D", which means increasing dimensionality from 3D to 5D has little improvement on the average data transfer time for AMG.
+However, there is obvious reduction in those outliers between 3D and 5D.
+When dimensionality increases up to 7D, the improvement in terms of average data transfer time is still small, but more outliers above the top whisker can be spotted.
 
-\subsection{Intra-Job Interference Analysis}
-\label{sec: introjob}
+The performance of CrystalRouter actually deteriorate as the torus dimensionality increasing, shown in Figure \ref{fig:samebd-cr}.
+This is due to the direct link bandwidth reduces from 4.67Gb/s to 2Gb/s, as dimension increases from 3D to 7D.
+CrystalRouter's large amount of intensive local data transfer benefit more from higher direct link bandwidth than its global transfer from lower diameter and shorter pair-wise distance between nodes.
 
-To study the intra-job interference of each job, we simulate each job running on 3D torus network with different allocation shapes.
-We assign each application with three different shapes allocation, i.e.
-3D balanced-cube, 3D unbalanced-cube, 2D mesh.
-Figure \ref{figs/cont/allocation} shows some possible allocations for a 64-node job.
-3D balanced allocation is $\{4,4,4\}$ as red, 3D unbalanced allocation is $\{8,4,2\}$ as green, 2D allocation is $\{8,8,1\}$ as blue.
+The performance of MultiGrid is gradually improved as the dimensionality of torus network increases, shown in Figure \ref{fig:samebd-mg}.
+The reduced diameter and shorten pair-wise distance between nodes would make the ``Many-to-Many" global data transfer in MultiGrid more efficient.
 
-Figure \ref{fig: differentshape-time} shows the data transfter time of three applications running with three different allocation ships.
+We can get the observation that higher dimensionality of torus network would improve the performance of application with ``Many-to-Many" communication patterns, while application with intensive local communication like ``Nearest Neighbor" won't benefit much from higher dimensionality.
 
-%There are lots of research work try to design fancy allocation algorithms to provide application with most compact allocation, like using space filling curve (SFC) to index high dimentional cube into one dimensional list.
-Allocation by cutting chunks of nodes from this one dimensional list will guarantee cubical allocation.
-However, try to provide cubical allocation without considering application's communication pattern could cause backfire.
-
-\begin{figure}[h!]
+\begin{figure*}[t!]
     \centering
-    \begin{subfigure}[t]{0.2\textwidth}
+    \begin{subfigure}[t]{0.32\textwidth}
         \centering
-        \includegraphics[height=1.2in]{figs/intra-job/shapestudy/avgtime}
-        \caption{Normalized Average Time.}
-        \label{fig:shape-sub1}
+        \includegraphics[height=1.5in]{figs/samebdw/amg}
+        \caption{AMG}
+        \label{fig:samebd-amg}
     \end{subfigure}%
     \hspace{1em}%
-    \begin{subfigure}[t]{0.2\textwidth}
+    \begin{subfigure}[t]{0.32\textwidth}
         \centering
-        \includegraphics[height=1.2in]{figs/intra-job/shapestudy/maxtime}
-        \caption{Normalized Max Time.}
-        \label{fig:shape-sub2}
+        \includegraphics[height=1.5in]{figs/samebdw/cr}
+        \caption{CrystalRouter}
+        \label{fig:samebd-cr}
     \end{subfigure}%
-   \caption{Normalized time spent to send data by AMG216, CrystalRouter100, MultiGrid125 with 2D, 3D-unbalanced, 3D-balanced allocation.}
-   \label{fig: differentshape-time}
-\end{figure}
+    \begin{subfigure}[t]{0.32\textwidth}
+        \centering
+        \includegraphics[height=1.5in]{figs/samebdw/mg}
+        \caption{MultiGrid}
+        \label{fig:samebd-mg}
+    \end{subfigure}%
+   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 3D, 5D and 7D torus network with same aggregated bandwidth. }
+   \label{fig: bandwidth-time-box}
+\end{figure*}
 
-%\begin{figure}[h!] 
-%  \centering
-%  \includegraphics[width=0.38\textwidth]{figs/shapestudy/amg216_rank_msgtime}
-%   \caption{(AMG216) Time spent by each rank to send data on 2D, 3D-unbalanced, 3D-balanced allocation. }
-%   \label{fig: shapestudy-amg216}
-%\end{figure}
-%
-%\begin{figure}[h!]
-%  \centering
-%  \includegraphics[width=0.38\textwidth]{figs/shapestudy/cr100_rank_msgtime}
-%   \caption{(CrystalRouter100) Time spent by each rank to send data on on 2D, 3D-unbalanced, 3D-balanced allocation.}
-%   \label{fig: shapestudy-cr100}
-%\end{figure}
-%
-%\begin{figure}[h!] 
-%  \centering
-%  \includegraphics[width=0.38\textwidth]{figs/shapestudy/mg125_rank_msgtime}
-%   \caption{(MultiGrid125) Time spent by each rank to send data on 2D, 3D-unbalanced, 3D-balanced allocation.}
-%   \label{fig: shapestudy-mg125}
-%\end{figure}
-Then, we simulate each application running on 3D torus with contiguous allocation and non-contiguous allocation.
-For non-contiguous allocation we set the chunk size to 10 in order to preserve application's locality.
-Figure \ref{fig: non-cont-vs-cont} shows the data transfer time of application running with contiguous and non-contiguous allocation.
+\subsection{Interference Analysis}
+\label{sec:interference}
 
-\begin{figure}[h!] 
-  \centering
-  \includegraphics[width=0.38\textwidth]{figs/intra-job/non-cont-vs-cont}
-   \caption{Three Applications running exclusively on 3D torus with contigous and non-contiguous allocation}
-   \label{fig: non-cont-vs-cont}
-\end{figure}
+Communication variability due to network sharing can cause application performance degradation in two ways, namely intra- and inter-job interference.
+The \textbf{\emph{intra-job interference}} refers to the network contention between ranks within each application.
+There are lots of research works\cite{jingjin}\cite{ozan} try to come up with better task mapping algorithms to alleviate the intra-job interference, which is out the scope of our work.
+In this work, we focus on the analysis of such interference with different allocations.
 
-\subsection{Inter-Job Interference Analysis}
-\label{sec: interjob ference study}
+\textbf{\emph{Inter-job interference}} is introduced by concurrently running jobs that adjacent to each other sharing network resources.
+Such interference has been identified as one of the major culprits for application's performance variability\cite{abhinav-sc13}\cite{skinner}\cite{rosenthal}.
+Inter-job interference is a more prominent issue for system adopted non-contiguous allocation policy than partition-based system.
+Application's communication time can vary from 36\% faster to 69\% slower due to inter-job interference when running on non-contiguous allocation\cite{abhinav-sc13}.
+
+In this section we will study both kinds of interference with with AMG, CrystalRouter and MultiGrid in their medium scale.
+To be specific, we choose AMG with 216 ranks, CrystalRouter with 100 ranks, and MultiGrid 125 ranks and simulate their running on 2K-node (16 $\times$ 16 $\times$ 8) 3D torus network.
+
+We can get observations that 
+\begin{enumerate}
+    \item The compact allocation can not guarantee the best performance for every application. 
+    \item The performance of application conforms to Nearest Neighbor communication pattern keep relatively stable with different shape allocations, while application conforms to ``Many-to-Many" communication pattern prefer compact allocation.
+    \item The allocation unit size should be decided according to application's communication pattern.
+Unit size that big enough to preserve the locality of the application will result better performance.
+\end{enumerate}
 
@@
         \includegraphics[height=1.2in]{figs/cont/allocation}
-        \caption{Contiguous}
+        \caption{ }
         \label{fig:cont_sub1}
@@
         \includegraphics[height=1.2in]{figs/inter-job/noncont/sharing}
-        \caption{Non-contiguous}
+        \caption{ }
         \label{fig:noncont_sub1}
     \end{subfigure}%
-%    \hspace{1em}%
-%    \begin{subfigure}[t]{0.2\textwidth}
-%        \centering
-%        \includegraphics[height=1.2in]{figs/inter-job/noncont/exclusive}
-%        \caption{"Exclusive Mode".}
-%        \label{fig:noncont_sub2}
-%    \end{subfigure}
-    \caption{(a) Illustrates the contiguous allocation in different shape. (b) Illustrates the non-contiguous allocation with chunk of size 4.}
+    \caption{(a) Illustrates the contiguous allocation for 64-node job in three different shapes.
+``Blue" is 2D mesh, ``Green" is 3D-unbalanced cube and ``Red" is 3D-balanced cube.
+(b) Illustrates the non-contiguous allocation.
+Different colors represent different jobs.
+The nodes assigned to different jobs are interleaved, the allocation unit is 4-node.}
 \end{figure}
 
-In order to analysis the interference between concurrently runing jobs, we simulate three application running concurrently on the same 3D network.
-We assign each application with a non-contiguous set of small chunk of nodes, 'chunk' refers to a small number(like 4) of adjacent nodes as shown in Figure \ref{fig:noncont_sub1}.
-The chunk belongs to one application is next to each other.
-The "chunk" size is critical to the application's performance.
-Figure\ref{fig: chunk_size} shows the results of applications' data transfer time with different chunk size.
+\subsubsection{Intra-Job Interference Analysis}
+\label{sec: introjob}
 
-\begin{figure}[h!] 
+To study the intra-job interference of each job, we simulate each job running on 3D torus network with different allocation shapes.
+We assign each application with three different shapes allocation, i.e.
+3D balanced-cube, 3D unbalanced-cube, 2D mesh.
+
+3D balanced-cube, shown as red in Figure \ref{fig:cont_sub1}, can guarantee the minimum \emph{average pair-wise distance} within the allocation.
+There are some research work like \cite{leung} \cite{abhinav-sc13} try to prove that compact allocation can guarantee job with better performance.
+They use many metrics to evaluate the compactness of the allocation, such as \emph{average pair-wise distance}, \emph{diameter} and \emph{contiguity}.
+In this work, we select 3D balanced-cube as the most compact allocation on a 3D torus network.
+
+3D unbalanced-cube, shown as green in Figure \ref{fig:cont_sub1}, is a rectangular prism, which is the possible allocation shape on system with asymmetric networks.
+For example, Blue Waters Cray XE6/XK7 system network is 3D torus with Gemini routers.
+The network connections in the \emph{y}direction have only half the bandwidth of the cables used in the \emph{x} and \emph{z} directions.
+In order to take advantage of the faster links in the \emph{x} and \emph{z} directions, job allocation is always start from X-Z plane, which leads to a rectangular prism shaped allocation\cite{RF}.
+
+2D mesh, shown as blue in Figure \ref{fig:cont_sub1}, can be cut out from a single layer of the 3D torus.
+2D mesh is a very common allocation shape in torus network for both partition-based and non-contiguous allocation strategies.
+For example, Cray Application Level Placement Scheduler indexes the nodes in torus network into a list and make allocation by simply works off the this list \cite{carl-cug}.
+When the list is obtained by sorting the nodes based on their spacial coordinates in the torus, allocation made off from this list will be 2D mesh.
+The IBM Blue/Gene Q supercomputer Mira at Argonne Leadership Computing Facility also allow its allocation partition configured into mesh \cite{zhou-ipdps}.
+
+There are lots of research work try to design fancy allocation algorithms to provide application with most compact allocation.
+However, providing cubical allocation without considering application's communication pattern won't guarantee the best performance for every application.
+Figure \ref{fig:shapstudy-amg} shows that for AMG, who has this symmetric local communication pattern through all ranks, spent less time when assigned with 3D-unbalanced allocation than 3D-balanced.
+And MultiGrid gets the best performance (shortest data transfer time) when running on 3D-balanced allocation, as shown in Figure \ref{fig:shapstudy-mg}.
+Since MultiGrid's communication pattern is many-to-many dominant, 3D-balanced is the most compact allocation with shortest pair-wise distance between nodes, which can reduce the aggregated hops for transferring message between ranks in MultiGrid.
+CrystalRouter has with both dominant local communication and multi-stage global data transfer, the 3D-balance is also the best allocation, but its advantage over 3D-unbalanced is not that obvious as to MultiGrid, as shown in Figure \ref{fig:shapstudy-cr}.
+
+\emph{OBSERVATION}
+Compact allocation should be provided to application with global data transfer, such as those conforms to ``Many-to-Many" communication pattern.
+Application with dominant communication pattern like Nearest Neighbor won't fully utilize the compact allocation and could get even better performance on a relaxed allocation shape.
+
+\begin{figure*}[t!]
     \centering
-  \includegraphics[width=0.38\textwidth]{figs/inter-job/noncont/chunk_size}
-   \caption{The Normalized Time spent to send data by AMG216, CrystalRouter100, MultiGrid125 with different chunk size.}
-   \label{fig: chunk_size}
-\end{figure}
+    \begin{subfigure}[t]{0.32\textwidth}
+        \centering
+        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/amg_box}
+        \caption{AMG}
+        \label{fig:shapstudy-amg}
+    \end{subfigure}%
+    \hspace{1em}%
+    \begin{subfigure}[t]{0.32\textwidth}
+        \centering
+        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/cr_box}
+        \caption{CrystalRouter}
+        \label{fig:shapstudy-cr}
+    \end{subfigure}%
+    \begin{subfigure}[t]{0.32\textwidth}
+        \centering
+        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/mg_box}
+        \caption{MultiGrid}
+        \label{fig:shapstudy-mg}
+    \end{subfigure}%
+   \caption{Intra-job Study: Data Transfer Time of 3 Apps on 3 different shapes allocation}
+   \label{fig: shapestudy}
+\end{figure*}
+
+In non-contiguous allocation systems, the nodes assigned to each job could be scatter anywhere in the system.
+Some HPC systems predefine the minimum number of nodes consist of the allocation unit.
+The nodes within each unit are contiguous, and units assigned to each job may have arbitrary locations.
+We assign each application with a non-contiguous set of allocation units, refers to a small number(like 4) of adjacent nodes as shown in Figure \ref{fig:noncont_sub1}.
+We compare application's performance when it runs with contiguous allocation and non-contiguous allocation with unit size of 16, as shown in Figure \ref{fig: cont-time-box}.
+
+Figure \ref{fig:cont-amg} shows that AMG is insensitive to the allocation's contiguity.
+Non-contiguous allocation can serve AMG equally well as contiguous allocation, as long as the allocation unit in non-contiguous allocation is big enough to preserve the locality of AMG.
+CrystalRouter shows preference in choosing between contiguous and non-contiguous allocation.
+As shown in Figure \ref{fig:cont-cr}, there is obvious degradation in terms data transfer time when CrystalRouter running with non-contiguous allocation.
+The global data transfer in CrystalRouter will take more hops on non-contiguous allocation, thus result in longer communication time.
+When it comes to MultiGrid shown in Figure \ref{fig:cont-mg}, this degradation is even worse.
+Since the global data transfer take a great portion of MultiGrid ``Many-to-Many" communication pattern.
+
+\emph{OBSERVATION}
+The performance of application conforms to Nearest Neighbor communication pattern keep relatively stable with different allocations.
+Non-contiguous allocation won't aggravate intra-job interference as long as the unit size can preserve the locality in application's communication topology graph.
+
+\begin{figure*}[t!]
+    \centering
+    \begin{subfigure}[t]{0.32\textwidth}
+        \centering
+        \includegraphics[height=1.5in]{figs/intra-job/cont/amg_box}
+        \caption{AMG}
+        \label{fig:cont-amg}
+    \end{subfigure}%
+    \hspace{1em}%
+    \begin{subfigure}[t]{0.32\textwidth}
+        \centering
+        \includegraphics[height=1.5in]{figs/intra-job/cont/cr_box}
+        \caption{CrystalRouter}
+        \label{fig:cont-cr}
+    \end{subfigure}%
+    \begin{subfigure}[t]{0.32\textwidth}
+        \centering
+        \includegraphics[height=1.5in]{figs/intra-job/cont/mg_box}
+        \caption{MultiGrid}
+        \label{fig:cont-mg}
+    \end{subfigure}%
+   \caption{Intra-job study: data transfer time of 3 apps with Non/Contiguous allocation}
+   \label{fig: cont-time-box}
+\end{figure*}
+
+%Apparently, the size of allocation unit can affect application's performance.
+The good allocation unit size for each application is decided by application's communication pattern.
+Once the locality of application's communication pattern could be identified, the size of allocation unit should be able to preserve the locality of application.
+Figure \ref{fig: chunk-time-box} shows the data transfer time of 3 App with different allocation unit size.
+%
+%For AMG, unit size of 16 and 8, are equally good, since AMG's locality is 6-rank based.
+Unit size of 2 can cause performance loss to AMG.
+For CrystalRouter data transfer time increases as the unit size decreases from 16 to 8 and 2.
+Unit size of 16 will preserve not only the 5-rank local community in CrystalRouter communication pattern, but also part of the diagonal global data transfer.
+Unit size of 8 can only guarantee the locality of that 5-rank community.
+And unit size of 2 will lose all the locality of CrystalRouter and make the intra-job communication less efficient.
+For MultiGrid, smaller unit size prolong the data transfer time.
+The difference between different unit size is not obvious, since MultiGrid is nearly all-to-all dominant.
+%
+%
+%\begin{figure*}[t!]
+%    \centering
+%    \begin{subfigure}[t]{0.3\textwidth}
+%        \centering
+%        \includegraphics[height=1.2in]{figs/intra-job/chunk/amg_box}
+%        \caption{AMG}
+%        \label{fig:chunk-amg-box}
+%    \end{subfigure}%
+%    \hspace{1em}%
+%    \begin{subfigure}[t]{0.3\textwidth}
+%        \centering
+%        \includegraphics[height=1.2in]{figs/intra-job/chunk/cr_box}
+%        \caption{CrystalRouter}
+%        \label{fig:chunk-cr-box}
+%    \end{subfigure}%
+%    \hspace{1em}%
+%    \begin{subfigure}[t]{0.3\textwidth}
+%        \centering
+%        \includegraphics[height=1.2in]{figs/intra-job/chunk/mg_box}
+%        \caption{MultiGrid}
+%        \label{fig:chunk-mg-box}
+%    \end{subfigure}%
+%   \caption{Data Transfer Time of 3 Apps with Different Unit Size}
+%   \label{fig: chunk-time-box}
+%\end{figure*}
+
+\subsubsection{Inter-Job Interference Analysis}
+\label{sec: interjob interference study}
+
+We simulate three application running concurrently with different allocation to analysis the interference between them.
+First, we assign each application with contiguous allocation side by side, and compare their performance with each application running exclusively.
+For AMG, sharing the network with other application won't affect its performance, since all the data transfer are between neighboring ranks.
+CrystalRouter and MultiGrid will suffer prolonged communication time when sharing network with others.
+The results are shown in Figure \ref{fig: concurrent-cont-time-box}.
+
+Then, we assign each application with non-contiguous allocation and make them running concurrently.
+The allocation unit belongs to different jobs are interleaved as shown in Figure \ref{fig:noncont_sub1}.
+The unit size is critical to the application's performance.
+Figure \ref{fig: 3apps interjob study} shows the results of each application data transfer time with different allocation unit size.
+
+The data transfer time of AMG in Figure \ref{fig:interjob-amg-box} keeps stable between allocation unit size of 16(bit) and 8(medium), since the locality of AMG is 6-rank based.
+When the unit size reduce to 2(small), AMG suffers prolong data transfer time by about 10\%.
+CrystalRouter is more sensitive to allocation unit size.
+The best unit size would be the one big enough to accommodate its 5-rank locality.
+Figure \ref{fig:interjob-cr-box} shows unit size of 16 and 8 can guarantee the same average data transfer time, while some rank spend more time with allocation unit size 8 than 16.
+When the unit size reduce to 2-node, the communication become less efficiency and take 15\% more times for transferring data.
+
+The data transfer time of MultiGrid with different allocation unit sizes doesn't show obvious variability in Figure \ref{fig:interjob-mg-box}.
+This is because even big allocation unit size like 16 will still fail to preserve MultiGrid's ``Many-to-Many" pattern.
+The data transfer time is almost doubled when MultiGrid running concurrently with allocation unit size of 16, as shown in Figure \ref{fig:interjob-mg-box} .
+As the unit size decreases, the data transfer time keep growing, but in a slow way, which means in terms of preserving the locality of MultiGrid, allocation unit of size 16, 8 and 2 serve equally bad.
+
+\emph{OBSERVATION}
+The unit size in non-contiguous allocation policy should be chosen according to application's communication pattern.
+Inter-job interference is inevitable in non-contiguous based systems, but unit size that big enough to preserve the locality of the application will alleviate such interference and improve job performance.
+
+\begin{figure*}[t!]
+    \centering
+    \begin{subfigure}[t]{0.32\textwidth}
+        \centering
+        \includegraphics[height=1.5in]{figs/inter-job/amg}
+        \caption{AMG}
+        \label{fig:interjob-amg-box}
+    \end{subfigure}%
+    \hspace{1em}%
+    \begin{subfigure}[t]{0.32\textwidth}
+        \centering
+        \includegraphics[height=1.5in]{figs/inter-job/cr}
+        \caption{CrystalRouter}
+        \label{fig:interjob-cr-box}
+    \end{subfigure}%
+    \begin{subfigure}[t]{0.32\textwidth}
+        \centering
+        \includegraphics[height=1.5in]{figs/inter-job/mg}
+        \caption{MultiGrid}
+        \label{fig:interjob-mg-box}
+    \end{subfigure}%
+   \caption{Inter-job interference study.
+``cont" indicates three applications running side by side concurrently on the same network with contiguous allocation.
+To study the impact of non-contiguous allocation, applications running concurrently with interleaved allocation of different unit sizes, which are 16-node, 8-node, 2-node.
+}
+   \label{fig: 3apps interjob study}
+\end{figure*}
+
+\section{Discussion}
+\label{sec:discussion}
+HPC system usually accommodate many parallel jobs running concurrently.
+Each job gets the required number of computing nodes by system resource manager.
+The allocated nodes for each job communicate with each other through system interconnect network in a sharing way, which cause interference as a negative side-effect.
+This interference can introduce performance degradation to both job and system, such as prolonged job running time and reduced system throughput.
+
+The knowledge about HPC job's interference is extremely useful for improving system and job's performance.
+By scrutinizing job's communication behavior, we can locate the critical path with job's communication graph, identify  job's dominant communication pattern.
+With such knowledge about job's communication characteristics, we can analysis the possible interference between jobs through simulation or even predict such interference when job running on different allocations.
+
+The most promising way to utilize such knowledge is to make smart allocation based on job's dominant communication pattern.
+Most existing batch schedulers take so few initiative because it been hold under the obscurantism policy.
+Whether it use partition-based or non-contiguous policy, means it chooses the benefit of either system or job over the other.
+Some propose compromised policies \cite{lo}\cite{pascual}\cite{yang-cluster14} to use fix allocation unit size and  make the allocation for job blindly without knowing job's communication information.
+
+Based on the findings in this work, we propose the idea of smart job allocation with consideration about job's communication pattern.
+As the dominant communication pattern of parallel applications on HPC system can be summarized into a couple of categories, the deep understanding of one application can be helpful to all other applications that conforms to the same communication pattern.
+When an application with dominant communication that is intensive all-to-all submitted to the system, it should be granted with compact allocation with exclusive network provision, so that it won't suffer performance loss due to interference.
+If job with dominant communication pattern like Nearest Neighbor or others with intensive small scale communication, rather than assign it with unnecessary compact allocation and exclusive network provision, non-contiguous allocation with proper unit size would serve it equally well.
+No degradation will be introduced to job's performance and system utilization will be improved.
+
+The advantage of making allocation with consideration about job's communication pattern is obvious.
+First, compared with partition-based policy, the new allocation strategy is more flexible, it doesn't require the system to provide a big partition that can accommodate the whole application, just small set of compact nodes enough for all the "local community" in the application.
+On the other hand, the new allocation strategy won't destroy the locality of application's communication pattern.
+On the contrary, the small compact node set (allocation unit) provided for the application's  ``local communities"  will preserve the communication locality in the maximum extent.
+
+%The off-line study of job's communication traces could be helpful to develop better allocation policies for batch scheduler to make on-line scheduling/allocating decisions.
+This is due to the set of applications running on those leadership computing facilities is usually stable, and the workloads from those HPC systems is usually has great repetitiveness.
+We analyzed the workload on Mira, an IBM Blue/Gene Q system at Argonne Leadership Computing Facility, found that a large portion of its workload are repetitive.
+Figure \ref{fig: repetitiveness of Mira} shows the 5 of most active users with repetitive submissions.
+The collected jobs information from history can be very useful for workload prediction and system optimization.
+%
+%
+%\begin{figure}[h!] 
+%  \centering
+%  \includegraphics[width=0.38\textwidth]{figs/repe-mira}
+%   \caption{The repetitiveness of Mira's workload. }
+%   \label{fig: repetitiveness of Mira}
+%\end{figure}
 
@@
 \label{sec:related_work}
+%\textbf{P1: tools for monitoring and profiling}\\
+There are some tools available for system monitoring and application profiling on HPC systems.
+Tools like TAU (Tuning and Analysis) \cite{tau}, SCALASCA\cite{scala} and mpiP\cite{mpip} can capture application's runtime information about application's communication and computation in event traces.
+However, recognizing the patterns about application's communication from those traces requires lots of extra effort.
+Some researchers work on the recognition and characterization of parallel application communication patterns.
+Oak Ridge National Laboratory has this ongoing project about developing a tool set named Oxbow, which can characterize the computation and communication behavior of 12 scientific applications and benchmarks\cite{oxbow}.
+In their recent work\cite{hpdc2015-ornl}, they demonstrate a new approach to automatically characterizing a parallel application's communication behavior.
 
-P1: tools for monitoring and profiling
+%\textbf{P2: communication characteristics}\\
+Some research institutes put great effort in characterization of scientific applications.
+The Department of Energy initiate this design forward project aims at the identification of the computational characteristics of the DOE MiniApps developed at various exascale co-design centers\cite{design forward webpage}.
+In this project, the communication patterns of several DOE full applications and associated mini-applications are studied to provide a more complete snapshot of the DOE workload.
+A joint project named CORAL from Oak Ridge, Argonne and Livermore provide a series of benchmarks to represent DOE workloads and technical requirements\cite{coral}.
+The CORAL project includes scalable science benchmarks, throughput benchmarks, data centric benchmarks, skeleton benchmarks and Micro benchmarks.
 
-P2: communication pattern
+%\textbf{P3: Interference between concurrently running jobs}\\
+The interference between concurrently running jobs on HPC system have been identified as major culprit for job's performance variability.
+Abhinav et al.
+found that concurrently running applications on HPC can cause interference to each other, and cause communication time varied from 36\% faster to 69\% slower.
+Such interference could come from Operating System noise, shape of the allocated partition and mainly other running jobs that sharing network resources\cite{abhinav-sc13}.
+Skinner et al.
+found that  there is a 2-3$\texttimes$ slowdown in MPI$\textunderscore$Allreduce due to network contention from other jobs\cite{skinner}.
+Rosenthal et al.
+found there is limited benefit for many applications by increasing network bandwidth.
+The applications that send mostly small messages or larger messages asynchronously are not bandwidth bounded, hence, benefit only slightly or not at all from increased bandwidth\cite{rosenthal}.
 
-P3: Interference between concurrently running jobs
+%\textbf{P4: Job allocation(little bit)}\\
+There are some research work focus on optimizing job allocation on HPC system to alleviate the interference between concurrently running jobs.
+Hoefler et al.
+propose to use performance modeling techniques to analysis factors that impact the performance of parallel scientific applications\cite{hoefler-modeling}.
+However, as the scale of HPC continue grows, the  interference of concurrently running jobs is getting worse, which is hard to be quantified by performance profiling tools.
+Bogdan et al provide a set of guidelines how to configure a network with Dragonfly topology for workload with Nearest Neighbor communication pattern\cite{Bogdan-hpdc14}.
+Dong et al describe IBM Blue Gene/Q's 5D torus interconnect network\cite{Dong-SC11}.
+They developed simple benchmarks that conforms to four different communication patterns, namely ping-pong, nearest neighbor, broadcast and allreduce, to demonstrate the effectiveness of this highly parallel 5D torus network.
 
-P4: Job allocation(little bit)
+Our work is different from all these works in the following ways.
+First, we focus on the dominant communication patterns rather than any specific application.
+We believe this can provide a guideline for other research work is the area.
+Secondly, we explored the inter-job interference between concurrently running jobs, while similar work such as \cite{abhinav-sc13} only focus on the single application's performance degradation due to network contention.
+Finally, we explored the impact of different allocation strategies to job's communication behavior.
+We identified the optimal allocation strategy for each application with specific dominant communication pattern.
+Based on our comprehensive exploration, we claim that better allocation strategy should take job's communication pattern into consideration for allocation decision making.
 
@@
 \label{sec:conclusion}
+In this work, we study the communication behavior of three parallel applications, namely AMG, CrystalRouter and MultiGrid.
+Each application has distinctive communication pattern, that can be representative for a group of jobs in HPC workload.
+We use a sophisticate simulation tool named CODES from Argonne National Laboratory to simulate the running of these three parallel applications on torus network.
+The torus network provided by CODES has good fidelity and scalability.
+We analyzed the performance of each application's communication in terms of data transfer time by simulating them running on torus networks with different bandwidth and dimensionality configurations.
+We found that higher dimensionality of torus network would improve the performance of application with ``Many-to-Many" communication patterns, while application with intensive local communication like ``Nearest Neighbor" won't benefit much from higher dimensionality.
+
+We also analysis the intra- and inter-job interference by simulating three applications running on 3D torus network.
+Based on our comprehensive experiments, we got three observations.
+1) The compact allocation can not guarantee the best performance for every application.
+2) The performance of application conforms to Nearest Neighbor communication pattern keep relatively stable with different shape allocations, while application conforms to ``Many-to-Many" communication pattern prefer compact allocation.
+3) The allocation unit size should be decided according to application's communication pattern.
+Unit size that big enough to preserve the locality of the application will result better performance.
+
+We believe that our finding in this work can provide guidance for HPC resource management to make flexible job allocations.
+Rather than use pre-defined partitions and reckless non-contiguous allocation, future HPC system should assign each job with preferable resources based on job's communication pattern.
 
@@
 \label{sec: ack}
-%We appreciate the valuable comments and suggestions from the anonymous reviewers. 
-The work at Illinois Institute of Technology is supported in part by
-US National Science Foundation grant CNS-1320125. The work at Argonne is
-supported in part by the U.S. Department of Energy (DOE), Office of Science,
-under Contract DE-AC02-06CH11357.
+The work at Illinois Institute of Technology is supported in part by U.S.
+National Science Foundation grants CNS-1320125 and CCF-1422009.
+This work is also supported by the U.S.
+Department of Energy, Office of Science, Advanced Scientific Computing Research, under Contract DE-AC02-06CH11357.
 
@@
  %=====start of scheduling papers=====
-\bibitem{feit}
-D.~Feitelson and A.~Weil.
-\newblock Utilization and predictability in scheduling the {IBM} {SP2} with
-  backfilling.
-\newblock In {\em International Parallel and Distributed Processing Symposium}, 1998.
 
-\bibitem{wei-ipdps2010}
-W.~Tang, N.~Desai, D.~Buettner, and Z.~Lan.
-\newblock Analyzing and adjusting user runtime estimates to improve job
-  scheduling on the {Blue} {Gene/P}.
-\newblock In {\em 2010 IEEE International Symposium on Parallel Distributed
-  Processing}, 2010.
-
-\bibitem{wei-ipdps2011}
-W.~Tang, Z.~Lan, N.~Desai, D.~Buettner, and Y.~Yu.
-\newblock Reducing fragmentation on torus-connected supercomputers.
-\newblock In {\em 2011 IEEE International Symposium on Parallel Distributed
-  Processing Symposium}, 2011.
-
-\bibitem{wei-jpdc2013}
-W.~Tang, N.~Desai, D.~Buettner, and Z.~Lan. 
-\newblock Job Scheduling With Adjusted Runtime Estimates on Production Supercomputers.
-\newblock Journal of Parallel and Distributed Computing (JPDC), 73(7):926-938, 2013.
-
-\bibitem{zhou}
-Z.~Zhou, Z.~Lan, W.~Tang, and N.~Desai.
-\newblock Reducing energy costs for {IBM Blue} {Gene/P} via {Power-Aware} job
-  scheduling.
-\newblock In {\em 17th Workshop on Job Scheduling Strategies for Parallel
-  Processing}, 2013.
-
-\bibitem{yang-sc13}
-X. ~Yang, Z.~Zhou, S.~Wallace Z.~Lan, W.~Tang, S. ~Coghlan and Mike. ~Papka.
-\newblock Integrating Dynamic Pricing of Electricity into Energy Aware Scheduling for HPC Systems. 
-\newblock In {Proceedings of the International Conference on High Performance Computing, Networking, Storage and Analysis(SC), 2013 International Conference for}
-\newblock SC '13,
-\newblock 2013,
-\newblock Salt Lake City, Utah
-
-\bibitem{Cqsim}
-{\textcolor{red} {Cqsim: An event-driven simulator}}.
-\newblock http://bluesky.cs.iit.edu/cqsim
+%\bibitem{yang-sc13}
+%X. ~Yang, Z.~Zhou, S.~Wallace Z.~Lan, W.~Tang, S. ~Coghlan and Mike. ~Papka.
+%\newblock Integrating Dynamic Pricing of Electricity into Energy Aware Scheduling for HPC Systems. 
+%\newblock In {Proceedings of the International Conference on High Performance Computing, Networking, Storage and Analysis(SC), 2013 International Conference for}
+%\newblock SC '13,
+%\newblock 2013,
+%\newblock Salt Lake City, Utah
 
@@
 \bibitem{leung}
-Leung, V.J.; Arkin, E.M.; Bender, M.A.; Bunde, D.; Johnston, J.; 
-Alok Lal; Mitchell, J.S.B.; Phillips, C.; Seiden, S.S., 
+V.J. Leung,  E.M. Arkin, M.A. Bender,  D. Bunde, J. Johnston, etc,
 \newblock Processor allocation on Cplant: achieving general processor locality using one-dimensional allocation strategies
@@
 \newblock pp.140-148
+
+\bibitem{lo}
+V. Lo,  K. Windisch, W.  Liu, and Nitzberg.
+\newblock Noncontiguous processor allocation algorithms for mesh-connected multicomputers.
+\newblock IEEE Transactions on Parallel and Distributed Systems 8 (1997). pages 712--726
+
+\bibitem{pascual} 
+JA. Pascual, J. Navaridas and  J. Miguel-Alonso. 
+\newblock Effects of Topology-Aware Allocation Policies on Scheduling Performance. 
+\newblock 14th Workshop on Job Scheduling Strategies for Parallel Processing. May 29, 2009.Rome, Italy.
 
@@
 \newblock Available: http://portal.nersc.gov/project/CAL/trace.htm
+\newblock (Accessed on: Sept 16th. 2015)
+
+\bibitem{coral}
+CORAL collaboration benchmark codes [Online].
+\newblock Available:https://asc.llnl.gov/CORAL-benchmarks/
 \newblock (Accessed on: Sept 16th. 2015)
@@
 %=====end of application/mapping papers=====
-\bibitem{Pjesivac}
-Pje\v{s}ivac-Grbovi\'{c}, Jelena and Angskun, Thara and Bosilca, George and Fagg, Graham E. and Gabriel, Edgar and Dongarra, Jack J.
-\newblock Performance Analysis of MPI Collective Operations,
-\newblock Cluster Computing,
-\newblock June 2007,
-\newblock volume 10,
-\newblock numpages 17,
 
-\bibitem{thakur}
-Thakur, Rajeev and Gropp, William D.
-\newblock Improving the performance of collective operations in MPICH,
-\newblock Recent Advances in Parallel Virtual Machine and Message Passing Interface,
-\newblock year 2003,
-\newblock Springer
+\bibitem{RF}
+Fiedler, Robert, and Stephen Whalen. 
+\newblock "Improving task placement for applications with 2D, 3D, and 4D virtual Cartesian topologies on 3D torus networks with service nodes." 
+\newblock Proc. Cray User’s Group (2013).
 
-\bibitem{abhinav-ipdps15}
-Abhinav Bhatele, Andrew R. Titus, Jayaraman J. Thiagarajan, Nikhil Jain, Todd Gamblin,
-Peer-Timo Bremer, Martin Schulz† and Laxmikant V. Kale,
-\newblock "Identifying the Culprits behind Network Congestion",
-\newblock In {\em 2015 IEEE International Symposium on Parallel Distributed Processing Symposium}, 2015.
+\bibitem{abhinav-sc13} 
+Bhatele, Abhinav and Mohror, Kathryn and Langer, Steven
+H. and Isaacs, Katherine E. 
+\newblock There Goes the Neighborhood: Performance Degradation Due to Nearby Jobs 
+\newblock In {Proceedings of SC13: International
+Conference for High Performance Computing, Networking, Storage and Analysis},
+2013
 
-\bibitem{langer}
-S. Langer, A. Bhatele, and C. H. Still,
-\newblock “pF3D simulations of laser-plasma interactions in National Ignition Facility experiments,” 
-\newblock Computing in Science and Engineering, vol. 99, Aug. 2014, lLNL- JRNL-648736. 
-\newblock [Online]. Available: http://doi.ieeecomputersociety.org/ 10.1109/MCSE.2014.79
+\bibitem{jose-ipdps15}
+Jokanovic, A.; Sancho, J.C.; Rodriguez, G.; Lucero, A.; Minkenberg, C.; Labarta, J.,
+\newblock "Quiet Neighborhoods: Key to Protect Job Performance Predictability," 
+\newblock in Parallel and Distributed Processing Symposium (IPDPS), 
+\newblock 2015 IEEE International , vol., no., pp.449-459, 25-29 May 2015
 
-%\bibitem{abhatele-ipdps08}
-%A. Bhatele, L.V.  Kale
-%\newblock Application-specific topology-aware mapping for three dimensional topologies.
-%\newblock In {\em Proceedings of Workshop on Large-Scale Parallel Processing (held as part of IPDPS' 08)} 2008
-%
-%\bibitem{abhinav-sc13} 
-%Bhatele, Abhinav and Mohror, Kathryn and Langer, Steven
-%H. and Isaacs, Katherine E. 
-%\newblock There Goes the Neighborhood: Performance Degradation Due to Nearby Jobs 
-%\newblock In {Proceedings of SC13: International
-%Conference for High Performance Computing, Networking, Storage and Analysis},
-%2013
+\bibitem{skinner}
+D. Skinner and W. Kramer. 
+\newblock Understanding the Causes of Performance Variability in HPC Workloads. 
+\newblock In Proceedings of the IEEE International Workload Characterization Symposium, 
+\newblock 2005, pages 137–149, 2005.
 
@@
 \newblock Salt Lake City, Utah
+
+\bibitem{ozan}
+Tuncer, Ozan, Vitus J Leung, Ayse K. Coskun,
+\newblock "PaCMap: Topology Mapping of Unstructured Communication Patterns onto Non-contiguous Allocations," 
+\newblock 29th International Conference on Supercomputing, June 2015.
 
@@
 
+\bibitem{oxbow}
+J. S. Vetter, S. Lee, D. Li, G. Marin, C. McCurdy, J. Meredith, P. C. Roth, and K. Spafford. 
+\newblock Quantifying architectural requirements of contemporary extreme-scale scientific applications. 
+\newblock In International Workshop on Performance Modeling, Benchmarking and Simulation of HPC Systems (PMBS13), 
+\newblock Denver, Colorado, 2013.
+
 \bibitem{Dong-SC11}
@@
 Jason Cope, Ning Liu, Sam Lang, Phil Carns, Chris Carothers, Robert Ross.
-\newblock Proceedings of the Workshop on Emerging Supercomputing Technologies 2011,
+\newblock CODES: Enabling Co-design of Multilayer Exascale Storage Architectures, 
+\newblock Proceedings of the Workshop on Emerging Supercomputing Technologies 2011, pp. 303-312
 \newblock 2011
@@
 
+\bibitem{codes-torus}
+Mubarak, Misbah, et al.
+\newblock "A case study in using massively parallel simulation for extreme-scale torus network codesign." 
+\newblock in Proceedings of the 2nd ACM SIGSIM/PADS conference on Principles of advanced discrete simulation. 
+\newblock ACM, 2014.
+
+\bibitem{bgq}
+C. Dong, N. Eisley, P. Heidelberger, R. Senger, Y. Sugawara, S. Kumar, V. Salapura, D. Satterfield, B. Steinmacher-Burow, and J. Parker, 
+\newblock “The IBM Blue Gene/Q interconnection fabric,” 
+\newblock Micro, IEEE, vol. 32, no. 1, pp. 32–43, 2012.
+
+\bibitem{tofu}
+Y. Ajima, Y. Takagi, T. Inoue, S. Hiramoto, and T. Shimizu, 
+\newblock “The Tofu Interconnect,” 
+\newblock in 2011 IEEE 19th Annual Symposium on High Performance Interconnects (HOTI), 2011, pp. 87–94.
+
+\bibitem{titan}
+Titan System Overview [online]
+\newblock Available: https://www.olcf.ornl.gov/kbarticles/titan-system-overview
+\newblock (Accessed on: OCT 4th, 2015)
+
 \bibitem{sst}
@@
 
+\bibitem{simgrid}
+SIMGRID Versatile Simulation of Distributed Systems [Online].
+\newblock Availabale: http://simgrid.gforge.inria.fr
+\newblock (Accessed on: Oct 4th, 2015)
+
+\bibitem{top500}
+“Top500 supercomputing web site.” [Online]. 
+\newblock Available: http://www.top500.org
+\newblock (Accessed on: Oct 4th, 2015)
+
 \end{thebibliography}
 
+\vspace{5\baselineskip}
+The submitted manuscript has been created by UChicago Argonne, LLC, Operator of Argonne National Laboratory ("Argonne").
+Argonne, a U.S.
+Department of Energy Office of Science laboratory, is operated under Contract No.
+DE-AC02-06CH11357.
+The U.S.
+Government retains for itself, and others acting on its behalf, a paid-up nonexclusive, irrevocable worldwide license in said article to reproduce, prepare derivative works, distribute copies to the public, and perform publicly and display publicly, by or on behalf of the Government.
+
 \end{document}

### 2015-10-13 submission.tex
chages based on MM's comments

@@
 
-It has been widely recognized that network contention between concurrently running jobs in HPC system is one of the primary causes of performance variability.
-Optimizing job allocation and avoiding network sharing are proven techniques to alleviating network contention and performance degradation.
-In this work, we propose a novel job communication pattern aware allocation strategy which serve as a critical module in our future HPC batch scheduler.
-The novelty of our new strategy is that it makes allocation decisions based on job’s communication pattern communication pattern in addition to system topology and resource availability information.
-The scheduler uses job communication patterns information to decide whether a job’s performance relies heavily on the network or not, and what the potential inter/intra-job interference would be for a given allocation.
+Network contention between concurrently running jobs on HPC system is one of the primary causes of performance variability.
+Optimizing job allocation and avoiding network sharing are proven techniques to alleviate network contention and performance degradation.
+In this work, we propose a novel job communication pattern aware allocation strategy which serves as a critical module in our future HPC batch scheduler.
+The novelty of our new strategy is that it makes allocation decisions based on job’s communication pattern in addition to system topology and resource availability information.
+The scheduler uses job communication patterns information to decide whether its performance relies heavily on the network or not, and what the potential inter/intra-job interference would be for a given allocation.
 Thus, the scheduler can pick a preferable resource allocation based on a specific job's communication pattern.
-Using traces have been collected from three representative parallel applications, we validate the effectiveness of our new design.
-The idea presented in this paper is widely applicable: HPC parallel applications have common communication patterns, and allocating resource with communication pattern awareness can improve performance on systems with a variety of network topologies.
+We validate the effectiveness of our new design using traces collected from three representative parallel applications.
+The idea presented in this paper is widely applicable: HPC parallel applications have common communication patterns, and allocating resource with communication pattern awareness can improve performance on these systems with a variety of network topologies.
 
@@
 The scale of supercomputers keeps growing to accommodate the the demand for computing power required by scientific research areas.
-Supercomputers contain hundreds of thousands of nodes, and serve as irreplaceable research vehicle for scientific problems with increasing size and complexity.
-Supercomputers are employed as shared resource to accommodate many jobs running concurrently\ref{fig: overview} and maintain high system utilization.
-These jobs share the system infrastructure such as network and I/O bandwidth, and inevitably there is contention over the shared resources.
+Supercomputers have hundreds of thousands of nodes, and serve as irreplaceable research vehicle for scientific problems with increasing size and complexity.
+Supercomputers are usually employed as a shared resource to accommodate many parallel applications (also known as jobs) running concurrently \cite{zhou-ipdps} and maintain high system utilization.
+These communication and I/O intensive jobs share the system infrastructure such as network and I/O bandwidth, and inevitably there is contention over these shared resources.
 As supercomputers continue to evolve, these shared resources are increasingly the bottleneck for performance.
@@
    \caption{Indication of multiple jobs running concurrently with different allocations.
-Different job represented by specific color.
-a) shows the effect of contiguous allocation, which would alleviate the inter-job interference.
-b) shows non-contiguous allocation, which definitely introduce both intra and inter-job interference.
+Each job is represented by a specific color.
+a) shows the effect of partition-based allocation, which reduce the inter-job interference.
+b) shows non-contiguous allocation, which introduces both intra and inter-job interferences.
 }
@@
 
-One of the most prominent  problems is network contention between concurrently running jobs.
-The network sharing between concurrently running jobs causes communication variability that results in jobs running slower\cite{abhinav-sc13}.
-The delay of currently running jobs increases the queueing time of the following submitted jobs, thus leading to low system throughput, low utilization\cite{jose-ipdps15}.
-
-This adverse effect of network sharing can be mitigated by providing jobs with better node allocations.
+One of the most prominent  problems is network contention among concurrently running jobs.
+The network sharing among concurrently running jobs causes communication variability that results in jobs running slower \cite{abhinav-sc13}.
+The delay of running jobs increases the queueing time of the following submitted jobs, thus leading to low system throughput and utilization \cite{jose-ipdps15}.
+
+This adverse effect of network sharing can be mitigated by providing jobs with isolated allocations and exclusive network resource.
 The batch scheduler is responsible for assigning job the amount of nodes it requires.
-Currently, two allocation strategies are commonly used on torus-connected systems.
-One is so called partition based allocation, where the scheduler assigns each user job a compact and contiguous set of computing nodes.
-IBM Blue Gene series systems use this strategy.
+On the widely used torus-connected HPC systems \cite{bgq}\cite{tofu}\cite{titan}, two allocation strategies are commonly used.
+In the partition based allocation, used in Blue Gene series systems, he scheduler assigns each user job a compact and contiguous set of computing nodes.
 This strategy favors application performance by providing application with isolated partitions and exclusive network connections, thus reducing network contention caused by concurrently running jobs.
-However, this strategy can cause internal fragmentation (when more nodes are allocated to a job than it requests) and external fragmentation (when sufficient nodes are available for a request, but they can not be allocated contiguously), therefore leading to low system utilization than otherwise possible.
-The other strategy is non-contiguous allocation, where free nodes are assigned to user jobs whether they are contiguous or not.
-Cray XT/XE series systems use this strategy.
+However, this strategy can cause internal fragmentation (when more nodes are allocated to a job than it requests) and external fragmentation (when sufficient nodes are available for a request, but they can not be allocated contiguously), therefore leading to lower system utilization that is otherwise possible.
+The other strategy is non-contiguous allocation, used by Cray XT/XE series, in which free nodes are assigned to user jobs whether they are contiguous or not.
 Non-contiguous allocation eliminates internal and external fragmentation as seen in partition-based systems, thereby leading to high system utilization.
 However, it introduces other problems, for example scattering application processes all over the system, causing both inter-job and intra-job contention.
-The non-contiguous node allocation can significantly reduce job performance, especially for those communication-intensive jobs.
+The non-contiguous node allocation can significantly reduce job performance, especially for communication-intensive jobs.
 
@@
 %=========come back and work from here=============
-Rather than choose blindly between partition-based or non-contiguous allocation, the new allocation strategy choses resources based on a job's communication pattern.
-With detail analysis of application's communication patterns, a clear view of its communication topology is obtained shows which processes are tightly connected and conduct intensive communication.
-The ``local community" consists of such tightly coupled processes should get a compact allocation.
-Then, the allocation strategy only needs to guarantee the locality of the subset of the nodes where those ``local communities" would reside.
-
-In this work, we focus on the study of impact of different allocation strategies on intra/inter-job interference on Torus network.
+Rather than choosing blindly between partition-based or non-contiguous allocation, the new allocation strategy choses resources based on a job's communication pattern.
+With detail analysis of application's communication patterns, our proposed allocation strategy obtains a clear view as to which processes are tightly connected and conduct intensive communication.
+The ``neighborhood communication" consists of such tightly coupled processes should get a compact allocation.
+Then, the allocation strategy only needs to guarantee the locality of the subset of the nodes where those ``neighborhood communication" would reside.
+
+In this work, we study the impact of different allocation strategies on intra/inter-job interference on Torus network.
 Torus networks have been extensively used in the current generation of supercomputers because of their linear scaling on per-node cost and competitive communication performance.
 On the recent Top500 list, 6 of the top 10 supercomputers use a high-radix torus-interconnected network \cite{top500}.
-The current generation of IBM Blue Gene/Q (BG/Q) supercomputer, such as Mira at Argonne Nation Lab and Sequoia at  Lawrence Livermore National Laboratory, has its nodes connected in a 5D torus network \cite{bgq}.
+The current generation of IBM Blue Gene/Q (BG/Q) supercomputer, such as Mira at Argonne Nation Laboratory and Sequoia at  Lawrence Livermore National Laboratory, has its nodes connected in a 5D torus network \cite{bgq}.
 The K computer from Japan uses the “Tofu” system, which is a 6D mesh/torus topology \cite{tofu}.
@@
 
-We selected three signature applications from DOE Design Forward Project\cite{design forward webpage} as examples to conduct deep study about their communication patterns.
-We use a sophisticated simulation toolkit named CODES, Co-Design of Multi-layer Exascale Storage Architecture\cite{Jason-2011}, from Argonne National Laboratory as research vehicle to evaluate the performance of our communication pattern aware allocation policy on torus networks.
-We analysis each application's performance when it running on torus networks with different dimensionality and interconnected bandwidth.
-Then, we extend CODES with a Job Mapping API that support different allocation policies on torus network.
+We selected three signature applications from DOE Design Forward Project\cite{design forward webpage} as examples to conduct detailed study about their communication patterns.
+We use a sophisticated simulation toolkit named CODES, Co-Design of Multi-layer Exascale Storage Architecture\cite{Jason-2011}, from Argonne National Laboratory as a research vehicle to evaluate the performance of our communication pattern aware allocation policy on torus networks.
+We analyze each application's performance running on torus networks with different dimensionality and interconnected bandwidth.
+Then, we extend CODES with a Job Mapping API that supports different allocation policies on torus network.
 Different allocation strategies on torus network are studied with the help of this new Job Mapping API.
-We analysis the intra/inter-job interference by simulating those three applications running exclusively or concurrently on torus network with different allocation strategies.
+We analyze the intra/inter-job interference by simulating those three applications running exclusively or concurrently on torus network with different allocation strategies.
 According to results of our comprehensive exploration, making allocation with consideration about job's communication pattern is a very promising allocation strategy.
@@
 Section \ref{sec:codes} talks about CODEs as research vehicle for our work.
-Section \ref{sec:torus} shows the performance analysis of three applications on different torus networks.
+Section \ref{sec:config study} shows the performance analysis of three applications on different torus networks.
 Section \ref{sec:interference} provides detailed analysis about the intra/inter-job interference between three applications on torus network.
@@
     \end{subfigure}
-    \caption{AMG. (a) rank-to-rank communication topology graph. (b) the data amount send/received by each rank in AMG.}
+    \caption{AMG. (a) rank-to-rank communication topology graph. (b) the data sent/received by each rank in AMG.}
 \end{figure}
@@
     \end{subfigure}
-    \caption{CrystalRouter. (a) rank-to-rank communication topology graph. (b) the data amount send/received by each rank in CrystalRouter. }
+    \caption{CrystalRouter. (a) rank-to-rank communication topology graph. (b) the data sent/received by each rank in CrystalRouter. }
 \end{figure}
@@
     \end{subfigure}
-    \caption{MultiGrid. (a) rank-to-rank communication topology graph. (b) the data amount send/received by each rank in MultiGrid.}
+    \caption{MultiGrid. (a) rank-to-rank communication topology graph. (b) the data sent/received by each rank in MultiGrid.}
 \end{figure}
@@
 \subsection{Configuration Study }
-\label{sec: config study}
+\label{sec:config study}
 

### 2015-10-13 submission.tex
minor change

@@
 
-A parallel application usually conforms to a combination of several basic communication patterns\cite{hpdc2015-ornl}.
+A parallel application usually conforms to a combination of several basic communication patterns \cite{hpdc2015-ornl}.
 At its different execution phases, the application's communication behavior may follow different basic patterns respectively.
@@
 Nearest Neighbor communication is commonly seen in lots of structured and unstructured mesh physics packages.
-There are many profiling tools available to capture communication patterns from parallel applications, such as Tuning and Analysis Utilities(TAU)\cite{tau}, mpiP\cite{mpip}.
+There are many profiling tools available to capture communication patterns from parallel applications, such as Tuning and Analysis Utilities(TAU) \cite{tau}, mpiP \cite{mpip}.
 They can help analysis parallel application, providing information like the percentage of different MPI operations of the applications, communication topology, the amount data transferred between processes.
@@
 In this work, we provide detailed analysis about the communication pattern of three DOE MiniApps.
-MiniApps are reduced proxy applications that encapsulate the salient performance of larger full size applications\cite{miniapp}.
+MiniApps are reduced proxy applications that encapsulate the salient performance of larger full size applications \cite{miniapp}.
 Since we only focus on the dominant communication pattern of parallel applications, these MiniApps are perfect candidates for analysis the application's performance on different allocations or on new network architectures.
@@
 Their representative communication patterns that are commonly seen in the HPC system workload.
-The communication pattern figures we present here are generated with the IPM\cite{ipm} data from \cite{design forward webpage}.
+The communication pattern figures we present here are generated with the IPM \cite{ipm} data from \cite{design forward webpage}.
 
@@
 AMG is the MiniApp of Algebraic Multigrid Solver, which is a parallel algebraic multigrid solver for linear systems arising from problems on unstructured mesh physics packages.
-It has been derived directly from the BoomerAMG solver that is being developed in the Center for Applied Scientific Computing (CASC) at LLNL\cite{amg}.
+It has been derived directly from the BoomerAMG solver that is being developed in the Center for Applied Scientific Computing (CASC) at LLNL \cite{amg}.
 The dominant communication pattern is this regional communication with decreasing message size for different parts of the multigrid v-cycle.
@@
 AMG represents a group of applications whose dominant communication pattern is Nearest Neighbor.
-Applications with such similar dominant communication pattern are like PARTISN, SNAP\cite{design forward webpage}.
+Applications with such similar dominant communication pattern are like PARTISN, SNAP \cite{design forward webpage}.
 
@@
 The second MiniApp we have is Crystal Router, which is the extracted communication kernel of the full application Nek5000.
-Nek5000\cite{crystalrouter} is a spectral element CFD application developed at Argonne National Laboratory.
+Nek5000 \cite{crystalrouter} is a spectral element CFD application developed at Argonne National Laboratory.
 It features spectral element multigrid solvers coupled with a highly scalable, parallel coarse-grid solver that widely used for projects including ocean current modeling, thermal hydraulics of reactor cores, and spatiotemporal chaos.
@@
 It is widely used for structured grid physics packages.
-Figure\ref{fig: mg communication topology} shows the communication topology of MultiGrid and Figure\ref{fig: mg data trans} shows the the amount of data transferred between ranks in MultiGrid.
+Figure \ref{fig: mg communication topology} shows the communication topology of MultiGrid and Figure\ref{fig: mg data trans} shows the the amount of data transferred between ranks in MultiGrid.
 
@@
 There are couple of simulation platform to study the behavior of large scale systems.
-SimGrid from Inria is a versatile simulation tool sets that provide simulation for Grids, Clouds and HPC systems\cite{simgrid}.
+SimGrid from Inria is a versatile simulation tool sets that provide simulation for Grids, Clouds and HPC systems \cite{simgrid}.
 The Structural Simulation Toolkit (SST) from Sandia National Lab was developed to explore innovations in highly concurrent systems where the ISA, microarchitecture, and memory interact with the programming model and communications system\cite{sst}.
 
-A simulation toolkit named CODES (Co-Design of Multi-layer Exascale Storage Architecture) from Argonne National Lab enables the exploration of simulating different HPC networks with high fidelity\cite{Jason-2011}\cite{mubarak-sc2012}.
-CODES is built on top of Rensselaer Optimistic Simulation System (ROSS) parallel discrete-event simulator, which is capable of processing billions of events per second on leadership-class supercomputers\cite{ross}.
+A simulation toolkit named CODES (Co-Design of Multi-layer Exascale Storage Architecture) from Argonne National Lab enables the exploration of simulating different HPC networks with high fidelity \cite{Jason-2011}\cite{mubarak-sc2012}.
+CODES is built on top of Rensselaer Optimistic Simulation System (ROSS) parallel discrete-event simulator, which is capable of processing billions of events per second on leadership-class supercomputers \cite{ross}.
 CODES support both torus and dragonfly network with high fidelity flit-level simulation.
@@
 CODES has this network workload component that capable of conducting trace-driven simulation.
-It can take real MPI application trace generated by SST DUMPI\cite{sst} to drive CODES network models.
+It can take real MPI application trace generated by SST DUMPI \cite{sst} to drive CODES network models.
 We implemented a Job Mapping API for CODES, which can provide flexible job allocation strategies.
@@
 
-CODES can provide Torus network model with good scalability and high fidelity\cite{codes-torus}.
+CODES can provide Torus network model with good scalability and high fidelity \cite{codes-torus}.
 In this section, we will first study the communication behavior of three applications on torus network with different dimensionality and bandwidth configurations.
@@
 The \textbf{\emph{intra-job interference}} refers to the network contention between ranks within each application.
-There are lots of research works\cite{jingjin}\cite{ozan} try to come up with better task mapping algorithms to alleviate the intra-job interference, which is out the scope of our work.
+There are lots of research works \cite{jingjin}\cite{ozan} try to come up with better task mapping algorithms to alleviate the intra-job interference, which is out the scope of our work.
 In this work, we focus on the analysis of such interference with different allocations.
@@
 \textbf{\emph{Inter-job interference}} is introduced by concurrently running jobs that adjacent to each other sharing network resources.
-Such interference has been identified as one of the major culprits for application's performance variability\cite{abhinav-sc13}\cite{skinner}\cite{rosenthal}.
+Such interference has been identified as one of the major culprits for application's performance variability \cite{abhinav-sc13}\cite{skinner}\cite{rosenthal}.
 Inter-job interference is a more prominent issue for system adopted non-contiguous allocation policy than partition-based system.
-Application's communication time can vary from 36\% faster to 69\% slower due to inter-job interference when running on non-contiguous allocation\cite{abhinav-sc13}.
+Application's communication time can vary from 36\% faster to 69\% slower due to inter-job interference when running on non-contiguous allocation \cite{abhinav-sc13}.
 
@@
 The network connections in the \emph{y}direction have only half the bandwidth of the cables used in the \emph{x} and \emph{z} directions.
-In order to take advantage of the faster links in the \emph{x} and \emph{z} directions, job allocation is always start from X-Z plane, which leads to a rectangular prism shaped allocation\cite{RF}.
+In order to take advantage of the faster links in the \emph{x} and \emph{z} directions, job allocation is always start from X-Z plane, which leads to a rectangular prism shaped allocation \cite{RF}.
 
@@
 
-\emph{OBSERVATION}
+%\emph{OBSERVATION}
 Compact allocation should be provided to application with global data transfer, such as those conforms to ``Many-to-Many" communication pattern.
@@
 
-\emph{OBSERVATION}
+%\emph{OBSERVATION}
 The performance of application conforms to Nearest Neighbor communication pattern keep relatively stable with different allocations.
@@
 
-\emph{OBSERVATION}
+%\emph{OBSERVATION}
 The unit size in non-contiguous allocation policy should be chosen according to application's communication pattern.
@@
 There are some tools available for system monitoring and application profiling on HPC systems.
-Tools like TAU (Tuning and Analysis) \cite{tau}, SCALASCA\cite{scala} and mpiP\cite{mpip} can capture application's runtime information about application's communication and computation in event traces.
+Tools like TAU (Tuning and Analysis) \cite{tau}, SCALASCA \cite{scala} and mpiP \cite{mpip} can capture application's runtime information about application's communication and computation in event traces.
 However, recognizing the patterns about application's communication from those traces requires lots of extra effort.
 Some researchers work on the recognition and characterization of parallel application communication patterns.
-Oak Ridge National Laboratory has this ongoing project about developing a tool set named Oxbow, which can characterize the computation and communication behavior of 12 scientific applications and benchmarks\cite{oxbow}.
-In their recent work\cite{hpdc2015-ornl}, they demonstrate a new approach to automatically characterizing a parallel application's communication behavior.
+Oak Ridge National Laboratory has this ongoing project about developing a tool set named Oxbow, which can characterize the computation and communication behavior of 12 scientific applications and benchmarks \cite{oxbow}.
+In their recent work \cite{hpdc2015-ornl}, they demonstrate a new approach to automatically characterizing a parallel application's communication behavior.
 
@@
 Some research institutes put great effort in characterization of scientific applications.
-The Department of Energy initiate this design forward project aims at the identification of the computational characteristics of the DOE MiniApps developed at various exascale co-design centers\cite{design forward webpage}.
+The Department of Energy initiate this design forward project aims at the identification of the computational characteristics of the DOE MiniApps developed at various exascale co-design centers \cite{design forward webpage}.
 In this project, the communication patterns of several DOE full applications and associated mini-applications are studied to provide a more complete snapshot of the DOE workload.
-A joint project named CORAL from Oak Ridge, Argonne and Livermore provide a series of benchmarks to represent DOE workloads and technical requirements\cite{coral}.
+A joint project named CORAL from Oak Ridge, Argonne and Livermore provide a series of benchmarks to represent DOE workloads and technical requirements \cite{coral}.
 The CORAL project includes scalable science benchmarks, throughput benchmarks, data centric benchmarks, skeleton benchmarks and Micro benchmarks.
@@
 found that concurrently running applications on HPC can cause interference to each other, and cause communication time varied from 36\% faster to 69\% slower.
-Such interference could come from Operating System noise, shape of the allocated partition and mainly other running jobs that sharing network resources\cite{abhinav-sc13}.
+Such interference could come from Operating System noise, shape of the allocated partition and mainly other running jobs that sharing network resources \cite{abhinav-sc13}.
 Skinner et al.
-found that  there is a 2-3$\texttimes$ slowdown in MPI$\textunderscore$Allreduce due to network contention from other jobs\cite{skinner}.
+found that  there is a 2-3$\texttimes$ slowdown in MPI$\textunderscore$Allreduce due to network contention from other jobs \cite{skinner}.
 Rosenthal et al.
 found there is limited benefit for many applications by increasing network bandwidth.
-The applications that send mostly small messages or larger messages asynchronously are not bandwidth bounded, hence, benefit only slightly or not at all from increased bandwidth\cite{rosenthal}.
+The applications that send mostly small messages or larger messages asynchronously are not bandwidth bounded, hence, benefit only slightly or not at all from increased bandwidth \cite{rosenthal}.
 
@@
 Hoefler et al.
-propose to use performance modeling techniques to analysis factors that impact the performance of parallel scientific applications\cite{hoefler-modeling}.
+propose to use performance modeling techniques to analysis factors that impact the performance of parallel scientific applications \cite{hoefler-modeling}.
 However, as the scale of HPC continue grows, the  interference of concurrently running jobs is getting worse, which is hard to be quantified by performance profiling tools.
-Bogdan et al provide a set of guidelines how to configure a network with Dragonfly topology for workload with Nearest Neighbor communication pattern\cite{Bogdan-hpdc14}.
-Dong et al describe IBM Blue Gene/Q's 5D torus interconnect network\cite{Dong-SC11}.
+Bogdan et al provide a set of guidelines how to configure a network with Dragonfly topology for workload with Nearest Neighbor communication pattern \cite{Bogdan-hpdc14}.
+Dong et al describe IBM Blue Gene/Q's 5D torus interconnect network \cite{Dong-SC11}.
 They developed simple benchmarks that conforms to four different communication patterns, namely ping-pong, nearest neighbor, broadcast and allreduce, to demonstrate the effectiveness of this highly parallel 5D torus network.

### 2015-10-14 submission.tex
changes based on Lan's comments

@@
 % can use linebreaks \\ within to get better formatting as desired
-\title{Study of Intra- and Inter-Job Interference for Improving Allocation on HPC Systems}
+\title{Study of Intra- and Inter-Job Interference with Different Job allocations}
 % author names and affiliations
@@
 However, it introduces other problems, for example scattering application processes all over the system, causing both inter-job and intra-job contention.
-The non-contiguous node allocation can significantly reduce job performance, especially for communication-intensive jobs.
+The non-contiguous node allocation can significantly reduce job performance, especially for communication-intensive ones.
 
@@
 
-%=========come back and work from here=============
-Rather than choosing blindly between partition-based or non-contiguous allocation, the new allocation strategy choses resources based on a job's communication pattern.
-With detail analysis of application's communication patterns, our proposed allocation strategy obtains a clear view as to which processes are tightly connected and conduct intensive communication.
-The ``neighborhood communication" consists of such tightly coupled processes should get a compact allocation.
-Then, the allocation strategy only needs to guarantee the locality of the subset of the nodes where those ``neighborhood communication" would reside.
-
-In this work, we study the impact of different allocation strategies on intra/inter-job interference on Torus network.
+We envision that future HPC will adopt a flexible job allocation mechanism which combines the best of both partition-based allocation and non-contiguous allocation.
+Such flexible mechanism would make allocation for jobs based on their communication patterns.
+With detail analysis of application's communication patterns, it can be identified which processes are tightly connected and conduct intensive communication within each job.
+Rather than assign each job a compact partition-based allocation or non-contiguous allocation, the ``neighborhood communication" consists of such tightly coupled processes should get a compact allocation.
+The new allocation mechanism only needs to guarantee the locality of the subset of the nodes where those ``neighborhood communication" would reside.
+
+In this work, we focus on an in-depth analysis of intra- and inter-job communication interferences with  different job allocations on torus-connected HPC systems.
 Torus networks have been extensively used in the current generation of supercomputers because of their linear scaling on per-node cost and competitive communication performance.
@@
 We selected three signature applications from DOE Design Forward Project\cite{design forward webpage} as examples to conduct detailed study about their communication patterns.
-We use a sophisticated simulation toolkit named CODES, Co-Design of Multi-layer Exascale Storage Architecture\cite{Jason-2011}, from Argonne National Laboratory as a research vehicle to evaluate the performance of our communication pattern aware allocation policy on torus networks.
-We analyze each application's performance running on torus networks with different dimensionality and interconnected bandwidth.
-Then, we extend CODES with a Job Mapping API that supports different allocation policies on torus network.
-Different allocation strategies on torus network are studied with the help of this new Job Mapping API.
-We analyze the intra/inter-job interference by simulating those three applications running exclusively or concurrently on torus network with different allocation strategies.
-According to results of our comprehensive exploration, making allocation with consideration about job's communication pattern is a very promising allocation strategy.
-We believe the idea presented in this work would be the guideline for the design of future HPC batch job scheduler and resource management module.
+We use a sophisticated simulation toolkit named CODES, Co-Design of Multi-layer Exascale Storage Architecture\cite{Jason-2011}, from Argonne National Laboratory as a research vehicle to evaluate the performance of these applications with various allocations.
+We analyze the intra- and inter-job interference by simulating applications running exclusively or concurrently with different allocations.
+
+\textbf{replace this by summarizing a couple of key findings!!} According to results of our comprehensive exploration, making allocation with consideration about job's communication pattern is a very promising allocation strategy.
 
@@
 At its different execution phases, the application's communication behavior may follow different basic patterns respectively.
-Collective communication is often used for data distribution and collection among processes.
-Nearest Neighbor communication is commonly seen in lots of structured and unstructured mesh physics packages.
-There are many profiling tools available to capture communication patterns from parallel applications, such as Tuning and Analysis Utilities(TAU) \cite{tau}, mpiP \cite{mpip}.
-They can help analysis parallel application, providing information like the percentage of different MPI operations of the applications, communication topology, the amount data transferred between processes.
-
-In this work, we provide detailed analysis about the communication pattern of three DOE MiniApps.
-MiniApps are reduced proxy applications that encapsulate the salient performance of larger full size applications \cite{miniapp}.
+There are many profiling tools \cite{tau}\cite{mpip}\cite{scala}\cite{sst}\cite{oxbow} available to capture information regarding communication patterns of parallel applications.
+They can help analyze parallel applications' communication behavior, providing information like the percentage of different MPI operations of the applications, communication topology, the amount data transferred between processes.
+
+There are many applications running on HPC systems like Mira \cite{bgq} and Titan \cite{titan}.
+Most of these applications are parallel and communication intensive, and conform to different communication patterns.
+In this work, we select three representative applications from DOE Design Forward project \cite{design forward webpage}.
+Each application conforms to a distinctive communication pattern that is commonly seen from applications in HPC workload.
+We believe these three applications can represent a big group of applications that conform to similar communication patterns.
+The applications studied in this work are Algebraic MultiGrid Solver(AMG), Geometric MultiGrid(MultiGrid) and CrystalRouter.
+The communication pattern figures we present in Section \ref{sec:amg}, \ref{sec:crystalrouter}, \ref{sec:multigrid} are generated with the IPM \cite{ipm} data from \cite{design forward webpage}.
+
+% MiniApps are reduced proxy applications that encapsulate the salient performance of larger full size applications \cite{miniapp}.
 Since we only focus on the dominant communication pattern of parallel applications, these MiniApps are perfect candidates for analysis the application's performance on different allocations or on new network architectures.
@@
 
-AMG is the MiniApp of Algebraic Multigrid Solver, which is a parallel algebraic multigrid solver for linear systems arising from problems on unstructured mesh physics packages.
+AMG is the MiniApp of Algebraic MultiGrid Solver, which is a parallel algebraic multi-grid solver for linear systems arising from problems on unstructured mesh physics packages.
 It has been derived directly from the BoomerAMG solver that is being developed in the Center for Applied Scientific Computing (CASC) at LLNL \cite{amg}.
-The dominant communication pattern is this regional communication with decreasing message size for different parts of the multigrid v-cycle.
-
-Figure \ref{fig: amg communication topology} shows the communication topology of a small scale AMG with 27 MPI Ranks.
-Here we prefer to show the communication topology with small scale version of the MiniApp because the dominant communication pattern of the application doesn't change with its scale.
-We can make the observation from the upper-left of the figure that each rank in AMG has intense communication between 3 other ranks, such as rank 0 between rank 1, 3 and 9, rank 1 between 2, 4 and 10, etc.
-And this relation is also symmetric along the main diagnal.
+The dominant communication pattern is this regional communication with decreasing message size for different parts of the multi-grid v-cycle.
+
+Figure \ref{fig: amg communication topology} shows the communication topology of a small scale AMG with 216 MPI Ranks.
+Here we show the communication topology with small scale version because the dominant communication pattern of the application doesn't change with scale.
+We can make the observation that AMG's dominant communication pattern is 3D Nearest Neighbor, each rank has intensive communication with with its 6 neighbors unless it is on the boundary.
 The dominant communication pattern is quite obvious when we identify this relation between ranks.
@@
 Figure \ref{fig: amg data trans} shows the amount of data each rank transferred in AMG.
-The blue line shows that total data amount transferred from each rank, while the red shows the amount of received data and green shows the amount of send data.
-The first thing we can spot from the figure is that for each rank, the amount of data sent and received are basically the same, which also well explain the symmetry of the application's dominant communication pattern.
-And the second thing we can find about AMG is that the data transfer amount also has a symmetric pattern.
-Rank 0 has the same amount of data transfer as Rank 26, Rank 1 as of Rank 25, etc.
+The red line shows the amount of data sent by each rank, the blue line shows the amount of received data.
+The data transferring of each rank is between 7MB to 16MB.
 
@@
 
-The second MiniApp we have is Crystal Router, which is the extracted communication kernel of the full application Nek5000.
-Nek5000 \cite{crystalrouter} is a spectral element CFD application developed at Argonne National Laboratory.
-It features spectral element multigrid solvers coupled with a highly scalable, parallel coarse-grid solver that widely used for projects including ocean current modeling, thermal hydraulics of reactor cores, and spatiotemporal chaos.
-The communication kernel of Nek5000, CrystalRouter demonstrates the "many-to-many" communication pattern through scalable multi-stage communication process.
-\textcolor{yellow}{The way CrystalRouter works is nodes compute for a while, then synchronize and communicate, continually alternating between these two types of activities.}
-
-As we can see from Figure \ref{fig: cr communication topology}, every 5-rank is a compact community, where data transfer is very intensive.
-Beside the "5-rank community", there is "pair-wise" global data transfer in CrystalRouter.
-Obviously, there is data transfer alone the diagnal.
-\textcolor{yellow}{This is because messages would be buffered on their node of origin until a synchronization point was reached when a single system call sent every message to its destination in one pass.} When the scale of Crystal Router increase to 100, the "5-rank community" is still the dominant pattern in a hierarchical manner.
-The same pattern can be observed from the global data transfer.
-Our observation about the dominant communication pattern of CrystalRouter is the multi-stage "5-rank community" and "pair-wise" global data transfer happen hierarchically as the scale of the application increase.
-
-Figure \ref{fig: cr data trans} shows the amount of data each rank transferred in Crystal Router.
-The locality pattern of "5-rank community" can still be spotted in this figure.
-The spiky shape occurs in the frequency of every 5 rank.
-Unlike AMG, the there is variance between the amount of data send and received for each rank.
-This is due to the specific feature of Crystal Router's communication pattern.
+The second MiniApp we have is CrystalRouter, the extracted communication kernel of the full application Nek5000 \cite{crystalrouter}, which is a spectral element CFD application developed at Argonne National Laboratory.
+It features spectral element multi-grid solvers coupled with a highly scalable, parallel coarse-grid solver that widely used for projects including ocean current modeling, thermal hydraulics of reactor cores, and spatiotemporal chaos.
+CrystalRouter demonstrates the "many-to-many" communication pattern through scalable multi-stage communication process.
+The way CrystalRouter works is nodes compute for a while, then synchronize and communicate, continually alternating between these two types of activities.
+
+The collective communication in CrystalRouter utilizing a recursive doubling approach.
+Ranks in CrystalRouter conform to a n-dimensional hypercube and recursively splitting into (n-1)-dimensional hypercubes, with communication happening with respect to the links that are severed.
+The pattern of this recursive communication can be found in Figure \ref{fig: cr communication topology}.
+By nature of the logarithmic distribution, more data transferring will be conducted within the local neighborhood, which consists of 5 ranks.
+
+Figure \ref{fig: cr data trans} shows the amount of data each rank transferred in CrystalRouter.
+The spiky shape indicates there is intensive data transferring in the local neighborhood.
+There is great variance of data transferring among the ranks in CrystalRouter, which is between 20MB to 36MB.
+This is due to the specific feature of CrystalRouter's communication pattern.
 
@@
 \label{sec:multigrid}
-Different from AMG, MultiGrid is geometric multigrid v-cycle from production elliptic solver BoxLib, a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes.
-MultiGrid conforms to Many-to-Many communication pattern with decreasing message size and collectives for different parts of the multigrid v-cycle.
-It is widely used for structured grid physics packages.
-Figure \ref{fig: mg communication topology} shows the communication topology of MultiGrid and Figure\ref{fig: mg data trans} shows the the amount of data transferred between ranks in MultiGrid.
 
@@
 
+MultiGrid is geometric multi-grid v-cycle from production elliptic solver BoxLib, a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes.
+MultiGrid conforms to Many-to-Many communication pattern with decreasing message size and collectives for different parts of the multi-grid v-cycle.
+It is widely used for structured grid physics packages.
+
+Figure \ref{fig: mg communication topology} shows the communication topology of MultiGrid with 125 ranks.
+We can see intensive communication along the diagnal that conforms to 3D Nearest Neighbor communication pattern.
+However, there are also considerable amount of data transferring on both sides of the diagnal, which means the dominant communication pattern of MultiGrid is ``Many-to-Many".
+
+Figure\ref{fig: mg data trans} shows the the amount of data transferred between ranks in MultiGrid. The data transferring of each rank is relatively small and stable, which is between 2.8MB to 2.9MB.
+
 MultiGrid represents a group of applications whose dominant communication pattern is Many-to-Many (some literatures also refer it as All-to-All \cite{hpdc2015-ornl}).
+Applications with such similar dominant communication pattern are like FillBoundary and MiniDFT \cite{design forward webpage}.
 
@@
 
-It is impossible to accurately study the interference between concurrently running jobs with different allocation strategies on production systems.
+It is difficult to accurately and flexibly experiment with concurrently running jobs in an HPC context.
 There are a couple of reasons for that.
@@
 
-There are couple of simulation platform to study the behavior of large scale systems.
+%There are couple of simulation platform to study the behavior of large scale systems.
 SimGrid from Inria is a versatile simulation tool sets that provide simulation for Grids, Clouds and HPC systems \cite{simgrid}.
@@
 CODES support both torus and dragonfly network with high fidelity flit-level simulation.
-In this work, we only use torus network to show the impact of node allocation to job with different dominant communication patterns.
 CODES has this network workload component that capable of conducting trace-driven simulation.
 It can take real MPI application trace generated by SST DUMPI \cite{sst} to drive CODES network models.
-We implemented a Job Mapping API for CODES, which can provide flexible job allocation strategies.
+In this work, we focus on an in-depth analysis of intra- and inter-job communication interferences with  different job allocations on torus-connected HPC systems.
+Since torus networks have been extensively used in the current generation of supercomputers because of their linear scaling on per-node cost and competitive communication performance \cite{mubarak-sc2012}\cite{zhou-ipdps}.
+%We implemented a Job Mapping API for CODES, which can provide flexible job allocation strategies.
 The new API can help us to explore the impact of different allocation strategies to those three applications in Section \ref{sec:application study}
 
-\section{Experiment}
-\label{sec:experiment results}
-
-CODES can provide Torus network model with good scalability and high fidelity \cite{codes-torus}.
-In this section, we will first study the communication behavior of three applications on torus network with different dimensionality and bandwidth configurations.
-Then, we study the intra- and inter-job interference with these three applications by running them exclusively and concurrently with different allocation strategies.
-
-\subsection{Configuration Study }
+\section{Configuration Study }
 \label{sec:config study}
+
+In this section, we study the communication behavior of three applications on torus network with different dimensionality and bandwidth configurations.
 
@@
 Figure \ref{fig:dimensionality-study} shows the performance of the AMG, CrystalRouter and MultiGrid on a 2K node torus network model with a 3D torus (16x16x8), a 5D torus (8x4x4x4x4), and a 7D torus (4x4x4x4x2x2x2).
-The bandwidth of direct link between nodes is 2GiB/s in each direction, thus, the aggregated bandwidth is 12GiB/s per node in 3D torus, 20GiB/s per node in 5D torus, and 28GiB/s per node in 7D torus.
+The bandwidth of direct link between nodes is 2GiB/s in each direction \cite{bgq}, thus, the aggregated bandwidth is 12GiB/s per node in 3D torus, 20GiB/s per node in 5D torus, and 28GiB/s per node in 7D torus.
 
 We can see from Figure \ref{fig:dimensionality-study} that the data transfer time of those three applications are greatly reduced as the dimensionality increases.
-The increased aggregated bandwidth of each node can accelerate the transfer of data.
-This can justify the fidelity and consistency of the torus network model.
+The increased aggregated bandwidth of each node can accelerate the data transfer.
 We allocate each rank within each application linearly on the 3D torus.
 However, due to the variance of data transfer volume of each rank, AMG's data transfer time has great dispersion, shown as the outliers above the boxes in Figure \ref{fig:dimen-amg}.
-There is no such great dispersion for CrystalRouter and MultiGrid, shown in Figure \ref{fig:dimen-cr} and \ref{fig:dimen-mg}, since no great variance of data transfer amount of each rank in these applications.
+There is no such great dispersion of MultiGrid, shown in Figure \ref{fig:dimen-mg}, since there is no great variance of data transfer of each rank.
+The data transfer of each rank in MultiGrid varies slightly  between 2.8MB to 2.9MB.
+\textcolor{red}{explain crystalrouter}.
 
@@
 
-\subsection{Interference Analysis}
+\section{Interference Analysis}
 \label{sec:interference}
@@
 Inter-job interference is a more prominent issue for system adopted non-contiguous allocation policy than partition-based system.
-Application's communication time can vary from 36\% faster to 69\% slower due to inter-job interference when running on non-contiguous allocation \cite{abhinav-sc13}.
+Application communication times have been demonstrated to vary from 36\% faster to 69\% slower due to inter-job interference when running on non-contiguous allocation \cite{abhinav-sc13}.
 
@@
 \begin{enumerate}
-    \item The compact allocation can not guarantee the best performance for every application. 
+    \item The compact allocation may not guarantee the best performance for every application. 
     \item The performance of application conforms to Nearest Neighbor communication pattern keep relatively stable with different shape allocations, while application conforms to ``Many-to-Many" communication pattern prefer compact allocation.
@@
 
-\subsubsection{Intra-Job Interference Analysis}
+\subsection{Intra-Job Interference Analysis}
 \label{sec: introjob}
@@
 
-\subsubsection{Inter-Job Interference Analysis}
+\subsection{Inter-Job Interference Analysis}
 \label{sec: interjob interference study}
@@
 
-The data transfer time of AMG in Figure \ref{fig:interjob-amg-box} keeps stable between allocation unit size of 16(bit) and 8(medium), since the locality of AMG is 6-rank based.
-When the unit size reduce to 2(small), AMG suffers prolong data transfer time by about 10\%.
+The data transfer time of AMG in Figure \ref{fig:interjob-amg-box} keeps stable between allocation unit size of 16 (big) and 8 (medium), since the locality of AMG is 6-rank based.
+When the unit size reduce to 2 (small), AMG suffers prolong data transfer time by about 10\%.
 CrystalRouter is more sensitive to allocation unit size.
@@
 Most existing batch schedulers take so few initiative because it been hold under the obscurantism policy.
-Whether it use partition-based or non-contiguous policy, means it chooses the benefit of either system or job over the other.
-Some propose compromised policies \cite{lo}\cite{pascual}\cite{yang-cluster14} to use fix allocation unit size and  make the allocation for job blindly without knowing job's communication information.
-
+Whether it uses partition-based or non-contiguous policy, it chooses the benefit of either system or job over the other.
+Some proposed compromised policies \cite{lo}\cite{pascual}\cite{yang-cluster14} to use fix allocation unit size and  make the allocation for job blindly without knowing job's communication information.
+
+providing insights for ....
 Based on the findings in this work, we propose the idea of smart job allocation with consideration about job's communication pattern.

### 2015-10-14 submission.tex
minor change

@@
 The novelty of our new strategy is that it makes allocation decisions based on job’s communication pattern in addition to system topology and resource availability information.
-The scheduler uses job communication patterns information to decide whether its performance relies heavily on the network or not, and what the potential inter/intra-job interference would be for a given allocation.
+The scheduler uses job communication patterns information to decide whether its performance relies heavily on the network or not, and what the potential intra- and inter-job interference would be for a given allocation.
 Thus, the scheduler can pick a preferable resource allocation based on a specific job's communication pattern.
@@
 %The off-line study of job's communication traces could be helpful to develop better allocation policies for batch scheduler to make on-line scheduling/allocating decisions.
-This is because the communication pattern of parallel applications on HPC system can be summarized into a couple of categories\cite{hpdc2015-ornl}.
+This is because the communication pattern of parallel applications on HPC system can be summarized into a couple of categories\cite{roth}.
 The deep understanding of one application can be helpful to all other applications that conforms to the same communication pattern.
@@
 
-%A parallel application usually conforms to a combination of several basic communication patterns\cite{hpdc2015-ornl}.
+%A parallel application usually conforms to a combination of several basic communication patterns\cite{roth}.
 At its different execution phases, the application's communication behavior may follow different basic patterns respectively.
@@
 
-A parallel application usually conforms to a combination of several basic communication patterns \cite{hpdc2015-ornl}.
+A parallel application usually conforms to a combination of several basic communication patterns \cite{roth}.
 At its different execution phases, the application's communication behavior may follow different basic patterns respectively.
@@
 It features spectral element multi-grid solvers coupled with a highly scalable, parallel coarse-grid solver that widely used for projects including ocean current modeling, thermal hydraulics of reactor cores, and spatiotemporal chaos.
-CrystalRouter demonstrates the "many-to-many" communication pattern through scalable multi-stage communication process.
+CrystalRouter demonstrates the ``Many-to-Many" communication pattern through scalable multi-stage communication process.
 The way CrystalRouter works is nodes compute for a while, then synchronize and communicate, continually alternating between these two types of activities.
@@
 The pattern of this recursive communication can be found in Figure \ref{fig: cr communication topology}.
-By nature of the logarithmic distribution, more data transferring will be conducted within the local neighborhood, which consists of 5 ranks.
+By nature of the logarithmic distribution, more data transferring are in the local neighborhood, which consists of 5 ranks.
 
@@
 
-MultiGrid represents a group of applications whose dominant communication pattern is Many-to-Many (some literatures also refer it as All-to-All \cite{hpdc2015-ornl}).
+MultiGrid represents a group of applications whose dominant communication pattern is Many-to-Many (some literatures also refer it as All-to-All \cite{roth}).
 Applications with such similar dominant communication pattern are like FillBoundary and MiniDFT \cite{design forward webpage}.
@@
 AMG represents a group of applications whose dominant communication pattern is Nearest Neighbor.
-MultiGrid represents applications whose dominant communication pattern is Many-to-Many (some literatures also refer it to All-to-All\cite{hpdc2015-ornl}).
+MultiGrid represents applications whose dominant communication pattern is Many-to-Many (some literatures also refer it to All-to-All\cite{roth}).
 CrystalRouter represents those applications whose dominant communication is a hybrid of multi-stage local communication and hierarchical global communication.
@@
 It is difficult to accurately and flexibly experiment with concurrently running jobs in an HPC context.
-There are a couple of reasons for that.
-First, the allocation strategy used on production machine is part of the system software, which can not be changed by user.
+One reason is that the allocation strategy used on production machine is part of the system software, which can not be changed by user.
 Even the system administrator is not authorized to make that change.
-Second, it is unrealistic to reserve the system exclusively to run the same job with desired allocation without interference then compare the results to that in the presence of interference.
-Due to those reasons, we resort to simulation for this work.
+Another reason is that it is unrealistic to reserve the system exclusively to run the same job with desired allocation without interference then compare the results with those in the presence of interference.
+Therefore, we resort to simulation for this work.
 
@@
 
-A simulation toolkit named CODES (Co-Design of Multi-layer Exascale Storage Architecture) from Argonne National Lab enables the exploration of simulating different HPC networks with high fidelity \cite{Jason-2011}\cite{mubarak-sc2012}.
+A simulation toolkit named CODES enables the exploration of simulating different HPC networks with high fidelity \cite{Jason-2011}\cite{mubarak-sc2012}.
 CODES is built on top of Rensselaer Optimistic Simulation System (ROSS) parallel discrete-event simulator, which is capable of processing billions of events per second on leadership-class supercomputers \cite{ross}.
@@
 The topology of torus network is k-ary n-cube, with $k^n$ nodes in total arranged in an n-dimensional grid having k nodes in each dimension.
-Each node has 2$\times$ n direct linked neighbor nodes.
+Each node has 2$\times$n direct linked neighbor nodes.
 The torus network performance is determined by its dimensionality and link bandwidth.
@@
 The increased aggregated bandwidth of each node will definitely reduce the data transfer time of each rank in the application.
-Figure \ref{fig:dimensionality-study} shows the performance of the AMG, CrystalRouter and MultiGrid on a 2K node torus network model with a 3D torus (16x16x8), a 5D torus (8x4x4x4x4), and a 7D torus (4x4x4x4x2x2x2).
+Figure \ref{fig:dimensionality-study} shows the performance of the AMG, CrystalRouter and MultiGrid on a 2K node torus network model with a 3D torus (16$\times$16$\times$8), a 5D torus (8$\times$4$\times$4$\times$4$\times$4), and a 7D torus (4$\times$4$\times$4$\times$4$\times$2$\times$2$\times$2).
 The bandwidth of direct link between nodes is 2GiB/s in each direction \cite{bgq}, thus, the aggregated bandwidth is 12GiB/s per node in 3D torus, 20GiB/s per node in 5D torus, and 28GiB/s per node in 7D torus.
@@
 However, due to the variance of data transfer volume of each rank, AMG's data transfer time has great dispersion, shown as the outliers above the boxes in Figure \ref{fig:dimen-amg}.
-There is no such great dispersion of MultiGrid, shown in Figure \ref{fig:dimen-mg}, since there is no great variance of data transfer of each rank.
-The data transfer of each rank in MultiGrid varies slightly  between 2.8MB to 2.9MB.
-\textcolor{red}{explain crystalrouter}.
+There is few dispersion of MultiGrid, shown in Figure \ref{fig:dimen-mg}, since data transfer of each rank in MultiGrid varies slightly  between 2.8MB to 2.9MB.
+CrystalRouter also has few dispersion as shown in Figure \ref{fig:dimen-cr}, this is due to the majority of data transferring is in the local neighborhood, which only take few hops.
 
@@
 
-Communication variability due to network sharing can cause application performance degradation in two ways, namely intra- and inter-job interference.
-The \textbf{\emph{intra-job interference}} refers to the network contention between ranks within each application.
-There are lots of research works \cite{jingjin}\cite{ozan} try to come up with better task mapping algorithms to alleviate the intra-job interference, which is out the scope of our work.
-In this work, we focus on the analysis of such interference with different allocations.
-
+Communication variability due to network sharing can cause application performance degradation.
+\textbf{\emph{Intra-job interference}} refers to the network contention between ranks within each application.
 \textbf{\emph{Inter-job interference}} is introduced by concurrently running jobs that adjacent to each other sharing network resources.
-Such interference has been identified as one of the major culprits for application's performance variability \cite{abhinav-sc13}\cite{skinner}\cite{rosenthal}.
-Inter-job interference is a more prominent issue for system adopted non-contiguous allocation policy than partition-based system.
-Application communication times have been demonstrated to vary from 36\% faster to 69\% slower due to inter-job interference when running on non-contiguous allocation \cite{abhinav-sc13}.
-
-In this section we will study both kinds of interference with with AMG, CrystalRouter and MultiGrid in their medium scale.
-To be specific, we choose AMG with 216 ranks, CrystalRouter with 100 ranks, and MultiGrid 125 ranks and simulate their running on 2K-node (16 $\times$ 16 $\times$ 8) 3D torus network.
-
-We can get observations that 
-\begin{enumerate}
-    \item The compact allocation may not guarantee the best performance for every application. 
-    \item The performance of application conforms to Nearest Neighbor communication pattern keep relatively stable with different shape allocations, while application conforms to ``Many-to-Many" communication pattern prefer compact allocation.
-    \item The allocation unit size should be decided according to application's communication pattern.
-Unit size that big enough to preserve the locality of the application will result better performance.
-\end{enumerate}
+
+In this section we will study both kinds of interference in the context of torus network. 
+Since application's communication pattern won't change with its scale, we choose AMG with 216 ranks, CrystalRouter with 100 ranks, and MultiGrid 125 ranks.
+And in order to accommodate three applications running concurrently with different allocations, we simulate a 2K-node (16 $\times$ 16 $\times$ 8) 3D torus network for the experiments.
 
@@
 (b) Illustrates the non-contiguous allocation.
-Different colors represent different jobs.
+Each job is represented by a specific color.
 The nodes assigned to different jobs are interleaved, the allocation unit is 4-node.}
@@
 
+There are lots of research works \cite{jingjin}\cite{ozan} try to come up with better task mapping algorithms to alleviate the intra-job interference, which is out the scope of our work.
+In this work, we focus on the analysis of such interference with different allocations.
 To study the intra-job interference of each job, we simulate each job running on 3D torus network with different allocation shapes.
@@
 And MultiGrid gets the best performance (shortest data transfer time) when running on 3D-balanced allocation, as shown in Figure \ref{fig:shapstudy-mg}.
-Since MultiGrid's communication pattern is many-to-many dominant, 3D-balanced is the most compact allocation with shortest pair-wise distance between nodes, which can reduce the aggregated hops for transferring message between ranks in MultiGrid.
+Since MultiGrid's communication pattern is ``Many-to-Many" dominant, 3D-balanced is the most compact allocation with shortest pair-wise distance between nodes, which can reduce the aggregated hops for transferring message between ranks in MultiGrid.
 CrystalRouter has with both dominant local communication and multi-stage global data transfer, the 3D-balance is also the best allocation, but its advantage over 3D-unbalanced is not that obvious as to MultiGrid, as shown in Figure \ref{fig:shapstudy-cr}.
@@
 
-We simulate three application running concurrently with different allocation to analysis the interference between them.
+%We simulate three application running concurrently with different allocation to analysis the interference between them.
 First, we assign each application with contiguous allocation side by side, and compare their performance with each application running exclusively.
@@
 The results are shown in Figure \ref{fig: concurrent-cont-time-box}.
-
-Then, we assign each application with non-contiguous allocation and make them running concurrently.
+Inter-job interference has been identified as one of the major culprits for application's performance variability \cite{abhinav-sc13}\cite{skinner}\cite{rosenthal}.
+Inter-job interference is a more prominent issue for system adopted non-contiguous allocation policy than partition-based system.
+Application communication times have been demonstrated to vary from 36\% faster to 69\% slower due to job interference when running on non-contiguous allocation \cite{abhinav-sc13}.
+
+We assign each application with non-contiguous allocation and run them concurrently on the same network.
 The allocation unit belongs to different jobs are interleaved as shown in Figure \ref{fig:noncont_sub1}.
@@
 
+\subsection{Summary}
+\label{sec:summary}
+
+Based on our comprehensive study, we can get observations that 
+\begin{enumerate}
+    \item The compact allocation may not guarantee the best performance for every application. 
+    \item The performance of application conforms to Nearest Neighbor communication pattern keep relatively stable with different shape allocations, while application conforms to ``Many-to-Many" communication pattern prefer compact allocation.
+    \item The allocation unit size should be decided according to application's communication pattern.
+Unit size that big enough to preserve the locality of the application will result better performance.
+\end{enumerate}
+
 \section{Discussion}
 \label{sec:discussion}
+
 HPC system usually accommodate many parallel jobs running concurrently.
@@
 The most promising way to utilize such knowledge is to make smart allocation based on job's dominant communication pattern.
-Most existing batch schedulers take so few initiative because it been hold under the obscurantism policy.
-Whether it uses partition-based or non-contiguous policy, it chooses the benefit of either system or job over the other.
-Some proposed compromised policies \cite{lo}\cite{pascual}\cite{yang-cluster14} to use fix allocation unit size and  make the allocation for job blindly without knowing job's communication information.
-
-providing insights for ....
+Most existing batch schedulers take so few initiative because it fixed-size partition-based or non-contiguous policy, which means it chooses the benefit of either system or job over the other.
+Some proposed compromised policies \cite{lo}\cite{pascual}\cite{yang-cluster14} to use various allocation unit sizes, but they still  make the allocation for job blindly without knowing job's communication information.
+
 Based on the findings in this work, we propose the idea of smart job allocation with consideration about job's communication pattern.
 As the dominant communication pattern of parallel applications on HPC system can be summarized into a couple of categories, the deep understanding of one application can be helpful to all other applications that conforms to the same communication pattern.
-When an application with dominant communication that is intensive all-to-all submitted to the system, it should be granted with compact allocation with exclusive network provision, so that it won't suffer performance loss due to interference.
+When an application with dominant communication that is intensive all-to-all being submitted to the system, it should be granted with compact allocation with exclusive network provision, so that it won't suffer performance loss due to interference.
 If job with dominant communication pattern like Nearest Neighbor or others with intensive small scale communication, rather than assign it with unnecessary compact allocation and exclusive network provision, non-contiguous allocation with proper unit size would serve it equally well.
@@
 Oak Ridge National Laboratory has this ongoing project about developing a tool set named Oxbow, which can characterize the computation and communication behavior of 12 scientific applications and benchmarks \cite{oxbow}.
-In their recent work \cite{hpdc2015-ornl}, they demonstrate a new approach to automatically characterizing a parallel application's communication behavior.
+In their recent work \cite{roth}, they demonstrate a new approach to automatically characterizing a parallel application's communication behavior.
 
@@
 Dong et al describe IBM Blue Gene/Q's 5D torus interconnect network \cite{Dong-SC11}.
-They developed simple benchmarks that conforms to four different communication patterns, namely ping-pong, nearest neighbor, broadcast and allreduce, to demonstrate the effectiveness of this highly parallel 5D torus network.
+They developed simple benchmarks that conforms to four different communication patterns, namely ping-pong, nearest neighbor, broadcast and all$\textunderscore$reduce, to demonstrate the effectiveness of this highly parallel 5D torus network.
 
@@
 
-\bibitem{hpdc14}
-Prisacari, Bogdan and Rodriguez, German and Heidelberger, Philip and Chen, Dong and Minkenberg, Cyriel and Hoefler, Torsten
-\newblock Efficient Task Placement and Routing of Nearest Neighbor Exchanges in Dragonfly Networks
-\newblock In Proceedings of the 23rd International Symposium on High-performance Parallel and Distributed Computing
-\newblock HPDC '14,
-\newblock 2014,
-\newblock Vancouver, BC, Canada
+%\bibitem{hpdc14}
+%Prisacari, Bogdan and Rodriguez, German and Heidelberger, Philip and Chen, Dong and Minkenberg, Cyriel and Hoefler, Torsten
+%\newblock Efficient Task Placement and Routing of Nearest Neighbor Exchanges in Dragonfly Networks
+%\newblock In Proceedings of the 23rd International Symposium on High-performance Parallel and Distributed Computing
+%\newblock HPDC '14,
+%\newblock 2014,
+%\newblock Vancouver, BC, Canada
 
@@
 \newblock  Vancouver, BC, Canada,
-
-\bibitem{hpdc2015-ornl}
-Philip. C.R, Jeremy S. M, etc
-\newblock "Automated Characterization of Parallel Application Communication Patterns"
-\newblock HPDC'15
 

### 2015-10-15 submission.tex
move reference in .bib file in bibtex form

@@
 \end{figure}
-
-%\begin{figure}[h!] 
-%  \centering
-%  \includegraphics[width=0.38\textwidth]{figs/mira}
-%   \caption{The activity of Mira, IBM BlueGene/Q system at Argonne National Lab. Many jobs runninng concurrently on the system. Different jobs represented by specific colors. }
-%   \label{fig: mira activity}
-%\end{figure}
 
@@
 
+\subsection{Intra-Job Interference Analysis}
+\label{sec: introjob}
+
+There are lots of research works \cite{jingjin}\cite{ozan} try to come up with better task mapping algorithms to alleviate the intra-job interference, which is out the scope of our work.
+In this work, we focus on the analysis of such interference with different allocations.
+To study the intra-job interference of each job, we simulate each job running on 3D torus network with different allocation shapes.
+We assign each application with three different shapes allocation, i.e.
+3D balanced-cube, 3D unbalanced-cube, 2D mesh.
+
+3D balanced-cube, shown as red in Figure \ref{fig:cont_sub1}, can guarantee the minimum \emph{average pair-wise distance} within the allocation.
+There are some research work like \cite{leung} \cite{abhinav-sc13} try to prove that compact allocation can guarantee job with better performance.
+They use many metrics to evaluate the compactness of the allocation, such as \emph{average pair-wise distance}, \emph{diameter} and \emph{contiguity}.
+In this work, we select 3D balanced-cube as the most compact allocation on a 3D torus network.
+
+3D unbalanced-cube, shown as green in Figure \ref{fig:cont_sub1}, is a rectangular prism, which is the possible allocation shape on system with asymmetric networks.
+For example, Blue Waters Cray XE6/XK7 system network is 3D torus with Gemini routers.
+The network connections in the \emph{y}direction have only half the bandwidth of the cables used in the \emph{x} and \emph{z} directions.
+In order to take advantage of the faster links in the \emph{x} and \emph{z} directions, job allocation is always start from X-Z plane, which leads to a rectangular prism shaped allocation \cite{RF}.
+
+2D mesh, shown as blue in Figure \ref{fig:cont_sub1}, can be cut out from a single layer of the 3D torus.
+2D mesh is a very common allocation shape in torus network for both partition-based and non-contiguous allocation strategies.
+For example, Cray Application Level Placement Scheduler indexes the nodes in torus network into a list and make allocation by simply works off the this list \cite{carl-cug}.
+When the list is obtained by sorting the nodes based on their spacial coordinates in the torus, allocation made off from this list will be 2D mesh.
+The IBM Blue/Gene Q supercomputer Mira at Argonne Leadership Computing Facility also allow its allocation partition configured into mesh \cite{zhou-ipdps}.
+
 \begin{figure}[t!]
@@
 
-\subsection{Intra-Job Interference Analysis}
-\label{sec: introjob}
-
-There are lots of research works \cite{jingjin}\cite{ozan} try to come up with better task mapping algorithms to alleviate the intra-job interference, which is out the scope of our work.
-In this work, we focus on the analysis of such interference with different allocations.
-To study the intra-job interference of each job, we simulate each job running on 3D torus network with different allocation shapes.
-We assign each application with three different shapes allocation, i.e.
-3D balanced-cube, 3D unbalanced-cube, 2D mesh.
-
-3D balanced-cube, shown as red in Figure \ref{fig:cont_sub1}, can guarantee the minimum \emph{average pair-wise distance} within the allocation.
-There are some research work like \cite{leung} \cite{abhinav-sc13} try to prove that compact allocation can guarantee job with better performance.
-They use many metrics to evaluate the compactness of the allocation, such as \emph{average pair-wise distance}, \emph{diameter} and \emph{contiguity}.
-In this work, we select 3D balanced-cube as the most compact allocation on a 3D torus network.
-
-3D unbalanced-cube, shown as green in Figure \ref{fig:cont_sub1}, is a rectangular prism, which is the possible allocation shape on system with asymmetric networks.
-For example, Blue Waters Cray XE6/XK7 system network is 3D torus with Gemini routers.
-The network connections in the \emph{y}direction have only half the bandwidth of the cables used in the \emph{x} and \emph{z} directions.
-In order to take advantage of the faster links in the \emph{x} and \emph{z} directions, job allocation is always start from X-Z plane, which leads to a rectangular prism shaped allocation \cite{RF}.
-
-2D mesh, shown as blue in Figure \ref{fig:cont_sub1}, can be cut out from a single layer of the 3D torus.
-2D mesh is a very common allocation shape in torus network for both partition-based and non-contiguous allocation strategies.
-For example, Cray Application Level Placement Scheduler indexes the nodes in torus network into a list and make allocation by simply works off the this list \cite{carl-cug}.
-When the list is obtained by sorting the nodes based on their spacial coordinates in the torus, allocation made off from this list will be 2D mesh.
-The IBM Blue/Gene Q supercomputer Mira at Argonne Leadership Computing Facility also allow its allocation partition configured into mesh \cite{zhou-ipdps}.
-
-There are lots of research work try to design fancy allocation algorithms to provide application with most compact allocation.
-However, providing cubical allocation without considering application's communication pattern won't guarantee the best performance for every application.
-Figure \ref{fig:shapstudy-amg} shows that for AMG, who has this symmetric local communication pattern through all ranks, spent less time when assigned with 3D-unbalanced allocation than 3D-balanced.
+Figure \ref{fig:shapstudy-amg} shows that for AMG, who conforms to 3D ``Nearest Neighbor" communication pattern, spent less time when being assigned with 3D-unbalanced allocation than 3D-balanced.
 And MultiGrid gets the best performance (shortest data transfer time) when running on 3D-balanced allocation, as shown in Figure \ref{fig:shapstudy-mg}.
@@
 
-%\emph{OBSERVATION}
+There are lots of research work try to design fancy allocation algorithms to provide application with most compact allocation.
+However, providing cubical allocation without considering application's communication pattern won't guarantee the best performance for every application.
 Compact allocation should be provided to application with global data transfer, such as those conforms to ``Many-to-Many" communication pattern.
-Application with dominant communication pattern like Nearest Neighbor won't fully utilize the compact allocation and could get even better performance on a relaxed allocation shape.
+Application with dominant communication pattern like ``Nearest Neighbor" won't fully utilize the compact allocation and could get even better performance on a relaxed allocation shape.
 
@@
 CrystalRouter is more sensitive to allocation unit size.
-The best unit size would be the one big enough to accommodate its 5-rank locality.
+The best unit size would be the one big enough to accommodate its local neighborhood.
 Figure \ref{fig:interjob-cr-box} shows unit size of 16 and 8 can guarantee the same average data transfer time, while some rank spend more time with allocation unit size 8 than 16.
-When the unit size reduce to 2-node, the communication become less efficiency and take 15\% more times for transferring data.
+When the unit size reduce to 2-node, the communication become less efficiency and takes about 15\% more times for transferring data.
 
 The data transfer time of MultiGrid with different allocation unit sizes doesn't show obvious variability in Figure \ref{fig:interjob-mg-box}.
-This is because even big allocation unit size like 16 will still fail to preserve MultiGrid's ``Many-to-Many" pattern.
+This is because even big allocation unit size like 16-node will still fail to preserve MultiGrid's ``Many-to-Many" pattern.
 The data transfer time is almost doubled when MultiGrid running concurrently with allocation unit size of 16, as shown in Figure \ref{fig:interjob-mg-box} .
@@
 
-%\emph{OBSERVATION}
 The unit size in non-contiguous allocation policy should be chosen according to application's communication pattern.
-Inter-job interference is inevitable in non-contiguous based systems, but unit size that big enough to preserve the locality of the application will alleviate such interference and improve job performance.
+Inter-job interference is inevitable in non-contiguous based systems, but unit size that big enough to preserve the neighborhood communication of the application will alleviate such interference and improve job performance.
 
@@
 ``cont" indicates three applications running side by side concurrently on the same network with contiguous allocation.
-To study the impact of non-contiguous allocation, applications running concurrently with interleaved allocation of different unit sizes, which are 16-node, 8-node, 2-node.
+To study the impact of non-contiguous allocation on inter-job interference, applications running concurrently with interleaved allocation of different unit sizes, which are 16-node, 8-node, 2-node.
 }
@@
 
-Based on our comprehensive study, we can get observations that 
+Based on our comprehensive study, we can get following observations 
 \begin{enumerate}
     \item The compact allocation may not guarantee the best performance for every application. 
-    \item The performance of application conforms to Nearest Neighbor communication pattern keep relatively stable with different shape allocations, while application conforms to ``Many-to-Many" communication pattern prefer compact allocation.
+    \item The performance of application conforms to Nearest Neighbor communication pattern keeps relatively stable with different shape allocations, while application conforms to ``Many-to-Many" communication pattern prefers compact allocation.
     \item The allocation unit size should be decided according to application's communication pattern.
-Unit size that big enough to preserve the locality of the application will result better performance.
+Unit size that big enough to preserve the neighborhood communication of the application will result in better performance.
 \end{enumerate}
@@
 
-HPC system usually accommodate many parallel jobs running concurrently.
+%HPC system usually accommodate many parallel jobs running concurrently.
 Each job gets the required number of computing nodes by system resource manager.
@@
 
-The knowledge about HPC job's interference is extremely useful for improving system and job's performance.
-By scrutinizing job's communication behavior, we can locate the critical path with job's communication graph, identify  job's dominant communication pattern.
-With such knowledge about job's communication characteristics, we can analysis the possible interference between jobs through simulation or even predict such interference when job running on different allocations.
-
-The most promising way to utilize such knowledge is to make smart allocation based on job's dominant communication pattern.
-Most existing batch schedulers take so few initiative because it fixed-size partition-based or non-contiguous policy, which means it chooses the benefit of either system or job over the other.
-Some proposed compromised policies \cite{lo}\cite{pascual}\cite{yang-cluster14} to use various allocation unit sizes, but they still  make the allocation for job blindly without knowing job's communication information.
-
-Based on the findings in this work, we propose the idea of smart job allocation with consideration about job's communication pattern.
-As the dominant communication pattern of parallel applications on HPC system can be summarized into a couple of categories, the deep understanding of one application can be helpful to all other applications that conforms to the same communication pattern.
-When an application with dominant communication that is intensive all-to-all being submitted to the system, it should be granted with compact allocation with exclusive network provision, so that it won't suffer performance loss due to interference.
-If job with dominant communication pattern like Nearest Neighbor or others with intensive small scale communication, rather than assign it with unnecessary compact allocation and exclusive network provision, non-contiguous allocation with proper unit size would serve it equally well.
-No degradation will be introduced to job's performance and system utilization will be improved.
+The observations we obtained by analyzing the intra- and inter-job interference can provide insights for the design of a smart and flexible job allocation strategy.
+By scrutinizing job's communication behavior, we can identify job's dominant communication pattern and pinpoint the ``neighborhood communication".
+With such knowledge about jobs' communication patterns, we can analyze the possible interference between jobs and take precautions to alleviate the negative effect when making allocation decisions.
+
+When an application with dominant communication that is intensive ``Many-to-Many" being submitted to the system, it should be granted with compact node allocation and exclusive network provision.
+The compact allocation can guarantee the shortest pair-wise distance between all the ranks, and make the data transfer between ranks take less hops.
+On the other hand, the exclusive network provision will prevent other adjacent applications from sharing network resources, thus eliminate the performance degradation due to interference.
+
+Not every application's preferable resource are compact node allocation and exclusive network provision.
+As we found in our comprehensive analysis, applications whose dominant communication patterns contain intensive ``neighborhood communication" like ``Nearest Neighbor" won't benefit from compact node allocation and exclusive network provision.
+Such applications can run with non-contiguous node allocation without performance degradation, as long as the allocation unit can accommodate the ``neighborhood communication".
+
+The allocation unit size shouldn't be fixed.
+Instead, the new allocation strategy should choose the proper unit size based on the scale of ``neighborhood communication" of each job.
+The best unit size should be neither too big nor too small, jut perfectly fit for the size of that ``neighborhood".
+Big unit size won't be fully utilized and results in fragmentation.
+Small unit size won't be able to accommodate the ``neighborhood communication" and makes the intra-job communication less efficient.
 
@@
 
-\begin{thebibliography}{10}
- %=====start of scheduling papers=====
-
-%\bibitem{yang-sc13}
-%X. ~Yang, Z.~Zhou, S.~Wallace Z.~Lan, W.~Tang, S. ~Coghlan and Mike. ~Papka.
-%\newblock Integrating Dynamic Pricing of Electricity into Energy Aware Scheduling for HPC Systems. 
-%\newblock In {Proceedings of the International Conference on High Performance Computing, Networking, Storage and Analysis(SC), 2013 International Conference for}
-%\newblock SC '13,
-%\newblock 2013,
-%\newblock Salt Lake City, Utah
-
-%=====start of allocation aware papers===== 
-\bibitem{leung}
-V.J. Leung,  E.M. Arkin, M.A. Bender,  D. Bunde, J. Johnston, etc,
-\newblock Processor allocation on Cplant: achieving general processor locality using one-dimensional allocation strategies
-\newblock In {Proceedings of 2002 IEEE International Conference on Cluster Computing, 2002}.  pages 296--304, 2002
-
-\bibitem{carl-cug}
-Carl Albing and Mark Baker
-\newblock ALPS,Topology, and Performance: A Comparison of Linear Orderings for 
-Application Placement in a 3D torus.
-\newblock Presented at the CUG 2010, Edinburgh, Scotland, UK, 2010.
-
-\bibitem{yang-cluster14}
-Xu Yang and Zhou Zhou and Wei Tang and Xingwu Zheng and Jia Wang and Zhiling Lan.
-\newblock Cluster Computing (CLUSTER), 2014 IEEE International Conference on,
-\newblock Balancing job performance with system performance via locality-aware scheduling on torus-connected systems
-\newblock Sept, 2014,
-\newblock pp.140-148
-
-\bibitem{lo}
-V. Lo,  K. Windisch, W.  Liu, and Nitzberg.
-\newblock Noncontiguous processor allocation algorithms for mesh-connected multicomputers.
-\newblock IEEE Transactions on Parallel and Distributed Systems 8 (1997). pages 712--726
-
-\bibitem{pascual} 
-JA. Pascual, J. Navaridas and  J. Miguel-Alonso. 
-\newblock Effects of Topology-Aware Allocation Policies on Scheduling Performance. 
-\newblock 14th Workshop on Job Scheduling Strategies for Parallel Processing. May 29, 2009.Rome, Italy.
-
-%+++++++++++++++++++++++
-\bibitem{mpip}
-\newblock mpiP: Lightweight, Scalable MPI Profiling.
-\newblock http://mpip.sourceforge.net, 
-\newblock 2013.
-
-\bibitem{tau}
-S. Shende and A. Malony. 
-\newblock The TAU parallel performance system. 
-\newblock International Journal of High Performance Computing Applications, 20(2):287–311, 2006.
-
-\bibitem{scala}
-X. Wu, F. Mueller, S. Pakin, 
-\newblock "Automatic Generation of Executable Communication Specifications from Parallel Applications", 
-\newblock In {Proceedings of the International Conference on High Performance Computing, Networking, Storage and Analysis(SC), 2011 International Conference for}
-\newblock SC '11,
-\newblock 2011,pages 12-21.
-
-\bibitem{design forward webpage}
-Department of Energy, “Characterization of the DOE Mini-apps.” [Online].
-\newblock Available: http://portal.nersc.gov/project/CAL/trace.htm
-\newblock (Accessed on: Sept 16th. 2015)
-
-\bibitem{coral}
-CORAL collaboration benchmark codes [Online].
-\newblock Available:https://asc.llnl.gov/CORAL-benchmarks/
-\newblock (Accessed on: Sept 16th. 2015)
-
-\bibitem{ipm}
-David Skinner, Nicholas Wright, Karl Fuerlinger and Kathy Yelick
-\newblock Integrated Performance Montoring(IPM)
-\newblock Available http://ipm-hpc.sourceforge.net/overview.html
-\newblock (Accessed on: Sept 16th. 2015)
-
-\bibitem{miniapp}
-Bronson Messer
-\newblock "Using a Developing MiniApp to Compare Platform Characteristics on Cray Systems"
-\newblock In Proceedings of Cray User Group 2012
-
-\bibitem{amg}
-V. E. Henson and U. M. Yang, 
-\newblock "BoomerAMG: A Parallel Algebraic Multigrid Solver and Preconditioner",
-\newblock Appl. Num. Math. 41 (2002), pp. 155-177. UCRL-JC-141495.
-
-\bibitem{crystalrouter}
-P. Fischer, J. Lottes, D. Pointer, and A. Siegel, 
-\newblock “Petascale algorithms for reactor hydrodynamics,”
-\newblock Journal of Physics: Conference Series, vol. 125, no. 1, p. 012076, 2008.
-
-%+++++++++++++++++
-
-%=====start of network papers===== 
-\bibitem{rosenthal}
-E. Rosenthal and E.A. Leon
-\newblock Characterizing Application Sensitivity to Network Performance.
-\newblock  In {Proceedings of SC14: International
-Conference for High Performance Computing, Networking, Storage and Analysis},
-2014
-
-\bibitem{hoefler-modeling}
-T. Hoefler and W. Gropp and M. Snir and W. Kramer
-\newblock Performance Modeling for Systematic Performance Tuning,
-\newblock 2011
-\newblock Nov.
-\newblock In {Proceedings of International Conference for High Performance Computing, Networking, Storage and Analysis (SC'11), SotP Session,
-}
-
-\bibitem{roth}
-P. Roth and J. Meredith and J. Vetter,
-\newblock "Automated Characterization of Parallel Application Communication Patterns",
-In Proceedings of the 25rd International Symposium on High-performance Parallel and Distributed Computing
-\newblock HPDC '15,
-\newblock 2015,
-
-%=====end of application/mapping papers=====
-
-\bibitem{RF}
-Fiedler, Robert, and Stephen Whalen. 
-\newblock "Improving task placement for applications with 2D, 3D, and 4D virtual Cartesian topologies on 3D torus networks with service nodes." 
-\newblock Proc. Cray User’s Group (2013).
-
-\bibitem{abhinav-sc13} 
-Bhatele, Abhinav and Mohror, Kathryn and Langer, Steven
-H. and Isaacs, Katherine E. 
-\newblock There Goes the Neighborhood: Performance Degradation Due to Nearby Jobs 
-\newblock In {Proceedings of SC13: International
-Conference for High Performance Computing, Networking, Storage and Analysis},
-2013
-
-\bibitem{jose-ipdps15}
-Jokanovic, A.; Sancho, J.C.; Rodriguez, G.; Lucero, A.; Minkenberg, C.; Labarta, J.,
-\newblock "Quiet Neighborhoods: Key to Protect Job Performance Predictability," 
-\newblock in Parallel and Distributed Processing Symposium (IPDPS), 
-\newblock 2015 IEEE International , vol., no., pp.449-459, 25-29 May 2015
-
-\bibitem{skinner}
-D. Skinner and W. Kramer. 
-\newblock Understanding the Causes of Performance Variability in HPC Workloads. 
-\newblock In Proceedings of the IEEE International Workload Characterization Symposium, 
-\newblock 2005, pages 137–149, 2005.
-
-%\bibitem{hpdc14}
-%Prisacari, Bogdan and Rodriguez, German and Heidelberger, Philip and Chen, Dong and Minkenberg, Cyriel and Hoefler, Torsten
-%\newblock Efficient Task Placement and Routing of Nearest Neighbor Exchanges in Dragonfly Networks
-%\newblock In Proceedings of the 23rd International Symposium on High-performance Parallel and Distributed Computing
-%\newblock HPDC '14,
-%\newblock 2014,
-%\newblock Vancouver, BC, Canada
-
-\bibitem{jingjin}
-Wu, Jingjin and Lan, Zhiling and Xiong, Xuanxing and Gnedin, Nickolay Y. and Kravtsov, Andrey V.
-\newblock Hierarchical Task Mapping of Cell-based AMR Cosmology Simulations
-\newblock Proceedings of the International Conference on High Performance Computing, Networking, Storage and Analysis
-\newblock SC '12,
-\newblock 2012,
-\newblock Salt Lake City, Utah
-
-\bibitem{ozan}
-Tuncer, Ozan, Vitus J Leung, Ayse K. Coskun,
-\newblock "PaCMap: Topology Mapping of Unstructured Communication Patterns onto Non-contiguous Allocations," 
-\newblock 29th International Conference on Supercomputing, June 2015.
-
-\bibitem{zhou-ipdps}
-Z. Zhou, X. Yang, Z. Lan, P. Rich, W. Tang, V. Morozov, and N. Desai.
-\newblock Improving Batch Scheduling on Blue Gene/Q by Relaxing 5D Torus Network Allocation Constraints,
-\newblock Proceedings of 29th IEEE International Parallel \& Distributed Processing 
-Symposium,
-\newblock IPDPS'15, 2015.
-\newblock Hyderabad, INDIA
-
-\bibitem{Bogdan-hpdc14}
-Prisacari, Bogdan and Rodriguez, German and Heidelberger, Philip and Chen, Dong and Minkenberg, Cyriel and Hoefler, Torsten
-\newblock Efficient Task Placement and Routing of Nearest Neighbor Exchanges in Dragonfly Networks,
-\newblock Proceedings of the 23rd International Symposium on High-performance Parallel and Distributed Computing,
-\newblock HPDC '14,
-\newblock 2014,
-\newblock  Vancouver, BC, Canada,
-
-\bibitem{oxbow}
-J. S. Vetter, S. Lee, D. Li, G. Marin, C. McCurdy, J. Meredith, P. C. Roth, and K. Spafford. 
-\newblock Quantifying architectural requirements of contemporary extreme-scale scientific applications. 
-\newblock In International Workshop on Performance Modeling, Benchmarking and Simulation of HPC Systems (PMBS13), 
-\newblock Denver, Colorado, 2013.
-
-\bibitem{Dong-SC11}
-Dong Chen; Eisley, N.A.; Heidelberger, P.; Senger, R.M.; Sugawara, Y.; Kumar,
-S.; Salapura, V.; Satterfield, D.L.; Steinmacher-Burow, B.; Parker, J.J., 
-\newblock "The IBM Blue Gene/Q interconnection network and message unit," 
-\newblock High Performance Computing, Networking, Storage and Analysis (SC), 2011 International 
-Conference,
-\newblock pp.1,10, 
-\newblock 12-18 Nov. 2011
-
-\bibitem{Jason-2011}
-Jason Cope, Ning Liu, Sam Lang, Phil Carns, Chris Carothers, Robert Ross.
-\newblock CODES: Enabling Co-design of Multilayer Exascale Storage Architectures, 
-\newblock Proceedings of the Workshop on Emerging Supercomputing Technologies 2011, pp. 303-312
-\newblock 2011
-
-%===simulation papers =====
-
-\bibitem{ross}
-P. D. Barnes, C. D. Carothers, D. R. Jefferson, and J. M. LaPre, 
-\newblock “Warp speed: executing time warp on 1,966,080 cores,”
-\newblock in Proc. of the 2013 ACM SIGSIM Conf. on Principles of Advanced Discrete Simulation (PADS),
-\newblock May 2013, pp. 327–336.
-
-\bibitem{mubarak-sc2012}
-Mubarak, M.; Carothers, C.D.; Ross, R.; Carns, P.,
-\newblock "Modeling a Million-Node Dragonfly Network Using Massively Parallel Discrete-Event Simulation," 
-\newblock High Performance Computing, Networking, Storage and Analysis (SCC), 
-\newblock 2012 SC Companion: 
-\newblock pp.366,376, 10-16 Nov. 2012
-
-\bibitem{codes-torus}
-Mubarak, Misbah, et al.
-\newblock "A case study in using massively parallel simulation for extreme-scale torus network codesign." 
-\newblock in Proceedings of the 2nd ACM SIGSIM/PADS conference on Principles of advanced discrete simulation. 
-\newblock ACM, 2014.
-
-\bibitem{bgq}
-C. Dong, N. Eisley, P. Heidelberger, R. Senger, Y. Sugawara, S. Kumar, V. Salapura, D. Satterfield, B. Steinmacher-Burow, and J. Parker, 
-\newblock “The IBM Blue Gene/Q interconnection fabric,” 
-\newblock Micro, IEEE, vol. 32, no. 1, pp. 32–43, 2012.
-
-\bibitem{tofu}
-Y. Ajima, Y. Takagi, T. Inoue, S. Hiramoto, and T. Shimizu, 
-\newblock “The Tofu Interconnect,” 
-\newblock in 2011 IEEE 19th Annual Symposium on High Performance Interconnects (HOTI), 2011, pp. 87–94.
-
-\bibitem{titan}
-Titan System Overview [online]
-\newblock Available: https://www.olcf.ornl.gov/kbarticles/titan-system-overview
-\newblock (Accessed on: OCT 4th, 2015)
-
-\bibitem{sst}
-Sandia National Labs, “SST DUMPI trace library.” [Online].
-\newblock Available: http://sst.sandia.gov/using dumpi.html 
-\newblock (Accessed on: Sept 3rd, 2015)
-
-\bibitem{simgrid}
-SIMGRID Versatile Simulation of Distributed Systems [Online].
-\newblock Availabale: http://simgrid.gforge.inria.fr
-\newblock (Accessed on: Oct 4th, 2015)
-
-\bibitem{top500}
-“Top500 supercomputing web site.” [Online]. 
-\newblock Available: http://www.top500.org
-\newblock (Accessed on: Oct 4th, 2015)
-
-\end{thebibliography}
+\bibliographystyle{abbrv}
+\bibliography{./reference}
 

### 2015-10-15 submission.tex
update the cited webpage in bib file

@@
 
-We selected three signature applications from DOE Design Forward Project\cite{design forward webpage} as examples to conduct detailed study about their communication patterns.
+We selected three signature applications from DOE Design Forward Project\cite{designforwardwebpage} as examples to conduct detailed study about their communication patterns.
 We use a sophisticated simulation toolkit named CODES, Co-Design of Multi-layer Exascale Storage Architecture\cite{Jason-2011}, from Argonne National Laboratory as a research vehicle to evaluate the performance of these applications with various allocations.
@@
 Most of these applications are parallel and communication intensive, and conform to different communication patterns.
-In this work, we select three representative applications from DOE Design Forward project \cite{design forward webpage}.
+In this work, we select three representative applications from DOE Design Forward project \cite{designforwardwebpage}.
 Each application conforms to a distinctive communication pattern that is commonly seen from applications in HPC workload.
@@
 The applications studied in this work are Algebraic MultiGrid Solver(AMG), Geometric MultiGrid(MultiGrid) and CrystalRouter.
-The communication pattern figures we present in Section \ref{sec:amg}, \ref{sec:crystalrouter}, \ref{sec:multigrid} are generated with the IPM \cite{ipm} data from \cite{design forward webpage}.
+The communication pattern figures we present in Section \ref{sec:amg}, \ref{sec:crystalrouter}, \ref{sec:multigrid} are generated with the IPM \cite{ipm} data from \cite{designforwardwebpage}.
 
@@
 Their representative communication patterns that are commonly seen in the HPC system workload.
-The communication pattern figures we present here are generated with the IPM \cite{ipm} data from \cite{design forward webpage}.
+The communication pattern figures we present here are generated with the IPM \cite{ipm} data from \cite{designforwardwebpage}.
 
@@
 AMG represents a group of applications whose dominant communication pattern is Nearest Neighbor.
-Applications with such similar dominant communication pattern are like PARTISN, SNAP \cite{design forward webpage}.
+Applications with such similar dominant communication pattern are like PARTISN, SNAP \cite{designforwardwebpage}.
 
@@
 MultiGrid represents a group of applications whose dominant communication pattern is Many-to-Many (some literatures also refer it as All-to-All \cite{roth}).
-Applications with such similar dominant communication pattern are like FillBoundary and MiniDFT \cite{design forward webpage}.
+Applications with such similar dominant communication pattern are like FillBoundary and MiniDFT \cite{designforwardwebpage}.
 
@@
 
-%We summarize the benchmarks and MiniApps on DOE design forward webpage\cite{design forward webpage}, as shown in Table \ref{tab: app summary}.
+%We summarize the benchmarks and MiniApps on DOE design forward webpage\cite{designforwardwebpage}, as shown in Table \ref{tab: app summary}.
 According to their dominant communication patterns, there are three categories, namely Nearest Neighbor, All-to-All and 3D Lattice.
@@
 Some research institutes put great effort in characterization of scientific applications.
-The Department of Energy initiate this design forward project aims at the identification of the computational characteristics of the DOE MiniApps developed at various exascale co-design centers \cite{design forward webpage}.
+The Department of Energy initiate this design forward project aims at the identification of the computational characteristics of the DOE MiniApps developed at various exascale co-design centers \cite{designforwardwebpage}.
 In this project, the communication patterns of several DOE full applications and associated mini-applications are studied to provide a more complete snapshot of the DOE workload.

### 2015-10-15 submission.tex
list key finds in introduction

@@
     \end{subfigure}%
-   \caption{Indication of multiple jobs running concurrently with different allocations.
+   \caption{Multiple jobs running concurrently with different allocations.
 Each job is represented by a specific color.
-a) shows the effect of partition-based allocation, which reduce the inter-job interference.
+a) shows the effect of contiguous allocation, which reduce the inter-job interference.
 b) shows non-contiguous allocation, which introduces both intra and inter-job interferences.
 }
-   \label{fig: overview}
+   \label{fig:overview}
 \end{figure}
 
+Users make job submissions to the HPC system with specification about job's required number of nodes and expected runtime.
+Batch scheduler will pick several jobs and dispatch them to the system for running.
+These jobs run concurrently and utilize the system in a space sharing way.
 One of the most prominent problems is network contention among concurrently running jobs.
 The network sharing among concurrently running jobs causes communication variability that results in jobs running slower \cite{abhinav-sc13}.
-The delay of running jobs increases the queueing time of the following submitted jobs, thus leading to low system throughput and utilization \cite{jose-ipdps15}.
+The delay of running jobs also increases the queueing time of the following submitted jobs, thus leading to low system throughput and utilization \cite{jose-ipdps15}.
+This adverse effect of network sharing can be mitigated by providing jobs with isolated allocations and exclusive network resource.
 
-This adverse effect of network sharing can be mitigated by providing jobs with isolated allocations and exclusive network resource.
-The batch scheduler is responsible for assigning job the amount of nodes it requires.
 On the widely used torus-connected HPC systems \cite{bgq}\cite{tofu}\cite{titan}, two allocation strategies are commonly used.
-In the partition based allocation, used in Blue Gene series systems, he scheduler assigns each user job a compact and contiguous set of computing nodes.
-This strategy favors application performance by providing application with isolated partitions and exclusive network connections, thus reducing network contention caused by concurrently running jobs.
+In the contiguous allocation, the scheduler assigns each user job a compact and contiguous set of computing nodes, as shown in Figure \ref{fig:overview_sub1}.
+The partition-based allocation used in Blue Gene series systems is one of this kind \cite{zhou-ipdps}.
+This strategy favors application performance by providing application with isolated partitions and exclusive network connections, thus reducing network contention from concurrently running jobs.
 However, this strategy can cause internal fragmentation (when more nodes are allocated to a job than it requests) and external fragmentation (when sufficient nodes are available for a request, but they can not be allocated contiguously), therefore leading to lower system utilization that is otherwise possible.
-The other strategy is non-contiguous allocation, used by Cray XT/XE series, in which free nodes are assigned to user jobs whether they are contiguous or not.
-Non-contiguous allocation eliminates internal and external fragmentation as seen in partition-based systems, thereby leading to high system utilization.
+The other strategy is non-contiguous allocation, used by Cray XT/XE series \cite{carl-cug}, in which free nodes are assigned to jobs whether they are contiguous or not.
+Figure \ref{fig:overview_sub2} shows the effect of non-contiguous allocation.
+Non-contiguous allocation eliminates internal and external fragmentation as seen in contiguous allocation systems, thereby leading to high system utilization.
 However, it introduces other problems, for example scattering application processes all over the system, causing both inter-job and intra-job contention.
@@
 
-%\textcolor{red}{
-%Partition-based allocation achieves good job performance by sacrificing system utilization, whereas non-contiguous allocation can result in better system utilization but could severely degrade performance of user jobs.
-What is needed is a more flexible strategy that provides allocations that minimize intra/inter-job contention but allow greater system utilization.
-In this work, we propose a guideline for future batch scheduling systems, a smart way to allocate resources with the awareness of jobs communication patterns.
-%
-%Most existing batch schedulers take so few initiative because it been hold under the obscurantism policy.
-Whether it use partition-based or non-contiguous policy, means it chooses the benefit of either system or job over the other.
-Some propose compromised policies like buddy system\cite{buddy system paper} still make the allocation for job blindly without knowing job's communication information.
-%
-%Although HPC applications complexity continue to grow, people could get more knowledge about the applications communication information with the help from various parallel application profiling tools, such as TAU\cite{tau}, MPIP\cite{mpip} and DUMPI\cite{sst}.
-Job's communication trace and performance analysis can be get by using such tool at runtime.
-%
-%The off-line study of job's communication traces could be helpful to develop better allocation policies for batch scheduler to make on-line scheduling/allocating decisions.
-This is because the communication pattern of parallel applications on HPC system can be summarized into a couple of categories\cite{roth}.
-The deep understanding of one application can be helpful to all other applications that conforms to the same communication pattern.
-Job's communication traces study would be extreme helpful to leadership computing facility too.
-The set of applications running on those leadership computing facilities is usually stable, and the workloads from those HPC systems is usually has great repetitiveness.
-We analyzed the workload on Mira, an IBM Blue/Gene Q system at Argonne Leadership Computing Facility, found that most jobs require 2k nodes and about 14 particular jobs consist of the majority its workload\ref{fig: repetitiveness of Mira}.
-In other word, Mira accommodates mostly some particular 14 jobs and most jobs have the same size.
-%}
-
-We envision that future HPC will adopt a flexible job allocation mechanism which combines the best of both partition-based allocation and non-contiguous allocation.
+We envision that future HPC will adopt a flexible job allocation mechanism which combines the best of both contiguous and non-contiguous allocation.
 Such flexible mechanism would make allocation for jobs based on their communication patterns.
 With detail analysis of application's communication patterns, it can be identified which processes are tightly connected and conduct intensive communication within each job.
-Rather than assign each job a compact partition-based allocation or non-contiguous allocation, the ``neighborhood communication" consists of such tightly coupled processes should get a compact allocation.
-The new allocation mechanism only needs to guarantee the locality of the subset of the nodes where those ``neighborhood communication" would reside.
+Rather than assign each job a whole chunk of compact contiguous allocation, only the ``neighborhood communication" consists of such tightly coupled processes need to get a compact allocation.
+The new allocation mechanism only needs to guarantee the compactness of the subset of the nodes where those ``neighborhood communication" would reside.
 
 In this work, we focus on an in-depth analysis of intra- and inter-job communication interferences with  different job allocations on torus-connected HPC systems.
-Torus networks have been extensively used in the current generation of supercomputers because of their linear scaling on per-node cost and competitive communication performance.
+Torus networks have been extensively used in the current generation of supercomputers because of their good scalability and competitive communication performance.
 On the recent Top500 list, 6 of the top 10 supercomputers use a high-radix torus-interconnected network \cite{top500}.
 The current generation of IBM Blue Gene/Q (BG/Q) supercomputer, such as Mira at Argonne Nation Laboratory and Sequoia at  Lawrence Livermore National Laboratory, has its nodes connected in a 5D torus network \cite{bgq}.
-The K computer from Japan uses the “Tofu” system, which is a 6D mesh/torus topology \cite{tofu}.
+The K computer from Japan uses the ``Tofu” system, which is a 6D mesh/torus topology \cite{tofu}.
 Titan, a Cray XK7 supercomputer located at the Oak Ridge Leadership Computing Facility (OLCF), has nodes connected in a 3D torus within the compute partition \cite{titan}.
@@
 We use a sophisticated simulation toolkit named CODES, Co-Design of Multi-layer Exascale Storage Architecture\cite{Jason-2011}, from Argonne National Laboratory as a research vehicle to evaluate the performance of these applications with various allocations.
-We analyze the intra- and inter-job interference by simulating applications running exclusively or concurrently with different allocations.
+We analyze the intra- and inter-job interference by simulating these applications running exclusively and concurrently with different allocations.
+The major finds and insights of this work are the following:
 
-\textbf{replace this by summarizing a couple of key findings!!} According to results of our comprehensive exploration, making allocation with consideration about job's communication pattern is a very promising allocation strategy.
+\begin{itemize}
+    \item We study the communication behaviors and identify the communication patterns of three representative HPC applications.
+    \item We conduct detailed analysis about the intra- and inter-job interference by running three applications with different allocations.
+And we find the relation between application's communication pattern and the cause of interference.
+    \item We show a novel and flexible way to make allocation with awareness of job's communication pattern, which  is every effective to alleviate the interference between concurrently running jobs. 
+\end{itemize}
 
-The rest of this paper organized as follows.
-Section \ref{sec:application study} gives a detailed study of the three applications from DOE Design Forward Project.
+We believe the finds and insights presented in this work would be very useful for the design of future HPC batch job scheduler and resource management module.
+
+The rest of this paper is organized as follows.
+Section \ref{sec:application study} gives a detailed study of the three representative applications from DOE Design Forward Project.
 Section \ref{sec:codes} talks about CODEs as research vehicle for our work.
 Section \ref{sec:config study} shows the performance analysis of three applications on different torus networks.
-Section \ref{sec:interference} provides detailed analysis about the intra/inter-job interference between three applications on torus network.
-Section \ref{sec:discussion} talks about the idea of smart allocation strategy with job communication pattern awareness.
+Section \ref{sec:interference} provides detailed analysis about the intra- and inter-job interference between three applications on torus network with different allocations.
+Section \ref{sec:discussion} talks about the idea of a novel and flexible allocation strategy with job communication pattern awareness.
 Section \ref{sec:related_work} talks about the related work in the area and conclusion will be presented in Section \ref{sec:conclusion}.
@@
 2D mesh, shown as blue in Figure \ref{fig:cont_sub1}, can be cut out from a single layer of the 3D torus.
-2D mesh is a very common allocation shape in torus network for both partition-based and non-contiguous allocation strategies.
+2D mesh is a very common allocation shape in torus network for both contiguous and non-contiguous allocation strategies.
 For example, Cray Application Level Placement Scheduler indexes the nodes in torus network into a list and make allocation by simply works off the this list \cite{carl-cug}.
@@
 Inter-job interference has been identified as one of the major culprits for application's performance variability \cite{abhinav-sc13}\cite{skinner}\cite{rosenthal}.
-Inter-job interference is a more prominent issue for system adopted non-contiguous allocation policy than partition-based system.
+Inter-job interference is a more prominent issue for system adopted non-contiguous allocation policy than system with contiguous allocation.
 Application communication times have been demonstrated to vary from 36\% faster to 69\% slower due to job interference when running on non-contiguous allocation \cite{abhinav-sc13}.
@@
 
-Based on our comprehensive study, we can get following observations 
-\begin{enumerate}
+Based on our comprehensive study, we can get following observations:
+\begin{itemize}
     \item The compact allocation may not guarantee the best performance for every application. 
@@
 Unit size that big enough to preserve the neighborhood communication of the application will result in better performance.
-\end{enumerate}
+\end{itemize}
 
@@
 The advantage of making allocation with consideration about job's communication pattern is obvious.
-First, compared with partition-based policy, the new allocation strategy is more flexible, it doesn't require the system to provide a big partition that can accommodate the whole application, just small set of compact nodes enough for all the "local community" in the application.
+First, compared with contiguous allocation policy, the new allocation strategy is more flexible, it doesn't require the system to provide a big contiguous partition that can accommodate the whole application, just small set of compact nodes enough for all the "local community" in the application.
 On the other hand, the new allocation strategy won't destroy the locality of application's communication pattern.

### 2015-10-15 submission.tex
cut the abstraction short

@@
 Optimizing job allocation and avoiding network sharing are proven techniques to alleviate network contention and performance degradation.
-In this work, we propose a novel job communication pattern aware allocation strategy which serves as a critical module in our future HPC batch scheduler.
-The novelty of our new strategy is that it makes allocation decisions based on job’s communication pattern in addition to system topology and resource availability information.
-The scheduler uses job communication patterns information to decide whether its performance relies heavily on the network or not, and what the potential intra- and inter-job interference would be for a given allocation.
-Thus, the scheduler can pick a preferable resource allocation based on a specific job's communication pattern.
-We validate the effectiveness of our new design using traces collected from three representative parallel applications.
-The idea presented in this paper is widely applicable: HPC parallel applications have common communication patterns, and allocating resource with communication pattern awareness can improve performance on these systems with a variety of network topologies.
+In this work, we choose three representative HPC applications from DOE Design Forward Project and conduct detailed analysis about intra- and inter-job interference with them.
+By scrutinizing the communication behaviors of these applications, we identify the relation between applications' communication pattern and the possible interference when they running with different allocations.
+We propose a novel and flexible strategy that makes allocation with awareness of job's communication pattern.
+The new strategy can effectively alleviate the performance degradation caused by interference.
 

### 2015-10-15 submission.tex
minor chage to some citations

@@
 Supercomputers have hundreds of thousands of nodes, and serve as irreplaceable research vehicle for scientific problems with increasing size and complexity.
-Supercomputers are usually employed as a shared resource to accommodate many parallel applications (also known as jobs) running concurrently \cite{zhou-ipdps} and maintain high system utilization.
+Supercomputers are usually employed as a shared resource to accommodate many parallel applications (also known as jobs) running concurrently \cite{zhou-ipdps}.
+To be effective, they must maintain high system utilization.
 These communication and I/O intensive jobs share the system infrastructure such as network and I/O bandwidth, and inevitably there is contention over these shared resources.
@@
 We selected three signature applications from DOE Design Forward Project\cite{designforwardwebpage} as examples to conduct detailed study about their communication patterns.
-We use a sophisticated simulation toolkit named CODES, Co-Design of Multi-layer Exascale Storage Architecture\cite{Jason-2011}, from Argonne National Laboratory as a research vehicle to evaluate the performance of these applications with various allocations.
+We use a sophisticated simulation toolkit named CODES, Co-Design of Multi-layer Exascale Storage Architecture\cite{Jason-2011} from Argonne National Laboratory, as a research vehicle to evaluate the performance of these applications with various allocations.
 We analyze the intra- and inter-job interference by simulating these applications running exclusively and concurrently with different allocations.
@@
 AMG represents a group of applications whose dominant communication pattern is Nearest Neighbor.
-Applications with such similar dominant communication pattern are like PARTISN, SNAP \cite{designforwardwebpage}.
+Applications with such similar dominant communication pattern are like PARTISN\cite{designforwardwebpage} and SNAP \cite{designforwardwebpage}.
 

### 2015-10-15 submission.tex
abstract tweaks

@@
 \usepackage{subcaption}
+\usepackage{soul}
 %\usepackage[labelfont=bf]{caption}
 \usepackage[font=bf,labelsep=space]{caption}
+
+\newcommand{\TODO}[1]{\hl{TODO: #1}}
+\newcommand{\NOTE}[1]{\hl{NOTE: #1}}
 
@@
 
-Network contention between concurrently running jobs on HPC system is one of the primary causes of performance variability.
-Optimizing job allocation and avoiding network sharing are proven techniques to alleviate network contention and performance degradation.
-In this work, we choose three representative HPC applications from DOE Design Forward Project and conduct detailed analysis about intra- and inter-job interference with them.
-By scrutinizing the communication behaviors of these applications, we identify the relation between applications' communication pattern and the possible interference when they running with different allocations.
+Network contention between concurrently running jobs on HPC systems is a primary cause of performance variability.
+Optimizing job allocation and avoiding network sharing are hence crucial to alleviate the potential degradation in performance.
+In order to do these effectively, an understanding of the interaction between applications, their communication patterns, and the system architecture are required.
+In this work, we choose three representative HPC applications from DOE Design Forward Project and conduct detailed simulations of high-performance network topologies to analyze both intra- and inter-job interference.
+By scrutinizing the communication behaviors of these applications, we identify relationships between these behaviors and the possible interference induced by different job allocation scenarios.
 We propose a novel and flexible strategy that makes allocation with awareness of job's communication pattern.
 The new strategy can effectively alleviate the performance degradation caused by interference.
+\TODO{We propose a flexible allocation strategy?
+If not, we want to weaken this, and instead say something like the results of this research illuminate a path towards access-pattern-aware HPC job allocation.}
 

### 2015-10-15 submission.tex
update fig8

@@
 
-\begin{figure}[t!]
+\begin{figure*}[t!]
     \centering
@@
         \centering
-        \includegraphics[height=1.2in]{figs/cont/allocation}
+        \includegraphics[height=1.2in]{figs/allocshape/allocation}
         \caption{ }
@@
         \centering
-        \includegraphics[height=1.2in]{figs/inter-job/noncont/sharing}
+        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit16}
         \caption{ }
@@
     \end{subfigure}%
-    \caption{(a) Illustrates the contiguous allocation for 64-node job in three different shapes.
+    \begin{subfigure}[t]{0.2\textwidth}
+        \centering
+        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit8}
+        \caption{ }
+        \label{fig:noncont_sub2}
+    \end{subfigure}%
+    \begin{subfigure}[t]{0.2\textwidth}
+        \centering
+        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit2}
+        \caption{ }
+        \label{fig:noncont_sub3}
+    \end{subfigure}%
+    \caption{(a) Contiguous allocation for 64-node job in three different shapes.
 ``Blue" is 2D mesh, ``Green" is 3D-unbalanced cube and ``Red" is 3D-balanced cube.
-(b) Illustrates the non-contiguous allocation.
+(b)-(d) Non-contiguous allocation.
 Each job is represented by a specific color.
-The nodes assigned to different jobs are interleaved, the allocation unit is 4-node.}
-\end{figure}
+The nodes assigned to different jobs are interleaved, the size of allocation unit are 16, 8 and 2.}
+\end{figure*}
 

### 2015-10-15 submission.tex
pass on introduction

@@
 \usepackage{soul}
+\usepackage{verbatim}
 %\usepackage[labelfont=bf]{caption}
@@
 
-The scale of supercomputers keeps growing to accommodate the the demand for computing power required by scientific research areas.
-Supercomputers have hundreds of thousands of nodes, and serve as irreplaceable research vehicle for scientific problems with increasing size and complexity.
-Supercomputers are usually employed as a shared resource to accommodate many parallel applications (also known as jobs) running concurrently \cite{zhou-ipdps}.
+The scale of supercomputers continually grows at a breakneck pace to accommodate the computing power requirements for scientific research.
+Supercomputers have hundreds of thousands of nodes \TODO{you mean tens of thousands?
+Or 100K cores?} and serve as an irreplaceable research vehicle for scientific problems with increasing size and complexity.
+Supercomputers are usually employed as a shared resource to accommodate many parallel applications (jobs) running concurrently \cite{zhou-ipdps} \TODO{cite doesn't fit here, should put somewhere else}.
 To be effective, they must maintain high system utilization.
@@
 a) shows the effect of contiguous allocation, which reduce the inter-job interference.
-b) shows non-contiguous allocation, which introduces both intra and inter-job interferences.
+b) shows non-contiguous allocation, which may introduce both intra and inter-job interferences.
 }
@@
 
-Users make job submissions to the HPC system with specification about job's required number of nodes and expected runtime.
-Batch scheduler will pick several jobs and dispatch them to the system for running.
-These jobs run concurrently and utilize the system in a space sharing way.
-One of the most prominent problems is network contention among concurrently running jobs.
-The network sharing among concurrently running jobs causes communication variability that results in jobs running slower \cite{abhinav-sc13}.
-The delay of running jobs also increases the queueing time of the following submitted jobs, thus leading to low system throughput and utilization \cite{jose-ipdps15}.
-This adverse effect of network sharing can be mitigated by providing jobs with isolated allocations and exclusive network resource.
+When submitting jobs to HPC systems, users request resources in terms of node counts and expected runtime.
+The batch scheduler is then responsible for dispatching them to the system.
+Typically, multiple jobs run concurrently, resulting in the shared use of resources, particularly network and storage.
+A prominent problem with this approach is network contention among concurrently running jobs.
+If the network resources of concurrently running jobs aren't isolated, the resulting sharing can cause communication variability that results in performance degradation for some or all such jobs \cite{abhinav-sc13}.
+This delay can propagate into the queueing time of the following submitted jobs, thus leading to low system throughput and utilization \cite{jose-ipdps15}.
+This adverse effect of network sharing can be mitigated by providing jobs with isolated allocations and exclusive network resources.
 
-On the widely used torus-connected HPC systems \cite{bgq}\cite{tofu}\cite{titan}, two allocation strategies are commonly used.
-In the contiguous allocation, the scheduler assigns each user job a compact and contiguous set of computing nodes, as shown in Figure \ref{fig:overview_sub1}.
-The partition-based allocation used in Blue Gene series systems is one of this kind \cite{zhou-ipdps}.
-This strategy favors application performance by providing application with isolated partitions and exclusive network connections, thus reducing network contention from concurrently running jobs.
-However, this strategy can cause internal fragmentation (when more nodes are allocated to a job than it requests) and external fragmentation (when sufficient nodes are available for a request, but they can not be allocated contiguously), therefore leading to lower system utilization that is otherwise possible.
-The other strategy is non-contiguous allocation, used by Cray XT/XE series \cite{carl-cug}, in which free nodes are assigned to jobs whether they are contiguous or not.
-Figure \ref{fig:overview_sub2} shows the effect of non-contiguous allocation.
-Non-contiguous allocation eliminates internal and external fragmentation as seen in contiguous allocation systems, thereby leading to high system utilization.
-However, it introduces other problems, for example scattering application processes all over the system, causing both inter-job and intra-job contention.
+On the widely used torus-connected HPC systems \cite{bgq,tofu,titan}, two allocation strategies are commonly used.
+The contiguous allocation strategy assigns to each job a compact and contiguous set of computing nodes, as shown in Figure \ref{fig:overview_sub1}.
+The partition-based allocation approach used in Blue Gene series systems is an example of such a strategy  \cite{zhou-ipdps} \TODO{wrong cite - this implies that Zhou designed the algorithm or the BG :) }.
+The contiguous strategy favors application performance through isolated networking within a partition and the locality that implies.
+However, this strategy can cause both internal fragmentation (when more nodes are allocated to a job than it requests) and external fragmentation (when sufficient nodes are available for a request, but they can not be allocated contiguously), therefore leading to lower system utilization than is otherwise possible.
+On the other side of the coin, the non-contiguous allocation strategy, used by the Cray XT/XE series \cite{carl-cug}, assigns free nodes to jobs regardless of contiguity, though of course efforts are made to maximize locality.
+Figure~\ref{fig:overview_sub2} shows an example non-contiguous allocation.
+While eliminating internal and external fragmentation as seen in contiguous allocation systems, in return non-contiguous allocations introduce contention between jobs due to the interleaving of job nodes.
 The non-contiguous node allocation can significantly reduce job performance, especially for communication-intensive ones.
+\TODO{a cite would be good here.}
 
-We envision that future HPC will adopt a flexible job allocation mechanism which combines the best of both contiguous and non-contiguous allocation.
-Such flexible mechanism would make allocation for jobs based on their communication patterns.
-With detail analysis of application's communication patterns, it can be identified which processes are tightly connected and conduct intensive communication within each job.
-Rather than assign each job a whole chunk of compact contiguous allocation, only the ``neighborhood communication" consists of such tightly coupled processes need to get a compact allocation.
-The new allocation mechanism only needs to guarantee the compactness of the subset of the nodes where those ``neighborhood communication" would reside.
+We envision that future HPC schedulers will adopt a flexible job allocation mechanism which combines the best of both contiguous and non-contiguous allocation strategies.
+Such a flexible mechanism would take shared resource needs of jobs into account when making allocation decisions (e.g., network).
+With knowledge and analysis of job communication patterns, it can be identified which jobs require network isolation and locality, and to what degree.
+Then, rather than allocating each job in a ``know-nothing'' manner, one may specialize allocation so that, for example, only the jobs with stringent network needs are given compact, isolated allocations, resulting in maximized utilization and minimized perceivable resource contention effects.
 
-In this work, we focus on an in-depth analysis of intra- and inter-job communication interferences with  different job allocations on torus-connected HPC systems.
-Torus networks have been extensively used in the current generation of supercomputers because of their good scalability and competitive communication performance.
-On the recent Top500 list, 6 of the top 10 supercomputers use a high-radix torus-interconnected network \cite{top500}.
-The current generation of IBM Blue Gene/Q (BG/Q) supercomputer, such as Mira at Argonne Nation Laboratory and Sequoia at  Lawrence Livermore National Laboratory, has its nodes connected in a 5D torus network \cite{bgq}.
-The K computer from Japan uses the ``Tofu” system, which is a 6D mesh/torus topology \cite{tofu}.
+In this work, we focus on an in-depth analysis of intra- and inter-job communication interference with different job allocation strategies on torus-connected HPC systems, which are used on six of the top 10 supercomputers on recent Top500 lists~\cite{top500} \TODO{need to say which date this applies to}.
+The current generation of IBM Blue Gene/Q (BG/Q) supercomputers, such as Mira at Argonne Nation Laboratory and Sequoia at Lawrence Livermore National Laboratory, has its nodes connected in a 5D torus network \cite{bgq}.
+The K computer from Japan uses the ``Tofu'' system, which is a 6D mesh/torus topology \cite{tofu}.
 Titan, a Cray XK7 supercomputer located at the Oak Ridge Leadership Computing Facility (OLCF), has nodes connected in a 3D torus within the compute partition \cite{titan}.
-Although the analysis are based on torus network, the idea conveyed from this work is applicable to networks with different topologies.
+Although our analyses are based on torus network, the ideas conveyed in this work is applicable to networks with different topologies.
 
-We selected three signature applications from DOE Design Forward Project\cite{designforwardwebpage} as examples to conduct detailed study about their communication patterns.
-We use a sophisticated simulation toolkit named CODES, Co-Design of Multi-layer Exascale Storage Architecture\cite{Jason-2011} from Argonne National Laboratory, as a research vehicle to evaluate the performance of these applications with various allocations.
-We analyze the intra- and inter-job interference by simulating these applications running exclusively and concurrently with different allocations.
-The major finds and insights of this work are the following:
+We selected three signature applications from DOE Design Forward Project~\cite{designforwardwebpage} as examples to conduct detailed study about their communication patterns \TODO{is there a better cite for this?}.
+We use a sophisticated simulation toolkit named CODES (standing for Co-Design of Multi-layer Exascale Storage Architectures)~\cite{Jason-2011} as a research vehicle to evaluate the performance of these applications with various allocations in a controlled environment.
+We then analyze the intra- and inter-job interference by simulating these applications running exclusively and concurrently with different allocations.
+We believe the insights presented in this work can be very useful for the design of future HPC batch job schedulers and resource managers.
+
+\NOTE{JJ removed the list of contributions, because they are exactly the same as the preceding paragraph.}
+The observations of this work are the following:
+\begin{comment}
 
@@
     \item We study the communication behaviors and identify the communication patterns of three representative HPC applications.
-    \item We conduct detailed analysis about the intra- and inter-job interference by running three applications with different allocations.
+    \item We conduct detailed analysis about the intra- and inter-job interference by running three applications with different allocation strategies.
 And we find the relation between application's communication pattern and the cause of interference.
     \item We show a novel and flexible way to make allocation with awareness of job's communication pattern, which  is every effective to alleviate the interference between concurrently running jobs.
+\TODO{do we actually do this, or provide the underpinnings for such a method?}
 \end{itemize}
@@
 
+\end{comment}
+
 The rest of this paper is organized as follows.
-Section \ref{sec:application study} gives a detailed study of the three representative applications from DOE Design Forward Project.
-Section \ref{sec:codes} talks about CODEs as research vehicle for our work.
-Section \ref{sec:config study} shows the performance analysis of three applications on different torus networks.
-Section \ref{sec:interference} provides detailed analysis about the intra- and inter-job interference between three applications on torus network with different allocations.
-Section \ref{sec:discussion} talks about the idea of a novel and flexible allocation strategy with job communication pattern awareness.
-Section \ref{sec:related_work} talks about the related work in the area and conclusion will be presented in Section \ref{sec:conclusion}.
+Section~\ref{sec:application study} describes the three representative applications from DOE Design Forward Project for our study.
+Section~\ref{sec:codes} talks about the use of CODES as research vehicle for our work.
+Section~\ref{sec:config study} shows the performance analysis of three applications on different torus networks.
+Section~\ref{sec:interference} provides detailed analysis about the intra- and inter-job interference between three applications on torus network with different allocation strategies.
+Section~\ref{sec:discussion} introduces a path toward communication-pattern aware allocation strategies, given the results of our analysis.
+Section~\ref{sec:related_work} discusses related work.
+Finally, the conclusion is presented in Section~\ref{sec:conclusion}.
 

### 2015-10-15 submission.tex
add separat figs for fig8a
pass on application study

@@
 At its different execution phases, the application's communication behavior may follow different basic patterns respectively.
-There are many profiling tools \cite{tau}\cite{mpip}\cite{scala}\cite{sst}\cite{oxbow} available to capture information regarding communication patterns of parallel applications.
-They can help analyze parallel applications' communication behavior, providing information like the percentage of different MPI operations of the applications, communication topology, the amount data transferred between processes.
+There are many profiling tools available to capture information regarding communication patterns of parallel applications~\cite{tau,mpip,scala,sst,oxbow}.
+They can provide information such as the percentage of different MPI operations of the applications, communication topology, the amount of data transferred between processes.
 
-There are many applications running on HPC systems like Mira \cite{bgq} and Titan \cite{titan}.
-Most of these applications are parallel and communication intensive, and conform to different communication patterns.
-In this work, we select three representative applications from DOE Design Forward project \cite{designforwardwebpage}.
-Each application conforms to a distinctive communication pattern that is commonly seen from applications in HPC workload.
+There are many applications running on HPC systems, such as Mira \cite{bgq} and Titan \cite{titan}, that are communication intensive and conform to various communication patterns.
+In this work, we select three representative applications from the DOE Design Forward project \cite{designforwardwebpage} \TODO{better cite for this?}.
+Each application exhibits a distinctive communication pattern that is commonly seen HPC applications.
 We believe these three applications can represent a big group of applications that conform to similar communication patterns.
-The applications studied in this work are Algebraic MultiGrid Solver(AMG), Geometric MultiGrid(MultiGrid) and CrystalRouter.
-The communication pattern figures we present in Section \ref{sec:amg}, \ref{sec:crystalrouter}, \ref{sec:multigrid} are generated with the IPM \cite{ipm} data from \cite{designforwardwebpage}.
+The applications studied in this work are the Algebraic MultiGrid Solver (AMG), Geometric MultiGrid (MultiGrid) and CrystalRouter.
+The communication pattern figures we present in Section~\ref{sec:amg}, \ref{sec:crystalrouter}, and \ref{sec:multigrid} are generated with the IPM \cite{ipm} data from publicly available Design Forward traces~\cite{designforwardwebpage}.
 
@@
         \caption{Communication Topology}
-        \label{fig: amg communication topology}
+        \label{fig:amg-communication-topology}
     \end{subfigure}%
@@
         \caption{Data Volume}
-        \label{fig: amg data trans}
+        \label{fig:amg-data-trans}
     \end{subfigure}
@@
 
-AMG is the MiniApp of Algebraic MultiGrid Solver, which is a parallel algebraic multi-grid solver for linear systems arising from problems on unstructured mesh physics packages.
+The Algebraic MultiGrid Solver, or AMG, is a parallel algebraic multi-grid solver for linear systems arising from problems on unstructured mesh physics packages.
 It has been derived directly from the BoomerAMG solver that is being developed in the Center for Applied Scientific Computing (CASC) at LLNL \cite{amg}.
-The dominant communication pattern is this regional communication with decreasing message size for different parts of the multi-grid v-cycle.
+The dominant communication pattern is regional communication with decreasing message size for different parts of the multi-grid v-cycle.
 
-Figure \ref{fig: amg communication topology} shows the communication topology of a small scale AMG with 216 MPI Ranks.
-Here we show the communication topology with small scale version because the dominant communication pattern of the application doesn't change with scale.
-We can make the observation that AMG's dominant communication pattern is 3D Nearest Neighbor, each rank has intensive communication with with its 6 neighbors unless it is on the boundary.
-The dominant communication pattern is quite obvious when we identify this relation between ranks.
+Figure~\ref{fig:amg-communication-topology} shows the communication topology of a small scale AMG execution with 216 MPI Ranks.
+Note that the dominant communication pattern of the application doesn't change with scale.
+We can make the observation that AMG's dominant communication pattern is 3D Nearest Neighbor, each rank has intensive communication with up to six neighbors, depending on the boundaries of the ranks.
+Applications with similar pattern include PARTISN\cite{designforwardwebpage} and SNAP \cite{designforwardwebpage} \TODO{need better cites for these...}.
 
-Figure \ref{fig: amg data trans} shows the amount of data each rank transferred in AMG.
-The red line shows the amount of data sent by each rank, the blue line shows the amount of received data.
-The data transferring of each rank is between 7MB to 16MB.
-
-AMG represents a group of applications whose dominant communication pattern is Nearest Neighbor.
-Applications with such similar dominant communication pattern are like PARTISN\cite{designforwardwebpage} and SNAP \cite{designforwardwebpage}.
+Figure~\ref{fig:amg-data-trans} shows the amount of data each rank transferred in the example execution.
+The data transferring of each rank is between 7MB to 16MB, corresponding to whether the rank is on a boundary or not.
 
@@
         \caption{Communication Topology}
-        \label{fig: cr communication topology}
+        \label{fig:cr-communication-topology}
     \end{subfigure}
@@
         \caption{Data Volume}
-        \label{fig: cr data trans}
+        \label{fig:cr-data-trans}
     \end{subfigure}
@@
 
-The second MiniApp we have is CrystalRouter, the extracted communication kernel of the full application Nek5000 \cite{crystalrouter}, which is a spectral element CFD application developed at Argonne National Laboratory.
-It features spectral element multi-grid solvers coupled with a highly scalable, parallel coarse-grid solver that widely used for projects including ocean current modeling, thermal hydraulics of reactor cores, and spatiotemporal chaos.
-CrystalRouter demonstrates the ``Many-to-Many" communication pattern through scalable multi-stage communication process.
+The second MiniApp covered is CrystalRouter, the extracted communication kernel of the full application Nek5000 \cite{crystalrouter}, which is a spectral element CFD application developed at Argonne National Laboratory \TODO{should cite the nek5000 website.
+couldn't find a canonical citation for the software}.
+It features spectral element multi-grid solvers coupled with a highly scalable, parallel coarse-grid solver that is widely used for projects including ocean current modeling, thermal hydraulics of reactor cores, and spatiotemporal chaos.
+CrystalRouter demonstrates the ``many-to-many'' communication pattern through a scalable multi-stage communication process.
 The way CrystalRouter works is nodes compute for a while, then synchronize and communicate, continually alternating between these two types of activities.
 
-The collective communication in CrystalRouter utilizing a recursive doubling approach.
-Ranks in CrystalRouter conform to a n-dimensional hypercube and recursively splitting into (n-1)-dimensional hypercubes, with communication happening with respect to the links that are severed.
-The pattern of this recursive communication can be found in Figure \ref{fig: cr communication topology}.
-By nature of the logarithmic distribution, more data transferring are in the local neighborhood, which consists of 5 ranks.
+The collective communication in CrystalRouter utilizes a recursive doubling approach.
+Ranks in CrystalRouter conform to a $n$-dimensional hypercube and recursively split into ($n$-1)-dimensional hypercubes, with communication occurring along the splitting plane.
+The pattern of this communication can be found in Figure~\ref{fig:cr-communication-topology}.
+By nature of the logarithmic splitting process, a substantial portion of the communication occurs in small neighborhoods of ranks.
+CrystalRouter represents a group of applications whose dominant communication is a hybrid of multi-stage local and hierarchical global communication, and shares similarities with most MPI collective communication implementations.
 
-Figure \ref{fig: cr data trans} shows the amount of data each rank transferred in CrystalRouter.
-The spiky shape indicates there is intensive data transferring in the local neighborhood.
-There is great variance of data transferring among the ranks in CrystalRouter, which is between 20MB to 36MB.
-This is due to the specific feature of CrystalRouter's communication pattern.
-
-CrystalRouter represents a group of applications whose dominant communication is a hybrid of multi-stage local and hierarchical global communication.
+Figure~\ref{fig:cr-data-trans} shows the amount of data each rank transferred in CrystalRouter, which ranges between 20MB to 36MB.
+This is due to the rank counts not matching one-for-one with points in the hypercube.
 
@@
         \caption{Communication Topology}
-        \label{fig: mg communication topology}
+        \label{fig:mg-communication-topology}
     \end{subfigure}
@@
         \caption{Data Volume}
-        \label{fig: mg data trans}
+        \label{fig:mg-data-trans}
     \end{subfigure}
@@
 
-MultiGrid is geometric multi-grid v-cycle from production elliptic solver BoxLib, a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes.
-MultiGrid conforms to Many-to-Many communication pattern with decreasing message size and collectives for different parts of the multi-grid v-cycle.
+MultiGrid is geometric multi-grid v-cycle from the production elliptic solver BoxLib, a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes \TODO{cite?}.
+MultiGrid conforms to many-to-many communication pattern with decreasing message size and collectives for different parts of the multi-grid v-cycle.
 It is widely used for structured grid physics packages.
 
-Figure \ref{fig: mg communication topology} shows the communication topology of MultiGrid with 125 ranks.
-We can see intensive communication along the diagnal that conforms to 3D Nearest Neighbor communication pattern.
-However, there are also considerable amount of data transferring on both sides of the diagnal, which means the dominant communication pattern of MultiGrid is ``Many-to-Many".
+Figure~\ref{fig:mg-communication-topology} shows the communication topology of MultiGrid with 125 ranks.
+We can see intensive communication along the diagonal that resembles nearest neighbor communication, similar to AMG.
+However, the communication topology leads to a greater ``spread'' of communication across the set of ranks, challenging the maximization of communication locality with respect to rank.
+In this sense, it can be considered a ``many-to-many'' pattern.
+Applications with such similar dominant communication pattern are like FillBoundary and MiniDFT \cite{designforwardwebpage}.
 
-Figure\ref{fig: mg data trans} shows the the amount of data transferred between ranks in MultiGrid. The data transferring of each rank is relatively small and stable, which is between 2.8MB to 2.9MB.
-
-MultiGrid represents a group of applications whose dominant communication pattern is Many-to-Many (some literatures also refer it as All-to-All \cite{roth}).
-Applications with such similar dominant communication pattern are like FillBoundary and MiniDFT \cite{designforwardwebpage}.
+Figure~\ref{fig:mg-data-trans} shows the amount of data transferred between ranks in MultiGrid. For this miniapp, The data transferred per rank is relatively small and stable, between 2.8MB to 2.9MB.
 
@@
     \end{subfigure}%
-    \caption{(a) Contiguous allocation for 64-node job in three different shapes.
+    \caption{(a) Contiguous allocation in three different shapes.
 ``Blue" is 2D mesh, ``Green" is 3D-unbalanced cube and ``Red" is 3D-balanced cube.

### 2015-10-16 submission.tex
update citations based on John's TODO

@@
 \newcommand{\NOTE}[1]{\hl{NOTE: #1}}
+\newcommand{\changed}[1]{\textcolor{red}{#1}}
 
@@
 % can use linebreaks \\ within to get better formatting as desired
-\title{Study of Intra- and Inter-Job Interference with Different Job allocations}
+\title{Study of Intra- and Inter-Job Interference with Different Job Allocations}
 % author names and affiliations
@@
 By scrutinizing the communication behaviors of these applications, we identify relationships between these behaviors and the possible interference induced by different job allocation scenarios.
-We propose a novel and flexible strategy that makes allocation with awareness of job's communication pattern.
-The new strategy can effectively alleviate the performance degradation caused by interference.
-\TODO{We propose a flexible allocation strategy?
+\changed{Our comprehensive analysis in this work illuminate a path towards job-communication-pattern-aware HPC resource allocation.} \TODO{We propose a flexible allocation strategy?
 If not, we want to weaken this, and instead say something like the results of this research illuminate a path towards access-pattern-aware HPC job allocation.}
@@
 The scale of supercomputers continually grows at a breakneck pace to accommodate the computing power requirements for scientific research.
-Supercomputers have hundreds of thousands of nodes \TODO{you mean tens of thousands?
-Or 100K cores?} and serve as an irreplaceable research vehicle for scientific problems with increasing size and complexity.
-Supercomputers are usually employed as a shared resource to accommodate many parallel applications (jobs) running concurrently \cite{zhou-ipdps} \TODO{cite doesn't fit here, should put somewhere else}.
+Supercomputers have \changed{tens of thousands} of nodes  and serve as an irreplaceable research vehicle for scientific problems with increasing size and complexity.
+Supercomputers are usually employed as a shared resource to accommodate many parallel applications (jobs) running concurrently \changed{\cite{jose-ipdps15}} \TODO{cite doesn't fit here, should put somewhere else}.
 To be effective, they must maintain high system utilization.
@@
 The contiguous allocation strategy assigns to each job a compact and contiguous set of computing nodes, as shown in Figure \ref{fig:overview_sub1}.
-The partition-based allocation approach used in Blue Gene series systems is an example of such a strategy  \cite{zhou-ipdps} \TODO{wrong cite - this implies that Zhou designed the algorithm or the BG :) }.
+The partition-based allocation approach used in Blue Gene series systems is an example of such a strategy \changed{\cite{bgloverview}} \TODO{wrong cite - this implies that Zhou designed the algorithm or the BG :) }.
 The contiguous strategy favors application performance through isolated networking within a partition and the locality that implies.
@@
 While eliminating internal and external fragmentation as seen in contiguous allocation systems, in return non-contiguous allocations introduce contention between jobs due to the interleaving of job nodes.
-The non-contiguous node allocation can significantly reduce job performance, especially for communication-intensive ones.
-\TODO{a cite would be good here.}
+The non-contiguous node allocation can significantly reduce job performance, especially for communication-intensive ones \changed{\cite{abhinav-sc13}.} \TODO{a cite would be good here.}
 
@@
 
-In this work, we focus on an in-depth analysis of intra- and inter-job communication interference with different job allocation strategies on torus-connected HPC systems, which are used on six of the top 10 supercomputers on recent Top500 lists~\cite{top500} \TODO{need to say which date this applies to}.
+In this work, we focus on an in-depth analysis of intra- and inter-job communication interference with different job allocation strategies on torus-connected HPC systems, which are used on six of the top 10 supercomputers on the \changed{June 2015} Top500 lists~\cite{top500} \TODO{need to say which date this applies to}.
 The current generation of IBM Blue Gene/Q (BG/Q) supercomputers, such as Mira at Argonne Nation Laboratory and Sequoia at Lawrence Livermore National Laboratory, has its nodes connected in a 5D torus network \cite{bgq}.
@@
 We selected three signature applications from DOE Design Forward Project~\cite{designforwardwebpage} as examples to conduct detailed study about their communication patterns \TODO{is there a better cite for this?}.
-We use a sophisticated simulation toolkit named CODES (standing for Co-Design of Multi-layer Exascale Storage Architectures)~\cite{Jason-2011} as a research vehicle to evaluate the performance of these applications with various allocations in a controlled environment.
+We use a sophisticated simulation toolkit named CODES (standing for Co-Design of Multi-layer Exascale Storage Architectures)  \cite{Jason-2011} as a research vehicle to evaluate the performance of these applications with various allocations in a controlled environment.
 We then analyze the intra- and inter-job interference by simulating these applications running exclusively and concurrently with different allocations.
@@
 There are many applications running on HPC systems, such as Mira \cite{bgq} and Titan \cite{titan}, that are communication intensive and conform to various communication patterns.
-In this work, we select three representative applications from the DOE Design Forward project \cite{designforwardwebpage} \TODO{better cite for this?}.
+In this work, we select three representative applications from the DOE Design Forward Project \cite{designforwardwebpage} \TODO{better cite for this?}.
 Each application exhibits a distinctive communication pattern that is commonly seen HPC applications.
@@
 The applications studied in this work are the Algebraic MultiGrid Solver (AMG), Geometric MultiGrid (MultiGrid) and CrystalRouter.
-The communication pattern figures we present in Section~\ref{sec:amg}, \ref{sec:crystalrouter}, and \ref{sec:multigrid} are generated with the IPM \cite{ipm} data from publicly available Design Forward traces~\cite{designforwardwebpage}.
+The communication pattern figures we present in Section~\ref{sec:amg}, \ref{sec:crystalrouter}, and \ref{sec:multigrid} are generated with the IPM~\cite{ipm} data from publicly available traces~\cite{designforwardwebpage}.
 
@@
 We can make the observation that AMG's dominant communication pattern is 3D Nearest Neighbor, each rank has intensive communication with up to six neighbors, depending on the boundaries of the ranks.
-Applications with similar pattern include PARTISN\cite{designforwardwebpage} and SNAP \cite{designforwardwebpage} \TODO{need better cites for these...}.
+Applications with similar pattern include PARTISN \changed{\cite{partisn}} and SNAP \changed{\cite{snap}} \TODO{need better cites for these...}.
 
@@
 
-The second MiniApp covered is CrystalRouter, the extracted communication kernel of the full application Nek5000 \cite{crystalrouter}, which is a spectral element CFD application developed at Argonne National Laboratory \TODO{should cite the nek5000 website.
+The second MiniApp covered is CrystalRouter, the extracted communication kernel of the full application Nek5000~\cite{crystalrouter}, which is a spectral element CFD application developed at Argonne National Laboratory \changed{\cite{nek5000}} \TODO{should cite the nek5000 website.
 couldn't find a canonical citation for the software}.
@@
 
-MultiGrid is geometric multi-grid v-cycle from the production elliptic solver BoxLib, a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes \TODO{cite?}.
+MultiGrid is geometric multi-grid v-cycle from the production elliptic solver BoxLib, a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes \changed{\cite{boxlib}} \TODO{cite?}.
 MultiGrid conforms to many-to-many communication pattern with decreasing message size and collectives for different parts of the multi-grid v-cycle.
@@
 In this sense, it can be considered a ``many-to-many'' pattern.
-Applications with such similar dominant communication pattern are like FillBoundary and MiniDFT \cite{designforwardwebpage}.
+Applications with such similar dominant communication pattern are like FillBoundary, another PDE solver code in \cite{boxlib}.
 

### 2015-10-16 submission.tex
update some todos, fixup some cites

@@
 By scrutinizing the communication behaviors of these applications, we identify relationships between these behaviors and the possible interference induced by different job allocation scenarios.
-\changed{Our comprehensive analysis in this work illuminate a path towards job-communication-pattern-aware HPC resource allocation.} \TODO{We propose a flexible allocation strategy?
-If not, we want to weaken this, and instead say something like the results of this research illuminate a path towards access-pattern-aware HPC job allocation.}
+\changed{Our comprehensive analyses in this work illuminate a path towards job-communication-pattern-aware HPC resource allocation.}
 
@@
 Supercomputers have \changed{tens of thousands} of nodes  and serve as an irreplaceable research vehicle for scientific problems with increasing size and complexity.
-Supercomputers are usually employed as a shared resource to accommodate many parallel applications (jobs) running concurrently \changed{\cite{jose-ipdps15}} \TODO{cite doesn't fit here, should put somewhere else}.
+Supercomputers are usually employed as a shared resource to accommodate many parallel applications (jobs) running concurrently \changed{\cite{jose-ipdps15}} \TODO{cite (still) doesn't fit here.
+I don't think you need to defend this sentence with a cite}.
 To be effective, they must maintain high system utilization.
@@
 The contiguous allocation strategy assigns to each job a compact and contiguous set of computing nodes, as shown in Figure \ref{fig:overview_sub1}.
-The partition-based allocation approach used in Blue Gene series systems is an example of such a strategy \changed{\cite{bgloverview}} \TODO{wrong cite - this implies that Zhou designed the algorithm or the BG :) }.
+The partition-based allocation approach used in Blue Gene series systems is an example of such a strategy \changed{\cite{bgloverview}} \TODO{this one is OK.
+Tried seeing if there were a more recent cite for this, but didn't come up with anything}.
 The contiguous strategy favors application performance through isolated networking within a partition and the locality that implies.
@@
 While eliminating internal and external fragmentation as seen in contiguous allocation systems, in return non-contiguous allocations introduce contention between jobs due to the interleaving of job nodes.
-The non-contiguous node allocation can significantly reduce job performance, especially for communication-intensive ones \changed{\cite{abhinav-sc13}.} \TODO{a cite would be good here.}
+The non-contiguous node allocation can significantly reduce job performance, especially for communication-intensive ones \changed{\cite{abhinav-sc13}.}
 
@@
 
-In this work, we focus on an in-depth analysis of intra- and inter-job communication interference with different job allocation strategies on torus-connected HPC systems, which are used on six of the top 10 supercomputers on the \changed{June 2015} Top500 lists~\cite{top500} \TODO{need to say which date this applies to}.
+In this work, we focus on an in-depth analysis of intra- and inter-job communication interference with different job allocation strategies on torus-connected HPC systems, which are used on six of the top 10 supercomputers on the \changed{June 2015} Top500 lists~\cite{top500}.
 The current generation of IBM Blue Gene/Q (BG/Q) supercomputers, such as Mira at Argonne Nation Laboratory and Sequoia at Lawrence Livermore National Laboratory, has its nodes connected in a 5D torus network \cite{bgq}.
@@
 Titan, a Cray XK7 supercomputer located at the Oak Ridge Leadership Computing Facility (OLCF), has nodes connected in a 3D torus within the compute partition \cite{titan}.
-Although our analyses are based on torus network, the ideas conveyed in this work is applicable to networks with different topologies.
+Although our analyses are based on torus networks, the ideas conveyed in this work is applicable to networks with different topologies.
 
-We selected three signature applications from DOE Design Forward Project~\cite{designforwardwebpage} as examples to conduct detailed study about their communication patterns \TODO{is there a better cite for this?}.
-We use a sophisticated simulation toolkit named CODES (standing for Co-Design of Multi-layer Exascale Storage Architectures)  \cite{Jason-2011} as a research vehicle to evaluate the performance of these applications with various allocations in a controlled environment.
+We selected three signature applications from the DOE Design Forward Project~\cite{designforwardwebpage} as examples to conduct detailed study about their communication patterns \TODO{checked with some folks and there actually isn't a better citation for now.}.
+We use a sophisticated simulation toolkit named CODES (standing for Co-Design of Multi-layer Exascale Storage Architectures)~\cite{Jason-2011} as a research vehicle to evaluate the performance of these applications with various allocations in a controlled environment.
 We then analyze the intra- and inter-job interference by simulating these applications running exclusively and concurrently with different allocations.
@@
 \NOTE{JJ removed the list of contributions, because they are exactly the same as the preceding paragraph.}
+\begin{comment}
 The observations of this work are the following:
-\begin{comment}
 
@@
 There are many applications running on HPC systems, such as Mira \cite{bgq} and Titan \cite{titan}, that are communication intensive and conform to various communication patterns.
-In this work, we select three representative applications from the DOE Design Forward Project \cite{designforwardwebpage} \TODO{better cite for this?}.
-Each application exhibits a distinctive communication pattern that is commonly seen HPC applications.
-We believe these three applications can represent a big group of applications that conform to similar communication patterns.
-The applications studied in this work are the Algebraic MultiGrid Solver (AMG), Geometric MultiGrid (MultiGrid) and CrystalRouter.
-The communication pattern figures we present in Section~\ref{sec:amg}, \ref{sec:crystalrouter}, and \ref{sec:multigrid} are generated with the IPM~\cite{ipm} data from publicly available traces~\cite{designforwardwebpage}.
+In this work, we select three representative applications from the DOE Design Forward Project.
+Each application exhibits a distinctive communication pattern that is commonly seen in HPC applications.
+We believe that the communication patterns of these applications are representative of a wide array of applications run on leadership-class machines.
+Specifically, we study the Algebraic MultiGrid Solver (AMG), Geometric MultiGrid (MultiGrid) and CrystalRouter MiniApps.
+The communication pattern figures we present in Section~\ref{sec:amg}, \ref{sec:crystalrouter}, and \ref{sec:multigrid} are generated with the IPM~\cite{ipm} data gathered from publicly available traces~\cite{designforwardwebpage}.
 
@@
     \centering
-    \begin{subfigure}[t]{0.22\textwidth}
+    \begin{subfigure}[t]{0.24\textwidth}
         \centering
@@
     ~
-    \begin{subfigure}[t]{0.22\textwidth}
+    \begin{subfigure}[t]{0.24\textwidth}
         \centering
@@
 We can make the observation that AMG's dominant communication pattern is 3D Nearest Neighbor, each rank has intensive communication with up to six neighbors, depending on the boundaries of the ranks.
-Applications with similar pattern include PARTISN \changed{\cite{partisn}} and SNAP \changed{\cite{snap}} \TODO{need better cites for these...}.
+Applications with similar pattern include PARTISN \changed{\cite{partisn}} and SNAP \changed{\cite{snap}}.
 
@@
 
-The second MiniApp covered is CrystalRouter, the extracted communication kernel of the full application Nek5000~\cite{crystalrouter}, which is a spectral element CFD application developed at Argonne National Laboratory \changed{\cite{nek5000}} \TODO{should cite the nek5000 website.
-couldn't find a canonical citation for the software}.
+The second MiniApp covered is CrystalRouter, the extracted communication kernel of the full application Nek5000~\cite{crystalrouter}, which is a spectral element CFD application developed at Argonne National Laboratory \changed{\cite{nek5000}}.
 It features spectral element multi-grid solvers coupled with a highly scalable, parallel coarse-grid solver that is widely used for projects including ocean current modeling, thermal hydraulics of reactor cores, and spatiotemporal chaos.
@@
 
-MultiGrid is geometric multi-grid v-cycle from the production elliptic solver BoxLib, a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes \changed{\cite{boxlib}} \TODO{cite?}.
+MultiGrid is geometric multi-grid v-cycle from the production elliptic solver BoxLib, a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes \changed{\cite{boxlib}}.
 MultiGrid conforms to many-to-many communication pattern with decreasing message size and collectives for different parts of the multi-grid v-cycle.
@@
 
-\bibliographystyle{abbrv}
+\bibliographystyle{IEEEtran}
 \bibliography{./reference}

### 2015-10-16 submission.tex
update fig10, remove check todos

@@
 Supercomputers have \changed{tens of thousands} of nodes  and serve as an irreplaceable research vehicle for scientific problems with increasing size and complexity.
-Supercomputers are usually employed as a shared resource to accommodate many parallel applications (jobs) running concurrently \changed{\cite{jose-ipdps15}} \TODO{cite (still) doesn't fit here.
-I don't think you need to defend this sentence with a cite}.
+Supercomputers are usually employed as a shared resource to accommodate many parallel applications (jobs) running concurrently.
 To be effective, they must maintain high system utilization.
@@
 The contiguous allocation strategy assigns to each job a compact and contiguous set of computing nodes, as shown in Figure \ref{fig:overview_sub1}.
-The partition-based allocation approach used in Blue Gene series systems is an example of such a strategy \changed{\cite{bgloverview}} \TODO{this one is OK.
-Tried seeing if there were a more recent cite for this, but didn't come up with anything}.
+The partition-based allocation approach used in Blue Gene series systems is an example of such a strategy \changed{\cite{bgloverview}}.
 The contiguous strategy favors application performance through isolated networking within a partition and the locality that implies.
@@
 
-We selected three signature applications from the DOE Design Forward Project~\cite{designforwardwebpage} as examples to conduct detailed study about their communication patterns \TODO{checked with some folks and there actually isn't a better citation for now.}.
+We selected three signature applications from the DOE Design Forward Project~\cite{designforwardwebpage} as examples to conduct detailed study about their communication patterns.
 We use a sophisticated simulation toolkit named CODES (standing for Co-Design of Multi-layer Exascale Storage Architectures)~\cite{Jason-2011} as a research vehicle to evaluate the performance of these applications with various allocations in a controlled environment.
@@
 
+\TODO{This par is OUTDATED}.
 In non-contiguous allocation systems, the nodes assigned to each job could be scatter anywhere in the system.
@@
 
-Figure \ref{fig:cont-amg} shows that AMG is insensitive to the allocation's contiguity.
+\TODO{This par is OUTDATED} Figure \ref{fig:cont-amg} shows that AMG is insensitive to the allocation's contiguity.
 Non-contiguous allocation can serve AMG equally well as contiguous allocation, as long as the allocation unit in non-contiguous allocation is big enough to preserve the locality of AMG.
@@
 
-%\emph{OBSERVATION}
-The performance of application conforms to Nearest Neighbor communication pattern keep relatively stable with different allocations.
+\TODO{This par is OUTDATED} The performance of application conforms to Nearest Neighbor communication pattern keep relatively stable with different allocations.
 Non-contiguous allocation won't aggravate intra-job interference as long as the unit size can preserve the locality in application's communication topology graph.
@@
         \centering
-        \includegraphics[height=1.5in]{figs/intra-job/cont/amg_box}
+        \includegraphics[height=1.5in]{figs/diffmapping/amg}
         \caption{AMG}
-        \label{fig:cont-amg}
+        \label{fig:diffmap-amg}
     \end{subfigure}%
@@
         \centering
-        \includegraphics[height=1.5in]{figs/intra-job/cont/cr_box}
+        \includegraphics[height=1.5in]{figs/diffmapping/cr}
         \caption{CrystalRouter}
-        \label{fig:cont-cr}
+        \label{fig:diffmap-cr}
     \end{subfigure}%
@@
         \centering
-        \includegraphics[height=1.5in]{figs/intra-job/cont/mg_box}
+        \includegraphics[height=1.5in]{figs/diffmapping/mg}
         \caption{MultiGrid}
-        \label{fig:cont-mg}
+        \label{fig:diffmap-mg}
     \end{subfigure}%
-   \caption{Intra-job study: data transfer time of 3 apps with Non/Contiguous allocation}
+   \caption{\TODO{This fig is UPdated} Intra-job study: data transfer time of 3 apps on 3D-balanced allocation with different mapping strategies.}
    \label{fig: cont-time-box}
 \end{figure*}
-
-%Apparently, the size of allocation unit can affect application's performance.
-The good allocation unit size for each application is decided by application's communication pattern.
-Once the locality of application's communication pattern could be identified, the size of allocation unit should be able to preserve the locality of application.
-Figure \ref{fig: chunk-time-box} shows the data transfer time of 3 App with different allocation unit size.
-%
-%For AMG, unit size of 16 and 8, are equally good, since AMG's locality is 6-rank based.
-Unit size of 2 can cause performance loss to AMG.
-For CrystalRouter data transfer time increases as the unit size decreases from 16 to 8 and 2.
-Unit size of 16 will preserve not only the 5-rank local community in CrystalRouter communication pattern, but also part of the diagonal global data transfer.
-Unit size of 8 can only guarantee the locality of that 5-rank community.
-And unit size of 2 will lose all the locality of CrystalRouter and make the intra-job communication less efficient.
-For MultiGrid, smaller unit size prolong the data transfer time.
-The difference between different unit size is not obvious, since MultiGrid is nearly all-to-all dominant.
-%
-%
-%\begin{figure*}[t!]
-%    \centering
-%    \begin{subfigure}[t]{0.3\textwidth}
-%        \centering
-%        \includegraphics[height=1.2in]{figs/intra-job/chunk/amg_box}
-%        \caption{AMG}
-%        \label{fig:chunk-amg-box}
-%    \end{subfigure}%
-%    \hspace{1em}%
-%    \begin{subfigure}[t]{0.3\textwidth}
-%        \centering
-%        \includegraphics[height=1.2in]{figs/intra-job/chunk/cr_box}
-%        \caption{CrystalRouter}
-%        \label{fig:chunk-cr-box}
-%    \end{subfigure}%
-%    \hspace{1em}%
-%    \begin{subfigure}[t]{0.3\textwidth}
-%        \centering
-%        \includegraphics[height=1.2in]{figs/intra-job/chunk/mg_box}
-%        \caption{MultiGrid}
-%        \label{fig:chunk-mg-box}
-%    \end{subfigure}%
-%   \caption{Data Transfer Time of 3 Apps with Different Unit Size}
-%   \label{fig: chunk-time-box}
-%\end{figure*}
 

### 2015-10-16 submission.tex

@@
 
-\section{Configuration Study }
+\section{Configuration Study}
 \label{sec:config study}
 
-In this section, we study the communication behavior of three applications on torus network with different dimensionality and bandwidth configurations.
+In this section, we study the communication behavior of three applications on torus networks with different dimensionality and bandwidth configurations.
+Torus networks are $k$-ary $n$-cubes, with $k^{n}$ nodes in total arranged in an $n$-dimensional grid having $k$ nodes in each dimension.
+Each node has $2\times n$ direct linked neighbor nodes with wraparound.
+In these and all following experiments, non-adaptive routing is used.
+While making specific recommendations as to an ``optimal'' network configuration for a given application is out of the scope of this study, the experiments do highlight the nuances inherent in design choices such as network diameter and bandwidth planning.
 
-The topology of torus network is k-ary n-cube, with $k^n$ nodes in total arranged in an n-dimensional grid having k nodes in each dimension.
-Each node has 2$\times$n direct linked neighbor nodes.
-The torus network performance is determined by its dimensionality and link bandwidth.
-As the increase of the dimensionality of torus network, so does the number of links connected with each node.
-The increased aggregated bandwidth of each node will definitely reduce the data transfer time of each rank in the application.
-Figure \ref{fig:dimensionality-study} shows the performance of the AMG, CrystalRouter and MultiGrid on a 2K node torus network model with a 3D torus (16$\times$16$\times$8), a 5D torus (8$\times$4$\times$4$\times$4$\times$4), and a 7D torus (4$\times$4$\times$4$\times$4$\times$2$\times$2$\times$2).
-The bandwidth of direct link between nodes is 2GiB/s in each direction \cite{bgq}, thus, the aggregated bandwidth is 12GiB/s per node in 3D torus, 20GiB/s per node in 5D torus, and 28GiB/s per node in 7D torus.
-
-We can see from Figure \ref{fig:dimensionality-study} that the data transfer time of those three applications are greatly reduced as the dimensionality increases.
-The increased aggregated bandwidth of each node can accelerate the data transfer.
-We allocate each rank within each application linearly on the 3D torus.
-However, due to the variance of data transfer volume of each rank, AMG's data transfer time has great dispersion, shown as the outliers above the boxes in Figure \ref{fig:dimen-amg}.
-There is few dispersion of MultiGrid, shown in Figure \ref{fig:dimen-mg}, since data transfer of each rank in MultiGrid varies slightly  between 2.8MB to 2.9MB.
-CrystalRouter also has few dispersion as shown in Figure \ref{fig:dimen-cr}, this is due to the majority of data transferring is in the local neighborhood, which only take few hops.
+We perform three sets of experiments.
+Each looks at per-rank distribution of performance of AMG, CrystalRouter and MultiGrid on a set of 2048-node torus models, one with $n=3$ ($16^{2} \times 8$), one with $n=5$ ($8\times 4^{4}$), and one with $n=7$ ($4^{4} \times 2^{3}$).
+Performance in this context is defined by the total message transfer time for each sending rank, which we are able to precisely measure due to the simulation environment providing us with an absolute notion of time.
+Note that we simply perform linear rank assignment, one per node -- later sections investigate rank and node mapping effects.
+In the first, shown in Figure~\ref{fig:dimensionality-study}, we keep the link bandwidth constant at 2GiB/s in each direction, corresponding to Blue Gene/Q systems~\cite{bgq}, leading to aggregate node bandwidths of 12GiB/s, 20GiB/s, and 28GiB/s for the 3D, 5D, and 7D tori, respectively.
+In the second, shown in Figure~\ref{fig:bandwidth-study}, we fix the torus dimensionality to 5D.
+In the third, shown in Figure~\ref{fig: bandwidth-time-box}, we instead fix the per-node aggregate bandwidth (and thus the system aggregate bandwidth) and modify the torus dimensionality, resulting in per-link bandwidths shown in Table~\ref{tab: fix-bandwidth}.
 
@@
     \end{subfigure}%
-   \caption{Data Transfer Time of AMG, CrystalRouter and MultiGrid on 3D, 5D and 7D torus network.
-The network bandwidth increases as the dimensionality grows.
-Due to the data volume sent/received by each rank greatly varies, there is great dispersion of AMG's data transfer time, shown as outliers above the boxes in Figure \ref{fig:dimen-amg}.}
+   \caption{Data Transfer Time of AMG, CrystalRouter and MultiGrid on 3D, 5D and 7D torus network, with constant per-link bandwidth.}
    \label{fig:dimensionality-study}
@@
 
-Then we use fixed dimensionality to study the impact of increased bandwidth.
-We run these three applications on 5D torus with direct link bandwidth increases from 2GiB/s, to 4GiB/s and 8GiB/s.
-As shown in Figure \ref{fig:bandwidth-study}, the data transfer time of each application can be greatly reduced as the bandwidth increases.
+In the fixed link bandwidth experiment, shown in Figure~\ref{fig:dimensionality-study}, the data transfer time of the three applications are significantly reduced as the dimensionality increases, so the applications can easily make use of the increased dimensionality (lower hops on average) and/or aggregate bandwidth (more links being concurrently used).
+Interestingly, AMG has the greatest degree of outliers compared to the others.
+We expect this to be a combination of non-optimal rank mappings causing a greater degree of variance due to the highly regular communication pattern and the distribution of message sizes.
+MultiGrid seems to benefit the most in terms of reducing the spread of the time distribution, though the median message time remains roughly within a factor of two.
+The time distribution in CrystalRouter seems to accrue the most benefits, likely due to the combination of cross-cutting and localized per-rank traffic.
 
@@
 
-We also study the impact of dimensionality with fixed aggregated bandwidth.
-We use 3D, 5D and 7D torus network with the same per-node aggregated bandwidth, which is 28Gb/s.
-With the same aggregated bandwidth, higher dimensionality means shorter diameter and pair-wise distance between nodes.
-The bandwidth between nodes in each torus network are shown in Table \ref{tab: fix-bandwidth}
-
-\begin{table}[ht]
-\begin{center}
-\caption{Torus networks with same aggregated bandwidth} 
-\label{tab: fix-bandwidth}
-\begin{tabular}{l c c c} 
-\toprule % Top horizontal line
-\toprule
-&\multicolumn{3}{c}{Bandwidth } \\
-\cmidrule(l){2-4}
-Dimension  & direct & aggregate &\\ % Column names row
-\midrule % In-table horizontal line
-3D      & 4.67Gb/s  & 28Gb/s  & \\  % Content row 1
-\midrule % In-table horizontal line
-5D    & 2.8Gb/s  & 28Gb/s  &\\ % Content row 1
-\midrule % In-table horizontal line
-7D    & 2Gb/s & 28Gb/s & \\
-\midrule
-\bottomrule % Bottom horizontal line
-\end{tabular}
-\end{center}
-\end{table}
-
-As we see from Figure \ref{fig:samebd-amg}, the quartiles of ``3D" box is a little bit higher (1\%) than that of ``5D", which means increasing dimensionality from 3D to 5D has little improvement on the average data transfer time for AMG.
-However, there is obvious reduction in those outliers between 3D and 5D.
-When dimensionality increases up to 7D, the improvement in terms of average data transfer time is still small, but more outliers above the top whisker can be spotted.
-
-The performance of CrystalRouter actually deteriorate as the torus dimensionality increasing, shown in Figure \ref{fig:samebd-cr}.
-This is due to the direct link bandwidth reduces from 4.67Gb/s to 2Gb/s, as dimension increases from 3D to 7D.
-CrystalRouter's large amount of intensive local data transfer benefit more from higher direct link bandwidth than its global transfer from lower diameter and shorter pair-wise distance between nodes.
-
-The performance of MultiGrid is gradually improved as the dimensionality of torus network increases, shown in Figure \ref{fig:samebd-mg}.
-The reduced diameter and shorten pair-wise distance between nodes would make the ``Many-to-Many" global data transfer in MultiGrid more efficient.
-
-We can get the observation that higher dimensionality of torus network would improve the performance of application with ``Many-to-Many" communication patterns, while application with intensive local communication like ``Nearest Neighbor" won't benefit much from higher dimensionality.
+In the fixed dimensionality experiment, shown in Figure~\ref{fig:bandwidth-study}, the increased per-link bandwidth, expectedly, benefits each application roughly equally, reducing the time distributions linearly.
 
@@
 \end{figure*}
+
+\begin{table}[ht]
+\begin{center}
+\caption{Torus networks with same aggregated bandwidth} 
+\label{tab: fix-bandwidth}
+\begin{tabular}{l c c c} 
+\toprule % Top horizontal line
+\toprule
+&\multicolumn{3}{c}{Bandwidth } \\
+\cmidrule(l){2-4}
+Dimension  & direct & aggregate &\\ % Column names row
+\midrule % In-table horizontal line
+3D      & 4.67Gb/s  & 28Gb/s  & \\  % Content row 1
+\midrule % In-table horizontal line
+5D    & 2.8Gb/s  & 28Gb/s  &\\ % Content row 1
+\midrule % In-table horizontal line
+7D    & 2Gb/s & 28Gb/s & \\
+\midrule
+\bottomrule % Bottom horizontal line
+\end{tabular}
+\end{center}
+\end{table}
+
+In the fixed aggregate bandwidth experiment, with configuration in Table~\ref{tab: fix-bandwidth} and results shown in Figure~\ref{fig: bandwidth-time-box}, the results are more nuanced.
+At least for these applications, the benefits from increased dimensionality do not outweigh the reduction in per-link bandwidth.
+Multigrid does see benefits in the increased dimensionality, particularly with respect to outliers, but has a similar median message time.
+CrystalRouter reduces performance as the dimensionality increases, but only by a small degree.
+AMG remains roughly the same, minus a dip in the maximum message time that could be due to rank-to-node mapping changes.
+
+\begin{comment}
+
+Then we use fixed dimensionality to study the impact of increased bandwidth.
+We run these three applications on 5D torus with direct link bandwidth increases from 2GiB/s, to 4GiB/s and 8GiB/s.
+As shown in Figure \ref{fig:bandwidth-study}, the data transfer time of each application can be greatly reduced as the bandwidth increases.
+
+We also study the impact of dimensionality with fixed aggregated bandwidth.
+We use 3D, 5D and 7D torus network with the same per-node aggregated bandwidth, which is 28Gb/s.
+With the same aggregated bandwidth, higher dimensionality means shorter diameter and pair-wise distance between nodes.
+The bandwidth between nodes in each torus network are shown in Table \ref{tab: fix-bandwidth}
+
+As we see from Figure \ref{fig:samebd-amg}, the quartiles of ``3D" box is a little bit higher (1\%) than that of ``5D", which means increasing dimensionality from 3D to 5D has little improvement on the average data transfer time for AMG.
+However, there is obvious reduction in those outliers between 3D and 5D.
+When dimensionality increases up to 7D, the improvement in terms of average data transfer time is still small, but more outliers above the top whisker can be spotted.
+
+The performance of CrystalRouter actually deteriorate as the torus dimensionality increasing, shown in Figure \ref{fig:samebd-cr}.
+This is due to the direct link bandwidth reduces from 4.67Gb/s to 2Gb/s, as dimension increases from 3D to 7D.
+CrystalRouter's large amount of intensive local data transfer benefit more from higher direct link bandwidth than its global transfer from lower diameter and shorter pair-wise distance between nodes.
+
+The performance of MultiGrid is gradually improved as the dimensionality of torus network increases, shown in Figure \ref{fig:samebd-mg}.
+The reduced diameter and shorten pair-wise distance between nodes would make the ``Many-to-Many" global data transfer in MultiGrid more efficient.
+
+We can get the observation that higher dimensionality of torus network would improve the performance of application with ``Many-to-Many" communication patterns, while application with intensive local communication like ``Nearest Neighbor" won't benefit much from higher dimensionality.
+
+\end{comment}
 

### 2015-10-16 submission.tex
update fig2,3,4 and fig10a

@@
     \centering
-    \begin{subfigure}[t]{0.24\textwidth}
+    \begin{subfigure}[t]{0.25\textwidth}
         \centering
-        \includegraphics[height=1.2in]{figs/appstudy/amg/amg_pip}
-        \caption{Communication Topology}
+        \includegraphics[height=1.5in]{figs/appstudy/amg/amg_pip}
+        \caption{}
         \label{fig:amg-communication-topology}
@@
         \includegraphics[height=1.2in]{figs/appstudy/amg/amg_data_transfer}
-        \caption{Data Volume}
+        \caption{}
         \label{fig:amg-data-trans}
     \end{subfigure}
-    \caption{AMG. (a) rank-to-rank communication topology graph. (b) the data sent/received by each rank in AMG.}
+    \caption{AMG.
+(a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in AMG.
+(b) The amount of data sent/received by each rank in AMG.
+Y-axis labels the amount of data transferred in MB, x-axis is the rank.
+}
 \end{figure}
@@
     \centering
-    \begin{subfigure}[t]{0.22\textwidth}
+    \begin{subfigure}[t]{0.25\textwidth}
         \centering
-        \includegraphics[height=1.2in]{figs/appstudy/cr/cr_pip}
+        \includegraphics[height=1.5in]{figs/appstudy/cr/cr_pip}
         \caption{Communication Topology}
@@
     \end{subfigure}
-    \caption{CrystalRouter. (a) rank-to-rank communication topology graph. (b) the data sent/received by each rank in CrystalRouter. }
+    \caption{CrystalRouter.
+(a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in CrystalRouter.
+(b) The amount of data sent/received by each rank in CrystalRouter.
+Y-axis labels the amount of data transferred in MB, x-axis is the rank.
+}
 \end{figure}
@@
     \centering
-    \begin{subfigure}[t]{0.22\textwidth}
+    \begin{subfigure}[t]{0.25\textwidth}
         \centering
-        \includegraphics[height=1.2in]{figs/appstudy/mg/mg_pip}
+        \includegraphics[height=1.5in]{figs/appstudy/mg/mg_pip}
         \caption{Communication Topology}
@@
     \end{subfigure}
-    \caption{MultiGrid. (a) rank-to-rank communication topology graph. (b) the data sent/received by each rank in MultiGrid.}
+    \caption{MultiGrid.
+(a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in MultiGrid.
+(b) The amount of data sent/received by each rank in MultiGrid.
+Y-axis labels the amount of data transferred in MB, x-axis is the rank.
+}
 \end{figure}

### 2015-10-17 submission.tex
integrated lan's commets for section1,2,3,4

@@
 By scrutinizing the communication behaviors of these applications, we identify relationships between these behaviors and the possible interference induced by different job allocation scenarios.
-\changed{Our comprehensive analyses in this work illuminate a path towards job-communication-pattern-aware HPC resource allocation.}
+Our comprehensive analyses in this work illuminate a path towards job-communication-pattern-aware HPC resource allocation.
 
@@
 The scale of supercomputers continually grows at a breakneck pace to accommodate the computing power requirements for scientific research.
-Supercomputers have \changed{tens of thousands} of nodes  and serve as an irreplaceable research vehicle for scientific problems with increasing size and complexity.
+Supercomputers have tens of thousands of nodes  and serve as an irreplaceable research vehicle for scientific problems with increasing size and complexity.
 Supercomputers are usually employed as a shared resource to accommodate many parallel applications (jobs) running concurrently.
-To be effective, they must maintain high system utilization.
-These communication and I/O intensive jobs share the system infrastructure such as network and I/O bandwidth, and inevitably there is contention over these shared resources.
+These parallel jobs share the system infrastructure such as network and I/O bandwidth, and inevitably there is contention over these shared resources.
 As supercomputers continue to evolve, these shared resources are increasingly the bottleneck for performance.
@@
 On the widely used torus-connected HPC systems \cite{bgq,tofu,titan}, two allocation strategies are commonly used.
-The contiguous allocation strategy assigns to each job a compact and contiguous set of computing nodes, as shown in Figure \ref{fig:overview_sub1}.
-The partition-based allocation approach used in Blue Gene series systems is an example of such a strategy \changed{\cite{bgloverview}}.
+The \emph{contiguous allocation} strategy assigns to each job a compact and contiguous set of computing nodes, as shown in Figure \ref{fig:overview_sub1}.
+The partition-based allocation approach used in Blue Gene series systems is an example of such a strategy \cite{bgloverview}.
 The contiguous strategy favors application performance through isolated networking within a partition and the locality that implies.
 However, this strategy can cause both internal fragmentation (when more nodes are allocated to a job than it requests) and external fragmentation (when sufficient nodes are available for a request, but they can not be allocated contiguously), therefore leading to lower system utilization than is otherwise possible.
-On the other side of the coin, the non-contiguous allocation strategy, used by the Cray XT/XE series \cite{carl-cug}, assigns free nodes to jobs regardless of contiguity, though of course efforts are made to maximize locality.
+On the other side of the coin, the \emph{non-contiguous allocation} strategy, used by the Cray XT/XE series \cite{carl-cug}, assigns free nodes to jobs regardless of contiguity, though of course efforts are made to maximize locality.
 Figure~\ref{fig:overview_sub2} shows an example non-contiguous allocation.
 While eliminating internal and external fragmentation as seen in contiguous allocation systems, in return non-contiguous allocations introduce contention between jobs due to the interleaving of job nodes.
-The non-contiguous node allocation can significantly reduce job performance, especially for communication-intensive ones \changed{\cite{abhinav-sc13}.}
+The non-contiguous node allocation can significantly reduce job performance, especially for communication-intensive ones \cite{abhinav-sc13}.
 
@@
 
-In this work, we focus on an in-depth analysis of intra- and inter-job communication interference with different job allocation strategies on torus-connected HPC systems, which are used on six of the top 10 supercomputers on the \changed{June 2015} Top500 lists~\cite{top500}.
+In this work, we focus on an in-depth analysis of intra- and inter-job communication interference with different job allocation strategies on torus-connected HPC systems.
+Torus-based networks are used on six of the top 10 supercomputers on the June 2015 Top500 lists~\cite{top500}.
 The current generation of IBM Blue Gene/Q (BG/Q) supercomputers, such as Mira at Argonne Nation Laboratory and Sequoia at Lawrence Livermore National Laboratory, has its nodes connected in a 5D torus network \cite{bgq}.
@@
 
-\NOTE{JJ removed the list of contributions, because they are exactly the same as the preceding paragraph.}
+%\NOTE{JJ removed the list of contributions, because they are exactly the same as the preceding paragraph.}
 \begin{comment}
@@
 Each application exhibits a distinctive communication pattern that is commonly seen in HPC applications.
-We believe that the communication patterns of these applications are representative of a wide array of applications run on leadership-class machines.
+We believe that the communication patterns of these applications are representative of a wide array of applications running on leadership-class machines.
 Specifically, we study the Algebraic MultiGrid Solver (AMG), Geometric MultiGrid (MultiGrid) and CrystalRouter MiniApps.
@@
 We can make the observation that AMG's dominant communication pattern is 3D Nearest Neighbor, each rank has intensive communication with up to six neighbors, depending on the boundaries of the ranks.
-Applications with similar pattern include PARTISN \changed{\cite{partisn}} and SNAP \changed{\cite{snap}}.
+Applications with similar pattern include PARTISN \cite{partisn} and SNAP \cite{snap}.
 
@@
     ~
-    \begin{subfigure}[t]{0.24\textwidth}
+    \begin{subfigure}[t]{0.22\textwidth}
         \centering
@@
 
-The second MiniApp covered is CrystalRouter, the extracted communication kernel of the full application Nek5000~\cite{crystalrouter}, which is a spectral element CFD application developed at Argonne National Laboratory \changed{\cite{nek5000}}.
+The second MiniApp covered is CrystalRouter, the extracted communication kernel of the full application Nek5000~\cite{crystalrouter}, which is a spectral element CFD application developed at Argonne National Laboratory \cite{nek5000}.
 It features spectral element multi-grid solvers coupled with a highly scalable, parallel coarse-grid solver that is widely used for projects including ocean current modeling, thermal hydraulics of reactor cores, and spatiotemporal chaos.
@@
 
-MultiGrid is geometric multi-grid v-cycle from the production elliptic solver BoxLib, a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes \changed{\cite{boxlib}}.
+MultiGrid is geometric multi-grid v-cycle from the production elliptic solver BoxLib, a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes \cite{boxlib}.
 MultiGrid conforms to many-to-many communication pattern with decreasing message size and collectives for different parts of the multi-grid v-cycle.
@@
 
-%\begin{figure*}[t!]
-%    \centering
-%    \begin{subfigure}[t]{0.32\textwidth}
-%        \centering
-%        \includegraphics[height=1.2in]{figs/appstudy/amg/amg_pip}
-%        \caption{AMG}
-%        \label{fig: amg communication topology}
-%    \end{subfigure}%
-%    ~ 
-%    \begin{subfigure}[t]{0.32\textwidth}
-%        \centering
-%        \includegraphics[height=1.2in]{figs/appstudy/cr/cr_pip}
-%        \caption{CrystalRouter}
-%        \label{fig: cr communication topology}
-%    \end{subfigure}
-%    ~
-%    \begin{subfigure}[t]{0.32\textwidth}
-%        \centering
-%        \includegraphics[height=1.2in]{figs/appstudy/mg/mg_pip}
-%        \caption{MultiGrid}
-%        \label{fig: mg communication topology}
-%    \end{subfigure}
-%    \caption{Caption place holder}
-%\end{figure*}
-%
-%\begin{figure*}[t!]
-%    \centering
-%    \begin{subfigure}[t]{0.32\textwidth}
-%        \centering
-%        \includegraphics[height=1.2in]{figs/appstudy/amg/amg_data_transfer}
-%        \caption{AMG DATA}
-%        \label{fig: amg data trans}
-%    \end{subfigure}
-%    ~
-%    \begin{subfigure}[t]{0.32\textwidth}
-%        \centering
-%        \includegraphics[height=1.2in]{figs/appstudy/cr/cr_data_transfer}
-%        \caption{CR DATA}
-%        \label{fig: cr data trans}
-%    \end{subfigure}
-%    ~
-%    \begin{subfigure}[t]{0.32\textwidth}
-%        \centering
-%        \includegraphics[height=1.2in]{figs/appstudy/mg/mg_data_transfer}
-%        \caption{MG DATA}
-%        \label{fig: mg data trans}
-%    \end{subfigure}
-%    \caption{Caption place holder}
-%\end{figure*}
-
-%We summarize the benchmarks and MiniApps on DOE design forward webpage\cite{designforwardwebpage}, as shown in Table \ref{tab: app summary}.
-According to their dominant communication patterns, there are three categories, namely Nearest Neighbor, All-to-All and 3D Lattice.
-The three applications studies in this work are representative for different categories.
-
-%\begin{table}[ht]
-%\begin{center}
-%\caption{ Summary of dominant communication pattern of MiniApps from DOE design forward project.} 
-%\label{tab: app summary}
-%\begin{tabular}{l c c c c} 
-%\toprule % Top horizontal line
-%\toprule
-%\multicolumn{2}{c}{Applications} \\
-%%\cmidrule(l){2-3}
-%\textbf{Dominant CP}  &  &  & &\\ % Column names row
-%\midrule % In-table horizontal line
-%\textbf{Nearest Neighbor}   &AMG    & SNAP  & PARTISN &\\ % Content row 1
-%
-%\midrule
-%\textbf{All-to-All}  &CrystalRouter    &BoxLib AMR    &MiniFE\\ % Content row 1
-%
-%\midrule
-%\textbf{3D Lattice}   &MultiGrid    & MiniDFT    &FillBoundary \\ % Content row 1
-%       
-%\midrule
-%\bottomrule % Bottom horizontal line
-%\end{tabular}
-%\end{center}
-%\end{table}
-
-%\textcolor{red}{Apps Summary} These three applications have representative communication patterns that are commonly seen in the HPC system workload.
-AMG represents a group of applications whose dominant communication pattern is Nearest Neighbor.
-MultiGrid represents applications whose dominant communication pattern is Many-to-Many (some literatures also refer it to All-to-All\cite{roth}).
-CrystalRouter represents those applications whose dominant communication is a hybrid of multi-stage local communication and hierarchical global communication.
-
 \section{Research Vehicle}
@@
 It is difficult to accurately and flexibly experiment with concurrently running jobs in an HPC context.
-One reason is that the allocation strategy used on production machine is part of the system software, which can not be changed by user.
+One reason is that the allocation strategy used on production machine is part of the system software, which can not be changed by users.
 Even the system administrator is not authorized to make that change.
-Another reason is that it is unrealistic to reserve the system exclusively to run the same job with desired allocation without interference then compare the results with those in the presence of interference.
+Another reason is that it is unrealistic to reserve the system exclusively to run the same job with desired allocation without interference and then compare the results with those in the presence of interference.
 Therefore, we resort to simulation for this work.
-
-%There are couple of simulation platform to study the behavior of large scale systems.
-SimGrid from Inria is a versatile simulation tool sets that provide simulation for Grids, Clouds and HPC systems \cite{simgrid}.
-The Structural Simulation Toolkit (SST) from Sandia National Lab was developed to explore innovations in highly concurrent systems where the ISA, microarchitecture, and memory interact with the programming model and communications system\cite{sst}.
 
@@
 CODES support both torus and dragonfly network with high fidelity flit-level simulation.
-CODES has this network workload component that capable of conducting trace-driven simulation.
-It can take real MPI application trace generated by SST DUMPI \cite{sst} to drive CODES network models.
+CODES has this network workload component that is capable of conducting trace-driven simulations.
+It can take real MPI application traces generated by SST DUMPI \cite{sst} to drive CODES network models.
 In this work, we focus on an in-depth analysis of intra- and inter-job communication interferences with  different job allocations on torus-connected HPC systems.
-Since torus networks have been extensively used in the current generation of supercomputers because of their linear scaling on per-node cost and competitive communication performance \cite{mubarak-sc2012}\cite{zhou-ipdps}.
-%We implemented a Job Mapping API for CODES, which can provide flexible job allocation strategies.
-The new API can help us to explore the impact of different allocation strategies to those three applications in Section \ref{sec:application study}
+Torus networks have been extensively used in the current generation of supercomputers because of their linear scaling on per-node cost and competitive communication performance \cite{mubarak-sc2012}\cite{zhou-ipdps}.
 
@@
     \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 5D torus network with 2Gb/s, 4Gb/s and 8Gb/s direct link bandwidth. }
+   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 5D torus network with 2GiB/s, 4GiB/s and GiB/s direct link bandwidth. }
    \label{fig:bandwidth-study}
@@
 \cmidrule(l){2-4}
-Dimension  & direct & aggregate &\\ % Column names row
+Dimension  & per-link & aggregate &\\ % Column names row
 \midrule % In-table horizontal line
-3D      & 4.67Gb/s  & 28Gb/s  & \\  % Content row 1
+3D      & 4.67GiB/s  & 28GiB/s  & \\  % Content row 1
 \midrule % In-table horizontal line
-5D    & 2.8Gb/s  & 28Gb/s  &\\ % Content row 1
+5D    & 2.8GiB/s  & 28GiB/s  &\\ % Content row 1
 \midrule % In-table horizontal line
-7D    & 2Gb/s & 28Gb/s & \\
+7D    & 2GiB/s & 28GiB/s & \\
 \midrule
@@
 At least for these applications, the benefits from increased dimensionality do not outweigh the reduction in per-link bandwidth.
-Multigrid does see benefits in the increased dimensionality, particularly with respect to outliers, but has a similar median message time.
+MultiGrid does see benefits in the increased dimensionality, particularly with respect to outliers, but has a similar median message time.
 CrystalRouter reduces performance as the dimensionality increases, but only by a small degree.
 AMG remains roughly the same, minus a dip in the maximum message time that could be due to rank-to-node mapping changes.
-
-\begin{comment}
-
-Then we use fixed dimensionality to study the impact of increased bandwidth.
-We run these three applications on 5D torus with direct link bandwidth increases from 2GiB/s, to 4GiB/s and 8GiB/s.
-As shown in Figure \ref{fig:bandwidth-study}, the data transfer time of each application can be greatly reduced as the bandwidth increases.
-
-We also study the impact of dimensionality with fixed aggregated bandwidth.
-We use 3D, 5D and 7D torus network with the same per-node aggregated bandwidth, which is 28Gb/s.
-With the same aggregated bandwidth, higher dimensionality means shorter diameter and pair-wise distance between nodes.
-The bandwidth between nodes in each torus network are shown in Table \ref{tab: fix-bandwidth}
-
-As we see from Figure \ref{fig:samebd-amg}, the quartiles of ``3D" box is a little bit higher (1\%) than that of ``5D", which means increasing dimensionality from 3D to 5D has little improvement on the average data transfer time for AMG.
-However, there is obvious reduction in those outliers between 3D and 5D.
-When dimensionality increases up to 7D, the improvement in terms of average data transfer time is still small, but more outliers above the top whisker can be spotted.
-
-The performance of CrystalRouter actually deteriorate as the torus dimensionality increasing, shown in Figure \ref{fig:samebd-cr}.
-This is due to the direct link bandwidth reduces from 4.67Gb/s to 2Gb/s, as dimension increases from 3D to 7D.
-CrystalRouter's large amount of intensive local data transfer benefit more from higher direct link bandwidth than its global transfer from lower diameter and shorter pair-wise distance between nodes.
-
-The performance of MultiGrid is gradually improved as the dimensionality of torus network increases, shown in Figure \ref{fig:samebd-mg}.
-The reduced diameter and shorten pair-wise distance between nodes would make the ``Many-to-Many" global data transfer in MultiGrid more efficient.
-
-We can get the observation that higher dimensionality of torus network would improve the performance of application with ``Many-to-Many" communication patterns, while application with intensive local communication like ``Nearest Neighbor" won't benefit much from higher dimensionality.
-
-\end{comment}
 
@@
 In this section we will study both kinds of interference in the context of torus network. 
-Since application's communication pattern won't change with its scale, we choose AMG with 216 ranks, CrystalRouter with 100 ranks, and MultiGrid 125 ranks.
-And in order to accommodate three applications running concurrently with different allocations, we simulate a 2K-node (16 $\times$ 16 $\times$ 8) 3D torus network for the experiments.
+Since application's communication pattern won't change with its scale, we use AMG with 216 ranks, CrystalRouter with 100 ranks, and MultiGrid 125 ranks.
+And in order to accommodate three applications running concurrently with different allocations, we simulate a 2K-node 3D torus ($16^{2} \times 8$) network for the experiments.
 
@@
 
-There are lots of research works \cite{jingjin}\cite{ozan} try to come up with better task mapping algorithms to alleviate the intra-job interference, which is out the scope of our work.
-In this work, we focus on the analysis of such interference with different allocations.
-To study the intra-job interference of each job, we simulate each job running on 3D torus network with different allocation shapes.
-We assign each application with three different shapes allocation, i.e.
-3D balanced-cube, 3D unbalanced-cube, 2D mesh.
+We design two sets of experiments to study the intra-job interference of each application.
+In the first, shown in Figure \ref{fig:shapestudy}, we assign each application with allocation in three different shapes.
+In the second, shown in Figure \ref{fig:diffmap}, we assign each application with 3D-balanced cube and use different mapping strategies for rank-to-node mapping.
+
+In the allocation shapes experiment, we select three possible shapes can can commonly seen on a 3D torus network. They are 3D balanced-cube, 3D unbalanced-cube, 2D mesh.
 
@@
 There are some research work like \cite{leung} \cite{abhinav-sc13} try to prove that compact allocation can guarantee job with better performance.
-They use many metrics to evaluate the compactness of the allocation, such as \emph{average pair-wise distance}, \emph{diameter} and \emph{contiguity}.
-In this work, we select 3D balanced-cube as the most compact allocation on a 3D torus network.
+In this experiment, 3D balanced-cube is the most compact allocation on a 3D torus network.
 
 3D unbalanced-cube, shown as green in Figure \ref{fig:cont_sub1}, is a rectangular prism, which is the possible allocation shape on system with asymmetric networks.
-For example, Blue Waters Cray XE6/XK7 system network is 3D torus with Gemini routers.
-The network connections in the \emph{y}direction have only half the bandwidth of the cables used in the \emph{x} and \emph{z} directions.
+The Blue Waters, a Cray XE6/XK7 system network is 3D torus with Gemini routers.
+The network connections in the \emph{y} direction have only half the bandwidth of the cables used in the \emph{x} and \emph{z} directions.
 In order to take advantage of the faster links in the \emph{x} and \emph{z} directions, job allocation is always start from X-Z plane, which leads to a rectangular prism shaped allocation \cite{RF}.
@@
     \end{subfigure}%
-   \caption{Intra-job Study: Data Transfer Time of 3 Apps on 3 different shapes allocation}
-   \label{fig: shapestudy}
+   \caption{Data Transfer Time of AMG, CrystalRouter and MultiGrid on 3D-balanced, 3D-unbalanced and 2D allocation.}
+   \label{fig:shapestudy}
 \end{figure*}
@@
     \end{subfigure}%
-   \caption{\TODO{This fig is UPdated} Intra-job study: data transfer time of 3 apps on 3D-balanced allocation with different mapping strategies.}
-   \label{fig: cont-time-box}
+   \caption{Data Transfer Time of AMG, CrystalRouter and MultiGrid on 3D-balanced allocation with different mapping strategies.}
+   \label{fig:diffmap}
 \end{figure*}

### 2015-10-17 submission.tex
almost there

@@
 \usepackage{amsmath}
+\usepackage{framed}
 \usepackage{textcomp}
@@
 (b) The amount of data sent/received by each rank in AMG.
-Y-axis labels the amount of data transferred in MB, x-axis is the rank.
+Y-axis labels the amount of data transferred in MB, x-axis labels 216 ranks of AMG.
 }
@@
 (b) The amount of data sent/received by each rank in CrystalRouter.
-Y-axis labels the amount of data transferred in MB, x-axis is the rank.
+Y-axis labels the amount of data transferred in MB, x-axis labels 100 ranks of CrystalRouter.
 }
@@
 (b) The amount of data sent/received by each rank in MultiGrid.
-Y-axis labels the amount of data transferred in MB, x-axis is the rank.
+Y-axis labels the amount of data transferred in MB, x-axis label 125 ranks of MultiGrid.
 }
@@
 We design two sets of experiments to study the intra-job interference of each application.
-In the first, shown in Figure \ref{fig:shapestudy}, we assign each application with allocation in three different shapes.
-In the second, shown in Figure \ref{fig:diffmap}, we assign each application with 3D-balanced cube and use different mapping strategies for rank-to-node mapping.
+In the first, shown in Figure \ref{fig:shapestudy}, we assign each application with allocations in \emph{three different shapes}.
+In the second, shown in Figure \ref{fig:diffmap}, we study the intra-job interference by using different mapping strategies for each application with given allocation shape.
 
-In the allocation shapes experiment, we select three possible shapes can can commonly seen on a 3D torus network. They are 3D balanced-cube, 3D unbalanced-cube, 2D mesh.
+In the allocation shapes experiment, we select three possible shapes can can commonly seen on the 3D torus network.
+They are 3D balanced-cube, 3D unbalanced-cube, 2D mesh, as shown in Figure \ref{fig:alloc-shapes}.
 
 3D balanced-cube, shown as red in Figure \ref{fig:cont_sub1}, can guarantee the minimum \emph{average pair-wise distance} within the allocation.
-There are some research work like \cite{leung} \cite{abhinav-sc13} try to prove that compact allocation can guarantee job with better performance.
-In this experiment, 3D balanced-cube is the most compact allocation on a 3D torus network.
+Some research studies like \cite{leung,abhinav-sc13} indicate that compact allocation can guarantee jobs with better performance.
+They use a variety of metrics to evaluate the compactness of the allocation, such as \emph{average pair-wise distance}, \emph{diameter} and \emph{contiguity}.
+In this work, we select 3D balanced-cube as the most compact allocation on a 3D torus network.
 
-3D unbalanced-cube, shown as green in Figure \ref{fig:cont_sub1}, is a rectangular prism, which is the possible allocation shape on system with asymmetric networks.
-The Blue Waters, a Cray XE6/XK7 system network is 3D torus with Gemini routers.
-The network connections in the \emph{y} direction have only half the bandwidth of the cables used in the \emph{x} and \emph{z} directions.
-In order to take advantage of the faster links in the \emph{x} and \emph{z} directions, job allocation is always start from X-Z plane, which leads to a rectangular prism shaped allocation \cite{RF}.
+3D unbalanced-cube, shown as green in Figure \ref{fig:cont_sub1}, is a rectangular prism, which is the possible allocation shape on the systems with asymmetric networks.
+For, example Cray XE6/XK7 systems are deployed 3D tori with Gemini routers.
+The network connections in the \emph{y}-direction have only half the bandwidth of the cables used in the \emph{x} and \emph{z} directions.
+In order to take advantage of the faster links in the \emph{x} and \emph{z} directions, job allocation starts from X-Z plane, which leads to a rectangular prism shaped allocation \cite{RF}.
 
@@
 2D mesh is a very common allocation shape in torus network for both contiguous and non-contiguous allocation strategies.
-For example, Cray Application Level Placement Scheduler indexes the nodes in torus network into a list and make allocation by simply works off the this list \cite{carl-cug}.
+For example, Cray Application Level Placement Scheduler (ALPS) indexes the nodes in torus network into a list and makes job allocation by simply going through the list \cite{carl-cug}.
 When the list is obtained by sorting the nodes based on their spacial coordinates in the torus, allocation made off from this list will be 2D mesh.
@@
 The nodes assigned to different jobs are interleaved, the size of allocation unit are 16, 8 and 2.}
+    \label{fig:alloc-shapes}
 \end{figure*}
@@
 And MultiGrid gets the best performance (shortest data transfer time) when running on 3D-balanced allocation, as shown in Figure \ref{fig:shapstudy-mg}.
-Since MultiGrid's communication pattern is ``Many-to-Many" dominant, 3D-balanced is the most compact allocation with shortest pair-wise distance between nodes, which can reduce the aggregated hops for transferring message between ranks in MultiGrid.
-CrystalRouter has with both dominant local communication and multi-stage global data transfer, the 3D-balance is also the best allocation, but its advantage over 3D-unbalanced is not that obvious as to MultiGrid, as shown in Figure \ref{fig:shapstudy-cr}.
+Since MultiGrid's communication pattern is ``Many-to-Many" dominant, 3D-balanced is the most compact allocation with shortest pair-wise distance between nodes, which can reduce the aggregated hops for transferring message among ranks in MultiGrid.
+As CrystalRouter has both dominant local communication and multi-stage global data transfer, the 3D-balance is also the best allocation, but its advantage over 3D-unbalanced is not that obvious as to MultiGrid, as shown in Figure \ref{fig:shapstudy-cr}.
 
-There are lots of research work try to design fancy allocation algorithms to provide application with most compact allocation.
-However, providing cubical allocation without considering application's communication pattern won't guarantee the best performance for every application.
+Lots of research studies have designed fancy allocation algorithms to provide application with the most compact allocation\cite{leung,LO}.
+However, as shown in our experiments, providing cubical allocation without considering application's communication pattern won't guarantee the best performance for every application.
 Compact allocation should be provided to application with global data transfer, such as those conforms to ``Many-to-Many" communication pattern.
@@
     \end{subfigure}%
-   \caption{Data Transfer Time of AMG, CrystalRouter and MultiGrid on 3D-balanced, 3D-unbalanced and 2D allocation.}
+   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 3D-balanced, 3D-unbalanced and 2D allocation.}
    \label{fig:shapestudy}
@@
 
-\TODO{This par is OUTDATED}.
-In non-contiguous allocation systems, the nodes assigned to each job could be scatter anywhere in the system.
-Some HPC systems predefine the minimum number of nodes consist of the allocation unit.
-The nodes within each unit are contiguous, and units assigned to each job may have arbitrary locations.
-We assign each application with a non-contiguous set of allocation units, refers to a small number(like 4) of adjacent nodes as shown in Figure \ref{fig:noncont_sub1}.
-We compare application's performance when it runs with contiguous allocation and non-contiguous allocation with unit size of 16, as shown in Figure \ref{fig: cont-time-box}.
+The rank-to-node mapping of parallel applications on HPC system could greatly impact applications' the performance.
+However, finding the optimal mapping solution for a given application is out the scope of this work.
+Our experiments aims to show how mapping strategies impact the intra-job interference of application with specific communication pattern.
 
-\TODO{This par is OUTDATED} Figure \ref{fig:cont-amg} shows that AMG is insensitive to the allocation's contiguity.
-Non-contiguous allocation can serve AMG equally well as contiguous allocation, as long as the allocation unit in non-contiguous allocation is big enough to preserve the locality of AMG.
-CrystalRouter shows preference in choosing between contiguous and non-contiguous allocation.
-As shown in Figure \ref{fig:cont-cr}, there is obvious degradation in terms data transfer time when CrystalRouter running with non-contiguous allocation.
-The global data transfer in CrystalRouter will take more hops on non-contiguous allocation, thus result in longer communication time.
-When it comes to MultiGrid shown in Figure \ref{fig:cont-mg}, this degradation is even worse.
-Since the global data transfer take a great portion of MultiGrid ``Many-to-Many" communication pattern.
+We provide AMG, CrystalRouter and MultiGrid with 3D-balance allocation and use three mapping strategies to do the rank-to-node mapping.
+The ``Linear" mapping, as we used in the previous experiments, is to map each rank according to the dimensional ordering of compute nodes.
+The ``Cube" mapping assigns ranks into consecutive $2^{3}$ cubes.
+The ``Random" mapping assigns ranks randomly within the allocation.
 
-\TODO{This par is OUTDATED} The performance of application conforms to Nearest Neighbor communication pattern keep relatively stable with different allocations.
-Non-contiguous allocation won't aggravate intra-job interference as long as the unit size can preserve the locality in application's communication topology graph.
+AMG remains roughly the same when it been mapped by ``Linear" and ``Cube", shown in Figure \ref{fig:diffmap-amg}.
+Although ``Linear" and ``Cube" mapping strategies cause some routing overlap for AMG's 3D ``nearest neighbor" communication, both mapping strategies still preserve the regularity of AMG's communication pattern.
+The ``Random" mapping totally destroys AMG's communication pattern, and results in expected intra-job interference among all the ranks in AMG.
+The performance degradation caused by using ``Random" mapping is up to 90\%.
+
+The ``Cube" mapping improves CrystalRouter's performance over the ``Linear" by reducing the pair-wise distance between ranks, shown in Figure \ref{fig:diffmap-cr}.
+The global data transfer in CrystalRouter will takes fewer hops with ``Cube" mapping.
+The ``Random" mapping scatters the ranks within the local neighborhood of CrystalRouter and make their communication less efficient.
+
+The ``Cube" mapping benefits MultiGrid's ``Many-to-Many" communication.
+However, due to the small amount of data transfer among ranks in MultiGrid, ``Cube" mapping doesn't exhibit a drastic advantage.
+Due to the same reason, ``Random" mapping neither cause too much degradation, shown in Figure \ref{fig:diffmap-mg}.
 
@@
     \end{subfigure}%
-   \caption{Data Transfer Time of AMG, CrystalRouter and MultiGrid on 3D-balanced allocation with different mapping strategies.}
+   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 3D-balanced allocation using  different mapping strategies.}
    \label{fig:diffmap}
@@
 
-%We simulate three application running concurrently with different allocation to analysis the interference between them.
-First, we assign each application with contiguous allocation side by side, and compare their performance with each application running exclusively.
-For AMG, sharing the network with other application won't affect its performance, since all the data transfer are between neighboring ranks.
-CrystalRouter and MultiGrid will suffer prolonged communication time when sharing network with others.
-The results are shown in Figure \ref{fig: concurrent-cont-time-box}.
-Inter-job interference has been identified as one of the major culprits for application's performance variability \cite{abhinav-sc13}\cite{skinner}\cite{rosenthal}.
+Inter-job interference has been identified as one of the major culprits for application's performance variability \cite{abhinav-sc13,skinner,rosenthal}.
 Inter-job interference is a more prominent issue for system adopted non-contiguous allocation policy than system with contiguous allocation.
@@
 We assign each application with non-contiguous allocation and run them concurrently on the same network.
-The allocation unit belongs to different jobs are interleaved as shown in Figure \ref{fig:noncont_sub1}.
-The unit size is critical to the application's performance.
-Figure \ref{fig: 3apps interjob study} shows the results of each application data transfer time with different allocation unit size.
+The allocation unit belongs to different jobs are interleaved.
+In order to study the impact of different allocation unit size on applications' inter-job interference, we conduct experiments with unit size of 16, 8 and 2, shown respectively in Figure \ref{fig:noncont_sub1}, \ref{fig:noncont_sub1} and \ref{fig:noncont_sub3}.
+Figure \ref{fig:interjobstudy} shows the results of each application data transfer time with different allocation unit size.
 
-The data transfer time of AMG in Figure \ref{fig:interjob-amg-box} keeps stable between allocation unit size of 16 (big) and 8 (medium), since the locality of AMG is 6-rank based.
-When the unit size reduce to 2 (small), AMG suffers prolong data transfer time by about 10\%.
+The data transfer time of AMG in Figure \ref{fig:interjob-amg} keeps stable between allocation unit size of 16 and 8, since the locality of AMG is 6-rank based.
+When the unit size reduce to 2, AMG suffers prolong data transfer time by about 10\%.
 CrystalRouter is more sensitive to allocation unit size.
 The best unit size would be the one big enough to accommodate its local neighborhood.
-Figure \ref{fig:interjob-cr-box} shows unit size of 16 and 8 can guarantee the same average data transfer time, while some rank spend more time with allocation unit size 8 than 16.
-When the unit size reduce to 2-node, the communication become less efficiency and takes about 15\% more times for transferring data.
+Figure \ref{fig:interjob-cr} shows unit size of 16 and 8 can guarantee the same average data transfer time, while some rank spend more time with allocation unit size 8 than 16.
+When the unit size reduce to 2, the communication become less efficiency and takes about 15\% more times for transferring data.
 
-The data transfer time of MultiGrid with different allocation unit sizes doesn't show obvious variability in Figure \ref{fig:interjob-mg-box}.
-This is because even big allocation unit size like 16-node will still fail to preserve MultiGrid's ``Many-to-Many" pattern.
-The data transfer time is almost doubled when MultiGrid running concurrently with allocation unit size of 16, as shown in Figure \ref{fig:interjob-mg-box} .
+The data transfer time of MultiGrid with different allocation unit sizes doesn't show obvious variability in Figure \ref{fig:interjob-mg}.
+This is because even big allocation unit size like 16 will still fail to preserve MultiGrid's ``Many-to-Many" pattern.
+The data transfer time is almost doubled when MultiGrid running concurrently with allocation unit size of 16, as shown in Figure \ref{fig:interjob-mg} .
 As the unit size decreases, the data transfer time keep growing, but in a slow way, which means in terms of preserving the locality of MultiGrid, allocation unit of size 16, 8 and 2 serve equally bad.
@@
         \caption{AMG}
-        \label{fig:interjob-amg-box}
+        \label{fig:interjob-amg}
     \end{subfigure}%
@@
         \caption{CrystalRouter}
-        \label{fig:interjob-cr-box}
+        \label{fig:interjob-cr}
     \end{subfigure}%
@@
         \caption{MultiGrid}
-        \label{fig:interjob-mg-box}
+        \label{fig:interjob-mg}
     \end{subfigure}%
@@
 }
-   \label{fig: 3apps interjob study}
+   \label{fig:interjobstudy}
 \end{figure*}
 
-\subsection{Summary}
+\subsection{Result Summary}
 \label{sec:summary}
 
-Based on our comprehensive study, we can get following observations:
+Based on the above comprehensive study, we can make following observations:
 \begin{itemize}
     \item The compact allocation may not guarantee the best performance for every application. 
-    \item The performance of application conforms to Nearest Neighbor communication pattern keeps relatively stable with different shape allocations, while application conforms to ``Many-to-Many" communication pattern prefers compact allocation.
-    \item The allocation unit size should be decided according to application's communication pattern.
-Unit size that big enough to preserve the neighborhood communication of the application will result in better performance.
+    \item The applications dominated with Nearest Neighbor communications exhibit relatively stable performance under different allocation shapes, while the applications dominated with Many-to-Many communication exhibit better performance with compact allocation (e.g.
+3D balanced).
+    \item A good rank-to-node mapping strategy can greatly improve application's performance when specific allocation is given.
+    \item An optimal size for allocation units should be determined according to an application's dominant communication pattern.
+In general, a unit size should be large enough to preserve neighboring communication of the application.
+    \item Inter-job interference is inevitable in non-contiguous allocation.
+However, choosing the proper allocation unit size with job's communication pattern awareness can alleviate such negative effect.
 \end{itemize}
@@
 
-%HPC system usually accommodate many parallel jobs running concurrently.
-Each job gets the required number of computing nodes by system resource manager.
-The allocated nodes for each job communicate with each other through system interconnect network in a sharing way, which cause interference as a negative side-effect.
-This interference can introduce performance degradation to both job and system, such as prolonged job running time and reduced system throughput.
-
-The observations we obtained by analyzing the intra- and inter-job interference can provide insights for the design of a smart and flexible job allocation strategy.
+The results shown in the previous section provide insights for the design of smart and flexible job allocation.
 By scrutinizing job's communication behavior, we can identify job's dominant communication pattern and pinpoint the ``neighborhood communication".
@@
 
-When an application with dominant communication that is intensive ``Many-to-Many" being submitted to the system, it should be granted with compact node allocation and exclusive network provision.
+When an application with dominant communication that is intensive ``Many-to-Many", the scheduler should grant it with compact node allocation and exclusive network provision.
 The compact allocation can guarantee the shortest pair-wise distance between all the ranks, and make the data transfer between ranks take less hops.
@@
 Not every application's preferable resource are compact node allocation and exclusive network provision.
-As we found in our comprehensive analysis, applications whose dominant communication patterns contain intensive ``neighborhood communication" like ``Nearest Neighbor" won't benefit from compact node allocation and exclusive network provision.
-Such applications can run with non-contiguous node allocation without performance degradation, as long as the allocation unit can accommodate the ``neighborhood communication".
+As we found in our analysis, applications whose dominant communication patterns contain intensive ``neighborhood communication" like ``Nearest Neighbor" don't benefit from compact node allocation and exclusive network provision.
+Applications, such as AMG, can run with non-contiguous node allocation without performance degradation, as long as the allocation unit can accommodate their ``neighborhood communication".
 
 The allocation unit size shouldn't be fixed.
-Instead, the new allocation strategy should choose the proper unit size based on the scale of ``neighborhood communication" of each job.
+Instead, the scheduler should choose a proper unit size based on the scale of ``neighborhood communication" of each job.
 The best unit size should be neither too big nor too small, jut perfectly fit for the size of that ``neighborhood".
-Big unit size won't be fully utilized and results in fragmentation.
+Big unit size won't be fully utilized and also cause fragmentation.
 Small unit size won't be able to accommodate the ``neighborhood communication" and makes the intra-job communication less efficient.
@@
 The advantage of making allocation with consideration about job's communication pattern is obvious.
-First, compared with contiguous allocation policy, the new allocation strategy is more flexible, it doesn't require the system to provide a big contiguous partition that can accommodate the whole application, just small set of compact nodes enough for all the "local community" in the application.
-On the other hand, the new allocation strategy won't destroy the locality of application's communication pattern.
-On the contrary, the small compact node set (allocation unit) provided for the application's  ``local communities"  will preserve the communication locality in the maximum extent.
-
-%The off-line study of job's communication traces could be helpful to develop better allocation policies for batch scheduler to make on-line scheduling/allocating decisions.
-This is due to the set of applications running on those leadership computing facilities is usually stable, and the workloads from those HPC systems is usually has great repetitiveness.
-We analyzed the workload on Mira, an IBM Blue/Gene Q system at Argonne Leadership Computing Facility, found that a large portion of its workload are repetitive.
-Figure \ref{fig: repetitiveness of Mira} shows the 5 of most active users with repetitive submissions.
-The collected jobs information from history can be very useful for workload prediction and system optimization.
-%
-%
-%\begin{figure}[h!] 
-%  \centering
-%  \includegraphics[width=0.38\textwidth]{figs/repe-mira}
-%   \caption{The repetitiveness of Mira's workload. }
-%   \label{fig: repetitiveness of Mira}
-%\end{figure}
+First, compared with contiguous allocation policy, the new scheduler outlined above is more flexible, it doesn't require the system to provide a big contiguous partition to accommodate the whole application, instead it only needs to provide a small set of compact nodes that is sufficient for all the "local community" in the application.
+On the other hand, the new scheduler outlined above won't destroy the locality of application's communication pattern.
+Indeed, the small compact node set (allocation unit) provided for the application's  ``local communities" can preserve the communication locality in the maximum extent.
+The design of such a smart scheduler is part of our future work.
 
@@
 \label{sec:related_work}
-%\textbf{P1: tools for monitoring and profiling}\\
-There are some tools available for system monitoring and application profiling on HPC systems.
+
+There are many tools available for system monitoring and application profiling.
 Tools like TAU (Tuning and Analysis) \cite{tau}, SCALASCA \cite{scala} and mpiP \cite{mpip} can capture application's runtime information about application's communication and computation in event traces.
-However, recognizing the patterns about application's communication from those traces requires lots of extra effort.
-Some researchers work on the recognition and characterization of parallel application communication patterns.
-Oak Ridge National Laboratory has this ongoing project about developing a tool set named Oxbow, which can characterize the computation and communication behavior of 12 scientific applications and benchmarks \cite{oxbow}.
-In their recent work \cite{roth}, they demonstrate a new approach to automatically characterizing a parallel application's communication behavior.
+However, recognizing the patterns about application's communication from those traces requires substantial effort.
+There are a number of studies on the recognition and characterization of parallel application communication patterns.
+Oak Ridge National Laboratory has its ongoing project about developing a tool set named Oxbow, which can characterize the computation and communication behavior of scientific applications and benchmarks \cite{oxbow}.
+In a recent work \cite{roth}, the authors demonstrate a new approach to automatically characterizing parallel applications' communication behaviors.
 
 %\textbf{P2: communication characteristics}\\
-Some research institutes put great effort in characterization of scientific applications.
-The Department of Energy initiate this design forward project aims at the identification of the computational characteristics of the DOE MiniApps developed at various exascale co-design centers \cite{designforwardwebpage}.
+
+Many research efforts have been conducted to characterize scientific applications.
+For instance, a DOE project aims at the identification of the computational characteristics of the DOE MiniApps developed at various exascale co-design centers \cite{designforwardwebpage}.
 In this project, the communication patterns of several DOE full applications and associated mini-applications are studied to provide a more complete snapshot of the DOE workload.
@@
 found that concurrently running applications on HPC can cause interference to each other, and cause communication time varied from 36\% faster to 69\% slower.
-Such interference could come from Operating System noise, shape of the allocated partition and mainly other running jobs that sharing network resources \cite{abhinav-sc13}.
+Such interference could come from Operating System noise, shape of the allocated partition , or other running jobs that sharing network resources \cite{abhinav-sc13}.
 Skinner et al.
-found that  there is a 2-3$\texttimes$ slowdown in MPI$\textunderscore$Allreduce due to network contention from other jobs \cite{skinner}.
+found that  there is a 2-3$\times$ slowdown in MPI$\textunderscore$Allreduce due to network contention from other jobs \cite{skinner}.
 Rosenthal et al.
-found there is limited benefit for many applications by increasing network bandwidth.
-The applications that send mostly small messages or larger messages asynchronously are not bandwidth bounded, hence, benefit only slightly or not at all from increased bandwidth \cite{rosenthal}.
+found increasing network bandwidth provide limited benefits to a number of applications.
+The applications that send mostly small messages or large messages asynchronously are not bandwidth bounded, hence, benefit only slightly or not at all from increased bandwidth \cite{rosenthal}.
 
 %\textbf{P4: Job allocation(little bit)}\\
-There are some research work focus on optimizing job allocation on HPC system to alleviate the interference between concurrently running jobs.
+Several research studies focus on optimizing job allocation on HPC system to alleviate the interference between concurrently running jobs.
 Hoefler et al.
-propose to use performance modeling techniques to analysis factors that impact the performance of parallel scientific applications \cite{hoefler-modeling}.
-However, as the scale of HPC continue grows, the  interference of concurrently running jobs is getting worse, which is hard to be quantified by performance profiling tools.
-Bogdan et al provide a set of guidelines how to configure a network with Dragonfly topology for workload with Nearest Neighbor communication pattern \cite{Bogdan-hpdc14}.
-Dong et al describe IBM Blue Gene/Q's 5D torus interconnect network \cite{Dong-SC11}.
-They developed simple benchmarks that conforms to four different communication patterns, namely ping-pong, nearest neighbor, broadcast and all$\textunderscore$reduce, to demonstrate the effectiveness of this highly parallel 5D torus network.
+proposed to use performance modeling techniques to analyze factors that impact the performance of parallel scientific applications \cite{hoefler-modeling}.
+However, as the scale of HPC continue to grow, the  interference of concurrently running jobs is getting worse, which is hard to be quantified by performance profiling tools alone.
+Bogdan et al.
+provide a set of guidelines of how to configure a network with Dragonfly topology for workload with Nearest Neighbor communication pattern \cite{Bogdan-hpdc14}.
+Dong et al.
+have developed simple benchmarks that conforms to four different communication patterns, namely ping-pong, nearest neighbor, broadcast and all$\textunderscore$reduce, to demonstrate the effectiveness of this highly parallel 5D torus network \cite{Dong-SC11}.
 
@@
 First, we focus on the dominant communication patterns rather than any specific application.
-We believe this can provide a guideline for other research work is the area.
-Secondly, we explored the inter-job interference between concurrently running jobs, while similar work such as \cite{abhinav-sc13} only focus on the single application's performance degradation due to network contention.
-Finally, we explored the impact of different allocation strategies to job's communication behavior.
-We identified the optimal allocation strategy for each application with specific dominant communication pattern.
-Based on our comprehensive exploration, we claim that better allocation strategy should take job's communication pattern into consideration for allocation decision making.
+We believe this can provide a guideline for other research work.
+Secondly, we explore the intra- and inter-job interference between concurrently running jobs, while similar work such as \cite{abhinav-sc13} only focuses on the single application's performance degradation due to network contention.
+Finally, we analyze the impact of different allocation strategies to job's communication behaviors.
+We further identify the optimal allocation strategy for each application with a specific dominant communication pattern.
+Based on our study, we claim that further scheduling should take job's communication pattern into consideration for its allocation decision making.
 
@@
 \label{sec:conclusion}
-In this work, we study the communication behavior of three parallel applications, namely AMG, CrystalRouter and MultiGrid.
+In this work, we have studied the communication behavior of three parallel applications, namely AMG, CrystalRouter and MultiGrid.
 Each application has distinctive communication pattern, that can be representative for a group of jobs in HPC workload.
-We use a sophisticate simulation tool named CODES from Argonne National Laboratory to simulate the running of these three parallel applications on torus network.
+We have used a sophisticate simulation tool named CODES from Argonne National Laboratory to simulate the running of these three parallel applications on torus network.
 The torus network provided by CODES has good fidelity and scalability.
-We analyzed the performance of each application's communication in terms of data transfer time by simulating them running on torus networks with different bandwidth and dimensionality configurations.
-We found that higher dimensionality of torus network would improve the performance of application with ``Many-to-Many" communication patterns, while application with intensive local communication like ``Nearest Neighbor" won't benefit much from higher dimensionality.
+We have analyzed the performance of each application's communication in terms of data transfer time by simulating them running on torus networks with different bandwidth and dimensionality configurations.
+We have found that higher dimensionality of torus network would improve the performance of application with ``Many-to-Many" communication patterns, while application with intensive local communication like ``Nearest Neighbor" won't benefit much from higher dimensionality.
 
-We also analysis the intra- and inter-job interference by simulating three applications running on 3D torus network.
-Based on our comprehensive experiments, we got three observations.
-1) The compact allocation can not guarantee the best performance for every application.
-2) The performance of application conforms to Nearest Neighbor communication pattern keep relatively stable with different shape allocations, while application conforms to ``Many-to-Many" communication pattern prefer compact allocation.
-3) The allocation unit size should be decided according to application's communication pattern.
-Unit size that big enough to preserve the locality of the application will result better performance.
+We have analyzed the intra- and inter-job interference by simulating three applications running on 3D torus network.
+Based on our comprehensive experiments, we got five observations.
+1) The compact allocation may not guarantee the best performance for every application.
+2)The applications dominated with Nearest Neighbor communications exhibit relatively stable performance under different allocation shapes, while the applications dominated with Many-to-Many communication exhibit better performance with compact allocation (e.g.
+3D balanced).
+3) A good rank-to-node mapping strategy can greatly improve application's performance when specific allocation is given.
+4) An optimal size for allocation units should be determined according to an application's dominant communication pattern.
+In general, a unit size should be large enough to preserve neighboring communication of the application.
+5) Inter-job interference is inevitable in non-contiguous allocation.
+However, choosing the proper allocation unit size with job's communication pattern awareness can alleviate such negative effect.
 
-We believe that our finding in this work can provide guidance for HPC resource management to make flexible job allocations.
-Rather than use pre-defined partitions and reckless non-contiguous allocation, future HPC system should assign each job with preferable resources based on job's communication pattern.
+We believe that our finding in this work can provide valuable guidance for HPC resource management to make flexible job allocations.
+Rather than using pre-defined partitions and non-contiguous allocation, future HPC systems should assign each job with preferable resources based on job's communication pattern.
 
@@
 \vspace{5\baselineskip}
+
+\begin{framed}
 The submitted manuscript has been created by UChicago Argonne, LLC, Operator of Argonne National Laboratory ("Argonne").
@@
 Government retains for itself, and others acting on its behalf, a paid-up nonexclusive, irrevocable worldwide license in said article to reproduce, prepare derivative works, distribute copies to the public, and perform publicly and display publicly, by or on behalf of the Government.
+\end{framed}
 

### 2015-10-17 submission.tex
minor teaks, almost there

@@
 In this work, we choose three representative HPC applications from DOE Design Forward Project and conduct detailed simulations of high-performance network topologies to analyze both intra- and inter-job interference.
-By scrutinizing the communication behaviors of these applications, we identify relationships between these behaviors and the possible interference induced by different job allocation scenarios.
+By scrutinizing the communication behaviors of these applications, we identify relationships between these behaviors and the possible interference introduced by different job allocation scenarios.
 Our comprehensive analyses in this work illuminate a path towards job-communication-pattern-aware HPC resource allocation.
@@
 When submitting jobs to HPC systems, users request resources in terms of node counts and expected runtime.
-The batch scheduler is then responsible for dispatching them to the system.
+The batch scheduler is then responsible for dispatching the submitted jobs to the system.
 Typically, multiple jobs run concurrently, resulting in the shared use of resources, particularly network and storage.
@@
 
-Communication variability due to network sharing can cause application performance degradation.
 \textbf{\emph{Intra-job interference}} refers to the network contention between ranks within each application.
 \textbf{\emph{Inter-job interference}} is introduced by concurrently running jobs that adjacent to each other sharing network resources.
+Communication variability due to such interference can cause application performance degradation.
 

### 2015-10-17 submission.tex
minor chages, ready for submission i wish

@@
 3D unbalanced-cube, shown as green in Figure \ref{fig:cont_sub1}, is a rectangular prism, which is the possible allocation shape on the systems with asymmetric networks.
-For, example Cray XE6/XK7 systems are deployed 3D tori with Gemini routers.
+For example, Cray XE6/XK7 systems are deployed 3D tori with Gemini routers.
 The network connections in the \emph{y}-direction have only half the bandwidth of the cables used in the \emph{x} and \emph{z} directions.
@@
 
-Figure \ref{fig:shapstudy-amg} shows that for AMG, who conforms to 3D ``Nearest Neighbor" communication pattern, spent less time when being assigned with 3D-unbalanced allocation than 3D-balanced.
+AMG with 3D ``Nearest Neighbor" communication pattern takes less time when being assigned with 3D-unbalanced allocation than 3D-balanced, as shown in Figure \ref{fig:shapstudy-amg}.
 And MultiGrid gets the best performance (shortest data transfer time) when running on 3D-balanced allocation, as shown in Figure \ref{fig:shapstudy-mg}.
@@
 
-The rank-to-node mapping of parallel applications on HPC system could greatly impact applications' the performance.
+The rank-to-node mapping of parallel applications on HPC system could greatly impact applications' performance.
 However, finding the optimal mapping solution for a given application is out the scope of this work.
@@
 AMG remains roughly the same when it been mapped by ``Linear" and ``Cube", shown in Figure \ref{fig:diffmap-amg}.
-Although ``Linear" and ``Cube" mapping strategies cause some routing overlap for AMG's 3D ``nearest neighbor" communication, both mapping strategies still preserve the regularity of AMG's communication pattern.
-The ``Random" mapping totally destroys AMG's communication pattern, and results in expected intra-job interference among all the ranks in AMG.
+Although ``Linear" and ``Cube" mapping strategies cause some routing overlap for AMG's communication, both mapping strategies still preserve the locality of AMG's 3D ``Nearest Neighbor" communication pattern.
+The ``Random" mapping totally destroys AMG's communication pattern, and results in intra-job interference among all the ranks in AMG.
 The performance degradation caused by using ``Random" mapping is up to 90\%.
@@
 However, due to the small amount of data transfer among ranks in MultiGrid, ``Cube" mapping doesn't exhibit a drastic advantage.
-Due to the same reason, ``Random" mapping neither cause too much degradation, shown in Figure \ref{fig:diffmap-mg}.
+``Random" mapping neither cause too much degradation for the same reason, shown in Figure \ref{fig:diffmap-mg}.
 

### 2016-07-07 app_study.tex
split into separate tex files for each section

--- 2015-10-17 submission.tex
@@
-\documentclass[conference]{IEEEtran}
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
-\else
-
-\fi
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-\usepackage{algorithm}
-%\usepackage{algorithm2e}
-\usepackage{algpseudocode}
-\usepackage{booktabs}
-\usepackage{flushend}
-\usepackage{url}
-\usepackage{xcolor}
-\usepackage{amsmath}
-\usepackage{framed}
-\usepackage{textcomp}
-%\usepackage{subfig}
-\usepackage{graphicx}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{soul}
-\usepackage{verbatim}
-%\usepackage[labelfont=bf]{caption}
-\usepackage[font=bf,labelsep=space]{caption}
-
-\newcommand{\TODO}[1]{\hl{TODO: #1}}
-\newcommand{\NOTE}[1]{\hl{NOTE: #1}}
-\newcommand{\changed}[1]{\textcolor{red}{#1}}
-
-\begin{document}
-%
-% paper title
-% can use linebreaks \\ within to get better formatting as desired
-\title{Study of Intra- and Inter-Job Interference with Different Job Allocations}
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
-\author{
-
-\IEEEauthorblockN{Xu Yang\IEEEauthorrefmark{1}, John Jenkins\IEEEauthorrefmark{2}, Misbah Mubarak\IEEEauthorrefmark{2}, Robert B.
-Ross\IEEEauthorrefmark{2}, Zhou Zhou\IEEEauthorrefmark{1}, Zhiling Lan\IEEEauthorrefmark{1}}
-
-\IEEEauthorblockA{\IEEEauthorrefmark{1}Department of Computer Science,
-Illinois Institute of Technology,
-Chicago, Illinois, USA 60616\\
-\{xyang56, zzhou\}@hawk.iit.edu, lan@iit.edu}
-
-\IEEEauthorblockA{\IEEEauthorrefmark{2}Mathematics and Computer Science Division, Argonne National Laboratory,
-Argonne, IL, USA 60439\\
-\{jenkins,rross\}@mcs.anl.gov, mmubarak@anl.gov}
-}
-
-% make the title area
-\maketitle
-
-\begin{abstract}
-
-Network contention between concurrently running jobs on HPC systems is a primary cause of performance variability.
-Optimizing job allocation and avoiding network sharing are hence crucial to alleviate the potential degradation in performance.
-In order to do these effectively, an understanding of the interaction between applications, their communication patterns, and the system architecture are required.
-In this work, we choose three representative HPC applications from DOE Design Forward Project and conduct detailed simulations of high-performance network topologies to analyze both intra- and inter-job interference.
-By scrutinizing the communication behaviors of these applications, we identify relationships between these behaviors and the possible interference introduced by different job allocation scenarios.
-Our comprehensive analyses in this work illuminate a path towards job-communication-pattern-aware HPC resource allocation.
-
-\end{abstract}
-
-\IEEEpeerreviewmaketitle
-
-\section{Introduction} 
-\label{sec: intro}
-
-The scale of supercomputers continually grows at a breakneck pace to accommodate the computing power requirements for scientific research.
-Supercomputers have tens of thousands of nodes and serve as an irreplaceable research vehicle for scientific problems with increasing size and complexity.
-Supercomputers are usually employed as a shared resource to accommodate many parallel applications (jobs) running concurrently.
-These parallel jobs share the system infrastructure such as network and I/O bandwidth, and inevitably there is contention over these shared resources.
-As supercomputers continue to evolve, these shared resources are increasingly the bottleneck for performance.
-
-\begin{figure}[h!]
-    \centering
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/goodallocation}
-        \caption{Contiguous}
-        \label{fig:overview_sub1}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/badallocation}
-        \caption{Non-contiguous}
-        \label{fig:overview_sub2}
-    \end{subfigure}%
-   \caption{Multiple jobs running concurrently with different allocations.
-Each job is represented by a specific color.
-a) shows the effect of contiguous allocation, which reduce the inter-job interference.
-b) shows non-contiguous allocation, which may introduce both intra and inter-job interferences.
-}
-   \label{fig:overview}
-\end{figure}
-
-When submitting jobs to HPC systems, users request resources in terms of node counts and expected runtime.
-The batch scheduler is then responsible for dispatching the submitted jobs to the system.
-Typically, multiple jobs run concurrently, resulting in the shared use of resources, particularly network and storage.
-A prominent problem with this approach is network contention among concurrently running jobs.
-If the network resources of concurrently running jobs aren't isolated, the resulting sharing can cause communication variability that results in performance degradation for some or all such jobs \cite{abhinav-sc13}.
-This delay can propagate into the queueing time of the following submitted jobs, thus leading to low system throughput and utilization \cite{jose-ipdps15}.
-This adverse effect of network sharing can be mitigated by providing jobs with isolated allocations and exclusive network resources.
-
-On the widely used torus-connected HPC systems \cite{bgq,tofu,titan}, two allocation strategies are commonly used.
-The \emph{contiguous allocation} strategy assigns to each job a compact and contiguous set of computing nodes, as shown in Figure \ref{fig:overview_sub1}.
-The partition-based allocation approach used in Blue Gene series systems is an example of such a strategy \cite{bgloverview}.
-The contiguous strategy favors application performance through isolated networking within a partition and the locality that implies.
-However, this strategy can cause both internal fragmentation (when more nodes are allocated to a job than it requests) and external fragmentation (when sufficient nodes are available for a request, but they can not be allocated contiguously), therefore leading to lower system utilization than is otherwise possible.
-On the other side of the coin, the \emph{non-contiguous allocation} strategy, used by the Cray XT/XE series \cite{carl-cug}, assigns free nodes to jobs regardless of contiguity, though of course efforts are made to maximize locality.
-Figure~\ref{fig:overview_sub2} shows an example non-contiguous allocation.
-While eliminating internal and external fragmentation as seen in contiguous allocation systems, in return non-contiguous allocations introduce contention between jobs due to the interleaving of job nodes.
-The non-contiguous node allocation can significantly reduce job performance, especially for communication-intensive ones \cite{abhinav-sc13}.
-
-We envision that future HPC schedulers will adopt a flexible job allocation mechanism which combines the best of both contiguous and non-contiguous allocation strategies.
-Such a flexible mechanism would take shared resource needs of jobs into account when making allocation decisions (e.g., network).
-With knowledge and analysis of job communication patterns, it can be identified which jobs require network isolation and locality, and to what degree.
-Then, rather than allocating each job in a ``know-nothing'' manner, one may specialize allocation so that, for example, only the jobs with stringent network needs are given compact, isolated allocations, resulting in maximized utilization and minimized perceivable resource contention effects.
-
-In this work, we focus on an in-depth analysis of intra- and inter-job communication interference with different job allocation strategies on torus-connected HPC systems.
-Torus-based networks are used on six of the top 10 supercomputers on the June 2015 Top500 lists~\cite{top500}.
-The current generation of IBM Blue Gene/Q (BG/Q) supercomputers, such as Mira at Argonne Nation Laboratory and Sequoia at Lawrence Livermore National Laboratory, has its nodes connected in a 5D torus network \cite{bgq}.
-The K computer from Japan uses the ``Tofu'' system, which is a 6D mesh/torus topology \cite{tofu}.
-Titan, a Cray XK7 supercomputer located at the Oak Ridge Leadership Computing Facility (OLCF), has nodes connected in a 3D torus within the compute partition \cite{titan}.
-Although our analyses are based on torus networks, the ideas conveyed in this work is applicable to networks with different topologies.
-
-We selected three signature applications from the DOE Design Forward Project~\cite{designforwardwebpage} as examples to conduct detailed study about their communication patterns.
-We use a sophisticated simulation toolkit named CODES (standing for Co-Design of Multi-layer Exascale Storage Architectures)~\cite{Jason-2011} as a research vehicle to evaluate the performance of these applications with various allocations in a controlled environment.
-We then analyze the intra- and inter-job interference by simulating these applications running exclusively and concurrently with different allocations.
-We believe the insights presented in this work can be very useful for the design of future HPC batch job schedulers and resource managers.
-
-%\NOTE{JJ removed the list of contributions, because they are exactly the same as the preceding paragraph.}
-\begin{comment}
-The observations of this work are the following:
-
-\begin{itemize}
-    \item We study the communication behaviors and identify the communication patterns of three representative HPC applications.
-    \item We conduct detailed analysis about the intra- and inter-job interference by running three applications with different allocation strategies.
-And we find the relation between application's communication pattern and the cause of interference.
-    \item We show a novel and flexible way to make allocation with awareness of job's communication pattern, which  is every effective to alleviate the interference between concurrently running jobs.
-\TODO{do we actually do this, or provide the underpinnings for such a method?}
-\end{itemize}
-
-We believe the finds and insights presented in this work would be very useful for the design of future HPC batch job scheduler and resource management module.
-
-\end{comment}
-
-The rest of this paper is organized as follows.
-Section~\ref{sec:application study} describes the three representative applications from DOE Design Forward Project for our study.
-Section~\ref{sec:codes} talks about the use of CODES as research vehicle for our work.
-Section~\ref{sec:config study} shows the performance analysis of three applications on different torus networks.
-Section~\ref{sec:interference} provides detailed analysis about the intra- and inter-job interference between three applications on torus network with different allocation strategies.
-Section~\ref{sec:discussion} introduces a path toward communication-pattern aware allocation strategies, given the results of our analysis.
-Section~\ref{sec:related_work} discusses related work.
-Finally, the conclusion is presented in Section~\ref{sec:conclusion}.
-
 \section{Application Study}
@@
 
+\begin{figure*}[htp]
+    \centering
+    \begin{subfigure}[t]{0.32\textwidth}
+        \centering
+        \includegraphics[height=1.5in]{figs/appstudy/amg/amg_pip}
+        \caption{AMG}
+        \label{fig:amg-communication-topology}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.32\textwidth}
+        \centering
+        \includegraphics[height=1.5in]{figs/appstudy/cr/cr_pip}
+        \caption{CrystalRouter}
+        \label{fig:cr-communication-topology}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.32\textwidth}
+        \centering
+        \includegraphics[height=1.5in]{figs/appstudy/mg/mg_pip}
+        \caption{MultiGrid}
+        \label{fig:mg-communication-topology}
+    \end{subfigure}
+    \caption{Applications rank-to-rank communication topology graph.}
+\end{figure*}
+
 \subsection{AMG}
@@
 
-\begin{figure}[t!]
-    \centering
-    \begin{subfigure}[t]{0.25\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/appstudy/amg/amg_pip}
-        \caption{}
-        \label{fig:amg-communication-topology}
-    \end{subfigure}%
-    ~
-    \begin{subfigure}[t]{0.24\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/appstudy/amg/amg_data_transfer}
-        \caption{}
-        \label{fig:amg-data-trans}
-    \end{subfigure}
-    \caption{AMG.
+%\begin{figure}[t!]
+%    \centering
+%    \begin{subfigure}[t]{0.25\textwidth}
+%        \centering
+%        \includegraphics[height=1.5in]{figs/appstudy/amg/amg_pip}
+%        \caption{}
+%        \label{fig:amg-communication-topology}
+%    \end{subfigure}%
+%    ~
+%    \begin{subfigure}[t]{0.24\textwidth}
+%        \centering
+%        \includegraphics[height=1.2in]{figs/appstudy/amg/amg_data_transfer}
+%        \caption{}
+%        \label{fig:amg-data-trans}
+%    \end{subfigure}
+%    \caption{AMG.
 (a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in AMG.
@@
 }
-\end{figure}
+%\end{figure}
 
@@
 
-Figure~\ref{fig:amg-data-trans} shows the amount of data each rank transferred in the example execution.
+%Figure~\ref{fig:amg-data-trans} shows the amount of data each rank transferred in the example execution.
 The data transferring of each rank is between 7MB to 16MB, corresponding to whether the rank is on a boundary or not.
@@
 
-\begin{figure}[t!]
-    \centering
-    \begin{subfigure}[t]{0.25\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/appstudy/cr/cr_pip}
-        \caption{Communication Topology}
-        \label{fig:cr-communication-topology}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[t]{0.22\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/appstudy/cr/cr_data_transfer}
-        \caption{Data Volume}
-        \label{fig:cr-data-trans}
-    \end{subfigure}
-    \caption{CrystalRouter.
+%\begin{figure}[t!]
+%    \centering
+%    \begin{subfigure}[t]{0.25\textwidth}
+%        \centering
+%        \includegraphics[height=1.5in]{figs/appstudy/cr/cr_pip}
+%        \caption{Communication Topology}
+%        \label{fig:cr-communication-topology}
+%    \end{subfigure}
+%    ~
+%    \begin{subfigure}[t]{0.22\textwidth}
+%        \centering
+%        \includegraphics[height=1.2in]{figs/appstudy/cr/cr_data_transfer}
+%        \caption{Data Volume}
+%        \label{fig:cr-data-trans}
+%    \end{subfigure}
+%    \caption{CrystalRouter.
 (a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in CrystalRouter.
@@
 
-Figure~\ref{fig:cr-data-trans} shows the amount of data each rank transferred in CrystalRouter, which ranges between 20MB to 36MB.
+%Figure~\ref{fig:cr-data-trans} shows the amount of data each rank transferred in CrystalRouter, which ranges between 20MB to 36MB.
 This is due to the rank counts not matching one-for-one with points in the hypercube.
@@
 
-\begin{figure}[t!]
-    \centering
-    \begin{subfigure}[t]{0.25\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/appstudy/mg/mg_pip}
-        \caption{Communication Topology}
-        \label{fig:mg-communication-topology}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[t]{0.22\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/appstudy/mg/mg_data_transfer}
-        \caption{Data Volume}
-        \label{fig:mg-data-trans}
-    \end{subfigure}
-    \caption{MultiGrid.
+%\begin{figure}[t!]
+%    \centering
+%    \begin{subfigure}[t]{0.25\textwidth}
+%        \centering
+%        \includegraphics[height=1.5in]{figs/appstudy/mg/mg_pip}
+%        \caption{Communication Topology}
+%        \label{fig:mg-communication-topology}
+%    \end{subfigure}
+%    ~
+%    \begin{subfigure}[t]{0.22\textwidth}
+%        \centering
+%        \includegraphics[height=1.2in]{figs/appstudy/mg/mg_data_transfer}
+%        \caption{Data Volume}
+%        \label{fig:mg-data-trans}
+%    \end{subfigure}
+%    \caption{MultiGrid.
 (a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in MultiGrid.
@@
 
-Figure~\ref{fig:mg-data-trans} shows the amount of data transferred between ranks in MultiGrid. For this miniapp, The data transferred per rank is relatively small and stable, between 2.8MB to 2.9MB.
-
-\section{Research Vehicle}
-\label{sec:codes}
-
-It is difficult to accurately and flexibly experiment with concurrently running jobs in an HPC context.
-One reason is that the allocation strategy used on production machine is part of the system software, which can not be changed by users.
-Even the system administrator is not authorized to make that change.
-Another reason is that it is unrealistic to reserve the system exclusively to run the same job with desired allocation without interference and then compare the results with those in the presence of interference.
-Therefore, we resort to simulation for this work.
-
-A simulation toolkit named CODES enables the exploration of simulating different HPC networks with high fidelity \cite{Jason-2011}\cite{mubarak-sc2012}.
-CODES is built on top of Rensselaer Optimistic Simulation System (ROSS) parallel discrete-event simulator, which is capable of processing billions of events per second on leadership-class supercomputers \cite{ross}.
-CODES support both torus and dragonfly network with high fidelity flit-level simulation.
-CODES has this network workload component that is capable of conducting trace-driven simulations.
-It can take real MPI application traces generated by SST DUMPI \cite{sst} to drive CODES network models.
-In this work, we focus on an in-depth analysis of intra- and inter-job communication interferences with  different job allocations on torus-connected HPC systems.
-Torus networks have been extensively used in the current generation of supercomputers because of their linear scaling on per-node cost and competitive communication performance \cite{mubarak-sc2012}\cite{zhou-ipdps}.
-
-\section{Configuration Study}
-\label{sec:config study}
-
-In this section, we study the communication behavior of three applications on torus networks with different dimensionality and bandwidth configurations.
-Torus networks are $k$-ary $n$-cubes, with $k^{n}$ nodes in total arranged in an $n$-dimensional grid having $k$ nodes in each dimension.
-Each node has $2\times n$ direct linked neighbor nodes with wraparound.
-In these and all following experiments, non-adaptive routing is used.
-While making specific recommendations as to an ``optimal'' network configuration for a given application is out of the scope of this study, the experiments do highlight the nuances inherent in design choices such as network diameter and bandwidth planning.
-
-We perform three sets of experiments.
-Each looks at per-rank distribution of performance of AMG, CrystalRouter and MultiGrid on a set of 2048-node torus models, one with $n=3$ ($16^{2} \times 8$), one with $n=5$ ($8\times 4^{4}$), and one with $n=7$ ($4^{4} \times 2^{3}$).
-Performance in this context is defined by the total message transfer time for each sending rank, which we are able to precisely measure due to the simulation environment providing us with an absolute notion of time.
-Note that we simply perform linear rank assignment, one per node -- later sections investigate rank and node mapping effects.
-In the first, shown in Figure~\ref{fig:dimensionality-study}, we keep the link bandwidth constant at 2GiB/s in each direction, corresponding to Blue Gene/Q systems~\cite{bgq}, leading to aggregate node bandwidths of 12GiB/s, 20GiB/s, and 28GiB/s for the 3D, 5D, and 7D tori, respectively.
-In the second, shown in Figure~\ref{fig:bandwidth-study}, we fix the torus dimensionality to 5D.
-In the third, shown in Figure~\ref{fig: bandwidth-time-box}, we instead fix the per-node aggregate bandwidth (and thus the system aggregate bandwidth) and modify the torus dimensionality, resulting in per-link bandwidths shown in Table~\ref{tab: fix-bandwidth}.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/dimenstudy/amg_box}
-        \caption{AMG}
-        \label{fig:dimen-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/dimenstudy/cr_box}
-        \caption{CrystalRouter}
-        \label{fig:dimen-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/dimenstudy/mg_box}
-        \caption{MultiGrid}
-        \label{fig:dimen-mg}
-    \end{subfigure}%
-   \caption{Data Transfer Time of AMG, CrystalRouter and MultiGrid on 3D, 5D and 7D torus network, with constant per-link bandwidth.}
-   \label{fig:dimensionality-study}
-\end{figure*}
-
-In the fixed link bandwidth experiment, shown in Figure~\ref{fig:dimensionality-study}, the data transfer time of the three applications are significantly reduced as the dimensionality increases, so the applications can easily make use of the increased dimensionality (lower hops on average) and/or aggregate bandwidth (more links being concurrently used).
-Interestingly, AMG has the greatest degree of outliers compared to the others.
-We expect this to be a combination of non-optimal rank mappings causing a greater degree of variance due to the highly regular communication pattern and the distribution of message sizes.
-MultiGrid seems to benefit the most in terms of reducing the spread of the time distribution, though the median message time remains roughly within a factor of two.
-The time distribution in CrystalRouter seems to accrue the most benefits, likely due to the combination of cross-cutting and localized per-rank traffic.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/bandwidthstudy/amg_bw_box}
-        \caption{AMG}
-        \label{fig:bdwstudy-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/bandwidthstudy/cr_bw_box}
-        \caption{CrystalRouter}
-        \label{fig:bdwstudy-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/bandwidthstudy/mg_bw_box}
-        \caption{MultiGrid}
-        \label{fig:bdwstudy-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 5D torus network with 2GiB/s, 4GiB/s and GiB/s direct link bandwidth. }
-   \label{fig:bandwidth-study}
-\end{figure*}
-
-In the fixed dimensionality experiment, shown in Figure~\ref{fig:bandwidth-study}, the increased per-link bandwidth, expectedly, benefits each application roughly equally, reducing the time distributions linearly.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/samebdw/amg}
-        \caption{AMG}
-        \label{fig:samebd-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/samebdw/cr}
-        \caption{CrystalRouter}
-        \label{fig:samebd-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/samebdw/mg}
-        \caption{MultiGrid}
-        \label{fig:samebd-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 3D, 5D and 7D torus network with same aggregated bandwidth. }
-   \label{fig: bandwidth-time-box}
-\end{figure*}
-
-\begin{table}[ht]
-\begin{center}
-\caption{Torus networks with same aggregated bandwidth} 
-\label{tab: fix-bandwidth}
-\begin{tabular}{l c c c} 
-\toprule % Top horizontal line
-\toprule
-&\multicolumn{3}{c}{Bandwidth } \\
-\cmidrule(l){2-4}
-Dimension  & per-link & aggregate &\\ % Column names row
-\midrule % In-table horizontal line
-3D      & 4.67GiB/s  & 28GiB/s  & \\  % Content row 1
-\midrule % In-table horizontal line
-5D    & 2.8GiB/s  & 28GiB/s  &\\ % Content row 1
-\midrule % In-table horizontal line
-7D    & 2GiB/s & 28GiB/s & \\
-\midrule
-\bottomrule % Bottom horizontal line
-\end{tabular}
-\end{center}
-\end{table}
-
-In the fixed aggregate bandwidth experiment, with configuration in Table~\ref{tab: fix-bandwidth} and results shown in Figure~\ref{fig: bandwidth-time-box}, the results are more nuanced.
-At least for these applications, the benefits from increased dimensionality do not outweigh the reduction in per-link bandwidth.
-MultiGrid does see benefits in the increased dimensionality, particularly with respect to outliers, but has a similar median message time.
-CrystalRouter reduces performance as the dimensionality increases, but only by a small degree.
-AMG remains roughly the same, minus a dip in the maximum message time that could be due to rank-to-node mapping changes.
-
-\section{Interference Analysis}
-\label{sec:interference}
-
-\textbf{\emph{Intra-job interference}} refers to the network contention between ranks within each application.
-\textbf{\emph{Inter-job interference}} is introduced by concurrently running jobs that adjacent to each other sharing network resources.
-Communication variability due to such interference can cause application performance degradation.
-
-In this section we will study both kinds of interference in the context of torus network. 
-Since application's communication pattern won't change with its scale, we use AMG with 216 ranks, CrystalRouter with 100 ranks, and MultiGrid 125 ranks.
-And in order to accommodate three applications running concurrently with different allocations, we simulate a 2K-node 3D torus ($16^{2} \times 8$) network for the experiments.
-
-\subsection{Intra-Job Interference Analysis}
-\label{sec: introjob}
-
-We design two sets of experiments to study the intra-job interference of each application.
-In the first, shown in Figure \ref{fig:shapestudy}, we assign each application with allocations in \emph{three different shapes}.
-In the second, shown in Figure \ref{fig:diffmap}, we study the intra-job interference by using different mapping strategies for each application with given allocation shape.
-
-In the allocation shapes experiment, we select three possible shapes can can commonly seen on the 3D torus network.
-They are 3D balanced-cube, 3D unbalanced-cube, 2D mesh, as shown in Figure \ref{fig:alloc-shapes}.
-
-3D balanced-cube, shown as red in Figure \ref{fig:cont_sub1}, can guarantee the minimum \emph{average pair-wise distance} within the allocation.
-Some research studies like \cite{leung,abhinav-sc13} indicate that compact allocation can guarantee jobs with better performance.
-They use a variety of metrics to evaluate the compactness of the allocation, such as \emph{average pair-wise distance}, \emph{diameter} and \emph{contiguity}.
-In this work, we select 3D balanced-cube as the most compact allocation on a 3D torus network.
-
-3D unbalanced-cube, shown as green in Figure \ref{fig:cont_sub1}, is a rectangular prism, which is the possible allocation shape on the systems with asymmetric networks.
-For example, Cray XE6/XK7 systems are deployed 3D tori with Gemini routers.
-The network connections in the \emph{y}-direction have only half the bandwidth of the cables used in the \emph{x} and \emph{z} directions.
-In order to take advantage of the faster links in the \emph{x} and \emph{z} directions, job allocation starts from X-Z plane, which leads to a rectangular prism shaped allocation \cite{RF}.
-
-2D mesh, shown as blue in Figure \ref{fig:cont_sub1}, can be cut out from a single layer of the 3D torus.
-2D mesh is a very common allocation shape in torus network for both contiguous and non-contiguous allocation strategies.
-For example, Cray Application Level Placement Scheduler (ALPS) indexes the nodes in torus network into a list and makes job allocation by simply going through the list \cite{carl-cug}.
-When the list is obtained by sorting the nodes based on their spacial coordinates in the torus, allocation made off from this list will be 2D mesh.
-The IBM Blue/Gene Q supercomputer Mira at Argonne Leadership Computing Facility also allow its allocation partition configured into mesh \cite{zhou-ipdps}.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/allocation}
-        \caption{ }
-        \label{fig:cont_sub1}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit16}
-        \caption{ }
-        \label{fig:noncont_sub1}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit8}
-        \caption{ }
-        \label{fig:noncont_sub2}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit2}
-        \caption{ }
-        \label{fig:noncont_sub3}
-    \end{subfigure}%
-    \caption{(a) Contiguous allocation in three different shapes.
-``Blue" is 2D mesh, ``Green" is 3D-unbalanced cube and ``Red" is 3D-balanced cube.
-(b)-(d) Non-contiguous allocation.
-Each job is represented by a specific color.
-The nodes assigned to different jobs are interleaved, the size of allocation unit are 16, 8 and 2.}
-    \label{fig:alloc-shapes}
-\end{figure*}
-
-AMG with 3D ``Nearest Neighbor" communication pattern takes less time when being assigned with 3D-unbalanced allocation than 3D-balanced, as shown in Figure \ref{fig:shapstudy-amg}.
-And MultiGrid gets the best performance (shortest data transfer time) when running on 3D-balanced allocation, as shown in Figure \ref{fig:shapstudy-mg}.
-Since MultiGrid's communication pattern is ``Many-to-Many" dominant, 3D-balanced is the most compact allocation with shortest pair-wise distance between nodes, which can reduce the aggregated hops for transferring message among ranks in MultiGrid.
-As CrystalRouter has both dominant local communication and multi-stage global data transfer, the 3D-balance is also the best allocation, but its advantage over 3D-unbalanced is not that obvious as to MultiGrid, as shown in Figure \ref{fig:shapstudy-cr}.
-
-Lots of research studies have designed fancy allocation algorithms to provide application with the most compact allocation\cite{leung,LO}.
-However, as shown in our experiments, providing cubical allocation without considering application's communication pattern won't guarantee the best performance for every application.
-Compact allocation should be provided to application with global data transfer, such as those conforms to ``Many-to-Many" communication pattern.
-Application with dominant communication pattern like ``Nearest Neighbor" won't fully utilize the compact allocation and could get even better performance on a relaxed allocation shape.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/amg_box}
-        \caption{AMG}
-        \label{fig:shapstudy-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/cr_box}
-        \caption{CrystalRouter}
-        \label{fig:shapstudy-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/mg_box}
-        \caption{MultiGrid}
-        \label{fig:shapstudy-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 3D-balanced, 3D-unbalanced and 2D allocation.}
-   \label{fig:shapestudy}
-\end{figure*}
-
-The rank-to-node mapping of parallel applications on HPC system could greatly impact applications' performance.
-However, finding the optimal mapping solution for a given application is out the scope of this work.
-Our experiments aims to show how mapping strategies impact the intra-job interference of application with specific communication pattern.
-
-We provide AMG, CrystalRouter and MultiGrid with 3D-balance allocation and use three mapping strategies to do the rank-to-node mapping.
-The ``Linear" mapping, as we used in the previous experiments, is to map each rank according to the dimensional ordering of compute nodes.
-The ``Cube" mapping assigns ranks into consecutive $2^{3}$ cubes.
-The ``Random" mapping assigns ranks randomly within the allocation.
-
-AMG remains roughly the same when it been mapped by ``Linear" and ``Cube", shown in Figure \ref{fig:diffmap-amg}.
-Although ``Linear" and ``Cube" mapping strategies cause some routing overlap for AMG's communication, both mapping strategies still preserve the locality of AMG's 3D ``Nearest Neighbor" communication pattern.
-The ``Random" mapping totally destroys AMG's communication pattern, and results in intra-job interference among all the ranks in AMG.
-The performance degradation caused by using ``Random" mapping is up to 90\%.
-
-The ``Cube" mapping improves CrystalRouter's performance over the ``Linear" by reducing the pair-wise distance between ranks, shown in Figure \ref{fig:diffmap-cr}.
-The global data transfer in CrystalRouter will takes fewer hops with ``Cube" mapping.
-The ``Random" mapping scatters the ranks within the local neighborhood of CrystalRouter and make their communication less efficient.
-
-The ``Cube" mapping benefits MultiGrid's ``Many-to-Many" communication.
-However, due to the small amount of data transfer among ranks in MultiGrid, ``Cube" mapping doesn't exhibit a drastic advantage.
-``Random" mapping neither cause too much degradation for the same reason, shown in Figure \ref{fig:diffmap-mg}.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/diffmapping/amg}
-        \caption{AMG}
-        \label{fig:diffmap-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/diffmapping/cr}
-        \caption{CrystalRouter}
-        \label{fig:diffmap-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/diffmapping/mg}
-        \caption{MultiGrid}
-        \label{fig:diffmap-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 3D-balanced allocation using  different mapping strategies.}
-   \label{fig:diffmap}
-\end{figure*}
-
-\subsection{Inter-Job Interference Analysis}
-\label{sec: interjob interference study}
-
-Inter-job interference has been identified as one of the major culprits for application's performance variability \cite{abhinav-sc13,skinner,rosenthal}.
-Inter-job interference is a more prominent issue for system adopted non-contiguous allocation policy than system with contiguous allocation.
-Application communication times have been demonstrated to vary from 36\% faster to 69\% slower due to job interference when running on non-contiguous allocation \cite{abhinav-sc13}.
-
-We assign each application with non-contiguous allocation and run them concurrently on the same network.
-The allocation unit belongs to different jobs are interleaved.
-In order to study the impact of different allocation unit size on applications' inter-job interference, we conduct experiments with unit size of 16, 8 and 2, shown respectively in Figure \ref{fig:noncont_sub1}, \ref{fig:noncont_sub1} and \ref{fig:noncont_sub3}.
-Figure \ref{fig:interjobstudy} shows the results of each application data transfer time with different allocation unit size.
-
-The data transfer time of AMG in Figure \ref{fig:interjob-amg} keeps stable between allocation unit size of 16 and 8, since the locality of AMG is 6-rank based.
-When the unit size reduce to 2, AMG suffers prolong data transfer time by about 10\%.
-CrystalRouter is more sensitive to allocation unit size.
-The best unit size would be the one big enough to accommodate its local neighborhood.
-Figure \ref{fig:interjob-cr} shows unit size of 16 and 8 can guarantee the same average data transfer time, while some rank spend more time with allocation unit size 8 than 16.
-When the unit size reduce to 2, the communication become less efficiency and takes about 15\% more times for transferring data.
-
-The data transfer time of MultiGrid with different allocation unit sizes doesn't show obvious variability in Figure \ref{fig:interjob-mg}.
-This is because even big allocation unit size like 16 will still fail to preserve MultiGrid's ``Many-to-Many" pattern.
-The data transfer time is almost doubled when MultiGrid running concurrently with allocation unit size of 16, as shown in Figure \ref{fig:interjob-mg} .
-As the unit size decreases, the data transfer time keep growing, but in a slow way, which means in terms of preserving the locality of MultiGrid, allocation unit of size 16, 8 and 2 serve equally bad.
-
-The unit size in non-contiguous allocation policy should be chosen according to application's communication pattern.
-Inter-job interference is inevitable in non-contiguous based systems, but unit size that big enough to preserve the neighborhood communication of the application will alleviate such interference and improve job performance.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/inter-job/amg}
-        \caption{AMG}
-        \label{fig:interjob-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/inter-job/cr}
-        \caption{CrystalRouter}
-        \label{fig:interjob-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/inter-job/mg}
-        \caption{MultiGrid}
-        \label{fig:interjob-mg}
-    \end{subfigure}%
-   \caption{Inter-job interference study.
-``cont" indicates three applications running side by side concurrently on the same network with contiguous allocation.
-To study the impact of non-contiguous allocation on inter-job interference, applications running concurrently with interleaved allocation of different unit sizes, which are 16-node, 8-node, 2-node.
-}
-   \label{fig:interjobstudy}
-\end{figure*}
-
-\subsection{Result Summary}
-\label{sec:summary}
-
-Based on the above comprehensive study, we can make following observations:
-\begin{itemize}
-    \item The compact allocation may not guarantee the best performance for every application. 
-    \item The applications dominated with Nearest Neighbor communications exhibit relatively stable performance under different allocation shapes, while the applications dominated with Many-to-Many communication exhibit better performance with compact allocation (e.g.
-3D balanced).
-    \item A good rank-to-node mapping strategy can greatly improve application's performance when specific allocation is given.
-    \item An optimal size for allocation units should be determined according to an application's dominant communication pattern.
-In general, a unit size should be large enough to preserve neighboring communication of the application.
-    \item Inter-job interference is inevitable in non-contiguous allocation.
-However, choosing the proper allocation unit size with job's communication pattern awareness can alleviate such negative effect.
-\end{itemize}
-
-\section{Discussion}
-\label{sec:discussion}
-
-The results shown in the previous section provide insights for the design of smart and flexible job allocation.
-By scrutinizing job's communication behavior, we can identify job's dominant communication pattern and pinpoint the ``neighborhood communication".
-With such knowledge about jobs' communication patterns, we can analyze the possible interference between jobs and take precautions to alleviate the negative effect when making allocation decisions.
-
-When an application with dominant communication that is intensive ``Many-to-Many", the scheduler should grant it with compact node allocation and exclusive network provision.
-The compact allocation can guarantee the shortest pair-wise distance between all the ranks, and make the data transfer between ranks take less hops.
-On the other hand, the exclusive network provision will prevent other adjacent applications from sharing network resources, thus eliminate the performance degradation due to interference.
-
-Not every application's preferable resource are compact node allocation and exclusive network provision.
-As we found in our analysis, applications whose dominant communication patterns contain intensive ``neighborhood communication" like ``Nearest Neighbor" don't benefit from compact node allocation and exclusive network provision.
-Applications, such as AMG, can run with non-contiguous node allocation without performance degradation, as long as the allocation unit can accommodate their ``neighborhood communication".
-
-The allocation unit size shouldn't be fixed.
-Instead, the scheduler should choose a proper unit size based on the scale of ``neighborhood communication" of each job.
-The best unit size should be neither too big nor too small, jut perfectly fit for the size of that ``neighborhood".
-Big unit size won't be fully utilized and also cause fragmentation.
-Small unit size won't be able to accommodate the ``neighborhood communication" and makes the intra-job communication less efficient.
-
-The advantage of making allocation with consideration about job's communication pattern is obvious.
-First, compared with contiguous allocation policy, the new scheduler outlined above is more flexible, it doesn't require the system to provide a big contiguous partition to accommodate the whole application, instead it only needs to provide a small set of compact nodes that is sufficient for all the "local community" in the application.
-On the other hand, the new scheduler outlined above won't destroy the locality of application's communication pattern.
-Indeed, the small compact node set (allocation unit) provided for the application's  ``local communities" can preserve the communication locality in the maximum extent.
-The design of such a smart scheduler is part of our future work.
-
-\section{Related Work}
-\label{sec:related_work}
-
-There are many tools available for system monitoring and application profiling.
-Tools like TAU (Tuning and Analysis) \cite{tau}, SCALASCA \cite{scala} and mpiP \cite{mpip} can capture application's runtime information about application's communication and computation in event traces.
-However, recognizing the patterns about application's communication from those traces requires substantial effort.
-There are a number of studies on the recognition and characterization of parallel application communication patterns.
-Oak Ridge National Laboratory has its ongoing project about developing a tool set named Oxbow, which can characterize the computation and communication behavior of scientific applications and benchmarks \cite{oxbow}.
-In a recent work \cite{roth}, the authors demonstrate a new approach to automatically characterizing parallel applications' communication behaviors.
-
-%\textbf{P2: communication characteristics}\\
-
-Many research efforts have been conducted to characterize scientific applications.
-For instance, a DOE project aims at the identification of the computational characteristics of the DOE MiniApps developed at various exascale co-design centers \cite{designforwardwebpage}.
-In this project, the communication patterns of several DOE full applications and associated mini-applications are studied to provide a more complete snapshot of the DOE workload.
-A joint project named CORAL from Oak Ridge, Argonne and Livermore provide a series of benchmarks to represent DOE workloads and technical requirements \cite{coral}.
-The CORAL project includes scalable science benchmarks, throughput benchmarks, data centric benchmarks, skeleton benchmarks and Micro benchmarks.
-
-%\textbf{P3: Interference between concurrently running jobs}\\
-The interference between concurrently running jobs on HPC system have been identified as major culprit for job's performance variability.
-Abhinav et al.
-found that concurrently running applications on HPC can cause interference to each other, and cause communication time varied from 36\% faster to 69\% slower.
-Such interference could come from Operating System noise, shape of the allocated partition , or other running jobs that sharing network resources \cite{abhinav-sc13}.
-Skinner et al.
-found that  there is a 2-3$\times$ slowdown in MPI$\textunderscore$Allreduce due to network contention from other jobs \cite{skinner}.
-Rosenthal et al.
-found increasing network bandwidth provide limited benefits to a number of applications.
-The applications that send mostly small messages or large messages asynchronously are not bandwidth bounded, hence, benefit only slightly or not at all from increased bandwidth \cite{rosenthal}.
-
-%\textbf{P4: Job allocation(little bit)}\\
-Several research studies focus on optimizing job allocation on HPC system to alleviate the interference between concurrently running jobs.
-Hoefler et al.
-proposed to use performance modeling techniques to analyze factors that impact the performance of parallel scientific applications \cite{hoefler-modeling}.
-However, as the scale of HPC continue to grow, the  interference of concurrently running jobs is getting worse, which is hard to be quantified by performance profiling tools alone.
-Bogdan et al.
-provide a set of guidelines of how to configure a network with Dragonfly topology for workload with Nearest Neighbor communication pattern \cite{Bogdan-hpdc14}.
-Dong et al.
-have developed simple benchmarks that conforms to four different communication patterns, namely ping-pong, nearest neighbor, broadcast and all$\textunderscore$reduce, to demonstrate the effectiveness of this highly parallel 5D torus network \cite{Dong-SC11}.
-
-Our work is different from all these works in the following ways.
-First, we focus on the dominant communication patterns rather than any specific application.
-We believe this can provide a guideline for other research work.
-Secondly, we explore the intra- and inter-job interference between concurrently running jobs, while similar work such as \cite{abhinav-sc13} only focuses on the single application's performance degradation due to network contention.
-Finally, we analyze the impact of different allocation strategies to job's communication behaviors.
-We further identify the optimal allocation strategy for each application with a specific dominant communication pattern.
-Based on our study, we claim that further scheduling should take job's communication pattern into consideration for its allocation decision making.
-
-\section{Conclusions}
-\label{sec:conclusion}
-In this work, we have studied the communication behavior of three parallel applications, namely AMG, CrystalRouter and MultiGrid.
-Each application has distinctive communication pattern, that can be representative for a group of jobs in HPC workload.
-We have used a sophisticate simulation tool named CODES from Argonne National Laboratory to simulate the running of these three parallel applications on torus network.
-The torus network provided by CODES has good fidelity and scalability.
-We have analyzed the performance of each application's communication in terms of data transfer time by simulating them running on torus networks with different bandwidth and dimensionality configurations.
-We have found that higher dimensionality of torus network would improve the performance of application with ``Many-to-Many" communication patterns, while application with intensive local communication like ``Nearest Neighbor" won't benefit much from higher dimensionality.
-
-We have analyzed the intra- and inter-job interference by simulating three applications running on 3D torus network.
-Based on our comprehensive experiments, we got five observations.
-1) The compact allocation may not guarantee the best performance for every application.
-2)The applications dominated with Nearest Neighbor communications exhibit relatively stable performance under different allocation shapes, while the applications dominated with Many-to-Many communication exhibit better performance with compact allocation (e.g.
-3D balanced).
-3) A good rank-to-node mapping strategy can greatly improve application's performance when specific allocation is given.
-4) An optimal size for allocation units should be determined according to an application's dominant communication pattern.
-In general, a unit size should be large enough to preserve neighboring communication of the application.
-5) Inter-job interference is inevitable in non-contiguous allocation.
-However, choosing the proper allocation unit size with job's communication pattern awareness can alleviate such negative effect.
-
-We believe that our finding in this work can provide valuable guidance for HPC resource management to make flexible job allocations.
-Rather than using pre-defined partitions and non-contiguous allocation, future HPC systems should assign each job with preferable resources based on job's communication pattern.
-
-\section*{Acknowledgment}
-\label{sec: ack}
-The work at Illinois Institute of Technology is supported in part by U.S.
-National Science Foundation grants CNS-1320125 and CCF-1422009.
-This work is also supported by the U.S.
-Department of Energy, Office of Science, Advanced Scientific Computing Research, under Contract DE-AC02-06CH11357.
-
-\bibliographystyle{IEEEtran}
-\bibliography{./reference}
-
-\vspace{5\baselineskip}
-
-\begin{framed}
-The submitted manuscript has been created by UChicago Argonne, LLC, Operator of Argonne National Laboratory ("Argonne").
-Argonne, a U.S.
-Department of Energy Office of Science laboratory, is operated under Contract No.
-DE-AC02-06CH11357.
-The U.S.
-Government retains for itself, and others acting on its behalf, a paid-up nonexclusive, irrevocable worldwide license in said article to reproduce, prepare derivative works, distribute copies to the public, and perform publicly and display publicly, by or on behalf of the Government.
-\end{framed}
-
-\end{document}
+%Figure~\ref{fig:mg-data-trans} shows the amount of data transferred between ranks in MultiGrid. For this miniapp, The data transferred per rank is relatively small and stable, between 2.8MB to 2.9MB.

### 2016-07-07 conclusion.tex
split into separate tex files for each section

--- 2015-10-17 submission.tex
@@
-\documentclass[conference]{IEEEtran}
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
-\else
-
-\fi
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-\usepackage{algorithm}
-%\usepackage{algorithm2e}
-\usepackage{algpseudocode}
-\usepackage{booktabs}
-\usepackage{flushend}
-\usepackage{url}
-\usepackage{xcolor}
-\usepackage{amsmath}
-\usepackage{framed}
-\usepackage{textcomp}
-%\usepackage{subfig}
-\usepackage{graphicx}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{soul}
-\usepackage{verbatim}
-%\usepackage[labelfont=bf]{caption}
-\usepackage[font=bf,labelsep=space]{caption}
-
-\newcommand{\TODO}[1]{\hl{TODO: #1}}
-\newcommand{\NOTE}[1]{\hl{NOTE: #1}}
-\newcommand{\changed}[1]{\textcolor{red}{#1}}
-
-\begin{document}
-%
-% paper title
-% can use linebreaks \\ within to get better formatting as desired
-\title{Study of Intra- and Inter-Job Interference with Different Job Allocations}
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
-\author{
-
-\IEEEauthorblockN{Xu Yang\IEEEauthorrefmark{1}, John Jenkins\IEEEauthorrefmark{2}, Misbah Mubarak\IEEEauthorrefmark{2}, Robert B.
-Ross\IEEEauthorrefmark{2}, Zhou Zhou\IEEEauthorrefmark{1}, Zhiling Lan\IEEEauthorrefmark{1}}
-
-\IEEEauthorblockA{\IEEEauthorrefmark{1}Department of Computer Science,
-Illinois Institute of Technology,
-Chicago, Illinois, USA 60616\\
-\{xyang56, zzhou\}@hawk.iit.edu, lan@iit.edu}
-
-\IEEEauthorblockA{\IEEEauthorrefmark{2}Mathematics and Computer Science Division, Argonne National Laboratory,
-Argonne, IL, USA 60439\\
-\{jenkins,rross\}@mcs.anl.gov, mmubarak@anl.gov}
-}
-
-% make the title area
-\maketitle
-
-\begin{abstract}
-
-Network contention between concurrently running jobs on HPC systems is a primary cause of performance variability.
-Optimizing job allocation and avoiding network sharing are hence crucial to alleviate the potential degradation in performance.
-In order to do these effectively, an understanding of the interaction between applications, their communication patterns, and the system architecture are required.
-In this work, we choose three representative HPC applications from DOE Design Forward Project and conduct detailed simulations of high-performance network topologies to analyze both intra- and inter-job interference.
-By scrutinizing the communication behaviors of these applications, we identify relationships between these behaviors and the possible interference introduced by different job allocation scenarios.
-Our comprehensive analyses in this work illuminate a path towards job-communication-pattern-aware HPC resource allocation.
-
-\end{abstract}
-
-\IEEEpeerreviewmaketitle
-
-\section{Introduction} 
-\label{sec: intro}
-
-The scale of supercomputers continually grows at a breakneck pace to accommodate the computing power requirements for scientific research.
-Supercomputers have tens of thousands of nodes and serve as an irreplaceable research vehicle for scientific problems with increasing size and complexity.
-Supercomputers are usually employed as a shared resource to accommodate many parallel applications (jobs) running concurrently.
-These parallel jobs share the system infrastructure such as network and I/O bandwidth, and inevitably there is contention over these shared resources.
-As supercomputers continue to evolve, these shared resources are increasingly the bottleneck for performance.
-
-\begin{figure}[h!]
-    \centering
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/goodallocation}
-        \caption{Contiguous}
-        \label{fig:overview_sub1}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/badallocation}
-        \caption{Non-contiguous}
-        \label{fig:overview_sub2}
-    \end{subfigure}%
-   \caption{Multiple jobs running concurrently with different allocations.
-Each job is represented by a specific color.
-a) shows the effect of contiguous allocation, which reduce the inter-job interference.
-b) shows non-contiguous allocation, which may introduce both intra and inter-job interferences.
-}
-   \label{fig:overview}
-\end{figure}
-
-When submitting jobs to HPC systems, users request resources in terms of node counts and expected runtime.
-The batch scheduler is then responsible for dispatching the submitted jobs to the system.
-Typically, multiple jobs run concurrently, resulting in the shared use of resources, particularly network and storage.
-A prominent problem with this approach is network contention among concurrently running jobs.
-If the network resources of concurrently running jobs aren't isolated, the resulting sharing can cause communication variability that results in performance degradation for some or all such jobs \cite{abhinav-sc13}.
-This delay can propagate into the queueing time of the following submitted jobs, thus leading to low system throughput and utilization \cite{jose-ipdps15}.
-This adverse effect of network sharing can be mitigated by providing jobs with isolated allocations and exclusive network resources.
-
-On the widely used torus-connected HPC systems \cite{bgq,tofu,titan}, two allocation strategies are commonly used.
-The \emph{contiguous allocation} strategy assigns to each job a compact and contiguous set of computing nodes, as shown in Figure \ref{fig:overview_sub1}.
-The partition-based allocation approach used in Blue Gene series systems is an example of such a strategy \cite{bgloverview}.
-The contiguous strategy favors application performance through isolated networking within a partition and the locality that implies.
-However, this strategy can cause both internal fragmentation (when more nodes are allocated to a job than it requests) and external fragmentation (when sufficient nodes are available for a request, but they can not be allocated contiguously), therefore leading to lower system utilization than is otherwise possible.
-On the other side of the coin, the \emph{non-contiguous allocation} strategy, used by the Cray XT/XE series \cite{carl-cug}, assigns free nodes to jobs regardless of contiguity, though of course efforts are made to maximize locality.
-Figure~\ref{fig:overview_sub2} shows an example non-contiguous allocation.
-While eliminating internal and external fragmentation as seen in contiguous allocation systems, in return non-contiguous allocations introduce contention between jobs due to the interleaving of job nodes.
-The non-contiguous node allocation can significantly reduce job performance, especially for communication-intensive ones \cite{abhinav-sc13}.
-
-We envision that future HPC schedulers will adopt a flexible job allocation mechanism which combines the best of both contiguous and non-contiguous allocation strategies.
-Such a flexible mechanism would take shared resource needs of jobs into account when making allocation decisions (e.g., network).
-With knowledge and analysis of job communication patterns, it can be identified which jobs require network isolation and locality, and to what degree.
-Then, rather than allocating each job in a ``know-nothing'' manner, one may specialize allocation so that, for example, only the jobs with stringent network needs are given compact, isolated allocations, resulting in maximized utilization and minimized perceivable resource contention effects.
-
-In this work, we focus on an in-depth analysis of intra- and inter-job communication interference with different job allocation strategies on torus-connected HPC systems.
-Torus-based networks are used on six of the top 10 supercomputers on the June 2015 Top500 lists~\cite{top500}.
-The current generation of IBM Blue Gene/Q (BG/Q) supercomputers, such as Mira at Argonne Nation Laboratory and Sequoia at Lawrence Livermore National Laboratory, has its nodes connected in a 5D torus network \cite{bgq}.
-The K computer from Japan uses the ``Tofu'' system, which is a 6D mesh/torus topology \cite{tofu}.
-Titan, a Cray XK7 supercomputer located at the Oak Ridge Leadership Computing Facility (OLCF), has nodes connected in a 3D torus within the compute partition \cite{titan}.
-Although our analyses are based on torus networks, the ideas conveyed in this work is applicable to networks with different topologies.
-
-We selected three signature applications from the DOE Design Forward Project~\cite{designforwardwebpage} as examples to conduct detailed study about their communication patterns.
-We use a sophisticated simulation toolkit named CODES (standing for Co-Design of Multi-layer Exascale Storage Architectures)~\cite{Jason-2011} as a research vehicle to evaluate the performance of these applications with various allocations in a controlled environment.
-We then analyze the intra- and inter-job interference by simulating these applications running exclusively and concurrently with different allocations.
-We believe the insights presented in this work can be very useful for the design of future HPC batch job schedulers and resource managers.
-
-%\NOTE{JJ removed the list of contributions, because they are exactly the same as the preceding paragraph.}
-\begin{comment}
-The observations of this work are the following:
-
-\begin{itemize}
-    \item We study the communication behaviors and identify the communication patterns of three representative HPC applications.
-    \item We conduct detailed analysis about the intra- and inter-job interference by running three applications with different allocation strategies.
-And we find the relation between application's communication pattern and the cause of interference.
-    \item We show a novel and flexible way to make allocation with awareness of job's communication pattern, which  is every effective to alleviate the interference between concurrently running jobs.
-\TODO{do we actually do this, or provide the underpinnings for such a method?}
-\end{itemize}
-
-We believe the finds and insights presented in this work would be very useful for the design of future HPC batch job scheduler and resource management module.
-
-\end{comment}
-
-The rest of this paper is organized as follows.
-Section~\ref{sec:application study} describes the three representative applications from DOE Design Forward Project for our study.
-Section~\ref{sec:codes} talks about the use of CODES as research vehicle for our work.
-Section~\ref{sec:config study} shows the performance analysis of three applications on different torus networks.
-Section~\ref{sec:interference} provides detailed analysis about the intra- and inter-job interference between three applications on torus network with different allocation strategies.
-Section~\ref{sec:discussion} introduces a path toward communication-pattern aware allocation strategies, given the results of our analysis.
-Section~\ref{sec:related_work} discusses related work.
-Finally, the conclusion is presented in Section~\ref{sec:conclusion}.
-
-\section{Application Study}
-\label{sec:application study}
-
-%A parallel application usually conforms to a combination of several basic communication patterns\cite{roth}.
-At its different execution phases, the application's communication behavior may follow different basic patterns respectively.
-\textcolor{blue}{When we look into the data flow during application execution, most parallel applications start with broadcast operation to distribute the data from root process to other processes\textcolor{red}{ref needed here}, followed by a series of computation and communication that conforms to certain pattern, which is usually the dominant part of application's execution.
-Before the application come to completion, all the working process will return their results to the root directly or hierarchically.} Parallel application's communication pattern usually involves lots of factors, such as communication intensity, operation-to-operation dependencies and the ``critical path" in its communication topology graph, etc.
-In this work, we focus on the locality of application's rank-to-rank communication topology graph.
-
-A parallel application usually conforms to a combination of several basic communication patterns \cite{roth}.
-At its different execution phases, the application's communication behavior may follow different basic patterns respectively.
-There are many profiling tools available to capture information regarding communication patterns of parallel applications~\cite{tau,mpip,scala,sst,oxbow}.
-They can provide information such as the percentage of different MPI operations of the applications, communication topology, the amount of data transferred between processes.
-
-There are many applications running on HPC systems, such as Mira \cite{bgq} and Titan \cite{titan}, that are communication intensive and conform to various communication patterns.
-In this work, we select three representative applications from the DOE Design Forward Project.
-Each application exhibits a distinctive communication pattern that is commonly seen in HPC applications.
-We believe that the communication patterns of these applications are representative of a wide array of applications running on leadership-class machines.
-Specifically, we study the Algebraic MultiGrid Solver (AMG), Geometric MultiGrid (MultiGrid) and CrystalRouter MiniApps.
-The communication pattern figures we present in Section~\ref{sec:amg}, \ref{sec:crystalrouter}, and \ref{sec:multigrid} are generated with the IPM~\cite{ipm} data gathered from publicly available traces~\cite{designforwardwebpage}.
-
-% MiniApps are reduced proxy applications that encapsulate the salient performance of larger full size applications \cite{miniapp}.
-Since we only focus on the dominant communication pattern of parallel applications, these MiniApps are perfect candidates for analysis the application's performance on different allocations or on new network architectures.
-We choose three MiniApps, namely, Algebraic Multigrid Solver(AMG), Geometric MultiGrid(MultiGrid) and CrystalRouter.
-Their representative communication patterns that are commonly seen in the HPC system workload.
-The communication pattern figures we present here are generated with the IPM \cite{ipm} data from \cite{designforwardwebpage}.
-
-\subsection{AMG}
-\label{sec:amg}
-
-\begin{figure}[t!]
-    \centering
-    \begin{subfigure}[t]{0.25\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/appstudy/amg/amg_pip}
-        \caption{}
-        \label{fig:amg-communication-topology}
-    \end{subfigure}%
-    ~
-    \begin{subfigure}[t]{0.24\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/appstudy/amg/amg_data_transfer}
-        \caption{}
-        \label{fig:amg-data-trans}
-    \end{subfigure}
-    \caption{AMG.
-(a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in AMG.
-(b) The amount of data sent/received by each rank in AMG.
-Y-axis labels the amount of data transferred in MB, x-axis labels 216 ranks of AMG.
-}
-\end{figure}
-
-The Algebraic MultiGrid Solver, or AMG, is a parallel algebraic multi-grid solver for linear systems arising from problems on unstructured mesh physics packages.
-It has been derived directly from the BoomerAMG solver that is being developed in the Center for Applied Scientific Computing (CASC) at LLNL \cite{amg}.
-The dominant communication pattern is regional communication with decreasing message size for different parts of the multi-grid v-cycle.
-
-Figure~\ref{fig:amg-communication-topology} shows the communication topology of a small scale AMG execution with 216 MPI Ranks.
-Note that the dominant communication pattern of the application doesn't change with scale.
-We can make the observation that AMG's dominant communication pattern is 3D Nearest Neighbor, each rank has intensive communication with up to six neighbors, depending on the boundaries of the ranks.
-Applications with similar pattern include PARTISN \cite{partisn} and SNAP \cite{snap}.
-
-Figure~\ref{fig:amg-data-trans} shows the amount of data each rank transferred in the example execution.
-The data transferring of each rank is between 7MB to 16MB, corresponding to whether the rank is on a boundary or not.
-
-\subsection{Crystal Router}
-\label{sec:crystalrouter}
-
-\begin{figure}[t!]
-    \centering
-    \begin{subfigure}[t]{0.25\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/appstudy/cr/cr_pip}
-        \caption{Communication Topology}
-        \label{fig:cr-communication-topology}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[t]{0.22\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/appstudy/cr/cr_data_transfer}
-        \caption{Data Volume}
-        \label{fig:cr-data-trans}
-    \end{subfigure}
-    \caption{CrystalRouter.
-(a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in CrystalRouter.
-(b) The amount of data sent/received by each rank in CrystalRouter.
-Y-axis labels the amount of data transferred in MB, x-axis labels 100 ranks of CrystalRouter.
-}
-\end{figure}
-
-The second MiniApp covered is CrystalRouter, the extracted communication kernel of the full application Nek5000~\cite{crystalrouter}, which is a spectral element CFD application developed at Argonne National Laboratory \cite{nek5000}.
-It features spectral element multi-grid solvers coupled with a highly scalable, parallel coarse-grid solver that is widely used for projects including ocean current modeling, thermal hydraulics of reactor cores, and spatiotemporal chaos.
-CrystalRouter demonstrates the ``many-to-many'' communication pattern through a scalable multi-stage communication process.
-The way CrystalRouter works is nodes compute for a while, then synchronize and communicate, continually alternating between these two types of activities.
-
-The collective communication in CrystalRouter utilizes a recursive doubling approach.
-Ranks in CrystalRouter conform to a $n$-dimensional hypercube and recursively split into ($n$-1)-dimensional hypercubes, with communication occurring along the splitting plane.
-The pattern of this communication can be found in Figure~\ref{fig:cr-communication-topology}.
-By nature of the logarithmic splitting process, a substantial portion of the communication occurs in small neighborhoods of ranks.
-CrystalRouter represents a group of applications whose dominant communication is a hybrid of multi-stage local and hierarchical global communication, and shares similarities with most MPI collective communication implementations.
-
-Figure~\ref{fig:cr-data-trans} shows the amount of data each rank transferred in CrystalRouter, which ranges between 20MB to 36MB.
-This is due to the rank counts not matching one-for-one with points in the hypercube.
-
-\subsection{MultiGrid}
-\label{sec:multigrid}
-
-\begin{figure}[t!]
-    \centering
-    \begin{subfigure}[t]{0.25\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/appstudy/mg/mg_pip}
-        \caption{Communication Topology}
-        \label{fig:mg-communication-topology}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[t]{0.22\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/appstudy/mg/mg_data_transfer}
-        \caption{Data Volume}
-        \label{fig:mg-data-trans}
-    \end{subfigure}
-    \caption{MultiGrid.
-(a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in MultiGrid.
-(b) The amount of data sent/received by each rank in MultiGrid.
-Y-axis labels the amount of data transferred in MB, x-axis label 125 ranks of MultiGrid.
-}
-\end{figure}
-
-MultiGrid is geometric multi-grid v-cycle from the production elliptic solver BoxLib, a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes \cite{boxlib}.
-MultiGrid conforms to many-to-many communication pattern with decreasing message size and collectives for different parts of the multi-grid v-cycle.
-It is widely used for structured grid physics packages.
-
-Figure~\ref{fig:mg-communication-topology} shows the communication topology of MultiGrid with 125 ranks.
-We can see intensive communication along the diagonal that resembles nearest neighbor communication, similar to AMG.
-However, the communication topology leads to a greater ``spread'' of communication across the set of ranks, challenging the maximization of communication locality with respect to rank.
-In this sense, it can be considered a ``many-to-many'' pattern.
-Applications with such similar dominant communication pattern are like FillBoundary, another PDE solver code in \cite{boxlib}.
-
-Figure~\ref{fig:mg-data-trans} shows the amount of data transferred between ranks in MultiGrid. For this miniapp, The data transferred per rank is relatively small and stable, between 2.8MB to 2.9MB.
-
-\section{Research Vehicle}
-\label{sec:codes}
-
-It is difficult to accurately and flexibly experiment with concurrently running jobs in an HPC context.
-One reason is that the allocation strategy used on production machine is part of the system software, which can not be changed by users.
-Even the system administrator is not authorized to make that change.
-Another reason is that it is unrealistic to reserve the system exclusively to run the same job with desired allocation without interference and then compare the results with those in the presence of interference.
-Therefore, we resort to simulation for this work.
-
-A simulation toolkit named CODES enables the exploration of simulating different HPC networks with high fidelity \cite{Jason-2011}\cite{mubarak-sc2012}.
-CODES is built on top of Rensselaer Optimistic Simulation System (ROSS) parallel discrete-event simulator, which is capable of processing billions of events per second on leadership-class supercomputers \cite{ross}.
-CODES support both torus and dragonfly network with high fidelity flit-level simulation.
-CODES has this network workload component that is capable of conducting trace-driven simulations.
-It can take real MPI application traces generated by SST DUMPI \cite{sst} to drive CODES network models.
-In this work, we focus on an in-depth analysis of intra- and inter-job communication interferences with  different job allocations on torus-connected HPC systems.
-Torus networks have been extensively used in the current generation of supercomputers because of their linear scaling on per-node cost and competitive communication performance \cite{mubarak-sc2012}\cite{zhou-ipdps}.
-
-\section{Configuration Study}
-\label{sec:config study}
-
-In this section, we study the communication behavior of three applications on torus networks with different dimensionality and bandwidth configurations.
-Torus networks are $k$-ary $n$-cubes, with $k^{n}$ nodes in total arranged in an $n$-dimensional grid having $k$ nodes in each dimension.
-Each node has $2\times n$ direct linked neighbor nodes with wraparound.
-In these and all following experiments, non-adaptive routing is used.
-While making specific recommendations as to an ``optimal'' network configuration for a given application is out of the scope of this study, the experiments do highlight the nuances inherent in design choices such as network diameter and bandwidth planning.
-
-We perform three sets of experiments.
-Each looks at per-rank distribution of performance of AMG, CrystalRouter and MultiGrid on a set of 2048-node torus models, one with $n=3$ ($16^{2} \times 8$), one with $n=5$ ($8\times 4^{4}$), and one with $n=7$ ($4^{4} \times 2^{3}$).
-Performance in this context is defined by the total message transfer time for each sending rank, which we are able to precisely measure due to the simulation environment providing us with an absolute notion of time.
-Note that we simply perform linear rank assignment, one per node -- later sections investigate rank and node mapping effects.
-In the first, shown in Figure~\ref{fig:dimensionality-study}, we keep the link bandwidth constant at 2GiB/s in each direction, corresponding to Blue Gene/Q systems~\cite{bgq}, leading to aggregate node bandwidths of 12GiB/s, 20GiB/s, and 28GiB/s for the 3D, 5D, and 7D tori, respectively.
-In the second, shown in Figure~\ref{fig:bandwidth-study}, we fix the torus dimensionality to 5D.
-In the third, shown in Figure~\ref{fig: bandwidth-time-box}, we instead fix the per-node aggregate bandwidth (and thus the system aggregate bandwidth) and modify the torus dimensionality, resulting in per-link bandwidths shown in Table~\ref{tab: fix-bandwidth}.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/dimenstudy/amg_box}
-        \caption{AMG}
-        \label{fig:dimen-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/dimenstudy/cr_box}
-        \caption{CrystalRouter}
-        \label{fig:dimen-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/dimenstudy/mg_box}
-        \caption{MultiGrid}
-        \label{fig:dimen-mg}
-    \end{subfigure}%
-   \caption{Data Transfer Time of AMG, CrystalRouter and MultiGrid on 3D, 5D and 7D torus network, with constant per-link bandwidth.}
-   \label{fig:dimensionality-study}
-\end{figure*}
-
-In the fixed link bandwidth experiment, shown in Figure~\ref{fig:dimensionality-study}, the data transfer time of the three applications are significantly reduced as the dimensionality increases, so the applications can easily make use of the increased dimensionality (lower hops on average) and/or aggregate bandwidth (more links being concurrently used).
-Interestingly, AMG has the greatest degree of outliers compared to the others.
-We expect this to be a combination of non-optimal rank mappings causing a greater degree of variance due to the highly regular communication pattern and the distribution of message sizes.
-MultiGrid seems to benefit the most in terms of reducing the spread of the time distribution, though the median message time remains roughly within a factor of two.
-The time distribution in CrystalRouter seems to accrue the most benefits, likely due to the combination of cross-cutting and localized per-rank traffic.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/bandwidthstudy/amg_bw_box}
-        \caption{AMG}
-        \label{fig:bdwstudy-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/bandwidthstudy/cr_bw_box}
-        \caption{CrystalRouter}
-        \label{fig:bdwstudy-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/bandwidthstudy/mg_bw_box}
-        \caption{MultiGrid}
-        \label{fig:bdwstudy-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 5D torus network with 2GiB/s, 4GiB/s and GiB/s direct link bandwidth. }
-   \label{fig:bandwidth-study}
-\end{figure*}
-
-In the fixed dimensionality experiment, shown in Figure~\ref{fig:bandwidth-study}, the increased per-link bandwidth, expectedly, benefits each application roughly equally, reducing the time distributions linearly.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/samebdw/amg}
-        \caption{AMG}
-        \label{fig:samebd-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/samebdw/cr}
-        \caption{CrystalRouter}
-        \label{fig:samebd-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/samebdw/mg}
-        \caption{MultiGrid}
-        \label{fig:samebd-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 3D, 5D and 7D torus network with same aggregated bandwidth. }
-   \label{fig: bandwidth-time-box}
-\end{figure*}
-
-\begin{table}[ht]
-\begin{center}
-\caption{Torus networks with same aggregated bandwidth} 
-\label{tab: fix-bandwidth}
-\begin{tabular}{l c c c} 
-\toprule % Top horizontal line
-\toprule
-&\multicolumn{3}{c}{Bandwidth } \\
-\cmidrule(l){2-4}
-Dimension  & per-link & aggregate &\\ % Column names row
-\midrule % In-table horizontal line
-3D      & 4.67GiB/s  & 28GiB/s  & \\  % Content row 1
-\midrule % In-table horizontal line
-5D    & 2.8GiB/s  & 28GiB/s  &\\ % Content row 1
-\midrule % In-table horizontal line
-7D    & 2GiB/s & 28GiB/s & \\
-\midrule
-\bottomrule % Bottom horizontal line
-\end{tabular}
-\end{center}
-\end{table}
-
-In the fixed aggregate bandwidth experiment, with configuration in Table~\ref{tab: fix-bandwidth} and results shown in Figure~\ref{fig: bandwidth-time-box}, the results are more nuanced.
-At least for these applications, the benefits from increased dimensionality do not outweigh the reduction in per-link bandwidth.
-MultiGrid does see benefits in the increased dimensionality, particularly with respect to outliers, but has a similar median message time.
-CrystalRouter reduces performance as the dimensionality increases, but only by a small degree.
-AMG remains roughly the same, minus a dip in the maximum message time that could be due to rank-to-node mapping changes.
-
-\section{Interference Analysis}
-\label{sec:interference}
-
-\textbf{\emph{Intra-job interference}} refers to the network contention between ranks within each application.
-\textbf{\emph{Inter-job interference}} is introduced by concurrently running jobs that adjacent to each other sharing network resources.
-Communication variability due to such interference can cause application performance degradation.
-
-In this section we will study both kinds of interference in the context of torus network. 
-Since application's communication pattern won't change with its scale, we use AMG with 216 ranks, CrystalRouter with 100 ranks, and MultiGrid 125 ranks.
-And in order to accommodate three applications running concurrently with different allocations, we simulate a 2K-node 3D torus ($16^{2} \times 8$) network for the experiments.
-
-\subsection{Intra-Job Interference Analysis}
-\label{sec: introjob}
-
-We design two sets of experiments to study the intra-job interference of each application.
-In the first, shown in Figure \ref{fig:shapestudy}, we assign each application with allocations in \emph{three different shapes}.
-In the second, shown in Figure \ref{fig:diffmap}, we study the intra-job interference by using different mapping strategies for each application with given allocation shape.
-
-In the allocation shapes experiment, we select three possible shapes can can commonly seen on the 3D torus network.
-They are 3D balanced-cube, 3D unbalanced-cube, 2D mesh, as shown in Figure \ref{fig:alloc-shapes}.
-
-3D balanced-cube, shown as red in Figure \ref{fig:cont_sub1}, can guarantee the minimum \emph{average pair-wise distance} within the allocation.
-Some research studies like \cite{leung,abhinav-sc13} indicate that compact allocation can guarantee jobs with better performance.
-They use a variety of metrics to evaluate the compactness of the allocation, such as \emph{average pair-wise distance}, \emph{diameter} and \emph{contiguity}.
-In this work, we select 3D balanced-cube as the most compact allocation on a 3D torus network.
-
-3D unbalanced-cube, shown as green in Figure \ref{fig:cont_sub1}, is a rectangular prism, which is the possible allocation shape on the systems with asymmetric networks.
-For example, Cray XE6/XK7 systems are deployed 3D tori with Gemini routers.
-The network connections in the \emph{y}-direction have only half the bandwidth of the cables used in the \emph{x} and \emph{z} directions.
-In order to take advantage of the faster links in the \emph{x} and \emph{z} directions, job allocation starts from X-Z plane, which leads to a rectangular prism shaped allocation \cite{RF}.
-
-2D mesh, shown as blue in Figure \ref{fig:cont_sub1}, can be cut out from a single layer of the 3D torus.
-2D mesh is a very common allocation shape in torus network for both contiguous and non-contiguous allocation strategies.
-For example, Cray Application Level Placement Scheduler (ALPS) indexes the nodes in torus network into a list and makes job allocation by simply going through the list \cite{carl-cug}.
-When the list is obtained by sorting the nodes based on their spacial coordinates in the torus, allocation made off from this list will be 2D mesh.
-The IBM Blue/Gene Q supercomputer Mira at Argonne Leadership Computing Facility also allow its allocation partition configured into mesh \cite{zhou-ipdps}.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/allocation}
-        \caption{ }
-        \label{fig:cont_sub1}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit16}
-        \caption{ }
-        \label{fig:noncont_sub1}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit8}
-        \caption{ }
-        \label{fig:noncont_sub2}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit2}
-        \caption{ }
-        \label{fig:noncont_sub3}
-    \end{subfigure}%
-    \caption{(a) Contiguous allocation in three different shapes.
-``Blue" is 2D mesh, ``Green" is 3D-unbalanced cube and ``Red" is 3D-balanced cube.
-(b)-(d) Non-contiguous allocation.
-Each job is represented by a specific color.
-The nodes assigned to different jobs are interleaved, the size of allocation unit are 16, 8 and 2.}
-    \label{fig:alloc-shapes}
-\end{figure*}
-
-AMG with 3D ``Nearest Neighbor" communication pattern takes less time when being assigned with 3D-unbalanced allocation than 3D-balanced, as shown in Figure \ref{fig:shapstudy-amg}.
-And MultiGrid gets the best performance (shortest data transfer time) when running on 3D-balanced allocation, as shown in Figure \ref{fig:shapstudy-mg}.
-Since MultiGrid's communication pattern is ``Many-to-Many" dominant, 3D-balanced is the most compact allocation with shortest pair-wise distance between nodes, which can reduce the aggregated hops for transferring message among ranks in MultiGrid.
-As CrystalRouter has both dominant local communication and multi-stage global data transfer, the 3D-balance is also the best allocation, but its advantage over 3D-unbalanced is not that obvious as to MultiGrid, as shown in Figure \ref{fig:shapstudy-cr}.
-
-Lots of research studies have designed fancy allocation algorithms to provide application with the most compact allocation\cite{leung,LO}.
-However, as shown in our experiments, providing cubical allocation without considering application's communication pattern won't guarantee the best performance for every application.
-Compact allocation should be provided to application with global data transfer, such as those conforms to ``Many-to-Many" communication pattern.
-Application with dominant communication pattern like ``Nearest Neighbor" won't fully utilize the compact allocation and could get even better performance on a relaxed allocation shape.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/amg_box}
-        \caption{AMG}
-        \label{fig:shapstudy-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/cr_box}
-        \caption{CrystalRouter}
-        \label{fig:shapstudy-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/mg_box}
-        \caption{MultiGrid}
-        \label{fig:shapstudy-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 3D-balanced, 3D-unbalanced and 2D allocation.}
-   \label{fig:shapestudy}
-\end{figure*}
-
-The rank-to-node mapping of parallel applications on HPC system could greatly impact applications' performance.
-However, finding the optimal mapping solution for a given application is out the scope of this work.
-Our experiments aims to show how mapping strategies impact the intra-job interference of application with specific communication pattern.
-
-We provide AMG, CrystalRouter and MultiGrid with 3D-balance allocation and use three mapping strategies to do the rank-to-node mapping.
-The ``Linear" mapping, as we used in the previous experiments, is to map each rank according to the dimensional ordering of compute nodes.
-The ``Cube" mapping assigns ranks into consecutive $2^{3}$ cubes.
-The ``Random" mapping assigns ranks randomly within the allocation.
-
-AMG remains roughly the same when it been mapped by ``Linear" and ``Cube", shown in Figure \ref{fig:diffmap-amg}.
-Although ``Linear" and ``Cube" mapping strategies cause some routing overlap for AMG's communication, both mapping strategies still preserve the locality of AMG's 3D ``Nearest Neighbor" communication pattern.
-The ``Random" mapping totally destroys AMG's communication pattern, and results in intra-job interference among all the ranks in AMG.
-The performance degradation caused by using ``Random" mapping is up to 90\%.
-
-The ``Cube" mapping improves CrystalRouter's performance over the ``Linear" by reducing the pair-wise distance between ranks, shown in Figure \ref{fig:diffmap-cr}.
-The global data transfer in CrystalRouter will takes fewer hops with ``Cube" mapping.
-The ``Random" mapping scatters the ranks within the local neighborhood of CrystalRouter and make their communication less efficient.
-
-The ``Cube" mapping benefits MultiGrid's ``Many-to-Many" communication.
-However, due to the small amount of data transfer among ranks in MultiGrid, ``Cube" mapping doesn't exhibit a drastic advantage.
-``Random" mapping neither cause too much degradation for the same reason, shown in Figure \ref{fig:diffmap-mg}.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/diffmapping/amg}
-        \caption{AMG}
-        \label{fig:diffmap-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/diffmapping/cr}
-        \caption{CrystalRouter}
-        \label{fig:diffmap-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/diffmapping/mg}
-        \caption{MultiGrid}
-        \label{fig:diffmap-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 3D-balanced allocation using  different mapping strategies.}
-   \label{fig:diffmap}
-\end{figure*}
-
-\subsection{Inter-Job Interference Analysis}
-\label{sec: interjob interference study}
-
-Inter-job interference has been identified as one of the major culprits for application's performance variability \cite{abhinav-sc13,skinner,rosenthal}.
-Inter-job interference is a more prominent issue for system adopted non-contiguous allocation policy than system with contiguous allocation.
-Application communication times have been demonstrated to vary from 36\% faster to 69\% slower due to job interference when running on non-contiguous allocation \cite{abhinav-sc13}.
-
-We assign each application with non-contiguous allocation and run them concurrently on the same network.
-The allocation unit belongs to different jobs are interleaved.
-In order to study the impact of different allocation unit size on applications' inter-job interference, we conduct experiments with unit size of 16, 8 and 2, shown respectively in Figure \ref{fig:noncont_sub1}, \ref{fig:noncont_sub1} and \ref{fig:noncont_sub3}.
-Figure \ref{fig:interjobstudy} shows the results of each application data transfer time with different allocation unit size.
-
-The data transfer time of AMG in Figure \ref{fig:interjob-amg} keeps stable between allocation unit size of 16 and 8, since the locality of AMG is 6-rank based.
-When the unit size reduce to 2, AMG suffers prolong data transfer time by about 10\%.
-CrystalRouter is more sensitive to allocation unit size.
-The best unit size would be the one big enough to accommodate its local neighborhood.
-Figure \ref{fig:interjob-cr} shows unit size of 16 and 8 can guarantee the same average data transfer time, while some rank spend more time with allocation unit size 8 than 16.
-When the unit size reduce to 2, the communication become less efficiency and takes about 15\% more times for transferring data.
-
-The data transfer time of MultiGrid with different allocation unit sizes doesn't show obvious variability in Figure \ref{fig:interjob-mg}.
-This is because even big allocation unit size like 16 will still fail to preserve MultiGrid's ``Many-to-Many" pattern.
-The data transfer time is almost doubled when MultiGrid running concurrently with allocation unit size of 16, as shown in Figure \ref{fig:interjob-mg} .
-As the unit size decreases, the data transfer time keep growing, but in a slow way, which means in terms of preserving the locality of MultiGrid, allocation unit of size 16, 8 and 2 serve equally bad.
-
-The unit size in non-contiguous allocation policy should be chosen according to application's communication pattern.
-Inter-job interference is inevitable in non-contiguous based systems, but unit size that big enough to preserve the neighborhood communication of the application will alleviate such interference and improve job performance.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/inter-job/amg}
-        \caption{AMG}
-        \label{fig:interjob-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/inter-job/cr}
-        \caption{CrystalRouter}
-        \label{fig:interjob-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/inter-job/mg}
-        \caption{MultiGrid}
-        \label{fig:interjob-mg}
-    \end{subfigure}%
-   \caption{Inter-job interference study.
-``cont" indicates three applications running side by side concurrently on the same network with contiguous allocation.
-To study the impact of non-contiguous allocation on inter-job interference, applications running concurrently with interleaved allocation of different unit sizes, which are 16-node, 8-node, 2-node.
-}
-   \label{fig:interjobstudy}
-\end{figure*}
-
-\subsection{Result Summary}
-\label{sec:summary}
-
-Based on the above comprehensive study, we can make following observations:
-\begin{itemize}
-    \item The compact allocation may not guarantee the best performance for every application. 
-    \item The applications dominated with Nearest Neighbor communications exhibit relatively stable performance under different allocation shapes, while the applications dominated with Many-to-Many communication exhibit better performance with compact allocation (e.g.
-3D balanced).
-    \item A good rank-to-node mapping strategy can greatly improve application's performance when specific allocation is given.
-    \item An optimal size for allocation units should be determined according to an application's dominant communication pattern.
-In general, a unit size should be large enough to preserve neighboring communication of the application.
-    \item Inter-job interference is inevitable in non-contiguous allocation.
-However, choosing the proper allocation unit size with job's communication pattern awareness can alleviate such negative effect.
-\end{itemize}
-
-\section{Discussion}
-\label{sec:discussion}
-
-The results shown in the previous section provide insights for the design of smart and flexible job allocation.
-By scrutinizing job's communication behavior, we can identify job's dominant communication pattern and pinpoint the ``neighborhood communication".
-With such knowledge about jobs' communication patterns, we can analyze the possible interference between jobs and take precautions to alleviate the negative effect when making allocation decisions.
-
-When an application with dominant communication that is intensive ``Many-to-Many", the scheduler should grant it with compact node allocation and exclusive network provision.
-The compact allocation can guarantee the shortest pair-wise distance between all the ranks, and make the data transfer between ranks take less hops.
-On the other hand, the exclusive network provision will prevent other adjacent applications from sharing network resources, thus eliminate the performance degradation due to interference.
-
-Not every application's preferable resource are compact node allocation and exclusive network provision.
-As we found in our analysis, applications whose dominant communication patterns contain intensive ``neighborhood communication" like ``Nearest Neighbor" don't benefit from compact node allocation and exclusive network provision.
-Applications, such as AMG, can run with non-contiguous node allocation without performance degradation, as long as the allocation unit can accommodate their ``neighborhood communication".
-
-The allocation unit size shouldn't be fixed.
-Instead, the scheduler should choose a proper unit size based on the scale of ``neighborhood communication" of each job.
-The best unit size should be neither too big nor too small, jut perfectly fit for the size of that ``neighborhood".
-Big unit size won't be fully utilized and also cause fragmentation.
-Small unit size won't be able to accommodate the ``neighborhood communication" and makes the intra-job communication less efficient.
-
-The advantage of making allocation with consideration about job's communication pattern is obvious.
-First, compared with contiguous allocation policy, the new scheduler outlined above is more flexible, it doesn't require the system to provide a big contiguous partition to accommodate the whole application, instead it only needs to provide a small set of compact nodes that is sufficient for all the "local community" in the application.
-On the other hand, the new scheduler outlined above won't destroy the locality of application's communication pattern.
-Indeed, the small compact node set (allocation unit) provided for the application's  ``local communities" can preserve the communication locality in the maximum extent.
-The design of such a smart scheduler is part of our future work.
-
-\section{Related Work}
-\label{sec:related_work}
-
-There are many tools available for system monitoring and application profiling.
-Tools like TAU (Tuning and Analysis) \cite{tau}, SCALASCA \cite{scala} and mpiP \cite{mpip} can capture application's runtime information about application's communication and computation in event traces.
-However, recognizing the patterns about application's communication from those traces requires substantial effort.
-There are a number of studies on the recognition and characterization of parallel application communication patterns.
-Oak Ridge National Laboratory has its ongoing project about developing a tool set named Oxbow, which can characterize the computation and communication behavior of scientific applications and benchmarks \cite{oxbow}.
-In a recent work \cite{roth}, the authors demonstrate a new approach to automatically characterizing parallel applications' communication behaviors.
-
-%\textbf{P2: communication characteristics}\\
-
-Many research efforts have been conducted to characterize scientific applications.
-For instance, a DOE project aims at the identification of the computational characteristics of the DOE MiniApps developed at various exascale co-design centers \cite{designforwardwebpage}.
-In this project, the communication patterns of several DOE full applications and associated mini-applications are studied to provide a more complete snapshot of the DOE workload.
-A joint project named CORAL from Oak Ridge, Argonne and Livermore provide a series of benchmarks to represent DOE workloads and technical requirements \cite{coral}.
-The CORAL project includes scalable science benchmarks, throughput benchmarks, data centric benchmarks, skeleton benchmarks and Micro benchmarks.
-
-%\textbf{P3: Interference between concurrently running jobs}\\
-The interference between concurrently running jobs on HPC system have been identified as major culprit for job's performance variability.
-Abhinav et al.
-found that concurrently running applications on HPC can cause interference to each other, and cause communication time varied from 36\% faster to 69\% slower.
-Such interference could come from Operating System noise, shape of the allocated partition , or other running jobs that sharing network resources \cite{abhinav-sc13}.
-Skinner et al.
-found that  there is a 2-3$\times$ slowdown in MPI$\textunderscore$Allreduce due to network contention from other jobs \cite{skinner}.
-Rosenthal et al.
-found increasing network bandwidth provide limited benefits to a number of applications.
-The applications that send mostly small messages or large messages asynchronously are not bandwidth bounded, hence, benefit only slightly or not at all from increased bandwidth \cite{rosenthal}.
-
-%\textbf{P4: Job allocation(little bit)}\\
-Several research studies focus on optimizing job allocation on HPC system to alleviate the interference between concurrently running jobs.
-Hoefler et al.
-proposed to use performance modeling techniques to analyze factors that impact the performance of parallel scientific applications \cite{hoefler-modeling}.
-However, as the scale of HPC continue to grow, the  interference of concurrently running jobs is getting worse, which is hard to be quantified by performance profiling tools alone.
-Bogdan et al.
-provide a set of guidelines of how to configure a network with Dragonfly topology for workload with Nearest Neighbor communication pattern \cite{Bogdan-hpdc14}.
-Dong et al.
-have developed simple benchmarks that conforms to four different communication patterns, namely ping-pong, nearest neighbor, broadcast and all$\textunderscore$reduce, to demonstrate the effectiveness of this highly parallel 5D torus network \cite{Dong-SC11}.
-
-Our work is different from all these works in the following ways.
-First, we focus on the dominant communication patterns rather than any specific application.
-We believe this can provide a guideline for other research work.
-Secondly, we explore the intra- and inter-job interference between concurrently running jobs, while similar work such as \cite{abhinav-sc13} only focuses on the single application's performance degradation due to network contention.
-Finally, we analyze the impact of different allocation strategies to job's communication behaviors.
-We further identify the optimal allocation strategy for each application with a specific dominant communication pattern.
-Based on our study, we claim that further scheduling should take job's communication pattern into consideration for its allocation decision making.
-
 \section{Conclusions}
@@
 Rather than using pre-defined partitions and non-contiguous allocation, future HPC systems should assign each job with preferable resources based on job's communication pattern.
-
-\section*{Acknowledgment}
-\label{sec: ack}
-The work at Illinois Institute of Technology is supported in part by U.S.
-National Science Foundation grants CNS-1320125 and CCF-1422009.
-This work is also supported by the U.S.
-Department of Energy, Office of Science, Advanced Scientific Computing Research, under Contract DE-AC02-06CH11357.
-
-\bibliographystyle{IEEEtran}
-\bibliography{./reference}
-
-\vspace{5\baselineskip}
-
-\begin{framed}
-The submitted manuscript has been created by UChicago Argonne, LLC, Operator of Argonne National Laboratory ("Argonne").
-Argonne, a U.S.
-Department of Energy Office of Science laboratory, is operated under Contract No.
-DE-AC02-06CH11357.
-The U.S.
-Government retains for itself, and others acting on its behalf, a paid-up nonexclusive, irrevocable worldwide license in said article to reproduce, prepare derivative works, distribute copies to the public, and perform publicly and display publicly, by or on behalf of the Government.
-\end{framed}
-
-\end{document}

### 2016-07-07 config_study.tex
split into separate tex files for each section

--- 2015-10-17 submission.tex
@@
-\documentclass[conference]{IEEEtran}
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
-\else
-
-\fi
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-\usepackage{algorithm}
-%\usepackage{algorithm2e}
-\usepackage{algpseudocode}
-\usepackage{booktabs}
-\usepackage{flushend}
-\usepackage{url}
-\usepackage{xcolor}
-\usepackage{amsmath}
-\usepackage{framed}
-\usepackage{textcomp}
-%\usepackage{subfig}
-\usepackage{graphicx}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{soul}
-\usepackage{verbatim}
-%\usepackage[labelfont=bf]{caption}
-\usepackage[font=bf,labelsep=space]{caption}
-
-\newcommand{\TODO}[1]{\hl{TODO: #1}}
-\newcommand{\NOTE}[1]{\hl{NOTE: #1}}
-\newcommand{\changed}[1]{\textcolor{red}{#1}}
-
-\begin{document}
-%
-% paper title
-% can use linebreaks \\ within to get better formatting as desired
-\title{Study of Intra- and Inter-Job Interference with Different Job Allocations}
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
-\author{
-
-\IEEEauthorblockN{Xu Yang\IEEEauthorrefmark{1}, John Jenkins\IEEEauthorrefmark{2}, Misbah Mubarak\IEEEauthorrefmark{2}, Robert B.
-Ross\IEEEauthorrefmark{2}, Zhou Zhou\IEEEauthorrefmark{1}, Zhiling Lan\IEEEauthorrefmark{1}}
-
-\IEEEauthorblockA{\IEEEauthorrefmark{1}Department of Computer Science,
-Illinois Institute of Technology,
-Chicago, Illinois, USA 60616\\
-\{xyang56, zzhou\}@hawk.iit.edu, lan@iit.edu}
-
-\IEEEauthorblockA{\IEEEauthorrefmark{2}Mathematics and Computer Science Division, Argonne National Laboratory,
-Argonne, IL, USA 60439\\
-\{jenkins,rross\}@mcs.anl.gov, mmubarak@anl.gov}
-}
-
-% make the title area
-\maketitle
-
-\begin{abstract}
-
-Network contention between concurrently running jobs on HPC systems is a primary cause of performance variability.
-Optimizing job allocation and avoiding network sharing are hence crucial to alleviate the potential degradation in performance.
-In order to do these effectively, an understanding of the interaction between applications, their communication patterns, and the system architecture are required.
-In this work, we choose three representative HPC applications from DOE Design Forward Project and conduct detailed simulations of high-performance network topologies to analyze both intra- and inter-job interference.
-By scrutinizing the communication behaviors of these applications, we identify relationships between these behaviors and the possible interference introduced by different job allocation scenarios.
-Our comprehensive analyses in this work illuminate a path towards job-communication-pattern-aware HPC resource allocation.
-
-\end{abstract}
-
-\IEEEpeerreviewmaketitle
-
-\section{Introduction} 
-\label{sec: intro}
-
-The scale of supercomputers continually grows at a breakneck pace to accommodate the computing power requirements for scientific research.
-Supercomputers have tens of thousands of nodes and serve as an irreplaceable research vehicle for scientific problems with increasing size and complexity.
-Supercomputers are usually employed as a shared resource to accommodate many parallel applications (jobs) running concurrently.
-These parallel jobs share the system infrastructure such as network and I/O bandwidth, and inevitably there is contention over these shared resources.
-As supercomputers continue to evolve, these shared resources are increasingly the bottleneck for performance.
-
-\begin{figure}[h!]
-    \centering
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/goodallocation}
-        \caption{Contiguous}
-        \label{fig:overview_sub1}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/badallocation}
-        \caption{Non-contiguous}
-        \label{fig:overview_sub2}
-    \end{subfigure}%
-   \caption{Multiple jobs running concurrently with different allocations.
-Each job is represented by a specific color.
-a) shows the effect of contiguous allocation, which reduce the inter-job interference.
-b) shows non-contiguous allocation, which may introduce both intra and inter-job interferences.
-}
-   \label{fig:overview}
-\end{figure}
-
-When submitting jobs to HPC systems, users request resources in terms of node counts and expected runtime.
-The batch scheduler is then responsible for dispatching the submitted jobs to the system.
-Typically, multiple jobs run concurrently, resulting in the shared use of resources, particularly network and storage.
-A prominent problem with this approach is network contention among concurrently running jobs.
-If the network resources of concurrently running jobs aren't isolated, the resulting sharing can cause communication variability that results in performance degradation for some or all such jobs \cite{abhinav-sc13}.
-This delay can propagate into the queueing time of the following submitted jobs, thus leading to low system throughput and utilization \cite{jose-ipdps15}.
-This adverse effect of network sharing can be mitigated by providing jobs with isolated allocations and exclusive network resources.
-
-On the widely used torus-connected HPC systems \cite{bgq,tofu,titan}, two allocation strategies are commonly used.
-The \emph{contiguous allocation} strategy assigns to each job a compact and contiguous set of computing nodes, as shown in Figure \ref{fig:overview_sub1}.
-The partition-based allocation approach used in Blue Gene series systems is an example of such a strategy \cite{bgloverview}.
-The contiguous strategy favors application performance through isolated networking within a partition and the locality that implies.
-However, this strategy can cause both internal fragmentation (when more nodes are allocated to a job than it requests) and external fragmentation (when sufficient nodes are available for a request, but they can not be allocated contiguously), therefore leading to lower system utilization than is otherwise possible.
-On the other side of the coin, the \emph{non-contiguous allocation} strategy, used by the Cray XT/XE series \cite{carl-cug}, assigns free nodes to jobs regardless of contiguity, though of course efforts are made to maximize locality.
-Figure~\ref{fig:overview_sub2} shows an example non-contiguous allocation.
-While eliminating internal and external fragmentation as seen in contiguous allocation systems, in return non-contiguous allocations introduce contention between jobs due to the interleaving of job nodes.
-The non-contiguous node allocation can significantly reduce job performance, especially for communication-intensive ones \cite{abhinav-sc13}.
-
-We envision that future HPC schedulers will adopt a flexible job allocation mechanism which combines the best of both contiguous and non-contiguous allocation strategies.
-Such a flexible mechanism would take shared resource needs of jobs into account when making allocation decisions (e.g., network).
-With knowledge and analysis of job communication patterns, it can be identified which jobs require network isolation and locality, and to what degree.
-Then, rather than allocating each job in a ``know-nothing'' manner, one may specialize allocation so that, for example, only the jobs with stringent network needs are given compact, isolated allocations, resulting in maximized utilization and minimized perceivable resource contention effects.
-
-In this work, we focus on an in-depth analysis of intra- and inter-job communication interference with different job allocation strategies on torus-connected HPC systems.
-Torus-based networks are used on six of the top 10 supercomputers on the June 2015 Top500 lists~\cite{top500}.
-The current generation of IBM Blue Gene/Q (BG/Q) supercomputers, such as Mira at Argonne Nation Laboratory and Sequoia at Lawrence Livermore National Laboratory, has its nodes connected in a 5D torus network \cite{bgq}.
-The K computer from Japan uses the ``Tofu'' system, which is a 6D mesh/torus topology \cite{tofu}.
-Titan, a Cray XK7 supercomputer located at the Oak Ridge Leadership Computing Facility (OLCF), has nodes connected in a 3D torus within the compute partition \cite{titan}.
-Although our analyses are based on torus networks, the ideas conveyed in this work is applicable to networks with different topologies.
-
-We selected three signature applications from the DOE Design Forward Project~\cite{designforwardwebpage} as examples to conduct detailed study about their communication patterns.
-We use a sophisticated simulation toolkit named CODES (standing for Co-Design of Multi-layer Exascale Storage Architectures)~\cite{Jason-2011} as a research vehicle to evaluate the performance of these applications with various allocations in a controlled environment.
-We then analyze the intra- and inter-job interference by simulating these applications running exclusively and concurrently with different allocations.
-We believe the insights presented in this work can be very useful for the design of future HPC batch job schedulers and resource managers.
-
-%\NOTE{JJ removed the list of contributions, because they are exactly the same as the preceding paragraph.}
-\begin{comment}
-The observations of this work are the following:
-
-\begin{itemize}
-    \item We study the communication behaviors and identify the communication patterns of three representative HPC applications.
-    \item We conduct detailed analysis about the intra- and inter-job interference by running three applications with different allocation strategies.
-And we find the relation between application's communication pattern and the cause of interference.
-    \item We show a novel and flexible way to make allocation with awareness of job's communication pattern, which  is every effective to alleviate the interference between concurrently running jobs.
-\TODO{do we actually do this, or provide the underpinnings for such a method?}
-\end{itemize}
-
-We believe the finds and insights presented in this work would be very useful for the design of future HPC batch job scheduler and resource management module.
-
-\end{comment}
-
-The rest of this paper is organized as follows.
-Section~\ref{sec:application study} describes the three representative applications from DOE Design Forward Project for our study.
-Section~\ref{sec:codes} talks about the use of CODES as research vehicle for our work.
-Section~\ref{sec:config study} shows the performance analysis of three applications on different torus networks.
-Section~\ref{sec:interference} provides detailed analysis about the intra- and inter-job interference between three applications on torus network with different allocation strategies.
-Section~\ref{sec:discussion} introduces a path toward communication-pattern aware allocation strategies, given the results of our analysis.
-Section~\ref{sec:related_work} discusses related work.
-Finally, the conclusion is presented in Section~\ref{sec:conclusion}.
-
-\section{Application Study}
-\label{sec:application study}
-
-%A parallel application usually conforms to a combination of several basic communication patterns\cite{roth}.
-At its different execution phases, the application's communication behavior may follow different basic patterns respectively.
-\textcolor{blue}{When we look into the data flow during application execution, most parallel applications start with broadcast operation to distribute the data from root process to other processes\textcolor{red}{ref needed here}, followed by a series of computation and communication that conforms to certain pattern, which is usually the dominant part of application's execution.
-Before the application come to completion, all the working process will return their results to the root directly or hierarchically.} Parallel application's communication pattern usually involves lots of factors, such as communication intensity, operation-to-operation dependencies and the ``critical path" in its communication topology graph, etc.
-In this work, we focus on the locality of application's rank-to-rank communication topology graph.
-
-A parallel application usually conforms to a combination of several basic communication patterns \cite{roth}.
-At its different execution phases, the application's communication behavior may follow different basic patterns respectively.
-There are many profiling tools available to capture information regarding communication patterns of parallel applications~\cite{tau,mpip,scala,sst,oxbow}.
-They can provide information such as the percentage of different MPI operations of the applications, communication topology, the amount of data transferred between processes.
-
-There are many applications running on HPC systems, such as Mira \cite{bgq} and Titan \cite{titan}, that are communication intensive and conform to various communication patterns.
-In this work, we select three representative applications from the DOE Design Forward Project.
-Each application exhibits a distinctive communication pattern that is commonly seen in HPC applications.
-We believe that the communication patterns of these applications are representative of a wide array of applications running on leadership-class machines.
-Specifically, we study the Algebraic MultiGrid Solver (AMG), Geometric MultiGrid (MultiGrid) and CrystalRouter MiniApps.
-The communication pattern figures we present in Section~\ref{sec:amg}, \ref{sec:crystalrouter}, and \ref{sec:multigrid} are generated with the IPM~\cite{ipm} data gathered from publicly available traces~\cite{designforwardwebpage}.
-
-% MiniApps are reduced proxy applications that encapsulate the salient performance of larger full size applications \cite{miniapp}.
-Since we only focus on the dominant communication pattern of parallel applications, these MiniApps are perfect candidates for analysis the application's performance on different allocations or on new network architectures.
-We choose three MiniApps, namely, Algebraic Multigrid Solver(AMG), Geometric MultiGrid(MultiGrid) and CrystalRouter.
-Their representative communication patterns that are commonly seen in the HPC system workload.
-The communication pattern figures we present here are generated with the IPM \cite{ipm} data from \cite{designforwardwebpage}.
-
-\subsection{AMG}
-\label{sec:amg}
-
-\begin{figure}[t!]
-    \centering
-    \begin{subfigure}[t]{0.25\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/appstudy/amg/amg_pip}
-        \caption{}
-        \label{fig:amg-communication-topology}
-    \end{subfigure}%
-    ~
-    \begin{subfigure}[t]{0.24\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/appstudy/amg/amg_data_transfer}
-        \caption{}
-        \label{fig:amg-data-trans}
-    \end{subfigure}
-    \caption{AMG.
-(a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in AMG.
-(b) The amount of data sent/received by each rank in AMG.
-Y-axis labels the amount of data transferred in MB, x-axis labels 216 ranks of AMG.
-}
-\end{figure}
-
-The Algebraic MultiGrid Solver, or AMG, is a parallel algebraic multi-grid solver for linear systems arising from problems on unstructured mesh physics packages.
-It has been derived directly from the BoomerAMG solver that is being developed in the Center for Applied Scientific Computing (CASC) at LLNL \cite{amg}.
-The dominant communication pattern is regional communication with decreasing message size for different parts of the multi-grid v-cycle.
-
-Figure~\ref{fig:amg-communication-topology} shows the communication topology of a small scale AMG execution with 216 MPI Ranks.
-Note that the dominant communication pattern of the application doesn't change with scale.
-We can make the observation that AMG's dominant communication pattern is 3D Nearest Neighbor, each rank has intensive communication with up to six neighbors, depending on the boundaries of the ranks.
-Applications with similar pattern include PARTISN \cite{partisn} and SNAP \cite{snap}.
-
-Figure~\ref{fig:amg-data-trans} shows the amount of data each rank transferred in the example execution.
-The data transferring of each rank is between 7MB to 16MB, corresponding to whether the rank is on a boundary or not.
-
-\subsection{Crystal Router}
-\label{sec:crystalrouter}
-
-\begin{figure}[t!]
-    \centering
-    \begin{subfigure}[t]{0.25\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/appstudy/cr/cr_pip}
-        \caption{Communication Topology}
-        \label{fig:cr-communication-topology}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[t]{0.22\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/appstudy/cr/cr_data_transfer}
-        \caption{Data Volume}
-        \label{fig:cr-data-trans}
-    \end{subfigure}
-    \caption{CrystalRouter.
-(a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in CrystalRouter.
-(b) The amount of data sent/received by each rank in CrystalRouter.
-Y-axis labels the amount of data transferred in MB, x-axis labels 100 ranks of CrystalRouter.
-}
-\end{figure}
-
-The second MiniApp covered is CrystalRouter, the extracted communication kernel of the full application Nek5000~\cite{crystalrouter}, which is a spectral element CFD application developed at Argonne National Laboratory \cite{nek5000}.
-It features spectral element multi-grid solvers coupled with a highly scalable, parallel coarse-grid solver that is widely used for projects including ocean current modeling, thermal hydraulics of reactor cores, and spatiotemporal chaos.
-CrystalRouter demonstrates the ``many-to-many'' communication pattern through a scalable multi-stage communication process.
-The way CrystalRouter works is nodes compute for a while, then synchronize and communicate, continually alternating between these two types of activities.
-
-The collective communication in CrystalRouter utilizes a recursive doubling approach.
-Ranks in CrystalRouter conform to a $n$-dimensional hypercube and recursively split into ($n$-1)-dimensional hypercubes, with communication occurring along the splitting plane.
-The pattern of this communication can be found in Figure~\ref{fig:cr-communication-topology}.
-By nature of the logarithmic splitting process, a substantial portion of the communication occurs in small neighborhoods of ranks.
-CrystalRouter represents a group of applications whose dominant communication is a hybrid of multi-stage local and hierarchical global communication, and shares similarities with most MPI collective communication implementations.
-
-Figure~\ref{fig:cr-data-trans} shows the amount of data each rank transferred in CrystalRouter, which ranges between 20MB to 36MB.
-This is due to the rank counts not matching one-for-one with points in the hypercube.
-
-\subsection{MultiGrid}
-\label{sec:multigrid}
-
-\begin{figure}[t!]
-    \centering
-    \begin{subfigure}[t]{0.25\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/appstudy/mg/mg_pip}
-        \caption{Communication Topology}
-        \label{fig:mg-communication-topology}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[t]{0.22\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/appstudy/mg/mg_data_transfer}
-        \caption{Data Volume}
-        \label{fig:mg-data-trans}
-    \end{subfigure}
-    \caption{MultiGrid.
-(a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in MultiGrid.
-(b) The amount of data sent/received by each rank in MultiGrid.
-Y-axis labels the amount of data transferred in MB, x-axis label 125 ranks of MultiGrid.
-}
-\end{figure}
-
-MultiGrid is geometric multi-grid v-cycle from the production elliptic solver BoxLib, a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes \cite{boxlib}.
-MultiGrid conforms to many-to-many communication pattern with decreasing message size and collectives for different parts of the multi-grid v-cycle.
-It is widely used for structured grid physics packages.
-
-Figure~\ref{fig:mg-communication-topology} shows the communication topology of MultiGrid with 125 ranks.
-We can see intensive communication along the diagonal that resembles nearest neighbor communication, similar to AMG.
-However, the communication topology leads to a greater ``spread'' of communication across the set of ranks, challenging the maximization of communication locality with respect to rank.
-In this sense, it can be considered a ``many-to-many'' pattern.
-Applications with such similar dominant communication pattern are like FillBoundary, another PDE solver code in \cite{boxlib}.
-
-Figure~\ref{fig:mg-data-trans} shows the amount of data transferred between ranks in MultiGrid. For this miniapp, The data transferred per rank is relatively small and stable, between 2.8MB to 2.9MB.
-
-\section{Research Vehicle}
-\label{sec:codes}
-
-It is difficult to accurately and flexibly experiment with concurrently running jobs in an HPC context.
-One reason is that the allocation strategy used on production machine is part of the system software, which can not be changed by users.
-Even the system administrator is not authorized to make that change.
-Another reason is that it is unrealistic to reserve the system exclusively to run the same job with desired allocation without interference and then compare the results with those in the presence of interference.
-Therefore, we resort to simulation for this work.
-
-A simulation toolkit named CODES enables the exploration of simulating different HPC networks with high fidelity \cite{Jason-2011}\cite{mubarak-sc2012}.
-CODES is built on top of Rensselaer Optimistic Simulation System (ROSS) parallel discrete-event simulator, which is capable of processing billions of events per second on leadership-class supercomputers \cite{ross}.
-CODES support both torus and dragonfly network with high fidelity flit-level simulation.
-CODES has this network workload component that is capable of conducting trace-driven simulations.
-It can take real MPI application traces generated by SST DUMPI \cite{sst} to drive CODES network models.
-In this work, we focus on an in-depth analysis of intra- and inter-job communication interferences with  different job allocations on torus-connected HPC systems.
-Torus networks have been extensively used in the current generation of supercomputers because of their linear scaling on per-node cost and competitive communication performance \cite{mubarak-sc2012}\cite{zhou-ipdps}.
-
 \section{Configuration Study}
@@
 AMG remains roughly the same, minus a dip in the maximum message time that could be due to rank-to-node mapping changes.
-
-\section{Interference Analysis}
-\label{sec:interference}
-
-\textbf{\emph{Intra-job interference}} refers to the network contention between ranks within each application.
-\textbf{\emph{Inter-job interference}} is introduced by concurrently running jobs that adjacent to each other sharing network resources.
-Communication variability due to such interference can cause application performance degradation.
-
-In this section we will study both kinds of interference in the context of torus network. 
-Since application's communication pattern won't change with its scale, we use AMG with 216 ranks, CrystalRouter with 100 ranks, and MultiGrid 125 ranks.
-And in order to accommodate three applications running concurrently with different allocations, we simulate a 2K-node 3D torus ($16^{2} \times 8$) network for the experiments.
-
-\subsection{Intra-Job Interference Analysis}
-\label{sec: introjob}
-
-We design two sets of experiments to study the intra-job interference of each application.
-In the first, shown in Figure \ref{fig:shapestudy}, we assign each application with allocations in \emph{three different shapes}.
-In the second, shown in Figure \ref{fig:diffmap}, we study the intra-job interference by using different mapping strategies for each application with given allocation shape.
-
-In the allocation shapes experiment, we select three possible shapes can can commonly seen on the 3D torus network.
-They are 3D balanced-cube, 3D unbalanced-cube, 2D mesh, as shown in Figure \ref{fig:alloc-shapes}.
-
-3D balanced-cube, shown as red in Figure \ref{fig:cont_sub1}, can guarantee the minimum \emph{average pair-wise distance} within the allocation.
-Some research studies like \cite{leung,abhinav-sc13} indicate that compact allocation can guarantee jobs with better performance.
-They use a variety of metrics to evaluate the compactness of the allocation, such as \emph{average pair-wise distance}, \emph{diameter} and \emph{contiguity}.
-In this work, we select 3D balanced-cube as the most compact allocation on a 3D torus network.
-
-3D unbalanced-cube, shown as green in Figure \ref{fig:cont_sub1}, is a rectangular prism, which is the possible allocation shape on the systems with asymmetric networks.
-For example, Cray XE6/XK7 systems are deployed 3D tori with Gemini routers.
-The network connections in the \emph{y}-direction have only half the bandwidth of the cables used in the \emph{x} and \emph{z} directions.
-In order to take advantage of the faster links in the \emph{x} and \emph{z} directions, job allocation starts from X-Z plane, which leads to a rectangular prism shaped allocation \cite{RF}.
-
-2D mesh, shown as blue in Figure \ref{fig:cont_sub1}, can be cut out from a single layer of the 3D torus.
-2D mesh is a very common allocation shape in torus network for both contiguous and non-contiguous allocation strategies.
-For example, Cray Application Level Placement Scheduler (ALPS) indexes the nodes in torus network into a list and makes job allocation by simply going through the list \cite{carl-cug}.
-When the list is obtained by sorting the nodes based on their spacial coordinates in the torus, allocation made off from this list will be 2D mesh.
-The IBM Blue/Gene Q supercomputer Mira at Argonne Leadership Computing Facility also allow its allocation partition configured into mesh \cite{zhou-ipdps}.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/allocation}
-        \caption{ }
-        \label{fig:cont_sub1}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit16}
-        \caption{ }
-        \label{fig:noncont_sub1}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit8}
-        \caption{ }
-        \label{fig:noncont_sub2}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit2}
-        \caption{ }
-        \label{fig:noncont_sub3}
-    \end{subfigure}%
-    \caption{(a) Contiguous allocation in three different shapes.
-``Blue" is 2D mesh, ``Green" is 3D-unbalanced cube and ``Red" is 3D-balanced cube.
-(b)-(d) Non-contiguous allocation.
-Each job is represented by a specific color.
-The nodes assigned to different jobs are interleaved, the size of allocation unit are 16, 8 and 2.}
-    \label{fig:alloc-shapes}
-\end{figure*}
-
-AMG with 3D ``Nearest Neighbor" communication pattern takes less time when being assigned with 3D-unbalanced allocation than 3D-balanced, as shown in Figure \ref{fig:shapstudy-amg}.
-And MultiGrid gets the best performance (shortest data transfer time) when running on 3D-balanced allocation, as shown in Figure \ref{fig:shapstudy-mg}.
-Since MultiGrid's communication pattern is ``Many-to-Many" dominant, 3D-balanced is the most compact allocation with shortest pair-wise distance between nodes, which can reduce the aggregated hops for transferring message among ranks in MultiGrid.
-As CrystalRouter has both dominant local communication and multi-stage global data transfer, the 3D-balance is also the best allocation, but its advantage over 3D-unbalanced is not that obvious as to MultiGrid, as shown in Figure \ref{fig:shapstudy-cr}.
-
-Lots of research studies have designed fancy allocation algorithms to provide application with the most compact allocation\cite{leung,LO}.
-However, as shown in our experiments, providing cubical allocation without considering application's communication pattern won't guarantee the best performance for every application.
-Compact allocation should be provided to application with global data transfer, such as those conforms to ``Many-to-Many" communication pattern.
-Application with dominant communication pattern like ``Nearest Neighbor" won't fully utilize the compact allocation and could get even better performance on a relaxed allocation shape.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/amg_box}
-        \caption{AMG}
-        \label{fig:shapstudy-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/cr_box}
-        \caption{CrystalRouter}
-        \label{fig:shapstudy-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/mg_box}
-        \caption{MultiGrid}
-        \label{fig:shapstudy-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 3D-balanced, 3D-unbalanced and 2D allocation.}
-   \label{fig:shapestudy}
-\end{figure*}
-
-The rank-to-node mapping of parallel applications on HPC system could greatly impact applications' performance.
-However, finding the optimal mapping solution for a given application is out the scope of this work.
-Our experiments aims to show how mapping strategies impact the intra-job interference of application with specific communication pattern.
-
-We provide AMG, CrystalRouter and MultiGrid with 3D-balance allocation and use three mapping strategies to do the rank-to-node mapping.
-The ``Linear" mapping, as we used in the previous experiments, is to map each rank according to the dimensional ordering of compute nodes.
-The ``Cube" mapping assigns ranks into consecutive $2^{3}$ cubes.
-The ``Random" mapping assigns ranks randomly within the allocation.
-
-AMG remains roughly the same when it been mapped by ``Linear" and ``Cube", shown in Figure \ref{fig:diffmap-amg}.
-Although ``Linear" and ``Cube" mapping strategies cause some routing overlap for AMG's communication, both mapping strategies still preserve the locality of AMG's 3D ``Nearest Neighbor" communication pattern.
-The ``Random" mapping totally destroys AMG's communication pattern, and results in intra-job interference among all the ranks in AMG.
-The performance degradation caused by using ``Random" mapping is up to 90\%.
-
-The ``Cube" mapping improves CrystalRouter's performance over the ``Linear" by reducing the pair-wise distance between ranks, shown in Figure \ref{fig:diffmap-cr}.
-The global data transfer in CrystalRouter will takes fewer hops with ``Cube" mapping.
-The ``Random" mapping scatters the ranks within the local neighborhood of CrystalRouter and make their communication less efficient.
-
-The ``Cube" mapping benefits MultiGrid's ``Many-to-Many" communication.
-However, due to the small amount of data transfer among ranks in MultiGrid, ``Cube" mapping doesn't exhibit a drastic advantage.
-``Random" mapping neither cause too much degradation for the same reason, shown in Figure \ref{fig:diffmap-mg}.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/diffmapping/amg}
-        \caption{AMG}
-        \label{fig:diffmap-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/diffmapping/cr}
-        \caption{CrystalRouter}
-        \label{fig:diffmap-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/diffmapping/mg}
-        \caption{MultiGrid}
-        \label{fig:diffmap-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 3D-balanced allocation using  different mapping strategies.}
-   \label{fig:diffmap}
-\end{figure*}
-
-\subsection{Inter-Job Interference Analysis}
-\label{sec: interjob interference study}
-
-Inter-job interference has been identified as one of the major culprits for application's performance variability \cite{abhinav-sc13,skinner,rosenthal}.
-Inter-job interference is a more prominent issue for system adopted non-contiguous allocation policy than system with contiguous allocation.
-Application communication times have been demonstrated to vary from 36\% faster to 69\% slower due to job interference when running on non-contiguous allocation \cite{abhinav-sc13}.
-
-We assign each application with non-contiguous allocation and run them concurrently on the same network.
-The allocation unit belongs to different jobs are interleaved.
-In order to study the impact of different allocation unit size on applications' inter-job interference, we conduct experiments with unit size of 16, 8 and 2, shown respectively in Figure \ref{fig:noncont_sub1}, \ref{fig:noncont_sub1} and \ref{fig:noncont_sub3}.
-Figure \ref{fig:interjobstudy} shows the results of each application data transfer time with different allocation unit size.
-
-The data transfer time of AMG in Figure \ref{fig:interjob-amg} keeps stable between allocation unit size of 16 and 8, since the locality of AMG is 6-rank based.
-When the unit size reduce to 2, AMG suffers prolong data transfer time by about 10\%.
-CrystalRouter is more sensitive to allocation unit size.
-The best unit size would be the one big enough to accommodate its local neighborhood.
-Figure \ref{fig:interjob-cr} shows unit size of 16 and 8 can guarantee the same average data transfer time, while some rank spend more time with allocation unit size 8 than 16.
-When the unit size reduce to 2, the communication become less efficiency and takes about 15\% more times for transferring data.
-
-The data transfer time of MultiGrid with different allocation unit sizes doesn't show obvious variability in Figure \ref{fig:interjob-mg}.
-This is because even big allocation unit size like 16 will still fail to preserve MultiGrid's ``Many-to-Many" pattern.
-The data transfer time is almost doubled when MultiGrid running concurrently with allocation unit size of 16, as shown in Figure \ref{fig:interjob-mg} .
-As the unit size decreases, the data transfer time keep growing, but in a slow way, which means in terms of preserving the locality of MultiGrid, allocation unit of size 16, 8 and 2 serve equally bad.
-
-The unit size in non-contiguous allocation policy should be chosen according to application's communication pattern.
-Inter-job interference is inevitable in non-contiguous based systems, but unit size that big enough to preserve the neighborhood communication of the application will alleviate such interference and improve job performance.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/inter-job/amg}
-        \caption{AMG}
-        \label{fig:interjob-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/inter-job/cr}
-        \caption{CrystalRouter}
-        \label{fig:interjob-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/inter-job/mg}
-        \caption{MultiGrid}
-        \label{fig:interjob-mg}
-    \end{subfigure}%
-   \caption{Inter-job interference study.
-``cont" indicates three applications running side by side concurrently on the same network with contiguous allocation.
-To study the impact of non-contiguous allocation on inter-job interference, applications running concurrently with interleaved allocation of different unit sizes, which are 16-node, 8-node, 2-node.
-}
-   \label{fig:interjobstudy}
-\end{figure*}
-
-\subsection{Result Summary}
-\label{sec:summary}
-
-Based on the above comprehensive study, we can make following observations:
-\begin{itemize}
-    \item The compact allocation may not guarantee the best performance for every application. 
-    \item The applications dominated with Nearest Neighbor communications exhibit relatively stable performance under different allocation shapes, while the applications dominated with Many-to-Many communication exhibit better performance with compact allocation (e.g.
-3D balanced).
-    \item A good rank-to-node mapping strategy can greatly improve application's performance when specific allocation is given.
-    \item An optimal size for allocation units should be determined according to an application's dominant communication pattern.
-In general, a unit size should be large enough to preserve neighboring communication of the application.
-    \item Inter-job interference is inevitable in non-contiguous allocation.
-However, choosing the proper allocation unit size with job's communication pattern awareness can alleviate such negative effect.
-\end{itemize}
-
-\section{Discussion}
-\label{sec:discussion}
-
-The results shown in the previous section provide insights for the design of smart and flexible job allocation.
-By scrutinizing job's communication behavior, we can identify job's dominant communication pattern and pinpoint the ``neighborhood communication".
-With such knowledge about jobs' communication patterns, we can analyze the possible interference between jobs and take precautions to alleviate the negative effect when making allocation decisions.
-
-When an application with dominant communication that is intensive ``Many-to-Many", the scheduler should grant it with compact node allocation and exclusive network provision.
-The compact allocation can guarantee the shortest pair-wise distance between all the ranks, and make the data transfer between ranks take less hops.
-On the other hand, the exclusive network provision will prevent other adjacent applications from sharing network resources, thus eliminate the performance degradation due to interference.
-
-Not every application's preferable resource are compact node allocation and exclusive network provision.
-As we found in our analysis, applications whose dominant communication patterns contain intensive ``neighborhood communication" like ``Nearest Neighbor" don't benefit from compact node allocation and exclusive network provision.
-Applications, such as AMG, can run with non-contiguous node allocation without performance degradation, as long as the allocation unit can accommodate their ``neighborhood communication".
-
-The allocation unit size shouldn't be fixed.
-Instead, the scheduler should choose a proper unit size based on the scale of ``neighborhood communication" of each job.
-The best unit size should be neither too big nor too small, jut perfectly fit for the size of that ``neighborhood".
-Big unit size won't be fully utilized and also cause fragmentation.
-Small unit size won't be able to accommodate the ``neighborhood communication" and makes the intra-job communication less efficient.
-
-The advantage of making allocation with consideration about job's communication pattern is obvious.
-First, compared with contiguous allocation policy, the new scheduler outlined above is more flexible, it doesn't require the system to provide a big contiguous partition to accommodate the whole application, instead it only needs to provide a small set of compact nodes that is sufficient for all the "local community" in the application.
-On the other hand, the new scheduler outlined above won't destroy the locality of application's communication pattern.
-Indeed, the small compact node set (allocation unit) provided for the application's  ``local communities" can preserve the communication locality in the maximum extent.
-The design of such a smart scheduler is part of our future work.
-
-\section{Related Work}
-\label{sec:related_work}
-
-There are many tools available for system monitoring and application profiling.
-Tools like TAU (Tuning and Analysis) \cite{tau}, SCALASCA \cite{scala} and mpiP \cite{mpip} can capture application's runtime information about application's communication and computation in event traces.
-However, recognizing the patterns about application's communication from those traces requires substantial effort.
-There are a number of studies on the recognition and characterization of parallel application communication patterns.
-Oak Ridge National Laboratory has its ongoing project about developing a tool set named Oxbow, which can characterize the computation and communication behavior of scientific applications and benchmarks \cite{oxbow}.
-In a recent work \cite{roth}, the authors demonstrate a new approach to automatically characterizing parallel applications' communication behaviors.
-
-%\textbf{P2: communication characteristics}\\
-
-Many research efforts have been conducted to characterize scientific applications.
-For instance, a DOE project aims at the identification of the computational characteristics of the DOE MiniApps developed at various exascale co-design centers \cite{designforwardwebpage}.
-In this project, the communication patterns of several DOE full applications and associated mini-applications are studied to provide a more complete snapshot of the DOE workload.
-A joint project named CORAL from Oak Ridge, Argonne and Livermore provide a series of benchmarks to represent DOE workloads and technical requirements \cite{coral}.
-The CORAL project includes scalable science benchmarks, throughput benchmarks, data centric benchmarks, skeleton benchmarks and Micro benchmarks.
-
-%\textbf{P3: Interference between concurrently running jobs}\\
-The interference between concurrently running jobs on HPC system have been identified as major culprit for job's performance variability.
-Abhinav et al.
-found that concurrently running applications on HPC can cause interference to each other, and cause communication time varied from 36\% faster to 69\% slower.
-Such interference could come from Operating System noise, shape of the allocated partition , or other running jobs that sharing network resources \cite{abhinav-sc13}.
-Skinner et al.
-found that  there is a 2-3$\times$ slowdown in MPI$\textunderscore$Allreduce due to network contention from other jobs \cite{skinner}.
-Rosenthal et al.
-found increasing network bandwidth provide limited benefits to a number of applications.
-The applications that send mostly small messages or large messages asynchronously are not bandwidth bounded, hence, benefit only slightly or not at all from increased bandwidth \cite{rosenthal}.
-
-%\textbf{P4: Job allocation(little bit)}\\
-Several research studies focus on optimizing job allocation on HPC system to alleviate the interference between concurrently running jobs.
-Hoefler et al.
-proposed to use performance modeling techniques to analyze factors that impact the performance of parallel scientific applications \cite{hoefler-modeling}.
-However, as the scale of HPC continue to grow, the  interference of concurrently running jobs is getting worse, which is hard to be quantified by performance profiling tools alone.
-Bogdan et al.
-provide a set of guidelines of how to configure a network with Dragonfly topology for workload with Nearest Neighbor communication pattern \cite{Bogdan-hpdc14}.
-Dong et al.
-have developed simple benchmarks that conforms to four different communication patterns, namely ping-pong, nearest neighbor, broadcast and all$\textunderscore$reduce, to demonstrate the effectiveness of this highly parallel 5D torus network \cite{Dong-SC11}.
-
-Our work is different from all these works in the following ways.
-First, we focus on the dominant communication patterns rather than any specific application.
-We believe this can provide a guideline for other research work.
-Secondly, we explore the intra- and inter-job interference between concurrently running jobs, while similar work such as \cite{abhinav-sc13} only focuses on the single application's performance degradation due to network contention.
-Finally, we analyze the impact of different allocation strategies to job's communication behaviors.
-We further identify the optimal allocation strategy for each application with a specific dominant communication pattern.
-Based on our study, we claim that further scheduling should take job's communication pattern into consideration for its allocation decision making.
-
-\section{Conclusions}
-\label{sec:conclusion}
-In this work, we have studied the communication behavior of three parallel applications, namely AMG, CrystalRouter and MultiGrid.
-Each application has distinctive communication pattern, that can be representative for a group of jobs in HPC workload.
-We have used a sophisticate simulation tool named CODES from Argonne National Laboratory to simulate the running of these three parallel applications on torus network.
-The torus network provided by CODES has good fidelity and scalability.
-We have analyzed the performance of each application's communication in terms of data transfer time by simulating them running on torus networks with different bandwidth and dimensionality configurations.
-We have found that higher dimensionality of torus network would improve the performance of application with ``Many-to-Many" communication patterns, while application with intensive local communication like ``Nearest Neighbor" won't benefit much from higher dimensionality.
-
-We have analyzed the intra- and inter-job interference by simulating three applications running on 3D torus network.
-Based on our comprehensive experiments, we got five observations.
-1) The compact allocation may not guarantee the best performance for every application.
-2)The applications dominated with Nearest Neighbor communications exhibit relatively stable performance under different allocation shapes, while the applications dominated with Many-to-Many communication exhibit better performance with compact allocation (e.g.
-3D balanced).
-3) A good rank-to-node mapping strategy can greatly improve application's performance when specific allocation is given.
-4) An optimal size for allocation units should be determined according to an application's dominant communication pattern.
-In general, a unit size should be large enough to preserve neighboring communication of the application.
-5) Inter-job interference is inevitable in non-contiguous allocation.
-However, choosing the proper allocation unit size with job's communication pattern awareness can alleviate such negative effect.
-
-We believe that our finding in this work can provide valuable guidance for HPC resource management to make flexible job allocations.
-Rather than using pre-defined partitions and non-contiguous allocation, future HPC systems should assign each job with preferable resources based on job's communication pattern.
-
-\section*{Acknowledgment}
-\label{sec: ack}
-The work at Illinois Institute of Technology is supported in part by U.S.
-National Science Foundation grants CNS-1320125 and CCF-1422009.
-This work is also supported by the U.S.
-Department of Energy, Office of Science, Advanced Scientific Computing Research, under Contract DE-AC02-06CH11357.
-
-\bibliographystyle{IEEEtran}
-\bibliography{./reference}
-
-\vspace{5\baselineskip}
-
-\begin{framed}
-The submitted manuscript has been created by UChicago Argonne, LLC, Operator of Argonne National Laboratory ("Argonne").
-Argonne, a U.S.
-Department of Energy Office of Science laboratory, is operated under Contract No.
-DE-AC02-06CH11357.
-The U.S.
-Government retains for itself, and others acting on its behalf, a paid-up nonexclusive, irrevocable worldwide license in said article to reproduce, prepare derivative works, distribute copies to the public, and perform publicly and display publicly, by or on behalf of the Government.
-\end{framed}
-
-\end{document}

### 2016-07-07 discussion.tex
split into separate tex files for each section

--- 2015-10-17 submission.tex
@@
-\documentclass[conference]{IEEEtran}
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
-\else
-
-\fi
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-\usepackage{algorithm}
-%\usepackage{algorithm2e}
-\usepackage{algpseudocode}
-\usepackage{booktabs}
-\usepackage{flushend}
-\usepackage{url}
-\usepackage{xcolor}
-\usepackage{amsmath}
-\usepackage{framed}
-\usepackage{textcomp}
-%\usepackage{subfig}
-\usepackage{graphicx}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{soul}
-\usepackage{verbatim}
-%\usepackage[labelfont=bf]{caption}
-\usepackage[font=bf,labelsep=space]{caption}
-
-\newcommand{\TODO}[1]{\hl{TODO: #1}}
-\newcommand{\NOTE}[1]{\hl{NOTE: #1}}
-\newcommand{\changed}[1]{\textcolor{red}{#1}}
-
-\begin{document}
-%
-% paper title
-% can use linebreaks \\ within to get better formatting as desired
-\title{Study of Intra- and Inter-Job Interference with Different Job Allocations}
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
-\author{
-
-\IEEEauthorblockN{Xu Yang\IEEEauthorrefmark{1}, John Jenkins\IEEEauthorrefmark{2}, Misbah Mubarak\IEEEauthorrefmark{2}, Robert B.
-Ross\IEEEauthorrefmark{2}, Zhou Zhou\IEEEauthorrefmark{1}, Zhiling Lan\IEEEauthorrefmark{1}}
-
-\IEEEauthorblockA{\IEEEauthorrefmark{1}Department of Computer Science,
-Illinois Institute of Technology,
-Chicago, Illinois, USA 60616\\
-\{xyang56, zzhou\}@hawk.iit.edu, lan@iit.edu}
-
-\IEEEauthorblockA{\IEEEauthorrefmark{2}Mathematics and Computer Science Division, Argonne National Laboratory,
-Argonne, IL, USA 60439\\
-\{jenkins,rross\}@mcs.anl.gov, mmubarak@anl.gov}
-}
-
-% make the title area
-\maketitle
-
-\begin{abstract}
-
-Network contention between concurrently running jobs on HPC systems is a primary cause of performance variability.
-Optimizing job allocation and avoiding network sharing are hence crucial to alleviate the potential degradation in performance.
-In order to do these effectively, an understanding of the interaction between applications, their communication patterns, and the system architecture are required.
-In this work, we choose three representative HPC applications from DOE Design Forward Project and conduct detailed simulations of high-performance network topologies to analyze both intra- and inter-job interference.
-By scrutinizing the communication behaviors of these applications, we identify relationships between these behaviors and the possible interference introduced by different job allocation scenarios.
-Our comprehensive analyses in this work illuminate a path towards job-communication-pattern-aware HPC resource allocation.
-
-\end{abstract}
-
-\IEEEpeerreviewmaketitle
-
-\section{Introduction} 
-\label{sec: intro}
-
-The scale of supercomputers continually grows at a breakneck pace to accommodate the computing power requirements for scientific research.
-Supercomputers have tens of thousands of nodes and serve as an irreplaceable research vehicle for scientific problems with increasing size and complexity.
-Supercomputers are usually employed as a shared resource to accommodate many parallel applications (jobs) running concurrently.
-These parallel jobs share the system infrastructure such as network and I/O bandwidth, and inevitably there is contention over these shared resources.
-As supercomputers continue to evolve, these shared resources are increasingly the bottleneck for performance.
-
-\begin{figure}[h!]
-    \centering
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/goodallocation}
-        \caption{Contiguous}
-        \label{fig:overview_sub1}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/badallocation}
-        \caption{Non-contiguous}
-        \label{fig:overview_sub2}
-    \end{subfigure}%
-   \caption{Multiple jobs running concurrently with different allocations.
-Each job is represented by a specific color.
-a) shows the effect of contiguous allocation, which reduce the inter-job interference.
-b) shows non-contiguous allocation, which may introduce both intra and inter-job interferences.
-}
-   \label{fig:overview}
-\end{figure}
-
-When submitting jobs to HPC systems, users request resources in terms of node counts and expected runtime.
-The batch scheduler is then responsible for dispatching the submitted jobs to the system.
-Typically, multiple jobs run concurrently, resulting in the shared use of resources, particularly network and storage.
-A prominent problem with this approach is network contention among concurrently running jobs.
-If the network resources of concurrently running jobs aren't isolated, the resulting sharing can cause communication variability that results in performance degradation for some or all such jobs \cite{abhinav-sc13}.
-This delay can propagate into the queueing time of the following submitted jobs, thus leading to low system throughput and utilization \cite{jose-ipdps15}.
-This adverse effect of network sharing can be mitigated by providing jobs with isolated allocations and exclusive network resources.
-
-On the widely used torus-connected HPC systems \cite{bgq,tofu,titan}, two allocation strategies are commonly used.
-The \emph{contiguous allocation} strategy assigns to each job a compact and contiguous set of computing nodes, as shown in Figure \ref{fig:overview_sub1}.
-The partition-based allocation approach used in Blue Gene series systems is an example of such a strategy \cite{bgloverview}.
-The contiguous strategy favors application performance through isolated networking within a partition and the locality that implies.
-However, this strategy can cause both internal fragmentation (when more nodes are allocated to a job than it requests) and external fragmentation (when sufficient nodes are available for a request, but they can not be allocated contiguously), therefore leading to lower system utilization than is otherwise possible.
-On the other side of the coin, the \emph{non-contiguous allocation} strategy, used by the Cray XT/XE series \cite{carl-cug}, assigns free nodes to jobs regardless of contiguity, though of course efforts are made to maximize locality.
-Figure~\ref{fig:overview_sub2} shows an example non-contiguous allocation.
-While eliminating internal and external fragmentation as seen in contiguous allocation systems, in return non-contiguous allocations introduce contention between jobs due to the interleaving of job nodes.
-The non-contiguous node allocation can significantly reduce job performance, especially for communication-intensive ones \cite{abhinav-sc13}.
-
-We envision that future HPC schedulers will adopt a flexible job allocation mechanism which combines the best of both contiguous and non-contiguous allocation strategies.
-Such a flexible mechanism would take shared resource needs of jobs into account when making allocation decisions (e.g., network).
-With knowledge and analysis of job communication patterns, it can be identified which jobs require network isolation and locality, and to what degree.
-Then, rather than allocating each job in a ``know-nothing'' manner, one may specialize allocation so that, for example, only the jobs with stringent network needs are given compact, isolated allocations, resulting in maximized utilization and minimized perceivable resource contention effects.
-
-In this work, we focus on an in-depth analysis of intra- and inter-job communication interference with different job allocation strategies on torus-connected HPC systems.
-Torus-based networks are used on six of the top 10 supercomputers on the June 2015 Top500 lists~\cite{top500}.
-The current generation of IBM Blue Gene/Q (BG/Q) supercomputers, such as Mira at Argonne Nation Laboratory and Sequoia at Lawrence Livermore National Laboratory, has its nodes connected in a 5D torus network \cite{bgq}.
-The K computer from Japan uses the ``Tofu'' system, which is a 6D mesh/torus topology \cite{tofu}.
-Titan, a Cray XK7 supercomputer located at the Oak Ridge Leadership Computing Facility (OLCF), has nodes connected in a 3D torus within the compute partition \cite{titan}.
-Although our analyses are based on torus networks, the ideas conveyed in this work is applicable to networks with different topologies.
-
-We selected three signature applications from the DOE Design Forward Project~\cite{designforwardwebpage} as examples to conduct detailed study about their communication patterns.
-We use a sophisticated simulation toolkit named CODES (standing for Co-Design of Multi-layer Exascale Storage Architectures)~\cite{Jason-2011} as a research vehicle to evaluate the performance of these applications with various allocations in a controlled environment.
-We then analyze the intra- and inter-job interference by simulating these applications running exclusively and concurrently with different allocations.
-We believe the insights presented in this work can be very useful for the design of future HPC batch job schedulers and resource managers.
-
-%\NOTE{JJ removed the list of contributions, because they are exactly the same as the preceding paragraph.}
-\begin{comment}
-The observations of this work are the following:
-
-\begin{itemize}
-    \item We study the communication behaviors and identify the communication patterns of three representative HPC applications.
-    \item We conduct detailed analysis about the intra- and inter-job interference by running three applications with different allocation strategies.
-And we find the relation between application's communication pattern and the cause of interference.
-    \item We show a novel and flexible way to make allocation with awareness of job's communication pattern, which  is every effective to alleviate the interference between concurrently running jobs.
-\TODO{do we actually do this, or provide the underpinnings for such a method?}
-\end{itemize}
-
-We believe the finds and insights presented in this work would be very useful for the design of future HPC batch job scheduler and resource management module.
-
-\end{comment}
-
-The rest of this paper is organized as follows.
-Section~\ref{sec:application study} describes the three representative applications from DOE Design Forward Project for our study.
-Section~\ref{sec:codes} talks about the use of CODES as research vehicle for our work.
-Section~\ref{sec:config study} shows the performance analysis of three applications on different torus networks.
-Section~\ref{sec:interference} provides detailed analysis about the intra- and inter-job interference between three applications on torus network with different allocation strategies.
-Section~\ref{sec:discussion} introduces a path toward communication-pattern aware allocation strategies, given the results of our analysis.
-Section~\ref{sec:related_work} discusses related work.
-Finally, the conclusion is presented in Section~\ref{sec:conclusion}.
-
-\section{Application Study}
-\label{sec:application study}
-
-%A parallel application usually conforms to a combination of several basic communication patterns\cite{roth}.
-At its different execution phases, the application's communication behavior may follow different basic patterns respectively.
-\textcolor{blue}{When we look into the data flow during application execution, most parallel applications start with broadcast operation to distribute the data from root process to other processes\textcolor{red}{ref needed here}, followed by a series of computation and communication that conforms to certain pattern, which is usually the dominant part of application's execution.
-Before the application come to completion, all the working process will return their results to the root directly or hierarchically.} Parallel application's communication pattern usually involves lots of factors, such as communication intensity, operation-to-operation dependencies and the ``critical path" in its communication topology graph, etc.
-In this work, we focus on the locality of application's rank-to-rank communication topology graph.
-
-A parallel application usually conforms to a combination of several basic communication patterns \cite{roth}.
-At its different execution phases, the application's communication behavior may follow different basic patterns respectively.
-There are many profiling tools available to capture information regarding communication patterns of parallel applications~\cite{tau,mpip,scala,sst,oxbow}.
-They can provide information such as the percentage of different MPI operations of the applications, communication topology, the amount of data transferred between processes.
-
-There are many applications running on HPC systems, such as Mira \cite{bgq} and Titan \cite{titan}, that are communication intensive and conform to various communication patterns.
-In this work, we select three representative applications from the DOE Design Forward Project.
-Each application exhibits a distinctive communication pattern that is commonly seen in HPC applications.
-We believe that the communication patterns of these applications are representative of a wide array of applications running on leadership-class machines.
-Specifically, we study the Algebraic MultiGrid Solver (AMG), Geometric MultiGrid (MultiGrid) and CrystalRouter MiniApps.
-The communication pattern figures we present in Section~\ref{sec:amg}, \ref{sec:crystalrouter}, and \ref{sec:multigrid} are generated with the IPM~\cite{ipm} data gathered from publicly available traces~\cite{designforwardwebpage}.
-
-% MiniApps are reduced proxy applications that encapsulate the salient performance of larger full size applications \cite{miniapp}.
-Since we only focus on the dominant communication pattern of parallel applications, these MiniApps are perfect candidates for analysis the application's performance on different allocations or on new network architectures.
-We choose three MiniApps, namely, Algebraic Multigrid Solver(AMG), Geometric MultiGrid(MultiGrid) and CrystalRouter.
-Their representative communication patterns that are commonly seen in the HPC system workload.
-The communication pattern figures we present here are generated with the IPM \cite{ipm} data from \cite{designforwardwebpage}.
-
-\subsection{AMG}
-\label{sec:amg}
-
-\begin{figure}[t!]
-    \centering
-    \begin{subfigure}[t]{0.25\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/appstudy/amg/amg_pip}
-        \caption{}
-        \label{fig:amg-communication-topology}
-    \end{subfigure}%
-    ~
-    \begin{subfigure}[t]{0.24\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/appstudy/amg/amg_data_transfer}
-        \caption{}
-        \label{fig:amg-data-trans}
-    \end{subfigure}
-    \caption{AMG.
-(a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in AMG.
-(b) The amount of data sent/received by each rank in AMG.
-Y-axis labels the amount of data transferred in MB, x-axis labels 216 ranks of AMG.
-}
-\end{figure}
-
-The Algebraic MultiGrid Solver, or AMG, is a parallel algebraic multi-grid solver for linear systems arising from problems on unstructured mesh physics packages.
-It has been derived directly from the BoomerAMG solver that is being developed in the Center for Applied Scientific Computing (CASC) at LLNL \cite{amg}.
-The dominant communication pattern is regional communication with decreasing message size for different parts of the multi-grid v-cycle.
-
-Figure~\ref{fig:amg-communication-topology} shows the communication topology of a small scale AMG execution with 216 MPI Ranks.
-Note that the dominant communication pattern of the application doesn't change with scale.
-We can make the observation that AMG's dominant communication pattern is 3D Nearest Neighbor, each rank has intensive communication with up to six neighbors, depending on the boundaries of the ranks.
-Applications with similar pattern include PARTISN \cite{partisn} and SNAP \cite{snap}.
-
-Figure~\ref{fig:amg-data-trans} shows the amount of data each rank transferred in the example execution.
-The data transferring of each rank is between 7MB to 16MB, corresponding to whether the rank is on a boundary or not.
-
-\subsection{Crystal Router}
-\label{sec:crystalrouter}
-
-\begin{figure}[t!]
-    \centering
-    \begin{subfigure}[t]{0.25\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/appstudy/cr/cr_pip}
-        \caption{Communication Topology}
-        \label{fig:cr-communication-topology}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[t]{0.22\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/appstudy/cr/cr_data_transfer}
-        \caption{Data Volume}
-        \label{fig:cr-data-trans}
-    \end{subfigure}
-    \caption{CrystalRouter.
-(a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in CrystalRouter.
-(b) The amount of data sent/received by each rank in CrystalRouter.
-Y-axis labels the amount of data transferred in MB, x-axis labels 100 ranks of CrystalRouter.
-}
-\end{figure}
-
-The second MiniApp covered is CrystalRouter, the extracted communication kernel of the full application Nek5000~\cite{crystalrouter}, which is a spectral element CFD application developed at Argonne National Laboratory \cite{nek5000}.
-It features spectral element multi-grid solvers coupled with a highly scalable, parallel coarse-grid solver that is widely used for projects including ocean current modeling, thermal hydraulics of reactor cores, and spatiotemporal chaos.
-CrystalRouter demonstrates the ``many-to-many'' communication pattern through a scalable multi-stage communication process.
-The way CrystalRouter works is nodes compute for a while, then synchronize and communicate, continually alternating between these two types of activities.
-
-The collective communication in CrystalRouter utilizes a recursive doubling approach.
-Ranks in CrystalRouter conform to a $n$-dimensional hypercube and recursively split into ($n$-1)-dimensional hypercubes, with communication occurring along the splitting plane.
-The pattern of this communication can be found in Figure~\ref{fig:cr-communication-topology}.
-By nature of the logarithmic splitting process, a substantial portion of the communication occurs in small neighborhoods of ranks.
-CrystalRouter represents a group of applications whose dominant communication is a hybrid of multi-stage local and hierarchical global communication, and shares similarities with most MPI collective communication implementations.
-
-Figure~\ref{fig:cr-data-trans} shows the amount of data each rank transferred in CrystalRouter, which ranges between 20MB to 36MB.
-This is due to the rank counts not matching one-for-one with points in the hypercube.
-
-\subsection{MultiGrid}
-\label{sec:multigrid}
-
-\begin{figure}[t!]
-    \centering
-    \begin{subfigure}[t]{0.25\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/appstudy/mg/mg_pip}
-        \caption{Communication Topology}
-        \label{fig:mg-communication-topology}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[t]{0.22\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/appstudy/mg/mg_data_transfer}
-        \caption{Data Volume}
-        \label{fig:mg-data-trans}
-    \end{subfigure}
-    \caption{MultiGrid.
-(a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in MultiGrid.
-(b) The amount of data sent/received by each rank in MultiGrid.
-Y-axis labels the amount of data transferred in MB, x-axis label 125 ranks of MultiGrid.
-}
-\end{figure}
-
-MultiGrid is geometric multi-grid v-cycle from the production elliptic solver BoxLib, a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes \cite{boxlib}.
-MultiGrid conforms to many-to-many communication pattern with decreasing message size and collectives for different parts of the multi-grid v-cycle.
-It is widely used for structured grid physics packages.
-
-Figure~\ref{fig:mg-communication-topology} shows the communication topology of MultiGrid with 125 ranks.
-We can see intensive communication along the diagonal that resembles nearest neighbor communication, similar to AMG.
-However, the communication topology leads to a greater ``spread'' of communication across the set of ranks, challenging the maximization of communication locality with respect to rank.
-In this sense, it can be considered a ``many-to-many'' pattern.
-Applications with such similar dominant communication pattern are like FillBoundary, another PDE solver code in \cite{boxlib}.
-
-Figure~\ref{fig:mg-data-trans} shows the amount of data transferred between ranks in MultiGrid. For this miniapp, The data transferred per rank is relatively small and stable, between 2.8MB to 2.9MB.
-
-\section{Research Vehicle}
-\label{sec:codes}
-
-It is difficult to accurately and flexibly experiment with concurrently running jobs in an HPC context.
-One reason is that the allocation strategy used on production machine is part of the system software, which can not be changed by users.
-Even the system administrator is not authorized to make that change.
-Another reason is that it is unrealistic to reserve the system exclusively to run the same job with desired allocation without interference and then compare the results with those in the presence of interference.
-Therefore, we resort to simulation for this work.
-
-A simulation toolkit named CODES enables the exploration of simulating different HPC networks with high fidelity \cite{Jason-2011}\cite{mubarak-sc2012}.
-CODES is built on top of Rensselaer Optimistic Simulation System (ROSS) parallel discrete-event simulator, which is capable of processing billions of events per second on leadership-class supercomputers \cite{ross}.
-CODES support both torus and dragonfly network with high fidelity flit-level simulation.
-CODES has this network workload component that is capable of conducting trace-driven simulations.
-It can take real MPI application traces generated by SST DUMPI \cite{sst} to drive CODES network models.
-In this work, we focus on an in-depth analysis of intra- and inter-job communication interferences with  different job allocations on torus-connected HPC systems.
-Torus networks have been extensively used in the current generation of supercomputers because of their linear scaling on per-node cost and competitive communication performance \cite{mubarak-sc2012}\cite{zhou-ipdps}.
-
-\section{Configuration Study}
-\label{sec:config study}
-
-In this section, we study the communication behavior of three applications on torus networks with different dimensionality and bandwidth configurations.
-Torus networks are $k$-ary $n$-cubes, with $k^{n}$ nodes in total arranged in an $n$-dimensional grid having $k$ nodes in each dimension.
-Each node has $2\times n$ direct linked neighbor nodes with wraparound.
-In these and all following experiments, non-adaptive routing is used.
-While making specific recommendations as to an ``optimal'' network configuration for a given application is out of the scope of this study, the experiments do highlight the nuances inherent in design choices such as network diameter and bandwidth planning.
-
-We perform three sets of experiments.
-Each looks at per-rank distribution of performance of AMG, CrystalRouter and MultiGrid on a set of 2048-node torus models, one with $n=3$ ($16^{2} \times 8$), one with $n=5$ ($8\times 4^{4}$), and one with $n=7$ ($4^{4} \times 2^{3}$).
-Performance in this context is defined by the total message transfer time for each sending rank, which we are able to precisely measure due to the simulation environment providing us with an absolute notion of time.
-Note that we simply perform linear rank assignment, one per node -- later sections investigate rank and node mapping effects.
-In the first, shown in Figure~\ref{fig:dimensionality-study}, we keep the link bandwidth constant at 2GiB/s in each direction, corresponding to Blue Gene/Q systems~\cite{bgq}, leading to aggregate node bandwidths of 12GiB/s, 20GiB/s, and 28GiB/s for the 3D, 5D, and 7D tori, respectively.
-In the second, shown in Figure~\ref{fig:bandwidth-study}, we fix the torus dimensionality to 5D.
-In the third, shown in Figure~\ref{fig: bandwidth-time-box}, we instead fix the per-node aggregate bandwidth (and thus the system aggregate bandwidth) and modify the torus dimensionality, resulting in per-link bandwidths shown in Table~\ref{tab: fix-bandwidth}.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/dimenstudy/amg_box}
-        \caption{AMG}
-        \label{fig:dimen-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/dimenstudy/cr_box}
-        \caption{CrystalRouter}
-        \label{fig:dimen-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/dimenstudy/mg_box}
-        \caption{MultiGrid}
-        \label{fig:dimen-mg}
-    \end{subfigure}%
-   \caption{Data Transfer Time of AMG, CrystalRouter and MultiGrid on 3D, 5D and 7D torus network, with constant per-link bandwidth.}
-   \label{fig:dimensionality-study}
-\end{figure*}
-
-In the fixed link bandwidth experiment, shown in Figure~\ref{fig:dimensionality-study}, the data transfer time of the three applications are significantly reduced as the dimensionality increases, so the applications can easily make use of the increased dimensionality (lower hops on average) and/or aggregate bandwidth (more links being concurrently used).
-Interestingly, AMG has the greatest degree of outliers compared to the others.
-We expect this to be a combination of non-optimal rank mappings causing a greater degree of variance due to the highly regular communication pattern and the distribution of message sizes.
-MultiGrid seems to benefit the most in terms of reducing the spread of the time distribution, though the median message time remains roughly within a factor of two.
-The time distribution in CrystalRouter seems to accrue the most benefits, likely due to the combination of cross-cutting and localized per-rank traffic.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/bandwidthstudy/amg_bw_box}
-        \caption{AMG}
-        \label{fig:bdwstudy-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/bandwidthstudy/cr_bw_box}
-        \caption{CrystalRouter}
-        \label{fig:bdwstudy-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/bandwidthstudy/mg_bw_box}
-        \caption{MultiGrid}
-        \label{fig:bdwstudy-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 5D torus network with 2GiB/s, 4GiB/s and GiB/s direct link bandwidth. }
-   \label{fig:bandwidth-study}
-\end{figure*}
-
-In the fixed dimensionality experiment, shown in Figure~\ref{fig:bandwidth-study}, the increased per-link bandwidth, expectedly, benefits each application roughly equally, reducing the time distributions linearly.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/samebdw/amg}
-        \caption{AMG}
-        \label{fig:samebd-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/samebdw/cr}
-        \caption{CrystalRouter}
-        \label{fig:samebd-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/samebdw/mg}
-        \caption{MultiGrid}
-        \label{fig:samebd-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 3D, 5D and 7D torus network with same aggregated bandwidth. }
-   \label{fig: bandwidth-time-box}
-\end{figure*}
-
-\begin{table}[ht]
-\begin{center}
-\caption{Torus networks with same aggregated bandwidth} 
-\label{tab: fix-bandwidth}
-\begin{tabular}{l c c c} 
-\toprule % Top horizontal line
-\toprule
-&\multicolumn{3}{c}{Bandwidth } \\
-\cmidrule(l){2-4}
-Dimension  & per-link & aggregate &\\ % Column names row
-\midrule % In-table horizontal line
-3D      & 4.67GiB/s  & 28GiB/s  & \\  % Content row 1
-\midrule % In-table horizontal line
-5D    & 2.8GiB/s  & 28GiB/s  &\\ % Content row 1
-\midrule % In-table horizontal line
-7D    & 2GiB/s & 28GiB/s & \\
-\midrule
-\bottomrule % Bottom horizontal line
-\end{tabular}
-\end{center}
-\end{table}
-
-In the fixed aggregate bandwidth experiment, with configuration in Table~\ref{tab: fix-bandwidth} and results shown in Figure~\ref{fig: bandwidth-time-box}, the results are more nuanced.
-At least for these applications, the benefits from increased dimensionality do not outweigh the reduction in per-link bandwidth.
-MultiGrid does see benefits in the increased dimensionality, particularly with respect to outliers, but has a similar median message time.
-CrystalRouter reduces performance as the dimensionality increases, but only by a small degree.
-AMG remains roughly the same, minus a dip in the maximum message time that could be due to rank-to-node mapping changes.
-
-\section{Interference Analysis}
-\label{sec:interference}
-
-\textbf{\emph{Intra-job interference}} refers to the network contention between ranks within each application.
-\textbf{\emph{Inter-job interference}} is introduced by concurrently running jobs that adjacent to each other sharing network resources.
-Communication variability due to such interference can cause application performance degradation.
-
-In this section we will study both kinds of interference in the context of torus network. 
-Since application's communication pattern won't change with its scale, we use AMG with 216 ranks, CrystalRouter with 100 ranks, and MultiGrid 125 ranks.
-And in order to accommodate three applications running concurrently with different allocations, we simulate a 2K-node 3D torus ($16^{2} \times 8$) network for the experiments.
-
-\subsection{Intra-Job Interference Analysis}
-\label{sec: introjob}
-
-We design two sets of experiments to study the intra-job interference of each application.
-In the first, shown in Figure \ref{fig:shapestudy}, we assign each application with allocations in \emph{three different shapes}.
-In the second, shown in Figure \ref{fig:diffmap}, we study the intra-job interference by using different mapping strategies for each application with given allocation shape.
-
-In the allocation shapes experiment, we select three possible shapes can can commonly seen on the 3D torus network.
-They are 3D balanced-cube, 3D unbalanced-cube, 2D mesh, as shown in Figure \ref{fig:alloc-shapes}.
-
-3D balanced-cube, shown as red in Figure \ref{fig:cont_sub1}, can guarantee the minimum \emph{average pair-wise distance} within the allocation.
-Some research studies like \cite{leung,abhinav-sc13} indicate that compact allocation can guarantee jobs with better performance.
-They use a variety of metrics to evaluate the compactness of the allocation, such as \emph{average pair-wise distance}, \emph{diameter} and \emph{contiguity}.
-In this work, we select 3D balanced-cube as the most compact allocation on a 3D torus network.
-
-3D unbalanced-cube, shown as green in Figure \ref{fig:cont_sub1}, is a rectangular prism, which is the possible allocation shape on the systems with asymmetric networks.
-For example, Cray XE6/XK7 systems are deployed 3D tori with Gemini routers.
-The network connections in the \emph{y}-direction have only half the bandwidth of the cables used in the \emph{x} and \emph{z} directions.
-In order to take advantage of the faster links in the \emph{x} and \emph{z} directions, job allocation starts from X-Z plane, which leads to a rectangular prism shaped allocation \cite{RF}.
-
-2D mesh, shown as blue in Figure \ref{fig:cont_sub1}, can be cut out from a single layer of the 3D torus.
-2D mesh is a very common allocation shape in torus network for both contiguous and non-contiguous allocation strategies.
-For example, Cray Application Level Placement Scheduler (ALPS) indexes the nodes in torus network into a list and makes job allocation by simply going through the list \cite{carl-cug}.
-When the list is obtained by sorting the nodes based on their spacial coordinates in the torus, allocation made off from this list will be 2D mesh.
-The IBM Blue/Gene Q supercomputer Mira at Argonne Leadership Computing Facility also allow its allocation partition configured into mesh \cite{zhou-ipdps}.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/allocation}
-        \caption{ }
-        \label{fig:cont_sub1}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit16}
-        \caption{ }
-        \label{fig:noncont_sub1}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit8}
-        \caption{ }
-        \label{fig:noncont_sub2}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit2}
-        \caption{ }
-        \label{fig:noncont_sub3}
-    \end{subfigure}%
-    \caption{(a) Contiguous allocation in three different shapes.
-``Blue" is 2D mesh, ``Green" is 3D-unbalanced cube and ``Red" is 3D-balanced cube.
-(b)-(d) Non-contiguous allocation.
-Each job is represented by a specific color.
-The nodes assigned to different jobs are interleaved, the size of allocation unit are 16, 8 and 2.}
-    \label{fig:alloc-shapes}
-\end{figure*}
-
-AMG with 3D ``Nearest Neighbor" communication pattern takes less time when being assigned with 3D-unbalanced allocation than 3D-balanced, as shown in Figure \ref{fig:shapstudy-amg}.
-And MultiGrid gets the best performance (shortest data transfer time) when running on 3D-balanced allocation, as shown in Figure \ref{fig:shapstudy-mg}.
-Since MultiGrid's communication pattern is ``Many-to-Many" dominant, 3D-balanced is the most compact allocation with shortest pair-wise distance between nodes, which can reduce the aggregated hops for transferring message among ranks in MultiGrid.
-As CrystalRouter has both dominant local communication and multi-stage global data transfer, the 3D-balance is also the best allocation, but its advantage over 3D-unbalanced is not that obvious as to MultiGrid, as shown in Figure \ref{fig:shapstudy-cr}.
-
-Lots of research studies have designed fancy allocation algorithms to provide application with the most compact allocation\cite{leung,LO}.
-However, as shown in our experiments, providing cubical allocation without considering application's communication pattern won't guarantee the best performance for every application.
-Compact allocation should be provided to application with global data transfer, such as those conforms to ``Many-to-Many" communication pattern.
-Application with dominant communication pattern like ``Nearest Neighbor" won't fully utilize the compact allocation and could get even better performance on a relaxed allocation shape.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/amg_box}
-        \caption{AMG}
-        \label{fig:shapstudy-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/cr_box}
-        \caption{CrystalRouter}
-        \label{fig:shapstudy-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/mg_box}
-        \caption{MultiGrid}
-        \label{fig:shapstudy-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 3D-balanced, 3D-unbalanced and 2D allocation.}
-   \label{fig:shapestudy}
-\end{figure*}
-
-The rank-to-node mapping of parallel applications on HPC system could greatly impact applications' performance.
-However, finding the optimal mapping solution for a given application is out the scope of this work.
-Our experiments aims to show how mapping strategies impact the intra-job interference of application with specific communication pattern.
-
-We provide AMG, CrystalRouter and MultiGrid with 3D-balance allocation and use three mapping strategies to do the rank-to-node mapping.
-The ``Linear" mapping, as we used in the previous experiments, is to map each rank according to the dimensional ordering of compute nodes.
-The ``Cube" mapping assigns ranks into consecutive $2^{3}$ cubes.
-The ``Random" mapping assigns ranks randomly within the allocation.
-
-AMG remains roughly the same when it been mapped by ``Linear" and ``Cube", shown in Figure \ref{fig:diffmap-amg}.
-Although ``Linear" and ``Cube" mapping strategies cause some routing overlap for AMG's communication, both mapping strategies still preserve the locality of AMG's 3D ``Nearest Neighbor" communication pattern.
-The ``Random" mapping totally destroys AMG's communication pattern, and results in intra-job interference among all the ranks in AMG.
-The performance degradation caused by using ``Random" mapping is up to 90\%.
-
-The ``Cube" mapping improves CrystalRouter's performance over the ``Linear" by reducing the pair-wise distance between ranks, shown in Figure \ref{fig:diffmap-cr}.
-The global data transfer in CrystalRouter will takes fewer hops with ``Cube" mapping.
-The ``Random" mapping scatters the ranks within the local neighborhood of CrystalRouter and make their communication less efficient.
-
-The ``Cube" mapping benefits MultiGrid's ``Many-to-Many" communication.
-However, due to the small amount of data transfer among ranks in MultiGrid, ``Cube" mapping doesn't exhibit a drastic advantage.
-``Random" mapping neither cause too much degradation for the same reason, shown in Figure \ref{fig:diffmap-mg}.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/diffmapping/amg}
-        \caption{AMG}
-        \label{fig:diffmap-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/diffmapping/cr}
-        \caption{CrystalRouter}
-        \label{fig:diffmap-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/diffmapping/mg}
-        \caption{MultiGrid}
-        \label{fig:diffmap-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 3D-balanced allocation using  different mapping strategies.}
-   \label{fig:diffmap}
-\end{figure*}
-
-\subsection{Inter-Job Interference Analysis}
-\label{sec: interjob interference study}
-
-Inter-job interference has been identified as one of the major culprits for application's performance variability \cite{abhinav-sc13,skinner,rosenthal}.
-Inter-job interference is a more prominent issue for system adopted non-contiguous allocation policy than system with contiguous allocation.
-Application communication times have been demonstrated to vary from 36\% faster to 69\% slower due to job interference when running on non-contiguous allocation \cite{abhinav-sc13}.
-
-We assign each application with non-contiguous allocation and run them concurrently on the same network.
-The allocation unit belongs to different jobs are interleaved.
-In order to study the impact of different allocation unit size on applications' inter-job interference, we conduct experiments with unit size of 16, 8 and 2, shown respectively in Figure \ref{fig:noncont_sub1}, \ref{fig:noncont_sub1} and \ref{fig:noncont_sub3}.
-Figure \ref{fig:interjobstudy} shows the results of each application data transfer time with different allocation unit size.
-
-The data transfer time of AMG in Figure \ref{fig:interjob-amg} keeps stable between allocation unit size of 16 and 8, since the locality of AMG is 6-rank based.
-When the unit size reduce to 2, AMG suffers prolong data transfer time by about 10\%.
-CrystalRouter is more sensitive to allocation unit size.
-The best unit size would be the one big enough to accommodate its local neighborhood.
-Figure \ref{fig:interjob-cr} shows unit size of 16 and 8 can guarantee the same average data transfer time, while some rank spend more time with allocation unit size 8 than 16.
-When the unit size reduce to 2, the communication become less efficiency and takes about 15\% more times for transferring data.
-
-The data transfer time of MultiGrid with different allocation unit sizes doesn't show obvious variability in Figure \ref{fig:interjob-mg}.
-This is because even big allocation unit size like 16 will still fail to preserve MultiGrid's ``Many-to-Many" pattern.
-The data transfer time is almost doubled when MultiGrid running concurrently with allocation unit size of 16, as shown in Figure \ref{fig:interjob-mg} .
-As the unit size decreases, the data transfer time keep growing, but in a slow way, which means in terms of preserving the locality of MultiGrid, allocation unit of size 16, 8 and 2 serve equally bad.
-
-The unit size in non-contiguous allocation policy should be chosen according to application's communication pattern.
-Inter-job interference is inevitable in non-contiguous based systems, but unit size that big enough to preserve the neighborhood communication of the application will alleviate such interference and improve job performance.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/inter-job/amg}
-        \caption{AMG}
-        \label{fig:interjob-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/inter-job/cr}
-        \caption{CrystalRouter}
-        \label{fig:interjob-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/inter-job/mg}
-        \caption{MultiGrid}
-        \label{fig:interjob-mg}
-    \end{subfigure}%
-   \caption{Inter-job interference study.
-``cont" indicates three applications running side by side concurrently on the same network with contiguous allocation.
-To study the impact of non-contiguous allocation on inter-job interference, applications running concurrently with interleaved allocation of different unit sizes, which are 16-node, 8-node, 2-node.
-}
-   \label{fig:interjobstudy}
-\end{figure*}
-
-\subsection{Result Summary}
-\label{sec:summary}
-
-Based on the above comprehensive study, we can make following observations:
-\begin{itemize}
-    \item The compact allocation may not guarantee the best performance for every application. 
-    \item The applications dominated with Nearest Neighbor communications exhibit relatively stable performance under different allocation shapes, while the applications dominated with Many-to-Many communication exhibit better performance with compact allocation (e.g.
-3D balanced).
-    \item A good rank-to-node mapping strategy can greatly improve application's performance when specific allocation is given.
-    \item An optimal size for allocation units should be determined according to an application's dominant communication pattern.
-In general, a unit size should be large enough to preserve neighboring communication of the application.
-    \item Inter-job interference is inevitable in non-contiguous allocation.
-However, choosing the proper allocation unit size with job's communication pattern awareness can alleviate such negative effect.
-\end{itemize}
-
 \section{Discussion}
@@
 The design of such a smart scheduler is part of our future work.
-
-\section{Related Work}
-\label{sec:related_work}
-
-There are many tools available for system monitoring and application profiling.
-Tools like TAU (Tuning and Analysis) \cite{tau}, SCALASCA \cite{scala} and mpiP \cite{mpip} can capture application's runtime information about application's communication and computation in event traces.
-However, recognizing the patterns about application's communication from those traces requires substantial effort.
-There are a number of studies on the recognition and characterization of parallel application communication patterns.
-Oak Ridge National Laboratory has its ongoing project about developing a tool set named Oxbow, which can characterize the computation and communication behavior of scientific applications and benchmarks \cite{oxbow}.
-In a recent work \cite{roth}, the authors demonstrate a new approach to automatically characterizing parallel applications' communication behaviors.
-
-%\textbf{P2: communication characteristics}\\
-
-Many research efforts have been conducted to characterize scientific applications.
-For instance, a DOE project aims at the identification of the computational characteristics of the DOE MiniApps developed at various exascale co-design centers \cite{designforwardwebpage}.
-In this project, the communication patterns of several DOE full applications and associated mini-applications are studied to provide a more complete snapshot of the DOE workload.
-A joint project named CORAL from Oak Ridge, Argonne and Livermore provide a series of benchmarks to represent DOE workloads and technical requirements \cite{coral}.
-The CORAL project includes scalable science benchmarks, throughput benchmarks, data centric benchmarks, skeleton benchmarks and Micro benchmarks.
-
-%\textbf{P3: Interference between concurrently running jobs}\\
-The interference between concurrently running jobs on HPC system have been identified as major culprit for job's performance variability.
-Abhinav et al.
-found that concurrently running applications on HPC can cause interference to each other, and cause communication time varied from 36\% faster to 69\% slower.
-Such interference could come from Operating System noise, shape of the allocated partition , or other running jobs that sharing network resources \cite{abhinav-sc13}.
-Skinner et al.
-found that  there is a 2-3$\times$ slowdown in MPI$\textunderscore$Allreduce due to network contention from other jobs \cite{skinner}.
-Rosenthal et al.
-found increasing network bandwidth provide limited benefits to a number of applications.
-The applications that send mostly small messages or large messages asynchronously are not bandwidth bounded, hence, benefit only slightly or not at all from increased bandwidth \cite{rosenthal}.
-
-%\textbf{P4: Job allocation(little bit)}\\
-Several research studies focus on optimizing job allocation on HPC system to alleviate the interference between concurrently running jobs.
-Hoefler et al.
-proposed to use performance modeling techniques to analyze factors that impact the performance of parallel scientific applications \cite{hoefler-modeling}.
-However, as the scale of HPC continue to grow, the  interference of concurrently running jobs is getting worse, which is hard to be quantified by performance profiling tools alone.
-Bogdan et al.
-provide a set of guidelines of how to configure a network with Dragonfly topology for workload with Nearest Neighbor communication pattern \cite{Bogdan-hpdc14}.
-Dong et al.
-have developed simple benchmarks that conforms to four different communication patterns, namely ping-pong, nearest neighbor, broadcast and all$\textunderscore$reduce, to demonstrate the effectiveness of this highly parallel 5D torus network \cite{Dong-SC11}.
-
-Our work is different from all these works in the following ways.
-First, we focus on the dominant communication patterns rather than any specific application.
-We believe this can provide a guideline for other research work.
-Secondly, we explore the intra- and inter-job interference between concurrently running jobs, while similar work such as \cite{abhinav-sc13} only focuses on the single application's performance degradation due to network contention.
-Finally, we analyze the impact of different allocation strategies to job's communication behaviors.
-We further identify the optimal allocation strategy for each application with a specific dominant communication pattern.
-Based on our study, we claim that further scheduling should take job's communication pattern into consideration for its allocation decision making.
-
-\section{Conclusions}
-\label{sec:conclusion}
-In this work, we have studied the communication behavior of three parallel applications, namely AMG, CrystalRouter and MultiGrid.
-Each application has distinctive communication pattern, that can be representative for a group of jobs in HPC workload.
-We have used a sophisticate simulation tool named CODES from Argonne National Laboratory to simulate the running of these three parallel applications on torus network.
-The torus network provided by CODES has good fidelity and scalability.
-We have analyzed the performance of each application's communication in terms of data transfer time by simulating them running on torus networks with different bandwidth and dimensionality configurations.
-We have found that higher dimensionality of torus network would improve the performance of application with ``Many-to-Many" communication patterns, while application with intensive local communication like ``Nearest Neighbor" won't benefit much from higher dimensionality.
-
-We have analyzed the intra- and inter-job interference by simulating three applications running on 3D torus network.
-Based on our comprehensive experiments, we got five observations.
-1) The compact allocation may not guarantee the best performance for every application.
-2)The applications dominated with Nearest Neighbor communications exhibit relatively stable performance under different allocation shapes, while the applications dominated with Many-to-Many communication exhibit better performance with compact allocation (e.g.
-3D balanced).
-3) A good rank-to-node mapping strategy can greatly improve application's performance when specific allocation is given.
-4) An optimal size for allocation units should be determined according to an application's dominant communication pattern.
-In general, a unit size should be large enough to preserve neighboring communication of the application.
-5) Inter-job interference is inevitable in non-contiguous allocation.
-However, choosing the proper allocation unit size with job's communication pattern awareness can alleviate such negative effect.
-
-We believe that our finding in this work can provide valuable guidance for HPC resource management to make flexible job allocations.
-Rather than using pre-defined partitions and non-contiguous allocation, future HPC systems should assign each job with preferable resources based on job's communication pattern.
-
-\section*{Acknowledgment}
-\label{sec: ack}
-The work at Illinois Institute of Technology is supported in part by U.S.
-National Science Foundation grants CNS-1320125 and CCF-1422009.
-This work is also supported by the U.S.
-Department of Energy, Office of Science, Advanced Scientific Computing Research, under Contract DE-AC02-06CH11357.
-
-\bibliographystyle{IEEEtran}
-\bibliography{./reference}
-
-\vspace{5\baselineskip}
-
-\begin{framed}
-The submitted manuscript has been created by UChicago Argonne, LLC, Operator of Argonne National Laboratory ("Argonne").
-Argonne, a U.S.
-Department of Energy Office of Science laboratory, is operated under Contract No.
-DE-AC02-06CH11357.
-The U.S.
-Government retains for itself, and others acting on its behalf, a paid-up nonexclusive, irrevocable worldwide license in said article to reproduce, prepare derivative works, distribute copies to the public, and perform publicly and display publicly, by or on behalf of the Government.
-\end{framed}
-
-\end{document}

### 2016-07-07 interference_study.tex
split into separate tex files for each section

--- 2015-10-17 submission.tex
@@
-\documentclass[conference]{IEEEtran}
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
-\else
-
-\fi
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-\usepackage{algorithm}
-%\usepackage{algorithm2e}
-\usepackage{algpseudocode}
-\usepackage{booktabs}
-\usepackage{flushend}
-\usepackage{url}
-\usepackage{xcolor}
-\usepackage{amsmath}
-\usepackage{framed}
-\usepackage{textcomp}
-%\usepackage{subfig}
-\usepackage{graphicx}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{soul}
-\usepackage{verbatim}
-%\usepackage[labelfont=bf]{caption}
-\usepackage[font=bf,labelsep=space]{caption}
-
-\newcommand{\TODO}[1]{\hl{TODO: #1}}
-\newcommand{\NOTE}[1]{\hl{NOTE: #1}}
-\newcommand{\changed}[1]{\textcolor{red}{#1}}
-
-\begin{document}
-%
-% paper title
-% can use linebreaks \\ within to get better formatting as desired
-\title{Study of Intra- and Inter-Job Interference with Different Job Allocations}
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
-\author{
-
-\IEEEauthorblockN{Xu Yang\IEEEauthorrefmark{1}, John Jenkins\IEEEauthorrefmark{2}, Misbah Mubarak\IEEEauthorrefmark{2}, Robert B.
-Ross\IEEEauthorrefmark{2}, Zhou Zhou\IEEEauthorrefmark{1}, Zhiling Lan\IEEEauthorrefmark{1}}
-
-\IEEEauthorblockA{\IEEEauthorrefmark{1}Department of Computer Science,
-Illinois Institute of Technology,
-Chicago, Illinois, USA 60616\\
-\{xyang56, zzhou\}@hawk.iit.edu, lan@iit.edu}
-
-\IEEEauthorblockA{\IEEEauthorrefmark{2}Mathematics and Computer Science Division, Argonne National Laboratory,
-Argonne, IL, USA 60439\\
-\{jenkins,rross\}@mcs.anl.gov, mmubarak@anl.gov}
-}
-
-% make the title area
-\maketitle
-
-\begin{abstract}
-
-Network contention between concurrently running jobs on HPC systems is a primary cause of performance variability.
-Optimizing job allocation and avoiding network sharing are hence crucial to alleviate the potential degradation in performance.
-In order to do these effectively, an understanding of the interaction between applications, their communication patterns, and the system architecture are required.
-In this work, we choose three representative HPC applications from DOE Design Forward Project and conduct detailed simulations of high-performance network topologies to analyze both intra- and inter-job interference.
-By scrutinizing the communication behaviors of these applications, we identify relationships between these behaviors and the possible interference introduced by different job allocation scenarios.
-Our comprehensive analyses in this work illuminate a path towards job-communication-pattern-aware HPC resource allocation.
-
-\end{abstract}
-
-\IEEEpeerreviewmaketitle
-
-\section{Introduction} 
-\label{sec: intro}
-
-The scale of supercomputers continually grows at a breakneck pace to accommodate the computing power requirements for scientific research.
-Supercomputers have tens of thousands of nodes and serve as an irreplaceable research vehicle for scientific problems with increasing size and complexity.
-Supercomputers are usually employed as a shared resource to accommodate many parallel applications (jobs) running concurrently.
-These parallel jobs share the system infrastructure such as network and I/O bandwidth, and inevitably there is contention over these shared resources.
-As supercomputers continue to evolve, these shared resources are increasingly the bottleneck for performance.
-
-\begin{figure}[h!]
-    \centering
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/goodallocation}
-        \caption{Contiguous}
-        \label{fig:overview_sub1}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/badallocation}
-        \caption{Non-contiguous}
-        \label{fig:overview_sub2}
-    \end{subfigure}%
-   \caption{Multiple jobs running concurrently with different allocations.
-Each job is represented by a specific color.
-a) shows the effect of contiguous allocation, which reduce the inter-job interference.
-b) shows non-contiguous allocation, which may introduce both intra and inter-job interferences.
-}
-   \label{fig:overview}
-\end{figure}
-
-When submitting jobs to HPC systems, users request resources in terms of node counts and expected runtime.
-The batch scheduler is then responsible for dispatching the submitted jobs to the system.
-Typically, multiple jobs run concurrently, resulting in the shared use of resources, particularly network and storage.
-A prominent problem with this approach is network contention among concurrently running jobs.
-If the network resources of concurrently running jobs aren't isolated, the resulting sharing can cause communication variability that results in performance degradation for some or all such jobs \cite{abhinav-sc13}.
-This delay can propagate into the queueing time of the following submitted jobs, thus leading to low system throughput and utilization \cite{jose-ipdps15}.
-This adverse effect of network sharing can be mitigated by providing jobs with isolated allocations and exclusive network resources.
-
-On the widely used torus-connected HPC systems \cite{bgq,tofu,titan}, two allocation strategies are commonly used.
-The \emph{contiguous allocation} strategy assigns to each job a compact and contiguous set of computing nodes, as shown in Figure \ref{fig:overview_sub1}.
-The partition-based allocation approach used in Blue Gene series systems is an example of such a strategy \cite{bgloverview}.
-The contiguous strategy favors application performance through isolated networking within a partition and the locality that implies.
-However, this strategy can cause both internal fragmentation (when more nodes are allocated to a job than it requests) and external fragmentation (when sufficient nodes are available for a request, but they can not be allocated contiguously), therefore leading to lower system utilization than is otherwise possible.
-On the other side of the coin, the \emph{non-contiguous allocation} strategy, used by the Cray XT/XE series \cite{carl-cug}, assigns free nodes to jobs regardless of contiguity, though of course efforts are made to maximize locality.
-Figure~\ref{fig:overview_sub2} shows an example non-contiguous allocation.
-While eliminating internal and external fragmentation as seen in contiguous allocation systems, in return non-contiguous allocations introduce contention between jobs due to the interleaving of job nodes.
-The non-contiguous node allocation can significantly reduce job performance, especially for communication-intensive ones \cite{abhinav-sc13}.
-
-We envision that future HPC schedulers will adopt a flexible job allocation mechanism which combines the best of both contiguous and non-contiguous allocation strategies.
-Such a flexible mechanism would take shared resource needs of jobs into account when making allocation decisions (e.g., network).
-With knowledge and analysis of job communication patterns, it can be identified which jobs require network isolation and locality, and to what degree.
-Then, rather than allocating each job in a ``know-nothing'' manner, one may specialize allocation so that, for example, only the jobs with stringent network needs are given compact, isolated allocations, resulting in maximized utilization and minimized perceivable resource contention effects.
-
-In this work, we focus on an in-depth analysis of intra- and inter-job communication interference with different job allocation strategies on torus-connected HPC systems.
-Torus-based networks are used on six of the top 10 supercomputers on the June 2015 Top500 lists~\cite{top500}.
-The current generation of IBM Blue Gene/Q (BG/Q) supercomputers, such as Mira at Argonne Nation Laboratory and Sequoia at Lawrence Livermore National Laboratory, has its nodes connected in a 5D torus network \cite{bgq}.
-The K computer from Japan uses the ``Tofu'' system, which is a 6D mesh/torus topology \cite{tofu}.
-Titan, a Cray XK7 supercomputer located at the Oak Ridge Leadership Computing Facility (OLCF), has nodes connected in a 3D torus within the compute partition \cite{titan}.
-Although our analyses are based on torus networks, the ideas conveyed in this work is applicable to networks with different topologies.
-
-We selected three signature applications from the DOE Design Forward Project~\cite{designforwardwebpage} as examples to conduct detailed study about their communication patterns.
-We use a sophisticated simulation toolkit named CODES (standing for Co-Design of Multi-layer Exascale Storage Architectures)~\cite{Jason-2011} as a research vehicle to evaluate the performance of these applications with various allocations in a controlled environment.
-We then analyze the intra- and inter-job interference by simulating these applications running exclusively and concurrently with different allocations.
-We believe the insights presented in this work can be very useful for the design of future HPC batch job schedulers and resource managers.
-
-%\NOTE{JJ removed the list of contributions, because they are exactly the same as the preceding paragraph.}
-\begin{comment}
-The observations of this work are the following:
-
-\begin{itemize}
-    \item We study the communication behaviors and identify the communication patterns of three representative HPC applications.
-    \item We conduct detailed analysis about the intra- and inter-job interference by running three applications with different allocation strategies.
-And we find the relation between application's communication pattern and the cause of interference.
-    \item We show a novel and flexible way to make allocation with awareness of job's communication pattern, which  is every effective to alleviate the interference between concurrently running jobs.
-\TODO{do we actually do this, or provide the underpinnings for such a method?}
-\end{itemize}
-
-We believe the finds and insights presented in this work would be very useful for the design of future HPC batch job scheduler and resource management module.
-
-\end{comment}
-
-The rest of this paper is organized as follows.
-Section~\ref{sec:application study} describes the three representative applications from DOE Design Forward Project for our study.
-Section~\ref{sec:codes} talks about the use of CODES as research vehicle for our work.
-Section~\ref{sec:config study} shows the performance analysis of three applications on different torus networks.
-Section~\ref{sec:interference} provides detailed analysis about the intra- and inter-job interference between three applications on torus network with different allocation strategies.
-Section~\ref{sec:discussion} introduces a path toward communication-pattern aware allocation strategies, given the results of our analysis.
-Section~\ref{sec:related_work} discusses related work.
-Finally, the conclusion is presented in Section~\ref{sec:conclusion}.
-
-\section{Application Study}
-\label{sec:application study}
-
-%A parallel application usually conforms to a combination of several basic communication patterns\cite{roth}.
-At its different execution phases, the application's communication behavior may follow different basic patterns respectively.
-\textcolor{blue}{When we look into the data flow during application execution, most parallel applications start with broadcast operation to distribute the data from root process to other processes\textcolor{red}{ref needed here}, followed by a series of computation and communication that conforms to certain pattern, which is usually the dominant part of application's execution.
-Before the application come to completion, all the working process will return their results to the root directly or hierarchically.} Parallel application's communication pattern usually involves lots of factors, such as communication intensity, operation-to-operation dependencies and the ``critical path" in its communication topology graph, etc.
-In this work, we focus on the locality of application's rank-to-rank communication topology graph.
-
-A parallel application usually conforms to a combination of several basic communication patterns \cite{roth}.
-At its different execution phases, the application's communication behavior may follow different basic patterns respectively.
-There are many profiling tools available to capture information regarding communication patterns of parallel applications~\cite{tau,mpip,scala,sst,oxbow}.
-They can provide information such as the percentage of different MPI operations of the applications, communication topology, the amount of data transferred between processes.
-
-There are many applications running on HPC systems, such as Mira \cite{bgq} and Titan \cite{titan}, that are communication intensive and conform to various communication patterns.
-In this work, we select three representative applications from the DOE Design Forward Project.
-Each application exhibits a distinctive communication pattern that is commonly seen in HPC applications.
-We believe that the communication patterns of these applications are representative of a wide array of applications running on leadership-class machines.
-Specifically, we study the Algebraic MultiGrid Solver (AMG), Geometric MultiGrid (MultiGrid) and CrystalRouter MiniApps.
-The communication pattern figures we present in Section~\ref{sec:amg}, \ref{sec:crystalrouter}, and \ref{sec:multigrid} are generated with the IPM~\cite{ipm} data gathered from publicly available traces~\cite{designforwardwebpage}.
-
-% MiniApps are reduced proxy applications that encapsulate the salient performance of larger full size applications \cite{miniapp}.
-Since we only focus on the dominant communication pattern of parallel applications, these MiniApps are perfect candidates for analysis the application's performance on different allocations or on new network architectures.
-We choose three MiniApps, namely, Algebraic Multigrid Solver(AMG), Geometric MultiGrid(MultiGrid) and CrystalRouter.
-Their representative communication patterns that are commonly seen in the HPC system workload.
-The communication pattern figures we present here are generated with the IPM \cite{ipm} data from \cite{designforwardwebpage}.
-
-\subsection{AMG}
-\label{sec:amg}
-
-\begin{figure}[t!]
-    \centering
-    \begin{subfigure}[t]{0.25\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/appstudy/amg/amg_pip}
-        \caption{}
-        \label{fig:amg-communication-topology}
-    \end{subfigure}%
-    ~
-    \begin{subfigure}[t]{0.24\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/appstudy/amg/amg_data_transfer}
-        \caption{}
-        \label{fig:amg-data-trans}
-    \end{subfigure}
-    \caption{AMG.
-(a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in AMG.
-(b) The amount of data sent/received by each rank in AMG.
-Y-axis labels the amount of data transferred in MB, x-axis labels 216 ranks of AMG.
-}
-\end{figure}
-
-The Algebraic MultiGrid Solver, or AMG, is a parallel algebraic multi-grid solver for linear systems arising from problems on unstructured mesh physics packages.
-It has been derived directly from the BoomerAMG solver that is being developed in the Center for Applied Scientific Computing (CASC) at LLNL \cite{amg}.
-The dominant communication pattern is regional communication with decreasing message size for different parts of the multi-grid v-cycle.
-
-Figure~\ref{fig:amg-communication-topology} shows the communication topology of a small scale AMG execution with 216 MPI Ranks.
-Note that the dominant communication pattern of the application doesn't change with scale.
-We can make the observation that AMG's dominant communication pattern is 3D Nearest Neighbor, each rank has intensive communication with up to six neighbors, depending on the boundaries of the ranks.
-Applications with similar pattern include PARTISN \cite{partisn} and SNAP \cite{snap}.
-
-Figure~\ref{fig:amg-data-trans} shows the amount of data each rank transferred in the example execution.
-The data transferring of each rank is between 7MB to 16MB, corresponding to whether the rank is on a boundary or not.
-
-\subsection{Crystal Router}
-\label{sec:crystalrouter}
-
-\begin{figure}[t!]
-    \centering
-    \begin{subfigure}[t]{0.25\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/appstudy/cr/cr_pip}
-        \caption{Communication Topology}
-        \label{fig:cr-communication-topology}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[t]{0.22\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/appstudy/cr/cr_data_transfer}
-        \caption{Data Volume}
-        \label{fig:cr-data-trans}
-    \end{subfigure}
-    \caption{CrystalRouter.
-(a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in CrystalRouter.
-(b) The amount of data sent/received by each rank in CrystalRouter.
-Y-axis labels the amount of data transferred in MB, x-axis labels 100 ranks of CrystalRouter.
-}
-\end{figure}
-
-The second MiniApp covered is CrystalRouter, the extracted communication kernel of the full application Nek5000~\cite{crystalrouter}, which is a spectral element CFD application developed at Argonne National Laboratory \cite{nek5000}.
-It features spectral element multi-grid solvers coupled with a highly scalable, parallel coarse-grid solver that is widely used for projects including ocean current modeling, thermal hydraulics of reactor cores, and spatiotemporal chaos.
-CrystalRouter demonstrates the ``many-to-many'' communication pattern through a scalable multi-stage communication process.
-The way CrystalRouter works is nodes compute for a while, then synchronize and communicate, continually alternating between these two types of activities.
-
-The collective communication in CrystalRouter utilizes a recursive doubling approach.
-Ranks in CrystalRouter conform to a $n$-dimensional hypercube and recursively split into ($n$-1)-dimensional hypercubes, with communication occurring along the splitting plane.
-The pattern of this communication can be found in Figure~\ref{fig:cr-communication-topology}.
-By nature of the logarithmic splitting process, a substantial portion of the communication occurs in small neighborhoods of ranks.
-CrystalRouter represents a group of applications whose dominant communication is a hybrid of multi-stage local and hierarchical global communication, and shares similarities with most MPI collective communication implementations.
-
-Figure~\ref{fig:cr-data-trans} shows the amount of data each rank transferred in CrystalRouter, which ranges between 20MB to 36MB.
-This is due to the rank counts not matching one-for-one with points in the hypercube.
-
-\subsection{MultiGrid}
-\label{sec:multigrid}
-
-\begin{figure}[t!]
-    \centering
-    \begin{subfigure}[t]{0.25\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/appstudy/mg/mg_pip}
-        \caption{Communication Topology}
-        \label{fig:mg-communication-topology}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[t]{0.22\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/appstudy/mg/mg_data_transfer}
-        \caption{Data Volume}
-        \label{fig:mg-data-trans}
-    \end{subfigure}
-    \caption{MultiGrid.
-(a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in MultiGrid.
-(b) The amount of data sent/received by each rank in MultiGrid.
-Y-axis labels the amount of data transferred in MB, x-axis label 125 ranks of MultiGrid.
-}
-\end{figure}
-
-MultiGrid is geometric multi-grid v-cycle from the production elliptic solver BoxLib, a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes \cite{boxlib}.
-MultiGrid conforms to many-to-many communication pattern with decreasing message size and collectives for different parts of the multi-grid v-cycle.
-It is widely used for structured grid physics packages.
-
-Figure~\ref{fig:mg-communication-topology} shows the communication topology of MultiGrid with 125 ranks.
-We can see intensive communication along the diagonal that resembles nearest neighbor communication, similar to AMG.
-However, the communication topology leads to a greater ``spread'' of communication across the set of ranks, challenging the maximization of communication locality with respect to rank.
-In this sense, it can be considered a ``many-to-many'' pattern.
-Applications with such similar dominant communication pattern are like FillBoundary, another PDE solver code in \cite{boxlib}.
-
-Figure~\ref{fig:mg-data-trans} shows the amount of data transferred between ranks in MultiGrid. For this miniapp, The data transferred per rank is relatively small and stable, between 2.8MB to 2.9MB.
-
-\section{Research Vehicle}
-\label{sec:codes}
-
-It is difficult to accurately and flexibly experiment with concurrently running jobs in an HPC context.
-One reason is that the allocation strategy used on production machine is part of the system software, which can not be changed by users.
-Even the system administrator is not authorized to make that change.
-Another reason is that it is unrealistic to reserve the system exclusively to run the same job with desired allocation without interference and then compare the results with those in the presence of interference.
-Therefore, we resort to simulation for this work.
-
-A simulation toolkit named CODES enables the exploration of simulating different HPC networks with high fidelity \cite{Jason-2011}\cite{mubarak-sc2012}.
-CODES is built on top of Rensselaer Optimistic Simulation System (ROSS) parallel discrete-event simulator, which is capable of processing billions of events per second on leadership-class supercomputers \cite{ross}.
-CODES support both torus and dragonfly network with high fidelity flit-level simulation.
-CODES has this network workload component that is capable of conducting trace-driven simulations.
-It can take real MPI application traces generated by SST DUMPI \cite{sst} to drive CODES network models.
-In this work, we focus on an in-depth analysis of intra- and inter-job communication interferences with  different job allocations on torus-connected HPC systems.
-Torus networks have been extensively used in the current generation of supercomputers because of their linear scaling on per-node cost and competitive communication performance \cite{mubarak-sc2012}\cite{zhou-ipdps}.
-
-\section{Configuration Study}
-\label{sec:config study}
-
-In this section, we study the communication behavior of three applications on torus networks with different dimensionality and bandwidth configurations.
-Torus networks are $k$-ary $n$-cubes, with $k^{n}$ nodes in total arranged in an $n$-dimensional grid having $k$ nodes in each dimension.
-Each node has $2\times n$ direct linked neighbor nodes with wraparound.
-In these and all following experiments, non-adaptive routing is used.
-While making specific recommendations as to an ``optimal'' network configuration for a given application is out of the scope of this study, the experiments do highlight the nuances inherent in design choices such as network diameter and bandwidth planning.
-
-We perform three sets of experiments.
-Each looks at per-rank distribution of performance of AMG, CrystalRouter and MultiGrid on a set of 2048-node torus models, one with $n=3$ ($16^{2} \times 8$), one with $n=5$ ($8\times 4^{4}$), and one with $n=7$ ($4^{4} \times 2^{3}$).
-Performance in this context is defined by the total message transfer time for each sending rank, which we are able to precisely measure due to the simulation environment providing us with an absolute notion of time.
-Note that we simply perform linear rank assignment, one per node -- later sections investigate rank and node mapping effects.
-In the first, shown in Figure~\ref{fig:dimensionality-study}, we keep the link bandwidth constant at 2GiB/s in each direction, corresponding to Blue Gene/Q systems~\cite{bgq}, leading to aggregate node bandwidths of 12GiB/s, 20GiB/s, and 28GiB/s for the 3D, 5D, and 7D tori, respectively.
-In the second, shown in Figure~\ref{fig:bandwidth-study}, we fix the torus dimensionality to 5D.
-In the third, shown in Figure~\ref{fig: bandwidth-time-box}, we instead fix the per-node aggregate bandwidth (and thus the system aggregate bandwidth) and modify the torus dimensionality, resulting in per-link bandwidths shown in Table~\ref{tab: fix-bandwidth}.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/dimenstudy/amg_box}
-        \caption{AMG}
-        \label{fig:dimen-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/dimenstudy/cr_box}
-        \caption{CrystalRouter}
-        \label{fig:dimen-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/dimenstudy/mg_box}
-        \caption{MultiGrid}
-        \label{fig:dimen-mg}
-    \end{subfigure}%
-   \caption{Data Transfer Time of AMG, CrystalRouter and MultiGrid on 3D, 5D and 7D torus network, with constant per-link bandwidth.}
-   \label{fig:dimensionality-study}
-\end{figure*}
-
-In the fixed link bandwidth experiment, shown in Figure~\ref{fig:dimensionality-study}, the data transfer time of the three applications are significantly reduced as the dimensionality increases, so the applications can easily make use of the increased dimensionality (lower hops on average) and/or aggregate bandwidth (more links being concurrently used).
-Interestingly, AMG has the greatest degree of outliers compared to the others.
-We expect this to be a combination of non-optimal rank mappings causing a greater degree of variance due to the highly regular communication pattern and the distribution of message sizes.
-MultiGrid seems to benefit the most in terms of reducing the spread of the time distribution, though the median message time remains roughly within a factor of two.
-The time distribution in CrystalRouter seems to accrue the most benefits, likely due to the combination of cross-cutting and localized per-rank traffic.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/bandwidthstudy/amg_bw_box}
-        \caption{AMG}
-        \label{fig:bdwstudy-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/bandwidthstudy/cr_bw_box}
-        \caption{CrystalRouter}
-        \label{fig:bdwstudy-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/bandwidthstudy/mg_bw_box}
-        \caption{MultiGrid}
-        \label{fig:bdwstudy-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 5D torus network with 2GiB/s, 4GiB/s and GiB/s direct link bandwidth. }
-   \label{fig:bandwidth-study}
-\end{figure*}
-
-In the fixed dimensionality experiment, shown in Figure~\ref{fig:bandwidth-study}, the increased per-link bandwidth, expectedly, benefits each application roughly equally, reducing the time distributions linearly.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/samebdw/amg}
-        \caption{AMG}
-        \label{fig:samebd-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/samebdw/cr}
-        \caption{CrystalRouter}
-        \label{fig:samebd-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/samebdw/mg}
-        \caption{MultiGrid}
-        \label{fig:samebd-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 3D, 5D and 7D torus network with same aggregated bandwidth. }
-   \label{fig: bandwidth-time-box}
-\end{figure*}
-
-\begin{table}[ht]
-\begin{center}
-\caption{Torus networks with same aggregated bandwidth} 
-\label{tab: fix-bandwidth}
-\begin{tabular}{l c c c} 
-\toprule % Top horizontal line
-\toprule
-&\multicolumn{3}{c}{Bandwidth } \\
-\cmidrule(l){2-4}
-Dimension  & per-link & aggregate &\\ % Column names row
-\midrule % In-table horizontal line
-3D      & 4.67GiB/s  & 28GiB/s  & \\  % Content row 1
-\midrule % In-table horizontal line
-5D    & 2.8GiB/s  & 28GiB/s  &\\ % Content row 1
-\midrule % In-table horizontal line
-7D    & 2GiB/s & 28GiB/s & \\
-\midrule
-\bottomrule % Bottom horizontal line
-\end{tabular}
-\end{center}
-\end{table}
-
-In the fixed aggregate bandwidth experiment, with configuration in Table~\ref{tab: fix-bandwidth} and results shown in Figure~\ref{fig: bandwidth-time-box}, the results are more nuanced.
-At least for these applications, the benefits from increased dimensionality do not outweigh the reduction in per-link bandwidth.
-MultiGrid does see benefits in the increased dimensionality, particularly with respect to outliers, but has a similar median message time.
-CrystalRouter reduces performance as the dimensionality increases, but only by a small degree.
-AMG remains roughly the same, minus a dip in the maximum message time that could be due to rank-to-node mapping changes.
-
 \section{Interference Analysis}
@@
 \end{itemize}
-
-\section{Discussion}
-\label{sec:discussion}
-
-The results shown in the previous section provide insights for the design of smart and flexible job allocation.
-By scrutinizing job's communication behavior, we can identify job's dominant communication pattern and pinpoint the ``neighborhood communication".
-With such knowledge about jobs' communication patterns, we can analyze the possible interference between jobs and take precautions to alleviate the negative effect when making allocation decisions.
-
-When an application with dominant communication that is intensive ``Many-to-Many", the scheduler should grant it with compact node allocation and exclusive network provision.
-The compact allocation can guarantee the shortest pair-wise distance between all the ranks, and make the data transfer between ranks take less hops.
-On the other hand, the exclusive network provision will prevent other adjacent applications from sharing network resources, thus eliminate the performance degradation due to interference.
-
-Not every application's preferable resource are compact node allocation and exclusive network provision.
-As we found in our analysis, applications whose dominant communication patterns contain intensive ``neighborhood communication" like ``Nearest Neighbor" don't benefit from compact node allocation and exclusive network provision.
-Applications, such as AMG, can run with non-contiguous node allocation without performance degradation, as long as the allocation unit can accommodate their ``neighborhood communication".
-
-The allocation unit size shouldn't be fixed.
-Instead, the scheduler should choose a proper unit size based on the scale of ``neighborhood communication" of each job.
-The best unit size should be neither too big nor too small, jut perfectly fit for the size of that ``neighborhood".
-Big unit size won't be fully utilized and also cause fragmentation.
-Small unit size won't be able to accommodate the ``neighborhood communication" and makes the intra-job communication less efficient.
-
-The advantage of making allocation with consideration about job's communication pattern is obvious.
-First, compared with contiguous allocation policy, the new scheduler outlined above is more flexible, it doesn't require the system to provide a big contiguous partition to accommodate the whole application, instead it only needs to provide a small set of compact nodes that is sufficient for all the "local community" in the application.
-On the other hand, the new scheduler outlined above won't destroy the locality of application's communication pattern.
-Indeed, the small compact node set (allocation unit) provided for the application's  ``local communities" can preserve the communication locality in the maximum extent.
-The design of such a smart scheduler is part of our future work.
-
-\section{Related Work}
-\label{sec:related_work}
-
-There are many tools available for system monitoring and application profiling.
-Tools like TAU (Tuning and Analysis) \cite{tau}, SCALASCA \cite{scala} and mpiP \cite{mpip} can capture application's runtime information about application's communication and computation in event traces.
-However, recognizing the patterns about application's communication from those traces requires substantial effort.
-There are a number of studies on the recognition and characterization of parallel application communication patterns.
-Oak Ridge National Laboratory has its ongoing project about developing a tool set named Oxbow, which can characterize the computation and communication behavior of scientific applications and benchmarks \cite{oxbow}.
-In a recent work \cite{roth}, the authors demonstrate a new approach to automatically characterizing parallel applications' communication behaviors.
-
-%\textbf{P2: communication characteristics}\\
-
-Many research efforts have been conducted to characterize scientific applications.
-For instance, a DOE project aims at the identification of the computational characteristics of the DOE MiniApps developed at various exascale co-design centers \cite{designforwardwebpage}.
-In this project, the communication patterns of several DOE full applications and associated mini-applications are studied to provide a more complete snapshot of the DOE workload.
-A joint project named CORAL from Oak Ridge, Argonne and Livermore provide a series of benchmarks to represent DOE workloads and technical requirements \cite{coral}.
-The CORAL project includes scalable science benchmarks, throughput benchmarks, data centric benchmarks, skeleton benchmarks and Micro benchmarks.
-
-%\textbf{P3: Interference between concurrently running jobs}\\
-The interference between concurrently running jobs on HPC system have been identified as major culprit for job's performance variability.
-Abhinav et al.
-found that concurrently running applications on HPC can cause interference to each other, and cause communication time varied from 36\% faster to 69\% slower.
-Such interference could come from Operating System noise, shape of the allocated partition , or other running jobs that sharing network resources \cite{abhinav-sc13}.
-Skinner et al.
-found that  there is a 2-3$\times$ slowdown in MPI$\textunderscore$Allreduce due to network contention from other jobs \cite{skinner}.
-Rosenthal et al.
-found increasing network bandwidth provide limited benefits to a number of applications.
-The applications that send mostly small messages or large messages asynchronously are not bandwidth bounded, hence, benefit only slightly or not at all from increased bandwidth \cite{rosenthal}.
-
-%\textbf{P4: Job allocation(little bit)}\\
-Several research studies focus on optimizing job allocation on HPC system to alleviate the interference between concurrently running jobs.
-Hoefler et al.
-proposed to use performance modeling techniques to analyze factors that impact the performance of parallel scientific applications \cite{hoefler-modeling}.
-However, as the scale of HPC continue to grow, the  interference of concurrently running jobs is getting worse, which is hard to be quantified by performance profiling tools alone.
-Bogdan et al.
-provide a set of guidelines of how to configure a network with Dragonfly topology for workload with Nearest Neighbor communication pattern \cite{Bogdan-hpdc14}.
-Dong et al.
-have developed simple benchmarks that conforms to four different communication patterns, namely ping-pong, nearest neighbor, broadcast and all$\textunderscore$reduce, to demonstrate the effectiveness of this highly parallel 5D torus network \cite{Dong-SC11}.
-
-Our work is different from all these works in the following ways.
-First, we focus on the dominant communication patterns rather than any specific application.
-We believe this can provide a guideline for other research work.
-Secondly, we explore the intra- and inter-job interference between concurrently running jobs, while similar work such as \cite{abhinav-sc13} only focuses on the single application's performance degradation due to network contention.
-Finally, we analyze the impact of different allocation strategies to job's communication behaviors.
-We further identify the optimal allocation strategy for each application with a specific dominant communication pattern.
-Based on our study, we claim that further scheduling should take job's communication pattern into consideration for its allocation decision making.
-
-\section{Conclusions}
-\label{sec:conclusion}
-In this work, we have studied the communication behavior of three parallel applications, namely AMG, CrystalRouter and MultiGrid.
-Each application has distinctive communication pattern, that can be representative for a group of jobs in HPC workload.
-We have used a sophisticate simulation tool named CODES from Argonne National Laboratory to simulate the running of these three parallel applications on torus network.
-The torus network provided by CODES has good fidelity and scalability.
-We have analyzed the performance of each application's communication in terms of data transfer time by simulating them running on torus networks with different bandwidth and dimensionality configurations.
-We have found that higher dimensionality of torus network would improve the performance of application with ``Many-to-Many" communication patterns, while application with intensive local communication like ``Nearest Neighbor" won't benefit much from higher dimensionality.
-
-We have analyzed the intra- and inter-job interference by simulating three applications running on 3D torus network.
-Based on our comprehensive experiments, we got five observations.
-1) The compact allocation may not guarantee the best performance for every application.
-2)The applications dominated with Nearest Neighbor communications exhibit relatively stable performance under different allocation shapes, while the applications dominated with Many-to-Many communication exhibit better performance with compact allocation (e.g.
-3D balanced).
-3) A good rank-to-node mapping strategy can greatly improve application's performance when specific allocation is given.
-4) An optimal size for allocation units should be determined according to an application's dominant communication pattern.
-In general, a unit size should be large enough to preserve neighboring communication of the application.
-5) Inter-job interference is inevitable in non-contiguous allocation.
-However, choosing the proper allocation unit size with job's communication pattern awareness can alleviate such negative effect.
-
-We believe that our finding in this work can provide valuable guidance for HPC resource management to make flexible job allocations.
-Rather than using pre-defined partitions and non-contiguous allocation, future HPC systems should assign each job with preferable resources based on job's communication pattern.
-
-\section*{Acknowledgment}
-\label{sec: ack}
-The work at Illinois Institute of Technology is supported in part by U.S.
-National Science Foundation grants CNS-1320125 and CCF-1422009.
-This work is also supported by the U.S.
-Department of Energy, Office of Science, Advanced Scientific Computing Research, under Contract DE-AC02-06CH11357.
-
-\bibliographystyle{IEEEtran}
-\bibliography{./reference}
-
-\vspace{5\baselineskip}
-
-\begin{framed}
-The submitted manuscript has been created by UChicago Argonne, LLC, Operator of Argonne National Laboratory ("Argonne").
-Argonne, a U.S.
-Department of Energy Office of Science laboratory, is operated under Contract No.
-DE-AC02-06CH11357.
-The U.S.
-Government retains for itself, and others acting on its behalf, a paid-up nonexclusive, irrevocable worldwide license in said article to reproduce, prepare derivative works, distribute copies to the public, and perform publicly and display publicly, by or on behalf of the Government.
-\end{framed}
-
-\end{document}

### 2016-07-07 intro.tex
split into separate tex files for each section

--- 2015-10-17 submission.tex
@@
-\documentclass[conference]{IEEEtran}
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
-\else
-
-\fi
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-\usepackage{algorithm}
-%\usepackage{algorithm2e}
-\usepackage{algpseudocode}
-\usepackage{booktabs}
-\usepackage{flushend}
-\usepackage{url}
-\usepackage{xcolor}
-\usepackage{amsmath}
-\usepackage{framed}
-\usepackage{textcomp}
-%\usepackage{subfig}
-\usepackage{graphicx}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{soul}
-\usepackage{verbatim}
-%\usepackage[labelfont=bf]{caption}
-\usepackage[font=bf,labelsep=space]{caption}
-
-\newcommand{\TODO}[1]{\hl{TODO: #1}}
-\newcommand{\NOTE}[1]{\hl{NOTE: #1}}
-\newcommand{\changed}[1]{\textcolor{red}{#1}}
-
-\begin{document}
-%
-% paper title
-% can use linebreaks \\ within to get better formatting as desired
-\title{Study of Intra- and Inter-Job Interference with Different Job Allocations}
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
-\author{
-
-\IEEEauthorblockN{Xu Yang\IEEEauthorrefmark{1}, John Jenkins\IEEEauthorrefmark{2}, Misbah Mubarak\IEEEauthorrefmark{2}, Robert B.
-Ross\IEEEauthorrefmark{2}, Zhou Zhou\IEEEauthorrefmark{1}, Zhiling Lan\IEEEauthorrefmark{1}}
-
-\IEEEauthorblockA{\IEEEauthorrefmark{1}Department of Computer Science,
-Illinois Institute of Technology,
-Chicago, Illinois, USA 60616\\
-\{xyang56, zzhou\}@hawk.iit.edu, lan@iit.edu}
-
-\IEEEauthorblockA{\IEEEauthorrefmark{2}Mathematics and Computer Science Division, Argonne National Laboratory,
-Argonne, IL, USA 60439\\
-\{jenkins,rross\}@mcs.anl.gov, mmubarak@anl.gov}
-}
-
-% make the title area
-\maketitle
-
-\begin{abstract}
-
-Network contention between concurrently running jobs on HPC systems is a primary cause of performance variability.
-Optimizing job allocation and avoiding network sharing are hence crucial to alleviate the potential degradation in performance.
-In order to do these effectively, an understanding of the interaction between applications, their communication patterns, and the system architecture are required.
-In this work, we choose three representative HPC applications from DOE Design Forward Project and conduct detailed simulations of high-performance network topologies to analyze both intra- and inter-job interference.
-By scrutinizing the communication behaviors of these applications, we identify relationships between these behaviors and the possible interference introduced by different job allocation scenarios.
-Our comprehensive analyses in this work illuminate a path towards job-communication-pattern-aware HPC resource allocation.
-
-\end{abstract}
-
-\IEEEpeerreviewmaketitle
-
 \section{Introduction} 
@@
 Finally, the conclusion is presented in Section~\ref{sec:conclusion}.
-
-\section{Application Study}
-\label{sec:application study}
-
-%A parallel application usually conforms to a combination of several basic communication patterns\cite{roth}.
-At its different execution phases, the application's communication behavior may follow different basic patterns respectively.
-\textcolor{blue}{When we look into the data flow during application execution, most parallel applications start with broadcast operation to distribute the data from root process to other processes\textcolor{red}{ref needed here}, followed by a series of computation and communication that conforms to certain pattern, which is usually the dominant part of application's execution.
-Before the application come to completion, all the working process will return their results to the root directly or hierarchically.} Parallel application's communication pattern usually involves lots of factors, such as communication intensity, operation-to-operation dependencies and the ``critical path" in its communication topology graph, etc.
-In this work, we focus on the locality of application's rank-to-rank communication topology graph.
-
-A parallel application usually conforms to a combination of several basic communication patterns \cite{roth}.
-At its different execution phases, the application's communication behavior may follow different basic patterns respectively.
-There are many profiling tools available to capture information regarding communication patterns of parallel applications~\cite{tau,mpip,scala,sst,oxbow}.
-They can provide information such as the percentage of different MPI operations of the applications, communication topology, the amount of data transferred between processes.
-
-There are many applications running on HPC systems, such as Mira \cite{bgq} and Titan \cite{titan}, that are communication intensive and conform to various communication patterns.
-In this work, we select three representative applications from the DOE Design Forward Project.
-Each application exhibits a distinctive communication pattern that is commonly seen in HPC applications.
-We believe that the communication patterns of these applications are representative of a wide array of applications running on leadership-class machines.
-Specifically, we study the Algebraic MultiGrid Solver (AMG), Geometric MultiGrid (MultiGrid) and CrystalRouter MiniApps.
-The communication pattern figures we present in Section~\ref{sec:amg}, \ref{sec:crystalrouter}, and \ref{sec:multigrid} are generated with the IPM~\cite{ipm} data gathered from publicly available traces~\cite{designforwardwebpage}.
-
-% MiniApps are reduced proxy applications that encapsulate the salient performance of larger full size applications \cite{miniapp}.
-Since we only focus on the dominant communication pattern of parallel applications, these MiniApps are perfect candidates for analysis the application's performance on different allocations or on new network architectures.
-We choose three MiniApps, namely, Algebraic Multigrid Solver(AMG), Geometric MultiGrid(MultiGrid) and CrystalRouter.
-Their representative communication patterns that are commonly seen in the HPC system workload.
-The communication pattern figures we present here are generated with the IPM \cite{ipm} data from \cite{designforwardwebpage}.
-
-\subsection{AMG}
-\label{sec:amg}
-
-\begin{figure}[t!]
-    \centering
-    \begin{subfigure}[t]{0.25\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/appstudy/amg/amg_pip}
-        \caption{}
-        \label{fig:amg-communication-topology}
-    \end{subfigure}%
-    ~
-    \begin{subfigure}[t]{0.24\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/appstudy/amg/amg_data_transfer}
-        \caption{}
-        \label{fig:amg-data-trans}
-    \end{subfigure}
-    \caption{AMG.
-(a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in AMG.
-(b) The amount of data sent/received by each rank in AMG.
-Y-axis labels the amount of data transferred in MB, x-axis labels 216 ranks of AMG.
-}
-\end{figure}
-
-The Algebraic MultiGrid Solver, or AMG, is a parallel algebraic multi-grid solver for linear systems arising from problems on unstructured mesh physics packages.
-It has been derived directly from the BoomerAMG solver that is being developed in the Center for Applied Scientific Computing (CASC) at LLNL \cite{amg}.
-The dominant communication pattern is regional communication with decreasing message size for different parts of the multi-grid v-cycle.
-
-Figure~\ref{fig:amg-communication-topology} shows the communication topology of a small scale AMG execution with 216 MPI Ranks.
-Note that the dominant communication pattern of the application doesn't change with scale.
-We can make the observation that AMG's dominant communication pattern is 3D Nearest Neighbor, each rank has intensive communication with up to six neighbors, depending on the boundaries of the ranks.
-Applications with similar pattern include PARTISN \cite{partisn} and SNAP \cite{snap}.
-
-Figure~\ref{fig:amg-data-trans} shows the amount of data each rank transferred in the example execution.
-The data transferring of each rank is between 7MB to 16MB, corresponding to whether the rank is on a boundary or not.
-
-\subsection{Crystal Router}
-\label{sec:crystalrouter}
-
-\begin{figure}[t!]
-    \centering
-    \begin{subfigure}[t]{0.25\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/appstudy/cr/cr_pip}
-        \caption{Communication Topology}
-        \label{fig:cr-communication-topology}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[t]{0.22\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/appstudy/cr/cr_data_transfer}
-        \caption{Data Volume}
-        \label{fig:cr-data-trans}
-    \end{subfigure}
-    \caption{CrystalRouter.
-(a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in CrystalRouter.
-(b) The amount of data sent/received by each rank in CrystalRouter.
-Y-axis labels the amount of data transferred in MB, x-axis labels 100 ranks of CrystalRouter.
-}
-\end{figure}
-
-The second MiniApp covered is CrystalRouter, the extracted communication kernel of the full application Nek5000~\cite{crystalrouter}, which is a spectral element CFD application developed at Argonne National Laboratory \cite{nek5000}.
-It features spectral element multi-grid solvers coupled with a highly scalable, parallel coarse-grid solver that is widely used for projects including ocean current modeling, thermal hydraulics of reactor cores, and spatiotemporal chaos.
-CrystalRouter demonstrates the ``many-to-many'' communication pattern through a scalable multi-stage communication process.
-The way CrystalRouter works is nodes compute for a while, then synchronize and communicate, continually alternating between these two types of activities.
-
-The collective communication in CrystalRouter utilizes a recursive doubling approach.
-Ranks in CrystalRouter conform to a $n$-dimensional hypercube and recursively split into ($n$-1)-dimensional hypercubes, with communication occurring along the splitting plane.
-The pattern of this communication can be found in Figure~\ref{fig:cr-communication-topology}.
-By nature of the logarithmic splitting process, a substantial portion of the communication occurs in small neighborhoods of ranks.
-CrystalRouter represents a group of applications whose dominant communication is a hybrid of multi-stage local and hierarchical global communication, and shares similarities with most MPI collective communication implementations.
-
-Figure~\ref{fig:cr-data-trans} shows the amount of data each rank transferred in CrystalRouter, which ranges between 20MB to 36MB.
-This is due to the rank counts not matching one-for-one with points in the hypercube.
-
-\subsection{MultiGrid}
-\label{sec:multigrid}
-
-\begin{figure}[t!]
-    \centering
-    \begin{subfigure}[t]{0.25\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/appstudy/mg/mg_pip}
-        \caption{Communication Topology}
-        \label{fig:mg-communication-topology}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[t]{0.22\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/appstudy/mg/mg_data_transfer}
-        \caption{Data Volume}
-        \label{fig:mg-data-trans}
-    \end{subfigure}
-    \caption{MultiGrid.
-(a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in MultiGrid.
-(b) The amount of data sent/received by each rank in MultiGrid.
-Y-axis labels the amount of data transferred in MB, x-axis label 125 ranks of MultiGrid.
-}
-\end{figure}
-
-MultiGrid is geometric multi-grid v-cycle from the production elliptic solver BoxLib, a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes \cite{boxlib}.
-MultiGrid conforms to many-to-many communication pattern with decreasing message size and collectives for different parts of the multi-grid v-cycle.
-It is widely used for structured grid physics packages.
-
-Figure~\ref{fig:mg-communication-topology} shows the communication topology of MultiGrid with 125 ranks.
-We can see intensive communication along the diagonal that resembles nearest neighbor communication, similar to AMG.
-However, the communication topology leads to a greater ``spread'' of communication across the set of ranks, challenging the maximization of communication locality with respect to rank.
-In this sense, it can be considered a ``many-to-many'' pattern.
-Applications with such similar dominant communication pattern are like FillBoundary, another PDE solver code in \cite{boxlib}.
-
-Figure~\ref{fig:mg-data-trans} shows the amount of data transferred between ranks in MultiGrid. For this miniapp, The data transferred per rank is relatively small and stable, between 2.8MB to 2.9MB.
-
-\section{Research Vehicle}
-\label{sec:codes}
-
-It is difficult to accurately and flexibly experiment with concurrently running jobs in an HPC context.
-One reason is that the allocation strategy used on production machine is part of the system software, which can not be changed by users.
-Even the system administrator is not authorized to make that change.
-Another reason is that it is unrealistic to reserve the system exclusively to run the same job with desired allocation without interference and then compare the results with those in the presence of interference.
-Therefore, we resort to simulation for this work.
-
-A simulation toolkit named CODES enables the exploration of simulating different HPC networks with high fidelity \cite{Jason-2011}\cite{mubarak-sc2012}.
-CODES is built on top of Rensselaer Optimistic Simulation System (ROSS) parallel discrete-event simulator, which is capable of processing billions of events per second on leadership-class supercomputers \cite{ross}.
-CODES support both torus and dragonfly network with high fidelity flit-level simulation.
-CODES has this network workload component that is capable of conducting trace-driven simulations.
-It can take real MPI application traces generated by SST DUMPI \cite{sst} to drive CODES network models.
-In this work, we focus on an in-depth analysis of intra- and inter-job communication interferences with  different job allocations on torus-connected HPC systems.
-Torus networks have been extensively used in the current generation of supercomputers because of their linear scaling on per-node cost and competitive communication performance \cite{mubarak-sc2012}\cite{zhou-ipdps}.
-
-\section{Configuration Study}
-\label{sec:config study}
-
-In this section, we study the communication behavior of three applications on torus networks with different dimensionality and bandwidth configurations.
-Torus networks are $k$-ary $n$-cubes, with $k^{n}$ nodes in total arranged in an $n$-dimensional grid having $k$ nodes in each dimension.
-Each node has $2\times n$ direct linked neighbor nodes with wraparound.
-In these and all following experiments, non-adaptive routing is used.
-While making specific recommendations as to an ``optimal'' network configuration for a given application is out of the scope of this study, the experiments do highlight the nuances inherent in design choices such as network diameter and bandwidth planning.
-
-We perform three sets of experiments.
-Each looks at per-rank distribution of performance of AMG, CrystalRouter and MultiGrid on a set of 2048-node torus models, one with $n=3$ ($16^{2} \times 8$), one with $n=5$ ($8\times 4^{4}$), and one with $n=7$ ($4^{4} \times 2^{3}$).
-Performance in this context is defined by the total message transfer time for each sending rank, which we are able to precisely measure due to the simulation environment providing us with an absolute notion of time.
-Note that we simply perform linear rank assignment, one per node -- later sections investigate rank and node mapping effects.
-In the first, shown in Figure~\ref{fig:dimensionality-study}, we keep the link bandwidth constant at 2GiB/s in each direction, corresponding to Blue Gene/Q systems~\cite{bgq}, leading to aggregate node bandwidths of 12GiB/s, 20GiB/s, and 28GiB/s for the 3D, 5D, and 7D tori, respectively.
-In the second, shown in Figure~\ref{fig:bandwidth-study}, we fix the torus dimensionality to 5D.
-In the third, shown in Figure~\ref{fig: bandwidth-time-box}, we instead fix the per-node aggregate bandwidth (and thus the system aggregate bandwidth) and modify the torus dimensionality, resulting in per-link bandwidths shown in Table~\ref{tab: fix-bandwidth}.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/dimenstudy/amg_box}
-        \caption{AMG}
-        \label{fig:dimen-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/dimenstudy/cr_box}
-        \caption{CrystalRouter}
-        \label{fig:dimen-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/dimenstudy/mg_box}
-        \caption{MultiGrid}
-        \label{fig:dimen-mg}
-    \end{subfigure}%
-   \caption{Data Transfer Time of AMG, CrystalRouter and MultiGrid on 3D, 5D and 7D torus network, with constant per-link bandwidth.}
-   \label{fig:dimensionality-study}
-\end{figure*}
-
-In the fixed link bandwidth experiment, shown in Figure~\ref{fig:dimensionality-study}, the data transfer time of the three applications are significantly reduced as the dimensionality increases, so the applications can easily make use of the increased dimensionality (lower hops on average) and/or aggregate bandwidth (more links being concurrently used).
-Interestingly, AMG has the greatest degree of outliers compared to the others.
-We expect this to be a combination of non-optimal rank mappings causing a greater degree of variance due to the highly regular communication pattern and the distribution of message sizes.
-MultiGrid seems to benefit the most in terms of reducing the spread of the time distribution, though the median message time remains roughly within a factor of two.
-The time distribution in CrystalRouter seems to accrue the most benefits, likely due to the combination of cross-cutting and localized per-rank traffic.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/bandwidthstudy/amg_bw_box}
-        \caption{AMG}
-        \label{fig:bdwstudy-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/bandwidthstudy/cr_bw_box}
-        \caption{CrystalRouter}
-        \label{fig:bdwstudy-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/bandwidthstudy/mg_bw_box}
-        \caption{MultiGrid}
-        \label{fig:bdwstudy-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 5D torus network with 2GiB/s, 4GiB/s and GiB/s direct link bandwidth. }
-   \label{fig:bandwidth-study}
-\end{figure*}
-
-In the fixed dimensionality experiment, shown in Figure~\ref{fig:bandwidth-study}, the increased per-link bandwidth, expectedly, benefits each application roughly equally, reducing the time distributions linearly.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/samebdw/amg}
-        \caption{AMG}
-        \label{fig:samebd-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/samebdw/cr}
-        \caption{CrystalRouter}
-        \label{fig:samebd-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/samebdw/mg}
-        \caption{MultiGrid}
-        \label{fig:samebd-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 3D, 5D and 7D torus network with same aggregated bandwidth. }
-   \label{fig: bandwidth-time-box}
-\end{figure*}
-
-\begin{table}[ht]
-\begin{center}
-\caption{Torus networks with same aggregated bandwidth} 
-\label{tab: fix-bandwidth}
-\begin{tabular}{l c c c} 
-\toprule % Top horizontal line
-\toprule
-&\multicolumn{3}{c}{Bandwidth } \\
-\cmidrule(l){2-4}
-Dimension  & per-link & aggregate &\\ % Column names row
-\midrule % In-table horizontal line
-3D      & 4.67GiB/s  & 28GiB/s  & \\  % Content row 1
-\midrule % In-table horizontal line
-5D    & 2.8GiB/s  & 28GiB/s  &\\ % Content row 1
-\midrule % In-table horizontal line
-7D    & 2GiB/s & 28GiB/s & \\
-\midrule
-\bottomrule % Bottom horizontal line
-\end{tabular}
-\end{center}
-\end{table}
-
-In the fixed aggregate bandwidth experiment, with configuration in Table~\ref{tab: fix-bandwidth} and results shown in Figure~\ref{fig: bandwidth-time-box}, the results are more nuanced.
-At least for these applications, the benefits from increased dimensionality do not outweigh the reduction in per-link bandwidth.
-MultiGrid does see benefits in the increased dimensionality, particularly with respect to outliers, but has a similar median message time.
-CrystalRouter reduces performance as the dimensionality increases, but only by a small degree.
-AMG remains roughly the same, minus a dip in the maximum message time that could be due to rank-to-node mapping changes.
-
-\section{Interference Analysis}
-\label{sec:interference}
-
-\textbf{\emph{Intra-job interference}} refers to the network contention between ranks within each application.
-\textbf{\emph{Inter-job interference}} is introduced by concurrently running jobs that adjacent to each other sharing network resources.
-Communication variability due to such interference can cause application performance degradation.
-
-In this section we will study both kinds of interference in the context of torus network. 
-Since application's communication pattern won't change with its scale, we use AMG with 216 ranks, CrystalRouter with 100 ranks, and MultiGrid 125 ranks.
-And in order to accommodate three applications running concurrently with different allocations, we simulate a 2K-node 3D torus ($16^{2} \times 8$) network for the experiments.
-
-\subsection{Intra-Job Interference Analysis}
-\label{sec: introjob}
-
-We design two sets of experiments to study the intra-job interference of each application.
-In the first, shown in Figure \ref{fig:shapestudy}, we assign each application with allocations in \emph{three different shapes}.
-In the second, shown in Figure \ref{fig:diffmap}, we study the intra-job interference by using different mapping strategies for each application with given allocation shape.
-
-In the allocation shapes experiment, we select three possible shapes can can commonly seen on the 3D torus network.
-They are 3D balanced-cube, 3D unbalanced-cube, 2D mesh, as shown in Figure \ref{fig:alloc-shapes}.
-
-3D balanced-cube, shown as red in Figure \ref{fig:cont_sub1}, can guarantee the minimum \emph{average pair-wise distance} within the allocation.
-Some research studies like \cite{leung,abhinav-sc13} indicate that compact allocation can guarantee jobs with better performance.
-They use a variety of metrics to evaluate the compactness of the allocation, such as \emph{average pair-wise distance}, \emph{diameter} and \emph{contiguity}.
-In this work, we select 3D balanced-cube as the most compact allocation on a 3D torus network.
-
-3D unbalanced-cube, shown as green in Figure \ref{fig:cont_sub1}, is a rectangular prism, which is the possible allocation shape on the systems with asymmetric networks.
-For example, Cray XE6/XK7 systems are deployed 3D tori with Gemini routers.
-The network connections in the \emph{y}-direction have only half the bandwidth of the cables used in the \emph{x} and \emph{z} directions.
-In order to take advantage of the faster links in the \emph{x} and \emph{z} directions, job allocation starts from X-Z plane, which leads to a rectangular prism shaped allocation \cite{RF}.
-
-2D mesh, shown as blue in Figure \ref{fig:cont_sub1}, can be cut out from a single layer of the 3D torus.
-2D mesh is a very common allocation shape in torus network for both contiguous and non-contiguous allocation strategies.
-For example, Cray Application Level Placement Scheduler (ALPS) indexes the nodes in torus network into a list and makes job allocation by simply going through the list \cite{carl-cug}.
-When the list is obtained by sorting the nodes based on their spacial coordinates in the torus, allocation made off from this list will be 2D mesh.
-The IBM Blue/Gene Q supercomputer Mira at Argonne Leadership Computing Facility also allow its allocation partition configured into mesh \cite{zhou-ipdps}.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/allocation}
-        \caption{ }
-        \label{fig:cont_sub1}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit16}
-        \caption{ }
-        \label{fig:noncont_sub1}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit8}
-        \caption{ }
-        \label{fig:noncont_sub2}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit2}
-        \caption{ }
-        \label{fig:noncont_sub3}
-    \end{subfigure}%
-    \caption{(a) Contiguous allocation in three different shapes.
-``Blue" is 2D mesh, ``Green" is 3D-unbalanced cube and ``Red" is 3D-balanced cube.
-(b)-(d) Non-contiguous allocation.
-Each job is represented by a specific color.
-The nodes assigned to different jobs are interleaved, the size of allocation unit are 16, 8 and 2.}
-    \label{fig:alloc-shapes}
-\end{figure*}
-
-AMG with 3D ``Nearest Neighbor" communication pattern takes less time when being assigned with 3D-unbalanced allocation than 3D-balanced, as shown in Figure \ref{fig:shapstudy-amg}.
-And MultiGrid gets the best performance (shortest data transfer time) when running on 3D-balanced allocation, as shown in Figure \ref{fig:shapstudy-mg}.
-Since MultiGrid's communication pattern is ``Many-to-Many" dominant, 3D-balanced is the most compact allocation with shortest pair-wise distance between nodes, which can reduce the aggregated hops for transferring message among ranks in MultiGrid.
-As CrystalRouter has both dominant local communication and multi-stage global data transfer, the 3D-balance is also the best allocation, but its advantage over 3D-unbalanced is not that obvious as to MultiGrid, as shown in Figure \ref{fig:shapstudy-cr}.
-
-Lots of research studies have designed fancy allocation algorithms to provide application with the most compact allocation\cite{leung,LO}.
-However, as shown in our experiments, providing cubical allocation without considering application's communication pattern won't guarantee the best performance for every application.
-Compact allocation should be provided to application with global data transfer, such as those conforms to ``Many-to-Many" communication pattern.
-Application with dominant communication pattern like ``Nearest Neighbor" won't fully utilize the compact allocation and could get even better performance on a relaxed allocation shape.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/amg_box}
-        \caption{AMG}
-        \label{fig:shapstudy-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/cr_box}
-        \caption{CrystalRouter}
-        \label{fig:shapstudy-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/mg_box}
-        \caption{MultiGrid}
-        \label{fig:shapstudy-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 3D-balanced, 3D-unbalanced and 2D allocation.}
-   \label{fig:shapestudy}
-\end{figure*}
-
-The rank-to-node mapping of parallel applications on HPC system could greatly impact applications' performance.
-However, finding the optimal mapping solution for a given application is out the scope of this work.
-Our experiments aims to show how mapping strategies impact the intra-job interference of application with specific communication pattern.
-
-We provide AMG, CrystalRouter and MultiGrid with 3D-balance allocation and use three mapping strategies to do the rank-to-node mapping.
-The ``Linear" mapping, as we used in the previous experiments, is to map each rank according to the dimensional ordering of compute nodes.
-The ``Cube" mapping assigns ranks into consecutive $2^{3}$ cubes.
-The ``Random" mapping assigns ranks randomly within the allocation.
-
-AMG remains roughly the same when it been mapped by ``Linear" and ``Cube", shown in Figure \ref{fig:diffmap-amg}.
-Although ``Linear" and ``Cube" mapping strategies cause some routing overlap for AMG's communication, both mapping strategies still preserve the locality of AMG's 3D ``Nearest Neighbor" communication pattern.
-The ``Random" mapping totally destroys AMG's communication pattern, and results in intra-job interference among all the ranks in AMG.
-The performance degradation caused by using ``Random" mapping is up to 90\%.
-
-The ``Cube" mapping improves CrystalRouter's performance over the ``Linear" by reducing the pair-wise distance between ranks, shown in Figure \ref{fig:diffmap-cr}.
-The global data transfer in CrystalRouter will takes fewer hops with ``Cube" mapping.
-The ``Random" mapping scatters the ranks within the local neighborhood of CrystalRouter and make their communication less efficient.
-
-The ``Cube" mapping benefits MultiGrid's ``Many-to-Many" communication.
-However, due to the small amount of data transfer among ranks in MultiGrid, ``Cube" mapping doesn't exhibit a drastic advantage.
-``Random" mapping neither cause too much degradation for the same reason, shown in Figure \ref{fig:diffmap-mg}.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/diffmapping/amg}
-        \caption{AMG}
-        \label{fig:diffmap-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/diffmapping/cr}
-        \caption{CrystalRouter}
-        \label{fig:diffmap-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/diffmapping/mg}
-        \caption{MultiGrid}
-        \label{fig:diffmap-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 3D-balanced allocation using  different mapping strategies.}
-   \label{fig:diffmap}
-\end{figure*}
-
-\subsection{Inter-Job Interference Analysis}
-\label{sec: interjob interference study}
-
-Inter-job interference has been identified as one of the major culprits for application's performance variability \cite{abhinav-sc13,skinner,rosenthal}.
-Inter-job interference is a more prominent issue for system adopted non-contiguous allocation policy than system with contiguous allocation.
-Application communication times have been demonstrated to vary from 36\% faster to 69\% slower due to job interference when running on non-contiguous allocation \cite{abhinav-sc13}.
-
-We assign each application with non-contiguous allocation and run them concurrently on the same network.
-The allocation unit belongs to different jobs are interleaved.
-In order to study the impact of different allocation unit size on applications' inter-job interference, we conduct experiments with unit size of 16, 8 and 2, shown respectively in Figure \ref{fig:noncont_sub1}, \ref{fig:noncont_sub1} and \ref{fig:noncont_sub3}.
-Figure \ref{fig:interjobstudy} shows the results of each application data transfer time with different allocation unit size.
-
-The data transfer time of AMG in Figure \ref{fig:interjob-amg} keeps stable between allocation unit size of 16 and 8, since the locality of AMG is 6-rank based.
-When the unit size reduce to 2, AMG suffers prolong data transfer time by about 10\%.
-CrystalRouter is more sensitive to allocation unit size.
-The best unit size would be the one big enough to accommodate its local neighborhood.
-Figure \ref{fig:interjob-cr} shows unit size of 16 and 8 can guarantee the same average data transfer time, while some rank spend more time with allocation unit size 8 than 16.
-When the unit size reduce to 2, the communication become less efficiency and takes about 15\% more times for transferring data.
-
-The data transfer time of MultiGrid with different allocation unit sizes doesn't show obvious variability in Figure \ref{fig:interjob-mg}.
-This is because even big allocation unit size like 16 will still fail to preserve MultiGrid's ``Many-to-Many" pattern.
-The data transfer time is almost doubled when MultiGrid running concurrently with allocation unit size of 16, as shown in Figure \ref{fig:interjob-mg} .
-As the unit size decreases, the data transfer time keep growing, but in a slow way, which means in terms of preserving the locality of MultiGrid, allocation unit of size 16, 8 and 2 serve equally bad.
-
-The unit size in non-contiguous allocation policy should be chosen according to application's communication pattern.
-Inter-job interference is inevitable in non-contiguous based systems, but unit size that big enough to preserve the neighborhood communication of the application will alleviate such interference and improve job performance.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/inter-job/amg}
-        \caption{AMG}
-        \label{fig:interjob-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/inter-job/cr}
-        \caption{CrystalRouter}
-        \label{fig:interjob-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/inter-job/mg}
-        \caption{MultiGrid}
-        \label{fig:interjob-mg}
-    \end{subfigure}%
-   \caption{Inter-job interference study.
-``cont" indicates three applications running side by side concurrently on the same network with contiguous allocation.
-To study the impact of non-contiguous allocation on inter-job interference, applications running concurrently with interleaved allocation of different unit sizes, which are 16-node, 8-node, 2-node.
-}
-   \label{fig:interjobstudy}
-\end{figure*}
-
-\subsection{Result Summary}
-\label{sec:summary}
-
-Based on the above comprehensive study, we can make following observations:
-\begin{itemize}
-    \item The compact allocation may not guarantee the best performance for every application. 
-    \item The applications dominated with Nearest Neighbor communications exhibit relatively stable performance under different allocation shapes, while the applications dominated with Many-to-Many communication exhibit better performance with compact allocation (e.g.
-3D balanced).
-    \item A good rank-to-node mapping strategy can greatly improve application's performance when specific allocation is given.
-    \item An optimal size for allocation units should be determined according to an application's dominant communication pattern.
-In general, a unit size should be large enough to preserve neighboring communication of the application.
-    \item Inter-job interference is inevitable in non-contiguous allocation.
-However, choosing the proper allocation unit size with job's communication pattern awareness can alleviate such negative effect.
-\end{itemize}
-
-\section{Discussion}
-\label{sec:discussion}
-
-The results shown in the previous section provide insights for the design of smart and flexible job allocation.
-By scrutinizing job's communication behavior, we can identify job's dominant communication pattern and pinpoint the ``neighborhood communication".
-With such knowledge about jobs' communication patterns, we can analyze the possible interference between jobs and take precautions to alleviate the negative effect when making allocation decisions.
-
-When an application with dominant communication that is intensive ``Many-to-Many", the scheduler should grant it with compact node allocation and exclusive network provision.
-The compact allocation can guarantee the shortest pair-wise distance between all the ranks, and make the data transfer between ranks take less hops.
-On the other hand, the exclusive network provision will prevent other adjacent applications from sharing network resources, thus eliminate the performance degradation due to interference.
-
-Not every application's preferable resource are compact node allocation and exclusive network provision.
-As we found in our analysis, applications whose dominant communication patterns contain intensive ``neighborhood communication" like ``Nearest Neighbor" don't benefit from compact node allocation and exclusive network provision.
-Applications, such as AMG, can run with non-contiguous node allocation without performance degradation, as long as the allocation unit can accommodate their ``neighborhood communication".
-
-The allocation unit size shouldn't be fixed.
-Instead, the scheduler should choose a proper unit size based on the scale of ``neighborhood communication" of each job.
-The best unit size should be neither too big nor too small, jut perfectly fit for the size of that ``neighborhood".
-Big unit size won't be fully utilized and also cause fragmentation.
-Small unit size won't be able to accommodate the ``neighborhood communication" and makes the intra-job communication less efficient.
-
-The advantage of making allocation with consideration about job's communication pattern is obvious.
-First, compared with contiguous allocation policy, the new scheduler outlined above is more flexible, it doesn't require the system to provide a big contiguous partition to accommodate the whole application, instead it only needs to provide a small set of compact nodes that is sufficient for all the "local community" in the application.
-On the other hand, the new scheduler outlined above won't destroy the locality of application's communication pattern.
-Indeed, the small compact node set (allocation unit) provided for the application's  ``local communities" can preserve the communication locality in the maximum extent.
-The design of such a smart scheduler is part of our future work.
-
-\section{Related Work}
-\label{sec:related_work}
-
-There are many tools available for system monitoring and application profiling.
-Tools like TAU (Tuning and Analysis) \cite{tau}, SCALASCA \cite{scala} and mpiP \cite{mpip} can capture application's runtime information about application's communication and computation in event traces.
-However, recognizing the patterns about application's communication from those traces requires substantial effort.
-There are a number of studies on the recognition and characterization of parallel application communication patterns.
-Oak Ridge National Laboratory has its ongoing project about developing a tool set named Oxbow, which can characterize the computation and communication behavior of scientific applications and benchmarks \cite{oxbow}.
-In a recent work \cite{roth}, the authors demonstrate a new approach to automatically characterizing parallel applications' communication behaviors.
-
-%\textbf{P2: communication characteristics}\\
-
-Many research efforts have been conducted to characterize scientific applications.
-For instance, a DOE project aims at the identification of the computational characteristics of the DOE MiniApps developed at various exascale co-design centers \cite{designforwardwebpage}.
-In this project, the communication patterns of several DOE full applications and associated mini-applications are studied to provide a more complete snapshot of the DOE workload.
-A joint project named CORAL from Oak Ridge, Argonne and Livermore provide a series of benchmarks to represent DOE workloads and technical requirements \cite{coral}.
-The CORAL project includes scalable science benchmarks, throughput benchmarks, data centric benchmarks, skeleton benchmarks and Micro benchmarks.
-
-%\textbf{P3: Interference between concurrently running jobs}\\
-The interference between concurrently running jobs on HPC system have been identified as major culprit for job's performance variability.
-Abhinav et al.
-found that concurrently running applications on HPC can cause interference to each other, and cause communication time varied from 36\% faster to 69\% slower.
-Such interference could come from Operating System noise, shape of the allocated partition , or other running jobs that sharing network resources \cite{abhinav-sc13}.
-Skinner et al.
-found that  there is a 2-3$\times$ slowdown in MPI$\textunderscore$Allreduce due to network contention from other jobs \cite{skinner}.
-Rosenthal et al.
-found increasing network bandwidth provide limited benefits to a number of applications.
-The applications that send mostly small messages or large messages asynchronously are not bandwidth bounded, hence, benefit only slightly or not at all from increased bandwidth \cite{rosenthal}.
-
-%\textbf{P4: Job allocation(little bit)}\\
-Several research studies focus on optimizing job allocation on HPC system to alleviate the interference between concurrently running jobs.
-Hoefler et al.
-proposed to use performance modeling techniques to analyze factors that impact the performance of parallel scientific applications \cite{hoefler-modeling}.
-However, as the scale of HPC continue to grow, the  interference of concurrently running jobs is getting worse, which is hard to be quantified by performance profiling tools alone.
-Bogdan et al.
-provide a set of guidelines of how to configure a network with Dragonfly topology for workload with Nearest Neighbor communication pattern \cite{Bogdan-hpdc14}.
-Dong et al.
-have developed simple benchmarks that conforms to four different communication patterns, namely ping-pong, nearest neighbor, broadcast and all$\textunderscore$reduce, to demonstrate the effectiveness of this highly parallel 5D torus network \cite{Dong-SC11}.
-
-Our work is different from all these works in the following ways.
-First, we focus on the dominant communication patterns rather than any specific application.
-We believe this can provide a guideline for other research work.
-Secondly, we explore the intra- and inter-job interference between concurrently running jobs, while similar work such as \cite{abhinav-sc13} only focuses on the single application's performance degradation due to network contention.
-Finally, we analyze the impact of different allocation strategies to job's communication behaviors.
-We further identify the optimal allocation strategy for each application with a specific dominant communication pattern.
-Based on our study, we claim that further scheduling should take job's communication pattern into consideration for its allocation decision making.
-
-\section{Conclusions}
-\label{sec:conclusion}
-In this work, we have studied the communication behavior of three parallel applications, namely AMG, CrystalRouter and MultiGrid.
-Each application has distinctive communication pattern, that can be representative for a group of jobs in HPC workload.
-We have used a sophisticate simulation tool named CODES from Argonne National Laboratory to simulate the running of these three parallel applications on torus network.
-The torus network provided by CODES has good fidelity and scalability.
-We have analyzed the performance of each application's communication in terms of data transfer time by simulating them running on torus networks with different bandwidth and dimensionality configurations.
-We have found that higher dimensionality of torus network would improve the performance of application with ``Many-to-Many" communication patterns, while application with intensive local communication like ``Nearest Neighbor" won't benefit much from higher dimensionality.
-
-We have analyzed the intra- and inter-job interference by simulating three applications running on 3D torus network.
-Based on our comprehensive experiments, we got five observations.
-1) The compact allocation may not guarantee the best performance for every application.
-2)The applications dominated with Nearest Neighbor communications exhibit relatively stable performance under different allocation shapes, while the applications dominated with Many-to-Many communication exhibit better performance with compact allocation (e.g.
-3D balanced).
-3) A good rank-to-node mapping strategy can greatly improve application's performance when specific allocation is given.
-4) An optimal size for allocation units should be determined according to an application's dominant communication pattern.
-In general, a unit size should be large enough to preserve neighboring communication of the application.
-5) Inter-job interference is inevitable in non-contiguous allocation.
-However, choosing the proper allocation unit size with job's communication pattern awareness can alleviate such negative effect.
-
-We believe that our finding in this work can provide valuable guidance for HPC resource management to make flexible job allocations.
-Rather than using pre-defined partitions and non-contiguous allocation, future HPC systems should assign each job with preferable resources based on job's communication pattern.
-
-\section*{Acknowledgment}
-\label{sec: ack}
-The work at Illinois Institute of Technology is supported in part by U.S.
-National Science Foundation grants CNS-1320125 and CCF-1422009.
-This work is also supported by the U.S.
-Department of Energy, Office of Science, Advanced Scientific Computing Research, under Contract DE-AC02-06CH11357.
-
-\bibliographystyle{IEEEtran}
-\bibliography{./reference}
-
-\vspace{5\baselineskip}
-
-\begin{framed}
-The submitted manuscript has been created by UChicago Argonne, LLC, Operator of Argonne National Laboratory ("Argonne").
-Argonne, a U.S.
-Department of Energy Office of Science laboratory, is operated under Contract No.
-DE-AC02-06CH11357.
-The U.S.
-Government retains for itself, and others acting on its behalf, a paid-up nonexclusive, irrevocable worldwide license in said article to reproduce, prepare derivative works, distribute copies to the public, and perform publicly and display publicly, by or on behalf of the Government.
-\end{framed}
-
-\end{document}

### 2016-07-07 related_work.tex
split into separate tex files for each section

--- 2015-10-17 submission.tex
@@
-\documentclass[conference]{IEEEtran}
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
-\else
-
-\fi
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-\usepackage{algorithm}
-%\usepackage{algorithm2e}
-\usepackage{algpseudocode}
-\usepackage{booktabs}
-\usepackage{flushend}
-\usepackage{url}
-\usepackage{xcolor}
-\usepackage{amsmath}
-\usepackage{framed}
-\usepackage{textcomp}
-%\usepackage{subfig}
-\usepackage{graphicx}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{soul}
-\usepackage{verbatim}
-%\usepackage[labelfont=bf]{caption}
-\usepackage[font=bf,labelsep=space]{caption}
-
-\newcommand{\TODO}[1]{\hl{TODO: #1}}
-\newcommand{\NOTE}[1]{\hl{NOTE: #1}}
-\newcommand{\changed}[1]{\textcolor{red}{#1}}
-
-\begin{document}
-%
-% paper title
-% can use linebreaks \\ within to get better formatting as desired
-\title{Study of Intra- and Inter-Job Interference with Different Job Allocations}
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
-\author{
-
-\IEEEauthorblockN{Xu Yang\IEEEauthorrefmark{1}, John Jenkins\IEEEauthorrefmark{2}, Misbah Mubarak\IEEEauthorrefmark{2}, Robert B.
-Ross\IEEEauthorrefmark{2}, Zhou Zhou\IEEEauthorrefmark{1}, Zhiling Lan\IEEEauthorrefmark{1}}
-
-\IEEEauthorblockA{\IEEEauthorrefmark{1}Department of Computer Science,
-Illinois Institute of Technology,
-Chicago, Illinois, USA 60616\\
-\{xyang56, zzhou\}@hawk.iit.edu, lan@iit.edu}
-
-\IEEEauthorblockA{\IEEEauthorrefmark{2}Mathematics and Computer Science Division, Argonne National Laboratory,
-Argonne, IL, USA 60439\\
-\{jenkins,rross\}@mcs.anl.gov, mmubarak@anl.gov}
-}
-
-% make the title area
-\maketitle
-
-\begin{abstract}
-
-Network contention between concurrently running jobs on HPC systems is a primary cause of performance variability.
-Optimizing job allocation and avoiding network sharing are hence crucial to alleviate the potential degradation in performance.
-In order to do these effectively, an understanding of the interaction between applications, their communication patterns, and the system architecture are required.
-In this work, we choose three representative HPC applications from DOE Design Forward Project and conduct detailed simulations of high-performance network topologies to analyze both intra- and inter-job interference.
-By scrutinizing the communication behaviors of these applications, we identify relationships between these behaviors and the possible interference introduced by different job allocation scenarios.
-Our comprehensive analyses in this work illuminate a path towards job-communication-pattern-aware HPC resource allocation.
-
-\end{abstract}
-
-\IEEEpeerreviewmaketitle
-
-\section{Introduction} 
-\label{sec: intro}
-
-The scale of supercomputers continually grows at a breakneck pace to accommodate the computing power requirements for scientific research.
-Supercomputers have tens of thousands of nodes and serve as an irreplaceable research vehicle for scientific problems with increasing size and complexity.
-Supercomputers are usually employed as a shared resource to accommodate many parallel applications (jobs) running concurrently.
-These parallel jobs share the system infrastructure such as network and I/O bandwidth, and inevitably there is contention over these shared resources.
-As supercomputers continue to evolve, these shared resources are increasingly the bottleneck for performance.
-
-\begin{figure}[h!]
-    \centering
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/goodallocation}
-        \caption{Contiguous}
-        \label{fig:overview_sub1}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/badallocation}
-        \caption{Non-contiguous}
-        \label{fig:overview_sub2}
-    \end{subfigure}%
-   \caption{Multiple jobs running concurrently with different allocations.
-Each job is represented by a specific color.
-a) shows the effect of contiguous allocation, which reduce the inter-job interference.
-b) shows non-contiguous allocation, which may introduce both intra and inter-job interferences.
-}
-   \label{fig:overview}
-\end{figure}
-
-When submitting jobs to HPC systems, users request resources in terms of node counts and expected runtime.
-The batch scheduler is then responsible for dispatching the submitted jobs to the system.
-Typically, multiple jobs run concurrently, resulting in the shared use of resources, particularly network and storage.
-A prominent problem with this approach is network contention among concurrently running jobs.
-If the network resources of concurrently running jobs aren't isolated, the resulting sharing can cause communication variability that results in performance degradation for some or all such jobs \cite{abhinav-sc13}.
-This delay can propagate into the queueing time of the following submitted jobs, thus leading to low system throughput and utilization \cite{jose-ipdps15}.
-This adverse effect of network sharing can be mitigated by providing jobs with isolated allocations and exclusive network resources.
-
-On the widely used torus-connected HPC systems \cite{bgq,tofu,titan}, two allocation strategies are commonly used.
-The \emph{contiguous allocation} strategy assigns to each job a compact and contiguous set of computing nodes, as shown in Figure \ref{fig:overview_sub1}.
-The partition-based allocation approach used in Blue Gene series systems is an example of such a strategy \cite{bgloverview}.
-The contiguous strategy favors application performance through isolated networking within a partition and the locality that implies.
-However, this strategy can cause both internal fragmentation (when more nodes are allocated to a job than it requests) and external fragmentation (when sufficient nodes are available for a request, but they can not be allocated contiguously), therefore leading to lower system utilization than is otherwise possible.
-On the other side of the coin, the \emph{non-contiguous allocation} strategy, used by the Cray XT/XE series \cite{carl-cug}, assigns free nodes to jobs regardless of contiguity, though of course efforts are made to maximize locality.
-Figure~\ref{fig:overview_sub2} shows an example non-contiguous allocation.
-While eliminating internal and external fragmentation as seen in contiguous allocation systems, in return non-contiguous allocations introduce contention between jobs due to the interleaving of job nodes.
-The non-contiguous node allocation can significantly reduce job performance, especially for communication-intensive ones \cite{abhinav-sc13}.
-
-We envision that future HPC schedulers will adopt a flexible job allocation mechanism which combines the best of both contiguous and non-contiguous allocation strategies.
-Such a flexible mechanism would take shared resource needs of jobs into account when making allocation decisions (e.g., network).
-With knowledge and analysis of job communication patterns, it can be identified which jobs require network isolation and locality, and to what degree.
-Then, rather than allocating each job in a ``know-nothing'' manner, one may specialize allocation so that, for example, only the jobs with stringent network needs are given compact, isolated allocations, resulting in maximized utilization and minimized perceivable resource contention effects.
-
-In this work, we focus on an in-depth analysis of intra- and inter-job communication interference with different job allocation strategies on torus-connected HPC systems.
-Torus-based networks are used on six of the top 10 supercomputers on the June 2015 Top500 lists~\cite{top500}.
-The current generation of IBM Blue Gene/Q (BG/Q) supercomputers, such as Mira at Argonne Nation Laboratory and Sequoia at Lawrence Livermore National Laboratory, has its nodes connected in a 5D torus network \cite{bgq}.
-The K computer from Japan uses the ``Tofu'' system, which is a 6D mesh/torus topology \cite{tofu}.
-Titan, a Cray XK7 supercomputer located at the Oak Ridge Leadership Computing Facility (OLCF), has nodes connected in a 3D torus within the compute partition \cite{titan}.
-Although our analyses are based on torus networks, the ideas conveyed in this work is applicable to networks with different topologies.
-
-We selected three signature applications from the DOE Design Forward Project~\cite{designforwardwebpage} as examples to conduct detailed study about their communication patterns.
-We use a sophisticated simulation toolkit named CODES (standing for Co-Design of Multi-layer Exascale Storage Architectures)~\cite{Jason-2011} as a research vehicle to evaluate the performance of these applications with various allocations in a controlled environment.
-We then analyze the intra- and inter-job interference by simulating these applications running exclusively and concurrently with different allocations.
-We believe the insights presented in this work can be very useful for the design of future HPC batch job schedulers and resource managers.
-
-%\NOTE{JJ removed the list of contributions, because they are exactly the same as the preceding paragraph.}
-\begin{comment}
-The observations of this work are the following:
-
-\begin{itemize}
-    \item We study the communication behaviors and identify the communication patterns of three representative HPC applications.
-    \item We conduct detailed analysis about the intra- and inter-job interference by running three applications with different allocation strategies.
-And we find the relation between application's communication pattern and the cause of interference.
-    \item We show a novel and flexible way to make allocation with awareness of job's communication pattern, which  is every effective to alleviate the interference between concurrently running jobs.
-\TODO{do we actually do this, or provide the underpinnings for such a method?}
-\end{itemize}
-
-We believe the finds and insights presented in this work would be very useful for the design of future HPC batch job scheduler and resource management module.
-
-\end{comment}
-
-The rest of this paper is organized as follows.
-Section~\ref{sec:application study} describes the three representative applications from DOE Design Forward Project for our study.
-Section~\ref{sec:codes} talks about the use of CODES as research vehicle for our work.
-Section~\ref{sec:config study} shows the performance analysis of three applications on different torus networks.
-Section~\ref{sec:interference} provides detailed analysis about the intra- and inter-job interference between three applications on torus network with different allocation strategies.
-Section~\ref{sec:discussion} introduces a path toward communication-pattern aware allocation strategies, given the results of our analysis.
-Section~\ref{sec:related_work} discusses related work.
-Finally, the conclusion is presented in Section~\ref{sec:conclusion}.
-
-\section{Application Study}
-\label{sec:application study}
-
-%A parallel application usually conforms to a combination of several basic communication patterns\cite{roth}.
-At its different execution phases, the application's communication behavior may follow different basic patterns respectively.
-\textcolor{blue}{When we look into the data flow during application execution, most parallel applications start with broadcast operation to distribute the data from root process to other processes\textcolor{red}{ref needed here}, followed by a series of computation and communication that conforms to certain pattern, which is usually the dominant part of application's execution.
-Before the application come to completion, all the working process will return their results to the root directly or hierarchically.} Parallel application's communication pattern usually involves lots of factors, such as communication intensity, operation-to-operation dependencies and the ``critical path" in its communication topology graph, etc.
-In this work, we focus on the locality of application's rank-to-rank communication topology graph.
-
-A parallel application usually conforms to a combination of several basic communication patterns \cite{roth}.
-At its different execution phases, the application's communication behavior may follow different basic patterns respectively.
-There are many profiling tools available to capture information regarding communication patterns of parallel applications~\cite{tau,mpip,scala,sst,oxbow}.
-They can provide information such as the percentage of different MPI operations of the applications, communication topology, the amount of data transferred between processes.
-
-There are many applications running on HPC systems, such as Mira \cite{bgq} and Titan \cite{titan}, that are communication intensive and conform to various communication patterns.
-In this work, we select three representative applications from the DOE Design Forward Project.
-Each application exhibits a distinctive communication pattern that is commonly seen in HPC applications.
-We believe that the communication patterns of these applications are representative of a wide array of applications running on leadership-class machines.
-Specifically, we study the Algebraic MultiGrid Solver (AMG), Geometric MultiGrid (MultiGrid) and CrystalRouter MiniApps.
-The communication pattern figures we present in Section~\ref{sec:amg}, \ref{sec:crystalrouter}, and \ref{sec:multigrid} are generated with the IPM~\cite{ipm} data gathered from publicly available traces~\cite{designforwardwebpage}.
-
-% MiniApps are reduced proxy applications that encapsulate the salient performance of larger full size applications \cite{miniapp}.
-Since we only focus on the dominant communication pattern of parallel applications, these MiniApps are perfect candidates for analysis the application's performance on different allocations or on new network architectures.
-We choose three MiniApps, namely, Algebraic Multigrid Solver(AMG), Geometric MultiGrid(MultiGrid) and CrystalRouter.
-Their representative communication patterns that are commonly seen in the HPC system workload.
-The communication pattern figures we present here are generated with the IPM \cite{ipm} data from \cite{designforwardwebpage}.
-
-\subsection{AMG}
-\label{sec:amg}
-
-\begin{figure}[t!]
-    \centering
-    \begin{subfigure}[t]{0.25\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/appstudy/amg/amg_pip}
-        \caption{}
-        \label{fig:amg-communication-topology}
-    \end{subfigure}%
-    ~
-    \begin{subfigure}[t]{0.24\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/appstudy/amg/amg_data_transfer}
-        \caption{}
-        \label{fig:amg-data-trans}
-    \end{subfigure}
-    \caption{AMG.
-(a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in AMG.
-(b) The amount of data sent/received by each rank in AMG.
-Y-axis labels the amount of data transferred in MB, x-axis labels 216 ranks of AMG.
-}
-\end{figure}
-
-The Algebraic MultiGrid Solver, or AMG, is a parallel algebraic multi-grid solver for linear systems arising from problems on unstructured mesh physics packages.
-It has been derived directly from the BoomerAMG solver that is being developed in the Center for Applied Scientific Computing (CASC) at LLNL \cite{amg}.
-The dominant communication pattern is regional communication with decreasing message size for different parts of the multi-grid v-cycle.
-
-Figure~\ref{fig:amg-communication-topology} shows the communication topology of a small scale AMG execution with 216 MPI Ranks.
-Note that the dominant communication pattern of the application doesn't change with scale.
-We can make the observation that AMG's dominant communication pattern is 3D Nearest Neighbor, each rank has intensive communication with up to six neighbors, depending on the boundaries of the ranks.
-Applications with similar pattern include PARTISN \cite{partisn} and SNAP \cite{snap}.
-
-Figure~\ref{fig:amg-data-trans} shows the amount of data each rank transferred in the example execution.
-The data transferring of each rank is between 7MB to 16MB, corresponding to whether the rank is on a boundary or not.
-
-\subsection{Crystal Router}
-\label{sec:crystalrouter}
-
-\begin{figure}[t!]
-    \centering
-    \begin{subfigure}[t]{0.25\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/appstudy/cr/cr_pip}
-        \caption{Communication Topology}
-        \label{fig:cr-communication-topology}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[t]{0.22\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/appstudy/cr/cr_data_transfer}
-        \caption{Data Volume}
-        \label{fig:cr-data-trans}
-    \end{subfigure}
-    \caption{CrystalRouter.
-(a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in CrystalRouter.
-(b) The amount of data sent/received by each rank in CrystalRouter.
-Y-axis labels the amount of data transferred in MB, x-axis labels 100 ranks of CrystalRouter.
-}
-\end{figure}
-
-The second MiniApp covered is CrystalRouter, the extracted communication kernel of the full application Nek5000~\cite{crystalrouter}, which is a spectral element CFD application developed at Argonne National Laboratory \cite{nek5000}.
-It features spectral element multi-grid solvers coupled with a highly scalable, parallel coarse-grid solver that is widely used for projects including ocean current modeling, thermal hydraulics of reactor cores, and spatiotemporal chaos.
-CrystalRouter demonstrates the ``many-to-many'' communication pattern through a scalable multi-stage communication process.
-The way CrystalRouter works is nodes compute for a while, then synchronize and communicate, continually alternating between these two types of activities.
-
-The collective communication in CrystalRouter utilizes a recursive doubling approach.
-Ranks in CrystalRouter conform to a $n$-dimensional hypercube and recursively split into ($n$-1)-dimensional hypercubes, with communication occurring along the splitting plane.
-The pattern of this communication can be found in Figure~\ref{fig:cr-communication-topology}.
-By nature of the logarithmic splitting process, a substantial portion of the communication occurs in small neighborhoods of ranks.
-CrystalRouter represents a group of applications whose dominant communication is a hybrid of multi-stage local and hierarchical global communication, and shares similarities with most MPI collective communication implementations.
-
-Figure~\ref{fig:cr-data-trans} shows the amount of data each rank transferred in CrystalRouter, which ranges between 20MB to 36MB.
-This is due to the rank counts not matching one-for-one with points in the hypercube.
-
-\subsection{MultiGrid}
-\label{sec:multigrid}
-
-\begin{figure}[t!]
-    \centering
-    \begin{subfigure}[t]{0.25\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/appstudy/mg/mg_pip}
-        \caption{Communication Topology}
-        \label{fig:mg-communication-topology}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[t]{0.22\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/appstudy/mg/mg_data_transfer}
-        \caption{Data Volume}
-        \label{fig:mg-data-trans}
-    \end{subfigure}
-    \caption{MultiGrid.
-(a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in MultiGrid.
-(b) The amount of data sent/received by each rank in MultiGrid.
-Y-axis labels the amount of data transferred in MB, x-axis label 125 ranks of MultiGrid.
-}
-\end{figure}
-
-MultiGrid is geometric multi-grid v-cycle from the production elliptic solver BoxLib, a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes \cite{boxlib}.
-MultiGrid conforms to many-to-many communication pattern with decreasing message size and collectives for different parts of the multi-grid v-cycle.
-It is widely used for structured grid physics packages.
-
-Figure~\ref{fig:mg-communication-topology} shows the communication topology of MultiGrid with 125 ranks.
-We can see intensive communication along the diagonal that resembles nearest neighbor communication, similar to AMG.
-However, the communication topology leads to a greater ``spread'' of communication across the set of ranks, challenging the maximization of communication locality with respect to rank.
-In this sense, it can be considered a ``many-to-many'' pattern.
-Applications with such similar dominant communication pattern are like FillBoundary, another PDE solver code in \cite{boxlib}.
-
-Figure~\ref{fig:mg-data-trans} shows the amount of data transferred between ranks in MultiGrid. For this miniapp, The data transferred per rank is relatively small and stable, between 2.8MB to 2.9MB.
-
-\section{Research Vehicle}
-\label{sec:codes}
-
-It is difficult to accurately and flexibly experiment with concurrently running jobs in an HPC context.
-One reason is that the allocation strategy used on production machine is part of the system software, which can not be changed by users.
-Even the system administrator is not authorized to make that change.
-Another reason is that it is unrealistic to reserve the system exclusively to run the same job with desired allocation without interference and then compare the results with those in the presence of interference.
-Therefore, we resort to simulation for this work.
-
-A simulation toolkit named CODES enables the exploration of simulating different HPC networks with high fidelity \cite{Jason-2011}\cite{mubarak-sc2012}.
-CODES is built on top of Rensselaer Optimistic Simulation System (ROSS) parallel discrete-event simulator, which is capable of processing billions of events per second on leadership-class supercomputers \cite{ross}.
-CODES support both torus and dragonfly network with high fidelity flit-level simulation.
-CODES has this network workload component that is capable of conducting trace-driven simulations.
-It can take real MPI application traces generated by SST DUMPI \cite{sst} to drive CODES network models.
-In this work, we focus on an in-depth analysis of intra- and inter-job communication interferences with  different job allocations on torus-connected HPC systems.
-Torus networks have been extensively used in the current generation of supercomputers because of their linear scaling on per-node cost and competitive communication performance \cite{mubarak-sc2012}\cite{zhou-ipdps}.
-
-\section{Configuration Study}
-\label{sec:config study}
-
-In this section, we study the communication behavior of three applications on torus networks with different dimensionality and bandwidth configurations.
-Torus networks are $k$-ary $n$-cubes, with $k^{n}$ nodes in total arranged in an $n$-dimensional grid having $k$ nodes in each dimension.
-Each node has $2\times n$ direct linked neighbor nodes with wraparound.
-In these and all following experiments, non-adaptive routing is used.
-While making specific recommendations as to an ``optimal'' network configuration for a given application is out of the scope of this study, the experiments do highlight the nuances inherent in design choices such as network diameter and bandwidth planning.
-
-We perform three sets of experiments.
-Each looks at per-rank distribution of performance of AMG, CrystalRouter and MultiGrid on a set of 2048-node torus models, one with $n=3$ ($16^{2} \times 8$), one with $n=5$ ($8\times 4^{4}$), and one with $n=7$ ($4^{4} \times 2^{3}$).
-Performance in this context is defined by the total message transfer time for each sending rank, which we are able to precisely measure due to the simulation environment providing us with an absolute notion of time.
-Note that we simply perform linear rank assignment, one per node -- later sections investigate rank and node mapping effects.
-In the first, shown in Figure~\ref{fig:dimensionality-study}, we keep the link bandwidth constant at 2GiB/s in each direction, corresponding to Blue Gene/Q systems~\cite{bgq}, leading to aggregate node bandwidths of 12GiB/s, 20GiB/s, and 28GiB/s for the 3D, 5D, and 7D tori, respectively.
-In the second, shown in Figure~\ref{fig:bandwidth-study}, we fix the torus dimensionality to 5D.
-In the third, shown in Figure~\ref{fig: bandwidth-time-box}, we instead fix the per-node aggregate bandwidth (and thus the system aggregate bandwidth) and modify the torus dimensionality, resulting in per-link bandwidths shown in Table~\ref{tab: fix-bandwidth}.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/dimenstudy/amg_box}
-        \caption{AMG}
-        \label{fig:dimen-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/dimenstudy/cr_box}
-        \caption{CrystalRouter}
-        \label{fig:dimen-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/dimenstudy/mg_box}
-        \caption{MultiGrid}
-        \label{fig:dimen-mg}
-    \end{subfigure}%
-   \caption{Data Transfer Time of AMG, CrystalRouter and MultiGrid on 3D, 5D and 7D torus network, with constant per-link bandwidth.}
-   \label{fig:dimensionality-study}
-\end{figure*}
-
-In the fixed link bandwidth experiment, shown in Figure~\ref{fig:dimensionality-study}, the data transfer time of the three applications are significantly reduced as the dimensionality increases, so the applications can easily make use of the increased dimensionality (lower hops on average) and/or aggregate bandwidth (more links being concurrently used).
-Interestingly, AMG has the greatest degree of outliers compared to the others.
-We expect this to be a combination of non-optimal rank mappings causing a greater degree of variance due to the highly regular communication pattern and the distribution of message sizes.
-MultiGrid seems to benefit the most in terms of reducing the spread of the time distribution, though the median message time remains roughly within a factor of two.
-The time distribution in CrystalRouter seems to accrue the most benefits, likely due to the combination of cross-cutting and localized per-rank traffic.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/bandwidthstudy/amg_bw_box}
-        \caption{AMG}
-        \label{fig:bdwstudy-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/bandwidthstudy/cr_bw_box}
-        \caption{CrystalRouter}
-        \label{fig:bdwstudy-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/bandwidthstudy/mg_bw_box}
-        \caption{MultiGrid}
-        \label{fig:bdwstudy-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 5D torus network with 2GiB/s, 4GiB/s and GiB/s direct link bandwidth. }
-   \label{fig:bandwidth-study}
-\end{figure*}
-
-In the fixed dimensionality experiment, shown in Figure~\ref{fig:bandwidth-study}, the increased per-link bandwidth, expectedly, benefits each application roughly equally, reducing the time distributions linearly.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/samebdw/amg}
-        \caption{AMG}
-        \label{fig:samebd-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/samebdw/cr}
-        \caption{CrystalRouter}
-        \label{fig:samebd-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/samebdw/mg}
-        \caption{MultiGrid}
-        \label{fig:samebd-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 3D, 5D and 7D torus network with same aggregated bandwidth. }
-   \label{fig: bandwidth-time-box}
-\end{figure*}
-
-\begin{table}[ht]
-\begin{center}
-\caption{Torus networks with same aggregated bandwidth} 
-\label{tab: fix-bandwidth}
-\begin{tabular}{l c c c} 
-\toprule % Top horizontal line
-\toprule
-&\multicolumn{3}{c}{Bandwidth } \\
-\cmidrule(l){2-4}
-Dimension  & per-link & aggregate &\\ % Column names row
-\midrule % In-table horizontal line
-3D      & 4.67GiB/s  & 28GiB/s  & \\  % Content row 1
-\midrule % In-table horizontal line
-5D    & 2.8GiB/s  & 28GiB/s  &\\ % Content row 1
-\midrule % In-table horizontal line
-7D    & 2GiB/s & 28GiB/s & \\
-\midrule
-\bottomrule % Bottom horizontal line
-\end{tabular}
-\end{center}
-\end{table}
-
-In the fixed aggregate bandwidth experiment, with configuration in Table~\ref{tab: fix-bandwidth} and results shown in Figure~\ref{fig: bandwidth-time-box}, the results are more nuanced.
-At least for these applications, the benefits from increased dimensionality do not outweigh the reduction in per-link bandwidth.
-MultiGrid does see benefits in the increased dimensionality, particularly with respect to outliers, but has a similar median message time.
-CrystalRouter reduces performance as the dimensionality increases, but only by a small degree.
-AMG remains roughly the same, minus a dip in the maximum message time that could be due to rank-to-node mapping changes.
-
-\section{Interference Analysis}
-\label{sec:interference}
-
-\textbf{\emph{Intra-job interference}} refers to the network contention between ranks within each application.
-\textbf{\emph{Inter-job interference}} is introduced by concurrently running jobs that adjacent to each other sharing network resources.
-Communication variability due to such interference can cause application performance degradation.
-
-In this section we will study both kinds of interference in the context of torus network. 
-Since application's communication pattern won't change with its scale, we use AMG with 216 ranks, CrystalRouter with 100 ranks, and MultiGrid 125 ranks.
-And in order to accommodate three applications running concurrently with different allocations, we simulate a 2K-node 3D torus ($16^{2} \times 8$) network for the experiments.
-
-\subsection{Intra-Job Interference Analysis}
-\label{sec: introjob}
-
-We design two sets of experiments to study the intra-job interference of each application.
-In the first, shown in Figure \ref{fig:shapestudy}, we assign each application with allocations in \emph{three different shapes}.
-In the second, shown in Figure \ref{fig:diffmap}, we study the intra-job interference by using different mapping strategies for each application with given allocation shape.
-
-In the allocation shapes experiment, we select three possible shapes can can commonly seen on the 3D torus network.
-They are 3D balanced-cube, 3D unbalanced-cube, 2D mesh, as shown in Figure \ref{fig:alloc-shapes}.
-
-3D balanced-cube, shown as red in Figure \ref{fig:cont_sub1}, can guarantee the minimum \emph{average pair-wise distance} within the allocation.
-Some research studies like \cite{leung,abhinav-sc13} indicate that compact allocation can guarantee jobs with better performance.
-They use a variety of metrics to evaluate the compactness of the allocation, such as \emph{average pair-wise distance}, \emph{diameter} and \emph{contiguity}.
-In this work, we select 3D balanced-cube as the most compact allocation on a 3D torus network.
-
-3D unbalanced-cube, shown as green in Figure \ref{fig:cont_sub1}, is a rectangular prism, which is the possible allocation shape on the systems with asymmetric networks.
-For example, Cray XE6/XK7 systems are deployed 3D tori with Gemini routers.
-The network connections in the \emph{y}-direction have only half the bandwidth of the cables used in the \emph{x} and \emph{z} directions.
-In order to take advantage of the faster links in the \emph{x} and \emph{z} directions, job allocation starts from X-Z plane, which leads to a rectangular prism shaped allocation \cite{RF}.
-
-2D mesh, shown as blue in Figure \ref{fig:cont_sub1}, can be cut out from a single layer of the 3D torus.
-2D mesh is a very common allocation shape in torus network for both contiguous and non-contiguous allocation strategies.
-For example, Cray Application Level Placement Scheduler (ALPS) indexes the nodes in torus network into a list and makes job allocation by simply going through the list \cite{carl-cug}.
-When the list is obtained by sorting the nodes based on their spacial coordinates in the torus, allocation made off from this list will be 2D mesh.
-The IBM Blue/Gene Q supercomputer Mira at Argonne Leadership Computing Facility also allow its allocation partition configured into mesh \cite{zhou-ipdps}.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/allocation}
-        \caption{ }
-        \label{fig:cont_sub1}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit16}
-        \caption{ }
-        \label{fig:noncont_sub1}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit8}
-        \caption{ }
-        \label{fig:noncont_sub2}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit2}
-        \caption{ }
-        \label{fig:noncont_sub3}
-    \end{subfigure}%
-    \caption{(a) Contiguous allocation in three different shapes.
-``Blue" is 2D mesh, ``Green" is 3D-unbalanced cube and ``Red" is 3D-balanced cube.
-(b)-(d) Non-contiguous allocation.
-Each job is represented by a specific color.
-The nodes assigned to different jobs are interleaved, the size of allocation unit are 16, 8 and 2.}
-    \label{fig:alloc-shapes}
-\end{figure*}
-
-AMG with 3D ``Nearest Neighbor" communication pattern takes less time when being assigned with 3D-unbalanced allocation than 3D-balanced, as shown in Figure \ref{fig:shapstudy-amg}.
-And MultiGrid gets the best performance (shortest data transfer time) when running on 3D-balanced allocation, as shown in Figure \ref{fig:shapstudy-mg}.
-Since MultiGrid's communication pattern is ``Many-to-Many" dominant, 3D-balanced is the most compact allocation with shortest pair-wise distance between nodes, which can reduce the aggregated hops for transferring message among ranks in MultiGrid.
-As CrystalRouter has both dominant local communication and multi-stage global data transfer, the 3D-balance is also the best allocation, but its advantage over 3D-unbalanced is not that obvious as to MultiGrid, as shown in Figure \ref{fig:shapstudy-cr}.
-
-Lots of research studies have designed fancy allocation algorithms to provide application with the most compact allocation\cite{leung,LO}.
-However, as shown in our experiments, providing cubical allocation without considering application's communication pattern won't guarantee the best performance for every application.
-Compact allocation should be provided to application with global data transfer, such as those conforms to ``Many-to-Many" communication pattern.
-Application with dominant communication pattern like ``Nearest Neighbor" won't fully utilize the compact allocation and could get even better performance on a relaxed allocation shape.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/amg_box}
-        \caption{AMG}
-        \label{fig:shapstudy-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/cr_box}
-        \caption{CrystalRouter}
-        \label{fig:shapstudy-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/mg_box}
-        \caption{MultiGrid}
-        \label{fig:shapstudy-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 3D-balanced, 3D-unbalanced and 2D allocation.}
-   \label{fig:shapestudy}
-\end{figure*}
-
-The rank-to-node mapping of parallel applications on HPC system could greatly impact applications' performance.
-However, finding the optimal mapping solution for a given application is out the scope of this work.
-Our experiments aims to show how mapping strategies impact the intra-job interference of application with specific communication pattern.
-
-We provide AMG, CrystalRouter and MultiGrid with 3D-balance allocation and use three mapping strategies to do the rank-to-node mapping.
-The ``Linear" mapping, as we used in the previous experiments, is to map each rank according to the dimensional ordering of compute nodes.
-The ``Cube" mapping assigns ranks into consecutive $2^{3}$ cubes.
-The ``Random" mapping assigns ranks randomly within the allocation.
-
-AMG remains roughly the same when it been mapped by ``Linear" and ``Cube", shown in Figure \ref{fig:diffmap-amg}.
-Although ``Linear" and ``Cube" mapping strategies cause some routing overlap for AMG's communication, both mapping strategies still preserve the locality of AMG's 3D ``Nearest Neighbor" communication pattern.
-The ``Random" mapping totally destroys AMG's communication pattern, and results in intra-job interference among all the ranks in AMG.
-The performance degradation caused by using ``Random" mapping is up to 90\%.
-
-The ``Cube" mapping improves CrystalRouter's performance over the ``Linear" by reducing the pair-wise distance between ranks, shown in Figure \ref{fig:diffmap-cr}.
-The global data transfer in CrystalRouter will takes fewer hops with ``Cube" mapping.
-The ``Random" mapping scatters the ranks within the local neighborhood of CrystalRouter and make their communication less efficient.
-
-The ``Cube" mapping benefits MultiGrid's ``Many-to-Many" communication.
-However, due to the small amount of data transfer among ranks in MultiGrid, ``Cube" mapping doesn't exhibit a drastic advantage.
-``Random" mapping neither cause too much degradation for the same reason, shown in Figure \ref{fig:diffmap-mg}.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/diffmapping/amg}
-        \caption{AMG}
-        \label{fig:diffmap-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/diffmapping/cr}
-        \caption{CrystalRouter}
-        \label{fig:diffmap-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/diffmapping/mg}
-        \caption{MultiGrid}
-        \label{fig:diffmap-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 3D-balanced allocation using  different mapping strategies.}
-   \label{fig:diffmap}
-\end{figure*}
-
-\subsection{Inter-Job Interference Analysis}
-\label{sec: interjob interference study}
-
-Inter-job interference has been identified as one of the major culprits for application's performance variability \cite{abhinav-sc13,skinner,rosenthal}.
-Inter-job interference is a more prominent issue for system adopted non-contiguous allocation policy than system with contiguous allocation.
-Application communication times have been demonstrated to vary from 36\% faster to 69\% slower due to job interference when running on non-contiguous allocation \cite{abhinav-sc13}.
-
-We assign each application with non-contiguous allocation and run them concurrently on the same network.
-The allocation unit belongs to different jobs are interleaved.
-In order to study the impact of different allocation unit size on applications' inter-job interference, we conduct experiments with unit size of 16, 8 and 2, shown respectively in Figure \ref{fig:noncont_sub1}, \ref{fig:noncont_sub1} and \ref{fig:noncont_sub3}.
-Figure \ref{fig:interjobstudy} shows the results of each application data transfer time with different allocation unit size.
-
-The data transfer time of AMG in Figure \ref{fig:interjob-amg} keeps stable between allocation unit size of 16 and 8, since the locality of AMG is 6-rank based.
-When the unit size reduce to 2, AMG suffers prolong data transfer time by about 10\%.
-CrystalRouter is more sensitive to allocation unit size.
-The best unit size would be the one big enough to accommodate its local neighborhood.
-Figure \ref{fig:interjob-cr} shows unit size of 16 and 8 can guarantee the same average data transfer time, while some rank spend more time with allocation unit size 8 than 16.
-When the unit size reduce to 2, the communication become less efficiency and takes about 15\% more times for transferring data.
-
-The data transfer time of MultiGrid with different allocation unit sizes doesn't show obvious variability in Figure \ref{fig:interjob-mg}.
-This is because even big allocation unit size like 16 will still fail to preserve MultiGrid's ``Many-to-Many" pattern.
-The data transfer time is almost doubled when MultiGrid running concurrently with allocation unit size of 16, as shown in Figure \ref{fig:interjob-mg} .
-As the unit size decreases, the data transfer time keep growing, but in a slow way, which means in terms of preserving the locality of MultiGrid, allocation unit of size 16, 8 and 2 serve equally bad.
-
-The unit size in non-contiguous allocation policy should be chosen according to application's communication pattern.
-Inter-job interference is inevitable in non-contiguous based systems, but unit size that big enough to preserve the neighborhood communication of the application will alleviate such interference and improve job performance.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/inter-job/amg}
-        \caption{AMG}
-        \label{fig:interjob-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/inter-job/cr}
-        \caption{CrystalRouter}
-        \label{fig:interjob-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/inter-job/mg}
-        \caption{MultiGrid}
-        \label{fig:interjob-mg}
-    \end{subfigure}%
-   \caption{Inter-job interference study.
-``cont" indicates three applications running side by side concurrently on the same network with contiguous allocation.
-To study the impact of non-contiguous allocation on inter-job interference, applications running concurrently with interleaved allocation of different unit sizes, which are 16-node, 8-node, 2-node.
-}
-   \label{fig:interjobstudy}
-\end{figure*}
-
-\subsection{Result Summary}
-\label{sec:summary}
-
-Based on the above comprehensive study, we can make following observations:
-\begin{itemize}
-    \item The compact allocation may not guarantee the best performance for every application. 
-    \item The applications dominated with Nearest Neighbor communications exhibit relatively stable performance under different allocation shapes, while the applications dominated with Many-to-Many communication exhibit better performance with compact allocation (e.g.
-3D balanced).
-    \item A good rank-to-node mapping strategy can greatly improve application's performance when specific allocation is given.
-    \item An optimal size for allocation units should be determined according to an application's dominant communication pattern.
-In general, a unit size should be large enough to preserve neighboring communication of the application.
-    \item Inter-job interference is inevitable in non-contiguous allocation.
-However, choosing the proper allocation unit size with job's communication pattern awareness can alleviate such negative effect.
-\end{itemize}
-
-\section{Discussion}
-\label{sec:discussion}
-
-The results shown in the previous section provide insights for the design of smart and flexible job allocation.
-By scrutinizing job's communication behavior, we can identify job's dominant communication pattern and pinpoint the ``neighborhood communication".
-With such knowledge about jobs' communication patterns, we can analyze the possible interference between jobs and take precautions to alleviate the negative effect when making allocation decisions.
-
-When an application with dominant communication that is intensive ``Many-to-Many", the scheduler should grant it with compact node allocation and exclusive network provision.
-The compact allocation can guarantee the shortest pair-wise distance between all the ranks, and make the data transfer between ranks take less hops.
-On the other hand, the exclusive network provision will prevent other adjacent applications from sharing network resources, thus eliminate the performance degradation due to interference.
-
-Not every application's preferable resource are compact node allocation and exclusive network provision.
-As we found in our analysis, applications whose dominant communication patterns contain intensive ``neighborhood communication" like ``Nearest Neighbor" don't benefit from compact node allocation and exclusive network provision.
-Applications, such as AMG, can run with non-contiguous node allocation without performance degradation, as long as the allocation unit can accommodate their ``neighborhood communication".
-
-The allocation unit size shouldn't be fixed.
-Instead, the scheduler should choose a proper unit size based on the scale of ``neighborhood communication" of each job.
-The best unit size should be neither too big nor too small, jut perfectly fit for the size of that ``neighborhood".
-Big unit size won't be fully utilized and also cause fragmentation.
-Small unit size won't be able to accommodate the ``neighborhood communication" and makes the intra-job communication less efficient.
-
-The advantage of making allocation with consideration about job's communication pattern is obvious.
-First, compared with contiguous allocation policy, the new scheduler outlined above is more flexible, it doesn't require the system to provide a big contiguous partition to accommodate the whole application, instead it only needs to provide a small set of compact nodes that is sufficient for all the "local community" in the application.
-On the other hand, the new scheduler outlined above won't destroy the locality of application's communication pattern.
-Indeed, the small compact node set (allocation unit) provided for the application's  ``local communities" can preserve the communication locality in the maximum extent.
-The design of such a smart scheduler is part of our future work.
-
 \section{Related Work}
@@
 Based on our study, we claim that further scheduling should take job's communication pattern into consideration for its allocation decision making.
-
-\section{Conclusions}
-\label{sec:conclusion}
-In this work, we have studied the communication behavior of three parallel applications, namely AMG, CrystalRouter and MultiGrid.
-Each application has distinctive communication pattern, that can be representative for a group of jobs in HPC workload.
-We have used a sophisticate simulation tool named CODES from Argonne National Laboratory to simulate the running of these three parallel applications on torus network.
-The torus network provided by CODES has good fidelity and scalability.
-We have analyzed the performance of each application's communication in terms of data transfer time by simulating them running on torus networks with different bandwidth and dimensionality configurations.
-We have found that higher dimensionality of torus network would improve the performance of application with ``Many-to-Many" communication patterns, while application with intensive local communication like ``Nearest Neighbor" won't benefit much from higher dimensionality.
-
-We have analyzed the intra- and inter-job interference by simulating three applications running on 3D torus network.
-Based on our comprehensive experiments, we got five observations.
-1) The compact allocation may not guarantee the best performance for every application.
-2)The applications dominated with Nearest Neighbor communications exhibit relatively stable performance under different allocation shapes, while the applications dominated with Many-to-Many communication exhibit better performance with compact allocation (e.g.
-3D balanced).
-3) A good rank-to-node mapping strategy can greatly improve application's performance when specific allocation is given.
-4) An optimal size for allocation units should be determined according to an application's dominant communication pattern.
-In general, a unit size should be large enough to preserve neighboring communication of the application.
-5) Inter-job interference is inevitable in non-contiguous allocation.
-However, choosing the proper allocation unit size with job's communication pattern awareness can alleviate such negative effect.
-
-We believe that our finding in this work can provide valuable guidance for HPC resource management to make flexible job allocations.
-Rather than using pre-defined partitions and non-contiguous allocation, future HPC systems should assign each job with preferable resources based on job's communication pattern.
-
-\section*{Acknowledgment}
-\label{sec: ack}
-The work at Illinois Institute of Technology is supported in part by U.S.
-National Science Foundation grants CNS-1320125 and CCF-1422009.
-This work is also supported by the U.S.
-Department of Energy, Office of Science, Advanced Scientific Computing Research, under Contract DE-AC02-06CH11357.
-
-\bibliographystyle{IEEEtran}
-\bibliography{./reference}
-
-\vspace{5\baselineskip}
-
-\begin{framed}
-The submitted manuscript has been created by UChicago Argonne, LLC, Operator of Argonne National Laboratory ("Argonne").
-Argonne, a U.S.
-Department of Energy Office of Science laboratory, is operated under Contract No.
-DE-AC02-06CH11357.
-The U.S.
-Government retains for itself, and others acting on its behalf, a paid-up nonexclusive, irrevocable worldwide license in said article to reproduce, prepare derivative works, distribute copies to the public, and perform publicly and display publicly, by or on behalf of the Government.
-\end{framed}
-
-\end{document}

### 2016-07-07 research_tool.tex
split into separate tex files for each section

--- 2015-10-17 submission.tex
@@
-\documentclass[conference]{IEEEtran}
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
-\else
-
-\fi
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-\usepackage{algorithm}
-%\usepackage{algorithm2e}
-\usepackage{algpseudocode}
-\usepackage{booktabs}
-\usepackage{flushend}
-\usepackage{url}
-\usepackage{xcolor}
-\usepackage{amsmath}
-\usepackage{framed}
-\usepackage{textcomp}
-%\usepackage{subfig}
-\usepackage{graphicx}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{soul}
-\usepackage{verbatim}
-%\usepackage[labelfont=bf]{caption}
-\usepackage[font=bf,labelsep=space]{caption}
-
-\newcommand{\TODO}[1]{\hl{TODO: #1}}
-\newcommand{\NOTE}[1]{\hl{NOTE: #1}}
-\newcommand{\changed}[1]{\textcolor{red}{#1}}
-
-\begin{document}
-%
-% paper title
-% can use linebreaks \\ within to get better formatting as desired
-\title{Study of Intra- and Inter-Job Interference with Different Job Allocations}
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
-\author{
-
-\IEEEauthorblockN{Xu Yang\IEEEauthorrefmark{1}, John Jenkins\IEEEauthorrefmark{2}, Misbah Mubarak\IEEEauthorrefmark{2}, Robert B.
-Ross\IEEEauthorrefmark{2}, Zhou Zhou\IEEEauthorrefmark{1}, Zhiling Lan\IEEEauthorrefmark{1}}
-
-\IEEEauthorblockA{\IEEEauthorrefmark{1}Department of Computer Science,
-Illinois Institute of Technology,
-Chicago, Illinois, USA 60616\\
-\{xyang56, zzhou\}@hawk.iit.edu, lan@iit.edu}
-
-\IEEEauthorblockA{\IEEEauthorrefmark{2}Mathematics and Computer Science Division, Argonne National Laboratory,
-Argonne, IL, USA 60439\\
-\{jenkins,rross\}@mcs.anl.gov, mmubarak@anl.gov}
-}
-
-% make the title area
-\maketitle
-
-\begin{abstract}
-
-Network contention between concurrently running jobs on HPC systems is a primary cause of performance variability.
-Optimizing job allocation and avoiding network sharing are hence crucial to alleviate the potential degradation in performance.
-In order to do these effectively, an understanding of the interaction between applications, their communication patterns, and the system architecture are required.
-In this work, we choose three representative HPC applications from DOE Design Forward Project and conduct detailed simulations of high-performance network topologies to analyze both intra- and inter-job interference.
-By scrutinizing the communication behaviors of these applications, we identify relationships between these behaviors and the possible interference introduced by different job allocation scenarios.
-Our comprehensive analyses in this work illuminate a path towards job-communication-pattern-aware HPC resource allocation.
-
-\end{abstract}
-
-\IEEEpeerreviewmaketitle
-
-\section{Introduction} 
-\label{sec: intro}
-
-The scale of supercomputers continually grows at a breakneck pace to accommodate the computing power requirements for scientific research.
-Supercomputers have tens of thousands of nodes and serve as an irreplaceable research vehicle for scientific problems with increasing size and complexity.
-Supercomputers are usually employed as a shared resource to accommodate many parallel applications (jobs) running concurrently.
-These parallel jobs share the system infrastructure such as network and I/O bandwidth, and inevitably there is contention over these shared resources.
-As supercomputers continue to evolve, these shared resources are increasingly the bottleneck for performance.
-
-\begin{figure}[h!]
-    \centering
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/goodallocation}
-        \caption{Contiguous}
-        \label{fig:overview_sub1}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/badallocation}
-        \caption{Non-contiguous}
-        \label{fig:overview_sub2}
-    \end{subfigure}%
-   \caption{Multiple jobs running concurrently with different allocations.
-Each job is represented by a specific color.
-a) shows the effect of contiguous allocation, which reduce the inter-job interference.
-b) shows non-contiguous allocation, which may introduce both intra and inter-job interferences.
-}
-   \label{fig:overview}
-\end{figure}
-
-When submitting jobs to HPC systems, users request resources in terms of node counts and expected runtime.
-The batch scheduler is then responsible for dispatching the submitted jobs to the system.
-Typically, multiple jobs run concurrently, resulting in the shared use of resources, particularly network and storage.
-A prominent problem with this approach is network contention among concurrently running jobs.
-If the network resources of concurrently running jobs aren't isolated, the resulting sharing can cause communication variability that results in performance degradation for some or all such jobs \cite{abhinav-sc13}.
-This delay can propagate into the queueing time of the following submitted jobs, thus leading to low system throughput and utilization \cite{jose-ipdps15}.
-This adverse effect of network sharing can be mitigated by providing jobs with isolated allocations and exclusive network resources.
-
-On the widely used torus-connected HPC systems \cite{bgq,tofu,titan}, two allocation strategies are commonly used.
-The \emph{contiguous allocation} strategy assigns to each job a compact and contiguous set of computing nodes, as shown in Figure \ref{fig:overview_sub1}.
-The partition-based allocation approach used in Blue Gene series systems is an example of such a strategy \cite{bgloverview}.
-The contiguous strategy favors application performance through isolated networking within a partition and the locality that implies.
-However, this strategy can cause both internal fragmentation (when more nodes are allocated to a job than it requests) and external fragmentation (when sufficient nodes are available for a request, but they can not be allocated contiguously), therefore leading to lower system utilization than is otherwise possible.
-On the other side of the coin, the \emph{non-contiguous allocation} strategy, used by the Cray XT/XE series \cite{carl-cug}, assigns free nodes to jobs regardless of contiguity, though of course efforts are made to maximize locality.
-Figure~\ref{fig:overview_sub2} shows an example non-contiguous allocation.
-While eliminating internal and external fragmentation as seen in contiguous allocation systems, in return non-contiguous allocations introduce contention between jobs due to the interleaving of job nodes.
-The non-contiguous node allocation can significantly reduce job performance, especially for communication-intensive ones \cite{abhinav-sc13}.
-
-We envision that future HPC schedulers will adopt a flexible job allocation mechanism which combines the best of both contiguous and non-contiguous allocation strategies.
-Such a flexible mechanism would take shared resource needs of jobs into account when making allocation decisions (e.g., network).
-With knowledge and analysis of job communication patterns, it can be identified which jobs require network isolation and locality, and to what degree.
-Then, rather than allocating each job in a ``know-nothing'' manner, one may specialize allocation so that, for example, only the jobs with stringent network needs are given compact, isolated allocations, resulting in maximized utilization and minimized perceivable resource contention effects.
-
-In this work, we focus on an in-depth analysis of intra- and inter-job communication interference with different job allocation strategies on torus-connected HPC systems.
-Torus-based networks are used on six of the top 10 supercomputers on the June 2015 Top500 lists~\cite{top500}.
-The current generation of IBM Blue Gene/Q (BG/Q) supercomputers, such as Mira at Argonne Nation Laboratory and Sequoia at Lawrence Livermore National Laboratory, has its nodes connected in a 5D torus network \cite{bgq}.
-The K computer from Japan uses the ``Tofu'' system, which is a 6D mesh/torus topology \cite{tofu}.
-Titan, a Cray XK7 supercomputer located at the Oak Ridge Leadership Computing Facility (OLCF), has nodes connected in a 3D torus within the compute partition \cite{titan}.
-Although our analyses are based on torus networks, the ideas conveyed in this work is applicable to networks with different topologies.
-
-We selected three signature applications from the DOE Design Forward Project~\cite{designforwardwebpage} as examples to conduct detailed study about their communication patterns.
-We use a sophisticated simulation toolkit named CODES (standing for Co-Design of Multi-layer Exascale Storage Architectures)~\cite{Jason-2011} as a research vehicle to evaluate the performance of these applications with various allocations in a controlled environment.
-We then analyze the intra- and inter-job interference by simulating these applications running exclusively and concurrently with different allocations.
-We believe the insights presented in this work can be very useful for the design of future HPC batch job schedulers and resource managers.
-
-%\NOTE{JJ removed the list of contributions, because they are exactly the same as the preceding paragraph.}
-\begin{comment}
-The observations of this work are the following:
-
-\begin{itemize}
-    \item We study the communication behaviors and identify the communication patterns of three representative HPC applications.
-    \item We conduct detailed analysis about the intra- and inter-job interference by running three applications with different allocation strategies.
-And we find the relation between application's communication pattern and the cause of interference.
-    \item We show a novel and flexible way to make allocation with awareness of job's communication pattern, which  is every effective to alleviate the interference between concurrently running jobs.
-\TODO{do we actually do this, or provide the underpinnings for such a method?}
-\end{itemize}
-
-We believe the finds and insights presented in this work would be very useful for the design of future HPC batch job scheduler and resource management module.
-
-\end{comment}
-
-The rest of this paper is organized as follows.
-Section~\ref{sec:application study} describes the three representative applications from DOE Design Forward Project for our study.
-Section~\ref{sec:codes} talks about the use of CODES as research vehicle for our work.
-Section~\ref{sec:config study} shows the performance analysis of three applications on different torus networks.
-Section~\ref{sec:interference} provides detailed analysis about the intra- and inter-job interference between three applications on torus network with different allocation strategies.
-Section~\ref{sec:discussion} introduces a path toward communication-pattern aware allocation strategies, given the results of our analysis.
-Section~\ref{sec:related_work} discusses related work.
-Finally, the conclusion is presented in Section~\ref{sec:conclusion}.
-
-\section{Application Study}
-\label{sec:application study}
-
-%A parallel application usually conforms to a combination of several basic communication patterns\cite{roth}.
-At its different execution phases, the application's communication behavior may follow different basic patterns respectively.
-\textcolor{blue}{When we look into the data flow during application execution, most parallel applications start with broadcast operation to distribute the data from root process to other processes\textcolor{red}{ref needed here}, followed by a series of computation and communication that conforms to certain pattern, which is usually the dominant part of application's execution.
-Before the application come to completion, all the working process will return their results to the root directly or hierarchically.} Parallel application's communication pattern usually involves lots of factors, such as communication intensity, operation-to-operation dependencies and the ``critical path" in its communication topology graph, etc.
-In this work, we focus on the locality of application's rank-to-rank communication topology graph.
-
-A parallel application usually conforms to a combination of several basic communication patterns \cite{roth}.
-At its different execution phases, the application's communication behavior may follow different basic patterns respectively.
-There are many profiling tools available to capture information regarding communication patterns of parallel applications~\cite{tau,mpip,scala,sst,oxbow}.
-They can provide information such as the percentage of different MPI operations of the applications, communication topology, the amount of data transferred between processes.
-
-There are many applications running on HPC systems, such as Mira \cite{bgq} and Titan \cite{titan}, that are communication intensive and conform to various communication patterns.
-In this work, we select three representative applications from the DOE Design Forward Project.
-Each application exhibits a distinctive communication pattern that is commonly seen in HPC applications.
-We believe that the communication patterns of these applications are representative of a wide array of applications running on leadership-class machines.
-Specifically, we study the Algebraic MultiGrid Solver (AMG), Geometric MultiGrid (MultiGrid) and CrystalRouter MiniApps.
-The communication pattern figures we present in Section~\ref{sec:amg}, \ref{sec:crystalrouter}, and \ref{sec:multigrid} are generated with the IPM~\cite{ipm} data gathered from publicly available traces~\cite{designforwardwebpage}.
-
-% MiniApps are reduced proxy applications that encapsulate the salient performance of larger full size applications \cite{miniapp}.
-Since we only focus on the dominant communication pattern of parallel applications, these MiniApps are perfect candidates for analysis the application's performance on different allocations or on new network architectures.
-We choose three MiniApps, namely, Algebraic Multigrid Solver(AMG), Geometric MultiGrid(MultiGrid) and CrystalRouter.
-Their representative communication patterns that are commonly seen in the HPC system workload.
-The communication pattern figures we present here are generated with the IPM \cite{ipm} data from \cite{designforwardwebpage}.
-
-\subsection{AMG}
-\label{sec:amg}
-
-\begin{figure}[t!]
-    \centering
-    \begin{subfigure}[t]{0.25\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/appstudy/amg/amg_pip}
-        \caption{}
-        \label{fig:amg-communication-topology}
-    \end{subfigure}%
-    ~
-    \begin{subfigure}[t]{0.24\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/appstudy/amg/amg_data_transfer}
-        \caption{}
-        \label{fig:amg-data-trans}
-    \end{subfigure}
-    \caption{AMG.
-(a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in AMG.
-(b) The amount of data sent/received by each rank in AMG.
-Y-axis labels the amount of data transferred in MB, x-axis labels 216 ranks of AMG.
-}
-\end{figure}
-
-The Algebraic MultiGrid Solver, or AMG, is a parallel algebraic multi-grid solver for linear systems arising from problems on unstructured mesh physics packages.
-It has been derived directly from the BoomerAMG solver that is being developed in the Center for Applied Scientific Computing (CASC) at LLNL \cite{amg}.
-The dominant communication pattern is regional communication with decreasing message size for different parts of the multi-grid v-cycle.
-
-Figure~\ref{fig:amg-communication-topology} shows the communication topology of a small scale AMG execution with 216 MPI Ranks.
-Note that the dominant communication pattern of the application doesn't change with scale.
-We can make the observation that AMG's dominant communication pattern is 3D Nearest Neighbor, each rank has intensive communication with up to six neighbors, depending on the boundaries of the ranks.
-Applications with similar pattern include PARTISN \cite{partisn} and SNAP \cite{snap}.
-
-Figure~\ref{fig:amg-data-trans} shows the amount of data each rank transferred in the example execution.
-The data transferring of each rank is between 7MB to 16MB, corresponding to whether the rank is on a boundary or not.
-
-\subsection{Crystal Router}
-\label{sec:crystalrouter}
-
-\begin{figure}[t!]
-    \centering
-    \begin{subfigure}[t]{0.25\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/appstudy/cr/cr_pip}
-        \caption{Communication Topology}
-        \label{fig:cr-communication-topology}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[t]{0.22\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/appstudy/cr/cr_data_transfer}
-        \caption{Data Volume}
-        \label{fig:cr-data-trans}
-    \end{subfigure}
-    \caption{CrystalRouter.
-(a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in CrystalRouter.
-(b) The amount of data sent/received by each rank in CrystalRouter.
-Y-axis labels the amount of data transferred in MB, x-axis labels 100 ranks of CrystalRouter.
-}
-\end{figure}
-
-The second MiniApp covered is CrystalRouter, the extracted communication kernel of the full application Nek5000~\cite{crystalrouter}, which is a spectral element CFD application developed at Argonne National Laboratory \cite{nek5000}.
-It features spectral element multi-grid solvers coupled with a highly scalable, parallel coarse-grid solver that is widely used for projects including ocean current modeling, thermal hydraulics of reactor cores, and spatiotemporal chaos.
-CrystalRouter demonstrates the ``many-to-many'' communication pattern through a scalable multi-stage communication process.
-The way CrystalRouter works is nodes compute for a while, then synchronize and communicate, continually alternating between these two types of activities.
-
-The collective communication in CrystalRouter utilizes a recursive doubling approach.
-Ranks in CrystalRouter conform to a $n$-dimensional hypercube and recursively split into ($n$-1)-dimensional hypercubes, with communication occurring along the splitting plane.
-The pattern of this communication can be found in Figure~\ref{fig:cr-communication-topology}.
-By nature of the logarithmic splitting process, a substantial portion of the communication occurs in small neighborhoods of ranks.
-CrystalRouter represents a group of applications whose dominant communication is a hybrid of multi-stage local and hierarchical global communication, and shares similarities with most MPI collective communication implementations.
-
-Figure~\ref{fig:cr-data-trans} shows the amount of data each rank transferred in CrystalRouter, which ranges between 20MB to 36MB.
-This is due to the rank counts not matching one-for-one with points in the hypercube.
-
-\subsection{MultiGrid}
-\label{sec:multigrid}
-
-\begin{figure}[t!]
-    \centering
-    \begin{subfigure}[t]{0.25\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/appstudy/mg/mg_pip}
-        \caption{Communication Topology}
-        \label{fig:mg-communication-topology}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[t]{0.22\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/appstudy/mg/mg_data_transfer}
-        \caption{Data Volume}
-        \label{fig:mg-data-trans}
-    \end{subfigure}
-    \caption{MultiGrid.
-(a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in MultiGrid.
-(b) The amount of data sent/received by each rank in MultiGrid.
-Y-axis labels the amount of data transferred in MB, x-axis label 125 ranks of MultiGrid.
-}
-\end{figure}
-
-MultiGrid is geometric multi-grid v-cycle from the production elliptic solver BoxLib, a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes \cite{boxlib}.
-MultiGrid conforms to many-to-many communication pattern with decreasing message size and collectives for different parts of the multi-grid v-cycle.
-It is widely used for structured grid physics packages.
-
-Figure~\ref{fig:mg-communication-topology} shows the communication topology of MultiGrid with 125 ranks.
-We can see intensive communication along the diagonal that resembles nearest neighbor communication, similar to AMG.
-However, the communication topology leads to a greater ``spread'' of communication across the set of ranks, challenging the maximization of communication locality with respect to rank.
-In this sense, it can be considered a ``many-to-many'' pattern.
-Applications with such similar dominant communication pattern are like FillBoundary, another PDE solver code in \cite{boxlib}.
-
-Figure~\ref{fig:mg-data-trans} shows the amount of data transferred between ranks in MultiGrid. For this miniapp, The data transferred per rank is relatively small and stable, between 2.8MB to 2.9MB.
-
 \section{Research Vehicle}
@@
 Torus networks have been extensively used in the current generation of supercomputers because of their linear scaling on per-node cost and competitive communication performance \cite{mubarak-sc2012}\cite{zhou-ipdps}.
-
-\section{Configuration Study}
-\label{sec:config study}
-
-In this section, we study the communication behavior of three applications on torus networks with different dimensionality and bandwidth configurations.
-Torus networks are $k$-ary $n$-cubes, with $k^{n}$ nodes in total arranged in an $n$-dimensional grid having $k$ nodes in each dimension.
-Each node has $2\times n$ direct linked neighbor nodes with wraparound.
-In these and all following experiments, non-adaptive routing is used.
-While making specific recommendations as to an ``optimal'' network configuration for a given application is out of the scope of this study, the experiments do highlight the nuances inherent in design choices such as network diameter and bandwidth planning.
-
-We perform three sets of experiments.
-Each looks at per-rank distribution of performance of AMG, CrystalRouter and MultiGrid on a set of 2048-node torus models, one with $n=3$ ($16^{2} \times 8$), one with $n=5$ ($8\times 4^{4}$), and one with $n=7$ ($4^{4} \times 2^{3}$).
-Performance in this context is defined by the total message transfer time for each sending rank, which we are able to precisely measure due to the simulation environment providing us with an absolute notion of time.
-Note that we simply perform linear rank assignment, one per node -- later sections investigate rank and node mapping effects.
-In the first, shown in Figure~\ref{fig:dimensionality-study}, we keep the link bandwidth constant at 2GiB/s in each direction, corresponding to Blue Gene/Q systems~\cite{bgq}, leading to aggregate node bandwidths of 12GiB/s, 20GiB/s, and 28GiB/s for the 3D, 5D, and 7D tori, respectively.
-In the second, shown in Figure~\ref{fig:bandwidth-study}, we fix the torus dimensionality to 5D.
-In the third, shown in Figure~\ref{fig: bandwidth-time-box}, we instead fix the per-node aggregate bandwidth (and thus the system aggregate bandwidth) and modify the torus dimensionality, resulting in per-link bandwidths shown in Table~\ref{tab: fix-bandwidth}.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/dimenstudy/amg_box}
-        \caption{AMG}
-        \label{fig:dimen-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/dimenstudy/cr_box}
-        \caption{CrystalRouter}
-        \label{fig:dimen-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/dimenstudy/mg_box}
-        \caption{MultiGrid}
-        \label{fig:dimen-mg}
-    \end{subfigure}%
-   \caption{Data Transfer Time of AMG, CrystalRouter and MultiGrid on 3D, 5D and 7D torus network, with constant per-link bandwidth.}
-   \label{fig:dimensionality-study}
-\end{figure*}
-
-In the fixed link bandwidth experiment, shown in Figure~\ref{fig:dimensionality-study}, the data transfer time of the three applications are significantly reduced as the dimensionality increases, so the applications can easily make use of the increased dimensionality (lower hops on average) and/or aggregate bandwidth (more links being concurrently used).
-Interestingly, AMG has the greatest degree of outliers compared to the others.
-We expect this to be a combination of non-optimal rank mappings causing a greater degree of variance due to the highly regular communication pattern and the distribution of message sizes.
-MultiGrid seems to benefit the most in terms of reducing the spread of the time distribution, though the median message time remains roughly within a factor of two.
-The time distribution in CrystalRouter seems to accrue the most benefits, likely due to the combination of cross-cutting and localized per-rank traffic.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/bandwidthstudy/amg_bw_box}
-        \caption{AMG}
-        \label{fig:bdwstudy-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/bandwidthstudy/cr_bw_box}
-        \caption{CrystalRouter}
-        \label{fig:bdwstudy-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/bandwidthstudy/mg_bw_box}
-        \caption{MultiGrid}
-        \label{fig:bdwstudy-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 5D torus network with 2GiB/s, 4GiB/s and GiB/s direct link bandwidth. }
-   \label{fig:bandwidth-study}
-\end{figure*}
-
-In the fixed dimensionality experiment, shown in Figure~\ref{fig:bandwidth-study}, the increased per-link bandwidth, expectedly, benefits each application roughly equally, reducing the time distributions linearly.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/samebdw/amg}
-        \caption{AMG}
-        \label{fig:samebd-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/samebdw/cr}
-        \caption{CrystalRouter}
-        \label{fig:samebd-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/samebdw/mg}
-        \caption{MultiGrid}
-        \label{fig:samebd-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 3D, 5D and 7D torus network with same aggregated bandwidth. }
-   \label{fig: bandwidth-time-box}
-\end{figure*}
-
-\begin{table}[ht]
-\begin{center}
-\caption{Torus networks with same aggregated bandwidth} 
-\label{tab: fix-bandwidth}
-\begin{tabular}{l c c c} 
-\toprule % Top horizontal line
-\toprule
-&\multicolumn{3}{c}{Bandwidth } \\
-\cmidrule(l){2-4}
-Dimension  & per-link & aggregate &\\ % Column names row
-\midrule % In-table horizontal line
-3D      & 4.67GiB/s  & 28GiB/s  & \\  % Content row 1
-\midrule % In-table horizontal line
-5D    & 2.8GiB/s  & 28GiB/s  &\\ % Content row 1
-\midrule % In-table horizontal line
-7D    & 2GiB/s & 28GiB/s & \\
-\midrule
-\bottomrule % Bottom horizontal line
-\end{tabular}
-\end{center}
-\end{table}
-
-In the fixed aggregate bandwidth experiment, with configuration in Table~\ref{tab: fix-bandwidth} and results shown in Figure~\ref{fig: bandwidth-time-box}, the results are more nuanced.
-At least for these applications, the benefits from increased dimensionality do not outweigh the reduction in per-link bandwidth.
-MultiGrid does see benefits in the increased dimensionality, particularly with respect to outliers, but has a similar median message time.
-CrystalRouter reduces performance as the dimensionality increases, but only by a small degree.
-AMG remains roughly the same, minus a dip in the maximum message time that could be due to rank-to-node mapping changes.
-
-\section{Interference Analysis}
-\label{sec:interference}
-
-\textbf{\emph{Intra-job interference}} refers to the network contention between ranks within each application.
-\textbf{\emph{Inter-job interference}} is introduced by concurrently running jobs that adjacent to each other sharing network resources.
-Communication variability due to such interference can cause application performance degradation.
-
-In this section we will study both kinds of interference in the context of torus network. 
-Since application's communication pattern won't change with its scale, we use AMG with 216 ranks, CrystalRouter with 100 ranks, and MultiGrid 125 ranks.
-And in order to accommodate three applications running concurrently with different allocations, we simulate a 2K-node 3D torus ($16^{2} \times 8$) network for the experiments.
-
-\subsection{Intra-Job Interference Analysis}
-\label{sec: introjob}
-
-We design two sets of experiments to study the intra-job interference of each application.
-In the first, shown in Figure \ref{fig:shapestudy}, we assign each application with allocations in \emph{three different shapes}.
-In the second, shown in Figure \ref{fig:diffmap}, we study the intra-job interference by using different mapping strategies for each application with given allocation shape.
-
-In the allocation shapes experiment, we select three possible shapes can can commonly seen on the 3D torus network.
-They are 3D balanced-cube, 3D unbalanced-cube, 2D mesh, as shown in Figure \ref{fig:alloc-shapes}.
-
-3D balanced-cube, shown as red in Figure \ref{fig:cont_sub1}, can guarantee the minimum \emph{average pair-wise distance} within the allocation.
-Some research studies like \cite{leung,abhinav-sc13} indicate that compact allocation can guarantee jobs with better performance.
-They use a variety of metrics to evaluate the compactness of the allocation, such as \emph{average pair-wise distance}, \emph{diameter} and \emph{contiguity}.
-In this work, we select 3D balanced-cube as the most compact allocation on a 3D torus network.
-
-3D unbalanced-cube, shown as green in Figure \ref{fig:cont_sub1}, is a rectangular prism, which is the possible allocation shape on the systems with asymmetric networks.
-For example, Cray XE6/XK7 systems are deployed 3D tori with Gemini routers.
-The network connections in the \emph{y}-direction have only half the bandwidth of the cables used in the \emph{x} and \emph{z} directions.
-In order to take advantage of the faster links in the \emph{x} and \emph{z} directions, job allocation starts from X-Z plane, which leads to a rectangular prism shaped allocation \cite{RF}.
-
-2D mesh, shown as blue in Figure \ref{fig:cont_sub1}, can be cut out from a single layer of the 3D torus.
-2D mesh is a very common allocation shape in torus network for both contiguous and non-contiguous allocation strategies.
-For example, Cray Application Level Placement Scheduler (ALPS) indexes the nodes in torus network into a list and makes job allocation by simply going through the list \cite{carl-cug}.
-When the list is obtained by sorting the nodes based on their spacial coordinates in the torus, allocation made off from this list will be 2D mesh.
-The IBM Blue/Gene Q supercomputer Mira at Argonne Leadership Computing Facility also allow its allocation partition configured into mesh \cite{zhou-ipdps}.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/allocation}
-        \caption{ }
-        \label{fig:cont_sub1}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit16}
-        \caption{ }
-        \label{fig:noncont_sub1}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit8}
-        \caption{ }
-        \label{fig:noncont_sub2}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit2}
-        \caption{ }
-        \label{fig:noncont_sub3}
-    \end{subfigure}%
-    \caption{(a) Contiguous allocation in three different shapes.
-``Blue" is 2D mesh, ``Green" is 3D-unbalanced cube and ``Red" is 3D-balanced cube.
-(b)-(d) Non-contiguous allocation.
-Each job is represented by a specific color.
-The nodes assigned to different jobs are interleaved, the size of allocation unit are 16, 8 and 2.}
-    \label{fig:alloc-shapes}
-\end{figure*}
-
-AMG with 3D ``Nearest Neighbor" communication pattern takes less time when being assigned with 3D-unbalanced allocation than 3D-balanced, as shown in Figure \ref{fig:shapstudy-amg}.
-And MultiGrid gets the best performance (shortest data transfer time) when running on 3D-balanced allocation, as shown in Figure \ref{fig:shapstudy-mg}.
-Since MultiGrid's communication pattern is ``Many-to-Many" dominant, 3D-balanced is the most compact allocation with shortest pair-wise distance between nodes, which can reduce the aggregated hops for transferring message among ranks in MultiGrid.
-As CrystalRouter has both dominant local communication and multi-stage global data transfer, the 3D-balance is also the best allocation, but its advantage over 3D-unbalanced is not that obvious as to MultiGrid, as shown in Figure \ref{fig:shapstudy-cr}.
-
-Lots of research studies have designed fancy allocation algorithms to provide application with the most compact allocation\cite{leung,LO}.
-However, as shown in our experiments, providing cubical allocation without considering application's communication pattern won't guarantee the best performance for every application.
-Compact allocation should be provided to application with global data transfer, such as those conforms to ``Many-to-Many" communication pattern.
-Application with dominant communication pattern like ``Nearest Neighbor" won't fully utilize the compact allocation and could get even better performance on a relaxed allocation shape.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/amg_box}
-        \caption{AMG}
-        \label{fig:shapstudy-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/cr_box}
-        \caption{CrystalRouter}
-        \label{fig:shapstudy-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/mg_box}
-        \caption{MultiGrid}
-        \label{fig:shapstudy-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 3D-balanced, 3D-unbalanced and 2D allocation.}
-   \label{fig:shapestudy}
-\end{figure*}
-
-The rank-to-node mapping of parallel applications on HPC system could greatly impact applications' performance.
-However, finding the optimal mapping solution for a given application is out the scope of this work.
-Our experiments aims to show how mapping strategies impact the intra-job interference of application with specific communication pattern.
-
-We provide AMG, CrystalRouter and MultiGrid with 3D-balance allocation and use three mapping strategies to do the rank-to-node mapping.
-The ``Linear" mapping, as we used in the previous experiments, is to map each rank according to the dimensional ordering of compute nodes.
-The ``Cube" mapping assigns ranks into consecutive $2^{3}$ cubes.
-The ``Random" mapping assigns ranks randomly within the allocation.
-
-AMG remains roughly the same when it been mapped by ``Linear" and ``Cube", shown in Figure \ref{fig:diffmap-amg}.
-Although ``Linear" and ``Cube" mapping strategies cause some routing overlap for AMG's communication, both mapping strategies still preserve the locality of AMG's 3D ``Nearest Neighbor" communication pattern.
-The ``Random" mapping totally destroys AMG's communication pattern, and results in intra-job interference among all the ranks in AMG.
-The performance degradation caused by using ``Random" mapping is up to 90\%.
-
-The ``Cube" mapping improves CrystalRouter's performance over the ``Linear" by reducing the pair-wise distance between ranks, shown in Figure \ref{fig:diffmap-cr}.
-The global data transfer in CrystalRouter will takes fewer hops with ``Cube" mapping.
-The ``Random" mapping scatters the ranks within the local neighborhood of CrystalRouter and make their communication less efficient.
-
-The ``Cube" mapping benefits MultiGrid's ``Many-to-Many" communication.
-However, due to the small amount of data transfer among ranks in MultiGrid, ``Cube" mapping doesn't exhibit a drastic advantage.
-``Random" mapping neither cause too much degradation for the same reason, shown in Figure \ref{fig:diffmap-mg}.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/diffmapping/amg}
-        \caption{AMG}
-        \label{fig:diffmap-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/diffmapping/cr}
-        \caption{CrystalRouter}
-        \label{fig:diffmap-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/diffmapping/mg}
-        \caption{MultiGrid}
-        \label{fig:diffmap-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 3D-balanced allocation using  different mapping strategies.}
-   \label{fig:diffmap}
-\end{figure*}
-
-\subsection{Inter-Job Interference Analysis}
-\label{sec: interjob interference study}
-
-Inter-job interference has been identified as one of the major culprits for application's performance variability \cite{abhinav-sc13,skinner,rosenthal}.
-Inter-job interference is a more prominent issue for system adopted non-contiguous allocation policy than system with contiguous allocation.
-Application communication times have been demonstrated to vary from 36\% faster to 69\% slower due to job interference when running on non-contiguous allocation \cite{abhinav-sc13}.
-
-We assign each application with non-contiguous allocation and run them concurrently on the same network.
-The allocation unit belongs to different jobs are interleaved.
-In order to study the impact of different allocation unit size on applications' inter-job interference, we conduct experiments with unit size of 16, 8 and 2, shown respectively in Figure \ref{fig:noncont_sub1}, \ref{fig:noncont_sub1} and \ref{fig:noncont_sub3}.
-Figure \ref{fig:interjobstudy} shows the results of each application data transfer time with different allocation unit size.
-
-The data transfer time of AMG in Figure \ref{fig:interjob-amg} keeps stable between allocation unit size of 16 and 8, since the locality of AMG is 6-rank based.
-When the unit size reduce to 2, AMG suffers prolong data transfer time by about 10\%.
-CrystalRouter is more sensitive to allocation unit size.
-The best unit size would be the one big enough to accommodate its local neighborhood.
-Figure \ref{fig:interjob-cr} shows unit size of 16 and 8 can guarantee the same average data transfer time, while some rank spend more time with allocation unit size 8 than 16.
-When the unit size reduce to 2, the communication become less efficiency and takes about 15\% more times for transferring data.
-
-The data transfer time of MultiGrid with different allocation unit sizes doesn't show obvious variability in Figure \ref{fig:interjob-mg}.
-This is because even big allocation unit size like 16 will still fail to preserve MultiGrid's ``Many-to-Many" pattern.
-The data transfer time is almost doubled when MultiGrid running concurrently with allocation unit size of 16, as shown in Figure \ref{fig:interjob-mg} .
-As the unit size decreases, the data transfer time keep growing, but in a slow way, which means in terms of preserving the locality of MultiGrid, allocation unit of size 16, 8 and 2 serve equally bad.
-
-The unit size in non-contiguous allocation policy should be chosen according to application's communication pattern.
-Inter-job interference is inevitable in non-contiguous based systems, but unit size that big enough to preserve the neighborhood communication of the application will alleviate such interference and improve job performance.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/inter-job/amg}
-        \caption{AMG}
-        \label{fig:interjob-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/inter-job/cr}
-        \caption{CrystalRouter}
-        \label{fig:interjob-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/inter-job/mg}
-        \caption{MultiGrid}
-        \label{fig:interjob-mg}
-    \end{subfigure}%
-   \caption{Inter-job interference study.
-``cont" indicates three applications running side by side concurrently on the same network with contiguous allocation.
-To study the impact of non-contiguous allocation on inter-job interference, applications running concurrently with interleaved allocation of different unit sizes, which are 16-node, 8-node, 2-node.
-}
-   \label{fig:interjobstudy}
-\end{figure*}
-
-\subsection{Result Summary}
-\label{sec:summary}
-
-Based on the above comprehensive study, we can make following observations:
-\begin{itemize}
-    \item The compact allocation may not guarantee the best performance for every application. 
-    \item The applications dominated with Nearest Neighbor communications exhibit relatively stable performance under different allocation shapes, while the applications dominated with Many-to-Many communication exhibit better performance with compact allocation (e.g.
-3D balanced).
-    \item A good rank-to-node mapping strategy can greatly improve application's performance when specific allocation is given.
-    \item An optimal size for allocation units should be determined according to an application's dominant communication pattern.
-In general, a unit size should be large enough to preserve neighboring communication of the application.
-    \item Inter-job interference is inevitable in non-contiguous allocation.
-However, choosing the proper allocation unit size with job's communication pattern awareness can alleviate such negative effect.
-\end{itemize}
-
-\section{Discussion}
-\label{sec:discussion}
-
-The results shown in the previous section provide insights for the design of smart and flexible job allocation.
-By scrutinizing job's communication behavior, we can identify job's dominant communication pattern and pinpoint the ``neighborhood communication".
-With such knowledge about jobs' communication patterns, we can analyze the possible interference between jobs and take precautions to alleviate the negative effect when making allocation decisions.
-
-When an application with dominant communication that is intensive ``Many-to-Many", the scheduler should grant it with compact node allocation and exclusive network provision.
-The compact allocation can guarantee the shortest pair-wise distance between all the ranks, and make the data transfer between ranks take less hops.
-On the other hand, the exclusive network provision will prevent other adjacent applications from sharing network resources, thus eliminate the performance degradation due to interference.
-
-Not every application's preferable resource are compact node allocation and exclusive network provision.
-As we found in our analysis, applications whose dominant communication patterns contain intensive ``neighborhood communication" like ``Nearest Neighbor" don't benefit from compact node allocation and exclusive network provision.
-Applications, such as AMG, can run with non-contiguous node allocation without performance degradation, as long as the allocation unit can accommodate their ``neighborhood communication".
-
-The allocation unit size shouldn't be fixed.
-Instead, the scheduler should choose a proper unit size based on the scale of ``neighborhood communication" of each job.
-The best unit size should be neither too big nor too small, jut perfectly fit for the size of that ``neighborhood".
-Big unit size won't be fully utilized and also cause fragmentation.
-Small unit size won't be able to accommodate the ``neighborhood communication" and makes the intra-job communication less efficient.
-
-The advantage of making allocation with consideration about job's communication pattern is obvious.
-First, compared with contiguous allocation policy, the new scheduler outlined above is more flexible, it doesn't require the system to provide a big contiguous partition to accommodate the whole application, instead it only needs to provide a small set of compact nodes that is sufficient for all the "local community" in the application.
-On the other hand, the new scheduler outlined above won't destroy the locality of application's communication pattern.
-Indeed, the small compact node set (allocation unit) provided for the application's  ``local communities" can preserve the communication locality in the maximum extent.
-The design of such a smart scheduler is part of our future work.
-
-\section{Related Work}
-\label{sec:related_work}
-
-There are many tools available for system monitoring and application profiling.
-Tools like TAU (Tuning and Analysis) \cite{tau}, SCALASCA \cite{scala} and mpiP \cite{mpip} can capture application's runtime information about application's communication and computation in event traces.
-However, recognizing the patterns about application's communication from those traces requires substantial effort.
-There are a number of studies on the recognition and characterization of parallel application communication patterns.
-Oak Ridge National Laboratory has its ongoing project about developing a tool set named Oxbow, which can characterize the computation and communication behavior of scientific applications and benchmarks \cite{oxbow}.
-In a recent work \cite{roth}, the authors demonstrate a new approach to automatically characterizing parallel applications' communication behaviors.
-
-%\textbf{P2: communication characteristics}\\
-
-Many research efforts have been conducted to characterize scientific applications.
-For instance, a DOE project aims at the identification of the computational characteristics of the DOE MiniApps developed at various exascale co-design centers \cite{designforwardwebpage}.
-In this project, the communication patterns of several DOE full applications and associated mini-applications are studied to provide a more complete snapshot of the DOE workload.
-A joint project named CORAL from Oak Ridge, Argonne and Livermore provide a series of benchmarks to represent DOE workloads and technical requirements \cite{coral}.
-The CORAL project includes scalable science benchmarks, throughput benchmarks, data centric benchmarks, skeleton benchmarks and Micro benchmarks.
-
-%\textbf{P3: Interference between concurrently running jobs}\\
-The interference between concurrently running jobs on HPC system have been identified as major culprit for job's performance variability.
-Abhinav et al.
-found that concurrently running applications on HPC can cause interference to each other, and cause communication time varied from 36\% faster to 69\% slower.
-Such interference could come from Operating System noise, shape of the allocated partition , or other running jobs that sharing network resources \cite{abhinav-sc13}.
-Skinner et al.
-found that  there is a 2-3$\times$ slowdown in MPI$\textunderscore$Allreduce due to network contention from other jobs \cite{skinner}.
-Rosenthal et al.
-found increasing network bandwidth provide limited benefits to a number of applications.
-The applications that send mostly small messages or large messages asynchronously are not bandwidth bounded, hence, benefit only slightly or not at all from increased bandwidth \cite{rosenthal}.
-
-%\textbf{P4: Job allocation(little bit)}\\
-Several research studies focus on optimizing job allocation on HPC system to alleviate the interference between concurrently running jobs.
-Hoefler et al.
-proposed to use performance modeling techniques to analyze factors that impact the performance of parallel scientific applications \cite{hoefler-modeling}.
-However, as the scale of HPC continue to grow, the  interference of concurrently running jobs is getting worse, which is hard to be quantified by performance profiling tools alone.
-Bogdan et al.
-provide a set of guidelines of how to configure a network with Dragonfly topology for workload with Nearest Neighbor communication pattern \cite{Bogdan-hpdc14}.
-Dong et al.
-have developed simple benchmarks that conforms to four different communication patterns, namely ping-pong, nearest neighbor, broadcast and all$\textunderscore$reduce, to demonstrate the effectiveness of this highly parallel 5D torus network \cite{Dong-SC11}.
-
-Our work is different from all these works in the following ways.
-First, we focus on the dominant communication patterns rather than any specific application.
-We believe this can provide a guideline for other research work.
-Secondly, we explore the intra- and inter-job interference between concurrently running jobs, while similar work such as \cite{abhinav-sc13} only focuses on the single application's performance degradation due to network contention.
-Finally, we analyze the impact of different allocation strategies to job's communication behaviors.
-We further identify the optimal allocation strategy for each application with a specific dominant communication pattern.
-Based on our study, we claim that further scheduling should take job's communication pattern into consideration for its allocation decision making.
-
-\section{Conclusions}
-\label{sec:conclusion}
-In this work, we have studied the communication behavior of three parallel applications, namely AMG, CrystalRouter and MultiGrid.
-Each application has distinctive communication pattern, that can be representative for a group of jobs in HPC workload.
-We have used a sophisticate simulation tool named CODES from Argonne National Laboratory to simulate the running of these three parallel applications on torus network.
-The torus network provided by CODES has good fidelity and scalability.
-We have analyzed the performance of each application's communication in terms of data transfer time by simulating them running on torus networks with different bandwidth and dimensionality configurations.
-We have found that higher dimensionality of torus network would improve the performance of application with ``Many-to-Many" communication patterns, while application with intensive local communication like ``Nearest Neighbor" won't benefit much from higher dimensionality.
-
-We have analyzed the intra- and inter-job interference by simulating three applications running on 3D torus network.
-Based on our comprehensive experiments, we got five observations.
-1) The compact allocation may not guarantee the best performance for every application.
-2)The applications dominated with Nearest Neighbor communications exhibit relatively stable performance under different allocation shapes, while the applications dominated with Many-to-Many communication exhibit better performance with compact allocation (e.g.
-3D balanced).
-3) A good rank-to-node mapping strategy can greatly improve application's performance when specific allocation is given.
-4) An optimal size for allocation units should be determined according to an application's dominant communication pattern.
-In general, a unit size should be large enough to preserve neighboring communication of the application.
-5) Inter-job interference is inevitable in non-contiguous allocation.
-However, choosing the proper allocation unit size with job's communication pattern awareness can alleviate such negative effect.
-
-We believe that our finding in this work can provide valuable guidance for HPC resource management to make flexible job allocations.
-Rather than using pre-defined partitions and non-contiguous allocation, future HPC systems should assign each job with preferable resources based on job's communication pattern.
-
-\section*{Acknowledgment}
-\label{sec: ack}
-The work at Illinois Institute of Technology is supported in part by U.S.
-National Science Foundation grants CNS-1320125 and CCF-1422009.
-This work is also supported by the U.S.
-Department of Energy, Office of Science, Advanced Scientific Computing Research, under Contract DE-AC02-06CH11357.
-
-\bibliographystyle{IEEEtran}
-\bibliography{./reference}
-
-\vspace{5\baselineskip}
-
-\begin{framed}
-The submitted manuscript has been created by UChicago Argonne, LLC, Operator of Argonne National Laboratory ("Argonne").
-Argonne, a U.S.
-Department of Energy Office of Science laboratory, is operated under Contract No.
-DE-AC02-06CH11357.
-The U.S.
-Government retains for itself, and others acting on its behalf, a paid-up nonexclusive, irrevocable worldwide license in said article to reproduce, prepare derivative works, distribute copies to the public, and perform publicly and display publicly, by or on behalf of the Government.
-\end{framed}
-
-\end{document}

### 2016-07-07 submission.tex
split into separate tex files for each section

@@
 
-\section{Introduction} 
-\label{sec: intro}
-
-The scale of supercomputers continually grows at a breakneck pace to accommodate the computing power requirements for scientific research.
-Supercomputers have tens of thousands of nodes and serve as an irreplaceable research vehicle for scientific problems with increasing size and complexity.
-Supercomputers are usually employed as a shared resource to accommodate many parallel applications (jobs) running concurrently.
-These parallel jobs share the system infrastructure such as network and I/O bandwidth, and inevitably there is contention over these shared resources.
-As supercomputers continue to evolve, these shared resources are increasingly the bottleneck for performance.
-
-\begin{figure}[h!]
-    \centering
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/goodallocation}
-        \caption{Contiguous}
-        \label{fig:overview_sub1}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/badallocation}
-        \caption{Non-contiguous}
-        \label{fig:overview_sub2}
-    \end{subfigure}%
-   \caption{Multiple jobs running concurrently with different allocations.
-Each job is represented by a specific color.
-a) shows the effect of contiguous allocation, which reduce the inter-job interference.
-b) shows non-contiguous allocation, which may introduce both intra and inter-job interferences.
-}
-   \label{fig:overview}
-\end{figure}
-
-When submitting jobs to HPC systems, users request resources in terms of node counts and expected runtime.
-The batch scheduler is then responsible for dispatching the submitted jobs to the system.
-Typically, multiple jobs run concurrently, resulting in the shared use of resources, particularly network and storage.
-A prominent problem with this approach is network contention among concurrently running jobs.
-If the network resources of concurrently running jobs aren't isolated, the resulting sharing can cause communication variability that results in performance degradation for some or all such jobs \cite{abhinav-sc13}.
-This delay can propagate into the queueing time of the following submitted jobs, thus leading to low system throughput and utilization \cite{jose-ipdps15}.
-This adverse effect of network sharing can be mitigated by providing jobs with isolated allocations and exclusive network resources.
-
-On the widely used torus-connected HPC systems \cite{bgq,tofu,titan}, two allocation strategies are commonly used.
-The \emph{contiguous allocation} strategy assigns to each job a compact and contiguous set of computing nodes, as shown in Figure \ref{fig:overview_sub1}.
-The partition-based allocation approach used in Blue Gene series systems is an example of such a strategy \cite{bgloverview}.
-The contiguous strategy favors application performance through isolated networking within a partition and the locality that implies.
-However, this strategy can cause both internal fragmentation (when more nodes are allocated to a job than it requests) and external fragmentation (when sufficient nodes are available for a request, but they can not be allocated contiguously), therefore leading to lower system utilization than is otherwise possible.
-On the other side of the coin, the \emph{non-contiguous allocation} strategy, used by the Cray XT/XE series \cite{carl-cug}, assigns free nodes to jobs regardless of contiguity, though of course efforts are made to maximize locality.
-Figure~\ref{fig:overview_sub2} shows an example non-contiguous allocation.
-While eliminating internal and external fragmentation as seen in contiguous allocation systems, in return non-contiguous allocations introduce contention between jobs due to the interleaving of job nodes.
-The non-contiguous node allocation can significantly reduce job performance, especially for communication-intensive ones \cite{abhinav-sc13}.
-
-We envision that future HPC schedulers will adopt a flexible job allocation mechanism which combines the best of both contiguous and non-contiguous allocation strategies.
-Such a flexible mechanism would take shared resource needs of jobs into account when making allocation decisions (e.g., network).
-With knowledge and analysis of job communication patterns, it can be identified which jobs require network isolation and locality, and to what degree.
-Then, rather than allocating each job in a ``know-nothing'' manner, one may specialize allocation so that, for example, only the jobs with stringent network needs are given compact, isolated allocations, resulting in maximized utilization and minimized perceivable resource contention effects.
-
-In this work, we focus on an in-depth analysis of intra- and inter-job communication interference with different job allocation strategies on torus-connected HPC systems.
-Torus-based networks are used on six of the top 10 supercomputers on the June 2015 Top500 lists~\cite{top500}.
-The current generation of IBM Blue Gene/Q (BG/Q) supercomputers, such as Mira at Argonne Nation Laboratory and Sequoia at Lawrence Livermore National Laboratory, has its nodes connected in a 5D torus network \cite{bgq}.
-The K computer from Japan uses the ``Tofu'' system, which is a 6D mesh/torus topology \cite{tofu}.
-Titan, a Cray XK7 supercomputer located at the Oak Ridge Leadership Computing Facility (OLCF), has nodes connected in a 3D torus within the compute partition \cite{titan}.
-Although our analyses are based on torus networks, the ideas conveyed in this work is applicable to networks with different topologies.
-
-We selected three signature applications from the DOE Design Forward Project~\cite{designforwardwebpage} as examples to conduct detailed study about their communication patterns.
-We use a sophisticated simulation toolkit named CODES (standing for Co-Design of Multi-layer Exascale Storage Architectures)~\cite{Jason-2011} as a research vehicle to evaluate the performance of these applications with various allocations in a controlled environment.
-We then analyze the intra- and inter-job interference by simulating these applications running exclusively and concurrently with different allocations.
-We believe the insights presented in this work can be very useful for the design of future HPC batch job schedulers and resource managers.
-
-%\NOTE{JJ removed the list of contributions, because they are exactly the same as the preceding paragraph.}
-\begin{comment}
-The observations of this work are the following:
-
-\begin{itemize}
-    \item We study the communication behaviors and identify the communication patterns of three representative HPC applications.
-    \item We conduct detailed analysis about the intra- and inter-job interference by running three applications with different allocation strategies.
-And we find the relation between application's communication pattern and the cause of interference.
-    \item We show a novel and flexible way to make allocation with awareness of job's communication pattern, which  is every effective to alleviate the interference between concurrently running jobs.
-\TODO{do we actually do this, or provide the underpinnings for such a method?}
-\end{itemize}
-
-We believe the finds and insights presented in this work would be very useful for the design of future HPC batch job scheduler and resource management module.
-
-\end{comment}
-
-The rest of this paper is organized as follows.
-Section~\ref{sec:application study} describes the three representative applications from DOE Design Forward Project for our study.
-Section~\ref{sec:codes} talks about the use of CODES as research vehicle for our work.
-Section~\ref{sec:config study} shows the performance analysis of three applications on different torus networks.
-Section~\ref{sec:interference} provides detailed analysis about the intra- and inter-job interference between three applications on torus network with different allocation strategies.
-Section~\ref{sec:discussion} introduces a path toward communication-pattern aware allocation strategies, given the results of our analysis.
-Section~\ref{sec:related_work} discusses related work.
-Finally, the conclusion is presented in Section~\ref{sec:conclusion}.
-
-\section{Application Study}
-\label{sec:application study}
-
-%A parallel application usually conforms to a combination of several basic communication patterns\cite{roth}.
-At its different execution phases, the application's communication behavior may follow different basic patterns respectively.
-\textcolor{blue}{When we look into the data flow during application execution, most parallel applications start with broadcast operation to distribute the data from root process to other processes\textcolor{red}{ref needed here}, followed by a series of computation and communication that conforms to certain pattern, which is usually the dominant part of application's execution.
-Before the application come to completion, all the working process will return their results to the root directly or hierarchically.} Parallel application's communication pattern usually involves lots of factors, such as communication intensity, operation-to-operation dependencies and the ``critical path" in its communication topology graph, etc.
-In this work, we focus on the locality of application's rank-to-rank communication topology graph.
-
-A parallel application usually conforms to a combination of several basic communication patterns \cite{roth}.
-At its different execution phases, the application's communication behavior may follow different basic patterns respectively.
-There are many profiling tools available to capture information regarding communication patterns of parallel applications~\cite{tau,mpip,scala,sst,oxbow}.
-They can provide information such as the percentage of different MPI operations of the applications, communication topology, the amount of data transferred between processes.
-
-There are many applications running on HPC systems, such as Mira \cite{bgq} and Titan \cite{titan}, that are communication intensive and conform to various communication patterns.
-In this work, we select three representative applications from the DOE Design Forward Project.
-Each application exhibits a distinctive communication pattern that is commonly seen in HPC applications.
-We believe that the communication patterns of these applications are representative of a wide array of applications running on leadership-class machines.
-Specifically, we study the Algebraic MultiGrid Solver (AMG), Geometric MultiGrid (MultiGrid) and CrystalRouter MiniApps.
-The communication pattern figures we present in Section~\ref{sec:amg}, \ref{sec:crystalrouter}, and \ref{sec:multigrid} are generated with the IPM~\cite{ipm} data gathered from publicly available traces~\cite{designforwardwebpage}.
-
-% MiniApps are reduced proxy applications that encapsulate the salient performance of larger full size applications \cite{miniapp}.
-Since we only focus on the dominant communication pattern of parallel applications, these MiniApps are perfect candidates for analysis the application's performance on different allocations or on new network architectures.
-We choose three MiniApps, namely, Algebraic Multigrid Solver(AMG), Geometric MultiGrid(MultiGrid) and CrystalRouter.
-Their representative communication patterns that are commonly seen in the HPC system workload.
-The communication pattern figures we present here are generated with the IPM \cite{ipm} data from \cite{designforwardwebpage}.
-
-\subsection{AMG}
-\label{sec:amg}
-
-\begin{figure}[t!]
-    \centering
-    \begin{subfigure}[t]{0.25\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/appstudy/amg/amg_pip}
-        \caption{}
-        \label{fig:amg-communication-topology}
-    \end{subfigure}%
-    ~
-    \begin{subfigure}[t]{0.24\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/appstudy/amg/amg_data_transfer}
-        \caption{}
-        \label{fig:amg-data-trans}
-    \end{subfigure}
-    \caption{AMG.
-(a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in AMG.
-(b) The amount of data sent/received by each rank in AMG.
-Y-axis labels the amount of data transferred in MB, x-axis labels 216 ranks of AMG.
-}
-\end{figure}
-
-The Algebraic MultiGrid Solver, or AMG, is a parallel algebraic multi-grid solver for linear systems arising from problems on unstructured mesh physics packages.
-It has been derived directly from the BoomerAMG solver that is being developed in the Center for Applied Scientific Computing (CASC) at LLNL \cite{amg}.
-The dominant communication pattern is regional communication with decreasing message size for different parts of the multi-grid v-cycle.
-
-Figure~\ref{fig:amg-communication-topology} shows the communication topology of a small scale AMG execution with 216 MPI Ranks.
-Note that the dominant communication pattern of the application doesn't change with scale.
-We can make the observation that AMG's dominant communication pattern is 3D Nearest Neighbor, each rank has intensive communication with up to six neighbors, depending on the boundaries of the ranks.
-Applications with similar pattern include PARTISN \cite{partisn} and SNAP \cite{snap}.
-
-Figure~\ref{fig:amg-data-trans} shows the amount of data each rank transferred in the example execution.
-The data transferring of each rank is between 7MB to 16MB, corresponding to whether the rank is on a boundary or not.
-
-\subsection{Crystal Router}
-\label{sec:crystalrouter}
-
-\begin{figure}[t!]
-    \centering
-    \begin{subfigure}[t]{0.25\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/appstudy/cr/cr_pip}
-        \caption{Communication Topology}
-        \label{fig:cr-communication-topology}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[t]{0.22\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/appstudy/cr/cr_data_transfer}
-        \caption{Data Volume}
-        \label{fig:cr-data-trans}
-    \end{subfigure}
-    \caption{CrystalRouter.
-(a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in CrystalRouter.
-(b) The amount of data sent/received by each rank in CrystalRouter.
-Y-axis labels the amount of data transferred in MB, x-axis labels 100 ranks of CrystalRouter.
-}
-\end{figure}
-
-The second MiniApp covered is CrystalRouter, the extracted communication kernel of the full application Nek5000~\cite{crystalrouter}, which is a spectral element CFD application developed at Argonne National Laboratory \cite{nek5000}.
-It features spectral element multi-grid solvers coupled with a highly scalable, parallel coarse-grid solver that is widely used for projects including ocean current modeling, thermal hydraulics of reactor cores, and spatiotemporal chaos.
-CrystalRouter demonstrates the ``many-to-many'' communication pattern through a scalable multi-stage communication process.
-The way CrystalRouter works is nodes compute for a while, then synchronize and communicate, continually alternating between these two types of activities.
-
-The collective communication in CrystalRouter utilizes a recursive doubling approach.
-Ranks in CrystalRouter conform to a $n$-dimensional hypercube and recursively split into ($n$-1)-dimensional hypercubes, with communication occurring along the splitting plane.
-The pattern of this communication can be found in Figure~\ref{fig:cr-communication-topology}.
-By nature of the logarithmic splitting process, a substantial portion of the communication occurs in small neighborhoods of ranks.
-CrystalRouter represents a group of applications whose dominant communication is a hybrid of multi-stage local and hierarchical global communication, and shares similarities with most MPI collective communication implementations.
-
-Figure~\ref{fig:cr-data-trans} shows the amount of data each rank transferred in CrystalRouter, which ranges between 20MB to 36MB.
-This is due to the rank counts not matching one-for-one with points in the hypercube.
-
-\subsection{MultiGrid}
-\label{sec:multigrid}
-
-\begin{figure}[t!]
-    \centering
-    \begin{subfigure}[t]{0.25\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/appstudy/mg/mg_pip}
-        \caption{Communication Topology}
-        \label{fig:mg-communication-topology}
-    \end{subfigure}
-    ~
-    \begin{subfigure}[t]{0.22\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/appstudy/mg/mg_data_transfer}
-        \caption{Data Volume}
-        \label{fig:mg-data-trans}
-    \end{subfigure}
-    \caption{MultiGrid.
-(a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in MultiGrid.
-(b) The amount of data sent/received by each rank in MultiGrid.
-Y-axis labels the amount of data transferred in MB, x-axis label 125 ranks of MultiGrid.
-}
-\end{figure}
-
-MultiGrid is geometric multi-grid v-cycle from the production elliptic solver BoxLib, a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes \cite{boxlib}.
-MultiGrid conforms to many-to-many communication pattern with decreasing message size and collectives for different parts of the multi-grid v-cycle.
-It is widely used for structured grid physics packages.
-
-Figure~\ref{fig:mg-communication-topology} shows the communication topology of MultiGrid with 125 ranks.
-We can see intensive communication along the diagonal that resembles nearest neighbor communication, similar to AMG.
-However, the communication topology leads to a greater ``spread'' of communication across the set of ranks, challenging the maximization of communication locality with respect to rank.
-In this sense, it can be considered a ``many-to-many'' pattern.
-Applications with such similar dominant communication pattern are like FillBoundary, another PDE solver code in \cite{boxlib}.
-
-Figure~\ref{fig:mg-data-trans} shows the amount of data transferred between ranks in MultiGrid. For this miniapp, The data transferred per rank is relatively small and stable, between 2.8MB to 2.9MB.
-
-\section{Research Vehicle}
-\label{sec:codes}
-
-It is difficult to accurately and flexibly experiment with concurrently running jobs in an HPC context.
-One reason is that the allocation strategy used on production machine is part of the system software, which can not be changed by users.
-Even the system administrator is not authorized to make that change.
-Another reason is that it is unrealistic to reserve the system exclusively to run the same job with desired allocation without interference and then compare the results with those in the presence of interference.
-Therefore, we resort to simulation for this work.
-
-A simulation toolkit named CODES enables the exploration of simulating different HPC networks with high fidelity \cite{Jason-2011}\cite{mubarak-sc2012}.
-CODES is built on top of Rensselaer Optimistic Simulation System (ROSS) parallel discrete-event simulator, which is capable of processing billions of events per second on leadership-class supercomputers \cite{ross}.
-CODES support both torus and dragonfly network with high fidelity flit-level simulation.
-CODES has this network workload component that is capable of conducting trace-driven simulations.
-It can take real MPI application traces generated by SST DUMPI \cite{sst} to drive CODES network models.
-In this work, we focus on an in-depth analysis of intra- and inter-job communication interferences with  different job allocations on torus-connected HPC systems.
-Torus networks have been extensively used in the current generation of supercomputers because of their linear scaling on per-node cost and competitive communication performance \cite{mubarak-sc2012}\cite{zhou-ipdps}.
-
-\section{Configuration Study}
-\label{sec:config study}
-
-In this section, we study the communication behavior of three applications on torus networks with different dimensionality and bandwidth configurations.
-Torus networks are $k$-ary $n$-cubes, with $k^{n}$ nodes in total arranged in an $n$-dimensional grid having $k$ nodes in each dimension.
-Each node has $2\times n$ direct linked neighbor nodes with wraparound.
-In these and all following experiments, non-adaptive routing is used.
-While making specific recommendations as to an ``optimal'' network configuration for a given application is out of the scope of this study, the experiments do highlight the nuances inherent in design choices such as network diameter and bandwidth planning.
-
-We perform three sets of experiments.
-Each looks at per-rank distribution of performance of AMG, CrystalRouter and MultiGrid on a set of 2048-node torus models, one with $n=3$ ($16^{2} \times 8$), one with $n=5$ ($8\times 4^{4}$), and one with $n=7$ ($4^{4} \times 2^{3}$).
-Performance in this context is defined by the total message transfer time for each sending rank, which we are able to precisely measure due to the simulation environment providing us with an absolute notion of time.
-Note that we simply perform linear rank assignment, one per node -- later sections investigate rank and node mapping effects.
-In the first, shown in Figure~\ref{fig:dimensionality-study}, we keep the link bandwidth constant at 2GiB/s in each direction, corresponding to Blue Gene/Q systems~\cite{bgq}, leading to aggregate node bandwidths of 12GiB/s, 20GiB/s, and 28GiB/s for the 3D, 5D, and 7D tori, respectively.
-In the second, shown in Figure~\ref{fig:bandwidth-study}, we fix the torus dimensionality to 5D.
-In the third, shown in Figure~\ref{fig: bandwidth-time-box}, we instead fix the per-node aggregate bandwidth (and thus the system aggregate bandwidth) and modify the torus dimensionality, resulting in per-link bandwidths shown in Table~\ref{tab: fix-bandwidth}.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/dimenstudy/amg_box}
-        \caption{AMG}
-        \label{fig:dimen-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/dimenstudy/cr_box}
-        \caption{CrystalRouter}
-        \label{fig:dimen-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/dimenstudy/mg_box}
-        \caption{MultiGrid}
-        \label{fig:dimen-mg}
-    \end{subfigure}%
-   \caption{Data Transfer Time of AMG, CrystalRouter and MultiGrid on 3D, 5D and 7D torus network, with constant per-link bandwidth.}
-   \label{fig:dimensionality-study}
-\end{figure*}
-
-In the fixed link bandwidth experiment, shown in Figure~\ref{fig:dimensionality-study}, the data transfer time of the three applications are significantly reduced as the dimensionality increases, so the applications can easily make use of the increased dimensionality (lower hops on average) and/or aggregate bandwidth (more links being concurrently used).
-Interestingly, AMG has the greatest degree of outliers compared to the others.
-We expect this to be a combination of non-optimal rank mappings causing a greater degree of variance due to the highly regular communication pattern and the distribution of message sizes.
-MultiGrid seems to benefit the most in terms of reducing the spread of the time distribution, though the median message time remains roughly within a factor of two.
-The time distribution in CrystalRouter seems to accrue the most benefits, likely due to the combination of cross-cutting and localized per-rank traffic.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/bandwidthstudy/amg_bw_box}
-        \caption{AMG}
-        \label{fig:bdwstudy-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/bandwidthstudy/cr_bw_box}
-        \caption{CrystalRouter}
-        \label{fig:bdwstudy-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/bandwidthstudy/mg_bw_box}
-        \caption{MultiGrid}
-        \label{fig:bdwstudy-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 5D torus network with 2GiB/s, 4GiB/s and GiB/s direct link bandwidth. }
-   \label{fig:bandwidth-study}
-\end{figure*}
-
-In the fixed dimensionality experiment, shown in Figure~\ref{fig:bandwidth-study}, the increased per-link bandwidth, expectedly, benefits each application roughly equally, reducing the time distributions linearly.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/samebdw/amg}
-        \caption{AMG}
-        \label{fig:samebd-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/samebdw/cr}
-        \caption{CrystalRouter}
-        \label{fig:samebd-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/samebdw/mg}
-        \caption{MultiGrid}
-        \label{fig:samebd-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 3D, 5D and 7D torus network with same aggregated bandwidth. }
-   \label{fig: bandwidth-time-box}
-\end{figure*}
-
-\begin{table}[ht]
-\begin{center}
-\caption{Torus networks with same aggregated bandwidth} 
-\label{tab: fix-bandwidth}
-\begin{tabular}{l c c c} 
-\toprule % Top horizontal line
-\toprule
-&\multicolumn{3}{c}{Bandwidth } \\
-\cmidrule(l){2-4}
-Dimension  & per-link & aggregate &\\ % Column names row
-\midrule % In-table horizontal line
-3D      & 4.67GiB/s  & 28GiB/s  & \\  % Content row 1
-\midrule % In-table horizontal line
-5D    & 2.8GiB/s  & 28GiB/s  &\\ % Content row 1
-\midrule % In-table horizontal line
-7D    & 2GiB/s & 28GiB/s & \\
-\midrule
-\bottomrule % Bottom horizontal line
-\end{tabular}
-\end{center}
-\end{table}
-
-In the fixed aggregate bandwidth experiment, with configuration in Table~\ref{tab: fix-bandwidth} and results shown in Figure~\ref{fig: bandwidth-time-box}, the results are more nuanced.
-At least for these applications, the benefits from increased dimensionality do not outweigh the reduction in per-link bandwidth.
-MultiGrid does see benefits in the increased dimensionality, particularly with respect to outliers, but has a similar median message time.
-CrystalRouter reduces performance as the dimensionality increases, but only by a small degree.
-AMG remains roughly the same, minus a dip in the maximum message time that could be due to rank-to-node mapping changes.
-
-\section{Interference Analysis}
-\label{sec:interference}
-
-\textbf{\emph{Intra-job interference}} refers to the network contention between ranks within each application.
-\textbf{\emph{Inter-job interference}} is introduced by concurrently running jobs that adjacent to each other sharing network resources.
-Communication variability due to such interference can cause application performance degradation.
-
-In this section we will study both kinds of interference in the context of torus network. 
-Since application's communication pattern won't change with its scale, we use AMG with 216 ranks, CrystalRouter with 100 ranks, and MultiGrid 125 ranks.
-And in order to accommodate three applications running concurrently with different allocations, we simulate a 2K-node 3D torus ($16^{2} \times 8$) network for the experiments.
-
-\subsection{Intra-Job Interference Analysis}
-\label{sec: introjob}
-
-We design two sets of experiments to study the intra-job interference of each application.
-In the first, shown in Figure \ref{fig:shapestudy}, we assign each application with allocations in \emph{three different shapes}.
-In the second, shown in Figure \ref{fig:diffmap}, we study the intra-job interference by using different mapping strategies for each application with given allocation shape.
-
-In the allocation shapes experiment, we select three possible shapes can can commonly seen on the 3D torus network.
-They are 3D balanced-cube, 3D unbalanced-cube, 2D mesh, as shown in Figure \ref{fig:alloc-shapes}.
-
-3D balanced-cube, shown as red in Figure \ref{fig:cont_sub1}, can guarantee the minimum \emph{average pair-wise distance} within the allocation.
-Some research studies like \cite{leung,abhinav-sc13} indicate that compact allocation can guarantee jobs with better performance.
-They use a variety of metrics to evaluate the compactness of the allocation, such as \emph{average pair-wise distance}, \emph{diameter} and \emph{contiguity}.
-In this work, we select 3D balanced-cube as the most compact allocation on a 3D torus network.
-
-3D unbalanced-cube, shown as green in Figure \ref{fig:cont_sub1}, is a rectangular prism, which is the possible allocation shape on the systems with asymmetric networks.
-For example, Cray XE6/XK7 systems are deployed 3D tori with Gemini routers.
-The network connections in the \emph{y}-direction have only half the bandwidth of the cables used in the \emph{x} and \emph{z} directions.
-In order to take advantage of the faster links in the \emph{x} and \emph{z} directions, job allocation starts from X-Z plane, which leads to a rectangular prism shaped allocation \cite{RF}.
-
-2D mesh, shown as blue in Figure \ref{fig:cont_sub1}, can be cut out from a single layer of the 3D torus.
-2D mesh is a very common allocation shape in torus network for both contiguous and non-contiguous allocation strategies.
-For example, Cray Application Level Placement Scheduler (ALPS) indexes the nodes in torus network into a list and makes job allocation by simply going through the list \cite{carl-cug}.
-When the list is obtained by sorting the nodes based on their spacial coordinates in the torus, allocation made off from this list will be 2D mesh.
-The IBM Blue/Gene Q supercomputer Mira at Argonne Leadership Computing Facility also allow its allocation partition configured into mesh \cite{zhou-ipdps}.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/allocation}
-        \caption{ }
-        \label{fig:cont_sub1}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit16}
-        \caption{ }
-        \label{fig:noncont_sub1}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit8}
-        \caption{ }
-        \label{fig:noncont_sub2}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.2\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit2}
-        \caption{ }
-        \label{fig:noncont_sub3}
-    \end{subfigure}%
-    \caption{(a) Contiguous allocation in three different shapes.
-``Blue" is 2D mesh, ``Green" is 3D-unbalanced cube and ``Red" is 3D-balanced cube.
-(b)-(d) Non-contiguous allocation.
-Each job is represented by a specific color.
-The nodes assigned to different jobs are interleaved, the size of allocation unit are 16, 8 and 2.}
-    \label{fig:alloc-shapes}
-\end{figure*}
-
-AMG with 3D ``Nearest Neighbor" communication pattern takes less time when being assigned with 3D-unbalanced allocation than 3D-balanced, as shown in Figure \ref{fig:shapstudy-amg}.
-And MultiGrid gets the best performance (shortest data transfer time) when running on 3D-balanced allocation, as shown in Figure \ref{fig:shapstudy-mg}.
-Since MultiGrid's communication pattern is ``Many-to-Many" dominant, 3D-balanced is the most compact allocation with shortest pair-wise distance between nodes, which can reduce the aggregated hops for transferring message among ranks in MultiGrid.
-As CrystalRouter has both dominant local communication and multi-stage global data transfer, the 3D-balance is also the best allocation, but its advantage over 3D-unbalanced is not that obvious as to MultiGrid, as shown in Figure \ref{fig:shapstudy-cr}.
-
-Lots of research studies have designed fancy allocation algorithms to provide application with the most compact allocation\cite{leung,LO}.
-However, as shown in our experiments, providing cubical allocation without considering application's communication pattern won't guarantee the best performance for every application.
-Compact allocation should be provided to application with global data transfer, such as those conforms to ``Many-to-Many" communication pattern.
-Application with dominant communication pattern like ``Nearest Neighbor" won't fully utilize the compact allocation and could get even better performance on a relaxed allocation shape.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/amg_box}
-        \caption{AMG}
-        \label{fig:shapstudy-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/cr_box}
-        \caption{CrystalRouter}
-        \label{fig:shapstudy-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/mg_box}
-        \caption{MultiGrid}
-        \label{fig:shapstudy-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 3D-balanced, 3D-unbalanced and 2D allocation.}
-   \label{fig:shapestudy}
-\end{figure*}
-
-The rank-to-node mapping of parallel applications on HPC system could greatly impact applications' performance.
-However, finding the optimal mapping solution for a given application is out the scope of this work.
-Our experiments aims to show how mapping strategies impact the intra-job interference of application with specific communication pattern.
-
-We provide AMG, CrystalRouter and MultiGrid with 3D-balance allocation and use three mapping strategies to do the rank-to-node mapping.
-The ``Linear" mapping, as we used in the previous experiments, is to map each rank according to the dimensional ordering of compute nodes.
-The ``Cube" mapping assigns ranks into consecutive $2^{3}$ cubes.
-The ``Random" mapping assigns ranks randomly within the allocation.
-
-AMG remains roughly the same when it been mapped by ``Linear" and ``Cube", shown in Figure \ref{fig:diffmap-amg}.
-Although ``Linear" and ``Cube" mapping strategies cause some routing overlap for AMG's communication, both mapping strategies still preserve the locality of AMG's 3D ``Nearest Neighbor" communication pattern.
-The ``Random" mapping totally destroys AMG's communication pattern, and results in intra-job interference among all the ranks in AMG.
-The performance degradation caused by using ``Random" mapping is up to 90\%.
-
-The ``Cube" mapping improves CrystalRouter's performance over the ``Linear" by reducing the pair-wise distance between ranks, shown in Figure \ref{fig:diffmap-cr}.
-The global data transfer in CrystalRouter will takes fewer hops with ``Cube" mapping.
-The ``Random" mapping scatters the ranks within the local neighborhood of CrystalRouter and make their communication less efficient.
-
-The ``Cube" mapping benefits MultiGrid's ``Many-to-Many" communication.
-However, due to the small amount of data transfer among ranks in MultiGrid, ``Cube" mapping doesn't exhibit a drastic advantage.
-``Random" mapping neither cause too much degradation for the same reason, shown in Figure \ref{fig:diffmap-mg}.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/diffmapping/amg}
-        \caption{AMG}
-        \label{fig:diffmap-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/diffmapping/cr}
-        \caption{CrystalRouter}
-        \label{fig:diffmap-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/diffmapping/mg}
-        \caption{MultiGrid}
-        \label{fig:diffmap-mg}
-    \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 3D-balanced allocation using  different mapping strategies.}
-   \label{fig:diffmap}
-\end{figure*}
-
-\subsection{Inter-Job Interference Analysis}
-\label{sec: interjob interference study}
-
-Inter-job interference has been identified as one of the major culprits for application's performance variability \cite{abhinav-sc13,skinner,rosenthal}.
-Inter-job interference is a more prominent issue for system adopted non-contiguous allocation policy than system with contiguous allocation.
-Application communication times have been demonstrated to vary from 36\% faster to 69\% slower due to job interference when running on non-contiguous allocation \cite{abhinav-sc13}.
-
-We assign each application with non-contiguous allocation and run them concurrently on the same network.
-The allocation unit belongs to different jobs are interleaved.
-In order to study the impact of different allocation unit size on applications' inter-job interference, we conduct experiments with unit size of 16, 8 and 2, shown respectively in Figure \ref{fig:noncont_sub1}, \ref{fig:noncont_sub1} and \ref{fig:noncont_sub3}.
-Figure \ref{fig:interjobstudy} shows the results of each application data transfer time with different allocation unit size.
-
-The data transfer time of AMG in Figure \ref{fig:interjob-amg} keeps stable between allocation unit size of 16 and 8, since the locality of AMG is 6-rank based.
-When the unit size reduce to 2, AMG suffers prolong data transfer time by about 10\%.
-CrystalRouter is more sensitive to allocation unit size.
-The best unit size would be the one big enough to accommodate its local neighborhood.
-Figure \ref{fig:interjob-cr} shows unit size of 16 and 8 can guarantee the same average data transfer time, while some rank spend more time with allocation unit size 8 than 16.
-When the unit size reduce to 2, the communication become less efficiency and takes about 15\% more times for transferring data.
-
-The data transfer time of MultiGrid with different allocation unit sizes doesn't show obvious variability in Figure \ref{fig:interjob-mg}.
-This is because even big allocation unit size like 16 will still fail to preserve MultiGrid's ``Many-to-Many" pattern.
-The data transfer time is almost doubled when MultiGrid running concurrently with allocation unit size of 16, as shown in Figure \ref{fig:interjob-mg} .
-As the unit size decreases, the data transfer time keep growing, but in a slow way, which means in terms of preserving the locality of MultiGrid, allocation unit of size 16, 8 and 2 serve equally bad.
-
-The unit size in non-contiguous allocation policy should be chosen according to application's communication pattern.
-Inter-job interference is inevitable in non-contiguous based systems, but unit size that big enough to preserve the neighborhood communication of the application will alleviate such interference and improve job performance.
-
-\begin{figure*}[t!]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/inter-job/amg}
-        \caption{AMG}
-        \label{fig:interjob-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/inter-job/cr}
-        \caption{CrystalRouter}
-        \label{fig:interjob-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/inter-job/mg}
-        \caption{MultiGrid}
-        \label{fig:interjob-mg}
-    \end{subfigure}%
-   \caption{Inter-job interference study.
-``cont" indicates three applications running side by side concurrently on the same network with contiguous allocation.
-To study the impact of non-contiguous allocation on inter-job interference, applications running concurrently with interleaved allocation of different unit sizes, which are 16-node, 8-node, 2-node.
-}
-   \label{fig:interjobstudy}
-\end{figure*}
-
-\subsection{Result Summary}
-\label{sec:summary}
-
-Based on the above comprehensive study, we can make following observations:
-\begin{itemize}
-    \item The compact allocation may not guarantee the best performance for every application. 
-    \item The applications dominated with Nearest Neighbor communications exhibit relatively stable performance under different allocation shapes, while the applications dominated with Many-to-Many communication exhibit better performance with compact allocation (e.g.
-3D balanced).
-    \item A good rank-to-node mapping strategy can greatly improve application's performance when specific allocation is given.
-    \item An optimal size for allocation units should be determined according to an application's dominant communication pattern.
-In general, a unit size should be large enough to preserve neighboring communication of the application.
-    \item Inter-job interference is inevitable in non-contiguous allocation.
-However, choosing the proper allocation unit size with job's communication pattern awareness can alleviate such negative effect.
-\end{itemize}
-
-\section{Discussion}
-\label{sec:discussion}
-
-The results shown in the previous section provide insights for the design of smart and flexible job allocation.
-By scrutinizing job's communication behavior, we can identify job's dominant communication pattern and pinpoint the ``neighborhood communication".
-With such knowledge about jobs' communication patterns, we can analyze the possible interference between jobs and take precautions to alleviate the negative effect when making allocation decisions.
-
-When an application with dominant communication that is intensive ``Many-to-Many", the scheduler should grant it with compact node allocation and exclusive network provision.
-The compact allocation can guarantee the shortest pair-wise distance between all the ranks, and make the data transfer between ranks take less hops.
-On the other hand, the exclusive network provision will prevent other adjacent applications from sharing network resources, thus eliminate the performance degradation due to interference.
-
-Not every application's preferable resource are compact node allocation and exclusive network provision.
-As we found in our analysis, applications whose dominant communication patterns contain intensive ``neighborhood communication" like ``Nearest Neighbor" don't benefit from compact node allocation and exclusive network provision.
-Applications, such as AMG, can run with non-contiguous node allocation without performance degradation, as long as the allocation unit can accommodate their ``neighborhood communication".
-
-The allocation unit size shouldn't be fixed.
-Instead, the scheduler should choose a proper unit size based on the scale of ``neighborhood communication" of each job.
-The best unit size should be neither too big nor too small, jut perfectly fit for the size of that ``neighborhood".
-Big unit size won't be fully utilized and also cause fragmentation.
-Small unit size won't be able to accommodate the ``neighborhood communication" and makes the intra-job communication less efficient.
-
-The advantage of making allocation with consideration about job's communication pattern is obvious.
-First, compared with contiguous allocation policy, the new scheduler outlined above is more flexible, it doesn't require the system to provide a big contiguous partition to accommodate the whole application, instead it only needs to provide a small set of compact nodes that is sufficient for all the "local community" in the application.
-On the other hand, the new scheduler outlined above won't destroy the locality of application's communication pattern.
-Indeed, the small compact node set (allocation unit) provided for the application's  ``local communities" can preserve the communication locality in the maximum extent.
-The design of such a smart scheduler is part of our future work.
-
-\section{Related Work}
-\label{sec:related_work}
-
-There are many tools available for system monitoring and application profiling.
-Tools like TAU (Tuning and Analysis) \cite{tau}, SCALASCA \cite{scala} and mpiP \cite{mpip} can capture application's runtime information about application's communication and computation in event traces.
-However, recognizing the patterns about application's communication from those traces requires substantial effort.
-There are a number of studies on the recognition and characterization of parallel application communication patterns.
-Oak Ridge National Laboratory has its ongoing project about developing a tool set named Oxbow, which can characterize the computation and communication behavior of scientific applications and benchmarks \cite{oxbow}.
-In a recent work \cite{roth}, the authors demonstrate a new approach to automatically characterizing parallel applications' communication behaviors.
-
-%\textbf{P2: communication characteristics}\\
-
-Many research efforts have been conducted to characterize scientific applications.
-For instance, a DOE project aims at the identification of the computational characteristics of the DOE MiniApps developed at various exascale co-design centers \cite{designforwardwebpage}.
-In this project, the communication patterns of several DOE full applications and associated mini-applications are studied to provide a more complete snapshot of the DOE workload.
-A joint project named CORAL from Oak Ridge, Argonne and Livermore provide a series of benchmarks to represent DOE workloads and technical requirements \cite{coral}.
-The CORAL project includes scalable science benchmarks, throughput benchmarks, data centric benchmarks, skeleton benchmarks and Micro benchmarks.
-
-%\textbf{P3: Interference between concurrently running jobs}\\
-The interference between concurrently running jobs on HPC system have been identified as major culprit for job's performance variability.
-Abhinav et al.
-found that concurrently running applications on HPC can cause interference to each other, and cause communication time varied from 36\% faster to 69\% slower.
-Such interference could come from Operating System noise, shape of the allocated partition , or other running jobs that sharing network resources \cite{abhinav-sc13}.
-Skinner et al.
-found that  there is a 2-3$\times$ slowdown in MPI$\textunderscore$Allreduce due to network contention from other jobs \cite{skinner}.
-Rosenthal et al.
-found increasing network bandwidth provide limited benefits to a number of applications.
-The applications that send mostly small messages or large messages asynchronously are not bandwidth bounded, hence, benefit only slightly or not at all from increased bandwidth \cite{rosenthal}.
-
-%\textbf{P4: Job allocation(little bit)}\\
-Several research studies focus on optimizing job allocation on HPC system to alleviate the interference between concurrently running jobs.
-Hoefler et al.
-proposed to use performance modeling techniques to analyze factors that impact the performance of parallel scientific applications \cite{hoefler-modeling}.
-However, as the scale of HPC continue to grow, the  interference of concurrently running jobs is getting worse, which is hard to be quantified by performance profiling tools alone.
-Bogdan et al.
-provide a set of guidelines of how to configure a network with Dragonfly topology for workload with Nearest Neighbor communication pattern \cite{Bogdan-hpdc14}.
-Dong et al.
-have developed simple benchmarks that conforms to four different communication patterns, namely ping-pong, nearest neighbor, broadcast and all$\textunderscore$reduce, to demonstrate the effectiveness of this highly parallel 5D torus network \cite{Dong-SC11}.
-
-Our work is different from all these works in the following ways.
-First, we focus on the dominant communication patterns rather than any specific application.
-We believe this can provide a guideline for other research work.
-Secondly, we explore the intra- and inter-job interference between concurrently running jobs, while similar work such as \cite{abhinav-sc13} only focuses on the single application's performance degradation due to network contention.
-Finally, we analyze the impact of different allocation strategies to job's communication behaviors.
-We further identify the optimal allocation strategy for each application with a specific dominant communication pattern.
-Based on our study, we claim that further scheduling should take job's communication pattern into consideration for its allocation decision making.
-
-\section{Conclusions}
-\label{sec:conclusion}
-In this work, we have studied the communication behavior of three parallel applications, namely AMG, CrystalRouter and MultiGrid.
-Each application has distinctive communication pattern, that can be representative for a group of jobs in HPC workload.
-We have used a sophisticate simulation tool named CODES from Argonne National Laboratory to simulate the running of these three parallel applications on torus network.
-The torus network provided by CODES has good fidelity and scalability.
-We have analyzed the performance of each application's communication in terms of data transfer time by simulating them running on torus networks with different bandwidth and dimensionality configurations.
-We have found that higher dimensionality of torus network would improve the performance of application with ``Many-to-Many" communication patterns, while application with intensive local communication like ``Nearest Neighbor" won't benefit much from higher dimensionality.
-
-We have analyzed the intra- and inter-job interference by simulating three applications running on 3D torus network.
-Based on our comprehensive experiments, we got five observations.
-1) The compact allocation may not guarantee the best performance for every application.
-2)The applications dominated with Nearest Neighbor communications exhibit relatively stable performance under different allocation shapes, while the applications dominated with Many-to-Many communication exhibit better performance with compact allocation (e.g.
-3D balanced).
-3) A good rank-to-node mapping strategy can greatly improve application's performance when specific allocation is given.
-4) An optimal size for allocation units should be determined according to an application's dominant communication pattern.
-In general, a unit size should be large enough to preserve neighboring communication of the application.
-5) Inter-job interference is inevitable in non-contiguous allocation.
-However, choosing the proper allocation unit size with job's communication pattern awareness can alleviate such negative effect.
-
-We believe that our finding in this work can provide valuable guidance for HPC resource management to make flexible job allocations.
-Rather than using pre-defined partitions and non-contiguous allocation, future HPC systems should assign each job with preferable resources based on job's communication pattern.
+\input{intro}
+\input{app_study}
+\input{research_tool}
+%\input{config_study}
+\input{interference_study}
+\input{discussion}
+\input{related_work}
+\input{conclusion}
 

### 2016-07-08 submission.tex
change some reference format

@@
 
-Network contention between concurrently running jobs on HPC systems is a primary cause of performance variability.
-Optimizing job allocation and avoiding network sharing are hence crucial to alleviate the potential degradation in performance.
-In order to do these effectively, an understanding of the interaction between applications, their communication patterns, and the system architecture are required.
-In this work, we choose three representative HPC applications from DOE Design Forward Project and conduct detailed simulations of high-performance network topologies to analyze both intra- and inter-job interference.
-By scrutinizing the communication behaviors of these applications, we identify relationships between these behaviors and the possible interference introduced by different job allocation scenarios.
-Our comprehensive analyses in this work illuminate a path towards job-communication-pattern-aware HPC resource allocation.
+Network contention between concurrently running jobs on HPC systems 
+is a primary cause of performance variability. 
+Optimizing job allocation and avoiding network sharing are 
+hence crucial to alleviate the potential performance degradation. 
+In order to do these effectively, 
+an understanding of the interference among the concurrently running jobs, 
+their communication patterns, and the network contention are required. 
+In this work, we choose three representative HPC applications from 
+DOE Design Forward Project and conduct detailed simulations of 
+on a Torus network model to analyze both intra- and inter-job interference. 
+By scrutinizing the communication behaviors of these applications, 
+we identify relationships between these behaviors and 
+the possible interference introduced by different job placement policies. 
+Our comprehensive analyses in this work illuminate a path 
+towards job-communication-pattern-aware HPC resource allocation.
 
@@
 \bibliographystyle{IEEEtran}
-\bibliography{./reference}
+\bibliography{IEEEabrv, reference}
 

### 2016-07-08 app_study.tex
pass on Introduction, remove some citations

@@
 
-A parallel application usually conforms to a combination of several basic communication patterns \cite{roth}.
-At its different execution phases, the application's communication behavior may follow different basic patterns respectively.
-There are many profiling tools available to capture information regarding communication patterns of parallel applications~\cite{tau,mpip,scala,sst,oxbow}.
-They can provide information such as the percentage of different MPI operations of the applications, communication topology, the amount of data transferred between processes.
+A parallel application usually conforms to a combination of several basic communication patterns~\cite{roth}. 
+At its different execution phases, 
+the application's communication behavior may follow certain basic patterns respectively. 
+There are many profiling tools available to capture 
+information regarding communication patterns of parallel applications~\cite{tau,mpip,sst,oxbow}. 
+They can provide information such as the percentage of different MPI operations, 
+communication topology, the amount of data transferred between processes, etc.
 
-There are many applications running on HPC systems, such as Mira \cite{bgq} and Titan \cite{titan}, that are communication intensive and conform to various communication patterns.
+There are many applications running on HPC systems such as Mira~\cite{bgq} and Titan~\cite{titan}. 
+These applications are communication intensive and conform to various communication patterns. 
 In this work, we select three representative applications from the DOE Design Forward Project. 
 Each application exhibits a distinctive communication pattern that is commonly seen in HPC applications. 
-We believe that the communication patterns of these applications are representative of a wide array of applications running on leadership-class machines.
-Specifically, we study the Algebraic MultiGrid Solver (AMG), Geometric MultiGrid (MultiGrid) and CrystalRouter MiniApps.
-The communication pattern figures we present in Section~\ref{sec:amg}, \ref{sec:crystalrouter}, and \ref{sec:multigrid} are generated with the IPM~\cite{ipm} data gathered from publicly available traces~\cite{designforwardwebpage}.
+We believe that the communication patterns of these applications 
+are representative of a wide array of applications running on leadership-class machines. 
+Specifically, we study the Algebraic MultiGrid Solver (AMG), 
+Geometric MultiGrid (MultiGrid) and CrystalRouter MiniApps. 
+The communication matrix presented in Figure~\ref{fig:apps_communication_matrix}
+are generated with the IPM~\cite{ipm} data gathered from publicly available traces~\cite{designforwardwebpage}.
 
@@
     \caption{Applications rank-to-rank communication topology graph.}
+    \label{fig:apps_communication_matrix}
 \end{figure*}

### 2016-07-08 discussion.tex
pass on Introduction, remove some citations

@@
 The advantage of making allocation with consideration about job's communication pattern is obvious.
-First, compared with contiguous allocation policy, the new scheduler outlined above is more flexible, it doesn't require the system to provide a big contiguous partition to accommodate the whole application, instead it only needs to provide a small set of compact nodes that is sufficient for all the "local community" in the application.
+First, compared with contiguous allocation policy, the new scheduler outlined above is more flexible, without requiring the system to provide a big contiguous partition to accommodate the whole application, it only needs to provide a small set of compact nodes that is sufficient for all the "local community" in the application.
 On the other hand, the new scheduler outlined above won't destroy the locality of application's communication pattern.

### 2016-07-08 interference_study.tex
pass on Introduction, remove some citations

@@
 
-Lots of research studies have designed fancy allocation algorithms to provide application with the most compact allocation\cite{leung,LO}.
-However, as shown in our experiments, providing cubical allocation without considering application's communication pattern won't guarantee the best performance for every application.
+Lots of research studies have designed complex allocation algorithms to provide application with the most compact allocation\cite{leung,LO}.
+However, as shown in our experiments, providing cubical allocation without considering application's communication pattern can not guarantee the best performance for every application.
 Compact allocation should be provided to application with global data transfer, such as those conforms to ``Many-to-Many" communication pattern.
@@
 The ``Cube" mapping benefits MultiGrid's ``Many-to-Many" communication.
-However, due to the small amount of data transfer among ranks in MultiGrid, ``Cube" mapping doesn't exhibit a drastic advantage.
+However, due to the small amount of data transfer among ranks in MultiGrid, ``Cube" mapping fails to exhibit a drastic advantage.
 ``Random" mapping neither cause too much degradation for the same reason, shown in Figure \ref{fig:diffmap-mg}.

### 2016-07-08 intro.tex
pass on Introduction, remove some citations

@@
 
-The scale of supercomputers continually grows at a breakneck pace to accommodate the computing power requirements for scientific research.
-Supercomputers have tens of thousands of nodes and serve as an irreplaceable research vehicle for scientific problems with increasing size and complexity.
-Supercomputers are usually employed as a shared resource to accommodate many parallel applications (jobs) running concurrently.
-These parallel jobs share the system infrastructure such as network and I/O bandwidth, and inevitably there is contention over these shared resources.
+The scale of supercomputers continually grows at a breakneck pace to 
+accommodate the computing power requirements for scientific research. 
+Supercomputers have tens of thousands of nodes and serve as an irreplaceable 
+research vehicle for scientific problems with increasing size and complexity. 
+Supercomputers are usually employed as a shared resource to 
+accommodate many parallel applications (jobs) running concurrently. 
+These parallel jobs share the system infrastructure such as network and I/O bandwidth, 
+and inevitably there is contention over these shared resources. 
 As supercomputers continue to evolve, these shared resources are increasingly the bottleneck for performance.
@@
     \centering
-    \begin{subfigure}[t]{0.2\textwidth}
+    \begin{subfigure}[t]{0.22\textwidth}
         \centering
@@
     \hspace{1em}%
-    \begin{subfigure}[t]{0.2\textwidth}
+    \begin{subfigure}[t]{0.22\textwidth}
         \centering
@@
    Each job is represented by a specific color. 
-a) shows the effect of contiguous allocation, which reduce the inter-job interference.
-b) shows non-contiguous allocation, which may introduce both intra and inter-job interferences.
+   a) shows the effect of contiguous allocation, 
+   which could possibly reduce the inter-job interference. 
+   b) shows non-contiguous allocation, 
+   which may introduce both intra and inter-job interference. 
    }
@@
 
-When submitting jobs to HPC systems, users request resources in terms of node counts and expected runtime.
+When submitting jobs to HPC systems, 
+users request system resources by specifying the number of compute nodes and expected runtime. 
 The batch scheduler is then responsible for dispatching the submitted jobs to the system. 
-Typically, multiple jobs run concurrently, resulting in the shared use of resources, particularly network and storage.
-A prominent problem with this approach is network contention among concurrently running jobs.
-If the network resources of concurrently running jobs aren't isolated, the resulting sharing can cause communication variability that results in performance degradation for some or all such jobs \cite{abhinav-sc13}.
-This delay can propagate into the queueing time of the following submitted jobs, thus leading to low system throughput and utilization \cite{jose-ipdps15}.
-This adverse effect of network sharing can be mitigated by providing jobs with isolated allocations and exclusive network resources.
+Typically, there are multiple jobs running concurrently on the system,
+resulting in the shared use of resources, particularly network links. 
+A prominent problem with this network sharing is the contention over
+the network resource among those concurrently running jobs. 
+The network contention can cause communication variability and 
+performance degradation to those jobs~\cite{abhinav-sc13}. 
+This performance degradation can propagate into the queueing time of the following submitted jobs, 
+thus leading to low system throughput and utilization~\cite{jose-ipdps15}. 
+This adverse effect of network sharing can be mitigated by 
+providing jobs with isolated allocations and exclusive network resources.
 
-On the widely used torus-connected HPC systems \cite{bgq,tofu,titan}, two allocation strategies are commonly used.
-The \emph{contiguous allocation} strategy assigns to each job a compact and contiguous set of computing nodes, as shown in Figure \ref{fig:overview_sub1}.
-The partition-based allocation approach used in Blue Gene series systems is an example of such a strategy \cite{bgloverview}.
-The contiguous strategy favors application performance through isolated networking within a partition and the locality that implies.
-However, this strategy can cause both internal fragmentation (when more nodes are allocated to a job than it requests) and external fragmentation (when sufficient nodes are available for a request, but they can not be allocated contiguously), therefore leading to lower system utilization than is otherwise possible.
-On the other side of the coin, the \emph{non-contiguous allocation} strategy, used by the Cray XT/XE series \cite{carl-cug}, assigns free nodes to jobs regardless of contiguity, though of course efforts are made to maximize locality.
-Figure~\ref{fig:overview_sub2} shows an example non-contiguous allocation.
-While eliminating internal and external fragmentation as seen in contiguous allocation systems, in return non-contiguous allocations introduce contention between jobs due to the interleaving of job nodes.
-The non-contiguous node allocation can significantly reduce job performance, especially for communication-intensive ones \cite{abhinav-sc13}.
+On the widely used torus-connected HPC systems~\cite{bgq,tofu,titan}, 
+two job placement policies are commonly used. 
+In \emph{contiguous placement} policy, 
+each job gets a compact and contiguous set of computing nodes, 
+as shown in Figure~\ref{fig:overview_sub1}. 
+The partition-based placement adopted on the Blue Gene 
+series systems is an example of contiguous placement policy~\cite{bgloverview}. 
+The contiguous placement favors application performance 
+through exclusive networking within a compact partition 
+and the locality that implies. 
+However, contiguous placement can cause both internal fragmentation 
+(when more nodes are allocated to a job than it requests) 
+and external fragmentation 
+(when sufficient nodes are available for a request, but they can not be allocated contiguously), 
+therefore leading to lower system utilization than is otherwise possible. 
+On the other hand, the \emph{non-contiguous placement} policy, 
+adopted by the Cray XT/XE series~\cite{carl-cug}, 
+assigns free nodes to jobs regardless of contiguity, 
+though of course efforts are made to maximize locality. 
+Figure~\ref{fig:overview_sub2} shows the effect of non-contiguous placement. 
+While eliminating internal and external fragmentation as seen in contiguous allocation systems, 
+in return non-contiguous allocations introduce contention between jobs due to the interleaving of job nodes. 
+The non-contiguous job placement can significantly reduce job performance, 
+especially for communication-intensive ones~\cite{abhinav-sc13}.
 
-We envision that future HPC schedulers will adopt a flexible job allocation mechanism which combines the best of both contiguous and non-contiguous allocation strategies.
-Such a flexible mechanism would take shared resource needs of jobs into account when making allocation decisions (e.g., network).
-With knowledge and analysis of job communication patterns, it can be identified which jobs require network isolation and locality, and to what degree.
-Then, rather than allocating each job in a ``know-nothing'' manner, one may specialize allocation so that, for example, only the jobs with stringent network needs are given compact, isolated allocations, resulting in maximized utilization and minimized perceivable resource contention effects.
+We envision that future HPC systems will be equipped with 
+a flexible job placement mechanism which combines the 
+best of both contiguous and non-contiguous placement. 
+Such a flexible job placement mechanism should take 
+shared resource needs (e.g., network resources) of jobs into account 
+when making scheduling and allocation decisions. 
+With the knowledge and analysis of job communication patterns, 
+it can be identified which jobs require exclusive network and allocation locality, and to what degree. 
+Then, rather than allocating each job in a ``know-nothing'' manner, 
+one may specialize allocation so that, for example, 
+only the jobs with stringent network needs are given compact, isolated allocations, 
+resulting in maximized utilization and minimized perceivable resource contention effects.
 
-In this work, we focus on an in-depth analysis of intra- and inter-job communication interference with different job allocation strategies on torus-connected HPC systems.
-Torus-based networks are used on six of the top 10 supercomputers on the June 2015 Top500 lists~\cite{top500}.
-The current generation of IBM Blue Gene/Q (BG/Q) supercomputers, such as Mira at Argonne Nation Laboratory and Sequoia at Lawrence Livermore National Laboratory, has its nodes connected in a 5D torus network \cite{bgq}.
-The K computer from Japan uses the ``Tofu'' system, which is a 6D mesh/torus topology \cite{tofu}.
-Titan, a Cray XK7 supercomputer located at the Oak Ridge Leadership Computing Facility (OLCF), has nodes connected in a 3D torus within the compute partition \cite{titan}.
-Although our analyses are based on torus networks, the ideas conveyed in this work is applicable to networks with different topologies.
+In this work, we focus on an in-depth analysis of intra- and inter-job 
+communication interference with different job placement on torus-connected HPC systems. 
+Torus-based networks are used on six of the top 10 supercomputers 
+on the June 2015 Top500 lists~\cite{top500}. 
+The current generation of IBM Blue Gene/Q (BG/Q) supercomputers, 
+such as Mira at Argonne Nation Laboratory and 
+Sequoia at Lawrence Livermore National Laboratory, 
+have their compute nodes connected by a 5D torus network~\cite{bgq}. 
+The K computer from Japan uses the ``Tofu'' interconnected system, 
+which has a 6D mesh/torus topology~\cite{tofu}. 
+Titan, a Cray XK7 supercomputer located at the Oak Ridge Leadership Computing Facility (OLCF), 
+has nodes connected in a 3D torus within the compute partition~\cite{titan}. 
+Although our analyses are based on torus networks, 
+the ideas conveyed in this work is applicable to networks with different topologies.
 
-We selected three signature applications from the DOE Design Forward Project~\cite{designforwardwebpage} as examples to conduct detailed study about their communication patterns.
-We use a sophisticated simulation toolkit named CODES (standing for Co-Design of Multi-layer Exascale Storage Architectures)~\cite{Jason-2011} as a research vehicle to evaluate the performance of these applications with various allocations in a controlled environment.
-We then analyze the intra- and inter-job interference by simulating these applications running exclusively and concurrently with different allocations.
-We believe the insights presented in this work can be very useful for the design of future HPC batch job schedulers and resource managers.
-
-%\NOTE{JJ removed the list of contributions, because they are exactly the same as the preceding paragraph.}
-\begin{comment}
-The observations of this work are the following:
-
-\begin{itemize}
-    \item We study the communication behaviors and identify the communication patterns of three representative HPC applications.
-    \item We conduct detailed analysis about the intra- and inter-job interference by running three applications with different allocation strategies.
-And we find the relation between application's communication pattern and the cause of interference.
-    \item We show a novel and flexible way to make allocation with awareness of job's communication pattern, which  is every effective to alleviate the interference between concurrently running jobs.
-\TODO{do we actually do this, or provide the underpinnings for such a method?}
-\end{itemize}
-
-We believe the finds and insights presented in this work would be very useful for the design of future HPC batch job scheduler and resource management module.
-
-\end{comment}
+We selected three signature applications from 
+the DOE Design Forward Project~\cite{designforwardwebpage} as examples 
+to conduct detailed study about the communication pattern of parallel applications. 
+We use a sophisticated simulation toolkit named CODES 
+(standing for Co-Design of Multi-layer Exascale Storage Architectures)~\cite{Jason-2011} 
+as a research vehicle to evaluate the performance of these applications 
+with various allocations in a controlled environment. 
+We then analyze the intra- and inter-job interference 
+by simulating these applications running concurrently with different allocations. 
+We believe the insights presented in this work can be very useful 
+for the design of future HPC batch schedulers and resource managers.
 
 The rest of this paper is organized as follows. 
-Section~\ref{sec:application study} describes the three representative applications from DOE Design Forward Project for our study.
+Section~\ref{sec:application study} describes the three representative applications 
+from the DOE Design Forward Project for our study. 
 Section~\ref{sec:codes} talks about the use of CODES as research vehicle for our work. 
-Section~\ref{sec:config study} shows the performance analysis of three applications on different torus networks.
-Section~\ref{sec:interference} provides detailed analysis about the intra- and inter-job interference between three applications on torus network with different allocation strategies.
-Section~\ref{sec:discussion} introduces a path toward communication-pattern aware allocation strategies, given the results of our analysis.
+%Section~\ref{sec:config study} shows the performance analysis of three applications on different torus networks. 
+Section~\ref{sec:interference} provides detailed analysis about the intra- and inter-job interference 
+among three applications on torus network with different allocations. 
+Section~\ref{sec:discussion} introduces a path toward communication-pattern aware allocation strategies, 
+given the results of our analysis. 
 Section~\ref{sec:related_work} discusses related work. 

### 2016-07-08 related_work.tex
pass on Introduction, remove some citations

@@
 
-There are many tools available for system monitoring and application profiling.
-Tools like TAU (Tuning and Analysis) \cite{tau}, SCALASCA \cite{scala} and mpiP \cite{mpip} can capture application's runtime information about application's communication and computation in event traces.
+There are many tools available for system monitoring and application profiling. Tools like TAU (Tuning and Analysis)~\cite{tau}, 
+and mpiP~\cite{mpip} can capture application's runtime information about application's communication and computation in event traces.
 However, recognizing the patterns about application's communication from those traces requires substantial effort.
 There are a number of studies on the recognition and characterization of parallel application communication patterns.
-Oak Ridge National Laboratory has its ongoing project about developing a tool set named Oxbow, which can characterize the computation and communication behavior of scientific applications and benchmarks \cite{oxbow}.
-In a recent work \cite{roth}, the authors demonstrate a new approach to automatically characterizing parallel applications' communication behaviors.
+Oak Ridge National Laboratory has its ongoing project about developing a tool set named Oxbow, which can characterize the computation and communication behavior of scientific applications and benchmarks~\cite{oxbow}.
+In a recent work~\cite{roth}, the authors demonstrate a new approach to automatically characterizing parallel applications' communication behaviors.
 

### 2016-07-08 app_study.tex
pass on Application Study

@@
 are generated with the IPM~\cite{ipm} data gathered from publicly available traces~\cite{designforwardwebpage}.
-
-% MiniApps are reduced proxy applications that encapsulate the salient performance of larger full size applications \cite{miniapp}.
-Since we only focus on the dominant communication pattern of parallel applications, these MiniApps are perfect candidates for analysis the application's performance on different allocations or on new network architectures.
-We choose three MiniApps, namely, Algebraic Multigrid Solver(AMG), Geometric MultiGrid(MultiGrid) and CrystalRouter.
-Their representative communication patterns that are commonly seen in the HPC system workload.
-The communication pattern figures we present here are generated with the IPM \cite{ipm} data from \cite{designforwardwebpage}.
 
@@
 
-The Algebraic MultiGrid Solver, or AMG, is a parallel algebraic multi-grid solver for linear systems arising from problems on unstructured mesh physics packages.
-It has been derived directly from the BoomerAMG solver that is being developed in the Center for Applied Scientific Computing (CASC) at LLNL \cite{amg}.
-The dominant communication pattern is regional communication with decreasing message size for different parts of the multi-grid v-cycle.
+The Algebraic MultiGrid Solver, or AMG, 
+is a parallel algebraic multi-grid solver for linear systems arising from problems on unstructured mesh physics packages. 
+It has been derived directly from the BoomerAMG solver 
+that is being developed in the Center for Applied Scientific Computing (CASC) at LLNL \cite{amg}. 
+The dominant communication pattern is regional communication 
+with decreasing message size for different parts of the multi-grid v-cycle.
 
-Figure~\ref{fig:amg-communication-topology} shows the communication topology of a small scale AMG execution with 216 MPI Ranks.
+Figure~\ref{fig:amg-communication-topology} shows the communication matrix 
+of a small scale AMG execution with 216 MPI Ranks. 
 Note that the dominant communication pattern of the application doesn't change with scale. 
-We can make the observation that AMG's dominant communication pattern is 3D Nearest Neighbor, each rank has intensive communication with up to six neighbors, depending on the boundaries of the ranks.
-Applications with similar pattern include PARTISN \cite{partisn} and SNAP \cite{snap}.
-
-%Figure~\ref{fig:amg-data-trans} shows the amount of data each rank transferred in the example execution.
-The data transferring of each rank is between 7MB to 16MB, corresponding to whether the rank is on a boundary or not.
+We can make the observation that AMG's dominant communication pattern is 3D Nearest Neighbor, 
+each rank has intensive communication with up to six neighbors, 
+depending on the boundaries of the ranks. 
+Applications with similar pattern include PARTISN~\cite{partisn} and SNAP~\cite{snap}.
 
@@
 
-The second MiniApp covered is CrystalRouter, the extracted communication kernel of the full application Nek5000~\cite{crystalrouter}, which is a spectral element CFD application developed at Argonne National Laboratory \cite{nek5000}.
-It features spectral element multi-grid solvers coupled with a highly scalable, parallel coarse-grid solver that is widely used for projects including ocean current modeling, thermal hydraulics of reactor cores, and spatiotemporal chaos.
-CrystalRouter demonstrates the ``many-to-many'' communication pattern through a scalable multi-stage communication process.
-The way CrystalRouter works is nodes compute for a while, then synchronize and communicate, continually alternating between these two types of activities.
+The second MiniApp covered is CrystalRouter, 
+the extracted communication kernel of the full application Nek5000~\cite{nek5000}, 
+which is a spectral element CFD application developed at Argonne National Laboratory. 
+It features spectral element multi-grid solvers coupled with a highly scalable, 
+parallel coarse-grid solver that is widely used for projects including ocean current modeling, 
+thermal hydraulics of reactor cores, and spatiotemporal chaos. 
+CrystalRouter demonstrates the ``many-to-many'' communication pattern 
+through a scalable multi-stage communication process. 
+The way CrystalRouter works is the nodes compute for a while, 
+then synchronize and communicate, continually alternating between these two types of activities.
 
 The collective communication in CrystalRouter utilizes a recursive doubling approach. 
-Ranks in CrystalRouter conform to a $n$-dimensional hypercube and recursively split into ($n$-1)-dimensional hypercubes, with communication occurring along the splitting plane.
+Ranks in CrystalRouter conform to a $n$-dimensional hypercube 
+and recursively split into ($n$-1)-dimensional hypercubes, 
+with communication occurring along the splitting plane. 
 The pattern of this communication can be found in Figure~\ref{fig:cr-communication-topology}. 
-By nature of the logarithmic splitting process, a substantial portion of the communication occurs in small neighborhoods of ranks.
-CrystalRouter represents a group of applications whose dominant communication is a hybrid of multi-stage local and hierarchical global communication, and shares similarities with most MPI collective communication implementations.
-
-%Figure~\ref{fig:cr-data-trans} shows the amount of data each rank transferred in CrystalRouter, which ranges between 20MB to 36MB.
-This is due to the rank counts not matching one-for-one with points in the hypercube.
+By nature of the logarithmic splitting process, 
+a substantial portion of the communication occurs in small neighborhoods of ranks. 
+CrystalRouter represents a group of applications whose dominant communication 
+is a hybrid of multi-stage local and hierarchical global communication, 
+and shares similarities with most MPI collective communication implementations.
 
@@
 
-MultiGrid is geometric multi-grid v-cycle from the production elliptic solver BoxLib, a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes \cite{boxlib}.
-MultiGrid conforms to many-to-many communication pattern with decreasing message size and collectives for different parts of the multi-grid v-cycle.
+MultiGrid is geometric multi-grid v-cycle from the production elliptic solver BoxLib, 
+a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes~\cite{boxlib}. 
+MultiGrid conforms to many-to-many communication pattern with decreasing message size and 
+collectives for different parts of the multi-grid v-cycle. 
 It is widely used for structured grid physics packages.
 
-Figure~\ref{fig:mg-communication-topology} shows the communication topology of MultiGrid with 125 ranks.
+Figure~\ref{fig:mg-communication-topology} shows the communication matrix of MultiGrid with 125 ranks. 
 We can see intensive communication along the diagonal that resembles nearest neighbor communication, similar to AMG. 
-However, the communication topology leads to a greater ``spread'' of communication across the set of ranks, challenging the maximization of communication locality with respect to rank.
+However, the communication topology leads to a greater ``spread'' of communication across the set of ranks, 
+challenging the maximization of communication locality with respect to ranks. 
 In this sense, it can be considered a ``many-to-many'' pattern. 
-Applications with such similar dominant communication pattern are like FillBoundary, another PDE solver code in \cite{boxlib}.
-
-%Figure~\ref{fig:mg-data-trans} shows the amount of data transferred between ranks in MultiGrid. For this miniapp, The data transferred per rank is relatively small and stable, between 2.8MB to 2.9MB.
+Applications with such similar dominant communication pattern are like FillBoundary, another PDE solver code in~\cite{boxlib}.

### 2016-07-08 research_tool.tex
pass on Research Vehicle

@@
 It is difficult to accurately and flexibly experiment with concurrently running jobs in an HPC context. 
-One reason is that the allocation strategy used on production machine is part of the system software, which can not be changed by users.
+One reason is that the allocation strategy used on production machine is part of the system software, 
+which can not be changed by users. 
 Even the system administrator is not authorized to make that change. 
-Another reason is that it is unrealistic to reserve the system exclusively to run the same job with desired allocation without interference and then compare the results with those in the presence of interference.
+Another reason is that it is unrealistic to reserve the system exclusively 
+to run the same job with desired allocation without interference and 
+then compare the results with those in the presence of interference. 
 Therefore, we resort to simulation for this work.
 
-A simulation toolkit named CODES enables the exploration of simulating different HPC networks with high fidelity \cite{Jason-2011}\cite{mubarak-sc2012}.
-CODES is built on top of Rensselaer Optimistic Simulation System (ROSS) parallel discrete-event simulator, which is capable of processing billions of events per second on leadership-class supercomputers \cite{ross}.
+A simulation toolkit named CODES enables the exploration of simulating 
+different HPC networks at flit-level with high fidelity~\cite{Jason-2011, mubarak-sc2012}. 
+CODES is built on top of Rensselaer Optimistic Simulation System (ROSS) parallel discrete-event simulator~\cite{ross}, 
+which is capable of processing billions of events per second on leadership-class supercomputers. 
 CODES support both torus and dragonfly network with high fidelity flit-level simulation. 
 CODES has this network workload component that is capable of conducting trace-driven simulations. 
-It can take real MPI application traces generated by SST DUMPI \cite{sst} to drive CODES network models.
-In this work, we focus on an in-depth analysis of intra- and inter-job communication interferences with  different job allocations on torus-connected HPC systems.
-Torus networks have been extensively used in the current generation of supercomputers because of their linear scaling on per-node cost and competitive communication performance \cite{mubarak-sc2012}\cite{zhou-ipdps}.
+It can take MPI application traces generated by SST DUMPI \cite{sst} to drive CODES network models. 
+In this work, we focus on an in-depth analysis of intra- and inter-job 
+communication interferences with different job allocations on torus-connected HPC systems. 
+Torus networks have been extensively used in the current generation of supercomputers 
+because of their linear scaling on per-node cost and competitive communication performance.

### 2016-07-10 interference_study.tex
pass on Interference Study

@@
 
-\textbf{\emph{Intra-job interference}} refers to the network contention between ranks within each application.
-\textbf{\emph{Inter-job interference}} is introduced by concurrently running jobs that adjacent to each other sharing network resources.
-Communication variability due to such interference can cause application performance degradation.
+\textbf{\emph{Intra-job interference}} refers to the network 
+contention between ranks within each application. 
+\textbf{\emph{Inter-job interference}} is introduced by concurrently 
+running jobs that adjacent to each other sharing network resources. 
+Communication variability due to such interference can 
+cause application performance degradation.
 
-In this section we will study both kinds of interference in the context of torus network. 
-Since application's communication pattern won't change with its scale, we use AMG with 216 ranks, CrystalRouter with 100 ranks, and MultiGrid 125 ranks.
-And in order to accommodate three applications running concurrently with different allocations, we simulate a 2K-node 3D torus ($16^{2} \times 8$) network for the experiments.
+In this section we will study both kinds of interference 
+in the context of a torus network. 
+Since application's communication pattern won't change with its scale, 
+we use AMG with 216 ranks, CrystalRouter with 100 ranks, and MultiGrid 125 ranks. 
+And in order to accommodate three applications running concurrently with different allocations, 
+we simulate a 2K-node 3D torus ($16^{2} \times 8$) network for the experiments.
 
@@
 
-We design two sets of experiments to study the intra-job interference of each application.
-In the first, shown in Figure \ref{fig:shapestudy}, we assign each application with allocations in \emph{three different shapes}.
-In the second, shown in Figure \ref{fig:diffmap}, we study the intra-job interference by using different mapping strategies for each application with given allocation shape.
+We design two sets of experiments to study the 
+intra-job interference of each application. 
+In the first, shown in Figure \ref{fig:shapestudy}, 
+we assign each application with allocations in \emph{three different shapes}. 
+In the second, shown in Figure \ref{fig:diffmap}, 
+we study the intra-job interference by using different 
+mapping strategies for each application with given allocation shape.
 
-In the allocation shapes experiment, we select three possible shapes can can commonly seen on the 3D torus network.
-They are 3D balanced-cube, 3D unbalanced-cube, 2D mesh, as shown in Figure \ref{fig:alloc-shapes}.
+In the allocation shapes experiment, 
+we select three possible shapes commonly seen on the 3D torus network. 
+They are 3D balanced-cube, 3D unbalanced-cube and 2D mesh, 
+as shown in Figure \ref{fig:alloc-shapes}.
 
-3D balanced-cube, shown as red in Figure \ref{fig:cont_sub1}, can guarantee the minimum \emph{average pair-wise distance} within the allocation.
-Some research studies like \cite{leung,abhinav-sc13} indicate that compact allocation can guarantee jobs with better performance.
-They use a variety of metrics to evaluate the compactness of the allocation, such as \emph{average pair-wise distance}, \emph{diameter} and \emph{contiguity}.
+3D balanced-cube, shown as red in Figure \ref{fig:cont_sub1}, 
+can guarantee the minimum \emph{average pair-wise distance} within the allocation. 
+Some research studies~\cite{leung,abhinav-sc13} indicate that 
+compact allocation can guarantee jobs with better performance. 
+They use a variety of metrics to evaluate the compactness of the allocation, 
+such as \emph{average pair-wise distance}, \emph{diameter} and \emph{contiguity}. 
 In this work, we select 3D balanced-cube as the most compact allocation on a 3D torus network.
 
-3D unbalanced-cube, shown as green in Figure \ref{fig:cont_sub1}, is a rectangular prism, which is the possible allocation shape on the systems with asymmetric networks.
+3D unbalanced-cube, shown in green color in Figure \ref{fig:cont_sub1}, 
+is a rectangular prism, which is the possible allocation shape 
+on the systems with asymmetric networks. 
 For example, Cray XE6/XK7 systems are deployed 3D tori with Gemini routers. 
-The network connections in the \emph{y}-direction have only half the bandwidth of the cables used in the \emph{x} and \emph{z} directions.
-In order to take advantage of the faster links in the \emph{x} and \emph{z} directions, job allocation starts from X-Z plane, which leads to a rectangular prism shaped allocation \cite{RF}.
+The network connections in the \emph{y}-direction have 
+only half the bandwidth of the cables used in the \emph{x} and \emph{z} directions. 
+In order to take advantage of the faster links in the \emph{x} and \emph{z} directions, 
+job allocation starts from X-Z plane, 
+which leads to a rectangular prism shaped allocation~\cite{RF}.
 
-2D mesh, shown as blue in Figure \ref{fig:cont_sub1}, can be cut out from a single layer of the 3D torus.
-2D mesh is a very common allocation shape in torus network for both contiguous and non-contiguous allocation strategies.
-For example, Cray Application Level Placement Scheduler (ALPS) indexes the nodes in torus network into a list and makes job allocation by simply going through the list \cite{carl-cug}.
-When the list is obtained by sorting the nodes based on their spacial coordinates in the torus, allocation made off from this list will be 2D mesh.
-The IBM Blue/Gene Q supercomputer Mira at Argonne Leadership Computing Facility also allow its allocation partition configured into mesh \cite{zhou-ipdps}.
+2D mesh, shown as blue in Figure \ref{fig:cont_sub1}, 
+can be cut out from a single layer of the 3D torus. 
+2D mesh is a very common allocation shape on the torus network 
+for both contiguous and non-contiguous placement policies. 
+For example, Cray Application Level Placement Scheduler (ALPS) indexes 
+all compute nodes in a torus network into a list and 
+makes job allocation by simply going through that list~\cite{carl-cug}. 
+When the list is obtained by sorting the nodes based on 
+their spacial coordinates in the torus, 
+allocation made off from this list will be 2D mesh. 
+The IBM Blue/Gene Q supercomputer Mira at Argonne Leadership Computing Facility 
+also allow its allocation partition configured into mesh~\cite{zhou-ipdps}.
 
@@
     \end{subfigure}%
-    \begin{subfigure}[t]{0.2\textwidth}
+    \hspace{1em}%
+    \begin{subfigure}[t]{0.22\textwidth}
         \centering
@@
     \end{subfigure}%
-    \caption{(a) Contiguous allocation in three different shapes.
-``Blue" is 2D mesh, ``Green" is 3D-unbalanced cube and ``Red" is 3D-balanced cube.
-(b)-(d) Non-contiguous allocation.
-Each job is represented by a specific color.
-The nodes assigned to different jobs are interleaved, the size of allocation unit are 16, 8 and 2.}
+    \caption{
+    (a) Contiguous allocation in three different shapes. 
+    ``Blue" is a 2D mesh, ``Green" is a 3D-unbalanced cube and ``Red" is a 3D-balanced cube. 
+    (b)-(d) Non-contiguous allocation. Each job is represented by a specific color. 
+    The nodes assigned to different jobs are interleaved, 
+    the size of allocation unit are 16, 8 and 2.
+    }
     \label{fig:alloc-shapes}
@@
 
-AMG with 3D ``Nearest Neighbor" communication pattern takes less time when being assigned with 3D-unbalanced allocation than 3D-balanced, as shown in Figure \ref{fig:shapstudy-amg}.
-And MultiGrid gets the best performance (shortest data transfer time) when running on 3D-balanced allocation, as shown in Figure \ref{fig:shapstudy-mg}.
-Since MultiGrid's communication pattern is ``Many-to-Many" dominant, 3D-balanced is the most compact allocation with shortest pair-wise distance between nodes, which can reduce the aggregated hops for transferring message among ranks in MultiGrid.
-As CrystalRouter has both dominant local communication and multi-stage global data transfer, the 3D-balance is also the best allocation, but its advantage over 3D-unbalanced is not that obvious as to MultiGrid, as shown in Figure \ref{fig:shapstudy-cr}.
+AMG with 3D ``Nearest Neighbor" communication pattern takes less time 
+when being assigned with 3D-unbalanced allocation than 3D-balanced, 
+as shown in Figure~\ref{fig:shapstudy-amg}. 
+And MultiGrid gets the best performance (shortest data transfer time) 
+when running on 3D-balanced allocation, as shown in Figure~\ref{fig:shapstudy-mg}. 
+Since MultiGrid's communication pattern is ``Many-to-Many" dominant, 
+3D-balanced is the most compact allocation with shortest pair-wise distance between nodes, 
+which can reduce the aggregated hops for transferring message among ranks in MultiGrid. 
+Since CrystalRouter has both dominant local communication and multi-stage global data transfer, 
+the 3D-balance is also the best allocation, but its advantage over 3D-unbalanced is not 
+that obvious as to MultiGrid, as shown in Figure~\ref{fig:shapstudy-cr}.
 
-Lots of research studies have designed complex allocation algorithms to provide application with the most compact allocation\cite{leung,LO}.
-However, as shown in our experiments, providing cubical allocation without considering application's communication pattern can not guarantee the best performance for every application.
-Compact allocation should be provided to application with global data transfer, such as those conforms to ``Many-to-Many" communication pattern.
-Application with dominant communication pattern like ``Nearest Neighbor" won't fully utilize the compact allocation and could get even better performance on a relaxed allocation shape.
+Lots of research studies have designed complex placement algorithms 
+to provide application with the most compact allocation~\cite{leung,LO}. 
+However, as shown in our experiments, providing cubic allocation without 
+considering application's communication pattern can not guarantee 
+the best performance for every application. 
+Compact allocation should be provided to applications with intensive global data transfer, 
+such as those conforms to ``Many-to-Many" communication pattern. 
+Applications with dominant communication pattern like ``Nearest Neighbor" 
+can not fully utilize the compact allocation and 
+may get even better performance on a relaxed allocation shape.
 
@@
     \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 3D-balanced, 3D-unbalanced and 2D allocation.}
+   \caption{
+   Data transfer time of AMG, CrystalRouter and MultiGrid on 3D-balanced, 
+   3D-unbalanced and 2D allocation.
+   }
    \label{fig:shapestudy}
@@
 
-The rank-to-node mapping of parallel applications on HPC system could greatly impact applications' performance.
-However, finding the optimal mapping solution for a given application is out the scope of this work.
-Our experiments aims to show how mapping strategies impact the intra-job interference of application with specific communication pattern.
+The rank-to-node mapping of parallel applications on HPC systems 
+could greatly impact applications' performance. 
+However, finding the optimal mapping solution for a given 
+application is out the scope of this work. 
+Our experiments aims to show how mapping strategies impact the 
+intra-job interference of application with specific communication pattern.
 
-We provide AMG, CrystalRouter and MultiGrid with 3D-balance allocation and use three mapping strategies to do the rank-to-node mapping.
-The ``Linear" mapping, as we used in the previous experiments, is to map each rank according to the dimensional ordering of compute nodes.
+We provide AMG, CrystalRouter and MultiGrid with 3D-balance allocation and 
+use three mapping strategies to do the rank-to-node mapping. 
+The ``Linear" mapping, as we used in the previous experiments, 
+is to map each rank according to the dimensional ordering of compute nodes. 
 The ``Cube" mapping assigns ranks into consecutive $2^{3}$ cubes. 
@@
 
-AMG remains roughly the same when it been mapped by ``Linear" and ``Cube", shown in Figure \ref{fig:diffmap-amg}.
-Although ``Linear" and ``Cube" mapping strategies cause some routing overlap for AMG's communication, both mapping strategies still preserve the locality of AMG's 3D ``Nearest Neighbor" communication pattern.
-The ``Random" mapping totally destroys AMG's communication pattern, and results in intra-job interference among all the ranks in AMG.
+AMG remains roughly the same when it been mapped by ``Linear" and ``Cube", 
+shown in Figure~\ref{fig:diffmap-amg}. 
+Although ``Linear" and ``Cube" mapping strategies cause some routing overlap for AMG's communication, 
+both mapping strategies still preserve the locality of AMG's ``3D Nearest Neighbor" communication pattern. 
+The ``Random" mapping totally destroys AMG's communication pattern, 
+and results in intra-job interference among all the ranks in AMG. 
 The performance degradation caused by using ``Random" mapping is up to 90\%.
 
-The ``Cube" mapping improves CrystalRouter's performance over the ``Linear" by reducing the pair-wise distance between ranks, shown in Figure \ref{fig:diffmap-cr}.
+The ``Cube" mapping improves CrystalRouter's performance over the ``Linear" 
+by reducing the pair-wise distance between ranks, shown in Figure~\ref{fig:diffmap-cr}. 
 The global data transfer in CrystalRouter will takes fewer hops with ``Cube" mapping. 
-The ``Random" mapping scatters the ranks within the local neighborhood of CrystalRouter and make their communication less efficient.
+The ``Random" mapping scatters the ranks within the local neighborhood 
+of CrystalRouter and make their communication less efficient.
 
 The ``Cube" mapping benefits MultiGrid's ``Many-to-Many" communication. 
-However, due to the small amount of data transfer among ranks in MultiGrid, ``Cube" mapping fails to exhibit a drastic advantage.
-``Random" mapping neither cause too much degradation for the same reason, shown in Figure \ref{fig:diffmap-mg}.
+However, due to the small amount of data transfer among ranks in MultiGrid, 
+``Cube" mapping fails to exhibit a drastic advantage. 
+``Random" mapping neither cause too much degradation for the same reason, 
+shown in Figure~\ref{fig:diffmap-mg}.
 
@@
     \end{subfigure}%
-   \caption{Data transfer time of AMG, CrystalRouter and MultiGrid on 3D-balanced allocation using  different mapping strategies.}
+   \caption{
+   Data transfer time of AMG, CrystalRouter and MultiGrid 
+   on 3D-balanced allocation using  different mapping strategies.
+   }
    \label{fig:diffmap}
@@
 
-Inter-job interference has been identified as one of the major culprits for application's performance variability \cite{abhinav-sc13,skinner,rosenthal}.
-Inter-job interference is a more prominent issue for system adopted non-contiguous allocation policy than system with contiguous allocation.
-Application communication times have been demonstrated to vary from 36\% faster to 69\% slower due to job interference when running on non-contiguous allocation \cite{abhinav-sc13}.
+Inter-job interference has been identified as one of the major culprits 
+for application's performance variability~\cite{abhinav-sc13,skinner,rosenthal}. 
+Inter-job interference is a more prominent issue for system adopted 
+non-contiguous placement policy than systems with contiguous policy. 
+Applications communication time have been demonstrated to 
+vary from 36\% faster to 69\% slower due to job interference 
+when they are running concurrently with non-contiguous allocations~\cite{abhinav-sc13}.
 
-We assign each application with non-contiguous allocation and run them concurrently on the same network.
+We assign each application with non-contiguous allocation and 
+run them concurrently on the same network. 
 The allocation unit belongs to different jobs are interleaved. 
-In order to study the impact of different allocation unit size on applications' inter-job interference, we conduct experiments with unit size of 16, 8 and 2, shown respectively in Figure \ref{fig:noncont_sub1}, \ref{fig:noncont_sub1} and \ref{fig:noncont_sub3}.
-Figure \ref{fig:interjobstudy} shows the results of each application data transfer time with different allocation unit size.
+In order to study the impact of different allocation unit size on 
+applications' inter-job interference, 
+we conduct experiments with unit size of 16, 8 and 2, 
+shown respectively in Figure~\ref{fig:noncont_sub1}, 
+~\ref{fig:noncont_sub1} and~\ref{fig:noncont_sub3}. 
+Figure~\ref{fig:interjobstudy} shows the results of each application 
+data transfer time with different allocation unit size.
 
-The data transfer time of AMG in Figure \ref{fig:interjob-amg} keeps stable between allocation unit size of 16 and 8, since the locality of AMG is 6-rank based.
+The data transfer time of AMG in Figure~\ref{fig:interjob-amg} keeps stable 
+between allocation unit size of 16 and 8, since the locality of AMG is 6-rank based. 
 When the unit size reduce to 2, AMG suffers prolong data transfer time by about 10\%. 
@@
 The best unit size would be the one big enough to accommodate its local neighborhood. 
-Figure \ref{fig:interjob-cr} shows unit size of 16 and 8 can guarantee the same average data transfer time, while some rank spend more time with allocation unit size 8 than 16.
-When the unit size reduce to 2, the communication become less efficiency and takes about 15\% more times for transferring data.
+Figure~\ref{fig:interjob-cr} shows that unit size of 16 and 8 can guarantee 
+the same average data transfer time, 
+while some rank spend more time with allocation unit size 8 than 16. 
+When the unit size reduce to 2, the communication become less efficiency 
+and takes about 15\% more time for transferring data.
 
@@
 
-The unit size in non-contiguous allocation policy should be chosen according to application's communication pattern.
-Inter-job interference is inevitable in non-contiguous based systems, but unit size that big enough to preserve the neighborhood communication of the application will alleviate such interference and improve job performance.
+Choosing the proper unit size in non-contiguous placement policy should 
+also consider application's communication pattern. 
+Inter-job interference is inevitable in non-contiguous based systems, 
+but unit size that big enough to preserve the neighborhood communication of the application 
+will alleviate such interference and improve job performance.
 
@@
     \end{subfigure}%
-   \caption{Inter-job interference study.
+   \caption{
+   Inter-job interference study. 
    ``cont" indicates three applications running side by side concurrently on the same network with contiguous allocation. 
-To study the impact of non-contiguous allocation on inter-job interference, applications running concurrently with interleaved allocation of different unit sizes, which are 16-node, 8-node, 2-node.
+   To study the impact of non-contiguous allocation on inter-job interference, 
+   applications running concurrently with interleaved allocation of different unit sizes,
+    which are 16-node, 8-node, 2-node. 
     }
@@
 Based on the above comprehensive study, we can make following observations:
+
 \begin{itemize}
+
     \item The compact allocation may not guarantee the best performance for every application.
-    \item The applications dominated with Nearest Neighbor communications exhibit relatively stable performance under different allocation shapes, while the applications dominated with Many-to-Many communication exhibit better performance with compact allocation (e.g.
-3D balanced).
-    \item A good rank-to-node mapping strategy can greatly improve application's performance when specific allocation is given.
-    \item An optimal size for allocation units should be determined according to an application's dominant communication pattern.
-In general, a unit size should be large enough to preserve neighboring communication of the application.
+
+    \item The applications dominated with ``Nearest Neighbor" communications exhibit 
+    relatively stable performance under different allocation shapes, 
+    while the applications dominated with ``Many-to-Many" communication exhibit 
+    better performance with compact allocation (e.g, 3D balanced).
+
+    \item A good rank-to-node mapping strategy can greatly improve 
+    application's performance when specific allocation is given.
+
+    \item An optimal size for allocation units should be determined 
+    according to an application's dominant communication pattern. 
+    In general, a unit size should be large enough to accommodate  
+    neighboring communication of the application.
+
     \item Inter-job interference is inevitable in non-contiguous allocation. 
-However, choosing the proper allocation unit size with job's communication pattern awareness can alleviate such negative effect.
+    However, choosing the proper allocation unit size with job's communication 
+    pattern awareness can alleviate such negative effect.
+
 \end{itemize}

### 2016-07-10 discussion.tex
pass on discussion

@@
 
-The results shown in the previous section provide insights for the design of smart and flexible job allocation.
-By scrutinizing job's communication behavior, we can identify job's dominant communication pattern and pinpoint the ``neighborhood communication".
-With such knowledge about jobs' communication patterns, we can analyze the possible interference between jobs and take precautions to alleviate the negative effect when making allocation decisions.
+The results shown in the previous section provide insights 
+for the design of smart and flexible job placement policy. 
+By scrutinizing job's communication behavior, 
+we can identify job's dominant communication pattern 
+and pinpoint the ``neighborhood communication". 
+With such knowledge about jobs' communication patterns, 
+we can analyze the possible interference between jobs and 
+take precautions to alleviate the negative effect when making placement decisions.
 
-When an application with dominant communication that is intensive ``Many-to-Many", the scheduler should grant it with compact node allocation and exclusive network provision.
-The compact allocation can guarantee the shortest pair-wise distance between all the ranks, and make the data transfer between ranks take less hops.
-On the other hand, the exclusive network provision will prevent other adjacent applications from sharing network resources, thus eliminate the performance degradation due to interference.
+When an application with dominant communication that is intensive ``Many-to-Many", 
+the scheduler should grant it with compact node allocation and exclusive network provision. 
+The compact allocation can guarantee the shortest pair-wise distance between all the ranks, 
+and make the data transfer between ranks take less hops. 
+On the other hand, 
+the exclusive network provision will prevent other adjacent applications from sharing network resources, 
+thus eliminate the performance degradation due to interference.
 
-Not every application's preferable resource are compact node allocation and exclusive network provision.
-As we found in our analysis, applications whose dominant communication patterns contain intensive ``neighborhood communication" like ``Nearest Neighbor" don't benefit from compact node allocation and exclusive network provision.
-Applications, such as AMG, can run with non-contiguous node allocation without performance degradation, as long as the allocation unit can accommodate their ``neighborhood communication".
+Not every application's preferable resource are compact node allocation 
+and exclusive network provision. 
+As we demonstrate in our analysis, 
+applications whose dominant communication pattern contains intensive 
+``neighborhood communication" like ``Nearest Neighbor" do not benefit 
+from compact node allocation and exclusive network provision. 
+Applications, such as AMG, can run with non-contiguous node 
+allocation without performance degradation, 
+as long as the allocation unit can accommodate their ``neighborhood communication".
 
 The allocation unit size shouldn't be fixed. 
-Instead, the scheduler should choose a proper unit size based on the scale of ``neighborhood communication" of each job.
-The best unit size should be neither too big nor too small, jut perfectly fit for the size of that ``neighborhood".
-Big unit size won't be fully utilized and also cause fragmentation.
-Small unit size won't be able to accommodate the ``neighborhood communication" and makes the intra-job communication less efficient.
+Instead, the scheduler should choose a proper unit size 
+based on the scale of ``neighborhood communication" of each job. 
+The best unit size should be neither too big nor too small, 
+just perfectly fit for the size of that ``neighborhood". 
+Big unit size can not be fully utilized and thus cause fragmentation. 
+Small unit size will not be able to accommodate the ``neighborhood communication", 
+making the intra-job communication less efficient.
 
-The advantage of making allocation with consideration about job's communication pattern is obvious.
-First, compared with contiguous allocation policy, the new scheduler outlined above is more flexible, without requiring the system to provide a big contiguous partition to accommodate the whole application, it only needs to provide a small set of compact nodes that is sufficient for all the "local community" in the application.
-On the other hand, the new scheduler outlined above won't destroy the locality of application's communication pattern.
-Indeed, the small compact node set (allocation unit) provided for the application's  ``local communities" can preserve the communication locality in the maximum extent.
-The design of such a smart scheduler is part of our future work.
+The advantage of making job placement with consideration about job's communication pattern is obvious. 
+First, compared with contiguous allocation policy, 
+the new scheduler with job's communication pattern awareness can be more flexible, 
+without requiring the system to provide a big contiguous partition to accommodate the whole application, 
+it only needs to provide a small set of compact nodes that is 
+sufficient for all the "local community" in the application. 
+On the other hand, 
+the new scheduler is able to preserve the locality of application's communication pattern. 
+Indeed, the small compact node set (allocation unit) provided for the application's 
+``local communities" can preserve the communication locality in the maximum extent. 
+The design of such a new scheduler with job communication pattern awareness 
+is part of our future work.

### 2016-07-10 related_work.tex
pass on discussion

@@
 
-There are many tools available for system monitoring and application profiling. Tools like TAU (Tuning and Analysis)~\cite{tau}, 
-and mpiP~\cite{mpip} can capture application's runtime information about application's communication and computation in event traces.
+There are many tools available for system monitoring and application profiling. 
+Tools like TAU (Tuning and Analysis)~\cite{tau} and mpiP~\cite{mpip} can capture 
+application's runtime information and keep records in event traces. 
 However, recognizing the patterns about application's communication from those traces requires substantial effort.

### 2016-07-10 related_work.tex
pass on realted work

@@
 There are a number of studies on the recognition and characterization of parallel application communication patterns. 
-Oak Ridge National Laboratory has its ongoing project about developing a tool set named Oxbow, which can characterize the computation and communication behavior of scientific applications and benchmarks~\cite{oxbow}.
-In a recent work~\cite{roth}, the authors demonstrate a new approach to automatically characterizing parallel applications' communication behaviors.
+Oak Ridge National Laboratory has its ongoing project about developing a tool set named Oxbow, 
+which can characterize the computation and communication behavior of scientific applications and benchmarks~\cite{oxbow}. 
+In a recent work~\cite{roth}, 
+the authors demonstrate a new approach to automatically characterizing parallel applications communication behaviors.
 
@@
 Many research efforts have been conducted to characterize scientific applications. 
-For instance, a DOE project aims at the identification of the computational characteristics of the DOE MiniApps developed at various exascale co-design centers \cite{designforwardwebpage}.
-In this project, the communication patterns of several DOE full applications and associated mini-applications are studied to provide a more complete snapshot of the DOE workload.
-A joint project named CORAL from Oak Ridge, Argonne and Livermore provide a series of benchmarks to represent DOE workloads and technical requirements \cite{coral}.
-The CORAL project includes scalable science benchmarks, throughput benchmarks, data centric benchmarks, skeleton benchmarks and Micro benchmarks.
+For instance, a DOE project aims at the identification of the computational characteristics of 
+the DOE MiniApps developed at various exascale co-design centers~\cite{designforwardwebpage}. 
+In this project, the communication patterns of several DOE full applications and associated 
+mini-applications are studied to provide a more complete snapshot of the DOE workload. 
+A joint project named CORAL from Oak Ridge, Argonne and Livermore provides a series of 
+benchmarks to represent DOE workloads and technical requirements~\cite{coral}. 
+The CORAL project includes scalable science benchmarks, throughput benchmarks, 
+data centric benchmarks, skeleton benchmarks and Micro benchmarks.
 
 %\textbf{P3: Interference between concurrently running jobs}\\
-The interference between concurrently running jobs on HPC system have been identified as major culprit for job's performance variability.
-Abhinav et al.
-found that concurrently running applications on HPC can cause interference to each other, and cause communication time varied from 36\% faster to 69\% slower.
-Such interference could come from Operating System noise, shape of the allocated partition , or other running jobs that sharing network resources \cite{abhinav-sc13}.
-Skinner et al.
-found that  there is a 2-3$\times$ slowdown in MPI$\textunderscore$Allreduce due to network contention from other jobs \cite{skinner}.
-Rosenthal et al.
-found increasing network bandwidth provide limited benefits to a number of applications.
-The applications that send mostly small messages or large messages asynchronously are not bandwidth bounded, hence, benefit only slightly or not at all from increased bandwidth \cite{rosenthal}.
+The interference among concurrently running jobs on HPC systems 
+have been identified as major culprit for job's performance variability. 
+Bhatel et al. found that concurrently running applications interfere with each other, 
+and cause their communication time varied from 
+36\% faster to 69\% slower on different HPC systems~\cite{abhinav-sc13}. 
+Skinner et al. found that  there is a 2-3$\times$ slowdown in MPI$\textunderscore$Allreduce 
+due to network contention from other concurrently running jobs~\cite{skinner}. 
+Rosenthal et al. found increasing network bandwidth provide limited benefits to a number of applications. 
+The applications that send mostly small messages or large messages asynchronously 
+are not bandwidth bounded, hence, benefit only slightly or not at all from increased bandwidth~\cite{rosenthal}.
 
 %\textbf{P4: Job allocation(little bit)}\\
-Several research studies focus on optimizing job allocation on HPC system to alleviate the interference between concurrently running jobs.
-Hoefler et al.
-proposed to use performance modeling techniques to analyze factors that impact the performance of parallel scientific applications \cite{hoefler-modeling}.
-However, as the scale of HPC continue to grow, the interference of concurrently running jobs is getting worse, which is hard to be quantified by performance profiling tools alone.
-Bogdan et al.
-provide a set of guidelines of how to configure a network with Dragonfly topology for workload with Nearest Neighbor communication pattern \cite{Bogdan-hpdc14}.
-Dong et al.
-have developed simple benchmarks that conforms to four different communication patterns, namely ping-pong, nearest neighbor, broadcast and all$\textunderscore$reduce, to demonstrate the effectiveness of this highly parallel 5D torus network \cite{Dong-SC11}.
+Several research studies focus on optimizing job allocation on HPC systems 
+to alleviate the interference between concurrently running jobs. 
+Hoefler et al. proposed to use performance modeling techniques to analyze factors that 
+impact the performance of parallel scientific applications~\cite{hoefler-modeling}. 
+However, as the scale of HPC systems continue to grow, 
+the interference of concurrently running jobs is getting worse, 
+which is hard to be quantified by performance profiling tools alone. 
+Bogdan et al. provide a set of guidelines of how to configure a network with Dragonfly topology 
+for workload with Nearest Neighbor communication pattern~\cite{Bogdan-hpdc14}. 
+Dong et al. have developed simple benchmarks that conforms to four different communication patterns, 
+namely ping-pong, nearest neighbor, broadcast and all$\textunderscore$reduce, 
+to demonstrate the effectiveness of this highly parallel 5D torus network~\cite{Dong-SC11}.
 
@@
 We believe this can provide a guideline for other research work. 
-Secondly, we explore the intra- and inter-job interference between concurrently running jobs, while similar work such as \cite{abhinav-sc13} only focuses on the single application's performance degradation due to network contention.
+Secondly, we explore the intra- and inter-job interference between concurrently running jobs, 
+while similar work such as~\cite{abhinav-sc13} only focuses 
+on the single application's performance degradation due to network contention. 
 Finally, we analyze the impact of different allocation strategies to job's communication behaviors. 
-We further identify the optimal allocation strategy for each application with a specific dominant communication pattern.
-Based on our study, we claim that further scheduling should take job's communication pattern into consideration for its allocation decision making.
+We further identify the optimal allocation strategy for 
+each application with a specific dominant communication pattern. 
+Based on our study, we claim that further batch scheduler should take 
+job's communication pattern into consideration for its placement decision making.

### 2016-07-11 app_study.tex
some minor tweaks

@@
 At its different execution phases, 
-the application's communication behavior may follow certain basic patterns respectively. 
+the application's communication behavior may follow certain basic pattern respectively. 
 There are many profiling tools available to capture 
@@
 Geometric MultiGrid (MultiGrid) and CrystalRouter MiniApps. 
-The communication matrix presented in Figure~\ref{fig:apps_communication_matrix}
-are generated with the IPM~\cite{ipm} data gathered from publicly available traces~\cite{designforwardwebpage}.
+The communication matrix of each application is presented in Figure~\ref{fig:apps_communication_matrix}. 
+%are generated with the IPM~\cite{ipm} data gathered from publicly available traces~\cite{designforwardwebpage}.
 
@@
 Figure~\ref{fig:amg-communication-topology} shows the communication matrix 
-of a small scale AMG execution with 216 MPI Ranks. 
+of a small scale AMG execution with 216 MPI ranks. 
 Note that the dominant communication pattern of the application doesn't change with scale. 
-We can make the observation that AMG's dominant communication pattern is 3D Nearest Neighbor, 
+We can make the observation that AMG's dominant communication pattern is 3D nearest neighbor, 
 each rank has intensive communication with up to six neighbors, 
@@
 thermal hydraulics of reactor cores, and spatiotemporal chaos. 
-CrystalRouter demonstrates the ``many-to-many'' communication pattern 
+CrystalRouter demonstrates the many-to-many communication pattern 
 through a scalable multi-stage communication process. 
@@
 challenging the maximization of communication locality with respect to ranks. 
-In this sense, it can be considered a ``many-to-many'' pattern. 
+In this sense, it can be considered as many-to-many pattern. 
 Applications with such similar dominant communication pattern are like FillBoundary, another PDE solver code in~\cite{boxlib}.

### 2016-07-11 conclusion.tex
some minor tweaks

@@
 \label{sec:conclusion}
-In this work, we have studied the communication behavior of three parallel applications, namely AMG, CrystalRouter and MultiGrid.
-Each application has distinctive communication pattern, that can be representative for a group of jobs in HPC workload.
-We have used a sophisticate simulation tool named CODES from Argonne National Laboratory to simulate the running of these three parallel applications on torus network.
+In this work, we have studied the communication behavior of three parallel applications, 
+namely AMG, CrystalRouter and MultiGrid. 
+Each application has distinctive communication pattern, 
+that can be representative for a group of jobs in HPC workload. 
+We have used a sophisticated simulation tool named CODES from Argonne National Laboratory 
+to simulate the running of these three parallel applications on a torus network. 
 The torus network provided by CODES has good fidelity and scalability. 
-We have analyzed the performance of each application's communication in terms of data transfer time by simulating them running on torus networks with different bandwidth and dimensionality configurations.
+%We have analyzed the performance of each application's communication in terms of data transfer time 
+%by simulating them running on torus networks with different bandwidth and dimensionality configurations.
 We have found that higher dimensionality of torus network would improve the performance of application with ``Many-to-Many" communication patterns, while application with intensive local communication like ``Nearest Neighbor" won't benefit much from higher dimensionality.
 
-We have analyzed the intra- and inter-job interference by simulating three applications running on 3D torus network.
+We have analyzed the intra- and inter-job interference by simulating three applications running on the torus network. 
 Based on our comprehensive experiments, we got five observations. 
 1) The compact allocation may not guarantee the best performance for every application. 
-2)The applications dominated with Nearest Neighbor communications exhibit relatively stable performance under different allocation shapes, while the applications dominated with Many-to-Many communication exhibit better performance with compact allocation (e.g.
-3D balanced).
+2) The applications dominated with nearest neighbor communications exhibit 
+relatively stable performance under different allocation shapes, 
+while the applications dominated with many-to-many communication exhibit 
+better performance with compact allocation (e.g. 3D balanced). 
 3) A good rank-to-node mapping strategy can greatly improve application's performance when specific allocation is given. 
-4) An optimal size for allocation units should be determined according to an application's dominant communication pattern.
+4) An optimal size for allocation units should be determined according to 
+an application's dominant communication pattern. 
 In general, a unit size should be large enough to preserve neighboring communication of the application. 
@@
 
-We believe that our finding in this work can provide valuable guidance for HPC resource management to make flexible job allocations.
-Rather than using pre-defined partitions and non-contiguous allocation, future HPC systems should assign each job with preferable resources based on job's communication pattern.
+We believe that the findings in this work can provide valuable guidance 
+for HPC batch scheduler and resource manager to make flexible job allocations. 
+Rather than using pre-defined partitions and non-contiguous placement policy, 
+future HPC systems should assign each job with preferable resources based on job's communication pattern.

### 2016-07-11 discussion.tex
some minor tweaks

@@
 
-When an application with dominant communication that is intensive ``Many-to-Many", 
-the scheduler should grant it with compact node allocation and exclusive network provision. 
+When an application with dominant communication that is intensive many-to-many, 
+the scheduler should provide it with compact node allocation and exclusive network provision. 
 The compact allocation can guarantee the shortest pair-wise distance between all the ranks, 
@@
 
-Not every application's preferable resource are compact node allocation 
+Not every application's preferable resource are compact allocation 
 and exclusive network provision. 
-As we demonstrate in our analysis, 
+As we demonstrate in our study, 
 applications whose dominant communication pattern contains intensive 
-``neighborhood communication" like ``Nearest Neighbor" do not benefit 
-from compact node allocation and exclusive network provision. 
+``neighborhood communication" like nearest Neighbor do not benefit 
+from compact allocation and exclusive network provision. 
 Applications, such as AMG, can run with non-contiguous node 
@@
 
-The allocation unit size shouldn't be fixed. 
+The size of allocation unit shouldn't be fixed. 
 Instead, the scheduler should choose a proper unit size 
@@
 The best unit size should be neither too big nor too small, 
-just perfectly fit for the size of that ``neighborhood". 
+just perfectly fit for the scale of that ``neighborhood". 
 Big unit size can not be fully utilized and thus cause fragmentation. 

### 2016-07-11 interference_study.tex
some minor tweaks

@@
 They are 3D balanced-cube, 3D unbalanced-cube and 2D mesh, 
-as shown in Figure \ref{fig:alloc-shapes}.
-
+as shown in Figure \ref{fig:cont_sub1}.
 3D balanced-cube, shown as red in Figure \ref{fig:cont_sub1}, 
@@
 3D unbalanced-cube, shown in green color in Figure \ref{fig:cont_sub1}, 
-is a rectangular prism, which is the possible allocation shape 
+is a rectangular prism, which is a possible allocation shape 
 on the systems with asymmetric networks. 
@@
 
-AMG with 3D ``Nearest Neighbor" communication pattern takes less time 
+AMG with 3D nearest neighbor communication pattern takes less time 
 when being assigned with 3D-unbalanced allocation than 3D-balanced, 
@@
 when running on 3D-balanced allocation, as shown in Figure~\ref{fig:shapstudy-mg}. 
-Since MultiGrid's communication pattern is ``Many-to-Many" dominant, 
+Since MultiGrid's communication pattern is many-to-many dominant, 
 3D-balanced is the most compact allocation with shortest pair-wise distance between nodes, 
@@
 Lots of research studies have designed complex placement algorithms 
-to provide application with the most compact allocation~\cite{leung,LO}. 
+to provide applications with the most compact allocation~\cite{leung,LO}. 
 However, as shown in our experiments, providing cubic allocation without 
@@
 Compact allocation should be provided to applications with intensive global data transfer, 
-such as those conforms to ``Many-to-Many" communication pattern. 
-Applications with dominant communication pattern like ``Nearest Neighbor" 
+such as those conforms to many-to-many communication pattern. 
+Applications with dominant communication pattern like nearest neighbor 
 can not fully utilize the compact allocation and 
@@
 Although ``Linear" and ``Cube" mapping strategies cause some routing overlap for AMG's communication, 
-both mapping strategies still preserve the locality of AMG's ``3D Nearest Neighbor" communication pattern. 
+both mapping strategies still preserve the locality of AMG's 3D nearest neighbor communication pattern. 
 The ``Random" mapping totally destroys AMG's communication pattern, 
@@
 of CrystalRouter and make their communication less efficient. 
-
-The ``Cube" mapping benefits MultiGrid's ``Many-to-Many" communication. 
+The ``Cube" mapping benefits MultiGrid's many-to-many communication. 
 However, due to the small amount of data transfer among ranks in MultiGrid, 
@@
 Inter-job interference has been identified as one of the major culprits 
-for application's performance variability~\cite{abhinav-sc13,skinner,rosenthal}. 
+for application's performance variability~\cite{abhinav-sc13,skinner}. 
 Inter-job interference is a more prominent issue for system adopted 
@@
 
-The data transfer time of MultiGrid with different allocation unit sizes doesn't show obvious variability in Figure \ref{fig:interjob-mg}.
-This is because even big allocation unit size like 16 will still fail to preserve MultiGrid's ``Many-to-Many" pattern.
-The data transfer time is almost doubled when MultiGrid running concurrently with allocation unit size of 16, as shown in Figure \ref{fig:interjob-mg} .
-As the unit size decreases, the data transfer time keep growing, but in a slow way, which means in terms of preserving the locality of MultiGrid, allocation unit of size 16, 8 and 2 serve equally bad.
+The data transfer time of MultiGrid with different allocation unit sizes 
+doesn't show obvious variability in Figure~\ref{fig:interjob-mg}. 
+This is because even big allocation unit size like 16 will still fail to preserve MultiGrid's many-to-many pattern. 
+The data transfer time is almost doubled when MultiGrid running concurrently with allocation unit size of 16, 
+as shown in Figure~\ref{fig:interjob-mg}. 
+As the unit size decreases, the data transfer time keep growing, 
+but in a slow way, which means in terms of preserving the locality of MultiGrid, 
+allocation unit of size 16, 8 and 2 serve equally bad.
 
@@
 
-    \item The applications dominated with ``Nearest Neighbor" communications exhibit 
+    \item The applications dominated with nearest neighbor communications exhibit 
     relatively stable performance under different allocation shapes, 
-    while the applications dominated with ``Many-to-Many" communication exhibit 
+    while the applications dominated with many-to-many communication exhibit 
     better performance with compact allocation (e.g, 3D balanced).

### 2016-07-11 intro.tex
some minor tweaks

@@
    Each job is represented by a specific color. 
-   a) shows the effect of contiguous allocation, 
+   a) shows the effect of contiguous job placement, 
    which could possibly reduce the inter-job interference. 
-   b) shows non-contiguous allocation, 
+   b) shows non-contiguous job placement, 
    which may introduce both intra and inter-job interference. 
@@
 A prominent problem with this network sharing is the contention over
-the network resource among those concurrently running jobs. 
+the network among those concurrently running jobs. 
 The network contention can cause communication variability and 
@@
 Figure~\ref{fig:overview_sub2} shows the effect of non-contiguous placement. 
-While eliminating internal and external fragmentation as seen in contiguous allocation systems, 
-in return non-contiguous allocations introduce contention between jobs due to the interleaving of job nodes. 
+While eliminating internal and external fragmentation as seen in contiguous placement systems, 
+in return non-contiguous placement policy introduces contention between jobs due to the interleaving of job nodes. 
 The non-contiguous job placement can significantly reduce job performance, 
@@
 a flexible job placement mechanism which combines the 
-best of both contiguous and non-contiguous placement. 
+best of both contiguous and non-contiguous policy. 
 Such a flexible job placement mechanism should take 
@@
 With the knowledge and analysis of job communication patterns, 
-it can be identified which jobs require exclusive network and allocation locality, and to what degree. 
+it can be identified which jobs require exclusive network and compact allocation, and to what degree. 
 Then, rather than allocating each job in a ``know-nothing'' manner, 
-one may specialize allocation so that, for example, 
+one may specialize job placement policy so that, for example, 
 only the jobs with stringent network needs are given compact, isolated allocations, 
@@
 In this work, we focus on an in-depth analysis of intra- and inter-job 
-communication interference with different job placement on torus-connected HPC systems. 
+communication interference with different job placements on torus-connected HPC systems. 
 Torus-based networks are used on six of the top 10 supercomputers 

### 2016-07-11 related_work.tex
some minor tweaks

@@
 have been identified as major culprit for job's performance variability. 
-Bhatel et al. found that concurrently running applications interfere with each other, 
+Bhatele et al. found that concurrently running applications interfere with each other, 
 and cause their communication time varied from 
 36\% faster to 69\% slower on different HPC systems~\cite{abhinav-sc13}. 
-Skinner et al. found that  there is a 2-3$\times$ slowdown in MPI$\textunderscore$Allreduce 
+Skinner et al. found that  there is a 2-3 times of slowdown in MPI$\textunderscore$Allreduce 
 due to network contention from other concurrently running jobs~\cite{skinner}. 
-Rosenthal et al. found increasing network bandwidth provide limited benefits to a number of applications. 
-The applications that send mostly small messages or large messages asynchronously 
-are not bandwidth bounded, hence, benefit only slightly or not at all from increased bandwidth~\cite{rosenthal}.
+%Rosenthal et al. found increasing network bandwidth provide limited benefits to a number of applications. 
+%The applications that send mostly small messages or large messages asynchronously 
+%are not bandwidth bounded, hence, benefit only slightly or not at all from increased bandwidth~\cite{rosenthal}.
 
@@
 which is hard to be quantified by performance profiling tools alone. 
-Bogdan et al. provide a set of guidelines of how to configure a network with Dragonfly topology 
-for workload with Nearest Neighbor communication pattern~\cite{Bogdan-hpdc14}. 
+Bogdan et al. provide a set of guidelines of how to configure a Dragonfly network 
+for workload with nearest neighbor communication pattern~\cite{Bogdan-hpdc14}. 
 Dong et al. have developed simple benchmarks that conforms to four different communication patterns, 
@@
 on the single application's performance degradation due to network contention. 
-Finally, we analyze the impact of different allocation strategies to job's communication behaviors. 
-We further identify the optimal allocation strategy for 
+Finally, we analyze the impact of different placement strategies to job's communication behaviors. 
+We further identify the optimal placement strategy for 
 each application with a specific dominant communication pattern. 

### 2016-07-11 submission.tex
some minor tweaks

@@
 In this work, we choose three representative HPC applications from 
-DOE Design Forward Project and conduct detailed simulations of 
-on a Torus network model to analyze both intra- and inter-job interference. 
+DOE Design Forward Project and conduct detailed simulations  
+on a torus network model to analyze both intra- and inter-job interference. 
 By scrutinizing the communication behaviors of these applications, 
@@
 Our comprehensive analyses in this work illuminate a path 
-towards job-communication-pattern-aware HPC resource allocation.
+towards job communication pattern aware job placement on HPC systems.
 

### 2016-07-13 submission.tex
change title, add keywords
pass on Abstract

--- 2016-07-08 submission.tex
@@
 % can use linebreaks \\ within to get better formatting as desired
-\title{Study of Intra- and Inter-Job Interference with Different Job Allocations}
+\title{Study of Intra- and Inter-Job Interference on Torus Networks}
 % author names and affiliations
@@
 hence crucial to alleviate the potential performance degradation. 
-In order to do these effectively, 
-an understanding of the interference among the concurrently running jobs, 
-their communication patterns, and the network contention are required. 
+In order to do so effectively, 
+an understanding of the interference among concurrently running jobs, 
+their communication patterns, and contention in the network are required. 
 In this work, we choose three representative HPC applications from 
-DOE Design Forward Project and conduct detailed simulations of 
-on a Torus network model to analyze both intra- and inter-job interference. 
+the DOE Design Forward Project and conduct detailed simulations  
+on a torus network model to analyze both intra- and inter-job interference. 
 By scrutinizing the communication behaviors of these applications, 
@@
 the possible interference introduced by different job placement policies. 
-Our comprehensive analyses in this work illuminate a path 
-towards job-communication-pattern-aware HPC resource allocation.
+Our analyses in this work illuminate a path 
+towards communication pattern awareness in job placement on HPC systems.
 
@@
 \IEEEpeerreviewmaketitle
+
+\begin{IEEEkeywords} 
+HPC systems, Torus, Interference, Job Placement
+
+\end{IEEEkeywords}
 

### 2016-07-13 intro.tex
pass on Introduction

@@
 
-The scale of supercomputers continually grows at a breakneck pace to 
-accommodate the computing power requirements for scientific research. 
+The scale of supercomputer continually grows at a breakneck pace to 
+accommodate the computing requirements for scientific research. 
 Supercomputers have tens of thousands of nodes and serve as an irreplaceable 
@@
 accommodate many parallel applications (jobs) running concurrently. 
-These parallel jobs share the system infrastructure such as network and I/O bandwidth, 
+These jobs share the system infrastructure such as network and I/O bandwidth, 
 and inevitably there is contention over these shared resources. 
@@
 resulting in the shared use of resources, particularly network links. 
-A prominent problem with this network sharing is the contention over
-the network among those concurrently running jobs. 
-The network contention can cause communication variability and 
-performance degradation to those jobs~\cite{abhinav-sc13}. 
+A prominent problem with this network sharing is the resulting contention,
+which can cause communication variability and 
+performance degradation in affected jobs~\cite{abhinav-sc13}. 
 This performance degradation can propagate into the queueing time of the following submitted jobs, 
-thus leading to low system throughput and utilization~\cite{jose-ipdps15}. 
-This adverse effect of network sharing can be mitigated by 
-providing jobs with isolated allocations and exclusive network resources.
+thus leading to lower system throughput and utilization~\cite{jose-ipdps15}.
 
-On the widely used torus-connected HPC systems~\cite{bgq,tofu,titan}, 
+On the widely used torus network topology~\cite{bgq,tofu,titan}, 
 two job placement policies are commonly used. 
@@
 as shown in Figure~\ref{fig:overview_sub1}. 
-The partition-based placement adopted on the Blue Gene 
-series systems is an example of contiguous placement policy~\cite{bgloverview}. 
-The contiguous placement favors application performance 
-through exclusive networking within a compact partition 
-and the locality that implies. 
+The partition-based placement adopted in the Blue Gene 
+series supercomputers is a prominent example~\cite{bgloverview}. 
+Contiguous placement favors application performance 
+through exclusive resource allocation and the locality that implies. 
 However, contiguous placement can cause both internal fragmentation 
@@
 though of course efforts are made to maximize locality. 
-Figure~\ref{fig:overview_sub2} shows the effect of non-contiguous placement. 
+Figure~\ref{fig:overview_sub2} shows an example non-contiguous placement. 
 While eliminating internal and external fragmentation as seen in contiguous placement systems, 
-in return non-contiguous placement policy introduces contention between jobs due to the interleaving of job nodes. 
-The non-contiguous job placement can significantly reduce job performance, 
-especially for communication-intensive ones~\cite{abhinav-sc13}.
+in return non-contiguous placement policy introduces contention among jobs due to the interleaving of jobs' nodes, 
+particularly for communication-intensive jobs~\cite{abhinav-sc13}.
 
 We envision that future HPC systems will be equipped with 
-a flexible job placement mechanism which combines the 
-best of both contiguous and non-contiguous policy. 
-Such a flexible job placement mechanism should take 
+flexible job placement mechanisms that combine the 
+best of both contiguous and non-contiguous policies. 
+Such a mechanism would take 
 shared resource needs (e.g., network resources) of jobs into account 
@@
 communication interference with different job placements on torus-connected HPC systems. 
+\TODO{Move to Torus discussion.}
 Torus-based networks are used on six of the top 10 supercomputers 
@@
 We selected three signature applications from 
-the DOE Design Forward Project~\cite{designforwardwebpage} as examples 
-to conduct detailed study about the communication pattern of parallel applications. 
+the DOE Design Forward Project~\cite{designforwardwebpage}  
+to conduct a detailed study about the communication behavior of parallel applications. 
 We use a sophisticated simulation toolkit named CODES 
@@
 by simulating these applications running concurrently with different allocations. 
-We believe the insights presented in this work can be very useful 
+We believe the insights presented in this work can be useful 
 for the design of future HPC batch schedulers and resource managers.
@@
 from the DOE Design Forward Project for our study. 
-Section~\ref{sec:codes} talks about the use of CODES as research vehicle for our work. 
+Section~\ref{sec:codes} talks about the use of CODES as a research vehicle for our work. 
 %Section~\ref{sec:config study} shows the performance analysis of three applications on different torus networks. 
-Section~\ref{sec:interference} provides detailed analysis about the intra- and inter-job interference 
-among three applications on torus network with different allocations. 
+Section~\ref{sec:interference} provides detailed analyses about the intra- and inter-job interference 
+among the applications on a torus network with different allocations. 
 Section~\ref{sec:discussion} introduces a path toward communication-pattern aware allocation strategies, 

### 2016-07-13 app_study.tex
pass on app study

@@
 
-%A parallel application usually conforms to a combination of several basic communication patterns\cite{roth}.
-At its different execution phases, the application's communication behavior may follow different basic patterns respectively.
-\textcolor{blue}{When we look into the data flow during application execution, most parallel applications start with broadcast operation to distribute the data from root process to other processes\textcolor{red}{ref needed here}, followed by a series of computation and communication that conforms to certain pattern, which is usually the dominant part of application's execution.
-Before the application come to completion, all the working process will return their results to the root directly or hierarchically.} Parallel application's communication pattern usually involves lots of factors, such as communication intensity, operation-to-operation dependencies and the ``critical path" in its communication topology graph, etc.
-In this work, we focus on the locality of application's rank-to-rank communication topology graph.
-
+\begin{comment}
 A parallel application usually conforms to a combination of several basic communication patterns~\cite{roth}. 
@@
 These applications are communication intensive and conform to various communication patterns. 
-In this work, we select three representative applications from the DOE Design Forward Project. 
+\end{comment}
+
+In this work, we select three applications from the DOE Design Forward Project. 
 Each application exhibits a distinctive communication pattern that is commonly seen in HPC applications. 
@@
 Geometric MultiGrid (MultiGrid) and CrystalRouter MiniApps. 
-The communication matrix of each application is presented in Figure~\ref{fig:apps_communication_matrix}. 
+The communication matrixes of each application presented in Figure~\ref{fig:amg-communication-topology}
+,\ref{fig:cr-communication-topology} and \ref{fig:mg-communication-topology} are generated by
+the IPM data collected from~\cite{designforwardwebpage}.
 %are generated with the IPM~\cite{ipm} data gathered from publicly available traces~\cite{designforwardwebpage}.
 
-\begin{figure*}[htp]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/appstudy/amg/amg_pip}
-        \caption{AMG}
-        \label{fig:amg-communication-topology}
-    \end{subfigure}
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/appstudy/cr/cr_pip}
-        \caption{CrystalRouter}
-        \label{fig:cr-communication-topology}
-    \end{subfigure}
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/appstudy/mg/mg_pip}
-        \caption{MultiGrid}
-        \label{fig:mg-communication-topology}
-    \end{subfigure}
-    \caption{Applications rank-to-rank communication topology graph.}
-    \label{fig:apps_communication_matrix}
-\end{figure*}
+%\begin{figure*}[htp]
+%    \centering
+%    \begin{subfigure}[t]{0.32\textwidth}
+%        \centering
+%        \includegraphics[height=1.5in]{figs/appstudy/amg/amg_pip}
+%        \caption{AMG}
+%        \label{fig:amg-communication-topology}
+%    \end{subfigure}
+%    \begin{subfigure}[t]{0.32\textwidth}
+%        \centering
+%        \includegraphics[height=1.5in]{figs/appstudy/cr/cr_pip}
+%        \caption{CrystalRouter}
+%        \label{fig:cr-communication-topology}
+%    \end{subfigure}
+%    \begin{subfigure}[t]{0.32\textwidth}
+%        \centering
+%        \includegraphics[height=1.5in]{figs/appstudy/mg/mg_pip}
+%        \caption{MultiGrid}
+%        \label{fig:mg-communication-topology}
+%    \end{subfigure}
+%    \caption{Applications rank-to-rank communication topology graph.}
+%    \label{fig:apps_communication_matrix}
+%\end{figure*}
 
@@
 
-%\begin{figure}[t!]
-%    \centering
-%    \begin{subfigure}[t]{0.25\textwidth}
-%        \centering
-%        \includegraphics[height=1.5in]{figs/appstudy/amg/amg_pip}
-%        \caption{}
-%        \label{fig:amg-communication-topology}
-%    \end{subfigure}%
-%    ~
-%    \begin{subfigure}[t]{0.24\textwidth}
-%        \centering
-%        \includegraphics[height=1.2in]{figs/appstudy/amg/amg_data_transfer}
-%        \caption{}
-%        \label{fig:amg-data-trans}
-%    \end{subfigure}
-%    \caption{AMG.
-(a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in AMG.
-(b) The amount of data sent/received by each rank in AMG.
-Y-axis labels the amount of data transferred in MB, x-axis labels 216 ranks of AMG.
+\begin{figure}[htp]
+    \centering
+    \includegraphics[height=2.5in]{figs/appstudy/amg/amg_pip}
+    \caption{AMG Communication Matrix. 
+        The label of both x and y axis is the index of MPI rank in AMG. 
+        The legend bar on the right indicates the data transfer 
+        amount between ranks.
         }
-%\end{figure}
+    \label{fig:amg-communication-topology}
+\end{figure}
 
 The Algebraic MultiGrid Solver, or AMG, 
-is a parallel algebraic multi-grid solver for linear systems arising from problems on unstructured mesh physics packages. 
+is a parallel algebraic multi-grid solver for linear systems arising 
+from problems on unstructured mesh physics packages. 
 It has been derived directly from the BoomerAMG solver 
@@
 Note that the dominant communication pattern of the application doesn't change with scale. 
-We can make the observation that AMG's dominant communication pattern is 3D nearest neighbor, 
-each rank has intensive communication with up to six neighbors, 
-depending on the boundaries of the ranks. 
-Applications with similar pattern include PARTISN~\cite{partisn} and SNAP~\cite{snap}.
+We observe that AMG's dominant communication pattern is 3D nearest neighbor - 
+each rank has intensive communication with up to six neighbors, depending on rank boundaries. 
+Applications with similar patterns include PARTISN~\cite{partisn} and SNAP~\cite{snap}.
 
@@
 
-%\begin{figure}[t!]
-%    \centering
-%    \begin{subfigure}[t]{0.25\textwidth}
-%        \centering
-%        \includegraphics[height=1.5in]{figs/appstudy/cr/cr_pip}
-%        \caption{Communication Topology}
-%        \label{fig:cr-communication-topology}
-%    \end{subfigure}
-%    ~
-%    \begin{subfigure}[t]{0.22\textwidth}
-%        \centering
-%        \includegraphics[height=1.2in]{figs/appstudy/cr/cr_data_transfer}
-%        \caption{Data Volume}
-%        \label{fig:cr-data-trans}
-%    \end{subfigure}
-%    \caption{CrystalRouter.
-(a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in CrystalRouter.
-(b) The amount of data sent/received by each rank in CrystalRouter.
-Y-axis labels the amount of data transferred in MB, x-axis labels 100 ranks of CrystalRouter.
+\begin{figure}[htp]
+    \centering
+    \includegraphics[height=2.5in]{figs/appstudy/cr/cr_pip}
+    \caption{CrystalRouter Communication Matrix. 
+        The label of both x and y axis is the index of MPI rank in CrystalRouter. 
+        The legend bar on the right indicates the data transfer 
+        amount between ranks.
         }
-%\end{figure}
+    \label{fig:cr-communication-topology}
+\end{figure}
 
@@
 through a scalable multi-stage communication process.
-The way CrystalRouter works is the nodes compute for a while, 
-then synchronize and communicate, continually alternating between these two types of activities.
 
@@
 
-%\begin{figure}[t!]
-%    \centering
-%    \begin{subfigure}[t]{0.25\textwidth}
-%        \centering
-%        \includegraphics[height=1.5in]{figs/appstudy/mg/mg_pip}
-%        \caption{Communication Topology}
-%        \label{fig:mg-communication-topology}
-%    \end{subfigure}
-%    ~
-%    \begin{subfigure}[t]{0.22\textwidth}
-%        \centering
-%        \includegraphics[height=1.2in]{figs/appstudy/mg/mg_data_transfer}
-%        \caption{Data Volume}
-%        \label{fig:mg-data-trans}
-%    \end{subfigure}
-%    \caption{MultiGrid.
-(a) Rank-to-rank communication topology graph, shows the intensity of communication between ranks in MultiGrid.
-(b) The amount of data sent/received by each rank in MultiGrid.
-Y-axis labels the amount of data transferred in MB, x-axis label 125 ranks of MultiGrid.
+\begin{figure}[htp]
+    \centering
+    \includegraphics[height=2.5in]{figs/appstudy/mg/mg_pip}
+    \caption{MultiGrid Communication Matrix. 
+    The label of both x and y axis is the index of MPI rank in MultiGrid. 
+    The legend bar on the right indicates the data transfer amount between ranks.
     }
-%\end{figure}
+    \label{fig:mg-communication-topology}
+\end{figure}
 
@@
 In this sense, it can be considered as many-to-many pattern. 
-Applications with such similar dominant communication pattern are like FillBoundary, another PDE solver code in~\cite{boxlib}.
+Applications with such similar dominant communication patterns are like FillBoundary, another PDE solver code in~\cite{boxlib}.

### 2016-07-13 research_tool.tex
pass on Section 3

@@
 which can not be changed by users. 
-Even the system administrator is not authorized to make that change. 
+Even system administrators may not be able to make such changes. 
 Another reason is that it is unrealistic to reserve the system exclusively 
@@
 different HPC networks at flit-level with high fidelity~\cite{Jason-2011, mubarak-sc2012}. 
-CODES is built on top of Rensselaer Optimistic Simulation System (ROSS) parallel discrete-event simulator~\cite{ross}, 
+CODES is built on top of the Rensselaer Optimistic Simulation System (ROSS) parallel discrete-event simulator~\cite{ross}, 
 which is capable of processing billions of events per second on leadership-class supercomputers. 
 CODES support both torus and dragonfly network with high fidelity flit-level simulation.
-CODES has this network workload component that is capable of conducting trace-driven simulations. 
-It can take MPI application traces generated by SST DUMPI \cite{sst} to drive CODES network models. 
-In this work, we focus on an in-depth analysis of intra- and inter-job 
-communication interferences with different job allocations on torus-connected HPC systems. 
-Torus networks have been extensively used in the current generation of supercomputers 
-because of their linear scaling on per-node cost and competitive communication performance.
+CODES additionally has the capability of replaying MPI application traces, 
+gathered via the SST DUMPI profiler~\cite{sst}.
+%CODES has this network workload component that is capable of conducting trace-driven simulations. 
+%It can take MPI application traces generated by SST DUMPI  to drive CODES network models. 
+%In this work, we focus on an in-depth analysis of intra- and inter-job 
+%communication interferences with different job allocations on torus-connected HPC systems. 
+%Torus networks have been extensively used in the current generation of supercomputers 
+%because of their linear scaling on per-node cost and competitive communication performance.
+%

### 2016-07-13 interference_study.tex
pass on inteference study

@@
 \textbf{\emph{Intra-job interference}} refers to the network 
-contention between ranks within each application. 
+contention between the ranks within each application. 
 \textbf{\emph{Inter-job interference}} is introduced by concurrently 
-running jobs that adjacent to each other sharing network resources. 
+running jobs share network resources. 
 Communication variability due to such interference can 
@@
 in the context of a torus network. 
-Since application's communication pattern won't change with its scale, 
-we use AMG with 216 ranks, CrystalRouter with 100 ranks, and MultiGrid 125 ranks. 
-And in order to accommodate three applications running concurrently with different allocations, 
-we simulate a 2K-node 3D torus ($16^{2} \times 8$) network for the experiments.
+Since the applications communication patterns do not change with the scale,
+\TODO{we perform our experiments with comparatively (to leadership scale) small scale,
+something a 2048 rank cluster to simplify analysis.  }
+%we use AMG with 216 ranks, CrystalRouter with 100 ranks, and MultiGrid 125 ranks. 
+%And in order to accommodate three applications running concurrently with different allocations, 
+%we simulate a 2K-node 3D torus ($16^{2} \times 8$) network for the experiments.
 
@@
 intra-job interference of each application. 
-In the first, shown in Figure \ref{fig:shapestudy}, 
+In the first, 
 we assign each application with allocations in \emph{three different shapes}. 
-In the second, shown in Figure \ref{fig:diffmap}, 
+In the second, 
 we study the intra-job interference by using different 
-mapping strategies for each application with given allocation shape.
+mapping strategies for each application with a given allocation shape.
+
+\subsubsection{Allocation Shape Study}
 
@@
 as shown in Figure \ref{fig:cont_sub1}.
-3D balanced-cube, shown as red in Figure \ref{fig:cont_sub1}, 
+3D balanced-cube, shown in red in Figure \ref{fig:cont_sub1}, 
 can guarantee the minimum \emph{average pair-wise distance} within the allocation. 
@@
 
-3D unbalanced-cube, shown in green color in Figure \ref{fig:cont_sub1}, 
+3D unbalanced-cube, shown in green in Figure~\ref{fig:cont_sub1}, 
 is a rectangular prism, which is a possible allocation shape 
@@
 In order to take advantage of the faster links in the \emph{x} and \emph{z} directions, 
-job allocation starts from X-Z plane, 
-which leads to a rectangular prism shaped allocation~\cite{RF}.
+job allocation favors the X-Z plane~\cite{RF}. Our torus network in this study is symmetric.
 
-2D mesh, shown as blue in Figure \ref{fig:cont_sub1}, 
+2D mesh, shown in blue in Figure \ref{fig:cont_sub1}, 
 can be cut out from a single layer of the 3D torus. 
-2D mesh is a very common allocation shape on the torus network 
+2D mesh is a common allocation shape on the torus network 
 for both contiguous and non-contiguous placement policies. 
-For example, Cray Application Level Placement Scheduler (ALPS) indexes 
-all compute nodes in a torus network into a list and 
-makes job allocation by simply going through that list~\cite{carl-cug}. 
+For example, Cray's Application Level Placement Scheduler (ALPS) indexes 
+all compute nodes in the torus into a list and 
+allocates by simply going through that list~\cite{carl-cug}. 
 When the list is obtained by sorting the nodes based on 
 their spacial coordinates in the torus, 
-allocation made off from this list will be 2D mesh. 
+the resulting allocations form 2D meshes. 
 The IBM Blue/Gene Q supercomputer Mira at Argonne Leadership Computing Facility 
-also allow its allocation partition configured into mesh~\cite{zhou-ipdps}.
+also allows its allocation partition to be configured as a mesh~\cite{zhou-ipdps}.
 
-\begin{figure*}[t!]
+\begin{figure*}[htp]
     \centering
@@
 
-AMG with 3D nearest neighbor communication pattern takes less time 
-when being assigned with 3D-unbalanced allocation than 3D-balanced, 
-as shown in Figure~\ref{fig:shapstudy-amg}. 
-And MultiGrid gets the best performance (shortest data transfer time) 
+Figure~\ref{fig:shapstudy-amg} shows 
+AMG with 3D nearest neighbor communication pattern takes slightly less time 
+when being assigned with 3D-unbalanced allocation than 3D-balanced. 
+And MultiGrid performs best (shortest data transfer time) 
 when running on 3D-balanced allocation, as shown in Figure~\ref{fig:shapstudy-mg}. 
@@
 which can reduce the aggregated hops for transferring message among ranks in MultiGrid. 
-Since CrystalRouter has both dominant local communication and multi-stage global data transfer, 
+Since CrystalRouter exibits both local and global rank-to-rank data transfers, 
 the 3D-balance is also the best allocation, but its advantage over 3D-unbalanced is not 
@@
 
-Lots of research studies have designed complex placement algorithms 
-to provide applications with the most compact allocation~\cite{leung,LO}. 
-However, as shown in our experiments, providing cubic allocation without 
-considering application's communication pattern can not guarantee 
-the best performance for every application. 
-Compact allocation should be provided to applications with intensive global data transfer, 
-such as those conforms to many-to-many communication pattern. 
-Applications with dominant communication pattern like nearest neighbor 
-can not fully utilize the compact allocation and 
-may get even better performance on a relaxed allocation shape.
-
-\begin{figure*}[t!]
+\begin{figure*}[htp]
     \centering
@@
 
+A number of studies have designed complex placement algorithms 
+to provide applications with the most compact allocation~\cite{leung,LO}. 
+However, as shown in our experiments, providing cubic allocation without 
+considering application's communication pattern may not guarantee 
+the best performance for every application. 
+Compact allocation should be provided to applications with intensive global data transfer, 
+such as those exhibiting a many-to-many pattern.
+
+\subsubsection{Mapping Strategy Study}
 The rank-to-node mapping of parallel applications on HPC systems 
@@
 The ``Linear" mapping, as we used in the previous experiments, 
-is to map each rank according to the dimensional ordering of compute nodes. 
+maps each rank according to the dimensional ordering of compute nodes. 
 The ``Cube" mapping assigns ranks into consecutive $2^{3}$ cubes. 
@@
 
-AMG remains roughly the same when it been mapped by ``Linear" and ``Cube", 
+AMG \TODO{remains} roughly the same when it been mapped by ``Linear" and ``Cube", 
 shown in Figure~\ref{fig:diffmap-amg}. 
@@
 both mapping strategies still preserve the locality of AMG's 3D nearest neighbor communication pattern. 
-The ``Random" mapping totally destroys AMG's communication pattern, 
-and results in intra-job interference among all the ranks in AMG. 
+The ``Random" mapping disrupts AMG's communication pattern, 
+and results in intra-job interference among the ranks. 
 The performance degradation caused by using ``Random" mapping is up to 90\%.
@@
 The ``Cube" mapping improves CrystalRouter's performance over the ``Linear" 
-by reducing the pair-wise distance between ranks, shown in Figure~\ref{fig:diffmap-cr}. 
-The global data transfer in CrystalRouter will takes fewer hops with ``Cube" mapping. 
-The ``Random" mapping scatters the ranks within the local neighborhood 
+by up to 10\% on average, shown in Figure~\ref{fig:diffmap-cr}. 
+This is due to the global data transfer in CrystalRouter takes fewer hops with ``Cube" mapping. 
+The ``Random" mapping for CrystalRouter results in poor locality across all ranks on average. 
 of CrystalRouter and make their communication less efficient. 
 The ``Cube" mapping benefits MultiGrid's many-to-many communication. 
-However, due to the small amount of data transfer among ranks in MultiGrid, 
+However, due to the small amount of data transfered among ranks in MultiGrid, 
 ``Cube" mapping fails to exhibit a drastic advantage. 
-``Random" mapping neither cause too much degradation for the same reason, 
+``Random" mapping causes little degradation for the same reason, 
 shown in Figure~\ref{fig:diffmap-mg}.
 
-\begin{figure*}[t!]
+\begin{figure*}[htp]
     \centering
@@
 for application's performance variability~\cite{abhinav-sc13,skinner}. 
-Inter-job interference is a more prominent issue for system adopted 
-non-contiguous placement policy than systems with contiguous policy. 
-Applications communication time have been demonstrated to 
-vary from 36\% faster to 69\% slower due to job interference 
+Inter-job interference is a more prominent issue for systems adopting 
+non-contiguous placement policies than systems with contiguous policy. 
+Application communication times have been demonstrated to 
+vary from 36\% shorter to 69\% longer due to job interference 
 when they are running concurrently with non-contiguous allocations~\cite{abhinav-sc13}.
 
-We assign each application with non-contiguous allocation and 
+We allocate each application with non-contiguous policy and 
 run them concurrently on the same network. 
-The allocation unit belongs to different jobs are interleaved. 
-In order to study the impact of different allocation unit size on 
+The compute nodes belonging to different jobs are interleaved. 
+In order to study the impact of different allocation unit sizes on 
 applications' inter-job interference, 
@@
 shown respectively in Figure~\ref{fig:noncont_sub1}, 
-~\ref{fig:noncont_sub1} and~\ref{fig:noncont_sub3}. 
+~\ref{fig:noncont_sub2} and~\ref{fig:noncont_sub3}. 
 Figure~\ref{fig:interjobstudy} shows the results of each application 
@@
 The data transfer time of AMG in Figure~\ref{fig:interjob-amg} keeps stable 
-between allocation unit size of 16 and 8, since the locality of AMG is 6-rank based. 
-When the unit size reduce to 2, AMG suffers prolong data transfer time by about 10\%. 
+between allocation unit size of 16 and 8 because of the 
+nearest neighbor pattern of AMG. 
+When the unit size is reduced to 2, AMG suffers prolong data transfer time by about 10\% on average. 
 CrystalRouter is more sensitive to allocation unit size. 
-The best unit size would be the one big enough to accommodate its local neighborhood. 
 Figure~\ref{fig:interjob-cr} shows that unit size of 16 and 8 can guarantee 
 the same average data transfer time, 
-while some rank spend more time with allocation unit size 8 than 16. 
-When the unit size reduce to 2, the communication become less efficiency 
-and takes about 15\% more time for transferring data.
+while some ranks spend more time with allocation unit size 8 than 16. 
+When the unit size is reduced to 2, the communication becomes less efficient 
+and takes about 15\% more time on average for transferring data.
 
@@
 as shown in Figure~\ref{fig:interjob-mg}. 
-As the unit size decreases, the data transfer time keep growing, 
-but in a slow way, which means in terms of preserving the locality of MultiGrid, 
-allocation unit of size 16, 8 and 2 serve equally bad.
+Further unit size decreases result in roughly similar average communication times though.
 
 Choosing the proper unit size in non-contiguous placement policy should 
-also consider application's communication pattern. 
+consider application's communication patterns. 
 Inter-job interference is inevitable in non-contiguous based systems, 
-but unit size that big enough to preserve the neighborhood communication of the application 
+but unit sizes big enough to preserve the neighborhood communication of the application 
 will alleviate such interference and improve job performance.
@@
 
-Based on the above comprehensive study, we can make following observations:
+Based on our simulation study, we can make following observations:
 
@@
 
-    \item The compact allocation may not guarantee the best performance for every application.
+    \item Compact allocation may not be necessary for every application.
 
-    \item The applications dominated with nearest neighbor communications exhibit 
-    relatively stable performance under different allocation shapes, 
-    while the applications dominated with many-to-many communication exhibit 
-    better performance with compact allocation (e.g, 3D balanced).
+    \item The applications dominated by nearest neighbor communications exhibit 
+    relatively stable performance under different allocation shapes as long as
+    the allocation exhibit some degree of locality.
+
+    \item The applications dominated with many-to-many communication exhibit 
+    better performance with more compact allocations (e.g, 3D balanced).
 
     \item A good rank-to-node mapping strategy can greatly improve 
-    application's performance when specific allocation is given.
+    an application's performance when a specific allocation is given.
 
@@
     \item Inter-job interference is inevitable in non-contiguous allocation. 
-    However, choosing the proper allocation unit size with job's communication 
-    pattern awareness can alleviate such negative effect.
+    However, choosing the proper allocation unit size with communication 
+    pattern awareness can help alleviate the resulting negative effects.
 

### 2016-07-13 discussion.tex
pass on discussion

@@
 
-The results shown in the previous section provide insights 
-for the design of smart and flexible job placement policy. 
-By scrutinizing job's communication behavior, 
-we can identify job's dominant communication pattern 
-and pinpoint the ``neighborhood communication". 
-With such knowledge about jobs' communication patterns, 
+The results shown in the previous sections provide insights 
+for the design of a smart and flexible job placement policy. 
+By scrutinizing the communication behavior of jobs, 
+we can identify their dominant communication patterns 
+and pinpoint locality needs. 
+With such knowledge about communication patterns, 
 we can analyze the possible interference between jobs and 
-take precautions to alleviate the negative effect when making placement decisions.
+take precautions to alleviate interference when making placement decisions.
 
-When an application with dominant communication that is intensive many-to-many, 
-the scheduler should provide it with compact node allocation and exclusive network provision. 
-The compact allocation can guarantee the shortest pair-wise distance between all the ranks, 
-and make the data transfer between ranks take less hops. 
-On the other hand, 
-the exclusive network provision will prevent other adjacent applications from sharing network resources, 
-thus eliminate the performance degradation due to interference.
+When an application with intensive many-to-many communication is submitted to the system, 
+the scheduler should provide it with compact node allocation and exclusive network resources. 
+Compact allocation can guarantee the shortest pair-wise distance between all the ranks. 
+Additionally, the exclusive network provision will prevent other jobs from sharing network resources, 
+thus eliminating the performance degradation due to interference.
 
-Not every application's preferable resource are compact allocation 
-and exclusive network provision. 
+Not every application requires compact allocation and exclusive network provisioning. 
 As we demonstrate in our study, 
 applications whose dominant communication pattern contains intensive 
-``neighborhood communication" like nearest Neighbor do not benefit 
+``neighborhood communication" like nearest neighbor may not benefit 
 from compact allocation and exclusive network provision. 
 Applications, such as AMG, can run with non-contiguous node 
-allocation without performance degradation, 
-as long as the allocation unit can accommodate their ``neighborhood communication".
+allocation without significant performance degradation, 
+as long as the allocation unit can accommodate their rank-rank locality.
 
@@
 Instead, the scheduler should choose a proper unit size 
-based on the scale of ``neighborhood communication" of each job. 
-The best unit size should be neither too big nor too small, 
-just perfectly fit for the scale of that ``neighborhood". 
-Big unit size can not be fully utilized and thus cause fragmentation. 
-Small unit size will not be able to accommodate the ``neighborhood communication", 
-making the intra-job communication less efficient.
+based on the granularity of communication locality of each job.   
+Large unit sizes can not be fully utilized and thus cause fragmentation. 
+Small unit sizes will not be able to accommodate the communication locality, 
+resulting in less efficient intra-job communication.
 
-The advantage of making job placement with consideration about job's communication pattern is obvious. 
-First, compared with contiguous allocation policy, 
-the new scheduler with job's communication pattern awareness can be more flexible, 
-without requiring the system to provide a big contiguous partition to accommodate the whole application, 
-it only needs to provide a small set of compact nodes that is 
-sufficient for all the "local community" in the application. 
-On the other hand, 
-the new scheduler is able to preserve the locality of application's communication pattern. 
-Indeed, the small compact node set (allocation unit) provided for the application's 
-``local communities" can preserve the communication locality in the maximum extent. 
+The advantage of performing job placement with consideration 
+of job communication patterns is straightforward,
+given our experimental analyses. 
+Compared with contiguous placement policy, 
+schedulers with communication pattern awareness can be more flexible, 
+relaxing the need to provide contiguous partitions to accommodate the whole application,
+avoiding fragmentation issues inherent in contiguous placement.
+Indeed, smaller compact node sets (allocation units) provided for application's
+``local communication" can help preserve the communication performance to a sufficient extent. 
 The design of such a new scheduler with job communication pattern awareness 

### 2016-07-13 related_work.tex
pass on related work

@@
 There are many tools available for system monitoring and application profiling. 
-Tools like TAU (Tuning and Analysis)~\cite{tau} and mpiP~\cite{mpip} can capture 
-application's runtime information and keep records in event traces. 
-However, recognizing the patterns about application's communication from those traces requires substantial effort. 
+Tools such as TAU (Tuning and Analysis)~\cite{tau} and mpiP~\cite{mpip} can capture 
+application runtime information, keeping record in event traces. 
+However, recognizing communication patterns from those traces require substantial effort. 
 There are a number of studies on the recognition and characterization of parallel application communication patterns. 
-Oak Ridge National Laboratory has its ongoing project about developing a tool set named Oxbow, 
+Oak Ridge National Laboratory has an ongoing project about developing a tool set named Oxbow, 
 which can characterize the computation and communication behavior of scientific applications and benchmarks~\cite{oxbow}. 
@@
 Many research efforts have been conducted to characterize scientific applications. 
-For instance, a DOE project aims at the identification of the computational characteristics of 
-the DOE MiniApps developed at various exascale co-design centers~\cite{designforwardwebpage}. 
+For instance, the DOE Design Forward Project aims to identify the 
+computation and communication characteristics of a collection of relevant 
+MiniApps developed at a number of exascale co-design centers~\cite{designforwardwebpage}. 
 In this project, the communication patterns of several DOE full applications and associated 
-mini-applications are studied to provide a more complete snapshot of the DOE workload. 
+mini-applications are studied to provide a more complete snapshot of DOE workloads. 
 A joint project named CORAL from Oak Ridge, Argonne and Livermore provides a series of 
@@
 The CORAL project includes scalable science benchmarks, throughput benchmarks, 
-data centric benchmarks, skeleton benchmarks and Micro benchmarks.
+data centric benchmarks, skeleton benchmarks and micro benchmarks.
 
@@
 The interference among concurrently running jobs on HPC systems 
-have been identified as major culprit for job's performance variability. 
+have been identified as major a culprit for job's performance variability. 
 Bhatele et al. found that concurrently running applications interfere with each other, 
 and cause their communication time varied from 
-36\% faster to 69\% slower on different HPC systems~\cite{abhinav-sc13}. 
+36\% shorter to 69\% longer on different HPC systems~\cite{abhinav-sc13}. 
 Skinner et al. found that  there is a 2-3 times of slowdown in MPI$\textunderscore$Allreduce 
@@
 the interference of concurrently running jobs is getting worse, 
-which is hard to be quantified by performance profiling tools alone. 
-Bogdan et al. provide a set of guidelines of how to configure a Dragonfly network 
+which is hard to quantify with performance profiling tools alone. 
+Bogdan et al. provide a set of guidelines on how to configure a Dragonfly network 
 for workload with nearest neighbor communication pattern~\cite{Bogdan-hpdc14}. 
@@
 namely ping-pong, nearest neighbor, broadcast and all$\textunderscore$reduce, 
-to demonstrate the effectiveness of this highly parallel 5D torus network~\cite{Dong-SC11}.
+to demonstrate the effectiveness of 5D torus networks~\cite{Dong-SC11}.
 
-Our work is different from all these works in the following ways. 
+We differentiate our work from these works in the following ways. 
 First, we focus on the dominant communication patterns rather than any specific application. 
 We believe this can provide a guideline for other research work. 
-Secondly, we explore the intra- and inter-job interference between concurrently running jobs, 
-while similar work such as~\cite{abhinav-sc13} only focuses 
-on the single application's performance degradation due to network contention. 
-Finally, we analyze the impact of different placement strategies to job's communication behaviors. 
-We further identify the optimal placement strategy for 
+Secondly, we explore \TODO{the} intra- and inter-job interference between concurrently running jobs, 
+while similar work such as~\cite{abhinav-sc13} focuses 
+on one single application's performance degradation due to network contention. 
+Finally, we analyze the impact of different placement strategies to job's communication behaviors, 
+identifying preferred placement strategies for 
 each application with a specific dominant communication pattern. 
-Based on our study, we claim that further batch scheduler should take 
-job's communication pattern into consideration for its placement decision making.
+Based on our study, we claim that future batch schedulers should take 
+job communication patterns into consideration for placement decision making.

### 2016-07-13 app_study.tex
final pass

@@
 ,\ref{fig:cr-communication-topology} and \ref{fig:mg-communication-topology} are generated by
-the IPM data collected from~\cite{designforwardwebpage}.
+the IPM data collected from the DOE Design Forward Project~\cite{designforwardwebpage}.
 %are generated with the IPM~\cite{ipm} data gathered from publicly available traces~\cite{designforwardwebpage}.

### 2016-07-13 conclusion.tex
final pass

@@
 namely AMG, CrystalRouter and MultiGrid. 
-Each application has distinctive communication pattern, 
-that can be representative for a group of jobs in HPC workload. 
-We have used a sophisticated simulation tool named CODES from Argonne National Laboratory 
-to simulate the running of these three parallel applications on a torus network. 
-The torus network provided by CODES has good fidelity and scalability. 
-%We have analyzed the performance of each application's communication in terms of data transfer time 
-%by simulating them running on torus networks with different bandwidth and dimensionality configurations.
-We have found that higher dimensionality of torus network would improve the performance of application with ``Many-to-Many" communication patterns, while application with intensive local communication like ``Nearest Neighbor" won't benefit much from higher dimensionality.
+Each application has a distinctive communication pattern, 
+which can be representative for a whole range of jobs commonly seen in HPC environment. 
+We have used the CODES simulation toolkit  
+to simulate the running of these applications on a torus network.
 
-We have analyzed the intra- and inter-job interference by simulating three applications running on the torus network. 
-Based on our comprehensive experiments, we got five observations. 
-1) The compact allocation may not guarantee the best performance for every application. 
-2) The applications dominated with nearest neighbor communications exhibit 
-relatively stable performance under different allocation shapes, 
-while the applications dominated with many-to-many communication exhibit 
-better performance with compact allocation (e.g. 3D balanced). 
-3) A good rank-to-node mapping strategy can greatly improve application's performance when specific allocation is given. 
-4) An optimal size for allocation units should be determined according to 
-an application's dominant communication pattern. 
-In general, a unit size should be large enough to preserve neighboring communication of the application. 
-5) Inter-job interference is inevitable in non-contiguous allocation. 
-However, choosing the proper allocation unit size with job's communication pattern awareness can alleviate such negative effect.
+We have analyzed the intra- and inter-job interference by simulating three applications
+running both independently and concurrently.
+Based on our comprehensive experiments, we made six observations.
+
+1) Compact allocation may not be necessary for every application.   
+2) The applications dominated by nearest neighbor communications exhibit 
+    relatively stable performance under different allocation shapes as long as
+    the allocation exhibit some degree of locality.
+
+3) The applications dominated with many-to-many communication exhibit 
+    better performance with more compact allocations (e.g, 3D balanced).
+
+4) A good rank-to-node mapping strategy can greatly improve 
+    an application performance when a specific allocation is given.
+
+5) An optimal size for allocation units should be determined 
+    according to an application's dominant communication pattern. 
+    In general, a unit size should be large enough to accommodate  
+    neighboring communication in the application.
+
+6) Inter-job interference is inevitable in non-contiguous allocation. 
+    However, choosing the proper allocation unit size with communication 
+    pattern awareness can help alleviate the resulting negative effects.
 
@@
 for HPC batch scheduler and resource manager to make flexible job allocations. 
-Rather than using pre-defined partitions and non-contiguous placement policy, 
-future HPC systems should assign each job with preferable resources based on job's communication pattern.
+Rather than using pre-defined partitions or non-contiguous placement policy, 
+future HPC systems should assign resources to each job based on job communication patterns.

### 2016-07-13 interference_study.tex
final pass

@@
 in the context of a torus network. 
+The current generation of IBM Blue Gene/Q (BG/Q) supercomputers, 
+such as Mira at Argonne Nation Laboratory and 
+Sequoia at Lawrence Livermore National Laboratory, 
+have their compute nodes connected by a 5D torus network~\cite{bgq}. 
+The K computer from Japan uses the ``Tofu'' interconnected system, 
+which has a 6D mesh/torus topology~\cite{tofu}. 
+Titan, a Cray XK7 supercomputer located at the Oak Ridge Leadership Computing Facility (OLCF), 
+has nodes connected in a 3D torus within the compute partition~\cite{titan}. 
 Since the applications communication patterns do not change with the scale,
-\TODO{we perform our experiments with comparatively (to leadership scale) small scale,
-something a 2048 rank cluster to simplify analysis.  }
+we perform our experiments with comparatively (to leadership class) modest scale,
+simulating a 3D torus network with 2048 nodes ($16\times 16 \times 8$) to simplify analysis.
 %we use AMG with 216 ranks, CrystalRouter with 100 ranks, and MultiGrid 125 ranks. 
@@
 \subsubsection{Allocation Shape Study}
+
+\begin{figure}[t]
+    \centering
+        \includegraphics[height=2in]{figs/allocshape/allocation}
+        \caption{Contiguous allocation in three different shapes. 
+    ``Blue" is a 2D mesh, ``Green" is a 3D-unbalanced cube and ``Red" is a 3D-balanced cube. }
+        \label{fig:cont_sub1}
+\end{figure}
 
@@
 
-\begin{figure*}[htp]
+\begin{figure*}[t]
     \centering
-    \begin{subfigure}[t]{0.22\textwidth}
+    \begin{subfigure}[t]{0.32\textwidth}
         \centering
-        \includegraphics[height=1.2in]{figs/allocshape/allocation}
-        \caption{ }
-        \label{fig:cont_sub1}
+        \includegraphics[height=1.8in]{figs/intra-job/shapestudy/amg_box}
+        \caption{AMG}
+        \label{fig:shapstudy-amg}
     \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.22\textwidth}
+    \begin{subfigure}[t]{0.32\textwidth}
         \centering
-        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit16}
-        \caption{ }
-        \label{fig:noncont_sub1}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.22\textwidth}
+        \includegraphics[height=1.8in]{figs/intra-job/shapestudy/cr_box}
+        \caption{CrystalRouter}
+        \label{fig:shapstudy-cr}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.32\textwidth}
         \centering
-        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit8}
-        \caption{ }
-        \label{fig:noncont_sub2}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.22\textwidth}
-        \centering
-        \includegraphics[height=1.2in]{figs/allocshape/unitsize/unit2}
-        \caption{ }
-        \label{fig:noncont_sub3}
+        \includegraphics[height=1.8in]{figs/intra-job/shapestudy/mg_box}
+        \caption{MultiGrid}
+        \label{fig:shapstudy-mg}
     \end{subfigure}%
    \caption{
-    (a) Contiguous allocation in three different shapes. 
-    ``Blue" is a 2D mesh, ``Green" is a 3D-unbalanced cube and ``Red" is a 3D-balanced cube. 
-    (b)-(d) Non-contiguous allocation. Each job is represented by a specific color. 
-    The nodes assigned to different jobs are interleaved, 
-    the size of allocation unit are 16, 8 and 2.
+   Data transfer time of AMG, CrystalRouter and MultiGrid on 3D-balanced, 
+   3D-unbalanced and 2D allocation.
    }
-    \label{fig:alloc-shapes}
+   \label{fig:shapestudy}
 \end{figure*}
@@
 that obvious as to MultiGrid, as shown in Figure~\ref{fig:shapstudy-cr}.
-
-\begin{figure*}[htp]
-    \centering
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/amg_box}
-        \caption{AMG}
-        \label{fig:shapstudy-amg}
-    \end{subfigure}%
-    \hspace{1em}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/cr_box}
-        \caption{CrystalRouter}
-        \label{fig:shapstudy-cr}
-    \end{subfigure}%
-    \begin{subfigure}[t]{0.32\textwidth}
-        \centering
-        \includegraphics[height=1.5in]{figs/intra-job/shapestudy/mg_box}
-        \caption{MultiGrid}
-        \label{fig:shapstudy-mg}
-    \end{subfigure}%
-   \caption{
-   Data transfer time of AMG, CrystalRouter and MultiGrid on 3D-balanced, 
-   3D-unbalanced and 2D allocation.
-   }
-   \label{fig:shapestudy}
-\end{figure*}
 
@@
 
-AMG \TODO{remains} roughly the same when it been mapped by ``Linear" and ``Cube", 
+AMG behavior remains roughly the same when it been mapped by ``Linear" and ``Cube", 
 shown in Figure~\ref{fig:diffmap-amg}. 
@@
 
-\begin{figure*}[htp]
+\begin{figure*}[t]
     \centering
@@
         \centering
-        \includegraphics[height=1.5in]{figs/diffmapping/amg}
+        \includegraphics[height=1.8in]{figs/diffmapping/amg}
         \caption{AMG}
@@
         \centering
-        \includegraphics[height=1.5in]{figs/diffmapping/cr}
+        \includegraphics[height=1.8in]{figs/diffmapping/cr}
         \caption{CrystalRouter}
@@
         \centering
-        \includegraphics[height=1.5in]{figs/diffmapping/mg}
+        \includegraphics[height=1.8in]{figs/diffmapping/mg}
         \caption{MultiGrid}
@@
 
+\begin{figure*}[t]
+    \centering
+%    \begin{subfigure}[t]{0.22\textwidth}
+%        \centering
+%        \includegraphics[height=1.2in]{figs/allocshape/allocation}
+%        \caption{ }
+%        \label{fig:cont_sub1}
+%    \end{subfigure}%
+%    \hspace{1em}%
+    \begin{subfigure}[t]{0.32\textwidth}
+        \centering
+        \includegraphics[height=1.8in]{figs/allocshape/unitsize/unit16}
+        \caption{ }
+        \label{fig:noncont_sub1}
+    \end{subfigure}%
+    \hspace{1em}%
+    \begin{subfigure}[t]{0.32\textwidth}
+        \centering
+        \includegraphics[height=1.8in]{figs/allocshape/unitsize/unit8}
+        \caption{ }
+        \label{fig:noncont_sub2}
+    \end{subfigure}%
+    \hspace{1em}%
+    \begin{subfigure}[t]{0.32\textwidth}
+        \centering
+        \includegraphics[height=1.8in]{figs/allocshape/unitsize/unit2}
+        \caption{ }
+        \label{fig:noncont_sub3}
+    \end{subfigure}%
+    \caption{
+%    (a) Contiguous allocation in three different shapes. 
+%    ``Blue" is a 2D mesh, ``Green" is a 3D-unbalanced cube and ``Red" is a 3D-balanced cube. 
+    Non-contiguous allocation. Each job is represented by a specific color. 
+    The nodes assigned to different jobs are interleaved, 
+    the size of allocation unit are 16, 8 and 2.
+    }
+    \label{fig:alloc-shapes}
+\end{figure*}
+
 The data transfer time of AMG in Figure~\ref{fig:interjob-amg} keeps stable 
@@
 
-\begin{figure*}[t!]
+\begin{figure*}[t]
     \centering
@@
         \centering
-        \includegraphics[height=1.5in]{figs/inter-job/amg}
+        \includegraphics[height=1.8in]{figs/inter-job/amg}
         \caption{AMG}
@@
         \centering
-        \includegraphics[height=1.5in]{figs/inter-job/cr}
+        \includegraphics[height=1.8in]{figs/inter-job/cr}
         \caption{CrystalRouter}
@@
         \centering
-        \includegraphics[height=1.5in]{figs/inter-job/mg}
+        \includegraphics[height=1.8in]{figs/inter-job/mg}
         \caption{MultiGrid}
@@
     \item A good rank-to-node mapping strategy can greatly improve 
-    an application's performance when a specific allocation is given.
+    an application performance when a specific allocation is given.
 
@@
     In general, a unit size should be large enough to accommodate  
-    neighboring communication of the application.
+    neighboring communication in the application.
 

### 2016-07-13 intro.tex
final pass

@@
 communication interference with different job placements on torus-connected HPC systems. 
-\TODO{Move to Torus discussion.}
-Torus-based networks are used on six of the top 10 supercomputers 
-on the June 2015 Top500 lists~\cite{top500}. 
-The current generation of IBM Blue Gene/Q (BG/Q) supercomputers, 
-such as Mira at Argonne Nation Laboratory and 
-Sequoia at Lawrence Livermore National Laboratory, 
-have their compute nodes connected by a 5D torus network~\cite{bgq}. 
-The K computer from Japan uses the ``Tofu'' interconnected system, 
-which has a 6D mesh/torus topology~\cite{tofu}. 
-Titan, a Cray XK7 supercomputer located at the Oak Ridge Leadership Computing Facility (OLCF), 
-has nodes connected in a 3D torus within the compute partition~\cite{titan}. 
 Although our analyses are based on torus networks, 

### 2016-07-13 related_work.tex
final pass

@@
 We believe this can provide a guideline for other research work. 
-Secondly, we explore \TODO{the} intra- and inter-job interference between concurrently running jobs, 
+Secondly, we explore both intra- and inter-job interference between concurrently running jobs, 
 while similar work such as~\cite{abhinav-sc13} focuses 
